### 2025-04-23 literature/paper-outline.Rmd
Initial commit for new repo

---
title: "Outline - Covariance Estimation for Hierarchical/Grouped Time Series Using the NOVELIST Estimator"
author: "Vincent Su"
date: "`r Sys.Date()`"
output: 
  pdf_document
header-includes:
  - \usepackage{bm}
  - \usepackage{titlesec}
  - \titlespacing{\section}{0pt}{30pt plus 12pt minus 1pt}{0pt plus 1pt minus 1pt}
# bibliography: bibliography.bib
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
```

# Notation

-   Scalar $y_t$

-   Vector $\bm{y}_t$

-   Matrix $\bm{S}$

-   Covariance matrix of in-sample 1-step-ahead base forecast errors $\hat{\bm{W}}_1$

-   Its shrinkage estimator with diagonal target $\hat{\bm{W}}^{shr}_{1, D}$

-   Its NOVELIST estimator $\hat{\bm{W}}^{shr}_{1, thr}$

# Abstract

[ This is pasted from the Project Description ]{style="color:red;"}

A collection of time series connected via a set of linear constraints is known as hierarchical time series. Forecasting these structures is challenging because the individual forecasts do not satisfy the linear constraints present in the data. To mitigate this issue, various forecast reconciliation approaches have been proposed in the literature, where the individual forecasts are adjusted by minimizing the mean squared reconciliation errors. Among these, the MinT (Minimum Trace) approach is widely used. However, this method requires an estimate of the covariance matrix of the base forecast errors. Although the sample covariance matrix is a natural choice, its performance is greatly impacted by the high-dimensional nature of hierarchical time series encountered in practice. Most published research uses shrinkage-type estimators that shrink the sample covariance matrix toward a diagonal matrix.

In this project, we aim to assess the forecasting performance of MinT when different covariance estimators are used, with a primary focus on the NOVELIST (NOVEL Integration of the Sample and Thresholded Covariance) estimator, which shrinks the sample covariance toward its thresholded version. A thorough literature review will also identify other high-dimensional covariance estimators relevant to this project.

# 1. Introduction

-   **Context & Motivation:**

    Hierarchical (and grouped) time series are collections of time series data connected via linear constraints (linear combinations?). Forecast reconciliation ensures that forecasts satisfy these constraints.

    One of the widely used appoarch is the MinT approach. It minimises the trace of reconciled forecast errors covariance matrix (minimum variance) and depends critically on the covariance matrix of h-step ahead base forecast errors $W_h$.

-   **Problem Statement:**\
    The sample covariance matrix, although natural, suffers in high-dimensional settings. Especially when the number of series $p$ is huge and larger than the time dimension $T$, the sample covariance matrix is non-positive definite (rank T if p\>T).

    The shrinkage estimators come in to tackle this issue. The shrinkage estimator with diagonal target (often shrinking toward a diagonal matrix) is proven to produce a guaranteed PD matrix (Schäfer & Strimmer, 2005). However, as it shrinks the covariance matrix toward a diagonal one, it does not have flexibility and might neglect the prominent structure presented in the covariance matrix.

    An alternative approach is to perform shrinkage of the sample covariance towards its thresholded version, instead of a diagonal matrix. This is the NOVELIST (NOVEL Integration of the Sample and Thresholded covariance estimators) method proposed by Huang & Fryzlewicz (2019). They introduced thresholding functions applied only to off-diagonal elements, allowing for more flexibility in the estimation.

    [ ... can include more estimators ... ]{style="color:red;"}

-   **Research Aim:**\
    This paper assesses the reconciled forecasting performance of MinT approach using various covariance estimators, with a focus on the NOVELIST estimator.

-   **Paper Outline:**\
    The paper is structured as follows:

    -   A literature review of forecast reconciliation and covariance estimation.
    -   A description of the methodology, including the NOVELIST estimator and its principal-component-adjusted variant.
    -   An experimental design using both synthetic and real hierarchical time series.
    -   Empirical results and discussion.
    -   Conclusions and suggestions for future work.

# 2. Literature Review

## 2.1 Forecast Reconciliation in Hierarchical and Grouped Time Series

-   Brief history of reconciliation methods (bottom-up, top-down, GLS, MinT).
-   Statistical challenges and practical applications of MinT.

## 2.2 Covariance Estimation in High Dimensions

-   Limitations of the sample covariance matrix.
-   Estimators used by Wickramasuriya et al. (2019).
-   Shrinkage estimators:
    -   Diagonal shrinkage (e.g., Schäfer & Strimmer, Ledoit & Wolf).
    -   NOVELIST estimator and its Cross-validation & PC-adjusted variant.
    -   ...

## 2.3 Relevance to Forecast Reconciliation

-   Discuss how covariance estimation affects MinT performance.
-   Identify research gaps.

# 3. Methodology

## 3.1 Theoretical Framework

### Hierarchical and Grouped Time Series Structure

Hierarchical/Grouped time series satisfy a set of linear constraints, often represented as: $$
y_t = S b_t,
$$ where $S$ is a summing matrix of order $m x n$ which aggregates the bottom-level series $b_t$ ($n$-vector) to the series at aggregation levels above. The $m$-vector $y_t$ contains all observations at time $t$

Let $\hat{y}_t(h)$ be a vector of h-step-ahead base forecasts for each time series in the collection, using information up to and including time $t$, and stacked in the same order as $y_t$. Then we can rewrite the linear reconciliation forecast equation as:

$$
\tilde{y}_t(h) = SP\hat{y}_t(h)
$$ where $P$ is an $n$x$p$ matrix that maps the base forecasts into bottom-level disaggregated forecasts $\hat{b}_t$.

-   Expand on this, include graphs, examples...

### The Minimum Trace (MinT) Reconciliation

*Lemma 1*: ... the covariance matrix of the h-step-ahead reconciled forecast errors is given by

$$
var[y_{t+h} - \tilde{y}_t(h) | F_t] = SPW_hP'S'
$$ where $W_h = E[\hat{e}_t(h) \hat{e}'_t(h) | F_t]$ is the covariance matrix of the h-step-ahead base forecast errors.

The MinT reconciled forecast is obtained by minimizing the trace of the above covariance matrix: $$
\min \; tr(SPW_hP'S') \quad \text{subject to } PS = I.
$$

which gives the optimal reconciliation matrix: $$
P = (S' W_h^{-1} S)^{-1} S' W_h^{-1}
$$

## 3.2 Covariance Estimation Approaches

-   $W_h$ is challenging to estimate, when $h$ \> 1, we need alternative estimates.

For forecast errors $\hat{e}_{t|t-1}$, the unbiased sample covariance of in-sample one-step-ahead base forecast errors is: $$
\hat{W}_1 = \frac{1}{T} \sum_{t=1}^{T} \hat{e}_{t|t-1} \hat{e}_{t|t-1}^{\top}.
$$

We reconstruct estimator of $W_h$ as $\hat{W}_h = k_h \, g(\hat{W}_1)$, where $k_h > 0$ simplifying the computation.

We will look into a few "sparse" + "non-sparse" approaches that apply on $\hat{W}_1$.

### The Shrinkage Estimator

$$
g(\hat{W}_1) = \hat{W}^{shr}_{1, D} = \lambda_D \hat{W}_{1, D} + (1 - \lambda_D) \hat{W}_1
$$ is the shrinkage estimator with diagonal target, proposed by Schäfer & Strimmer (2005). $\hat{W}_{1, D}$ is a diagonal matrix comprising the diagonal entries of $\hat{W}_1$. This approach will shrink the covariance matrix $\hat{W}_1$ towards its diagonal version, meaning the off-diagonal elements are shrunk towards zero while the diagonal ones remain unchanged.

They also proposed an optimal shrinkage intensity parameter $\lambda_D$ for this setting, assuming the variances are constant:

$$
\hat{\lambda}_D = \frac{\sum_{i \neq j} \widehat{Var}(\hat{r}_{ij})} {\sum_{i \neq j} \hat{r}_{ij}}
$$

-   (need more explanation on how to obtain optimal param)

-   Walkthrough the shrinkage estimator method

### The NOVELIST Estimator

$$
g(\hat{W}_1) = \hat{W}^{shr}_{1, thr} = \lambda_{\delta} \hat{W}_{1, \delta} + (1 - \lambda_{\delta}) \hat{W}_1
$$

is the NOVELIST shrinkage estimator, proposed by Huang & Fryzlewicz (2019). By convenient setting, we rewrite as sample correlation:

$$
\hat{R}^{shr}_{1, thr} = \lambda_{\delta} \hat{R}_{1,\delta} + (1 - \lambda_{\delta}) \hat{R}_1,
$$

$\hat{R}_{1, \delta}$ is a thresholded correlation matrix, in which thresholding is applied only to each off-diagonal element. This approach will shrink the sample correlation matrix $\hat{R}_1$ towards its thresholded version. There are various choices for the thresholding function, in this work, we use the soft-thresholding operator, defined as:

$$
\hat{r}_{1,ij}^\delta = \operatorname{sign}(\hat{r}_{1,ij}) \, (|\hat{r}_{1,ij}| - \delta)_+.
$$

After calculated the NOVELIST correlation matrix, we can re-obtain the covariance matrix $\hat{W}^{shr}_{1, thr} = \hat{D}_1^{1/2} \hat{R}^{shr}_{1, thr} \; \hat{D}_1^{1/2}$, where $\hat{D}_1 = diag(\hat{W}_1)$ is the diagonal matrix and the elements are given by the variances of the 1-step-ahead base forecast errors.

For a given threshold $\delta$, the optimal shrinkage intensity parameter $\lambda(\delta)$ can be estimated as:

$$
\hat{\lambda}(\delta) = \frac{
  \sum_{i \neq j} \widehat{Var}(\hat{r}_{1,ij}) \; I(|\hat{r}_{1,ij}| \leq \delta)
} {
  \sum_{i \neq j} (\hat{r}_{1,ij} - \hat{r}_{1,ij}^\delta)^2
}
$$

The formula derived using Ledoit-Wolf's lemma (Ledoit and Wolf, 2003) by Huang & Fryzlewicz. For the threshold parameter $\delta$, we use a cross-validation to select the optimal value

-   ALGORITHM WALKTHROUGH

**Principal-Component Adjustment**

When a factor structure is present, the procedure is:

<!-- 1. *Standardize Forecast Errors:* -->

<!-- $$ -->

<!-- \hat{z}_{t|t-1} = D_1^{-1/2} \hat{e}_{t|t-1}. -->

<!-- $$ -->

<!-- 2. *Construct the Data Matrix:* -->

<!-- $$ -->

<!-- \hat{Z} = [\hat{z}_{1|0}, \hat{z}_{2|1}, \dots, \hat{z}_{T|T-1}] -->

<!-- $$ -->

<!-- 3. *Extract Principal Components:* -->

<!-- Compute the top K eigenvectors to form: -->

<!-- $$ -->

<!-- \hat{B} = T^{-1} \hat{Z} \hat{V}_K, -->

<!-- $$ -->

<!-- where $\hat{V}_K$ contains the K largest eigenvectors. -->

<!-- 4. *Residual Component:* -->

<!-- $$ -->

<!-- \hat{Z}_{\text{rem}} = \hat{Z} - \hat{B}\hat{V}_K^\top. -->

<!-- $$ -->

<!-- Then compute the residual correlation: -->

<!-- $$ -->

<!-- \tilde{R}_1 = T^{-1\prime} (\hat{Z} - \hat{B}\hat{V}_K^\top)(\hat{Z} - \hat{B}\hat{V}_K^\top)^\top. -->

<!-- $$ -->

<!-- 5. *Apply NOVELIST to Residuals:* -->

<!-- $$ -->

<!-- \hat{R}_{K,\text{shr}}^{1,\text{thresh}} = \lambda \tilde{R}_\delta^1 + (1 - \lambda) \tilde{R}_1. -->

<!-- $$ -->

<!-- 6. *Reconstruct Full Estimator:* -->

<!-- $$ -->

<!-- \hat{R}_{1,K} = \hat{B}\hat{B}^\top + \hat{R}_{K,\text{shr}}^{1,\text{thresh}}. -->

<!-- $$ -->

<!-- **Shrinkage Intensity \(\lambda\):**   -->

<!-- Estimated as -->

<!-- $$ -->

<!-- \hat{\lambda} = \max\Big(0, \min\Big(1, \frac{\sum_{i \neq j} \operatorname{Var}(\hat{r}_{1,ij})}{\sum_{i \neq j} \hat{r}_{1,ij}^2}\Big)\Big). -->

<!-- $$ -->

<!-- **Threshold \(\delta\):**   -->

<!-- Selected via grid search or cross-validation. -->

### Alternative Covariance Estimators

Brief overview of other high-dimensional estimators used in the literature for benchmarking.

Others:

-   Principal Orthogonal complEment Thresholding (POET) by Fan et al. (2013)

    A Low-rank + Sparse method.\
    It decompose the covariance matrix into a prominent principle components part (Low-rank) and a orthogonal complement part $R_K$. Then apply thresholding to $R_K$.\
    This is similar to the NOVELIST estimator with PC-adjusted, but the difference is that POET apply thresholding to $R_K$, not a NOVELIST function.

# 4. Simulation

-   Talk about description of the hierarchical time series data set (e.g., economic, financial, or synthetic data).
-   Characteristics such as dimensionality, frequency, and hierarchical structure.

**Experimental Design**

-   Design different cases of simulation studies and real-data experiments.
-   Metrics: Forecast accuracy (e.g., RMSE, MAE), reconciliation error reduction, and matrix stability...
-   Visualisation...

**General Set-up I'm currently working on**

The designed data generating process for bottom-level series is a stationary VAR(1) process, with the following structure:

$$
\bm{y}_t = \bm{A} \bm{y}_{t-1} + \bm{e}_t,
$$

where $\bm{A}$ is a $p \times p$ block diagonal matrix of autoregressive coefficients $\bm{A} = diag(\bm{A}_1, \ldots, \bm{A}_m)$, with each $\bm{A}_i$ being a $p_i \times p_i$ matrix. The block diagonal structure ensures that the time series are grouped into $m$ groups, with each group having its own autoregressive coefficients. This aim to simulate the interdependencies between the time series within each group, where reconciliation will be better performed than the usual base forecasts.

The model is added with a Gaussian innovation process $\bm{e}_t$, with covariance matrix $\Sigma$. The covariance matrix $\Sigma$ is generated specifically in the following way:

1. A compound symmetric correlation matrix is used for each block of size $p_i$ in $\bm{A}_i$, where the coefficients are sampled from a uniform distribution.

2. The correlations between different blocks are imposed using the Algorithm 1 in Hardin, Garcia & Golan (2013).

3. The covariance matrix $\Sigma$ is then constructed by uniform sampling of standard deviations for all $p$ series.

We have an option to randomly flip the signs of the covariance elements, which will create a more realistic structure in the innovation process. This is also to simulate the real-world scenario where the observed covariance matrix is not necessarily positive definite.

# 5. Empirical Analysis

-   Comparative analysis of forecast performance using different covariance estimators on real-life dataset.

# 6. Discussion and Conclusion

-   Evaluate how covariance estimation impacts the MinT reconciliation.

-   Advantages and limitations of the NOVELIST estimator (with different ways of choosing threshold parameter).

-   Practical considerations: Computational efficiency, robustness, and ease of implementation.

-   Limitations of the research

# References

### 2025-04-23 literature/pseudo_novelist_cv.Rmd
Initial commit for new repo

---
title: "pseudo_novelist_cv"
author: "Vincent Su"
date: "2025-04-04"
output: pdf_document
header-includes:
  - \usepackage{bm}
---

## Procedure:

**Preliminary**

- Fit the model on $(y_1, \ldots, y_T)$ for each of the $p$ time series.
- Compute in-sample 1-step-ahead forecast errors $\hat{e}_{t|t-1}$ for $t=2,3,\ldots,T$

(in case such as AR(2) should we exclude the first 2 obs?)

\vspace{2em}

**Algorithm**

**Input**: matrices $\bm{y}_t, \hat{\bm{e}}_{t|t-1} \in \mathbb{R}^{T-1 \times p}$  
**Output**: optimal NOVELIST matrix $\hat{W}_1^* \in \mathbb{R}^{p \times p}$, threshold scalar $\delta^*$

for $i = n:T-1$ do  
\hspace*{1em} Training set $= \{ \hat{\bm{e}}_{t|t-1} : t = i-n+1, \dots, i \}$  
\hspace*{1em} for $\delta = 0:1$ by $x$ step do  
\hspace*{2em} Compute $\hat{R}_i$ using training set  
\hspace*{2em} Compute thresholded $\hat{R}_{i,\delta}$  
\hspace*{2em} Compute $\lambda_{i,\delta}$ using $\hat{R}_{i,\delta}$  
\hspace*{2em} $\hat{R}^{N}_{i,\delta} \leftarrow \lambda_{i,\delta} \hat{R}_{i,\delta} + (1 - \lambda_{i,\delta}) \hat{R}_i$  
\hspace*{2em} $\hat{W}^{N}_{i,\delta} \leftarrow \hat{D}^{1/2} \hat{R}^{N}_{i,\delta} \hat{D}^{1/2}$  
\hspace*{2em} Compute 1-step-ahead reconciled forecasts $\tilde {\bm{y}}_{i+1,\delta}$  
\hspace*{2em} $\tilde {\bm{e}}_{i+1,\delta} \leftarrow {\bm{y}}_{i+1} - \tilde {\bm{y}}_{i+1,\delta}$  
\hspace*{1em} end  
end

<!-- TODO: different function of error,MSE RMSE -->
$\bar {\bm{e}}_\delta \leftarrow \text{mean}(\tilde {\bm{e}}_{i+1,\delta})$, for $i = n:T-1$  
$\delta^* \leftarrow \arg\min_{\delta \in \{0, x, 2x, \dots, 1\}} \bar {\bm{e}}_\delta$  
Compute $\lambda^*$ using $\delta^*$ on all training data  
Compute $\hat{W}_1^*$ using $\delta^*$ and $\lambda^*$ on all training data

**Return** $\delta^*, \hat{W}_1^*$

<!-- 
We only fit the model once, on all available T by p data.
The fitted model will then compute in-sample 1-step-ahead base forecasts y^
Then compute in-sample 1-step-ahead base forecast error e^
-->

### 2025-04-24 literature/paper_draft.Rmd
writing simulation and literature review

--- 2025-04-23 literature/paper-outline.Rmd
@@
 ---
-title: "Outline - Covariance Estimation for Hierarchical/Grouped Time Series Using the NOVELIST Estimator"
+title: "Draft - Enhancing Forecast Reconciliation: A Study of Covariance Estimation"
 author: "Vincent Su"
@@
 To mitigate this issue, various forecast reconciliation approaches have been proposed in the literature, where the individual forecasts are adjusted by minimizing the mean squared reconciliation errors.
-Among these, the MinT (Minimum Trace) approach is widely used.
-However, this method requires an estimate of the covariance matrix of the base forecast errors.
-Although the sample covariance matrix is a natural choice, its performance is greatly impacted by the high-dimensional nature of hierarchical time series encountered in practice.
+Among these, the MinT (Minimum Trace) approach is widely used, and this method requires an estimate of the covariance matrix of the base forecast errors.
+
+<!-- Although the sample covariance matrix is a natural choice, its performance is greatly impacted by the high-dimensional nature of hierarchical time series encountered in practice.
 Most published research uses shrinkage-type estimators that shrink the sample covariance matrix toward a diagonal matrix.
+-->
 
-In this project, we aim to assess the forecasting performance of MinT when different covariance estimators are used, with a primary focus on the NOVELIST (NOVEL Integration of the Sample and Thresholded Covariance) estimator, which shrinks the sample covariance toward its thresholded version.
+In this project, we aim to assess the forecasting performance of MinT when different covariance estimators are used, and compare with the proposed estimators from published paper.
+with a primary focus on the NOVELIST (NOVEL Integration of the Sample and Thresholded Covariance) estimator, which shrinks the sample covariance toward its thresholded version.
 A thorough literature review will also identify other high-dimensional covariance estimators relevant to this project.
@@
 
--   **Problem Statement:**\
+-   **Problem Statement:**
+
     The sample covariance matrix, although natural, suffers in high-dimensional settings.
@@
 
--   **Research Aim:**\
+-   **Research Aim:**
+
     This paper assesses the reconciled forecasting performance of MinT approach using various covariance estimators, with a focus on the NOVELIST estimator.
 
--   **Paper Outline:**\
+-   **Paper Outline:**
+
     The paper is structured as follows:
@@
 
--   Brief history of reconciliation methods (bottom-up, top-down, GLS, MinT).
--   Statistical challenges and practical applications of MinT.
+Forecast reconciliation converts a collection of independent base forecasts into a set of coherent forecasts that respect the linear constraints defining a hierarchical or grouped time-series system.
+Early work focused on heuristic single-level strategies, including bottom-up, top-down, and middle-out (...), each of which exploits only part of the information in the hierarchy and can induce bias or high variance.
+
+- Cite the single level approach
+- Athanasopoulos et al., 2024
+
+Hyndman et al.
+(2011) first showed that all single-level methods can be written as $\tilde{y} = SG\hat{y}$, where $S$ is the summing matrix and $G$ is a matrix that maps base forecasts $\hat{y}$ to into the bottom level.
+Treating reconciliation as a GLS regression problem, Hyndman et al.
+(2011) found that it yields a solution for $G$, but the required covariance matrix of reconcilation error is not identifiable in practice (Wickramasuriya et al., 2019).
+
+- (Talk more about how others transform it to OLS, WLS,..)
+- Di Fonzo and Marini (2011)
+- Athanasopoulos et al. (2009)
+
+Wickramasuriya, Athanasopoulos & Hyndman (2019) reframed the problem by taking an optimisation approach rather than the regression.
+They formulated the problem as minimising the variances of all reconciled forecasts, which happens to be equivalent to minimising the trace of the covariance matrix (sum of the diagonal elements).
+This is known as the Minimum Trace (MinT) reconciliation method.
+The MinT solution is given by $G_h = (S'W_h^{-1}S)^{-1}S'W_h^{-1}$, and $W_h$ is the covariance matrix of the h-step-ahead base forecast errors.
+
+The MinT approach is an algebraical generalisation of the GLS, and the OLS and WLS methods are special cases of MinT when $W_h$ is a diagonal or identity matrix, respectively.
+However, the MinT solution hinges on a reliable estimate of the h-step-ahead base forecast error covariance $W_h$.
+In high-dimensional setting, the usual sample covariance matrix is unstable, thus we need alternative covariance estimators.
+
+- Structural Scaling, based only on the struc- ture of the hierarchy (Athanasopoulos et al., 2017)
+- Shrinkage estimators (Schäfer & Strimmer, 2005; Ledoit & Wolf, 2004)
+
+Empirical evaluations have demonstrated that MinT with an appropriate covariance estimate often outperforms earlier methods in both simulation and real data studies.
+
+<!-- Wickramasuriya et al.
+(2019) adopt the diagonal-target Ledoit–Wolf approach and show analytically that any scalar multiple $k_h$ from $W_h = k_hW_1$ suffices for point-forecast reconciliation.
+However, simulation and empirical evidence suggest that ignoring cross-series error correlation can leave accuracy untapped, especially in highly-coupled systems (energy, retail SKUs).
+Their subsequent work imposes non-negativity constraints via quadratic programming citeturn0file2 and studies bias-robust variants (MinT-U).
+-->
+
+<!-- Empirical findings to date   -->
+<!-- Athanasopoulos et al.
+(2024) synthesise the accumulated evidence: MinT dominates single-level baselines in most studies; WLS or diagonal-shrinkage MinT is the current industry default; but gains plateau when off-diagonals are weak citeturn0file9.
+Only a handful of papers explore richer covariance structures (e.g., block or factor shrinkage), and none benchmark them systematically against NOVELIST-type estimators.
+-->
 
@@
 
--   $W_h$ is challenging to estimate, when $h$ \> 1, we need alternative estimates.
+-   $W_h$ is challenging to estimate, when $h>1$, we need alternative estimates.
 

### 2025-04-28 literature/paper_draft.Rmd
small fixes in notation

@@
 
-Hierarchical/Grouped time series satisfy a set of linear constraints, often represented as: $$
-y_t = S b_t,
-$$ where $S$ is a summing matrix of order $m x n$ which aggregates the bottom-level series $b_t$ ($n$-vector) to the series at aggregation levels above.
-The $m$-vector $y_t$ contains all observations at time $t$
+Hierarchical/Grouped time series satisfy a set of linear constraints, often represented as: 
+$$
+\bm{y}_t = \bm{S} \bm{b}_t,
+$$ 
+where $\bm{S}$ is a summing matrix of order $m x n$ which aggregates the bottom-level series $\bm{b}_t$ ($n$-vector) to the series at aggregation levels above.
+The $m$-vector $\bm{y}_t$ contains all observations at time $t$
 

### 2025-04-28 research_proposal/research_proposal.qmd
writing research proposal

---
title: "Enhancing Forecast Reconciliation: A Study of Alternative Covariance Estimators"
format: 
  pdf:
    include-in-header: 
      text: |
        \usepackage{bm}
---

## Objectives

When forecasting sales for items in a cafe, such as matcha latte and mocha, the forecasts for each of these drinks (110 matcha lattes and 90 mochas) are often not consistent with the overall sales forecast for the cafe (180 drinks). This is an example of hierarchical time series, where the individual forecasts do not satisfy the linear constraints of different levels of aggregation. Forecast reconciliation comes in to solve this problem, where the individual forecasts are adjusted to satisfy the linear constraints. Among the various reconciliation methods, the **MinT** (Minimum Trace) is considered the optimal approach. However, this method requires an estimate of the covariance matrix of the base forecast errors. In this project, we aim to assess the forecasting performance of MinT when different covariance estimators are used, namely NOVELIST (NOVEL Integration of the Sample and Thresholded Covariance), POET (Principal Orthogonal complEment Thresholding), and others.

## Background

In time series forecasting, aggregation occurs in a variety of settings. A concrete example of a hierarchy would be electricity demand forecasting, where the national demand is the sum of the demands for each state, and demand for each state comes from many regions within the states. Forecasting national tourism or Gross Domestic Product (GDP) is another example of hierarchical/grouped time series. The impact of methods for forecasting hierarchical time series has not been limited to academia, with industry also showing a strong interest. Many companies have adopted these methods in practice, including Amazon, the International Monetary Fund, IBM, SAP, and more. (Athanasopolous, 2024)

The hierarchical structure can be represented as a tree, as shown in figure 1. The top level of the tree represents the total forecast, while the lower levels represent the individual forecasts. When there are attributes of interest that are crossed, such as the forecast for electricity demand can be broken down by usage purposes (e.g., residential and commercial), it will become a grouped time series.

![**Fig. 1.** Diagram of 2-level hierarchical tree structure](D:/Github/Recon_Honours_Thesis/research_proposal/figs/hierarchical_structure.png)

Both of these structure can be represented using matrix algebra:

$$
\bm{y}_t = \bm{S} \bm{b}_t,
$$

where $\bm{S}$ is a summing matrix of order $n /times n_b$ which aggregates the bottom-level series $\bm{b}_t$ ($n_b$-vector) to the series at aggregation levels above. The $n$-vector $\bm{y}_t$ contains all observations at time $t$.

## Methodology

Hierarchical Structure

MinT

Covariance Estimators

-   Shrinkage
-   Novelist
-   POET
-   Novelist with PC-adjusted

## Experimental Design

## Timeline & Milestones

## Expected Contributions

## REferences

### 2025-04-28 research_proposal/research_proposal.qmd
fix YAML header

@@
   pdf:
-    include-in-header: 
-      text: |
-        \usepackage{bm}
+    # include-in-header: 
+    #   text: |
+    #     \usepackage{bm}
+    #     \usepackage{amsmath}
+    mathspec: true
+    number-sections: true
 ---

### 2025-04-28 research_proposal/research_proposal.qmd
refine

@@
 When forecasting sales for items in a cafe, such as matcha latte and mocha, the forecasts for each of these drinks (110 matcha lattes and 90 mochas) are often not consistent with the overall sales forecast for the cafe (180 drinks).
-This is an example of hierarchical time series, where the individual forecasts do not satisfy the linear constraints of different levels of aggregation.
-Forecast reconciliation comes in to solve this problem, where the individual forecasts are adjusted to satisfy the linear constraints.
+This is a problem of forecasting hierarchical time series, where the individual forecasts do not satisfy the linear constraints of different levels of aggregation.
+Forecast reconciliation comes in to solve this problem, where the individual forecasts are adjusted to satisfy the given constraints.
 Among the various reconciliation methods, the **MinT** (Minimum Trace) is considered the optimal approach.
 However, this method requires an estimate of the covariance matrix of the base forecast errors.
+The current practice is to use the shrinkage estimator (often shrinking toward a diagonal matrix), but it lacks flexibility and might neglect the prominent structure presented.
 In this project, we aim to assess the forecasting performance of MinT when different covariance estimators are used, namely NOVELIST (NOVEL Integration of the Sample and Thresholded Covariance), POET (Principal Orthogonal complEment Thresholding), and others.
@@
 
-![**Fig. 1.** Diagram of 2-level hierarchical tree structure](D:/Github/Recon_Honours_Thesis/research_proposal/figs/hierarchical_structure.png)
+![Diagram of 2-level hierarchical tree structure](D:/Github/Recon_Honours_Thesis/research_proposal/figs/hierarchical_structure.png){width="40%"}
 

### 2025-04-28 research_proposal/research_proposal.qmd
write up background knowledge

@@
 $$
-\bm{y}_t = \bm{S} \bm{b}_t,
+\boldsymbol{y}_t = \boldsymbol{S} \boldsymbol{b}_t,
+$$ {#eq-1}
+
+where $\boldsymbol{S}$ is a summing matrix of order $n \times n_b$ which aggregates the bottom-level series $\boldsymbol{b}_t$ ($n_b$-vector) to the series at aggregation levels above.
+The $n$-vector $\boldsymbol{y}_t$ contains all observations at time $t$.
+
+The example summing matrix $\boldsymbol{S}$ for the tree structure in *Figure 1* is:
+
+$$
+\boldsymbol{S} = 
+\left[
+\begin{array}{cccc}
+1 & 1 & 1 & 1 \\
+1 & 1 & 0 & 0 \\
+0 & 0 & 1 & 1 \\
+\multicolumn{4}{c}{\boldsymbol{I_n}}
+\end{array}
+\right],
 $$
 
-where $\bm{S}$ is a summing matrix of order $n /times n_b$ which aggregates the bottom-level series $\bm{b}_t$ ($n_b$-vector) to the series at aggregation levels above.
-The $n$-vector $\bm{y}_t$ contains all observations at time $t$.
+The MinT method is a popular choice for estimating $\boldsymbol{G}_h$, as it minimizes the trace of the covariance matrix of the forecast errors, which is a measure of the uncertainty in the forecasts.
+The MinT method requires an estimate of the covariance matrix of the base forecast errors, which is typically obtained using a shrinkage estimator.
+However, this approach has limitations, as it may not fully capture the structure of the data.
 

### 2025-04-28 research_proposal/research_proposal.qmd
min trace

@@
 
-Hierarchical Structure
+If we use equation @eq-1 to forecast the time series, we would not be utilising all the information in the data, since we only forecast the bottom-level series $\boldsymbol{b}_t$ then aggregate to the higher-level.
+Thus, Hyndman et al.
+(2011) showed that existing methods could be expressed as:
 
-MinT
+$$
+\tilde{\boldsymbol{y}}_h = \boldsymbol{S} \boldsymbol{G}_h \hat{\boldsymbol{y}}_h,
+$$ {#eq-2}
+
+for a suitably $n_b \times n$ matrix $\boldsymbol{G}_h$ (we can drop the subscript $h$ when $G$ does not depend on the forecast horizon $h$).
+$\boldsymbol{G}_h$ maps the base forecasts of all levels $\hat{\boldsymbol{y}}_h$ down into the bottom series, which is then aggregated to the higher levels by $\boldsymbol{S}$.
+Note that any method may have been used to produce the base forecasts.
+
+From this we can see the importance of the matrix $\boldsymbol{G}_h$, and the choice of it determines the performance of reconciled forecasts $\tilde{\boldsymbol{y}}_h$.
+Methods are developed to estimate $\boldsymbol{G}_h$, including the OLS and WLS, proposed by Hyndman et al.
+(2011) and Hyndman et al.
+(2016) respectively.
+
+### Mininum Trace (MinT) Reconciliation
+
+Wickramasuriya et al.
+(2019) reframed the problem by taking an optimisation approach rather than the regression.
+They formulated the problem as minimising the variances of all reconciled forecasts from equation @eq-2, which happens to be equivalent to the trace of the covariance matrix (sum of the diagonal elements).
+This is known as the Minimum Trace (MinT) reconciliation method.
+The MinT solution is given by $G_h = (S'W_h^{-1}S)^{-1}S'W_h^{-1}$, and $W_h$ is the covariance matrix of the h-step-ahead base forecast errors.
+
+The MinT approach is an algebraical generalisation of the GLS, and the OLS and WLS methods are special cases of MinT when $W_h$ is a diagonal or identity matrix, respectively.
+However, the MinT solution hinges on a reliable estimate of the h-step-ahead base forecast error covariance $W_h$.
+In high-dimensional setting, the usual sample covariance matrix is unstable, thus we need alternative covariance estimators.
+
+### Alternative Covariance Estimators
+
+***Shrinkage***
 

### 2025-04-28 research_proposal/research_proposal.qmd
write up the shrinkage

@@
     #   text: |
-    #     \usepackage{bm}
     #     \usepackage{amsmath}
+    #     \usepackage{mathspec}
+    #     \newcommand{\b}[1]{\boldsymbol{#1}}
     mathspec: true
@@
 
-The MinT approach is an algebraical generalisation of the GLS, and the OLS and WLS methods are special cases of MinT when $W_h$ is a diagonal or identity matrix, respectively.
-However, the MinT solution hinges on a reliable estimate of the h-step-ahead base forecast error covariance $W_h$.
+$$
+\boldsymbol{G}_h = (\boldsymbol{S}' \boldsymbol{W}_h^{-1} \boldsymbol{S})^{-1}
+\boldsymbol{S}' \boldsymbol{W}_h^{-1}
+$$
+
+and $\boldsymbol{W}_h$ is the covariance matrix of the h-step-ahead base forecast errors.
+
+The MinT approach is an algebraical generalisation of the GLS, and the OLS and WLS methods are special cases of MinT when $\boldsymbol{W}_h$ is a diagonal or identity matrix, respectively.
+However, the MinT solution hinges on a reliable estimate of $\boldsymbol{W}_h$.
 In high-dimensional setting, the usual sample covariance matrix is unstable, thus we need alternative covariance estimators.
@@
 
-Covariance Estimators
+The proposed MinT approach by Wickramasuriya et al. (2019) uses the shrinkage estimator from Schäfer and Strimmer (2005). The shrinkage estimator is given by:
 
--   Shrinkage
+$$
+\hat{\boldsymbol{W}}^{shr}_{1, D} = \lambda_D \hat{\boldsymbol{W}}_{1, D} + (1 - \lambda_D) \hat{\boldsymbol{W}}_1
+$$
+
+and $\hat{\boldsymbol{W}}_1 = \frac{1}{T} \sum_{t=1}^{T} \hat{\boldsymbol{e}}_{t|t-1} \hat{\boldsymbol{e}}_{t|t-1}'$ is the unbiased sample covariance of in-sample one-step-ahead base forecast errors.
+
+$\hat{\boldsymbol{W}}_{1, D}$ is a diagonal matrix comprising the diagonal entries of $\hat{\boldsymbol{W}}_1$.
+This approach will shrink the covariance matrix $\hat{\boldsymbol{W}}_1$ towards its diagonal version, meaning the off-diagonal elements are shrunk towards zero while the diagonal ones remain unchanged.
+
+Schäfer and Strimmer (2005) also proposed an optimal shrinkage intensity parameter $\lambda_D$ for this setting, assuming the variances are constant:
+
+$$
+\hat{\lambda}_D = \frac{\sum_{i \neq j} \widehat{Var}(\hat{r}_{ij})} {\sum_{i \neq j} \hat{r}_{ij}}
+$$
+
+where $\hat{r}_{ij}$ is the $ij$th element of $\hat{\boldsymbol{R}}_1$, the 1-step-ahead sample correlation matrix (obtained from $\hat{\boldsymbol{W}}_1$) to shrink it toward an identity matrix
+
 -   Novelist

### 2025-04-28 literature/paper_draft.Rmd
cleanup

@@
 
-For forecast errors $\hat{e}_{t|t-1}$, the unbiased sample covariance of in-sample one-step-ahead base forecast errors is: $$
+For forecast errors $\hat{e}_{t|t-1}$, the unbiased sample covariance of in-sample one-step-ahead base forecast errors is:
+
+$$
 \hat{W}_1 = \frac{1}{T} \sum_{t=1}^{T} \hat{e}_{t|t-1} \hat{e}_{t|t-1}^{\top}.
@@
 $$
+
+where $\hat{r}_{ij}$ is the $ij$th element of $\hat{\boldsymbol{R}}_1$, the 1-step-ahead sample correlation matrix (obtained from $\hat{\boldsymbol{W}}_1$) to shrink it toward an identity matrix
 

### 2025-04-28 research_proposal/research_proposal.qmd
cleanup

@@
 
-![Diagram of 2-level hierarchical tree structure](D:/Github/Recon_Honours_Thesis/research_proposal/figs/hierarchical_structure.png){width="40%"}
+![Diagram of 2-level hierarchical tree structure](D:/Github/Recon_Honours_Thesis/research_proposal/figs/hierarchical_structure.png){#fig-hierarchy width="40%"}
 
@@
 
-The example summing matrix $\boldsymbol{S}$ for the tree structure in *Figure 1* is:
+The example summing matrix $\boldsymbol{S}$ for the tree structure in @fig-hierarchy is:
 
@@
 
-If we use equation @eq-1 to forecast the time series, we would not be utilising all the information in the data, since we only forecast the bottom-level series $\boldsymbol{b}_t$ then aggregate to the higher-level.
+If we use @eq-1 to forecast the time series, we would not be utilising all the information in the data, since we only forecast the bottom-level series $\boldsymbol{b}_t$ then aggregate to the higher-level.
 Thus, Hyndman et al.
@@
 (2019) reframed the problem by taking an optimisation approach rather than the regression.
-They formulated the problem as minimising the variances of all reconciled forecasts from equation @eq-2, which happens to be equivalent to the trace of the covariance matrix (sum of the diagonal elements).
+They formulated the problem as minimising the variances of all reconciled forecasts from @eq-2, which happens to be equivalent to the trace of the covariance matrix (sum of the diagonal elements).
 This is known as the Minimum Trace (MinT) reconciliation method.
-The MinT solution is given by $G_h = (S'W_h^{-1}S)^{-1}S'W_h^{-1}$, and $W_h$ is the covariance matrix of the h-step-ahead base forecast errors.
+The MinT solution is given by
 

### 2025-04-29 research_proposal/research_proposal.qmd
expand on the link to alternative estimators

@@
 The MinT approach is an algebraical generalisation of the GLS, and the OLS and WLS methods are special cases of MinT when $\boldsymbol{W}_h$ is a diagonal or identity matrix, respectively.
-However, the MinT solution hinges on a reliable estimate of $\boldsymbol{W}_h$.
-In high-dimensional setting, the usual sample covariance matrix is unstable, thus we need alternative covariance estimators.
+However, the MinT solution hinges on a reliable estimate of $\boldsymbol{W}_h$, which is challenging to estimate in high-dimensional setting.
+Therefore, we need alternative covariance estimators.
 
 ### Alternative Covariance Estimators
+
+We reconstruct estimator of $\boldsymbol{W}_h$ as $\hat{\boldsymbol{W}}_h = k_h \, g(\hat{\boldsymbol{W}}_1)$, and $g(.)$ is an estimator function of the unbiased sample covariance of in-sample one-step-ahead base forecast errors $\hat{\boldsymbol{W}}_1 = \frac{1}{T} \sum_{t=1}^{T} \hat{\boldsymbol{e}}_{t|t-1} \hat{\boldsymbol{e}}_{t|t-1}'$
 
@@
 $$
-
-and $\hat{\boldsymbol{W}}_1 = \frac{1}{T} \sum_{t=1}^{T} \hat{\boldsymbol{e}}_{t|t-1} \hat{\boldsymbol{e}}_{t|t-1}'$ is the unbiased sample covariance of in-sample one-step-ahead base forecast errors.
 

### 2025-04-29 research_proposal/research_proposal.qmd
shortcoming of shrinkage and intro to NOVELIST

@@
 
--   Novelist
+The optimal lambda is obtained by minimising the $MSE(\hat{\boldsymbol{W}}_1) = Bias(\hat{\boldsymbol{W}}_1)^2 + Var(\hat{\boldsymbol{W}}_1)$.
+More specifically, we trade the unbiasedness of the sample covariance matrix for a lower variance.
+The objective function itself does not take into account any possible principal components structure in the data, and is not flexible enough since it shrinks all off-diagonal elements equally towards zeros.
+
+***NOVELIST***
+
+The NOVELIST (NOVEL Integration of the Sample and Thresholded Covariance) estimator, proposed by Huang & Fryzlewicz (2019), is currently the main focus of this research project.
+It is based on the idea of soft-thresholding the sample covariance matrix, then performing shrinkage towards this thresholded version.
+This introduces an extra parameter, the threshold $\delta$, which is used to control the amount of soft-thresholding, offering more flexibility.
+The NOVELIST estimator is given by:
+
 -   POET

### 2025-04-29 research_proposal/research_proposal.qmd
novelist approach

@@
 
-for a suitably $n_b \times n$ matrix $\boldsymbol{G}_h$ (we can drop the subscript $h$ when $G$ does not depend on the forecast horizon $h$).
+for a suitable $n_b \times n$ matrix $\boldsymbol{G}_h$ (we can drop the subscript $h$ when $G$ does not depend on the forecast horizon $h$).
 $\boldsymbol{G}_h$ maps the base forecasts of all levels $\hat{\boldsymbol{y}}_h$ down into the bottom series, which is then aggregated to the higher levels by $\boldsymbol{S}$.
@@
 
-***Shrinkage***
+***(a) Shrinkage***
 
@@
 
-***NOVELIST***
+***(b) NOVELIST***
 
@@
 The NOVELIST estimator is given by:
+
+$$
+\hat{\boldsymbol{W}}^{N}_{1} = \lambda_{\delta} \hat{\boldsymbol{W}}_{1, \delta} + (1 - \lambda_{\delta}) \hat{\boldsymbol{W}}_1
+$$
+
+By convenient setting, we rewrite it as in correlation matrix:
+
+$$
+\hat{\boldsymbol{R}}^{N}_{1} = \lambda_{\delta} \hat{\boldsymbol{R}}_{1,\delta} + (1 - \lambda_{\delta}) \hat{\boldsymbol{R}}_1,
+$$
+
+In this setting, $\hat{\boldsymbol{R}}_{1,\delta}$ is the thresholded sample correlation matrix, where each element is regularised by:
+
+$$
+\hat{r}_{1,ij}^\delta = \operatorname{sign}(\hat{r}_{1,ij}) \, (|\hat{r}_{1,ij}| - \delta)_+.
+$$
+
+For a given threshold $\delta$, Huang & Fryzlewicz derived the optimal shrinkage intensity parameter $\lambda(\delta)$ using Ledoit-Wolf's lemma (Ledoit and Wolf, 2003). It can be computed as:
+
+$$
+\hat{\lambda}(\delta) = \frac{
+  \sum_{i \neq j} \widehat{Var}(\hat{r}_{1,ij}) \; I(|\hat{r}_{1,ij}| \leq \delta)
+} {
+  \sum_{i \neq j} (\hat{r}_{1,ij} - \hat{r}_{1,ij}^\delta)^2
+}
+$$
+
+On the other hand, the optimal threshold $\delta^*$ does not have a closed-form solution, and is typically obtained by executing a rolling-window cross-validation procedure.
+The idea is to, for each rolling-window set, loop through a range of threshold values, compute the corresponding $\hat{\lambda}(\delta)$ and $\hat{\boldsymbol{R}}^{N}_{1}$.
+Then we select the one that minimises the average out-of-sample reconciled forecast errors.
+Although it is not required to fit forecasting models multiple time, the cross-validation procedure is still computationally expensive as it computes the NOVELIST estimator and perform reconciliation for each threshold value.
+The formal algorithm is given in @pseudo_novelist_cv, in the Appendix section.
 

### 2025-04-29 research_proposal/research_proposal.qmd
poet and novelist-pc approach

@@
 
--   POET
--   Novelist with PC-adjusted
+***(c) POET***
+
+The POET (Principal Orthogonal complEment Thresholding) estimator, proposed by Fan et al.
+(2013), is another "sparse" + "non-sparse" covariance estimator.
+It starts by decomposing the correlation matrix $\hat{\boldsymbol{R}}_1$ into a prominent principle components part (low-rank) and a orthogonal complement part $\hat{\boldsymbol{R}}_{1, K}$ (the correlation matrix after removing the first $K$ principal components).
+Then it applies thresholding to $\hat{\boldsymbol{R}}_{1, K}$.
+The POET estimator is given by:
+
+$$
+\hat{\boldsymbol{R}}^{K}_{1} = \sum_{k = 1}^K \hat{\gamma}_k \hat{\boldsymbol{\xi}}_k \hat{\boldsymbol{\xi}}_k' + T(\hat{\boldsymbol{R}}_{1, K})
+$$
+
+where $\hat{\gamma}_k$ and $\hat{\boldsymbol{\xi}}_k$ are the $k$th eigenvalue and eigenvector of the sample covariance matrix.
+$T(.)$ is the thresholding function, which can be either soft-thresholding, hard-thresholding, or others.
+
+***(d) NOVELIST with PC-adjusted***
+
+This approach is similar to the POET estimator.
+The difference is that POET apply thresholding to the orthogonal complement part $\hat{\boldsymbol{R}}_{1, K}$, meanwhile we apply NOVELIST.
+It can be formulated as:
+
+$$
+\hat{\boldsymbol{R}}^{N, K}_{1} = \sum_{k = 1}^K \hat{\gamma}_k \hat{\boldsymbol{\xi}}_k \hat{\boldsymbol{\xi}}_k' + \hat{\boldsymbol{R}}^{N}_{1, K}
+$$
+
+and $\hat{\boldsymbol{R}}^{N}_{1, K}$ is the NOVELIST estimator applied to the orthogonal complement part $\hat{\boldsymbol{R}}_{1, K}$.
 

### 2025-04-29 research_proposal/research_proposal_draft.qmd
rename

--- 2025-04-29 research_proposal/research_proposal.qmd
@@
 
+The experimental design is to simulate a hierarchical time series data set, then apply the MinT reconciliation method with different covariance estimators.
+The data set will be split into training and test sets.
+In case of cross-validation, the training set will be further split into training and validation sets.
+
+          --------- put these into appendix -----------
+
+The designed data generating process for bottom-level series is a stationary VAR(1) process, with the following structure:
+
+$$
+\boldsymbol{y}_t = \boldsymbol{A} \boldsymbol{y}_{t-1} + \boldsymbol{e}_t,
+$$
+
+where $\boldsymbol{A}$ is a $p \times p$ block diagonal matrix of autoregressive coefficients $\boldsymbol{A} = diag(\boldsymbol{A}_1, \ldots, \boldsymbol{A}_m)$, with each $\boldsymbol{A}_i$ being a $p_i \times p_i$ matrix.
+The block diagonal structure ensures that the time series are grouped into $m$ groups, with each group having its own autoregressive coefficients.
+This aim to simulate the interdependencies between the time series within each group, where reconciliation will be better performed than the usual base forecasts.
+
+The model is added with a Gaussian innovation process $\boldsymbol{e}_t$, with covariance matrix $\Sigma$. The covariance matrix $\Sigma$ is generated specifically in the following way:
+
+1. A compound symmetric correlation matrix is used for each block of size $p_i$ in $\boldsymbol{A}_i$, where the coefficients are sampled from a uniform distribution.
+
+2. The correlations between different blocks are imposed using the Algorithm 1 in Hardin, Garcia & Golan (2013).
+
+3. The covariance matrix $\Sigma$ is then constructed by uniform sampling of standard deviations for all $p$ series.
+
+We have an option to randomly flip the signs of the covariance elements, which will create a more realistic structure in the innovation process.
+This is also to simulate the real-world scenario where the observed covariance matrix is not necessarily positive definite.
+
+              ----------------------------------------
+
+- Put some graphs here
+
 ## Timeline & Milestones
@@
 
-## REferences
+## References
+
+## Appendix
+
+Algorithm 1

### 2025-04-30 research_proposal/research_proposal_draft.qmd
add algorithm in appendix

@@
   pdf:
-    # include-in-header:
-    #   text: |
-    #     \usepackage{amsmath}
-    #     \usepackage{mathspec}
-    #     \newcommand{\b}[1]{\boldsymbol{#1}}
+    include-in-header:
+      text: |
+        <!-- \usepackage{amsmath} -->
+        <!-- \usepackage{mathspec} -->
+        <!-- \newcommand{\b}[1]{\boldsymbol{#1}} -->
+        \usepackage{algorithm}
+        \usepackage{algpseudocode}
     mathspec: true
@@
 Although it is not required to fit forecasting models multiple time, the cross-validation procedure is still computationally expensive as it computes the NOVELIST estimator and perform reconciliation for each threshold value.
-The formal algorithm is given in @pseudo_novelist_cv, in the Appendix section.
+The formal algorithm is given in @fig-novelist_cv, in the @sec-appendix Appendix.
 
@@
 
-## Appendix
+\pagebreak
 
-Algorithm 1
+## Appendix {#sec-appendix}
+
+#### Algorithm: NOVELIST cross-validation for optimal threshold $\delta^*$ {-}
+
+![NOVELIST cross-validation procedure](D:/Github/Recon_Honours_Thesis/research_proposal/figs/novelist_cv_algo.png){#fig-novelist_cv height=70%}
+
+#### Data generating progress design {-}
+
+<!-- ```{=latex} -->
+<!-- \begin{algorithm} -->
+<!-- \caption{Cross-validation procedure to select optimal threshold $\delta^*$ for NOVELIST} -->
+<!-- \begin{algorithmic}[1] -->
+<!-- \State \textbf{Input:} Residual matrix $E$ of dimension $T \times p$, grid of threshold candidates $\{\delta_j\}$, window size $n$ -->
+<!-- \For{each threshold candidate $\delta_j$} -->
+<!--     \For{$t = n+1$ to $T$} -->
+<!--         \State Define in-sample window: $E_{1:n} = \{e_{t-n}, ..., e_{t-1}\}$ -->
+<!--         \State Compute sample covariance $\hat{\Sigma}_t = \frac{1}{n} E_{1:n}^\top E_{1:n}$ -->
+<!--         \State Compute thresholded covariance: $\hat{\Sigma}^{\text{thr}}_t = \text{Threshold}(\hat{\Sigma}_t, \delta_j)$ -->
+<!--         \State Compute NOVELIST: $\hat{\Sigma}^{\text{nov}}_t = (1 - \lambda) \hat{\Sigma}_t + \lambda \hat{\Sigma}^{\text{thr}}_t$ -->
+<!--         \State Compute inverse: $W_t = (\hat{\Sigma}^{\text{nov}}_t)^{-1}$ -->
+<!--         \State Compute MinT reconciliation matrix: $P_t = (S^\top W_t S)^{-1} S^\top W_t$ -->
+<!--         \State Compute forecast error: $e_t = y_t - S P_t \hat{y}_t$ -->
+<!--     \EndFor -->
+<!--     \State Compute total CV loss $L_j = \sum_t \|e_t\|^2$ -->
+<!-- \EndFor -->
+<!-- \State \textbf{Output:} $\delta^* = \arg\min_{\delta_j} L_j$ -->
+<!-- \end{algorithmic} -->
+<!-- \end{algorithm} -->
+<!-- ``` -->

### 2025-04-30 research_proposal/research_proposal_draft.qmd
Update research_proposal_draft.qmd

@@
 Although it is not required to fit forecasting models multiple time, the cross-validation procedure is still computationally expensive as it computes the NOVELIST estimator and perform reconciliation for each threshold value.
-The formal algorithm is given in @fig-novelist_cv, in the @sec-appendix Appendix.
+The formal algorithm is given in the @sec-novelist_cv Appendix.
 
@@
 
-#### Algorithm: NOVELIST cross-validation for optimal threshold $\delta^*$ {-}
+### Algorithm: NOVELIST cross-validation for optimal threshold $\delta^*$ {#sec-novelist_cv}
 
@@
 
-#### Data generating progress design {-}
+### Data generating progress design {#sec-dgp}
 

### 2025-04-30 research_proposal/research_proposal_draft.qmd
adding timeline and working on referencing

@@
         <!-- \usepackage{mathspec} -->
-        <!-- \newcommand{\b}[1]{\boldsymbol{#1}} -->
-        \usepackage{algorithm}
-        \usepackage{algpseudocode}
+        <!-- \usepackage{algorithm} -->
+        <!-- \usepackage{algpseudocode} -->
+        \usepackage[colorlinks=true, linkcolor=blue, citecolor=blue, urlcolor=blue]{hyperref}
     mathspec: true
     number-sections: true
+execute:
+  cache: true
+bibliography: D:/Github/Recon_Honours_Thesis/references.bib
+# bibliographystyle: apa
+csl: D:/Github/Recon_Honours_Thesis/apa.csl
 ---
 
-## Objectives
+# Objectives
 
@@
 
-## Background
+# Background
 
@@
 
-The MinT method is a popular choice for estimating $\boldsymbol{G}_h$, as it minimizes the trace of the covariance matrix of the forecast errors, which is a measure of the uncertainty in the forecasts.
-The MinT method requires an estimate of the covariance matrix of the base forecast errors, which is typically obtained using a shrinkage estimator.
-However, this approach has limitations, as it may not fully capture the structure of the data.
-
-## Methodology
+# Methodology
 
@@
 
-From this we can see the importance of the matrix $\boldsymbol{G}_h$, and the choice of it determines the performance of reconciled forecasts $\tilde{\boldsymbol{y}}_h$.
+The choice of $\boldsymbol{G}_h$ determines the performance of reconciled forecasts $\tilde{\boldsymbol{y}}_h$.
 Methods are developed to estimate $\boldsymbol{G}_h$, including the OLS and WLS, proposed by Hyndman et al.
@@
 
-### Mininum Trace (MinT) Reconciliation
+## Mininum Trace (MinT) Reconciliation
 
-Wickramasuriya et al.
-(2019) reframed the problem by taking an optimisation approach rather than the regression.
+@Wickramasuriya2019-xq reframed the problem by taking an optimisation approach rather than the regression.
 They formulated the problem as minimising the variances of all reconciled forecasts from @eq-2, which happens to be equivalent to the trace of the covariance matrix (sum of the diagonal elements).
@@
 
-The MinT approach is an algebraical generalisation of the GLS, and the OLS and WLS methods are special cases of MinT when $\boldsymbol{W}_h$ is a diagonal or identity matrix, respectively.
+Wickramasuriya et al.
+(2019) also showed that MinT is an algebraical generalisation of the GLS, and the OLS and WLS methods are special cases of MinT when $\boldsymbol{W}_h$ is a diagonal or identity matrix, respectively.
 However, the MinT solution hinges on a reliable estimate of $\boldsymbol{W}_h$, which is challenging to estimate in high-dimensional setting.
@@
 
-### Alternative Covariance Estimators
+## Alternative Covariance Estimators
 
@@
 
-## Experimental Design
+# Experimental Design
 
@@
 In case of cross-validation, the training set will be further split into training and validation sets.
+The data generating process is described in @sec-dgp Appendix.
 
-          --------- put these into appendix -----------
+- Put some graphs here
+
+# Timeline & Milestones
+
+```{r timeline}
+#| echo: false
+#| message: false
+#| warning: false
+library(tibble)
+library(knitr)
+
+tibble(
+  Period = c(
+    "March - April", 
+    "May", 
+    "June - July", 
+    "August",
+    "September - October"
+  ),
+  Task = c(
+    "Literature review, methods implementation (shrinkage, NOVELIST) + simulation framework",
+    "Simulation + NOVELIST assessment",
+    "Experiments with POET and PC-adjusted NOVELIST",
+    "Real-world application + evaluation",
+    "Final writing + submission"
+  ),
+  Deliverable = c(
+    "Main codes + paper draft",
+    "Results: NOVELIST",
+    "Results: POET & PC-adjusted NOVELIST",
+    "Application completed",
+    "Thesis manuscript"
+  )
+) %>%
+  kable(caption = "Project Timeline and Key Milestones")
+```
+
+# Expected Contributions
+
+\pagebreak
+
+# Appendix {#sec-appendix}
+
+## Algorithm: NOVELIST cross-validation for optimal threshold $\delta^*$ {#sec-novelist_cv}
+
+![NOVELIST cross-validation procedure](D:/Github/Recon_Honours_Thesis/research_proposal/figs/novelist_cv_algo.png){#fig-novelist_cv height=70%}
+
+## Data generating progress design {#sec-dgp}
 
@@
 We have an option to randomly flip the signs of the covariance elements, which will create a more realistic structure in the innovation process.
-This is also to simulate the real-world scenario where the observed covariance matrix is not necessarily positive definite.
-
-              ----------------------------------------
-
-- Put some graphs here
-
-## Timeline & Milestones
-
-## Expected Contributions
-
-## References
-
-\pagebreak
-
-## Appendix {#sec-appendix}
-
-### Algorithm: NOVELIST cross-validation for optimal threshold $\delta^*$ {#sec-novelist_cv}
-
-![NOVELIST cross-validation procedure](D:/Github/Recon_Honours_Thesis/research_proposal/figs/novelist_cv_algo.png){#fig-novelist_cv height=70%}
-
-### Data generating progress design {#sec-dgp}
 
@@
 <!-- ``` -->
+
+\pagebreak
+
+# References

### 2025-05-02 research_proposal/research_proposal_draft.qmd
add vspace
refine and write more

@@
 
-When forecasting sales for items in a cafe, such as matcha latte and mocha, the forecasts for each of these drinks (110 matcha lattes and 90 mochas) are often not consistent with the overall sales forecast for the cafe (180 drinks).
-This is a problem of forecasting hierarchical time series, where the individual forecasts do not satisfy the linear constraints of different levels of aggregation.
-Forecast reconciliation comes in to solve this problem, where the individual forecasts are adjusted to satisfy the given constraints.
-Among the various reconciliation methods, the **MinT** (Minimum Trace) is considered the optimal approach.
-However, this method requires an estimate of the covariance matrix of the base forecast errors.
+\textcolor{red}{Globally, Starbucks manages thousands of outlets across dozens of countries, each projecting sales for dozens of beverages.
+Yet when regional, national, and brand-wide forecasts are aggregated, discrepancies of millions of dollars emerge simply because individual outlet forecasts fail to respect the hierarchical nature.
+Forecast reconciliation comes in to solve this huge problem, where the individual forecasts are adjusted to satisfy the aggregation constraints.} Among the various reconciliation methods, the **MinT** (Minimum Trace) is considered the optimal approach, however, it requires a good estimate of the covariance matrix of the base forecast errors.
 The current practice is to use the shrinkage estimator (often shrinking toward a diagonal matrix), but it lacks flexibility and might neglect the prominent structure presented.
@@
 A concrete example of a hierarchy would be electricity demand forecasting, where the national demand is the sum of the demands for each state, and demand for each state comes from many regions within the states.
-Forecasting national tourism or Gross Domestic Product (GDP) is another example of hierarchical/grouped time series.
+Forecasting national tourism or Gross Domestic Product (GDP) is another example of hierarchical time series.
 The impact of methods for forecasting hierarchical time series has not been limited to academia, with industry also showing a strong interest.
-Many companies have adopted these methods in practice, including Amazon, the International Monetary Fund, IBM, SAP, and more.
-(Athanasopolous, 2024)
+Many companies have adopted these methods in practice, including Amazon, the International Monetary Fund, IBM, SAP, and more [@Athanasopoulos2024-as].
 
-The hierarchical structure can be represented as a tree, as shown in figure 1.
-The top level of the tree represents the total forecast, while the lower levels represent the individual forecasts.
-When there are attributes of interest that are crossed, such as the forecast for electricity demand can be broken down by usage purposes (e.g., residential and commercial), it will become a grouped time series.
+The hierarchical structure can be represented as a tree, as shown in @fig-hierarchy.
+The top level of the tree represents the total value of all series, while the lower levels represent the individual series.
+When there are attributes of interest that are crossed, such as the electricity demand of any aggregation level (national, state, or regional) can be further broken down by usage purposes (e.g., residential, commercial), it will become a grouped time series.
 
@@
 
-Both of these structure can be represented using matrix algebra:
+We refer to both of these structures as hierarchical time series, and can be represented using matrix algebra:
 
@@
 \end{array}
-\right],
+\right]
 $$
@@
 
-If we use @eq-1 to forecast the time series, we would not be utilising all the information in the data, since we only forecast the bottom-level series $\boldsymbol{b}_t$ then aggregate to the higher-level.
-Thus, Hyndman et al.
-(2011) showed that existing methods could be expressed as:
+\textcolor{red}{If we produce base forecasts for the bottom-level series $\hat{\boldsymbol{b}}_h$ (any method may have been used to produce the base forecasts) and then reconcile them to all aggregated levels $\tilde{\boldsymbol{y}}_h$, we would not be utilising the information from those higher levels.
+This is similar to the top-down and middle-out approaches, with appropriate summing matrix $\boldsymbol{S}$.
+Another issue with expressing this way is that it restricts the reconciliation to only single-level approaches.
+Thus, @Hyndman2011-jv suggested a generalised expression for all existing reconciled methods, which also provides a framework for new methods to be developed:}
 
 $$
-\tilde{\boldsymbol{y}}_h = \boldsymbol{S} \boldsymbol{G}_h \hat{\boldsymbol{y}}_h,
+\tilde{\boldsymbol{y}}_h = \boldsymbol{S} \boldsymbol{G} \hat{\boldsymbol{y}}_h,
 $$ {#eq-2}
 
-for a suitable $n_b \times n$ matrix $\boldsymbol{G}_h$ (we can drop the subscript $h$ when $G$ does not depend on the forecast horizon $h$).
-$\boldsymbol{G}_h$ maps the base forecasts of all levels $\hat{\boldsymbol{y}}_h$ down into the bottom series, which is then aggregated to the higher levels by $\boldsymbol{S}$.
-Note that any method may have been used to produce the base forecasts.
+for a suitable $n_b \times n$ matrix $\boldsymbol{G}$.
+$\boldsymbol{G}$ maps the base forecasts of all levels $\hat{\boldsymbol{y}}_h$ down into the bottom series, which is then aggregated to the higher levels by $\boldsymbol{S}$.
 
-The choice of $\boldsymbol{G}_h$ determines the performance of reconciled forecasts $\tilde{\boldsymbol{y}}_h$.
-Methods are developed to estimate $\boldsymbol{G}_h$, including the OLS and WLS, proposed by Hyndman et al.
-(2011) and Hyndman et al.
-(2016) respectively.
+The choice of $\boldsymbol{G}$ determines the composition of reconciled forecasts $\tilde{\boldsymbol{y}}_h$.
+Methods are developed to estimate $\boldsymbol{G}$, including the OLS and WLS, proposed by @Hyndman2011-jv and @Hyndman2016-ic respectively.
+
+- expand on about OLS and WLS
 
@@
 @Wickramasuriya2019-xq reframed the problem by taking an optimisation approach rather than the regression.
-They formulated the problem as minimising the variances of all reconciled forecasts from @eq-2, which happens to be equivalent to the trace of the covariance matrix (sum of the diagonal elements).
+They formulated the problem as minimising the variances of all reconciled forecasts from @eq-2, which happens to be equivalent to minimising the trace of the covariance matrix (sum of the diagonal elements).
 This is known as the Minimum Trace (MinT) reconciliation method.
@@
 $$
-\boldsymbol{G}_h = (\boldsymbol{S}' \boldsymbol{W}_h^{-1} \boldsymbol{S})^{-1}
+\boldsymbol{G} = (\boldsymbol{S}' \boldsymbol{W}_h^{-1} \boldsymbol{S})^{-1}
 \boldsymbol{S}' \boldsymbol{W}_h^{-1}
@@
 
-Wickramasuriya et al.
-(2019) also showed that MinT is an algebraical generalisation of the GLS, and the OLS and WLS methods are special cases of MinT when $\boldsymbol{W}_h$ is a diagonal or identity matrix, respectively.
+@Wickramasuriya2019-xq also showed that MinT is an algebraical generalisation of the GLS, and the OLS and WLS methods are special cases of MinT when $\boldsymbol{W}_h$ is a diagonal or identity matrix, respectively.
 However, the MinT solution hinges on a reliable estimate of $\boldsymbol{W}_h$, which is challenging to estimate in high-dimensional setting.
-Therefore, we need alternative covariance estimators.
+Therefore, we will be exploring alternative covariance estimators.
 
@@
 
+\vspace{1em}
+
 ***(a) Shrinkage***
 
-The proposed MinT approach by Wickramasuriya et al. (2019) uses the shrinkage estimator from Schäfer and Strimmer (2005). The shrinkage estimator is given by:
+The proposed MinT approach by @Wickramasuriya2019-xq uses the shrinkage estimator from @Schafer2005-yw.
+\textcolor{red}{It guarantees positive definiteness and huge variance reduction when $p>T$}.
+The shrinkage estimator is given by:
 
@@
 
-Schäfer and Strimmer (2005) also proposed an optimal shrinkage intensity parameter $\lambda_D$ for this setting, assuming the variances are constant:
+@Schafer2005-yw also proposed an optimal shrinkage intensity parameter $\lambda_D$ for this setting, assuming the variances are constant:
 
@@
 
+\vspace{1em}
+
 ***(b) NOVELIST***
 
-The NOVELIST (NOVEL Integration of the Sample and Thresholded Covariance) estimator, proposed by Huang & Fryzlewicz (2019), is currently the main focus of this research project.
-It is based on the idea of soft-thresholding the sample covariance matrix, then performing shrinkage towards this thresholded version.
-This introduces an extra parameter, the threshold $\delta$, which is used to control the amount of soft-thresholding, offering more flexibility.
+The NOVELIST (NOVEL Integration of the Sample and Thresholded Covariance) estimator, proposed by @Huang2019-ua, is currently the main focus of this research project.
+\textcolor{red}{It introduces adaptive sparsity, retaining strong correlations while discarding weak, noisy links.
+NOVELIST offers more flexibility than the shrinkage estimator, however, it does not guarantee positive definiteness.}
+
+The method is based on the idea of soft-thresholding the sample covariance matrix, then performing shrinkage towards this thresholded version.
+This introduces an extra parameter, the threshold $\delta$, which is used to control the amount of soft-thresholding.
 The NOVELIST estimator is given by:
@@
 
-For a given threshold $\delta$, Huang & Fryzlewicz derived the optimal shrinkage intensity parameter $\lambda(\delta)$ using Ledoit-Wolf's lemma (Ledoit and Wolf, 2003). It can be computed as:
+For a given threshold $\delta$, @Huang2019-ua derived the optimal shrinkage intensity parameter $\lambda(\delta)$ using Ledoit-Wolf's lemma [@Ledoit2003-qv]. It can be computed as:
 
@@
 
+\textcolor{red}{Note that when $\delta \in \bigl[ \text{max}_{ij}|\hat{r}_{1,ij}|, \; 1 \bigr]$, the NOVELIST estimator collapses to the shrinkage estimator, and when $\delta = 0$, it becomes the sample covariance matrix.}
+
+\vspace{1em}
+
 ***(c) POET***
 
-The POET (Principal Orthogonal complEment Thresholding) estimator, proposed by Fan et al.
-(2013), is another "sparse" + "non-sparse" covariance estimator.
-It starts by decomposing the correlation matrix $\hat{\boldsymbol{R}}_1$ into a prominent principle components part (low-rank) and a orthogonal complement part $\hat{\boldsymbol{R}}_{1, K}$ (the correlation matrix after removing the first $K$ principal components).
+The POET (Principal Orthogonal complEment Thresholding) estimator, proposed by @Fan2013-jz, is another "sparse" + "non-sparse" covariance estimator.
+\textcolor{red}{It takes the latent factors into account, and is appealing when there are common drivers in the time series within the hierarchy.}
+
+The POET method starts by decomposing the correlation matrix $\hat{\boldsymbol{R}}_1$ into a prominent principle components part (low-rank) and a orthogonal complement part $\hat{\boldsymbol{R}}_{1, K}$ (the correlation matrix after removing the first $K$ principal components).
 Then it applies thresholding to $\hat{\boldsymbol{R}}_{1, K}$.
@@
 
-***(d) NOVELIST with PC-adjusted***
+\vspace{1em}
 
-This approach is similar to the POET estimator.
-The difference is that POET apply thresholding to the orthogonal complement part $\hat{\boldsymbol{R}}_{1, K}$, meanwhile we apply NOVELIST.
-It can be formulated as:
+***(d) PC-adjusted NOVELIST***
+
+\textcolor{red}{This approach is best of both worlds, leveraging the strengths of both NOVELIST and POET.
+The PC-adjusted NOVELIST overcomes the shortcomings of the current shrinkage estimator, taking prominent PCs into account while also offers extra flexibility.} The idea is to apply the NOVELIST estimator to the orthogonal complement part $\hat{\boldsymbol{R}}_{1, K}$, and then add the principal components part back.
+The PC-adjusted NOVELIST estimator is formulated as:
 
@@
 \hat{\boldsymbol{R}}^{N, K}_{1} = \sum_{k = 1}^K \hat{\gamma}_k \hat{\boldsymbol{\xi}}_k \hat{\boldsymbol{\xi}}_k' + \hat{\boldsymbol{R}}^{N}_{1, K}
+\; ,
 $$
@@
 and $\hat{\boldsymbol{R}}^{N}_{1, K}$ is the NOVELIST estimator applied to the orthogonal complement part $\hat{\boldsymbol{R}}_{1, K}$.
+Similar to the NOVELIST estimator, $\hat{\boldsymbol{R}}^{N, K}_{1}$ is not guaranteed to be positive definite.
 
@@
 
-2. The correlations between different blocks are imposed using the Algorithm 1 in Hardin, Garcia & Golan (2013).
+2. The correlations between different blocks are imposed using the Algorithm 1 in @Hardin2013-wu.
 

### 2025-05-05 research_proposal/research_proposal_draft.qmd
expand on OLS and WLS

@@
 
+\color{red}
+
 The choice of $\boldsymbol{G}$ determines the composition of reconciled forecasts $\tilde{\boldsymbol{y}}_h$.
-Methods are developed to estimate $\boldsymbol{G}$, including the OLS and WLS, proposed by @Hyndman2011-jv and @Hyndman2016-ic respectively.
+Methods are developed to estimate $\boldsymbol{G}$, the first attempts at least squares forecast reconciliation were made by @Hyndman2011-jv.
 
-- expand on about OLS and WLS
+The proposed the optimal $\boldsymbol{G}$ based on the regression model:
+
+$$
+\hat{\boldsymbol{y}}_h = \boldsymbol{S} \boldsymbol{\beta}_h + \boldsymbol{\epsilon}_h,
+$$
+
+where $\boldsymbol{\epsilon}_h = \tilde{\boldsymbol{y}}_h - \hat{\boldsymbol{y}}_h$ is the coherency error with variance $\boldsymbol{V}_h$.
+This led to the GLS solution: $\boldsymbol{G} = (\boldsymbol{S}' \boldsymbol{V}_h^{-1} \boldsymbol{S})^{-1} \boldsymbol{S}' \boldsymbol{V}_h^{-1}$
+
+However, the covariance matrix $\boldsymbol{V}_h$ is unknown (and later shown by @Wickramasuriya2019-xq to be unidentifiable), and replaced by an identity matrix.
+The method then collapses into an OLS solution: $\boldsymbol{G} = (\boldsymbol{S}' \boldsymbol{S})^{-1} \boldsymbol{S}'$.
+
+An obvious drawback of the OLS solution is that it weights all series in any level equally.
+The issue prompted @Hyndman2016-ic to propose a WLS solution, where the series are weighted by the inverse variances of the base forecasts.
+The WLS solution is $\boldsymbol{G} = (\boldsymbol{S}' \boldsymbol{\Lambda}_h^{-1} \boldsymbol{S})^{-1} \boldsymbol{S}' \boldsymbol{\Lambda}_h^{-1}$.
+$\boldsymbol{\Lambda}_h = \text{diag} (\boldsymbol{W}_h)$ and $\boldsymbol{W}_h = Var(\boldsymbol{y}_h - \hat{\boldsymbol{y}}_h)$.
+
+\color{black}
 
@@
 
-@Wickramasuriya2019-xq also showed that MinT is an algebraical generalisation of the GLS, and the OLS and WLS methods are special cases of MinT when $\boldsymbol{W}_h$ is a diagonal or identity matrix, respectively.
+@Wickramasuriya2019-xq also showed that MinT is an algebraical generalisation of the GLS, and the OLS and WLS methods are special cases of MinT when $\boldsymbol{W}_h$ is an identity and a diagonal matrix, respectively.
 However, the MinT solution hinges on a reliable estimate of $\boldsymbol{W}_h$, which is challenging to estimate in high-dimensional setting.

### 2025-05-05 research_proposal/research_proposal_draft.qmd
Added grouped fig, and small fix

@@
 The top level of the tree represents the total value of all series, while the lower levels represent the individual series.
-When there are attributes of interest that are crossed, such as the electricity demand of any aggregation level (national, state, or regional) can be further broken down by usage purposes (e.g., residential, commercial), it will become a grouped time series.
+When there are attributes of interest that are crossed, such as the electricity demand of any aggregation level (national, state, or regional) can be further broken down by usage purposes (e.g., residential, commercial), it will become a grouped time series (illustrated in @fig-grouped).
 
-![Diagram of 2-level hierarchical tree structure](D:/Github/Recon_Honours_Thesis/research_proposal/figs/hierarchical_structure.png){#fig-hierarchy width="40%"}
+![Diagram of 2-level hierarchical tree structure](D:/Github/Recon_Honours_Thesis/research_proposal/figs/hierarchical_structure.png){#fig-hierarchy width="40%" height="40%"}
+
+![Diagram of 2-level grouped structure, and can be considered as the union of two hierarchical trees with common top and bottom level series](D:/Github/Recon_Honours_Thesis/research_proposal/figs/grouped_structure.png){#fig-grouped width="90%" height="40%"}
 
@@
 The proposed MinT approach by @Wickramasuriya2019-xq uses the shrinkage estimator from @Schafer2005-yw.
-\textcolor{red}{It guarantees positive definiteness and huge variance reduction when $p>T$}.
+\textcolor{red}{It guarantees positive definiteness and variance reduction for the covariance matrix, especially when $p>T$}.
 The shrinkage estimator is given by:

### 2025-05-07 research_proposal/research_proposal_draft.qmd
change the subheading styles of covs
Update research_proposal_draft.qmd

@@
 
-\textcolor{red}{Globally, Starbucks manages thousands of outlets across dozens of countries, each projecting sales for dozens of beverages.
-Yet when regional, national, and brand-wide forecasts are aggregated, discrepancies of millions of dollars emerge simply because individual outlet forecasts fail to respect the hierarchical nature.
-Forecast reconciliation comes in to solve this huge problem, where the individual forecasts are adjusted to satisfy the aggregation constraints.} Among the various reconciliation methods, the **MinT** (Minimum Trace) is considered the optimal approach, however, it requires a good estimate of the covariance matrix of the base forecast errors.
+Globally, Starbucks manages thousands of outlets across dozens of countries, each projecting sales for dozens of beverages.
+Yet when regional, national, and brand-wide forecasts are generated, discrepancies of millions of dollars emerge simply because individual outlet forecasts fail to respect the hierarchical nature of the data.
+Forecast reconciliation comes in to solve this huge problem, where the individual forecasts are adjusted to satisfy the aggregation constraints.
+Among the various reconciliation methods, **MinT** (Minimum Trace) is considered the optimal approach, however, it requires a good estimate of the covariance matrix of the base forecast errors.
 The current practice is to use the shrinkage estimator (often shrinking toward a diagonal matrix), but it lacks flexibility and might neglect the prominent structure presented.
@@
 The hierarchical structure can be represented as a tree, as shown in @fig-hierarchy.
-The top level of the tree represents the total value of all series, while the lower levels represent the individual series.
-When there are attributes of interest that are crossed, such as the electricity demand of any aggregation level (national, state, or regional) can be further broken down by usage purposes (e.g., residential, commercial), it will become a grouped time series (illustrated in @fig-grouped).
+The top level of the tree represents the total value of all series, while the lower levels represent the series at different levels of disaggregation.
+When there are attributes of interest that are crossed, such as the electricity demand at any aggregation level (national, state, or regional) is also considered by usage purposes (e.g., residential, commercial), the structure is described as a grouped time series (illustrated in @fig-grouped).
 
@@
 
-![Diagram of 2-level grouped structure, and can be considered as the union of two hierarchical trees with common top and bottom level series](D:/Github/Recon_Honours_Thesis/research_proposal/figs/grouped_structure.png){#fig-grouped width="90%" height="40%"}
+![Diagram of 2-level grouped structure, which can be considered as the union of two hierarchical trees with common top and bottom level series](D:/Github/Recon_Honours_Thesis/research_proposal/figs/grouped_structure.png){#fig-grouped width="90%" height="40%"}
 
-We refer to both of these structures as hierarchical time series, and can be represented using matrix algebra:
+For simplicity, we refer to both of these structures as hierarchical time series, we will distinguish between them if and when it is necessary.
+All hierarchical structures can be represented using matrix algebra:
 
@@
 \boldsymbol{y}_t = \boldsymbol{S} \boldsymbol{b}_t,
-$$ {#eq-1}
+$$
 
@@
 
-The example summing matrix $\boldsymbol{S}$ for the tree structure in @fig-hierarchy is:
+The summing matrix $\boldsymbol{S}$ for the tree structure in @fig-hierarchy is:
 
@@
 \end{array}
-\right]
+\right].
 $$
@@
 
-\textcolor{red}{If we produce base forecasts for the bottom-level series $\hat{\boldsymbol{b}}_h$ (any method may have been used to produce the base forecasts) and then reconcile them to all aggregated levels $\tilde{\boldsymbol{y}}_h$, we would not be utilising the information from those higher levels.
-This is similar to the top-down and middle-out approaches, with appropriate summing matrix $\boldsymbol{S}$.
-Another issue with expressing this way is that it restricts the reconciliation to only single-level approaches.
-Thus, @Hyndman2011-jv suggested a generalised expression for all existing reconciled methods, which also provides a framework for new methods to be developed:}
+\color{red}
+
+Assume we produce h-step-ahead base forecasts $\hat{\boldsymbol{b}}_h$ for the bottom-level series, obtained by any prediction methods. Then pre-multiplying them by $\boldsymbol{S}$ we get:
+
+$$
+\tilde{\boldsymbol{y}}_h = \boldsymbol{S} \hat{\boldsymbol{b}}_h .
+$$ {#eq-sb}
+
+We refer to $\tilde{\boldsymbol{y}}_h$ as coherent forecasts, as they respect the aggregation structure.
+We also refer to this way of obtaining coherent forecasts by summing the bottom-level forecasts as the bottom-up approach.
+However, generating forecasts this way is anchored only to prediction models at a single level, and will not be utilising the information from other levels.
+This drawback applies to the top-down and middle-out approaches.
+For example, the bottom-level data can be very noisy or even intermittent, and the higher-level data might be smoother due to the aggregation.
+
+Another issue with expressing reconciled methods as in @eq-sb is that it restricts the reconciliation to only single-level approaches.
+Thus, @Hyndman2011-jv suggested a generalised expression for all existing methods, which also provides a framework for new methods to be developed:
+
+\color{black}
 
@@
 \tilde{\boldsymbol{y}}_h = \boldsymbol{S} \boldsymbol{G} \hat{\boldsymbol{y}}_h,
-$$ {#eq-2}
+$$ {#eq-sgy}
 
 for a suitable $n_b \times n$ matrix $\boldsymbol{G}$.
-$\boldsymbol{G}$ maps the base forecasts of all levels $\hat{\boldsymbol{y}}_h$ down into the bottom series, which is then aggregated to the higher levels by $\boldsymbol{S}$.
-
-\color{red}
+$\boldsymbol{G}$ maps the base forecasts of all levels $\hat{\boldsymbol{y}}_h$ down into the bottom level, which is then aggregated to the higher levels by $\boldsymbol{S}$.
 
@@
 
-The proposed the optimal $\boldsymbol{G}$ based on the regression model:
+They proposed the optimal $\boldsymbol{G}$ based on the regression model:
 
@@
 where $\boldsymbol{\epsilon}_h = \tilde{\boldsymbol{y}}_h - \hat{\boldsymbol{y}}_h$ is the coherency error with variance $\boldsymbol{V}_h$.
-This led to the GLS solution: $\boldsymbol{G} = (\boldsymbol{S}' \boldsymbol{V}_h^{-1} \boldsymbol{S})^{-1} \boldsymbol{S}' \boldsymbol{V}_h^{-1}$
+This led to the GLS solution: $\boldsymbol{G} = (\boldsymbol{S}' \boldsymbol{V}_h^{-1} \boldsymbol{S})^{-1} \boldsymbol{S}' \boldsymbol{V}_h^{-1}$.
 
@@
 
-An obvious drawback of the OLS solution is that it weights all series in any level equally.
-The issue prompted @Hyndman2016-ic to propose a WLS solution, where the series are weighted by the inverse variances of the base forecasts.
+An obvious drawback of the OLS solution is that it weights all series in any level equally, regardless of its scale or forecast error variance.
+The issue prompted @Hyndman2016-ic to propose a WLS solution, where the series are weighted by the inverse variances of the base forecast errors.
 The WLS solution is $\boldsymbol{G} = (\boldsymbol{S}' \boldsymbol{\Lambda}_h^{-1} \boldsymbol{S})^{-1} \boldsymbol{S}' \boldsymbol{\Lambda}_h^{-1}$.
@@
 
-\color{black}
+## Minimum Trace (MinT) Reconciliation
 
-## Mininum Trace (MinT) Reconciliation
-
-@Wickramasuriya2019-xq reframed the problem by taking an optimisation approach rather than the regression.
-They formulated the problem as minimising the variances of all reconciled forecasts from @eq-2, which happens to be equivalent to minimising the trace of the covariance matrix (sum of the diagonal elements).
+@Wickramasuriya2019-xq reframed the problem by taking an optimisation approach.
+They formulated the problem as minimising the variances of all reconciled forecasts from @eq-sgy.
+They showed that this is equivalent to minimising the trace of the base forecast error covariance matrix (sum of the diagonal elements - the variances).
 This is known as the Minimum Trace (MinT) reconciliation method.
@@
 
-\vspace{1em}
-
-***(a) Shrinkage***
+### (a) Shrinkage {-}
 
@@
 
-\vspace{1em}
-
-***(b) NOVELIST***
+### (b) NOVELIST {-}
 
@@
 
-\vspace{1em}
-
-***(c) POET***
+### (c) POET {-}
 
@@
 
-\vspace{1em}
-
-***(d) PC-adjusted NOVELIST***
+### (d) PC-adjusted NOVELIST {-}
 

### 2025-05-07 research_proposal/research_proposal_draft.qmd
small fixes and re render

@@
 
-@Wickramasuriya2019-xq also showed that MinT is an algebraical generalisation of the GLS, and the OLS and WLS methods are special cases of MinT when $\boldsymbol{W}_h$ is an identity and a diagonal matrix, respectively.
-However, the MinT solution hinges on a reliable estimate of $\boldsymbol{W}_h$, which is challenging to estimate in high-dimensional setting.
+@Wickramasuriya2019-xq also showed that MinT is an algebraic generalisation of the GLS, and the OLS and WLS methods are special cases of MinT when $\boldsymbol{W}_h$ is an identity and a diagonal matrix, respectively.
+
+The MinT solution hinges on a reliable, positive-definite estimate of $\boldsymbol{W}_h$, which is challenging to estimate in high-dimensional setting.
 Therefore, we will be exploring alternative covariance estimators.
@@
 $$
-\hat{\lambda}_D = \frac{\sum_{i \neq j} \widehat{Var}(\hat{r}_{ij})} {\sum_{i \neq j} \hat{r}_{ij}}
+\hat{\lambda}_D = \frac{\sum_{i \neq j} \widehat{Var}(\hat{r}_{ij})} 
+{\sum_{i \neq j} \hat{r}_{ij}^2}
 $$

### 2025-05-07 research_proposal/research_proposal_draft.qmd
Refine Alternative Covariance Estimators section

@@
 
-We reconstruct estimator of $\boldsymbol{W}_h$ as $\hat{\boldsymbol{W}}_h = k_h \, g(\hat{\boldsymbol{W}}_1)$, and $g(.)$ is an estimator function of the unbiased sample covariance of in-sample one-step-ahead base forecast errors $\hat{\boldsymbol{W}}_1 = \frac{1}{T} \sum_{t=1}^{T} \hat{\boldsymbol{e}}_{t|t-1} \hat{\boldsymbol{e}}_{t|t-1}'$
+We reconstruct the estimator of $\boldsymbol{W}_h$ as $\hat{\boldsymbol{W}}_h = k_h \, g(\hat{\boldsymbol{W}}_1)$, where $k_h > 0$.
+The function $g(.)$ is an estimator of the unbiased sample covariance matrix of the in-sample one-step-ahead base forecast errors $\hat{\boldsymbol{W}}_1 = \frac{1}{T} \sum_{t=1}^{T} \hat{\boldsymbol{e}}_{t|t-1} \hat{\boldsymbol{e}}_{t|t-1}'$, and $T$ is the length of series' time dimension.
 
@@
 The proposed MinT approach by @Wickramasuriya2019-xq uses the shrinkage estimator from @Schafer2005-yw.
-\textcolor{red}{It guarantees positive definiteness and variance reduction for the covariance matrix, especially when $p>T$}.
+It guarantees positive definiteness and variance reduction for the covariance matrix, especially when the total number of series $n>T$.
 The shrinkage estimator is given by:
@@
 $$
-\hat{\boldsymbol{W}}^{shr}_{1, D} = \lambda_D \hat{\boldsymbol{W}}_{1, D} + (1 - \lambda_D) \hat{\boldsymbol{W}}_1
+\hat{\boldsymbol{W}}^{shr}_{1} = \lambda_D \hat{\boldsymbol{W}}_{1, D} + (1 - \lambda_D) \hat{\boldsymbol{W}}_1,
 $$
 
-$\hat{\boldsymbol{W}}_{1, D}$ is a diagonal matrix comprising the diagonal entries of $\hat{\boldsymbol{W}}_1$.
-This approach will shrink the covariance matrix $\hat{\boldsymbol{W}}_1$ towards its diagonal version, meaning the off-diagonal elements are shrunk towards zero while the diagonal ones remain unchanged.
+where $\hat{\boldsymbol{W}}_{1, D}$ is a diagonal matrix comprising the diagonal entries of $\hat{\boldsymbol{W}}_1$.
+We refer to any $\lambda \in [0,1]$ as the shrinkage intensity parameter, the subscript specifies which estimator it belongs to.
+This approach shrinks the covariance matrix $\hat{\boldsymbol{W}}_1$ towards its diagonal matrix, meaning the off-diagonal elements are shrunk towards zero while the diagonal ones remain unchanged.
 
-@Schafer2005-yw also proposed an optimal shrinkage intensity parameter $\lambda_D$ for this setting, assuming the variances are constant:
+@Schafer2005-yw also proposed an estimate of the optimal shrinkage intensity parameter $\lambda_D$:
 
@@
 
-where $\hat{r}_{ij}$ is the $ij$th element of $\hat{\boldsymbol{R}}_1$, the 1-step-ahead sample correlation matrix (obtained from $\hat{\boldsymbol{W}}_1$) to shrink it toward an identity matrix
+where $\hat{r}_{ij}$ is the $ij$th element of $\hat{\boldsymbol{R}}_1$, the 1-step-ahead sample correlation matrix (obtained from $\hat{\boldsymbol{W}}_1$).
 
-The optimal lambda is obtained by minimising the $MSE(\hat{\boldsymbol{W}}_1) = Bias(\hat{\boldsymbol{W}}_1)^2 + Var(\hat{\boldsymbol{W}}_1)$.
+The optimal $\lambda_D$ is obtained by minimising the mean squared error of $\hat{\boldsymbol{W}}_1$: $MSE(\hat{\boldsymbol{W}}_1) = Bias(\hat{\boldsymbol{W}}_1)^2 + Var(\hat{\boldsymbol{W}}_1)$.
 More specifically, we trade the unbiasedness of the sample covariance matrix for a lower variance.
 The objective function itself does not take into account any possible principal components structure in the data, and is not flexible enough since it shrinks all off-diagonal elements equally towards zeros.
+
+- Screeplot of eigenvalues
 
@@
 The NOVELIST (NOVEL Integration of the Sample and Thresholded Covariance) estimator, proposed by @Huang2019-ua, is currently the main focus of this research project.
-\textcolor{red}{It introduces adaptive sparsity, retaining strong correlations while discarding weak, noisy links.
-NOVELIST offers more flexibility than the shrinkage estimator, however, it does not guarantee positive definiteness.}
+\textcolor{red}{It introduces adaptive sparsity, retaining strong correlations while discarding weak, noisy correlations.
+NOVELIST offers more flexibility than the shrinkage estimator, however, it does not guarantee to be positive definite.}
 
@@
 $$
-\hat{\boldsymbol{W}}^{N}_{1} = \lambda_{\delta} \hat{\boldsymbol{W}}_{1, \delta} + (1 - \lambda_{\delta}) \hat{\boldsymbol{W}}_1
+\hat{\boldsymbol{W}}^{N}_{1} = \lambda_{\delta} \hat{\boldsymbol{W}}_{1, \delta} + (1 - \lambda_{\delta}) \hat{\boldsymbol{W}}_1,
 $$
 
-By convenient setting, we rewrite it as in correlation matrix:
+where $\hat{\boldsymbol{W}}_{1, \delta}$ is the thresholded version of $\hat{\boldsymbol{W}}_{1}$. By convenient setting, we can rewrite it in terms of correlation:
 
@@
 
-In this setting, $\hat{\boldsymbol{R}}_{1,\delta}$ is the thresholded sample correlation matrix, where each element is regularised by:
+In this setting, $\hat{\boldsymbol{R}}_{1,\delta}$ is the thresholded correlation matrix, where each element is regularised by:
 
 $$
-\hat{r}_{1,ij}^\delta = \operatorname{sign}(\hat{r}_{1,ij}) \, (|\hat{r}_{1,ij}| - \delta)_+.
+\hat{r}_{1,ij}^\delta = \text{sign}(\hat{r}_{1,ij}) \, 
+\text{max}(|\hat{r}_{1,ij}| - \delta, \; 0),
 $$
 
-For a given threshold $\delta$, @Huang2019-ua derived the optimal shrinkage intensity parameter $\lambda(\delta)$ using Ledoit-Wolf's lemma [@Ledoit2003-qv]. It can be computed as:
+where $\delta \in [0,1]$ is the threshold parameter.
+For a given threshold $\delta$, @Huang2019-ua derived an analytical expression for the optimal shrinkage intensity parameter $\lambda(\delta)$ using Ledoit-Wolf's lemma [@Ledoit2003-qv].
+It can be computed as:
+
+- \textcolor{red}{@Huang2019-ua mentioned Ledoit-Wolf's lemma}
 
@@
 \hat{\lambda}(\delta) = \frac{
-  \sum_{i \neq j} \widehat{Var}(\hat{r}_{1,ij}) \; I(|\hat{r}_{1,ij}| \leq \delta)
+  \sum_{i \neq j} \widehat{Var}(\hat{r}_{1,ij}) \; \boldsymbol{1}(|\hat{r}_{1,ij}| \leq \delta)
 } {
   \sum_{i \neq j} (\hat{r}_{1,ij} - \hat{r}_{1,ij}^\delta)^2
-}
+},
 $$
 
+where $\boldsymbol{1}(.)$ is the indicator function.
+
 On the other hand, the optimal threshold $\delta^*$ does not have a closed-form solution, and is typically obtained by executing a rolling-window cross-validation procedure.
-The idea is to, for each rolling-window set, loop through a range of threshold values, compute the corresponding $\hat{\lambda}(\delta)$ and $\hat{\boldsymbol{R}}^{N}_{1}$.
-Then we select the one that minimises the average out-of-sample reconciled forecast errors.
-Although it is not required to fit forecasting models multiple time, the cross-validation procedure is still computationally expensive as it computes the NOVELIST estimator and perform reconciliation for each threshold value.
 The formal algorithm is given in the @sec-novelist_cv Appendix.
+Although it is not required to fit forecasting models multiple times, the cross-validation procedure is still computationally expensive as it computes the NOVELIST estimator and perform reconciliation for each threshold value.
 
-\textcolor{red}{Note that when $\delta \in \bigl[ \text{max}_{ij}|\hat{r}_{1,ij}|, \; 1 \bigr]$, the NOVELIST estimator collapses to the shrinkage estimator, and when $\delta = 0$, it becomes the sample covariance matrix.}
+<!-- The idea is to, for each rolling-window set, loop through a range of threshold values, compute the corresponding $\hat{\lambda}(\delta)$ and $\hat{\boldsymbol{R}}^{N}_{1}$.
+Then we select the one that minimises the average out-of-sample reconciled forecast errors -->
+
+Note that when $\delta \in \bigl[ \text{max}_{i \neq j}|\hat{r}_{1,ij}|, \; 1 \bigr]$, the NOVELIST estimator collapses to the shrinkage estimator, and when $\delta = 0$, it becomes the sample covariance matrix.
 
@@
 The POET (Principal Orthogonal complEment Thresholding) estimator, proposed by @Fan2013-jz, is another "sparse" + "non-sparse" covariance estimator.
-\textcolor{red}{It takes the latent factors into account, and is appealing when there are common drivers in the time series within the hierarchy.}
+It takes the latent factors into account, and is appealing when there are common drivers in the time series within the hierarchy.
 
@@
 
-where $\hat{\gamma}_k$ and $\hat{\boldsymbol{\xi}}_k$ are the $k$th eigenvalue and eigenvector of the sample covariance matrix.
-$T(.)$ is the thresholding function, which can be either soft-thresholding, hard-thresholding, or others.
+where $\hat{\gamma}_k$ and $\hat{\boldsymbol{\xi}}_k$ are the $k$th largest eigenvalue and the corresponding eigenvector of the sample covariance matrix, respectively, and $T(.)$ is the thresholding function, which can be either soft-thresholding, hard-thresholding, or others.
 
@@
 
-\textcolor{red}{This approach is best of both worlds, leveraging the strengths of both NOVELIST and POET.
-The PC-adjusted NOVELIST overcomes the shortcomings of the current shrinkage estimator, taking prominent PCs into account while also offers extra flexibility.} The idea is to apply the NOVELIST estimator to the orthogonal complement part $\hat{\boldsymbol{R}}_{1, K}$, and then add the principal components part back.
+This approach is best of both worlds, leveraging the strengths of both NOVELIST and POET.
+The PC-adjusted (Principal-Component-adjusted) NOVELIST overcomes the shortcomings of the current shrinkage estimator, taking prominent PCs into account while also offers extra flexibility.
+The idea is to apply the NOVELIST estimator to the orthogonal complement part $\hat{\boldsymbol{R}}_{1, K}$, and then add the principal components part back.
 The PC-adjusted NOVELIST estimator is formulated as:
@@
 
-and $\hat{\boldsymbol{R}}^{N}_{1, K}$ is the NOVELIST estimator applied to the orthogonal complement part $\hat{\boldsymbol{R}}_{1, K}$.
+where $\hat{\boldsymbol{R}}^{N}_{1, K}$ is the NOVELIST estimator applied to the orthogonal complement part $\hat{\boldsymbol{R}}_{1, K}$.
 Similar to the NOVELIST estimator, $\hat{\boldsymbol{R}}^{N, K}_{1}$ is not guaranteed to be positive definite.
+
+Methods to ensure positive definiteness of the NOVELIST estimator (and its PC-adjusted variant) will be explored and studied in the project. @Huang2019-ua proposed to diagonalise the NOVELIST
+estimator and replace any eigenvalues that fall under a certain small positive threshold by the value of that threshold.
+Alternatively, we can implement the algorithm of @Higham2002-te that computes the nearest positive definite matrix to a given matrix.
 

### 2025-05-07 research_proposal/research_proposal_draft.qmd
Refine timeline and DGP appendix

@@
     "May", 
-    "June - July", 
-    "August",
-    "September - October"
+    "June",
+    "July", 
+    "August - September",
+    "October"
   ),
@@
     "Experiments with POET and PC-adjusted NOVELIST",
+    "Exploration of other high-dimensional covariance estimators",
     "Real-world application + evaluation",
@@
     "Results: POET & PC-adjusted NOVELIST",
+    "Results: other estimators",
     "Application completed",
@@
 $$
-\boldsymbol{y}_t = \boldsymbol{A} \boldsymbol{y}_{t-1} + \boldsymbol{e}_t,
+\boldsymbol{b}_t = \boldsymbol{A} \boldsymbol{b}_{t-1} + \boldsymbol{\epsilon}_t,
 $$
 
-where $\boldsymbol{A}$ is a $p \times p$ block diagonal matrix of autoregressive coefficients $\boldsymbol{A} = diag(\boldsymbol{A}_1, \ldots, \boldsymbol{A}_m)$, with each $\boldsymbol{A}_i$ being a $p_i \times p_i$ matrix.
+where $\boldsymbol{A}$ is a $n_b \times n_b$ block diagonal matrix of autoregressive coefficients $\boldsymbol{A} = diag(\boldsymbol{A}_1, \ldots, \boldsymbol{A}_m)$, with each $\boldsymbol{A}_i$ being a $n_{b,i} \times n_{b,i}$ matrix.
 The block diagonal structure ensures that the time series are grouped into $m$ groups, with each group having its own autoregressive coefficients.
@@
 
-The model is added with a Gaussian innovation process $\boldsymbol{e}_t$, with covariance matrix $\Sigma$. The covariance matrix $\Sigma$ is generated specifically in the following way:
+The model is added with a Gaussian innovation process $\boldsymbol{\epsilon}_t$, with covariance matrix $\Sigma$. The covariance matrix $\Sigma$ is generated specifically in the following way:
 
-1. A compound symmetric correlation matrix is used for each block of size $p_i$ in $\boldsymbol{A}_i$, where the coefficients are sampled from a uniform distribution.
+1. A compound symmetric correlation matrix is used for each block of size $n_{b,i}$ in $\boldsymbol{A}_i$, where the coefficients are sampled from a uniform distribution between 0 and 1.
 
@@
 
-3. The covariance matrix $\Sigma$ is then constructed by uniform sampling of standard deviations for all $p$ series.
+3. The covariance matrix $\Sigma$ is then constructed by uniform sampling of standard deviations, in a range of $[\sqrt{2}, \sqrt{6}]$, for all $n_b$ series.
 
 We have an option to randomly flip the signs of the covariance elements, which will create a more realistic structure in the innovation process.
+This can be done by pre- and post-multiplying $\Sigma$ by a random diagonal matrix $V$ with entries sampled from $\{-1, 1\}$, yielding $\Sigma' = V \Sigma V$.
 

### 2025-05-08 research_proposal/research_proposal_draft.qmd
Pseudo algorithm

@@
         <!-- \usepackage{mathspec} -->
-        <!-- \usepackage{algorithm} -->
-        <!-- \usepackage{algpseudocode} -->
+        \usepackage{algorithm}
+        \usepackage{algpseudocode}
         \usepackage[colorlinks=true, linkcolor=blue, citecolor=blue, urlcolor=blue]{hyperref}
@@
 We reconstruct the estimator of $\boldsymbol{W}_h$ as $\hat{\boldsymbol{W}}_h = k_h \, g(\hat{\boldsymbol{W}}_1)$, where $k_h > 0$.
-The function $g(.)$ is an estimator of the unbiased sample covariance matrix of the in-sample one-step-ahead base forecast errors $\hat{\boldsymbol{W}}_1 = \frac{1}{T} \sum_{t=1}^{T} \hat{\boldsymbol{e}}_{t|t-1} \hat{\boldsymbol{e}}_{t|t-1}'$, and $T$ is the length of series' time dimension.
+The function $g(.)$ is an estimator of the unbiased sample covariance matrix of the in-sample one-step-ahead base forecast errors $\hat{\boldsymbol{W}}_1 = \frac{1}{T} \sum_{t=1}^{T} \hat{\boldsymbol{e}}_{t} \hat{\boldsymbol{e}}_{t}'$, and $T$ is the length of series' time dimension.
+
+<!-- $\boldsymbol{e}}_{t|t-1}$ we keep things simplified without the conditional subscript -->
 
@@
 \hat{\boldsymbol{W}}^{N}_{1} = \lambda_{\delta} \hat{\boldsymbol{W}}_{1, \delta} + (1 - \lambda_{\delta}) \hat{\boldsymbol{W}}_1,
-$$
+$$ {#eq-novelist-cov}
 
@@
 \hat{\boldsymbol{R}}^{N}_{1} = \lambda_{\delta} \hat{\boldsymbol{R}}_{1,\delta} + (1 - \lambda_{\delta}) \hat{\boldsymbol{R}}_1,
-$$
+$$ {#eq-novelist-cor}
 
@@
 \text{max}(|\hat{r}_{1,ij}| - \delta, \; 0),
-$$
+$$ {#eq-soft-thr}
 
@@
 },
-$$
+$$ {#eq-lambda-thr}
 
@@
 
-![NOVELIST cross-validation procedure](D:/Github/Recon_Honours_Thesis/research_proposal/figs/novelist_cv_algo.png){#fig-novelist_cv height=70%}
+```{=latex}
+\begin{algorithm}
+\caption{Cross-validation procedure}
+\begin{algorithmic}[1]
+
+\State \textbf{Input:} Observations and fitted values $\boldsymbol{y}_t, \hat{\boldsymbol{y}}_t \in \mathbb{R}^n$ for $t = 1,\dots,T$, set of threshold candidates $\Delta$, window size $v$.
+
+\State $\hat{\boldsymbol{e}}_t = \boldsymbol{y}_t - \hat{\boldsymbol{y}}_t$ for $t = 1,\dots,T$
+
+\For{$i = v:T-1$ do}
+    \State $j = i - v +1$
+    \State $\hat{\boldsymbol{W}}_j = \frac{1}{v} \sum_{t=j}^{i} \hat{\boldsymbol{e}}_{t} \hat{\boldsymbol{e}}_{t}'$
+    \State $\hat{\boldsymbol{D}}_j = \text{diag}(\hat{\boldsymbol{W}}_j)$
+    \State $\hat{\boldsymbol{R}}_j = \hat{\boldsymbol{D}}_j^{-1/2} \hat{\boldsymbol{W}}_j \hat{\boldsymbol{D}}_j^{-1/2}$
+    \For{$\delta \in \Delta$}
+        \State Compute thresholded correlation $\hat{\boldsymbol{R}}_{j,\delta}$ using Equation 5
+        \State Compute $\hat{\lambda}_{j,\delta}$ using Equation 6
+        \State Compute $\hat{\boldsymbol{R}}^{N}_{j,\delta}$ using Equation 4
+        \State $\hat{\boldsymbol{W}}^{N}_{j,\delta} = \hat{\boldsymbol{D}}_j^{1/2} \hat{\boldsymbol{R}}^{N}_{j,\delta} \hat{\boldsymbol{D}}_j^{1/2}$
+        \State $\boldsymbol{G} = (\boldsymbol{S}' \hat{\boldsymbol{W}}^{N^{-1}}_{j,\delta} \boldsymbol{S})^{-1} \boldsymbol{S}' \hat{\boldsymbol{W}}^{N^{-1}}_{j,\delta}$
+        \State Reconciled forecasts $\tilde{\boldsymbol{y}}_{i+1 | \delta} = \boldsymbol{S} \boldsymbol{G} \hat{\boldsymbol{y}}_{i+1}$
+        \State $\tilde{\boldsymbol{e}}_{i+1 | \delta} = \boldsymbol{y}_{i+1} - \tilde{\boldsymbol{y}}_{i+1 | \delta}$
+    \EndFor
+\EndFor
+
+\State $\text{MSE}_{\delta} = \frac{1}{T-v} \sum_{i=v}^{T-1} (\tilde{\boldsymbol{e}}_{i+1 | \delta})^2$ for each $\delta \in \Delta$
+\State $\hat{\delta}^* = \arg\min_{\delta \in \Delta} \text{MSE}_{\delta}$
+\State Compute $\hat{\lambda}^*$ on all training data using $\hat{\delta}^*$
+\State Compute $\hat{\boldsymbol{R}}_1^*$ using $\hat{\delta}^*$ and $\hat{\lambda}^*$ on all training data, using Equation 3
+
+\State \textbf{Output:} Estimate of optimal $\hat{\delta}^*$
+
+\end{algorithmic}
+\end{algorithm}
+```
+
+<!-- ![NOVELIST cross-validation procedure](D:/Github/Recon_Honours_Thesis/research_proposal/figs/novelist_cv_algo.png){#fig-novelist_cv height=70%} -->
 
@@
 
-<!-- ```{=latex} -->
-<!-- \begin{algorithm} -->
-<!-- \caption{Cross-validation procedure to select optimal threshold $\delta^*$ for NOVELIST} -->
-<!-- \begin{algorithmic}[1] -->
-<!-- \State \textbf{Input:} Residual matrix $E$ of dimension $T \times p$, grid of threshold candidates $\{\delta_j\}$, window size $n$ -->
-<!-- \For{each threshold candidate $\delta_j$} -->
-<!--     \For{$t = n+1$ to $T$} -->
-<!--         \State Define in-sample window: $E_{1:n} = \{e_{t-n}, ..., e_{t-1}\}$ -->
-<!--         \State Compute sample covariance $\hat{\Sigma}_t = \frac{1}{n} E_{1:n}^\top E_{1:n}$ -->
-<!--         \State Compute thresholded covariance: $\hat{\Sigma}^{\text{thr}}_t = \text{Threshold}(\hat{\Sigma}_t, \delta_j)$ -->
-<!--         \State Compute NOVELIST: $\hat{\Sigma}^{\text{nov}}_t = (1 - \lambda) \hat{\Sigma}_t + \lambda \hat{\Sigma}^{\text{thr}}_t$ -->
-<!--         \State Compute inverse: $W_t = (\hat{\Sigma}^{\text{nov}}_t)^{-1}$ -->
-<!--         \State Compute MinT reconciliation matrix: $P_t = (S^\top W_t S)^{-1} S^\top W_t$ -->
-<!--         \State Compute forecast error: $e_t = y_t - S P_t \hat{y}_t$ -->
-<!--     \EndFor -->
-<!--     \State Compute total CV loss $L_j = \sum_t \|e_t\|^2$ -->
-<!-- \EndFor -->
-<!-- \State \textbf{Output:} $\delta^* = \arg\min_{\delta_j} L_j$ -->
-<!-- \end{algorithmic} -->
-<!-- \end{algorithm} -->
-<!-- ``` -->
-
 \pagebreak

### 2025-05-09 research_proposal/table_function.R
function to generate kable

@@ -1,0 +1,130 @@
+library(tidyr)
+library(ggplot2)
+library(tsibble)
+library(dplyr)
+library(purrr)
+library(kableExtra)
+library(knitr)
+
+
+
+sim1 <- readRDS("D:/Github/Recon_Honours_Thesis/research_proposal/sim_results/S36-6-2-1_T100_M500_par.rds")
+sim2 <- readRDS("D:/Github/Recon_Honours_Thesis/research_proposal/sim_results/S36-6-2-1_T300_M500_par.rds")
+
+transfrom_sim_MSE <- function(MSE, to_ts = T) {
+  df <- imap_dfr(MSE, function(mat, model_name) {
+    as_tibble(mat) %>%
+      mutate(h = row_number()) %>%
+      pivot_longer(cols = -h, names_to = "series", values_to = "MSE") %>%
+      mutate(
+        .model = model_name,
+      ) %>%
+      select(.model, series, h, MSE)
+  })
+  if (to_ts) {
+    return(
+      as_tsibble(key = c(.model, series), index = h)
+    )
+  }
+  return(df)
+}
+
+
+MSE1 <- transfrom_sim_MSE(sim1$MSE, F)
+MSE2 <- transfrom_sim_MSE(sim2$MSE, F)
+
+hier_level <- MSE1$series |>
+  substring(1, 1) |>
+  unique() |> sort(decreasing = TRUE)
+
+MSE_full <- bind_rows(MSE1, MSE2, .id = "id") |>
+  mutate(
+    level = match(substring(series,1,1), hier_level)
+  )
+
+MSE_full |>
+  group_by(id, series, h) |>
+  mutate(
+    base_MSE = MSE[.model == "base"],
+    pct_change = ifelse(.model == "base",
+                        base_MSE,
+                        (MSE - base_MSE) / base_MSE * 100)
+  ) |>
+  ungroup()
+
+group_defs <- tibble(
+  h_group = c("h=1",  "h=1-8",   "h=1-16"),
+  h_max   = c(1,      8,         16)
+)
+
+MSE_tidy <- MSE_full %>%
+  # filter(h<=16) %>%
+  crossing(group_defs) %>%    # gives every row all 3 candidate groups
+  filter(h <= h_max) %>%      # keep only the overlapping ones
+  # filter(h>15)
+  group_by(id, .model, level, h_group) %>%
[…]

### 2025-05-09 research_proposal/research_proposal_draft.qmd
small fixes

@@
 
-![Diagram of 2-level hierarchical tree structure](D:/Github/Recon_Honours_Thesis/research_proposal/figs/hierarchical_structure.png){#fig-hierarchy width="40%" height="40%"}
+![A 2-level hierarchical tree structure](D:/Github/Recon_Honours_Thesis/research_proposal/figs/hierarchical_structure.png){#fig-hierarchy width="40%" height="40%"}
 
-![Diagram of 2-level grouped structure, which can be considered as the union of two hierarchical trees with common top and bottom level series](D:/Github/Recon_Honours_Thesis/research_proposal/figs/grouped_structure.png){#fig-grouped width="90%" height="40%"}
+![A 2-level grouped structure, which can be considered as the union of two hierarchical trees with common top and bottom level series](D:/Github/Recon_Honours_Thesis/research_proposal/figs/grouped_structure.png){#fig-grouped width="90%" height="40%"}
 
@@
 
-where $\boldsymbol{S}$ is a summing matrix of order $n \times n_b$ which aggregates the bottom-level series $\boldsymbol{b}_t$ ($n_b$-vector) to the series at aggregation levels above.
-The $n$-vector $\boldsymbol{y}_t$ contains all observations at time $t$.
+where $\boldsymbol{S}$ is a summing matrix of order $n \times n_b$ which aggregates the bottom-level series $\boldsymbol{b}_t \in \mathbb{R}^{n_b}$ to the series at aggregation levels above.
+The $n$-vector $\boldsymbol{y}_t \in \mathbb{R}^n$ contains all observations at time $t$.
 
@@
 0 & 0 & 1 & 1 \\
-\multicolumn{4}{c}{\boldsymbol{I_n}}
+\multicolumn{4}{c}{\boldsymbol{I_4}}
 \end{array}
@@
 Thus, @Hyndman2011-jv suggested a generalised expression for all existing methods, which also provides a framework for new methods to be developed:
-
-\color{black}
 
@@
 
-@Wickramasuriya2019-xq reframed the problem by taking an optimisation approach.
-They formulated the problem as minimising the variances of all reconciled forecasts from @eq-sgy.
+@Wickramasuriya2019-xq reframed the problem as minimising the variances of all reconciled forecasts from @eq-sgy.
 They showed that this is equivalent to minimising the trace of the base forecast error covariance matrix (sum of the diagonal elements - the variances).
@@
 The NOVELIST (NOVEL Integration of the Sample and Thresholded Covariance) estimator, proposed by @Huang2019-ua, is currently the main focus of this research project.
-\textcolor{red}{It introduces adaptive sparsity, retaining strong correlations while discarding weak, noisy correlations.
+\textcolor{red}{It introduces a way to control the target matrix's sparsity, retaining strong correlations while discarding weak, noisy correlations.
 NOVELIST offers more flexibility than the shrinkage estimator, however, it does not guarantee to be positive definite.}
@@
 where $\delta \in [0,1]$ is the threshold parameter.
-For a given threshold $\delta$, @Huang2019-ua derived an analytical expression for the optimal shrinkage intensity parameter $\lambda(\delta)$ using Ledoit-Wolf's lemma [@Ledoit2003-qv].
+For a given threshold $\delta$, @Huang2019-ua derived an analytical expression for the optimal shrinkage intensity parameter $\lambda(\delta)$ using Ledoit-Wolf's lemma [@Ledoit2003-qv], following similar logic to @Schafer2005-yw.
 It can be computed as:
 
-- \textcolor{red}{@Huang2019-ua mentioned Ledoit-Wolf's lemma}
+<!-- - \textcolor{red}{@Huang2019-ua mentioned Ledoit-Wolf's lemma} -->
 
@@
 On the other hand, the optimal threshold $\delta^*$ does not have a closed-form solution, and is typically obtained by executing a rolling-window cross-validation procedure.
+The idea is to find the threshold $\hat{\delta^*}$, with the corresponding $\hat{\lambda}^*$ and $\hat{\boldsymbol{R}}^{N}_{1}(\hat{\delta^*}, \hat{\lambda}^*)$, that minimises the average out-of-sample reconciled forecast errors.
 The formal algorithm is given in the @sec-novelist_cv Appendix.
 Although it is not required to fit forecasting models multiple times, the cross-validation procedure is still computationally expensive as it computes the NOVELIST estimator and perform reconciliation for each threshold value.
-
-<!-- The idea is to, for each rolling-window set, loop through a range of threshold values, compute the corresponding $\hat{\lambda}(\delta)$ and $\hat{\boldsymbol{R}}^{N}_{1}$.
-Then we select the one that minimises the average out-of-sample reconciled forecast errors -->
 

### 2025-05-09 research_proposal/research_proposal_draft.qmd
writing sim section and put table

@@
 
-The experimental design is to simulate a hierarchical time series data set, then apply the MinT reconciliation method with different covariance estimators.
+The experiment starts from simulating a hierarchical time series data set, then applying the MinT reconciliation method with different covariance estimators.
 The data set will be split into training and test sets.
-In case of cross-validation, the training set will be further split into training and validation sets.
-The data generating process is described in @sec-dgp Appendix.
+In the case of performing NOVELIST cross-validation, the training set will be further split into training and validation sets equally.
+
+In this preliminary experiment, we consider a slightly large hierarchical structure.
+The structure consists of 3 levels with a total of 45 series.
+The bottom level has 36 series and are aggregated in groups of size six to form six level-1 series.
+The level-1 series are then aggregated in groups of size three to form two level-2 series, which are finally aggregated to form the top level.
+The observations at bottom level is generated from a VAR(1) process.
+A more detailed data generating process is described in @sec-dgp Appendix.
+
+For each series, $T = 116$ or $316$ observations are generated.
+The first 100 or 300 observations are used for training, and the last 16 observations are used for testing.
+The remaining data is used to compute the best fitted ARIMA models by minimising the AICc criterion.
+We refer to them as base models, and their base forecasts are then reconciled using the MinT with different covariance estimators.
+These include using the unbiased sample covariance matrix - MinT(Sample), the shrinkage estimator - MinT(Shrink), and the NOVELIST estimator - MinT(N).
+The Monte Carlo simulation is repeated 500 times, in which the parameters for data generating process is also randomly generated.
+The results are presented in @tbl-sim-summary.
+
+```{r tbl-sim-summary}
+#| echo: false
+#| message: false
+#| warning: false
+
+# Code --------------------------------------
+library(tidyr)
+library(ggplot2)
+library(tsibble)
+library(dplyr)
+library(purrr)
+library(kableExtra)
+library(knitr)
+
+sim1 <- readRDS("D:/Github/Recon_Honours_Thesis/research_proposal/sim_results/S36-6-2-1_T100_M500_par.rds")
+sim2 <- readRDS("D:/Github/Recon_Honours_Thesis/research_proposal/sim_results/S36-6-2-1_T300_M500_par.rds")
+
+transfrom_sim_MSE <- function(MSE, to_ts = T) {
+  df <- imap_dfr(MSE, function(mat, model_name) {
+    as_tibble(mat) %>%
+      mutate(h = row_number()) %>%
+      pivot_longer(cols = -h, names_to = "series", values_to = "MSE") %>%
+      mutate(
+        .model = model_name,
+      ) %>%
+      select(.model, series, h, MSE)
+  })
+  if (to_ts) {
+    return(
+      as_tsibble(key = c(.model, series), index = h)
+    )
+  }
+  return(df)
+}
+
+MSE1 <- transfrom_sim_MSE(sim1$MSE, F)
+MSE2 <- transfrom_sim_MSE(sim2$MSE, F)
+
+hier_level <- MSE1$series |>
+  substring(1, 1) |>
+  unique() |> sort(decreasing = TRUE)
+
+MSE_full <- bind_rows(MSE1, MSE2, .id = "id") |>
+  mutate(
+    level = match(substring(series,1,1), hier_level)
+  )
+
+group_defs <- tibble(
+  h_group = c("h=1",  "h=1-8",   "h=1-16"),
+  h_max   = c(1,      8,         16)
+)
+
+MSE_tidy <- MSE_full %>%
+  # filter(h<=16) %>%
+  crossing(group_defs) %>%    # gives every row all 3 candidate groups
+  filter(h <= h_max) %>%      # keep only the overlapping ones
+  # filter(h>15)
+  group_by(id, .model, level, h_group) %>%
+  summarise(mean_MSE = mean(MSE), .groups = "drop")
+
+MSE_tidy <- MSE_tidy %>%
+  group_by(level, id, h_group) %>%
+  mutate(
+    base_MSE = mean_MSE[.model == "base"],
+    change = if_else(.model == "base",
+                     formatC(mean_MSE, digits = 1, format = "f"),
+                     formatC(100 * (mean_MSE - base_MSE) / base_MSE, digits = 1, format = "f"))
+  ) %>%
+  ungroup()
+
+MSE_tidy <- MSE_tidy %>%
+  pivot_wider(
+    id_cols = c(level, .model),
+    names_from = c(id, h_group),
+    values_from = change
+  ) %>%
+  mutate(
+    .model = recode(.model,
+                    "mint_sample" = "MinT(Sample)",
+                    "mint_shr"    = "MinT(Shrink)",
+                    "mint_n"      = "MinT(N)",
+                    "base"        = "Base"
+    ),
+    .model = factor(.model, levels = c("MinT(Sample)", "MinT(Shrink)", "MinT(N)", "Base"))
+  ) %>%
+  mutate(
+    level = recode(level,
+                   `1` = "Top level",
+                   `2` = "Level 1",
+                   `3` = "Level 2",
+                   `4` = "Bottom level"
+    ),
+    level = factor(level, levels = c("Top level", "Level 1", "Level 2", "Bottom level"))
+  ) %>%
+  arrange(level, .model)
+
+row_levels <- MSE_tidy$level
+MSE_display <- MSE_tidy %>% select(-level)
+
+# Kable --------------------------------------
+my_kable <- MSE_display %>%
+  mutate(across(where(is.numeric), ~formatC(.x, digits = 1, format = "f"))) %>%
+  kable(
+    booktabs = TRUE,
+    longtable = TRUE,
+    align = c("l", rep("r", 6)),  # now only 7 cols
+    caption = "Out-of-sample forecast results from simulating 45 series with 3 levels",
+    col.names = c(" ", rep(c("h=1", "h=1-8", "h=1-16"), 2)),
+    escape = FALSE
+  ) %>%
+  add_header_above(c(" " = 1, "T = 100" = 3, "T = 300" = 3)) %>%
+  {
+    tbl <- .
+    start <- 1
+    for (lvl in unique(row_levels)) {
+      end <- start + sum(row_levels == lvl) - 1
+      tbl <- tbl %>% pack_rows(lvl, start, end)
+      start <- end + 1
+    }
+    tbl
+  }
+
+my_kable
+```
+
+```{=latex}
+{\footnotesize
+\begin{center}
+  \textbf{NOTES:} Base shows the average MSE of the base forecasts. A negative (positive) entry above this row shows a percentage decrease (increase) in average MSE relative to the base forecasts.
+\end{center}
+}
+```
 

### 2025-05-09 research_proposal/research_proposal_draft.qmd
expanding sim

@@
 
+In this experiment, where the series are contemporaneously correlated, the Minimum Trace approach improves the out-of-sample forecast accuracy tremendously, especially for the levels closer to the aggregated top level.
+Both MinT(Shrink) and MinT(N) consistently outperform the MinT(Sample) at any level and forecasting horizon in terms of accuracy.
+
+The results also show that the MinT(Shrink) and MinT(N) methods are comparable, and almost identical in terms of accuracy.
+The MinT(Shrink) forecasts are slightly more accurate, but again, the difference is negligible.
+This suggests that the NOVELIST estimator might be shrinking the same way as the shrinkage estimator, and a more thorough investigation is needed to understand the differences between them.
+
+In this preliminary experiment, the observations are generated from a different data generating process for each simulation, attempting to assess the predictive accuracy between methods.
+In future experiments, we will consider a fixed data generating process, in order to assess the behaviors of NOVELIST's shrinkage intensity parameter $\lambda_{\delta}$ and threshold $\delta$.
+We will also explore the the effect of correlations and larger hierarchical structures, where the in-sample one-step-ahead base forecast error variance $\hat{\boldsymbol{W}}_1$ is not positive definite.
+
+Regarding other covariance estimators, we will implement the POET and PC-adjusted NOVELIST estimators, and experiment whether the prominent principal component structure can be exploited to improve the out-of-sample forecast performance.
+The project will continue to explore other high-dimensional covariance estimators, and their compatibility with the MinT reconciliation method.
+
 - Put some graphs here

### 2025-05-09 research_proposal/research_proposal_draft.qmd
Update research_proposal_draft.qmd
timeline and milestone

@@
         \usepackage[colorlinks=true, linkcolor=blue, citecolor=blue, urlcolor=blue]{hyperref}
+        \usepackage{array}
     mathspec: true
@@
 
-- Put some graphs here
+\pagebreak
 
@@
 library(knitr)
+library(kableExtra)
 
@@
   Period = c(
-    "March - April", 
+    "March",
+    "April", 
     "May", 
     "June",
-    "July", 
-    "August - September",
+    "July-August", 
+    "September",
     "October"
@@
   Task = c(
-    "Literature review, methods implementation (shrinkage, NOVELIST) + simulation framework",
-    "Simulation + NOVELIST assessment",
-    "Experiments with POET and PC-adjusted NOVELIST",
-    "Exploration of other high-dimensional covariance estimators",
-    "Real-world application + evaluation",
+    "Initial setup and understanding the research scope. Preliminary knowledge and Literature review",
+    "Implementing the Min Trace approach with the shrinkage and NOVELIST estimator.
+Implementing rolling-window cross-validation for choosing NOVELIST threshold parameter.
+Exploring simulation framework.",
+    "Executing simulation on different settings, in order to assess the performance and behaviours of NOVELIST under different conditions. Writing research proposal and presenting research project.",
+    "Implementing POET and PC-adjusted NOVELIST estimators. Assessing whether prominent principal components structure can be exploited to improve the out-of-sample forecast performance.",
+    "Exploring other high-dimensional covariance estimators",
+    "Real-world application + evaluation (e.g. company sales data or Australia tourism data)",
     "Final writing + submission"
@@
   Deliverable = c(
-    "Main codes + paper draft",
-    "Results: NOVELIST",
-    "Results: POET & PC-adjusted NOVELIST",
-    "Results: other estimators",
-    "Application completed",
+    "Literature review + paper draft",
+    "Main codes for estimators + simulation framework + cross-validation",
+    "NOVELIST simulation results + research proposal + presentation",
+    "POET & PC-adjusted NOVELIST simulation results",
+    "Other estimators simulation results",
+    "Application",
     "Thesis manuscript"
@@
 ) %>%
-  kable(caption = "Project Timeline and Key Milestones")
+  kable(
+    format = "latex",
+    booktabs = TRUE,
+    longtable = TRUE,
+    caption = "Project Timeline and Key Milestones",
+    escape = TRUE
+  ) %>%
+  column_spec(1, width = "1.9cm") %>%
+  column_spec(2, width = "8cm") %>%
+  column_spec(3, width = "4.8cm") %>%
+  kable_styling(latex_options = "hold_position", font_size = 10)
 ```
 
-# Expected Contributions
+<!-- # Expected Contributions -->
 

### 2025-05-09 research_proposal/research_proposal_draft.qmd
putting OLS and WLS into appendix

@@
 The choice of $\boldsymbol{G}$ determines the composition of reconciled forecasts $\tilde{\boldsymbol{y}}_h$.
-Methods are developed to estimate $\boldsymbol{G}$, the first attempts at least squares forecast reconciliation were made by @Hyndman2011-jv.
-
-They proposed the optimal $\boldsymbol{G}$ based on the regression model:
-
-$$
-\hat{\boldsymbol{y}}_h = \boldsymbol{S} \boldsymbol{\beta}_h + \boldsymbol{\epsilon}_h,
-$$
-
-where $\boldsymbol{\epsilon}_h = \tilde{\boldsymbol{y}}_h - \hat{\boldsymbol{y}}_h$ is the coherency error with variance $\boldsymbol{V}_h$.
-This led to the GLS solution: $\boldsymbol{G} = (\boldsymbol{S}' \boldsymbol{V}_h^{-1} \boldsymbol{S})^{-1} \boldsymbol{S}' \boldsymbol{V}_h^{-1}$.
-
-However, the covariance matrix $\boldsymbol{V}_h$ is unknown (and later shown by @Wickramasuriya2019-xq to be unidentifiable), and replaced by an identity matrix.
-The method then collapses into an OLS solution: $\boldsymbol{G} = (\boldsymbol{S}' \boldsymbol{S})^{-1} \boldsymbol{S}'$.
-
-An obvious drawback of the OLS solution is that it weights all series in any level equally, regardless of its scale or forecast error variance.
-The issue prompted @Hyndman2016-ic to propose a WLS solution, where the series are weighted by the inverse variances of the base forecast errors.
-The WLS solution is $\boldsymbol{G} = (\boldsymbol{S}' \boldsymbol{\Lambda}_h^{-1} \boldsymbol{S})^{-1} \boldsymbol{S}' \boldsymbol{\Lambda}_h^{-1}$.
-$\boldsymbol{\Lambda}_h = \text{diag} (\boldsymbol{W}_h)$ and $\boldsymbol{W}_h = Var(\boldsymbol{y}_h - \hat{\boldsymbol{y}}_h)$.
+Methods are developed to estimate $\boldsymbol{G}$, including the OLS and WLS developed by @Hyndman2011-jv and @Hyndman2016-ic, respectively.
+A more detailed explanation of these methods can be found in the @sec-ols-wls Appendix.
 
@@
 
+## OLS and WLS Reconciliation {#sec-ols-wls}
+
+The first attempts at least squares forecast reconciliation were made by @Hyndman2011-jv. They proposed the optimal $\boldsymbol{G}$ based on the regression model:
+
+$$
+\hat{\boldsymbol{y}}_h = \boldsymbol{S} \boldsymbol{\beta}_h + \boldsymbol{\epsilon}_h,
+$$
+
+where $\boldsymbol{\epsilon}_h = \tilde{\boldsymbol{y}}_h - \hat{\boldsymbol{y}}_h$ is the coherency error with variance $\boldsymbol{V}_h$.
+This led to the GLS solution: $\boldsymbol{G} = (\boldsymbol{S}' \boldsymbol{V}_h^{-1} \boldsymbol{S})^{-1} \boldsymbol{S}' \boldsymbol{V}_h^{-1}$.
+
+However, the covariance matrix $\boldsymbol{V}_h$ is unknown (and later shown by @Wickramasuriya2019-xq to be unidentifiable), and replaced by an identity matrix.
+The method then collapses into an OLS solution: $\boldsymbol{G} = (\boldsymbol{S}' \boldsymbol{S})^{-1} \boldsymbol{S}'$.
+
+An obvious drawback of the OLS solution is that it weights all series in any level equally, regardless of its scale or forecast error variance.
+The issue prompted @Hyndman2016-ic to propose a WLS solution, where the series are weighted by the inverse variances of the base forecast errors.
+The WLS solution is $\boldsymbol{G} = (\boldsymbol{S}' \boldsymbol{\Lambda}_h^{-1} \boldsymbol{S})^{-1} \boldsymbol{S}' \boldsymbol{\Lambda}_h^{-1}$.
+$\boldsymbol{\Lambda}_h = \text{diag} (\boldsymbol{W}_h)$ and $\boldsymbol{W}_h = Var(\boldsymbol{y}_h - \hat{\boldsymbol{y}}_h)$.
+
 ## Algorithm: NOVELIST cross-validation for optimal threshold $\delta^*$ {#sec-novelist_cv}

### 2025-05-09 research_proposal/tourism_eigen.R
Create tourism_eigen.R

@@ -1,0 +1,68 @@
+library(tidyverse)
+library(fpp3)
+
+tourism_full <- tourism |>
+  aggregate_key((State/Region), Trips = sum(Trips))
+  # aggregate_key((State/Region) * Purpose, Trips = sum(Trips))
+
+fit <- tourism_full |>
+  filter(year(Quarter) <= 2015) |>
+  model(base = ETS(Trips))
+
+augment_fit <- fit %>%
+  augment() %>%
+  filter(.model == "base")
+
+region <- augment_fit |>
+  filter(!is_aggregated(Region)) |>
+  as_tibble() |>
+  select(Quarter, Region, .resid) |>
+  pivot_wider(
+    names_from = Region,
+    values_from = .resid
+  )
+
+state <- augment_fit |>
+  filter(is_aggregated(Region), !is_aggregated(State)) |>
+  as_tibble() |>
+  select(Quarter, State, .resid) |>
+  pivot_wider(
+    names_from = State,
+    values_from = .resid
+  )
+
+total <- augment_fit |>
+  filter(is_aggregated(State)) |>
+  as_tibble() |>
+  select(Quarter, State, .resid) |>
+  pivot_wider(
+    names_from = State,
+    values_from = .resid
+  )
+
+full <- bind_cols(
+  total,
+  state |> select(-Quarter),
+  region |> select(-Quarter)
+)
+
+full |>
+  select(-Quarter) |>
+  cov() |> princomp()
+
+# prcomp(temp) |> screeplot(npcs=24, type="lines")
+
+eigen(temp)$value[1:20] |>
+  plot(
+    ylab = "Log of Eigenvalue",
+    xlab = " ",
+    log = "y",
+  ); abline(v=4.5, lty=2, col="purple")
+
+eigen(temp)$value[1:20] |>
+  plot(
+    ylab = "Log of Eigenvalue",
+    xlab = " ",
+    ylim = c(0, 1e4)
+    # log = "y",
+  ); abline(v=4.5, lty=2, col="purple")

### 2025-05-09 research_proposal/table_function.R
Update table_function.R

@@ -5,7 +5,7 @@ library(dplyr)
 library(purrr)
 library(kableExtra)
 library(knitr)
-
+library(webshot2)
 
 
 sim1 <- readRDS("D:/Github/Recon_Honours_Thesis/research_proposal/sim_results/S36-6-2-1_T100_M500_par.rds")
@@ -128,3 +128,9 @@ my_kable <- MSE_display %>%
   }
 
 my_kable
+
+save_kable(
+  my_kable,
+  file = "D:/Github/Recon_Honours_Thesis/research_proposal/my_kable.png",
+  keep_tex = TRUE
+)

### 2025-05-09 research_proposal/research_proposal_draft.qmd
Add eigenvalue graph and small fixes + add ref

@@
 # Methodology
-
-\color{red}
 
@@
 where $\hat{r}_{ij}$ is the $ij$th element of $\hat{\boldsymbol{R}}_1$, the 1-step-ahead sample correlation matrix (obtained from $\hat{\boldsymbol{W}}_1$).
+The optimal estimate is obtained by minimising $MSE(\hat{\boldsymbol{W}}_1) = Bias(\hat{\boldsymbol{W}}_1)^2 + Var(\hat{\boldsymbol{W}}_1)$.
+More specifically, we trade the unbiasedness of the sample covariance matrix for a lower variance.
 
-The optimal $\lambda_D$ is obtained by minimising the mean squared error of $\hat{\boldsymbol{W}}_1$: $MSE(\hat{\boldsymbol{W}}_1) = Bias(\hat{\boldsymbol{W}}_1)^2 + Var(\hat{\boldsymbol{W}}_1)$.
-More specifically, we trade the unbiasedness of the sample covariance matrix for a lower variance.
-The objective function itself does not take into account any possible principal components structure in the data, and is not flexible enough since it shrinks all off-diagonal elements equally towards zeros.
+However, the hierarchical time series data often exhibit a prominent principal components structure, which is not taken into account by the shrinkage estimator.
+Taking an example of the Australian domestic overnight trips data set [@tourismau-jk], where the national trips are disaggregated into states and territories, and further into regions.
+We then fit ETS models to all series, using the algorithm from Fabletools R package [@fabletools-fa], and compute the one-step-ahead in-sample base forecast error covariance matrix $\hat{\boldsymbol{W}}_1$.
+The twenty largest eigenvalues of the covariance matrix, which we scale by logarithm, are plotted in @fig-eigen.
+We can see that the first 4 eigenvalues are significantly larger than the others, indicating a prominent principal components structure.
 
-- Screeplot of eigenvalues
+![Twenty largest eigenvalues of one-step-ahead in-sample base forecast error covariance, Australian domestic overnight trips](D:/Github/Recon_Honours_Thesis/research_proposal/figs/tourism_eigen.png){#fig-eigen width="70%"}
+
+Additionally, the shrinkage estimator shrinks all off-diagonal elements towards zeros but not completely discard small, noisy correlations.
+This results in a covariance matrix that is not sparse.
+In the next sections, we will explore several options that can take the prominent structure into account, and/or also discard small correlations.
 
@@
 The NOVELIST (NOVEL Integration of the Sample and Thresholded Covariance) estimator, proposed by @Huang2019-ua, is currently the main focus of this research project.
-\textcolor{red}{It introduces a way to control the target matrix's sparsity, retaining strong correlations while discarding weak, noisy correlations.
-NOVELIST offers more flexibility than the shrinkage estimator, however, it does not guarantee to be positive definite.}
+It introduces a way to control the target matrix's sparsity, retaining strong correlations while discarding weak, noisy correlations.
+NOVELIST offers more flexibility than the shrinkage estimator, however, it does not guarantee to be positive definite.
 
@@
 The first 100 or 300 observations are used for training, and the last 16 observations are used for testing.
-The remaining data is used to compute the best fitted ARIMA models by minimising the AICc criterion.
+The remaining data is used to compute the best fitted ARIMA models by minimising the AICc criterion, in which we use the automatic algorithm from Fabletools R package [@fabletools-fa].
 We refer to them as base models, and their base forecasts are then reconciled using the MinT with different covariance estimators.
@@
 
+\pagebreak
+
 ## Algorithm: NOVELIST cross-validation for optimal threshold $\delta^*$ {#sec-novelist_cv}

### 2025-05-10 research_proposal/research_proposal_draft.qmd
super small fix
final touch

@@
 title: "Enhancing Forecast Reconciliation: A Study of Alternative Covariance Estimators"
+author: Vincent Su
 format: 
@@
 We then fit ETS models to all series, using the algorithm from Fabletools R package [@fabletools-fa], and compute the one-step-ahead in-sample base forecast error covariance matrix $\hat{\boldsymbol{W}}_1$.
-The twenty largest eigenvalues of the covariance matrix, which we scale by logarithm, are plotted in @fig-eigen.
-We can see that the first 4 eigenvalues are significantly larger than the others, indicating a prominent principal components structure.
+The twenty largest eigenvalues of the covariance matrix are plotted in @fig-eigen.
+We can see that the point of inflexion occurs at the component with 5th largest eigenvalue, indicating a prominent principal components structure.
 
-![Twenty largest eigenvalues of one-step-ahead in-sample base forecast error covariance, Australian domestic overnight trips](D:/Github/Recon_Honours_Thesis/research_proposal/figs/tourism_eigen.png){#fig-eigen width="70%"}
+![Twenty largest eigenvalues of one-step-ahead in-sample base forecast error covariance, Australian domestic overnight trips](D:/Github/Recon_Honours_Thesis/research_proposal/figs/eigen_tourism.png){#fig-eigen width="70%"}
 
@@
 
-\For{$i = v:T-1$ do}
+\For{$i = v:T-1$}
     \State $j = i - v +1$

### 2025-05-10 research_proposal/tourism_eigen.R

@@ -61,8 +61,9 @@ eigen(temp)$value[1:20] |>
 
 eigen(temp)$value[1:20] |>
   plot(
-    ylab = "Log of Eigenvalue",
-    xlab = " ",
-    ylim = c(0, 1e4)
+    ylab = "Eigenvalue",
+    xlab = "Component",
+    ylim = c(0, 1e5),
+    type = "b"
     # log = "y",
-  ); abline(v=4.5, lty=2, col="purple")
+  ); abline(v=5, lty=2, lwd=1.5, col="purple")

### 2025-05-21 research_proposal/research_proposal_draft.qmd
Refining and error correcting

@@
 In time series forecasting, aggregation occurs in a variety of settings.
-A concrete example of a hierarchy would be electricity demand forecasting, where the national demand is the sum of the demands for each state, and demand for each state comes from many regions within the states.
-Forecasting national tourism or Gross Domestic Product (GDP) is another example of hierarchical time series.
+A concrete example of a hierarchy would be the Starbucks sales data, where their total sales is the sum of all countries they are operating in, and each national sales is the sum for all cities, and the sales for each city come from many outlets.
+Data on national tourism, electricity demand, or Gross Domestic Product (GDP) are similar examples of hierarchical time series.
 The impact of methods for forecasting hierarchical time series has not been limited to academia, with industry also showing a strong interest.
@@
 The top level of the tree represents the total value of all series, while the lower levels represent the series at different levels of disaggregation.
-When there are attributes of interest that are crossed, such as the electricity demand at any aggregation level (national, state, or regional) is also considered by usage purposes (e.g., residential, commercial), the structure is described as a grouped time series (illustrated in @fig-grouped).
+When there are attributes of interest that are crossed, such as the Starbucks drinks sales at any aggregation level (brand-wise, national, city, or outlet) is also considered by kinds of drinks (e.g., coffees, refreshers), the structure is described as a grouped time series (illustrated in @fig-grouped).
 
@@
 
+<!-- \left[ -->
+<!-- \begin{array}{cccc} -->
+<!-- 1 & 1 & 1 & 1 \\ -->
+<!-- 1 & 1 & 0 & 0 \\ -->
+<!-- 0 & 0 & 1 & 1 \\ -->
+<!-- 1 & 0 & 0 & 0 \\ -->
+<!-- 0 & 1 & 0 & 0 \\ -->
+<!-- 0 & 0 & 1 & 0 \\ -->
+<!-- 0 & 0 & 0 & 1 \\ -->
+<!-- \end{array} -->
+<!-- \right] -->
+
 # Methodology
 
-Assume we produce h-step-ahead base forecasts $\hat{\boldsymbol{b}}_h$ for the bottom-level series, obtained by any prediction methods. Then pre-multiplying them by $\boldsymbol{S}$ we get:
+Assume we produce $h$-step-ahead base forecasts $\hat{\boldsymbol{b}}_h$ for the bottom-level series, obtained by any prediction methods. Then pre-multiplying them by $\boldsymbol{S}$ we get:
 
@@
 
-@Wickramasuriya2019-xq reframed the problem as minimising the variances of all reconciled forecasts from @eq-sgy.
-They showed that this is equivalent to minimising the trace of the base forecast error covariance matrix (sum of the diagonal elements - the variances).
+@Wickramasuriya2019-xq reframed the problem as minimising the variances of all reconciled forecast errors.
+They showed that this is equivalent to minimising the trace of the reconciled forecast error covariance matrix (sum of the diagonal elements - the variances).
 This is known as the Minimum Trace (MinT) reconciliation method.
@@
 
-and $\boldsymbol{W}_h$ is the covariance matrix of the h-step-ahead base forecast errors.
+and $\boldsymbol{W}_h$ is the positive definite covariance matrix of the $h$-step-ahead base forecast errors.
 
@@
 
-Additionally, the shrinkage estimator shrinks all off-diagonal elements towards zeros but not completely discard small, noisy correlations.
-This results in a covariance matrix that is not sparse.
-In the next sections, we will explore several options that can take the prominent structure into account, and/or also discard small correlations.
+Additionally, the shrinkage estimator shrinks all off-diagonal elements towards zeros with equal weights.
+We might prefer to better preserve strong signals, and largely discard the effects small, noisy correlations.
+In the next sections, we will explore several options that can take the prominent structure into account, and/or also largely discard the effects of small correlations.
 
@@
 The NOVELIST (NOVEL Integration of the Sample and Thresholded Covariance) estimator, proposed by @Huang2019-ua, is currently the main focus of this research project.
-It introduces a way to control the target matrix's sparsity, retaining strong correlations while discarding weak, noisy correlations.
-NOVELIST offers more flexibility than the shrinkage estimator, however, it does not guarantee to be positive definite.
+It introduces a way to control the target matrix's sparsity, retaining strong correlations while discarding weak, noisy effects.
+NOVELIST offers more flexibility than the shrinkage estimator, which is useful when we believe that only a few variables are truly correlated.
+However, it does not guarantee to be positive definite.
 
@@
 The POET (Principal Orthogonal complEment Thresholding) estimator, proposed by @Fan2013-jz, is another "sparse" + "non-sparse" covariance estimator.
-It takes the latent factors into account, and is appealing when there are common drivers in the time series within the hierarchy.
+It takes the latent factors directly into its construction, and is appealing when there are common drivers in the time series within the hierarchy, as we saw in the Australian tourism example.
 
@@
 
-An obvious drawback of the OLS solution is that it weights all series in any level equally, regardless of its scale or forecast error variance.
+An obvious drawback of the OLS solution is that it assumes an identity matrix for the covariance, ignoring information on the errors' scale and variance.
 The issue prompted @Hyndman2016-ic to propose a WLS solution, where the series are weighted by the inverse variances of the base forecast errors.

### 2025-05-22 research_proposal/research_proposal_draft.qmd
minor fix

@@
 
-![A 2-level hierarchical tree structure](D:/Github/Recon_Honours_Thesis/research_proposal/figs/hierarchical_structure.png){#fig-hierarchy width="40%" height="40%"}
+![A 2-level hierarchical tree structure](D:/Github/Recon_Honours_Thesis/research_proposal/figs/hierarchical_structure.png){#fig-hierarchy width="40%"}
 
-![A 2-level grouped structure, which can be considered as the union of two hierarchical trees with common top and bottom level series](D:/Github/Recon_Honours_Thesis/research_proposal/figs/grouped_structure.png){#fig-grouped width="90%" height="40%"}
+![A 2-level grouped structure, which can be considered as the union of two hierarchical trees with common top and bottom level series](D:/Github/Recon_Honours_Thesis/research_proposal/figs/grouped_structure.png){#fig-grouped width="90%"}
 

### 2025-06-17 literature/paper_draft.qmd
Create paper_draft.qmd

--- 2025-04-28 literature/paper_draft.Rmd
@@
 ---
-title: "Draft - Enhancing Forecast Reconciliation: A Study of Covariance Estimation"
-author: "Vincent Su"
-date: "`r Sys.Date()`"
-output: 
-  pdf_document
-header-includes:
-  - \usepackage{bm}
-  - \usepackage{titlesec}
-  - \titlespacing{\section}{0pt}{30pt plus 12pt minus 1pt}{0pt plus 1pt minus 1pt}
-# bibliography: bibliography.bib
+title: "Enhancing Forecast Reconciliation: A Study of Alternative Covariance Estimators"
+author: Vincent Su
+format: 
+  pdf:
+    include-in-header:
+      text: |
+        <!-- \usepackage{amsmath} -->
+        <!-- \usepackage{mathspec} -->
+        \usepackage{algorithm}
+        \usepackage{algpseudocode}
+        \usepackage[colorlinks=true, linkcolor=blue, citecolor=blue, urlcolor=blue]{hyperref}
+        \usepackage{array}
+    mathspec: true
+    number-sections: true
+execute:
+  cache: true
+bibliography: D:/Github/Recon_Honours_Thesis/references.bib
+# bibliographystyle: apa
+csl: D:/Github/Recon_Honours_Thesis/apa.csl
 ---
-
-```{r setup, include=FALSE}
-knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
-```
 
@@
 
--   Vector $\bm{y}_t$
+-   Vector $\boldsymbol{y}_t$
 
--   Matrix $\bm{S}$
+-   Matrix $\boldsymbol{S}$
 
--   Covariance matrix of in-sample 1-step-ahead base forecast errors $\hat{\bm{W}}_1$
+-   Covariance matrix of in-sample 1-step-ahead base forecast errors $\hat{\boldsymbol{W}}_1$
 
--   Its shrinkage estimator with diagonal target $\hat{\bm{W}}^{shr}_{1, D}$
+-   Its shrinkage estimator with diagonal target $\hat{\boldsymbol{W}}^{shr}_{1, D}$
 
--   Its NOVELIST estimator $\hat{\bm{W}}^{shr}_{1, thr}$
+-   Its NOVELIST estimator $\hat{\boldsymbol{W}}^{shr}_{1, thr}$
 
@@
 
-# 1. Introduction
+# Introduction
 
@@
 It minimises the trace of reconciled forecast errors covariance matrix (minimum variance) and depends critically on the covariance matrix of h-step ahead base forecast errors $W_h$.
+
+## Background
+
+In time series forecasting, aggregation occurs in a variety of settings.
+A concrete example of a hierarchy would be the Starbucks sales data, where their total sales is the sum of all countries they are operating in, and each national sales is the sum for all cities, and the sales for each city come from many outlets.
+Data on national tourism, electricity demand, or Gross Domestic Product (GDP) are similar examples of hierarchical time series.
+The impact of methods for forecasting hierarchical time series has not been limited to academia, with industry also showing a strong interest.
+Many companies have adopted these methods in practice, including Amazon, the International Monetary Fund, IBM, SAP, and more [@Athanasopoulos2024-as].
+
+The hierarchical structure can be represented as a tree, as shown in @fig-hierarchy.
+The top level of the tree represents the total value of all series, while the lower levels represent the series at different levels of disaggregation.
+When there are attributes of interest that are crossed, such as the Starbucks drinks sales at any aggregation level (brand-wise, national, city, or outlet) is also considered by kinds of drinks (e.g., coffees, refreshers), the structure is described as a grouped time series (illustrated in @fig-grouped).
+
+![A 2-level hierarchical tree structure](D:/Github/Recon_Honours_Thesis/research_proposal/figs/hierarchical_structure.png){#fig-hierarchy width="40%"}
+
+![A 2-level grouped structure, which can be considered as the union of two hierarchical trees with common top and bottom level series](D:/Github/Recon_Honours_Thesis/research_proposal/figs/grouped_structure.png){#fig-grouped width="90%"}
+
+For simplicity, we refer to both of these structures as hierarchical time series, we will distinguish between them if and when it is necessary.
+All hierarchical structures can be represented using matrix algebra:
+
+$$
+\boldsymbol{y}_t = \boldsymbol{S} \boldsymbol{b}_t,
+$$
+
+where $\boldsymbol{S}$ is a summing matrix of order $n \times n_b$ which aggregates the bottom-level series $\boldsymbol{b}_t \in \mathbb{R}^{n_b}$ to the series at aggregation levels above.
+The $n$-vector $\boldsymbol{y}_t \in \mathbb{R}^n$ contains all observations at time $t$.
+
+The summing matrix $\boldsymbol{S}$ for the tree structure in @fig-hierarchy is:
+
+$$
+\boldsymbol{S} = 
+\left[
+\begin{array}{cccc}
+1 & 1 & 1 & 1 \\
+1 & 1 & 0 & 0 \\
+0 & 0 & 1 & 1 \\
+\multicolumn{4}{c}{\boldsymbol{I_4}}
+\end{array}
+\right].
+$$
 
@@
 
-# 2. Literature Review
+# Literature Review
 
-## 2.1 Forecast Reconciliation in Hierarchical and Grouped Time Series
+## Forecast Reconciliation in Hierarchical and Grouped Time Series
 
@@
 
-## 2.2 Covariance Estimation in High Dimensions
+## Covariance Estimation in High Dimensions
 
@@
 
-## 2.3 Relevance to Forecast Reconciliation
+## Relevance to Forecast Reconciliation
 
@@
 
-# 3. Methodology
+# Methodology
 
-## 3.1 Theoretical Framework
+## Theoretical Framework
 
@@
 $$
-\bm{y}_t = \bm{S} \bm{b}_t,
+\boldsymbol{y}_t = \boldsymbol{S} \boldsymbol{b}_t,
 $$ 
-where $\bm{S}$ is a summing matrix of order $m x n$ which aggregates the bottom-level series $\bm{b}_t$ ($n$-vector) to the series at aggregation levels above.
-The $m$-vector $\bm{y}_t$ contains all observations at time $t$
+where $\boldsymbol{S}$ is a summing matrix of order $m x n$ which aggregates the bottom-level series $\boldsymbol{b}_t$ ($n$-vector) to the series at aggregation levels above.
+The $m$-vector $\boldsymbol{y}_t$ contains all observations at time $t$
 
@@
 $$
-var[y_{t+h} - \tilde{y}_t(h) | F_t] = SPW_hP'S'
-$$ where $W_h = E[\hat{e}_t(h) \hat{e}'_t(h) | F_t]$ is the covariance matrix of the h-step-ahead base forecast errors.
+Var[y_{t+h} - \tilde{y}_t(h) | F_t] = SPW_hP'S'
+$$
+
+where $W_h = E[\hat{e}_t(h) \hat{e}'_t(h) | F_t]$ is the covariance matrix of the h-step-ahead base forecast errors.
 
@@
 
-## 3.2 Covariance Estimation Approaches
+## Covariance Estimation Approaches
 
@@
 
-# 4. Simulation
+# Simulation
 
@@
 $$
-\bm{y}_t = \bm{A} \bm{y}_{t-1} + \bm{e}_t,
+\boldsymbol{b}_t = \boldsymbol{A} \boldsymbol{b}_{t-1} + \boldsymbol{\epsilon}_t,
 $$
 
-where $\bm{A}$ is a $p \times p$ block diagonal matrix of autoregressive coefficients $\bm{A} = diag(\bm{A}_1, \ldots, \bm{A}_m)$, with each $\bm{A}_i$ being a $p_i \times p_i$ matrix.
+where $\boldsymbol{A}$ is a $n_b \times n_b$ block diagonal matrix of autoregressive coefficients $\boldsymbol{A} = diag(\boldsymbol{A}_1, \ldots, \boldsymbol{A}_m)$, with each $\boldsymbol{A}_i$ being a $n_{b,i} \times n_{b,i}$ matrix.
 The block diagonal structure ensures that the time series are grouped into $m$ groups, with each group having its own autoregressive coefficients.
@@
 
-The model is added with a Gaussian innovation process $\bm{e}_t$, with covariance matrix $\Sigma$. The covariance matrix $\Sigma$ is generated specifically in the following way:
+The model is added with a Gaussian innovation process $\boldsymbol{\epsilon}_t$, with covariance matrix $\Sigma$. The covariance matrix $\Sigma$ is generated specifically in the following way:
 
-1. A compound symmetric correlation matrix is used for each block of size $p_i$ in $\bm{A}_i$, where the coefficients are sampled from a uniform distribution.
+1. A compound symmetric correlation matrix is used for each block of size $n_{b,i}$ in $\boldsymbol{A}_i$, where the coefficients are sampled from a uniform distribution between 0 and 1.
 
-2. The correlations between different blocks are imposed using the Algorithm 1 in Hardin, Garcia & Golan (2013).
+2. The correlations between different blocks are imposed using the Algorithm 1 in @Hardin2013-wu.
 
-3. The covariance matrix $\Sigma$ is then constructed by uniform sampling of standard deviations for all $p$ series.
+3. The covariance matrix $\Sigma$ is then constructed by uniform sampling of standard deviations, in a range of $[\sqrt{2}, \sqrt{6}]$, for all $n_b$ series.
 
 We have an option to randomly flip the signs of the covariance elements, which will create a more realistic structure in the innovation process.
-This is also to simulate the real-world scenario where the observed covariance matrix is not necessarily positive definite.
+This can be done by pre- and post-multiplying $\Sigma$ by a random diagonal matrix $\boldsymbol{V}$ with entries sampled from $\{-1, 1\}$, yielding $\Sigma^* = \boldsymbol{V} \Sigma \boldsymbol{V}$.
 
-# 5. Empirical Analysis
+For each series, $T = 116$ or $316$ observations are generated.
+The first 100 or 300 observations are used for training, and the last 16 observations are used for testing.
+The remaining data is used to compute the best fitted ARIMA models by minimising the AICc criterion, in which we use the automatic algorithm from Fabletools R package [@fabletools-fa].
+We refer to them as base models, and their base forecasts are then reconciled using the MinT with different covariance estimators.
+These include using the unbiased sample covariance matrix - MinT(Sample), the shrinkage estimator - MinT(Shrink), and the NOVELIST estimator - MinT(N).
+The Monte Carlo simulation is repeated $M = 10000$ times, in which the parameters for data generating process is fixed.
+
+# Empirical Analysis
 
@@
 
-# 6. Discussion and Conclusion
+# Discussion and Conclusion
 

### 2025-06-23 literature/paper_draft.qmd
Update paper_draft.qmd
edit the abstract

@@
 
-# Notation
+# Notation {-}
 
@@
 
-# Abstract
+# Abstract {-}
 
-[ This is pasted from the Project Description ]{style="color:red;"}
+[This is pasted from the Project Description]{style="color:red;"}
 
 A collection of time series connected via a set of linear constraints is known as hierarchical time series.
-Forecasting these structures is challenging because the individual forecasts do not satisfy the linear constraints present in the data.
-To mitigate this issue, various forecast reconciliation approaches have been proposed in the literature, where the individual forecasts are adjusted by minimizing the mean squared reconciliation errors.
-Among these, the MinT (Minimum Trace) approach is widely used, and this method requires an estimate of the covariance matrix of the base forecast errors.
+Forecasting these series without respecting the hierarchical nature of the data can lead to incoherent forecasts across aggregation levels and lower accuracy.
+To mitigate this issue, various forecast reconciliation approaches have been proposed in the literature, where the individual forecasts are adjusted to satisfy the aggregation constraints.
+Among these, **MinT** (Minimum Trace) is widely used, however, it requires a good estimate of the covariance matrix of the base forecast errors.
 
@@
 
-In this project, we aim to assess the forecasting performance of MinT when different covariance estimators are used, and compare with the proposed estimators from published paper.
-with a primary focus on the NOVELIST (NOVEL Integration of the Sample and Thresholded Covariance) estimator, which shrinks the sample covariance toward its thresholded version.
-A thorough literature review will also identify other high-dimensional covariance estimators relevant to this project.
+The current practice is to use the shrinkage estimator (often shrinking toward a diagonal matrix), but it lacks flexibility and might not fully utilise the prominent latent structure presented.
+In this project, we aim to assess the forecasting performance of MinT when different covariance estimators are used, namely NOVELIST (NOVEL Integration of the Sample and Thresholded Covariance), POET (Principal Orthogonal complEment Thresholding), and others.
 
@@
 
-    [ ... can include more estimators ... ]{style="color:red;"}
+    [... can include more estimators ...]{style="color:red;"}
 
@@
 <!-- Empirical findings to date   -->
+
 <!-- Athanasopoulos et al.
@@
 
-Hierarchical/Grouped time series satisfy a set of linear constraints, often represented as: 
-$$
+Hierarchical/Grouped time series satisfy a set of linear constraints, often represented as: $$
 \boldsymbol{y}_t = \boldsymbol{S} \boldsymbol{b}_t,
-$$ 
-where $\boldsymbol{S}$ is a summing matrix of order $m x n$ which aggregates the bottom-level series $\boldsymbol{b}_t$ ($n$-vector) to the series at aggregation levels above.
+$$ where $\boldsymbol{S}$ is a summing matrix of order $m x n$ which aggregates the bottom-level series $\boldsymbol{b}_t$ ($n$-vector) to the series at aggregation levels above.
 The $m$-vector $\boldsymbol{y}_t$ contains all observations at time $t$

### 2025-06-23 literature/paper_draft.qmd
abstract edit

@@
 Among these, **MinT** (Minimum Trace) is widely used, however, it requires a good estimate of the covariance matrix of the base forecast errors.
-
-<!-- Although the sample covariance matrix is a natural choice, its performance is greatly impacted by the high-dimensional nature of hierarchical time series encountered in practice.
-Most published research uses shrinkage-type estimators that shrink the sample covariance matrix toward a diagonal matrix.
--->
-
 The current practice is to use the shrinkage estimator (often shrinking toward a diagonal matrix), but it lacks flexibility and might not fully utilise the prominent latent structure presented.

### 2025-06-23 literature/paper_draft.qmd
notations and frameworks for hirerachy

@@
 
--   **Context & Motivation:**
-
-    Hierarchical (and grouped) time series are collections of time series data connected via linear constraints (linear combinations?).
-Forecast reconciliation ensures that forecasts satisfy these constraints.
-
-    One of the widely used appoarch is the MinT approach.
-It minimises the trace of reconciled forecast errors covariance matrix (minimum variance) and depends critically on the covariance matrix of h-step ahead base forecast errors $W_h$.
-
-## Background
+<!-- ############  Background  ############ -->
 
@@
 Many companies have adopted these methods in practice, including Amazon, the International Monetary Fund, IBM, SAP, and more [@Athanasopoulos2024-as].
-
-The hierarchical structure can be represented as a tree, as shown in @fig-hierarchy.
-The top level of the tree represents the total value of all series, while the lower levels represent the series at different levels of disaggregation.
-When there are attributes of interest that are crossed, such as the Starbucks drinks sales at any aggregation level (brand-wise, national, city, or outlet) is also considered by kinds of drinks (e.g., coffees, refreshers), the structure is described as a grouped time series (illustrated in @fig-grouped).
-
-![A 2-level hierarchical tree structure](D:/Github/Recon_Honours_Thesis/research_proposal/figs/hierarchical_structure.png){#fig-hierarchy width="40%"}
-
-![A 2-level grouped structure, which can be considered as the union of two hierarchical trees with common top and bottom level series](D:/Github/Recon_Honours_Thesis/research_proposal/figs/grouped_structure.png){#fig-grouped width="90%"}
-
-For simplicity, we refer to both of these structures as hierarchical time series, we will distinguish between them if and when it is necessary.
-All hierarchical structures can be represented using matrix algebra:
-
-$$
-\boldsymbol{y}_t = \boldsymbol{S} \boldsymbol{b}_t,
-$$
-
-where $\boldsymbol{S}$ is a summing matrix of order $n \times n_b$ which aggregates the bottom-level series $\boldsymbol{b}_t \in \mathbb{R}^{n_b}$ to the series at aggregation levels above.
-The $n$-vector $\boldsymbol{y}_t \in \mathbb{R}^n$ contains all observations at time $t$.
-
-The summing matrix $\boldsymbol{S}$ for the tree structure in @fig-hierarchy is:
-
-$$
-\boldsymbol{S} = 
-\left[
-\begin{array}{cccc}
-1 & 1 & 1 & 1 \\
-1 & 1 & 0 & 0 \\
-0 & 0 & 1 & 1 \\
-\multicolumn{4}{c}{\boldsymbol{I_4}}
-\end{array}
-\right].
-$$
 
@@
 
-# Methodology
+# Theoretical Framework
 
-## Theoretical Framework
+## Hierarchical tructure
 
-### Hierarchical and Grouped Time Series Structure
+The hierarchical structure can be represented as a tree, as shown in @fig-hierarchy.
+The top level of the tree represents the total value of all series, while the lower levels represent the series at different levels of disaggregation.
+When there are attributes of interest that are crossed, such as the Starbucks drinks sales at any aggregation level (brand-wise, national, city, or outlet) is also considered by kinds of drinks (e.g., coffees, refreshers), the structure is described as a grouped time series (illustrated in @fig-grouped).
 
-Hierarchical/Grouped time series satisfy a set of linear constraints, often represented as: $$
-\boldsymbol{y}_t = \boldsymbol{S} \boldsymbol{b}_t,
-$$ where $\boldsymbol{S}$ is a summing matrix of order $m x n$ which aggregates the bottom-level series $\boldsymbol{b}_t$ ($n$-vector) to the series at aggregation levels above.
-The $m$-vector $\boldsymbol{y}_t$ contains all observations at time $t$
+![A 2-level hierarchical tree structure](D:/Github/Recon_Honours_Thesis/research_proposal/figs/hierarchical_structure.png){#fig-hierarchy width="40%"}
 
-Let $\hat{y}_t(h)$ be a vector of h-step-ahead base forecasts for each time series in the collection, using information up to and including time $t$, and stacked in the same order as $y_t$.
-Then we can rewrite the linear reconciliation forecast equation as:
+![A 2-level grouped structure, which can be considered as the union of two hierarchical trees with common top and bottom level series](D:/Github/Recon_Honours_Thesis/research_proposal/figs/grouped_structure.png){#fig-grouped width="90%"}
+
+For simplicity, we refer to both of these structures as hierarchical time series, we will distinguish between them if and when it is necessary.
+All hierarchical structures can be represented using matrix algebra:
 
 $$
-\tilde{y}_t(h) = SP\hat{y}_t(h)
-$$ where $P$ is an $n$x$p$ matrix that maps the base forecasts into bottom-level disaggregated forecasts $\hat{b}_t$.
+\boldsymbol{y}_t = \boldsymbol{S} \boldsymbol{b}_t,
+$$
 
--   Expand on this, include graphs, examples...
+where $\boldsymbol{S}$ is a summing matrix of order $n \times n_b$ which aggregates the bottom-level series $\boldsymbol{b}_t \in \mathbb{R}^{n_b}$ to the series at aggregation levels above.
+The vector $\boldsymbol{y}_t \in \mathbb{R}^n$ contains all observations at time $t$.
+The summing matrix $\boldsymbol{S}$ for the tree structure in @fig-hierarchy is:
 
-### The Minimum Trace (MinT) Reconciliation
+$$
+\boldsymbol{S} = 
+\left[
+\begin{array}{cccc}
+1 & 1 & 1 & 1 \\
+1 & 1 & 0 & 0 \\
+0 & 0 & 1 & 1 \\
+\multicolumn{4}{c}{\boldsymbol{I_4}}
+\end{array}
+\right].
+$$
+
+Assume we produce $h$-step-ahead base forecasts $\hat{\boldsymbol{b}}_{t+h|t}$ for the bottom-level series, obtained by any prediction methods. Then pre-multiplying them by $\boldsymbol{S}$ we get:
+
+$$
+\tilde{\boldsymbol{y}}_{t+h|t} = \boldsymbol{S} \hat{\boldsymbol{b}}_{t+h|t} .
+$$ {#eq-sb}
+
+We refer to $\tilde{\boldsymbol{y}}_{t+h|t}$ as coherent forecasts, as they respect the aggregation structure.
+We also refer to this way of obtaining coherent forecasts by summing the bottom-level forecasts as the bottom-up approach.
+However, generating forecasts this way is anchored only to prediction models at a single level, and will not be utilising the information from other levels.
+This drawback applies to the top-down and middle-out approaches.
+For example, the bottom-level data can be very noisy or even intermittent, and the higher-level data might be smoother due to the aggregation.
+
+Another issue with expressing reconciled methods as in @eq-sb is that it restricts the reconciliation to only single-level approaches.
+Thus, @Hyndman2011-jv suggested a generalised expression for all existing methods, which also provides a framework for new methods to be developed:
+
+$$
+\tilde{\boldsymbol{y}}_{t+h|t} = \boldsymbol{S} \boldsymbol{G} \hat{\boldsymbol{y}}_{t+h|t},
+$$ {#eq-sgy}
+
+for a suitable $n_b \times n$ matrix $\boldsymbol{G}$.
+$\boldsymbol{G}$ maps the base forecasts of all levels $\hat{\boldsymbol{y}}_{t+h|t}$ down into the bottom level, which is then aggregated to the higher levels by $\boldsymbol{S}$.
+The choice of $\boldsymbol{G}$ determines the composition of reconciled forecasts $\tilde{\boldsymbol{y}}_{t+h|t}$, and modern methods are developed to estimate $\boldsymbol{G}$.
+
+<!-- including the OLS and WLS developed by @Hyndman2011-jv and @Hyndman2016-ic, respectively. A more detailed explanation of these methods can be found in the @sec-ols-wls Appendix. -->
+
+## The Minimum Trace (MinT) Reconciliation
 
@@
 
-## Covariance Estimation Approaches
+# Covariance Estimation Approaches
 
@@
 
-### The Shrinkage Estimator
+## The Shrinkage Estimator
 
@@
 
-### The NOVELIST Estimator
+## The NOVELIST Estimator
 

### 2025-06-24 literature/paper_draft.qmd
Expand MinT reconciliation and covariance estimation discussion
Extended the explanation of the Minimum Trace (MinT) reconciliation method, including its algebraic formulation, relationship to GLS, OLS, and WLS, and challenges in estimating the covariance matrix in high-dimensional settings. Added details on the diagonal-target shrinkage estimator, optimal shrinkage intensity, and the principal components structure in hierarchical time series data, with an illustrative example and figure. Set up motivation for exploring alternative covariance estimation approaches.

@@
 $\boldsymbol{G}$ maps the base forecasts of all levels $\hat{\boldsymbol{y}}_{t+h|t}$ down into the bottom level, which is then aggregated to the higher levels by $\boldsymbol{S}$.
-The choice of $\boldsymbol{G}$ determines the composition of reconciled forecasts $\tilde{\boldsymbol{y}}_{t+h|t}$, and modern methods are developed to estimate $\boldsymbol{G}$.
+The choice of $\boldsymbol{G}$ determines the composition of reconciled forecasts $\tilde{\boldsymbol{y}}_{t+h|t}$, and modern reconciliation methods are developed to estimate $\boldsymbol{G}$.
 
@@
 
-*Lemma 1*: ... the covariance matrix of the h-step-ahead reconciled forecast errors is given by
+@Wickramasuriya2019-xq framed the problem as minimising the variances of all reconciled forecast errors $\text{Var}[y_{t+h} - \tilde{y}_{t+h|t}] = \boldsymbol{S} \boldsymbol{G} \boldsymbol{W}_h \boldsymbol{G}' \boldsymbol{S}'$, where $\boldsymbol{W}_h = \mathbb{E}( \hat{\boldsymbol{e}}_{t+h|t} \; \hat{\boldsymbol{e}}_{t+h|t}' )$ is the positive definite covariance matrix of the $h$-step-ahead base forecast errors.
+They showed that this is equivalent to minimising the trace of the reconciled forecast error covariance matrix (sum of the diagonal elements - the variances).
+The Minimum Trace (MinT) solution is given by
 
 $$
-Var[y_{t+h} - \tilde{y}_t(h) | F_t] = SPW_hP'S'
+\boldsymbol{G} = (\boldsymbol{S}' \boldsymbol{W}_h^{-1} \boldsymbol{S})^{-1}
+\boldsymbol{S}' \boldsymbol{W}_h^{-1} .
 $$
 
-where $W_h = E[\hat{e}_t(h) \hat{e}'_t(h) | F_t]$ is the covariance matrix of the h-step-ahead base forecast errors.
+@Wickramasuriya2019-xq also showed that MinT is an algebraic generalisation of the GLS, and the OLS and WLS methods are special cases of MinT when $\boldsymbol{W}_h$ is an identity matrix $I_{n_b}$ and a diagonal matrix $\text{diag}( \boldsymbol{W}_h )$, respectively.
+In this paper, we place our main focus on the MinT method.
 
-The MinT reconciled forecast is obtained by minimizing the trace of the above covariance matrix: $$
-\min \; tr(SPW_hP'S') \quad \text{subject to } PS = I.
+The MinT solution hinges on a reliable, positive-definite estimate of $\boldsymbol{W}_h$, which is challenging to estimate in high-dimensional setting.
+The sample covariance matrix is unstable and non-positive-definite when the number of series $n$ is huge and larger than the time dimension $T$.
+To tackle this issue, the original paper @Wickramasuriya2019-xq adopted the diagonal-target shrinkage estimator from @Schafer2005-yw, given by
+
+$$
+\hat{\boldsymbol{W}}^{shr}_{1} = \lambda_D \hat{\boldsymbol{W}}_{1, D} + (1 - \lambda_D) \hat{\boldsymbol{W}}_1,
 $$
 
-which gives the optimal reconciliation matrix: $$
-P = (S' W_h^{-1} S)^{-1} S' W_h^{-1}
+where $\hat{\boldsymbol{W}}_{1, D}$ is a diagonal matrix comprising the diagonal entries  $\text{diag}( \hat{\boldsymbol{W}}_1 )$.
+We refer to any $\lambda \in [0,1]$ as the shrinkage intensity parameter, the subscript specifies which estimator it belongs to.
+This approach shrinks the covariance matrix $\hat{\boldsymbol{W}}_1$ towards its diagonal matrix, meaning the off-diagonal elements are shrunk towards zero while the diagonal ones remain unchanged.
+
+@Schafer2005-yw also proposed an estimate of the optimal shrinkage intensity parameter $\lambda_D$:
+
 $$
+\hat{\lambda}_D = \frac{\sum_{i \neq j} \widehat{Var}(\hat{r}_{ij})} 
+{\sum_{i \neq j} \hat{r}_{ij}^2}
+$$
+
+where $\hat{r}_{ij}$ is the $ij$th element of $\hat{\boldsymbol{R}}_1$, the 1-step-ahead sample correlation matrix (obtained from $\hat{\boldsymbol{W}}_1$).
+The optimal estimate is obtained by minimising $MSE(\hat{\boldsymbol{W}}_1) = Bias(\hat{\boldsymbol{W}}_1)^2 + Var(\hat{\boldsymbol{W}}_1)$.
+More specifically, we trade the unbiasedness of the sample covariance matrix for a lower variance.
+
+However, the hierarchical time series data often exhibit a prominent principal components structure, which is not fully taken advantage.
+Taking an example of the Australian domestic overnight trips data set [@tourismau-jk], where the national trips are disaggregated into states and territories, and further into regions.
+We then fit ETS models to all series, using the algorithm from Fabletools R package [@fabletools-fa], and compute the one-step-ahead in-sample base forecast error covariance matrix $\hat{\boldsymbol{W}}_1$.
+The twenty largest eigenvalues of the covariance matrix are plotted in @fig-eigen.
+We can see that the point of inflexion occurs at the component with 5th largest eigenvalue, indicating a prominent principal components structure.
+
+![Twenty largest eigenvalues of one-step-ahead in-sample base forecast error covariance, Australian domestic overnight trips](D:/Github/Recon_Honours_Thesis/research_proposal/figs/eigen_tourism.png){#fig-eigen width="70%"}
+
+Additionally, the shrinkage estimator shrinks all off-diagonal elements towards zeros with equal weights $\lambda_D$.
+We might prefer to better preserve strong signals, and largely reduce the effects of small, noisy correlations.
+In the next sections, we will explore several options that take these two issues into account.
 

### 2025-06-30 literature/paper_draft.qmd
an outline for literature review

@@
 Many companies have adopted these methods in practice, including Amazon, the International Monetary Fund, IBM, SAP, and more [@Athanasopoulos2024-as].
+<!-- ##########  Literature Review  ########## -->
+
+- Talk about the history and evolution of forecasting hierarchical time series, starting from the early heuristic methods to the modern statistical approaches.
+
+  - Heuristic methods
+  - Optimal combination methods (OLS, WLS)
+  - MinT
+  - Bayesian, Machine learning
+  - Probabilistic methods
+
+- Discuss the interests in MinT and how it has become a standard method for forecast reconciliation.
+
+- Discuss the MinT's reliance on a good estimate of the covariance matrix of base forecast errors and other gaps
+
+  - Empirical evidence of MinT under perform
+  - Comparison with other methods
+
+- Explain why this paper focus on exploring alternative covariance estimators for MinT. And is there any paper talk about this.
+
+  - Is better estimate of W_h really lead to better performance?
+
+- Talk about the paper outline and what will be covered in the following sections.
 
@@
 
-## Hierarchical tructure
+## Hierarchical tructure {#sec-hierarchy}
 
@@
 $$
-\tilde{\boldsymbol{y}}_{t+h|t} = \boldsymbol{S} \hat{\boldsymbol{b}}_{t+h|t} .
+\tilde{\boldsymbol{y}}_{t+h|t} = \boldsymbol{S} \hat{\boldsymbol{b}}_{t+h|t} \, .
 $$ {#eq-sb}
@@
 $$
-\tilde{\boldsymbol{y}}_{t+h|t} = \boldsymbol{S} \boldsymbol{G} \hat{\boldsymbol{y}}_{t+h|t},
+\tilde{\boldsymbol{y}}_{t+h|t} = \boldsymbol{S} \boldsymbol{G} \hat{\boldsymbol{y}}_{t+h|t} \; ,
 $$ {#eq-sgy}
@@
 $$
-\hat{\boldsymbol{W}}^{shr}_{1} = \lambda_D \hat{\boldsymbol{W}}_{1, D} + (1 - \lambda_D) \hat{\boldsymbol{W}}_1,
+\hat{\boldsymbol{W}}^{shr}_{1} = \lambda_D \hat{\boldsymbol{W}}_{1, D} + (1 - \lambda_D) \hat{\boldsymbol{W}}_1 \, ,
 $$
@@
 \hat{\lambda}_D = \frac{\sum_{i \neq j} \widehat{Var}(\hat{r}_{ij})} 
-{\sum_{i \neq j} \hat{r}_{ij}^2}
+{\sum_{i \neq j} \hat{r}_{ij}^2} \, ,
 $$
@@
 # Covariance Estimation Approaches
-
--   $W_h$ is challenging to estimate, when $h>1$, we need alternative estimates.
-
-For forecast errors $\hat{e}_{t|t-1}$, the unbiased sample covariance of in-sample one-step-ahead base forecast errors is:
-
-$$
-\hat{W}_1 = \frac{1}{T} \sum_{t=1}^{T} \hat{e}_{t|t-1} \hat{e}_{t|t-1}^{\top}.
-$$
-
-We reconstruct estimator of $W_h$ as $\hat{W}_h = k_h \, g(\hat{W}_1)$, where $k_h > 0$ simplifying the computation.
-
-We will look into a few "sparse" + "non-sparse" approaches that apply on $\hat{W}_1$.
-
-## The Shrinkage Estimator
-
-$$
-g(\hat{W}_1) = \hat{W}^{shr}_{1, D} = \lambda_D \hat{W}_{1, D} + (1 - \lambda_D) \hat{W}_1
-$$ is the shrinkage estimator with diagonal target, proposed by Schäfer & Strimmer (2005).
-$\hat{W}_{1, D}$ is a diagonal matrix comprising the diagonal entries of $\hat{W}_1$.
-This approach will shrink the covariance matrix $\hat{W}_1$ towards its diagonal version, meaning the off-diagonal elements are shrunk towards zero while the diagonal ones remain unchanged.
-
-They also proposed an optimal shrinkage intensity parameter $\lambda_D$ for this setting, assuming the variances are constant:
-
-$$
-\hat{\lambda}_D = \frac{\sum_{i \neq j} \widehat{Var}(\hat{r}_{ij})} {\sum_{i \neq j} \hat{r}_{ij}}
-$$
-
-where $\hat{r}_{ij}$ is the $ij$th element of $\hat{\boldsymbol{R}}_1$, the 1-step-ahead sample correlation matrix (obtained from $\hat{\boldsymbol{W}}_1$) to shrink it toward an identity matrix
-
--   (need more explanation on how to obtain optimal param)
-
--   Walkthrough the shrinkage estimator method
 

### 2025-06-30 literature/paper_draft.qmd
More detailed Intro

@@
 In time series forecasting, aggregation occurs in a variety of settings.
-A concrete example of a hierarchy would be the Starbucks sales data, where their total sales is the sum of all countries they are operating in, and each national sales is the sum for all cities, and the sales for each city come from many outlets.
-Data on national tourism, electricity demand, or Gross Domestic Product (GDP) are similar examples of hierarchical time series.
+While a formal definition of hierarchical time series can be found in @sec-hierarchy, we can think of Starbucks sales data as an illustrative example.
+Starbucks operates in many countries, and each country has multiple cities where they have outlets.
+The sales data is structured hierarchically: the top level is the total sales across all countries, followed by national sales for each country, then city sales for each city within a country, and finally outlet sales for each outlet in a city.
+As a result, there are over 40,000 individual outlet sales to forecast, plus additional series at higher levels of aggregation such as city and country.
+The hierarchy can be even more complex if we consider the sales of different kinds of drinks (e.g., coffees, teas, refreshers) at each aggregation level.
+
+This hierarchical structure is not unique to the Starbucks sales data; it can be found in many other domains, such as national tourism, electricity demand, or Gross Domestic Product (GDP).
 The impact of methods for forecasting hierarchical time series has not been limited to academia, with industry also showing a strong interest.
-Many companies have adopted these methods in practice, including Amazon, the International Monetary Fund, IBM, SAP, and more [@Athanasopoulos2024-as].
+Many companies and organisations have adopted these methods in practice, including Amazon, the International Monetary Fund, IBM, SAP, and more [@Athanasopoulos2024-as].
+
 <!-- ##########  Literature Review  ########## -->

### 2025-06-30 literature/paper_draft.qmd
Talk about single-level approaches

@@
 
-<!-- ############  Background  ############ -->
+<!-- ############  Intro  ############ -->
 
@@
 
-  - Heuristic methods
+  - Single level methods
   - Optimal combination methods (OLS, WLS)
@@
   - Probabilistic methods
+
+Traditionally, forecasting these hierarchical time series has been done using single-level methods, such as bottom-up, top-down, and middle-out approaches.
+Bottom-up methods involve generating forecasts for the bottom-level series and aggregating them to higher levels.
+Top-down methods start with forecasts for the only top level and disaggregate them down.
+Middle-out methods combine both approaches by forecasting a middle level and then aggregating or disaggregating as needed.
+Despite their simplicity, these methods only anchor forecasts to a single level, implying a large loss of information on the hierarchy's inherent correlation structure.
+Furthermore, as we will see in @sec-hierarchy, the disaggregation becomes more complex when structure is grouped since the disaggregation paths are not unique.
+
+<!-- Include more details on the gaps for each of these methods, such as bias, high variance? -->
 
@@
 The top level of the tree represents the total value of all series, while the lower levels represent the series at different levels of disaggregation.
-When there are attributes of interest that are crossed, such as the Starbucks drinks sales at any aggregation level (brand-wise, national, city, or outlet) is also considered by kinds of drinks (e.g., coffees, refreshers), the structure is described as a grouped time series (illustrated in @fig-grouped).
+When there are attributes of interest that are crossed, such as the Starbucks drinks sales at any aggregation level (brand-wise, national, city, or outlet) is also considered by kinds of drinks (e.g., coffees, refreshers), the structure is described as a grouped time series.
+As illustrated in @fig-grouped, the aggregation or disaggregation paths are not unique.
 
@@
 We also refer to this way of obtaining coherent forecasts by summing the bottom-level forecasts as the bottom-up approach.
-However, generating forecasts this way is anchored only to prediction models at a single level, and will not be utilising the information from other levels.
+However, generating forecasts this way is anchored only to prediction models at a single level, and will not be utilising the inherent information from other levels.
 This drawback applies to the top-down and middle-out approaches.

### 2025-07-04 literature/paper_draft.qmd
small fix

@@
 Despite their simplicity, these methods only anchor forecasts to a single level, implying a large loss of information on the hierarchy's inherent correlation structure.
-Furthermore, as we will see in @sec-hierarchy, the disaggregation becomes more complex when structure is grouped since the disaggregation paths are not unique.
+Furthermore, as we saw from the Starbucks example considering the sales of different kinds of drinks at each aggregation level -- formally defined as grouped structure in @sec-hierarchy -- the disaggregation becomes more complex since the disaggregation paths are not unique.
 

### 2025-07-18 literature/paper_draft.qmd
outline for simulation reports

@@
 Bottom-up methods involve generating forecasts for the bottom-level series and aggregating them to higher levels.
-Top-down methods start with forecasts for the only top level and disaggregate them down.
-Middle-out methods combine both approaches by forecasting a middle level and then aggregating or disaggregating as needed.
+Top-down methods start with forecasts for the only top-level series and disaggregate them down.
+Middle-out methods combine both approaches by forecasting middle-level series and then aggregating or disaggregating as needed.
 Despite their simplicity, these methods only anchor forecasts to a single level, implying a large loss of information on the hierarchy's inherent correlation structure.
+Additionally, the most disaggregated series often are very noisy or even intermittent, and the higher-level data might be smoother due to the aggregation.
 Furthermore, as we saw from the Starbucks example considering the sales of different kinds of drinks at each aggregation level -- formally defined as grouped structure in @sec-hierarchy -- the disaggregation becomes more complex since the disaggregation paths are not unique.
+Consequently, these single-level methods tend to give poor results across other levels of the hierarchy.
+
+To mitigate these issues, forecast reconciliation was introduced by @Hyndman2011-jv, and later developed by @Hyndman2016-ic, @Wickramasuriya2019-xq, and ...
+to achieve coherency in point forecasts and enhance accuracy.
+Forecast reconciliation projects a collection of independent base forecasts into a set of coherent forecasts that respect the linear constraints defining a hierarchical or grouped time-series system.
 
@@
 
--   Talk about description of the hierarchical time series data set (e.g., economic, financial, or synthetic data).
--   Characteristics such as dimensionality, frequency, and hierarchical structure.
+<!-- -   Talk about description of the hierarchical time series data set (e.g., economic, financial, or synthetic data). -->
+<!-- -   Characteristics such as dimensionality, frequency, and hierarchical structure. -->
 
-**Experimental Design**
+<!-- **Experimental Design** -->
 
--   Design different cases of simulation studies and real-data experiments.
--   Metrics: Forecast accuracy (e.g., RMSE, MAE), reconciliation error reduction, and matrix stability...
--   Visualisation...
+<!-- -   Design different cases of simulation studies and real-data experiments. -->
+<!-- -   Metrics: Forecast accuracy (e.g., RMSE, MAE), reconciliation error reduction, and matrix stability... -->
+<!-- -   Visualisation... -->
 
-**General Set-up I'm currently working on**
-
-The designed data generating process for bottom-level series is a stationary VAR(1) process, with the following structure:
+The general design of data generating process for bottom-level series is a stationary VAR(1) process, with the following structure:
 
@@
 The block diagonal structure ensures that the time series are grouped into $m$ groups, with each group having its own autoregressive coefficients.
-This aim to simulate the interdependencies between the time series within each group, where reconciliation will be better performed than the usual base forecasts.
+This aim to simulate the interdependencies between the time series within each group, where reconciliation will be expected to better performed than the usual base forecasts.
 
-The model is added with a Gaussian innovation process $\boldsymbol{\epsilon}_t$, with covariance matrix $\Sigma$. The covariance matrix $\Sigma$ is generated specifically in the following way:
+The model is added with a Gaussian innovation process $\boldsymbol{\epsilon}_t$, with covariance matrix $\boldsymbol{\Sigma}$.
+The covariance matrix $\boldsymbol{\Sigma}$ is generated specifically using the Algorithm 1 in @Hardin2013-wu:
 
-1.  A compound symmetric correlation matrix is used for each block of size $n_{b,i}$ in $\boldsymbol{A}_i$, where the coefficients are sampled from a uniform distribution between 0 and 1.
+1.
+A compound symmetric correlation matrix is used for each block of size $n_{b,i}$ in $\boldsymbol{A}_i$, where the entries $\rho_i$ for each block $i$ are sampled from a uniform distribution between 0 and 1.
+They are baseline correlations within group.
 
-2.  The correlations between different blocks are imposed using the Algorithm 1 in @Hardin2013-wu.
+2. A constant correlation, which is smaller than $\text{min} \{\rho_1, \rho_2, \dots, \rho_m \}$, is imposed on the entries between different blocks. It serves as baseline correlations between group.
 
-3.  The covariance matrix $\Sigma$ is then constructed by uniform sampling of standard deviations, in a range of $[\sqrt{2}, \sqrt{6}]$, for all $n_b$ series.
+3. The entry-wise random noise is added on top of the entire correlation matrix.
 
-We have an option to randomly flip the signs of the covariance elements, which will create a more realistic structure in the innovation process.
-This can be done by pre- and post-multiplying $\Sigma$ by a random diagonal matrix $\boldsymbol{V}$ with entries sampled from $\{-1, 1\}$, yielding $\Sigma^* = \boldsymbol{V} \Sigma \boldsymbol{V}$.
+4. The covariance matrix $\boldsymbol{\Sigma}$ is then constructed by uniform sampling of standard deviations, in a range of $[\sqrt{2}, \sqrt{6}]$, for all $n_b$ series.
+
+We will randomly flip the signs of the covariance elements, which will create a more realistic structure in the innovation process.
+This can be done by pre- and post-multiplying $\boldsymbol{\Sigma}$ by a random diagonal matrix $\boldsymbol{V}$ with diagonal entries sampled from $\{-1, 1\}$, yielding $\boldsymbol{\Sigma}^* = \boldsymbol{V} \boldsymbol{\Sigma} \boldsymbol{V}$.
 
@@
 The Monte Carlo simulation is repeated $M = 10000$ times, in which the parameters for data generating process is fixed.
+
+## Exploring Effects of Hierarchy's Size
+
+- 2x2, 6x6, 50x2
+
+- Finding: Larger size MinT_n performs well
+
+## Exploring Effects of Hierarchy's Aggregation
+
+- 50x2 with usual path and weird path
+
+- Finding: Complex aggregation MinT_n performs well
+
+## Exploring h-step-ahead forecast error metric in NOVELIST Cross-validation
+
+Instead of forecasting t+1 for each rolling window iteration, I forecast t+1 to t+h for each iteration and compute the MSE for each threshold value delta.
+
+Results barely show any difference with the original method.
+
+## Exploring the sparsity of the DGP covariance matrix
+
+Not much diff
+
+## Exploring Grouped Structure
+
+- small vs big
+
+- Finding: NOVELIST performs well in both cases
 

### 2025-07-21 literature/paper_draft.qmd
hierarchy size writing

@@
   cache: true
+  echo: false
+  warning: false
 bibliography: D:/Github/Recon_Honours_Thesis/references.bib
@@
 
-[This is pasted from the Project Description]{style="color:red;"}
+<!-- [This is pasted from the Project Description]{style="color:red;"} -->
 
@@
 
-To mitigate these issues, forecast reconciliation was introduced by @Hyndman2011-jv, and later developed by @Hyndman2016-ic, @Wickramasuriya2019-xq, and ...
+To overcome these issues, forecast reconciliation was introduced by @Hyndman2011-jv, and later developed by @Hyndman2016-ic, @Wickramasuriya2019-xq, and ...
 to achieve coherency in point forecasts and enhance accuracy.
@@
 
+Using the data generating process described above, we consider three different 2-level hierarchical structures, with the bottom-level series $n_b$ being 2 groups of 2 (4 series), 6 groups of 6 (36 series), and 2 groups of 50 (100 series), respectively.
+The first structure is exactly the same as the one in @fig-hierarchy, with 2 series at the level 1.
+The second structure has 36 bottom series and are aggregated in groups of size 6 to form 6 level-1 series, which are finally aggregated to form the top level.
+The third structure is larger, having 100 bottom series aggregated in groups of size 50 to form 2 level-1 series, then aggregated to get the top level.
+Because of this structure, the VAR(1) coefficients matrix would have entries between groups being zeros.
+To save space, we will only illustrate the settings of the second structure (6 groups of 6) in @fig-settings-6x6, and the first and third structure is similar.
+
+```{r load results}
+#| include: false
+library(ggplot2)
+library(patchwork)
+library(ReconCov)
+
+setwd(getwd())
+
+# sim1_1 <- readRDS("./lib/S4-2-1_T100_M10000_par.rds")
+# sim1_2 <- readRDS("./lib/S4-2-1_T300_M10000.rds")
+# 
+sim2_1 <- readRDS("./lib/S36-6-1_T100_M100_run1.rds")
+sim2_2 <- readRDS("./lib/S36-6-1_T300_M100_run2.rds")
+
+# sim3_1 <- readRDS("./lib/S100-2-1_T50_M200_run1.rds")
+# sim3_2 <- readRDS("./lib/S100-2-1_T100_M200_run1.rds")
+```
+
+```{r generate plots for A and Sigma}
+# p1_A <- plot_heatmap(sim1_1$A, T) + 
+#   ggtitle("2x2 VAR(1) Coefficients Matrix")
+# 
+# p1_Sigma <- plot_heatmap(cov2cor(sim1_1$Sigma), T) + 
+#   ggtitle("2x2 Correlation Matrix")
+# 
+p2_A <- plot_heatmap(sim2_1$A) +
+  ggtitle("VAR(1) Coefficients Matrix") + theme_void()
+
+p2_Sigma <- plot_heatmap(cov2cor(sim2_1$Sigma), T) +
+  ggtitle("Correlation Matrix") + theme_void()
+
+# p3_A <- plot_heatmap(sim3_1$A) +
+#   ggtitle("50x2 VAR(1) Coefficients Matrix")
+# 
+# p3_Sigma <- plot_heatmap(cov2cor(sim3_1$Sigma), T) +
+#   ggtitle("50x2 Correlation Matrix")
+```
+
+```{r}
+#| out.width: "30%"
+#| fig.align: "center"
+# (p1_A | p1_Sigma) +
+#   plot_layout(guides = "collect") & theme(legend.position = "right")
+```
+
+```{r}
+#| out.width: "60%"
+#| fig.align: "center"
+# (p2_A | p2_Sigma) +
+#   plot_layout(guides = "collect") & theme(legend.position = "right")
+```
+
+```{r}
+#| out.width: "100%"
+#| fig.align: "center"
+# (p3_A | p3_Sigma) +
+#   plot_layout(guides = "collect") & theme(legend.position = "right")
+```
+
+```{r fig-settings-6x6}
+#| label: fig-settings-6x6
+#| fig-cap: "Heatmaps of the VAR(1) coefficient matrix and correlation matrix for the 6×6 structure."
+#| out.width: "85%"
+#| fig.align: "center"
+
+p2_A + p2_Sigma + plot_layout(ncol = 2) & theme(legend.position = "bottom")
+```
+
 - Finding: Larger size MinT_n performs well
@@
 
-## Exploring h-step-ahead forecast error metric in NOVELIST Cross-validation
+## h-step-ahead forecast error metric in NOVELIST Cross-validation
 

### 2025-07-21 literature/paper_draft.qmd
codes for plotting sim results

@@
 csl: D:/Github/Recon_Honours_Thesis/apa.csl
+editor_options: 
+  chunk_output_type: console
 ---
@@
 
-For each series, $T = 116$ or $316$ observations are generated.
-The first 100 or 300 observations are used for training, and the last 16 observations are used for testing.
-The remaining data is used to compute the best fitted ARIMA models by minimising the AICc criterion, in which we use the automatic algorithm from Fabletools R package [@fabletools-fa].
-We refer to them as base models, and their base forecasts are then reconciled using the MinT with different covariance estimators.
-These include using the unbiased sample covariance matrix - MinT(Sample), the shrinkage estimator - MinT(Shrink), and the NOVELIST estimator - MinT(N).
-The Monte Carlo simulation is repeated $M = 10000$ times, in which the parameters for data generating process is fixed.
-
 ## Exploring Effects of Hierarchy's Size
@@
 library(ReconCov)
+library(dplyr)
 
-setwd(getwd())
+setwd("D:/Github/Recon_Honours_Thesis")
 
-# sim1_1 <- readRDS("./lib/S4-2-1_T100_M10000_par.rds")
-# sim1_2 <- readRDS("./lib/S4-2-1_T300_M10000.rds")
-# 
-sim2_1 <- readRDS("./lib/S36-6-1_T100_M100_run1.rds")
-sim2_2 <- readRDS("./lib/S36-6-1_T300_M100_run2.rds")
+sim1_1 <- readRDS("./literature/lib/S4-2-1_T100_M10000_par.rds")
+sim1_2 <- readRDS("./literature/lib/S4-2-1_T300_M10000.rds")
 
-# sim3_1 <- readRDS("./lib/S100-2-1_T50_M200_run1.rds")
-# sim3_2 <- readRDS("./lib/S100-2-1_T100_M200_run1.rds")
+sim2_1 <- readRDS("./literature/lib/S36-6-1_T100_M100_run1.rds")
+sim2_2 <- readRDS("./literature/lib/S36-6-1_T300_M100_run2.rds")
+
+sim3_1 <- readRDS("./literature/lib/S100-2-1_T50_M200_run1.rds")
+sim3_2 <- readRDS("./literature/lib/S100-2-1_T100_M200_run1.rds")
+
+MSE1_1 <- transform_sim_MSE(sim1_1$MSE, F)
+MSE1_2 <- transform_sim_MSE(sim1_2$MSE, F)
+
+MSE2_1 <- transform_sim_MSE(sim2_1$MSE, F)
+MSE2_2 <- transform_sim_MSE(sim2_2$MSE, F)
+
+MSE3_1 <- transform_sim_MSE(sim3_1$MSE, F)
+MSE3_2 <- transform_sim_MSE(sim3_2$MSE, F)
 ```
@@
 
-- Finding: Larger size MinT_n performs well
+For each series, $T = 54$ and $304$ observations are generated.
+The first 50 and 300 observations are used for training, and the last 4 observations are used for testing.
+The training data is used to compute the best fitted ARIMA models by minimising the AICc criterion, in which we use the automatic algorithm from Fabletools R package [@fabletools-fa].
+We refer to them as base models, and their base forecasts are then reconciled using the MinT with different covariance estimators.
+These include using the unbiased sample covariance matrix - mint_sample, the shrinkage estimator - mint_shr, and the NOVELIST estimator - mint_n.
+The Monte Carlo simulation is repeated $M = 200$ times, in which the parameters for data generating process is fixed.
+
+```{r}
+plot_relative_improvement <- function(MSE_data) {
+  MSE_data |>
+    filter(
+      .model != "mint_true",
+      h <= 4
+    ) |>
+    group_by(series, h) |>
+    mutate(
+      base_MSE = MSE[.model == "base"]
+    ) |>
+    ungroup() |>
+    group_by(.model, h) |>
+    summarise(
+      MSE = mean(MSE),
+      base_MSE = mean(base_MSE),
+      pct_change = (MSE - base_MSE) / base_MSE * 100
+    ) |>
+    ggplot(aes(x = h, y = pct_change, color = .model)) +
+    geom_line() +
+    labs(
+      x = "Forecast Horizon (h)",
+      y = "Percentage Change (%)"
+    ) +
+    # labs(x = element_blank(), y = element_blank()) +
+    theme_minimal()
+}
+```
+
+```{r fig-sim-results-1}
+#| label: fig-sim-results-1
+#| fig-cap: "Relative improvement of the MSE of reconciled forecasts over the base forecasts for the 2x2, 6x6, and 50x2 hierarchical structures, for 1 to 4 steps ahead forecasts, with 2 time series lengths (T = 54 and T = 304)."
+#| fig.width: 8
+#| fig.height: 6
+#| out.width: "100%"
+#| fig.align: "center"
+
+p_MSE1_1 <- MSE1_1 |> 
+  plot_relative_improvement() +
+    labs(title = "2x2 Hierarchy", subtitle = "T = 54")
+p_MSE2_1 <- MSE2_1 |> 
+  plot_relative_improvement() +
+    labs(title = "6x6 Hierarchy")
+p_MSE3_1 <- MSE3_1 |> 
+  plot_relative_improvement() +
+    labs(title = "50x2 Hierarchy")
+
+p_MSE1_2 <- MSE1_2 |> 
+  plot_relative_improvement() +
+    labs(subtitle = "T = 304")
+p_MSE2_2 <- MSE2_2 |> 
+  plot_relative_improvement()
+p_MSE3_2 <- MSE3_2 |> 
+  plot_relative_improvement()
+
+p_MSE1_1 + p_MSE2_1 + p_MSE3_1 + p_MSE1_2 + p_MSE2_2 + p_MSE3_2 +
+  plot_layout(ncol = 3, guides = "collect", axis_titles = "collect") &
+  theme(legend.position = "right")
+```
+
+In this setting, we aim to assess the performance of MinT with NOVELIST when the size of the hierarchy becomes larger. As shown in @fig-sim-results-1, ...
 

### 2025-07-22 literature/paper_draft.qmd
paste the cov ests section from research proposal

@@
 
-## The NOVELIST Estimator
+## NOVELIST Estimator
+
+The NOVELIST (NOVEL Integration of the Sample and Thresholded Covariance) estimator, proposed by @Huang2019-ua, is currently the main focus of this research project.
+It introduces a way to control the target matrix's sparsity, retaining strong correlations while discarding weak, noisy effects.
+NOVELIST offers more flexibility than the shrinkage estimator, which is useful when we believe that only a few variables are truly correlated.
+However, it does not guarantee to be positive definite.
+
+The method is based on the idea of soft-thresholding the sample covariance matrix, then performing shrinkage towards this thresholded version.
+This introduces an extra parameter, the threshold $\delta$, which is used to control the amount of soft-thresholding.
+The NOVELIST estimator is given by:
 
 $$
-g(\hat{W}_1) = \hat{W}^{shr}_{1, thr} = \lambda_{\delta} \hat{W}_{1, \delta} + (1 - \lambda_{\delta}) \hat{W}_1
-$$
+\hat{\boldsymbol{W}}^{N}_{1} = \lambda_{\delta} \hat{\boldsymbol{W}}_{1, \delta} + (1 - \lambda_{\delta}) \hat{\boldsymbol{W}}_1,
+$$ {#eq-novelist-cov}
 
-is the NOVELIST shrinkage estimator, proposed by Huang & Fryzlewicz (2019). By convenient setting, we rewrite as sample correlation:
+where $\hat{\boldsymbol{W}}_{1, \delta}$ is the thresholded version of $\hat{\boldsymbol{W}}_{1}$. By convenient setting, we can rewrite it in terms of correlation:
 
 $$
-\hat{R}^{shr}_{1, thr} = \lambda_{\delta} \hat{R}_{1,\delta} + (1 - \lambda_{\delta}) \hat{R}_1,
-$$
+\hat{\boldsymbol{R}}^{N}_{1} = \lambda_{\delta} \hat{\boldsymbol{R}}_{1,\delta} + (1 - \lambda_{\delta}) \hat{\boldsymbol{R}}_1,
+$$ {#eq-novelist-cor}
 
-$\hat{R}_{1, \delta}$ is a thresholded correlation matrix, in which thresholding is applied only to each off-diagonal element.
-This approach will shrink the sample correlation matrix $\hat{R}_1$ towards its thresholded version.
-There are various choices for the thresholding function, in this work, we use the soft-thresholding operator, defined as:
+In this setting, $\hat{\boldsymbol{R}}_{1,\delta}$ is the thresholded correlation matrix, where each element is regularised by:
 
 $$
-\hat{r}_{1,ij}^\delta = \operatorname{sign}(\hat{r}_{1,ij}) \, (|\hat{r}_{1,ij}| - \delta)_+.
-$$
+\hat{r}_{1,ij}^\delta = \text{sign}(\hat{r}_{1,ij}) \, 
+\text{max}(|\hat{r}_{1,ij}| - \delta, \; 0),
+$$ {#eq-soft-thr}
 
-After calculated the NOVELIST correlation matrix, we can re-obtain the covariance matrix $\hat{W}^{shr}_{1, thr} = \hat{D}_1^{1/2} \hat{R}^{shr}_{1, thr} \; \hat{D}_1^{1/2}$, where $\hat{D}_1 = diag(\hat{W}_1)$ is the diagonal matrix and the elements are given by the variances of the 1-step-ahead base forecast errors.
+where $\delta \in [0,1]$ is the threshold parameter.
+For a given threshold $\delta$, @Huang2019-ua derived an analytical expression for the optimal shrinkage intensity parameter $\lambda(\delta)$ using Ledoit-Wolf's lemma [@Ledoit2003-qv], following similar logic to @Schafer2005-yw.
+It can be computed as:
 
-For a given threshold $\delta$, the optimal shrinkage intensity parameter $\lambda(\delta)$ can be estimated as:
+<!-- - \textcolor{red}{@Huang2019-ua mentioned Ledoit-Wolf's lemma} -->
 
@@
 \hat{\lambda}(\delta) = \frac{
-  \sum_{i \neq j} \widehat{Var}(\hat{r}_{1,ij}) \; I(|\hat{r}_{1,ij}| \leq \delta)
+  \sum_{i \neq j} \widehat{Var}(\hat{r}_{1,ij}) \; \boldsymbol{1}(|\hat{r}_{1,ij}| \leq \delta)
 } {
   \sum_{i \neq j} (\hat{r}_{1,ij} - \hat{r}_{1,ij}^\delta)^2
-}
+},
+$$ {#eq-lambda-thr}
+
+where $\boldsymbol{1}(.)$ is the indicator function.
+
+On the other hand, the optimal threshold $\delta^*$ does not have a closed-form solution, and is typically obtained by executing a rolling-window cross-validation procedure.
+The idea is to find the threshold $\hat{\delta^*}$, with the corresponding $\hat{\lambda}^*$ and $\hat{\boldsymbol{R}}^{N}_{1}(\hat{\delta^*}, \hat{\lambda}^*)$, that minimises the average out-of-sample reconciled forecast errors.
+The formal algorithm is given in the @sec-novelist_cv Appendix.
+Although it is not required to fit forecasting models multiple times, the cross-validation procedure is still computationally expensive as it computes the NOVELIST estimator and perform reconciliation for each threshold value.
+
+Note that when $\delta \in \bigl[ \text{max}_{i \neq j}|\hat{r}_{1,ij}|, \; 1 \bigr]$, the NOVELIST estimator collapses to the shrinkage estimator, and when $\delta = 0$, it becomes the sample covariance matrix.
+
+## POET
+
+The POET (Principal Orthogonal complEment Thresholding) estimator, proposed by @Fan2013-jz, is another "sparse" + "non-sparse" covariance estimator.
+It takes the latent factors directly into its construction, and is appealing when there are common drivers in the time series within the hierarchy, as we saw in the Australian tourism example.
+
+The POET method starts by decomposing the correlation matrix $\hat{\boldsymbol{R}}_1$ into a prominent principle components part (low-rank) and a orthogonal complement part $\hat{\boldsymbol{R}}_{1, K}$ (the correlation matrix after removing the first $K$ principal components).
+Then it applies thresholding to $\hat{\boldsymbol{R}}_{1, K}$.
+The POET estimator is given by:
+
+$$
+\hat{\boldsymbol{R}}^{K}_{1} = \sum_{k = 1}^K \hat{\gamma}_k \hat{\boldsymbol{\xi}}_k \hat{\boldsymbol{\xi}}_k' + T(\hat{\boldsymbol{R}}_{1, K})
 $$
 
-The formula derived using Ledoit-Wolf's lemma (Ledoit and Wolf, 2003) by Huang & Fryzlewicz. For the threshold parameter $\delta$, we use a cross-validation to select the optimal value
+where $\hat{\gamma}_k$ and $\hat{\boldsymbol{\xi}}_k$ are the $k$th largest eigenvalue and the corresponding eigenvector of the sample covariance matrix, respectively, and $T(.)$ is the thresholding function, which can be either soft-thresholding, hard-thresholding, or others.
 
--   ALGORITHM WALKTHROUGH
+## PC-adjusted NOVELIST
 
-**Principal-Component Adjustment**
+This approach is best of both worlds, leveraging the strengths of both NOVELIST and POET.
+The PC-adjusted (Principal-Component-adjusted) NOVELIST overcomes the shortcomings of the current shrinkage estimator, taking prominent PCs into account while also offers extra flexibility.
+The idea is to apply the NOVELIST estimator to the orthogonal complement part $\hat{\boldsymbol{R}}_{1, K}$, and then add the principal components part back.
+The PC-adjusted NOVELIST estimator is formulated as:
 
-When a factor structure is present, the procedure is:
+$$
+\hat{\boldsymbol{R}}^{N, K}_{1} = \sum_{k = 1}^K \hat{\gamma}_k \hat{\boldsymbol{\xi}}_k \hat{\boldsymbol{\xi}}_k' + \hat{\boldsymbol{R}}^{N}_{1, K}
+\; ,
+$$
 
-<!-- 1. *Standardize Forecast Errors:* -->
+where $\hat{\boldsymbol{R}}^{N}_{1, K}$ is the NOVELIST estimator applied to the orthogonal complement part $\hat{\boldsymbol{R}}_{1, K}$.
+Similar to the NOVELIST estimator, $\hat{\boldsymbol{R}}^{N, K}_{1}$ is not guaranteed to be positive definite.
 
-<!-- $$ -->
-
-<!-- \hat{z}_{t|t-1} = D_1^{-1/2} \hat{e}_{t|t-1}. -->
-
-<!-- $$ -->
-
-<!-- 2. *Construct the Data Matrix:* -->
-
-<!-- $$ -->
-
-<!-- \hat{Z} = [\hat{z}_{1|0}, \hat{z}_{2|1}, \dots, \hat{z}_{T|T-1}] -->
-
-<!-- $$ -->
-
-<!-- 3. *Extract Principal Components:* -->
-
-<!-- Compute the top K eigenvectors to form: -->
-
-<!-- $$ -->
-
-<!-- \hat{B} = T^{-1} \hat{Z} \hat{V}_K, -->
-
-<!-- $$ -->
-
-<!-- where $\hat{V}_K$ contains the K largest eigenvectors. -->
-
-<!-- 4. *Residual Component:* -->
-
-<!-- $$ -->
-
-<!-- \hat{Z}_{\text{rem}} = \hat{Z} - \hat{B}\hat{V}_K^\top. -->
-
-<!-- $$ -->
-
-<!-- Then compute the residual correlation: -->
-
-<!-- $$ -->
-
-<!-- \tilde{R}_1 = T^{-1\prime} (\hat{Z} - \hat{B}\hat{V}_K^\top)(\hat{Z} - \hat{B}\hat{V}_K^\top)^\top. -->
-
-<!-- $$ -->
-
-<!-- 5. *Apply NOVELIST to Residuals:* -->
-
-<!-- $$ -->
-
-<!-- \hat{R}_{K,\text{shr}}^{1,\text{thresh}} = \lambda \tilde{R}_\delta^1 + (1 - \lambda) \tilde{R}_1. -->
-
-<!-- $$ -->
-
-<!-- 6. *Reconstruct Full Estimator:* -->
-
-<!-- $$ -->
-
-<!-- \hat{R}_{1,K} = \hat{B}\hat{B}^\top + \hat{R}_{K,\text{shr}}^{1,\text{thresh}}. -->
-
-<!-- $$ -->
-
-<!-- **Shrinkage Intensity \(\lambda\):**   -->
-
-<!-- Estimated as -->
-
-<!-- $$ -->
-
-<!-- \hat{\lambda} = \max\Big(0, \min\Big(1, \frac{\sum_{i \neq j} \operatorname{Var}(\hat{r}_{1,ij})}{\sum_{i \neq j} \hat{r}_{1,ij}^2}\Big)\Big). -->
-
-<!-- $$ -->
-
-<!-- **Threshold \(\delta\):**   -->
-
-<!-- Selected via grid search or cross-validation. -->
-
-## Alternative Covariance Estimators
-
-Brief overview of other high-dimensional estimators used in the literature for benchmarking.
-
-Others:
-
--   Principal Orthogonal complEment Thresholding (POET) by Fan et al. (2013)
-
-    A Low-rank + Sparse method.\
-    It decompose the covariance matrix into a prominent principle components part (Low-rank) and a orthogonal complement part $R_K$. Then apply thresholding to $R_K$.\
-    This is similar to the NOVELIST estimator with PC-adjusted, but the difference is that POET apply thresholding to $R_K$, not a NOVELIST function.
+Methods to ensure positive definiteness of the NOVELIST estimator (and its PC-adjusted variant) will be explored and studied in the project.
+@Huang2019-ua proposed to diagonalise the NOVELIST estimator and replace any eigenvalues that fall under a certain small positive threshold by the value of that threshold.
+Alternatively, we can implement the algorithm of @Higham2002-te that computes the nearest positive definite matrix to a given matrix.
 

### 2025-07-28 literature/paper_draft.qmd
trying to cross ref algo pseudo

@@
     number-sections: true
+# crossref:
+#   custom:
+#     - kind: float
+#       key: alg
+#       reference-prefix: Algorithm
+#       caption-prefix: Algorithm
+#       # for PDF output only, tell LaTeX to use the `algorithm` env:
+#       latex-env: algorithm
 execute:
@@
 
-# Notation {-}
+<!-- # Notation {-} -->
 
--   Scalar $y_t$
+<!-- -   Scalar $y_t$ -->
 
--   Vector $\boldsymbol{y}_t$
+<!-- -   Vector $\boldsymbol{y}_t$ -->
 
--   Matrix $\boldsymbol{S}$
+<!-- -   Matrix $\boldsymbol{S}$ -->
 
--   Covariance matrix of in-sample 1-step-ahead base forecast errors $\hat{\boldsymbol{W}}_1$
+<!-- -   Covariance matrix of in-sample 1-step-ahead base forecast errors $\hat{\boldsymbol{W}}_1$ -->
 
--   Its shrinkage estimator with diagonal target $\hat{\boldsymbol{W}}^{shr}_{1, D}$
+<!-- -   Its shrinkage estimator with diagonal target $\hat{\boldsymbol{W}}^{shr}_{1, D}$ -->
 
--   Its NOVELIST estimator $\hat{\boldsymbol{W}}^{shr}_{1, thr}$
+<!-- -   Its NOVELIST estimator $\hat{\boldsymbol{W}}^{shr}_{1, thr}$ -->
+
+```{r library}
+library(ggplot2)
+library(patchwork)
+library(ReconCov)
+library(dplyr)
+
+setwd("D:/Github/Recon_Honours_Thesis")
+```
 
@@
 ```{r load results}
-#| include: false
-library(ggplot2)
-library(patchwork)
-library(ReconCov)
-library(dplyr)
-
-setwd("D:/Github/Recon_Honours_Thesis")
-
 sim1_1 <- readRDS("./literature/lib/S4-2-1_T100_M10000_par.rds")
@@
 
+```{r}
+path <- "./literature/lib/hstep_CV_Novelist/"
+
+```
+
 ## Exploring the sparsity of the DGP covariance matrix
@@
 
+\pagebreak
+
+# Appendix {#sec-appendix}
+
+## Algorithm: NOVELIST cross-validation for optimal threshold $\delta^*$ {#sec-novelist_cv}
+
+```{=latex}
+\begin{algorithm}
+\caption{Cross-validation procedure}
+\begin{algorithmic}[1]
+
+\State \textbf{Input:} Observations and fitted values $\boldsymbol{y}_t, \hat{\boldsymbol{y}}_t \in \mathbb{R}^n$ for $t = 1,\dots,T$, set of threshold candidates $\Delta$, window size $v$.
+
+\State $\hat{\boldsymbol{e}}_t = \boldsymbol{y}_t - \hat{\boldsymbol{y}}_t$ for $t = 1,\dots,T$
+
+\For{$i = v:T-1$}
+    \State $j = i - v +1$
+    \State $\hat{\boldsymbol{W}}_j = \frac{1}{v} \sum_{t=j}^{i} \hat{\boldsymbol{e}}_{t} \hat{\boldsymbol{e}}_{t}'$
+    \State $\hat{\boldsymbol{D}}_j = \text{diag}(\hat{\boldsymbol{W}}_j)$
+    \State $\hat{\boldsymbol{R}}_j = \hat{\boldsymbol{D}}_j^{-1/2} \hat{\boldsymbol{W}}_j \hat{\boldsymbol{D}}_j^{-1/2}$
+    \For{$\delta \in \Delta$}
+        \State Compute thresholded correlation $\hat{\boldsymbol{R}}_{j,\delta}$ using Equation 5
+        \State Compute $\hat{\lambda}_{j,\delta}$ using Equation 6
+        \State Compute $\hat{\boldsymbol{R}}^{N}_{j,\delta}$ using Equation 4
+        \State $\hat{\boldsymbol{W}}^{N}_{j,\delta} = \hat{\boldsymbol{D}}_j^{1/2} \hat{\boldsymbol{R}}^{N}_{j,\delta} \hat{\boldsymbol{D}}_j^{1/2}$
+        \State $\boldsymbol{G} = (\boldsymbol{S}' \hat{\boldsymbol{W}}^{N^{-1}}_{j,\delta} \boldsymbol{S})^{-1} \boldsymbol{S}' \hat{\boldsymbol{W}}^{N^{-1}}_{j,\delta}$
+        \State Reconciled forecasts $\tilde{\boldsymbol{y}}_{i+1 | \delta} = \boldsymbol{S} \boldsymbol{G} \hat{\boldsymbol{y}}_{i+1}$
+        \State $\tilde{\boldsymbol{e}}_{i+1 | \delta} = \boldsymbol{y}_{i+1} - \tilde{\boldsymbol{y}}_{i+1 | \delta}$
+    \EndFor
+\EndFor
+
+\State $\text{MSE}_{\delta} = \frac{1}{T-v} \sum_{i=v}^{T-1} (\tilde{\boldsymbol{e}}_{i+1 | \delta})^2$ for each $\delta \in \Delta$
+\State $\hat{\delta}^* = \arg\min_{\delta \in \Delta} \text{MSE}_{\delta}$
+\State Compute $\hat{\lambda}^*$ on all training data using $\hat{\delta}^*$
+\State Compute $\hat{\boldsymbol{R}}_1^*$ using $\hat{\delta}^*$ and $\hat{\lambda}^*$ on all training data, using Equation 3
+
+\State \textbf{Output:} Estimate of optimal $\hat{\delta}^*$
+
+\end{algorithmic}
+\end{algorithm}
+```
+
+\pagebreak
+
 # References

### 2025-07-28 literature/paper_draft.qmd
outline for tourism section

@@
 
-# Empirical Analysis
+# Forecasting Australian Domestics Tourism
 
--   Comparative analysis of forecast performance using different covariance estimators on real-life dataset.
+- Description
+- Hierarchy Description
+- Models and Cross-validation 
+- Structural change after 2016 (slight upward trend)
+- Results over entire data set and excluding 2016 onwards
+- Next steps:
+  - Incorporate tourism setting into current DGP
+  - Explore how to tackle structural change in MinT
 

### 2025-07-28 literature/paper_draft.qmd
descipritn of data and hierarchy for tourism analys

@@
 
+Domestic tourism flows in any country naturally form a hierarchical structure.
+They are typically aggregated from the bottom level of individual regions to higher levels including zones, states, and the entire country.
+Our empirical analysis draws on the Australian domestic overnight tourism, a country with 7 states and territories, which are further divided into 27 zones and 77 regions.
+The data set also consider tourism flows by purposes of travel -- holiday, business, visiting friends and relatives, and other reason -- which can highlight the patterns of traveling and spending behaviours across these groups.
+Grouping the Australian domestic tourism by geographical divisions and purposes of travel forms a grouped structure comprising of 560 individual series in total (both aggregated and disaggregated).
+
+We measure the tourism demand using “visitor nights,” the total number of nights spent by Australians away from home.
+The data is collected via the National Visitor Survey, managed by Tourism Research Australia, using computer assisted telephone interviews from nearly 120,000 Australian residents aged 15 years and over [@tourismau-jk].
+The data are monthly time series spanning from January 1998 to December 2019, resulting in 264 observations per series, producing a canonical "$n \ll p$" setting for which high-dimensional covariance estimators were designed.
+This is particularly well-suited to evaluate forecast reconciliation methods, because the extreme dimensionality over sample size mirrors many comptemporary business problems, for instance, Starbucks drink sales.
+
+<!-- This dataset is particularly well‑suited to evaluating forecast‑reconciliation methods for three reasons.
+First, tourism demand is economically important and notoriously volatile; the heterogeneous regional and purpose‑specific patterns create a realistic stress‑test for reconciliation algorithms.
+Second, the clear structural shift towards stronger growth after 2016—visible in most aggregates—necessitates techniques capable of handling possible breaks or time‑varying error structures, making the exercise practically relevant.
+Third, the extreme dimensionality relative to sample size mirrors many contemporary business problems (e.g., SKU‑level retail sales), allowing insights to generalise beyond tourism.
+Using this panel therefore provides both a compelling policy context and a stringent statistical laboratory for comparing alternative reconciliation strategies, especially those that rely on regularised covariance estimation or exploit cross‑sectional information to stabilise forecasts when historical data are scant.
+-->
+
 # Discussion and Conclusion

### 2025-07-28 literature/paper_draft.qmd
expand on tourism section, included modelling and cross validation

@@
 
+<!-- Table to illustrate the hierarchy -->
+
 We measure the tourism demand using “visitor nights,” the total number of nights spent by Australians away from home.
 The data is collected via the National Visitor Survey, managed by Tourism Research Australia, using computer assisted telephone interviews from nearly 120,000 Australian residents aged 15 years and over [@tourismau-jk].
+
 The data are monthly time series spanning from January 1998 to December 2019, resulting in 264 observations per series, producing a canonical "$n \ll p$" setting for which high-dimensional covariance estimators were designed.
-This is particularly well-suited to evaluate forecast reconciliation methods, because the extreme dimensionality over sample size mirrors many comptemporary business problems, for instance, Starbucks drink sales.
+This is particularly well-suited to evaluate forecast reconciliation methods, because the extreme dimensionality over sample size mirrors many contemporary business problems, for instance, Starbucks drink sales.
+Tourism demand is also economically important and notoriously volatile; the heterogeneous regional and purpose‑specific patterns create a realistic stress‑test for reconciliation algorithms.
+Using this panel therefore provides both a compelling policy context and a stringent statistical laboratory for comparing between reconciliation methods, especially those that rely on regularised covariance estimation or exploit cross‑sectional information to stabilise forecasts when historical data are scant.
 
@@
 -->
+
+We evaluate the performance of reconcilation methods using a rolling window cross validation process.
+We set the training window to be 96 observations, which serves as a training set to obtain the best-fitted ARIMA models for each of 560 series by minimising AICc, in which we use the automatic algorithm from Fabletools R package [@fabletools-fa].
+The 1- to 12-step-ahead base forecasts are then generated by these ARIMA models.
+These base forecasts are combined or reconciled using alternative strategies to obtain reconciled forecasts.
+We repeat the procedure by rolling the training window by one observation at a time, until December 2018.
+This results in a total of 156 windows, with corresponding 156 sets of 1- to 12-step-ahead forecasts for each of 560 series.
+
+Regarding reconciliation strategies, we consider OLS, mint_shr (MinT with Shrinkage estimator), and mint_n (MinT with NOVELIST estimator).
+Note that the number of series is larger than the number of observations (560 compared to 96), hence the sample covariance matrix is not positive definite and will not be used.
 

### 2025-07-29 literature/paper_draft.qmd
Update paper_draft.qmd

@@
 library(dplyr)
+library(tidyr)
+library(fable)
+library(fabletools)
+library(feasts)
+library(tsibble)
 
-setwd("D:/Github/Recon_Honours_Thesis")
+path <- "D:/Github/Recon_Honours_Thesis"
+knitr::opts_knit$set(root.dir = path)
+# setwd(path)
 ```
@@
 ```{r load results}
-sim1_1 <- readRDS("./literature/lib/S4-2-1_T100_M10000_par.rds")
-sim1_2 <- readRDS("./literature/lib/S4-2-1_T300_M10000.rds")
+
+# sim1_1 <- readRDS("./literature/lib/S4-2-1_T100_M10000_par.rds")
+# sim1_2 <- readRDS("./literature/lib/S4-2-1_T300_M10000.rds")
 
@@
 
-sim3_1 <- readRDS("./literature/lib/S100-2-1_T50_M200_run1.rds")
-sim3_2 <- readRDS("./literature/lib/S100-2-1_T100_M200_run1.rds")
-
-MSE1_1 <- transform_sim_MSE(sim1_1$MSE, F)
-MSE1_2 <- transform_sim_MSE(sim1_2$MSE, F)
+# sim3_1 <- readRDS("./literature/lib/S100-2-1_T50_M200_run1.rds")
+# sim3_2 <- readRDS("./literature/lib/S100-2-1_T100_M200_run1.rds")
+# 
+# MSE1_1 <- transform_sim_MSE(sim1_1$MSE, F)
+# MSE1_2 <- transform_sim_MSE(sim1_2$MSE, F)
 
@@
 
-MSE3_1 <- transform_sim_MSE(sim3_1$MSE, F)
-MSE3_2 <- transform_sim_MSE(sim3_2$MSE, F)
+# MSE3_1 <- transform_sim_MSE(sim3_1$MSE, F)
+# MSE3_2 <- transform_sim_MSE(sim3_2$MSE, F)
 ```

### 2025-07-29 literature/paper_draft.qmd
talk about structural change

@@
 
+```{r}
+visnights_full <- readRDS("literature/lib/visnights_full.rds")
+```
+
+```{r}
+#| label: fig-STL-tourism
+#| fig-cap: "Monthly domestic overnight tourism nights in Australia by state, with STL decomposition to extract trend component"
+#| fig.width: 8
+#| fig.height: 6
+#| out.width: "100%"
+#| fig.align: "center"
+visnights_full %>% 
+  filter(
+    is_aggregated(Purpose) & is_aggregated(Zone) & is_aggregated(Region),
+    !is_aggregated(State)
+  ) %>% 
+  model(
+    STL(Nights)
+  ) %>% 
+  components() %>% 
+  pivot_longer(
+    cols = c(Nights, trend, season_year, remainder, season_adjust),
+    names_to = "component",
+    values_to = "value"
+  ) %>%
+  filter(
+    component == c("Nights", "trend"),
+    year(Month) >= 2000
+  ) %>%
+  as_tibble() %>%
+  mutate(component = recode(component,                     # give them nicer names
+                            Nights = "Observed Nights",
+                            trend  = "STL Decomposed Trend")) %>% 
+  ggplot(aes(Month, value, color = as.character(State))) +
+    geom_line() +
+    facet_wrap(~ component, ncol=1, scales = "free_y") +
+    labs(
+      x = "Month",
+      y = "Visitor Nights (in thousands)",
+      color = "State"
+    ) +
+    theme_bw()
+```
+
+Additionally, we notice a structural change in the tourism data after 2016, which is characterised by a slight upward trend in demand, as illustrated in @fig-STL-tourism.
+This trend only happened in New South Wales, Victoria, Queensland, and Western Australia.
+Our empirical results show that the base forecasts produced by the ARIMA models are not able to capture this trend, leading to even poorer performance of reconciliation methods.
+To account for this, we will also evaluate the performance of reconciliation methods on the data set excluding the period after 2016, which is from January 1998 to December 2016 (the last training window ends on December 2015).
+
 # Discussion and Conclusion

### 2025-07-29 literature/paper_draft.qmd
Update paper_draft.qmd

@@
 library(tsibble)
+library(lubridate)
 
@@
 
-# sim1_1 <- readRDS("./literature/lib/S4-2-1_T100_M10000_par.rds")
-# sim1_2 <- readRDS("./literature/lib/S4-2-1_T300_M10000.rds")
+sim1_1 <- readRDS("./literature/lib/S4-2-1_T100_M10000_par.rds")
+sim1_2 <- readRDS("./literature/lib/S4-2-1_T300_M10000.rds")
 
@@
 
-# sim3_1 <- readRDS("./literature/lib/S100-2-1_T50_M200_run1.rds")
-# sim3_2 <- readRDS("./literature/lib/S100-2-1_T100_M200_run1.rds")
-# 
-# MSE1_1 <- transform_sim_MSE(sim1_1$MSE, F)
-# MSE1_2 <- transform_sim_MSE(sim1_2$MSE, F)
+sim3_1 <- readRDS("./literature/lib/S100-2-1_T50_M200_run1.rds")
+sim3_2 <- readRDS("./literature/lib/S100-2-1_T100_M200_run1.rds")
+
+MSE1_1 <- transform_sim_MSE(sim1_1$MSE, F)
+MSE1_2 <- transform_sim_MSE(sim1_2$MSE, F)
 
@@
 
-# MSE3_1 <- transform_sim_MSE(sim3_1$MSE, F)
-# MSE3_2 <- transform_sim_MSE(sim3_2$MSE, F)
+MSE3_1 <- transform_sim_MSE(sim3_1$MSE, F)
+MSE3_2 <- transform_sim_MSE(sim3_2$MSE, F)
 ```

### 2025-07-29 literature/paper_draft.qmd
expand on trend after 2016

@@
 Additionally, we notice a structural change in the tourism data after 2016, which is characterised by a slight upward trend in demand, as illustrated in @fig-STL-tourism.
-This trend only happened in New South Wales, Victoria, Queensland, and Western Australia.
-Our empirical results show that the base forecasts produced by the ARIMA models are not able to capture this trend, leading to even poorer performance of reconciliation methods.
+The trend component is extracted using STL decomposition [@??] on the monthly domestic overnight tourism nights in Australia by state.
+This trend only happened in New South Wales, Victoria, Queensland, and Western Australia, which is likely due to the increasing popularity of tourism in these places, as well as the impact of various factors such as economic growth, population growth, and changes in consumer preferences.
+Our empirical analysis show that the base forecasts produced by the ARIMA models are not able to capture this trend, leading to even poorer performance of reconciliation methods.
 To account for this, we will also evaluate the performance of reconciliation methods on the data set excluding the period after 2016, which is from January 1998 to December 2016 (the last training window ends on December 2015).

### 2025-07-29 literature/paper_draft.qmd
Code for ploting MSE by each level for tourism results

@@
 
+```{r}
+library(stringr)
+
+desired_order <- c(
+  "Australia", "Australia_by_Purpose",
+  "State",     "State_by_Purpose",
+  "Zone",      "Zone_by_Purpose",
+  "Region",    "Region_by_Purpose"
+)
+```
+
+```{r}
+#| label: fig-tourism-results
+#| fig-cap: "Relative improvement of the MSE of reconciled forecasts over the base forecasts for the Australian domestic tourism flows"
+#| fig.width: 8
+#| fig.height: 10
+#| out.width: "100%"
+
+tourism_MSE <- readRDS("./literature/lib/tourism_MSE.rds")
+
+# % improv plot by level
+tourism_MSE |>
+  group_by(series, h) |>
+  mutate(
+    base_MSE = MSE[.model == "base"]
+  ) |>
+  ungroup() |>
+  group_by(.model, level, h) |>
+  summarise(
+    MSE = mean(MSE),
+    base_MSE = mean(base_MSE),
+    pct_change = (MSE - base_MSE) / base_MSE * 100
+  ) |>
+  filter(.model != "mint_sample") %>%
+  # Reorder the levels according to the desired order
+  mutate(
+    level = factor(
+      level,
+      levels = desired_order,
+      labels = str_replace_all(desired_order, "_", " ")  # replace "_" with " "
+    )
+  ) %>% 
+  ggplot(aes(x = h, y = pct_change, color = .model)) +
+    geom_line() +
+    labs(x = "Horizon", y = "% improvements") +
+    facet_wrap(~ level, ncol=2, scales = "free_y") +
+    theme_bw() +
+    scale_x_continuous(limits = c(1, 12), expand = c(0, 0)) +
+    theme(legend.position = "bottom", legend.title = element_blank())
+```
+
+```{r}
+#| label: fig-tourism-results-ex1026
+#| fig-cap: "Relative improvement of the MSE of reconciled forecasts over the base forecasts for the Australian domestic tourism flows, excluding the period after 2016"
+#| fig.width: 8
+#| fig.height: 10
+#| out.width: "100%"
+
+tourism_MSE_ex2016 <- readRDS("literature/lib/tourism_MSE_ex2016.rds")
+
+# % improv plot by level
+tourism_MSE_ex2016 |>
+  group_by(series, h) |>
+  mutate(
+    base_MSE = MSE[.model == "base"]
+  ) |>
+  ungroup() |>
+  group_by(.model, level, h) |>
+  summarise(
+    MSE = mean(MSE),
+    base_MSE = mean(base_MSE),
+    pct_change = (MSE - base_MSE) / base_MSE * 100
+  ) |>
+  filter(.model != "mint_sample") %>%
+  # Reorder the levels according to the desired order
+  mutate(
+    level = factor(
+      level,
+      levels = desired_order,
+      labels = str_replace_all(desired_order, "_", " ")  # replace "_" with " "
+    )
+  ) %>%
+  ggplot(aes(x = h, y = pct_change, color = .model)) +
+    geom_line() +
+    labs(x = "Horizon", y = "% improvements") +
+    facet_wrap(~ level, ncol=2, scales = "free_y") +
+    theme_bw() +
+    scale_x_continuous(limits = c(1, 12), expand = c(0, 0)) +
+    theme(legend.position = "bottom", legend.title = element_blank())
+```
+
 # Discussion and Conclusion

### 2025-07-29 literature/paper_draft.qmd
Update paper_draft.qmd

@@
 
+@Wickramasuriya2019-xq also argued that modelling spatial autocorrelations would be challenging as in the case of a large collection of time series.
+Reconciliation approaches have the advantage to implicitly model this spatial autocorreltion structure, especially when the MinT method is used.
+
 We evaluate the performance of reconcilation methods using a rolling window cross validation process.
@@
 Regarding reconciliation strategies, we consider OLS, mint_shr (MinT with Shrinkage estimator), and mint_n (MinT with NOVELIST estimator).
-Note that the number of series is larger than the number of observations (560 compared to 96), hence the sample covariance matrix is not positive definite and will not be used.
+Note that the number of series is larger than the number of observations (560 compared to 96), hence the sample covariance matrix is not positive definite and will not be considered
 

### 2025-07-30 literature/paper_draft.qmd
write on results empirical findings

@@
 
-Regarding reconciliation strategies, we consider OLS, mint_shr (MinT with Shrinkage estimator), and mint_n (MinT with NOVELIST estimator).
-Note that the number of series is larger than the number of observations (560 compared to 96), hence the sample covariance matrix is not positive definite and will not be considered
-
-```{r}
-visnights_full <- readRDS("literature/lib/visnights_full.rds")
-```
-
-```{r}
-#| label: fig-STL-tourism
-#| fig-cap: "Monthly domestic overnight tourism nights in Australia by state, with STL decomposition to extract trend component"
-#| fig.width: 8
-#| fig.height: 6
-#| out.width: "100%"
-#| fig.align: "center"
-visnights_full %>% 
-  filter(
-    is_aggregated(Purpose) & is_aggregated(Zone) & is_aggregated(Region),
-    !is_aggregated(State)
-  ) %>% 
-  model(
-    STL(Nights)
-  ) %>% 
-  components() %>% 
-  pivot_longer(
-    cols = c(Nights, trend, season_year, remainder, season_adjust),
-    names_to = "component",
-    values_to = "value"
-  ) %>%
-  filter(
-    component == c("Nights", "trend"),
-    year(Month) >= 2000
-  ) %>%
-  as_tibble() %>%
-  mutate(component = recode(component,                     # give them nicer names
-                            Nights = "Observed Nights",
-                            trend  = "STL Decomposed Trend")) %>% 
-  ggplot(aes(Month, value, color = as.character(State))) +
-    geom_line() +
-    facet_wrap(~ component, ncol=1, scales = "free_y") +
-    labs(
-      x = "Month",
-      y = "Visitor Nights (in thousands)",
-      color = "State"
-    ) +
-    theme_bw()
-```
-
-Additionally, we notice a structural change in the tourism data after 2016, which is characterised by a slight upward trend in demand, as illustrated in @fig-STL-tourism.
-The trend component is extracted using STL decomposition [@??] on the monthly domestic overnight tourism nights in Australia by state.
-This trend only happened in New South Wales, Victoria, Queensland, and Western Australia, which is likely due to the increasing popularity of tourism in these places, as well as the impact of various factors such as economic growth, population growth, and changes in consumer preferences.
-Our empirical analysis show that the base forecasts produced by the ARIMA models are not able to capture this trend, leading to even poorer performance of reconciliation methods.
-To account for this, we will also evaluate the performance of reconciliation methods on the data set excluding the period after 2016, which is from January 1998 to December 2016 (the last training window ends on December 2015).
-
 ```{r}
@@
 
-```{r}
+```{r fig tourism results}
 #| label: fig-tourism-results
@@
 
+```{r fig STL tourism}
+#| label: fig-STL-tourism
+#| fig-cap: "Monthly domestic overnight tourism nights in Australia by state, with STL decomposition to extract trend component"
+#| fig.width: 8
+#| fig.height: 6
+#| out.width: "100%"
+#| fig.align: "center"
+
+visnights_full <- readRDS("literature/lib/visnights_full.rds")
+
+visnights_full %>% 
+  filter(
+    is_aggregated(Purpose) & is_aggregated(Zone) & is_aggregated(Region),
+    !is_aggregated(State)
+  ) %>% 
+  model(
+    STL(Nights)
+  ) %>% 
+  components() %>% 
+  pivot_longer(
+    cols = c(Nights, trend, season_year, remainder, season_adjust),
+    names_to = "component",
+    values_to = "value"
+  ) %>%
+  filter(
+    component == c("Nights", "trend"),
+    year(Month) >= 2000
+  ) %>%
+  as_tibble() %>%
+  mutate(component = recode(component,                     # give them nicer names
+                            Nights = "Observed Nights",
+                            trend  = "STL Decomposed Trend")) %>% 
+  ggplot(aes(Month, value, color = as.character(State))) +
+    geom_line() +
+    facet_wrap(~ component, ncol=1, scales = "free_y") +
+    labs(
+      x = "Month",
+      y = "Visitor Nights (in thousands)",
+      color = "State"
+    ) +
+    theme_bw()
+```
+
+Regarding reconciliation strategies, we consider OLS, mint_shr (MinT with Shrinkage estimator), and mint_n (MinT with NOVELIST estimator).
+Note that the number of series is larger than the number of observations (560 compared to 96), hence the sample covariance matrix is not positive definite and will not be considered.
+
+The results are presented in @fig-tourism-results, which shows the relative improvement of the MSE of reconciled forecasts over the base forecasts for the Australian domestic tourism flows.
+The results are grouped by levels of aggregation.
+We can clearly see that MinT with either Shrinkage (mint_shr) or NOVELIST estimator (mint_n) improve the point forecast accuracy over the base ARIMA models for middle to lower aggregation levels.
+We hardly differentiate the point accuracy performance between mint_shr and mint_n in these level, except at the most disaggregated, where mint_shr has slightly larger improvement.
+
+However, surprisingly, MinT performs worse over OLS at higher aggregation level, and especially worse than the base models at country-wise.
+This contradicts with the empirical results from @Wickramasuriya2019-xq and others.
+Further inspection draws our attention to a structural change in the tourism data after 2016, which is characterised by a slight upward trend in demand, as illustrated in @fig-STL-tourism.
+The trend component is extracted using STL decomposition [@??] on the monthly domestic overnight tourism nights in Australia by state.
+This trend only happened in New South Wales, Victoria, Queensland, and Western Australia, which is likely due to the increasing popularity of tourism in these places, as well as the impact of various factors such as economic growth, population growth, and changes in consumer preferences.
+
 ```{r}
-#| label: fig-tourism-results-ex1026
+#| label: fig-tourism-results-ex2016
 #| fig-cap: "Relative improvement of the MSE of reconciled forecasts over the base forecasts for the Australian domestic tourism flows, excluding the period after 2016"
@@
 ```
+
+Our empirical analysis show that the base forecasts produced by the ARIMA models are not able to capture this trend, leading to even poorer performance of reconciliation methods.
+To account for this, we will also evaluate the performance of reconciliation methods on the data set excluding the period after the year 2016 (the last training window ends on December 2015), in which the results are shown in @fig-tourism-results-ex2016.
+
+In this case, we can clearly observe that both MinT methods return the most accurate forecasts.
+At higher aggregation levels (Australia, State, and Australia by Purposes levels), we no longer see the patterns as in @fig-tourism-results.
+In contrast, MinT performs drastically better than the base models.
+Additionally, the higher the aggregation level, the more accurate mint_n is over mint_shr.
+This is an advantage of NOVELIST estimator.
+At higher levels most series are combinations of many bottom cells, the base forecast error correlations are strong and genuinely useful.
+NOVELIST keeps the large cross-series correlations at those higher levels, while its thresholding feature minimises the effects of weak, noisy links.
+
+The Australian domestic tourism data shred lights on the potentials of NOVELIST estimator, having results that we did not produce in our previous simulation settings.
+Further ahead, we aim to incorporate the structure of the tourism flows into our data generating process settings, and further explore the cases where mint_n can take advantage of the features that mint_shr can not.
+Nevertheless, improving the forecasts of Australian tourism demand after the structural change occured is another question we should keep in mind.
 

### 2025-07-30 literature/paper_draft.qmd
cleanup and reword, refine the tourism part

@@
 execute:
-  cache: true
+  cache: false
   echo: false
@@
 
-- Description
-- Hierarchy Description
-- Models and Cross-validation 
-- Structural change after 2016 (slight upward trend)
-- Results over entire data set and excluding 2016 onwards
-- Next steps:
-  - Incorporate tourism setting into current DGP
-  - Explore how to tackle structural change in MinT
-
-Domestic tourism flows in any country naturally form a hierarchical structure.
-They are typically aggregated from the bottom level of individual regions to higher levels including zones, states, and the entire country.
-Our empirical analysis draws on the Australian domestic overnight tourism, a country with 7 states and territories, which are further divided into 27 zones and 77 regions.
-The data set also consider tourism flows by purposes of travel -- holiday, business, visiting friends and relatives, and other reason -- which can highlight the patterns of traveling and spending behaviours across these groups.
-Grouping the Australian domestic tourism by geographical divisions and purposes of travel forms a grouped structure comprising of 560 individual series in total (both aggregated and disaggregated).
+Domestic tourism flows in Australia exhibit a natural hierarchical and grouped structure, driven both by geography and by purpose of travel.
+At the top of this hierarchy lies the national total, which splits into the seven states and territories.
+Each state is further sub-divided into tourism “zones,” which in turn break down into 77 “regions.” Intersecting this geographic hierarchy is a second dimension--travel motive--partitioning tourism flows into four categories: holiday, business, visiting friends and relatives, and other.
+Altogether, this yields a grouped system of 560 series, from the most disaggregated regional-purpose cells up to the full national aggregate.
 
@@
 
-We measure the tourism demand using “visitor nights,” the total number of nights spent by Australians away from home.
+We quantify tourism demand via “visitor nights,” the total number of nights spent by Australians away from home.
 The data is collected via the National Visitor Survey, managed by Tourism Research Australia, using computer assisted telephone interviews from nearly 120,000 Australian residents aged 15 years and over [@tourismau-jk].
 
-The data are monthly time series spanning from January 1998 to December 2019, resulting in 264 observations per series, producing a canonical "$n \ll p$" setting for which high-dimensional covariance estimators were designed.
-This is particularly well-suited to evaluate forecast reconciliation methods, because the extreme dimensionality over sample size mirrors many contemporary business problems, for instance, Starbucks drink sales.
-Tourism demand is also economically important and notoriously volatile; the heterogeneous regional and purpose‑specific patterns create a realistic stress‑test for reconciliation algorithms.
-Using this panel therefore provides both a compelling policy context and a stringent statistical laboratory for comparing between reconciliation methods, especially those that rely on regularised covariance estimation or exploit cross‑sectional information to stabilise forecasts when historical data are scant.
+The data are monthly time series spanning from January 1998 to December 2019, resulting in 264 observations per series, producing a canonical "$n \ll p$" setting 
+which is ideal for evaluating reconciliation approaches that rely on high-dimensional covariance estimation.
+The extreme dimensionality over sample size mirrors many contemporary business problems, for instance, Starbucks drink sales.
+Tourism demand is also economically vital yet highly volatile, with geographical and purpose‑specific patterns create a realistic stress‑test for reconciliation algorithms.
+Consequently, this panel offers both a rich policy case study and a stringent statistical laboratory for comparing reconciliation strategies that exploit cross-series information to improve forecasts when historical data are scarce.
 
@@
 
-We evaluate the performance of reconcilation methods using a rolling window cross validation process.
-We set the training window to be 96 observations, which serves as a training set to obtain the best-fitted ARIMA models for each of 560 series by minimising AICc, in which we use the automatic algorithm from Fabletools R package [@fabletools-fa].
+To assess forecasting performance, we adopt a rolling-window cross-validation scheme.
+Beginning with the first 96 monthly observations (January 1998-December 2005) as the initial training set, we obtain the best-fitted ARIMA model for each of the 560 series via the automatic algorithm in the fabletools package, by minimising AICc [@fabletools-fa].
 The 1- to 12-step-ahead base forecasts are then generated by these ARIMA models.
-These base forecasts are combined or reconciled using alternative strategies to obtain reconciled forecasts.
-We repeat the procedure by rolling the training window by one observation at a time, until December 2018.
-This results in a total of 156 windows, with corresponding 156 sets of 1- to 12-step-ahead forecasts for each of 560 series.
+These incoherent base forecasts are reconciled via OLS, MinT with shrinkage (mint_shr), and MinT with the NOVELIST estimator (mint_n).
+We then roll the training window forward by one month and refit all models, rebuild reconciliations, and produce another batch of 1- to 12-step-ahead forecasts, repeating until December 2018.
+In total, this results in 156 out-of-sample windows and an equal number of forecast sets for each series.
 
@@
 
-Regarding reconciliation strategies, we consider OLS, mint_shr (MinT with Shrinkage estimator), and mint_n (MinT with NOVELIST estimator).
 Note that the number of series is larger than the number of observations (560 compared to 96), hence the sample covariance matrix is not positive definite and will not be considered.
@@
 The results are grouped by levels of aggregation.
-We can clearly see that MinT with either Shrinkage (mint_shr) or NOVELIST estimator (mint_n) improve the point forecast accuracy over the base ARIMA models for middle to lower aggregation levels.
-We hardly differentiate the point accuracy performance between mint_shr and mint_n in these level, except at the most disaggregated, where mint_shr has slightly larger improvement.
+MinT variants show improvement over the base ARIMA forecasts for middle to lower aggregation levels.
+We hardly differentiate the point accuracy performance between mint_shr and mint_n in these level, except at the most disaggregated, where mint_shr slightly edges out mint_n.
 
-However, surprisingly, MinT performs worse over OLS at higher aggregation level, and especially worse than the base models at country-wise.
-This contradicts with the empirical results from @Wickramasuriya2019-xq and others.
-Further inspection draws our attention to a structural change in the tourism data after 2016, which is characterised by a slight upward trend in demand, as illustrated in @fig-STL-tourism.
-The trend component is extracted using STL decomposition [@??] on the monthly domestic overnight tourism nights in Australia by state.
-This trend only happened in New South Wales, Victoria, Queensland, and Western Australia, which is likely due to the increasing popularity of tourism in these places, as well as the impact of various factors such as economic growth, population growth, and changes in consumer preferences.
+Surprisingly, however, MinT underperforms OLS at higher aggregation level and even the base forecasts at country-wise.
+This contradicts with the canonical results from @Wickramasuriya2019-xq and others, prompting further investigation.
+Applying STL decomposition to state-aggregated observations reveals a structural change around 2016: four large states (New South Wales, Victoria, Queensland, Western Australia) develop a modest upward trend in visitor nights, as shown in @fig-STL-tourism.
+The trend is likely due to the increasing popularity of tourism in these places, as well as the impact of various factors such as economic growth, population growth, and changes in consumer preferences.
 
@@
 
-Our empirical analysis show that the base forecasts produced by the ARIMA models are not able to capture this trend, leading to even poorer performance of reconciliation methods.
-To account for this, we will also evaluate the performance of reconciliation methods on the data set excluding the period after the year 2016 (the last training window ends on December 2015), in which the results are shown in @fig-tourism-results-ex2016.
+Neither ARIMA nor reconciliation strategies capture this shift, degrading accuracy at the most aggregated level.
+From the results in @fig-tourism-results-ex2016, excluding the post-2016 period (i.e., restricting the final training window to end on December 2015) restores reconciliation’s performance: both MinT methods now clearly outperform OLS and the base forecasts at all levels.
+Additionally, we notice that mint_n increasingly outperforms mint_shr at higher levels (Australia, State, and Australia by Purposes levels).
+This is an advantage of NOVELIST estimator.
+At higher levels most series are combinations of many bottom cells, their cross-series base forecast error correlations are strong and genuinely useful.
+NOVELIST keeps these strong links at those higher levels, while its thresholding feature minimises the effects of weak, noisy ones.
 
-In this case, we can clearly observe that both MinT methods return the most accurate forecasts.
-At higher aggregation levels (Australia, State, and Australia by Purposes levels), we no longer see the patterns as in @fig-tourism-results.
-In contrast, MinT performs drastically better than the base models.
-Additionally, the higher the aggregation level, the more accurate mint_n is over mint_shr.
-This is an advantage of NOVELIST estimator.
-At higher levels most series are combinations of many bottom cells, the base forecast error correlations are strong and genuinely useful.
-NOVELIST keeps the large cross-series correlations at those higher levels, while its thresholding feature minimises the effects of weak, noisy links.
+The empirical analysis on real, hierarchical Australian domestic tourism shreds lights on the strengths of the NOVELIST estimator--advantages that were not fully apparent in previous simulation studies.
+Going forward, we plan to embed this structure into our data-generating processes and to pinpoint contexts where NOVELIST’s thresholding yields gains beyond those of Shrinkage.
+Moreover, our findings underscore the need to accommodate structural changes--the post-2016 Australian domestic tourism demand shifts we encountered.
 
-The Australian domestic tourism data shred lights on the potentials of NOVELIST estimator, having results that we did not produce in our previous simulation settings.
-Further ahead, we aim to incorporate the structure of the tourism flows into our data generating process settings, and further explore the cases where mint_n can take advantage of the features that mint_shr can not.
-Nevertheless, improving the forecasts of Australian tourism demand after the structural change occured is another question we should keep in mind.
+<!-- perhaps via robust trend adjustment or state-space extensions that jointly model coherence and temporal dynamics. -->
 

### 2025-07-30 literature/paper_draft.qmd
cleanup

@@
 
+<!--
 -   **Problem Statement:**
@@
     -   Conclusions and suggestions for future work.
+-->
 
+<!--
 # Literature Review
@@
 
-<!-- Wickramasuriya et al.
+Wickramasuriya et al.
 (2019) adopt the diagonal-target Ledoit–Wolf approach and show analytically that any scalar multiple $k_h$ from $W_h = k_hW_1$ suffices for point-forecast reconciliation.
@@
 Their subsequent work imposes non-negativity constraints via quadratic programming citeturn0file2 and studies bias-robust variants (MinT-U).
--->
 
-<!-- Empirical findings to date   -->
+Empirical findings to date
 
-<!-- Athanasopoulos et al.
+Athanasopoulos et al.
 (2024) synthesise the accumulated evidence: MinT dominates single-level baselines in most studies; WLS or diagonal-shrinkage MinT is the current industry default; but gains plateau when off-diagonals are weak citeturn0file9.
 Only a handful of papers explore richer covariance structures (e.g., block or factor shrinkage), and none benchmark them systematically against NOVELIST-type estimators.
--->
 
@@
 -   Identify research gaps.
+-->
 
@@
 However, the hierarchical time series data often exhibit a prominent principal components structure, which is not fully taken advantage.
-Taking an example of the Australian domestic overnight trips data set [@tourismau-jk], where the national trips are disaggregated into states and territories, and further into regions.
-We then fit ETS models to all series, using the algorithm from Fabletools R package [@fabletools-fa], and compute the one-step-ahead in-sample base forecast error covariance matrix $\hat{\boldsymbol{W}}_1$.
+Taking an example of the Australian domestic overnight trips data set [@tourism-au], where the national trips are disaggregated into states and territories, and further into regions.
+We then fit ETS models to all series, using the algorithm from Fabletools R package [@O-Hara-Wild2024-we], and compute the one-step-ahead in-sample base forecast error covariance matrix $\hat{\boldsymbol{W}}_1$.
 The twenty largest eigenvalues of the covariance matrix are plotted in @fig-eigen.
@@
 The first 50 and 300 observations are used for training, and the last 4 observations are used for testing.
-The training data is used to compute the best fitted ARIMA models by minimising the AICc criterion, in which we use the automatic algorithm from Fabletools R package [@fabletools-fa].
+The training data is used to compute the best fitted ARIMA models by minimising the AICc criterion, in which we use the automatic algorithm from Fabletools R package [@O-Hara-Wild2024-we].
 We refer to them as base models, and their base forecasts are then reconciled using the MinT with different covariance estimators.
@@
 
+<!--
 ## Exploring Effects of Hierarchy's Aggregation
@@
 - Finding: Complex aggregation MinT_n performs well
-
-## h-step-ahead forecast error metric in NOVELIST Cross-validation
-
-Instead of forecasting t+1 for each rolling window iteration, I forecast t+1 to t+h for each iteration and compute the MSE for each threshold value delta.
-
-Results barely show any difference with the original method.
-
-```{r}
-path <- "./literature/lib/hstep_CV_Novelist/"
-
-```
+-->
 
@@
 At the top of this hierarchy lies the national total, which splits into the seven states and territories.
-Each state is further sub-divided into tourism “zones,” which in turn break down into 77 “regions.” Intersecting this geographic hierarchy is a second dimension--travel motive--partitioning tourism flows into four categories: holiday, business, visiting friends and relatives, and other.
+Each state is further sub-divided into tourism zones, which in turn break down into 77 regions.
+Intersecting this geographic hierarchy is a second dimension--travel motive--partitioning tourism flows into four categories: holiday, business, visiting friends and relatives, and other.
 Altogether, this yields a grouped system of 560 series, from the most disaggregated regional-purpose cells up to the full national aggregate.
 
-<!-- Table to illustrate the hierarchy -->
+- A table/figure to illustrate the hierarchy
 
-We quantify tourism demand via “visitor nights,” the total number of nights spent by Australians away from home.
-The data is collected via the National Visitor Survey, managed by Tourism Research Australia, using computer assisted telephone interviews from nearly 120,000 Australian residents aged 15 years and over [@tourismau-jk].
-
-The data are monthly time series spanning from January 1998 to December 2019, resulting in 264 observations per series, producing a canonical "$n \ll p$" setting 
-which is ideal for evaluating reconciliation approaches that rely on high-dimensional covariance estimation.
-The extreme dimensionality over sample size mirrors many contemporary business problems, for instance, Starbucks drink sales.
-Tourism demand is also economically vital yet highly volatile, with geographical and purpose‑specific patterns create a realistic stress‑test for reconciliation algorithms.
-Consequently, this panel offers both a rich policy case study and a stringent statistical laboratory for comparing reconciliation strategies that exploit cross-series information to improve forecasts when historical data are scarce.
-
-<!-- This dataset is particularly well‑suited to evaluating forecast‑reconciliation methods for three reasons.
-First, tourism demand is economically important and notoriously volatile; the heterogeneous regional and purpose‑specific patterns create a realistic stress‑test for reconciliation algorithms.
-Second, the clear structural shift towards stronger growth after 2016—visible in most aggregates—necessitates techniques capable of handling possible breaks or time‑varying error structures, making the exercise practically relevant.
-Third, the extreme dimensionality relative to sample size mirrors many contemporary business problems (e.g., SKU‑level retail sales), allowing insights to generalise beyond tourism.
-Using this panel therefore provides both a compelling policy context and a stringent statistical laboratory for comparing alternative reconciliation strategies, especially those that rely on regularised covariance estimation or exploit cross‑sectional information to stabilise forecasts when historical data are scant.
--->
-
-@Wickramasuriya2019-xq also argued that modelling spatial autocorrelations would be challenging as in the case of a large collection of time series.
-Reconciliation approaches have the advantage to implicitly model this spatial autocorreltion structure, especially when the MinT method is used.
-
-To assess forecasting performance, we adopt a rolling-window cross-validation scheme.
-Beginning with the first 96 monthly observations (January 1998-December 2005) as the initial training set, we obtain the best-fitted ARIMA model for each of the 560 series via the automatic algorithm in the fabletools package, by minimising AICc [@fabletools-fa].
-The 1- to 12-step-ahead base forecasts are then generated by these ARIMA models.
-These incoherent base forecasts are reconciled via OLS, MinT with shrinkage (mint_shr), and MinT with the NOVELIST estimator (mint_n).
-We then roll the training window forward by one month and refit all models, rebuild reconciliations, and produce another batch of 1- to 12-step-ahead forecasts, repeating until December 2018.
-In total, this results in 156 out-of-sample windows and an equal number of forecast sets for each series.
+We quantify tourism demand via "visitor nights", the total number of nights spent by Australians away from home.
+The data is collected via the National Visitor Survey, managed by Tourism Research Australia, using computer assisted telephone interviews from nearly 120,000 Australian residents aged 15 years and over [@tourism-au].
 
@@
 
+The data are monthly time series spanning from January 1998 to December 2019, resulting in 264 observations per series, producing a canonical "$n \ll p$" setting which is ideal for evaluating reconciliation approaches that rely on high-dimensional covariance estimation.
+The extreme dimensionality over sample size mirrors many contemporary business problems, for instance, Starbucks drink sales.
+Tourism demand is also economically vital yet highly volatile, with geographical and purpose‑specific patterns create a realistic stress‑test for reconciliation algorithms.
+Consequently, this panel offers both a rich policy case study and a stringent statistical laboratory for comparing reconciliation strategies that exploit cross-series information to improve forecasts when historical data are scarce.
+
+<!-- This dataset is particularly well‑suited to evaluating forecast‑reconciliation methods for three reasons.
+First, tourism demand is economically important and notoriously volatile; the heterogeneous regional and purpose‑specific patterns create a realistic stress‑test for reconciliation algorithms.
+Second, the clear structural shift towards stronger growth after 2016—visible in most aggregates—necessitates techniques capable of handling possible breaks or time‑varying error structures, making the exercise practically relevant.
+Third, the extreme dimensionality relative to sample size mirrors many contemporary business problems (e.g., SKU‑level retail sales), allowing insights to generalise beyond tourism.
+Using this panel therefore provides both a compelling policy context and a stringent statistical laboratory for comparing alternative reconciliation strategies, especially those that rely on regularised covariance estimation or exploit cross‑sectional information to stabilise forecasts when historical data are scant.
+-->
+
+@Wickramasuriya2019-xq also argued that modelling spatial autocorrelations would be challenging as in the case of a large collection of time series.
+Reconciliation approaches have the advantage to implicitly model this spatial autocorreltion structure, especially when the MinT method is used.
+
+To assess forecasting performance, we adopt a rolling-window cross-validation scheme.
+Beginning with the first 96 monthly observations (January 1998-December 2005) as the initial training set, we obtain the best-fitted ARIMA model for each of the 560 series via the automatic algorithm in the fabletools package, by minimising AICc [@O-Hara-Wild2024-we].
+The 1- to 12-step-ahead base forecasts are then generated by these ARIMA models.
+These incoherent base forecasts are reconciled via OLS, MinT with shrinkage (mint_shr), and MinT with the NOVELIST estimator (mint_n).
+We then roll the training window forward by one month and refit all models, rebuild reconciliations, and produce another batch of 1- to 12-step-ahead forecasts, repeating until December 2018.
+In total, this results in 156 out-of-sample windows and an equal number of forecast sets for each series.
+
+Note that the number of series is larger than the number of observations (560 compared to 96), hence the sample covariance matrix is not positive definite and will not be considered.
+
+The results are presented in @fig-tourism-results, which shows the relative improvement of the MSE of reconciled forecasts over the base forecasts for the Australian domestic tourism flows.
+The results are grouped by levels of aggregation.
+MinT variants show improvement over the base ARIMA forecasts for middle to lower aggregation levels.
+We hardly differentiate the point accuracy performance between mint_shr and mint_n in these level, except at the most disaggregated, where mint_shr slightly edges out mint_n.
+
 ```{r fig STL tourism}
@@
 
-Note that the number of series is larger than the number of observations (560 compared to 96), hence the sample covariance matrix is not positive definite and will not be considered.
-
-The results are presented in @fig-tourism-results, which shows the relative improvement of the MSE of reconciled forecasts over the base forecasts for the Australian domestic tourism flows.
-The results are grouped by levels of aggregation.
-MinT variants show improvement over the base ARIMA forecasts for middle to lower aggregation levels.
-We hardly differentiate the point accuracy performance between mint_shr and mint_n in these level, except at the most disaggregated, where mint_shr slightly edges out mint_n.
-
 Surprisingly, however, MinT underperforms OLS at higher aggregation level and even the base forecasts at country-wise.
-This contradicts with the canonical results from @Wickramasuriya2019-xq and others, prompting further investigation.
-Applying STL decomposition to state-aggregated observations reveals a structural change around 2016: four large states (New South Wales, Victoria, Queensland, Western Australia) develop a modest upward trend in visitor nights, as shown in @fig-STL-tourism.
+This contradicts with the canonical results from @Wickramasuriya2019-xq, prompting further investigation.
+Applying STL decomposition by @Bandara2022-og to state-aggregated observations reveals a structural change around 2016: four large states (New South Wales, Victoria, Queensland, Western Australia) develop a modest upward trend in visitor nights, as shown in @fig-STL-tourism.
 The trend is likely due to the increasing popularity of tourism in these places, as well as the impact of various factors such as economic growth, population growth, and changes in consumer preferences.
@@
 
+\pagebreak
+
+# Timeline & Milestones
+
+<!--
 # Discussion and Conclusion
@@
 -   Limitations of the research
+-->
 

### 2025-07-31 literature/paper_draft.qmd
update sim hierarchical section

@@
 
-- 2x2, 6x6, 50x2
-
 Using the data generating process described above, we consider three different 2-level hierarchical structures, with the bottom-level series $n_b$ being 2 groups of 2 (4 series), 6 groups of 6 (36 series), and 2 groups of 50 (100 series), respectively.
@@
 The third structure is larger, having 100 bottom series aggregated in groups of size 50 to form 2 level-1 series, then aggregated to get the top level.
+
   Because of this structure, the VAR(1) coefficients matrix would have entries between groups being zeros.
@@
 ```{r load results}
+sim1_1 <- readRDS("./literature/lib/S4-2-1_T50_M500.rds")
+sim1_2 <- readRDS("./literature/lib/S4-2-1_T300_M500.rds")
 
-sim1_1 <- readRDS("./literature/lib/S4-2-1_T100_M10000_par.rds")
-sim1_2 <- readRDS("./literature/lib/S4-2-1_T300_M10000.rds")
-
-sim2_1 <- readRDS("./literature/lib/S36-6-1_T100_M100_run1.rds")
-sim2_2 <- readRDS("./literature/lib/S36-6-1_T300_M100_run2.rds")
+sim2_1 <- readRDS("./literature/lib/S36-6-1_T50_M500.rds")
+sim2_2 <- readRDS("./literature/lib/S36-6-1_T300_M500.rds")
 
@@
 ```{r generate plots for A and Sigma}
-# p1_A <- plot_heatmap(sim1_1$A, T) + 
-#   ggtitle("2x2 VAR(1) Coefficients Matrix")
-# 
-# p1_Sigma <- plot_heatmap(cov2cor(sim1_1$Sigma), T) + 
-#   ggtitle("2x2 Correlation Matrix")
-# 
+p1_A <- plot_heatmap(sim1_1$A, T) + 
+  ggtitle("2x2 VAR(1) Coefficients Matrix")
+
+p1_Sigma <- plot_heatmap(cov2cor(sim1_1$Sigma), T) + 
+  ggtitle("2x2 Correlation Matrix")
+
 p2_A <- plot_heatmap(sim2_1$A) +
@@
 
-# p3_A <- plot_heatmap(sim3_1$A) +
-#   ggtitle("50x2 VAR(1) Coefficients Matrix")
-# 
-# p3_Sigma <- plot_heatmap(cov2cor(sim3_1$Sigma), T) +
-#   ggtitle("50x2 Correlation Matrix")
+p3_A <- plot_heatmap(sim3_1$A) +
+  ggtitle("50x2 VAR(1) Coefficients Matrix")
+
+p3_Sigma <- plot_heatmap(cov2cor(sim3_1$Sigma), T) +
+  ggtitle("50x2 Correlation Matrix")
 ```
 
-```{r}
-#| out.width: "30%"
+```{r fig settings 2x2}
+#| label: fig-settings-2x2
+#| fig-cap: "Heatmaps of the VAR(1) coefficient matrix and correlation matrix for the 2x2 structure."
+#| fig.height: 4
+#| fig.width: 6
+#| out.width: "55%"
 #| fig.align: "center"
-# (p1_A | p1_Sigma) +
-#   plot_layout(guides = "collect") & theme(legend.position = "right")
+
+p1_A + p1_Sigma + plot_layout(ncol = 2) & theme(legend.position = "bottom")
 ```
 
-```{r}
-#| out.width: "60%"
-#| fig.align: "center"
-# (p2_A | p2_Sigma) +
-#   plot_layout(guides = "collect") & theme(legend.position = "right")
-```
-
-```{r}
-#| out.width: "100%"
-#| fig.align: "center"
-# (p3_A | p3_Sigma) +
-#   plot_layout(guides = "collect") & theme(legend.position = "right")
-```
-
-```{r fig-settings-6x6}
+```{r fig settings 6x6}
 #| label: fig-settings-6x6
-#| fig-cap: "Heatmaps of the VAR(1) coefficient matrix and correlation matrix for the 6×6 structure."
+#| fig-cap: "Heatmaps of the VAR(1) coefficient matrix and correlation matrix for the 50x2 structure."
+#| fig.height: 4
+#| fig.width: 6
 #| out.width: "85%"
@@
 p2_A + p2_Sigma + plot_layout(ncol = 2) & theme(legend.position = "bottom")
+```
+
+```{r fig settings 50x2}
+#| label: fig-settings-50x2
+#| fig-cap: "Heatmaps of the VAR(1) coefficient matrix and correlation matrix for the 6×6 structure."
+#| fig.height: 12
+#| fig.width: 8
+#| out.width: "85%"
+#| fig.align: "center"
+
+p3_A + p3_Sigma + plot_layout(ncol = 1) & theme(legend.position = "right")
 ```

### 2025-07-31 literature/paper_draft.qmd
initial draft dense and sparse simulation

@@
 
-Not much diff
+```{r}
+sim_dense_50 <- readRDS("./literature/lib/S100-10-3-1_T50_M200_dense.rds")
+sim_dense_300 <- readRDS("./literature/lib/S100-10-3-1_T300_M200_dense.rds")
 
+sim_sparse_50 <- readRDS("./literature/lib/S100-10-3-1_T50_M500_sparsebwgrp.rds")
+sim_sparse_300 <- readRDS("./literature/lib/S100-10-3-1_T300_M200_sparsebwgrp.rds")
+```
+
+```{r}
+#| label: fig-Sigma-50x2
+#| fig-cap: "Heatmaps of the dense and sparse correlation matrix of the data generating process."
+#| fig.height: 12
+#| fig.width: 8
+#| out.width: "100%"
+#| fig.align: "center"
+p_dense_Sigma <- sim_dense_50$Sigma |>
+  cov2cor() |> 
+  plot_heatmap(T) +
+    labs(title = "Dense")
+
+p_sparse_Sigma <- sim_sparse_50$Sigma |>
+  cov2cor() |> 
+  plot_heatmap(T) +
+    labs(title = "Sparse")
+
+p_dense_Sigma + p_sparse_Sigma + 
+  plot_layout(ncol = 1, guides = "collect", axis_titles = "collect") & 
+  theme(legend.position = "right")
+```
+
+```{r}
+#| label: fig-sim-results-dsp
+#| fig-cap: "Relative improvement of the MSE of reconciled forecasts over the base forecasts for the 50x2 hierarchical structure with dense and sparse DGP's correlation matrix, for 1 to 4 steps ahead forecasts, with 2 time series lengths (T = 54 and T = 304)."
+#| fig.height: 10
+#| fig.width: 10
+#| out.width: "90%"
+#| fig.align: "center"
+p_dense_50 <- sim_dense_50$MSE |>
+  transform_sim_MSE(F) |> 
+  plot_relative_improvement() +
+    labs(title = "Dense DGP Covariance Matrix", subtitle = "T = 54")
+p_sparse_50 <- sim_sparse_50$MSE |>
+  transform_sim_MSE(F) |> 
+  plot_relative_improvement() +
+    labs(title = "Dense DGP Covariance Matrix")
+
+p_dense_300 <- sim_dense_300$MSE |>
+  transform_sim_MSE(F) |> 
+  filter(.model != "ols") |> 
+  plot_relative_improvement() +
+    labs(subtitle = "T = 304")
+p_sparse_300 <- sim_sparse_300$MSE |>
+  transform_sim_MSE(F) |> 
+  plot_relative_improvement()
+
+p_dense_50 + p_sparse_50 + p_dense_300 + p_sparse_300 +
+  plot_layout(ncol = 2, guides = "collect", axis_titles = "collect") &
+  theme(legend.position = "right")
+```
+
+In this section we explore the effects of "sparse" and "dense" correlation matrix in our data generating process for series at the bottom level.
+We consider a large structure, having 2 groups of 50 bottom series.
+The cross-series within-group correlations are strong, and the between-group correlations will be weaker for the "dense" case, and zero for the "sparse" case.
+The dense and sparse correlation matrices for the DGP of bottom series are illustrated in @fig-Sigma-50x2.
+Note that these 2 cases have the same VAR(1) structure, only the correlation matrix is different.
+
+However, rather than aggregating series within group to form 2 level-1 series as we did earlier, we combine them in groups of size 10 to obtain 10 level-1 series.
+These 10 series will be splitted into 3 groups, having 4, 3, and 4 series respectively, and are aggregated into 3 level-2 series.
+Finally, they are added to obtain the top level.
+The intention behind this unusual aggregation path is to create a more complex structure for higher levels, in which the level-2 series will correlate with the bottom series in the first or second group, or both.
+
+Results are presented in @fig-sim-results-dsp, with the graphs in the first column being the "dense" case and the ones in second column being "sparse" case.
+We consider two time series length--54 and 304 observations with the last 4 being the test set--presenting at the first and second row respectively.
+As shown in the plots, MinT with either Shrinkage (mint_shr) or NOVELIST estimator (mint_n) return significant improvements over the base forecasts, especially in the 1-step-ahead horizon.
+However, the difference between their accuracy is hardly noticeable, with mint_n slightly outperform at the 1-step-ahead and mint_shr takes over in further forecasting horizons.
+mint_sample performs worse than base ARIMA models in short time series length, only to improve drastically in longer length.
+It is due to the sample covariance matrix gets more accurate as sample size gets larger.
+
+These simulations
+
+<!--
 ## Exploring Grouped Structure
@@
 - Finding: NOVELIST performs well in both cases
+-->
 

### 2025-08-01 literature/paper_draft.qmd
move figures in 1st sim to appendix

@@
 Because of this structure, the VAR(1) coefficients matrix would have entries between groups being zeros.
-To save space, we will only illustrate the settings of the second structure (6 groups of 6) in @fig-settings-6x6, and the first and third structure is similar.
+To save space, we will only illustrate the settings of the second structure (6 groups of 6) in @fig-settings-6x6, and the illustrations of the first and third structure is attached in @sec-sim-supplement Appendix.
 
@@
 
-```{r fig settings 2x2}
-#| label: fig-settings-2x2
-#| fig-cap: "Heatmaps of the VAR(1) coefficient matrix and correlation matrix for the 2x2 structure."
-#| fig.height: 4
-#| fig.width: 6
-#| out.width: "55%"
-#| fig.align: "center"
-
-p1_A + p1_Sigma + plot_layout(ncol = 2) & theme(legend.position = "bottom")
-```
-
 ```{r fig settings 6x6}
@@
 p2_A + p2_Sigma + plot_layout(ncol = 2) & theme(legend.position = "bottom")
-```
-
-```{r fig settings 50x2}
-#| label: fig-settings-50x2
-#| fig-cap: "Heatmaps of the VAR(1) coefficient matrix and correlation matrix for the 6×6 structure."
-#| fig.height: 12
-#| fig.width: 8
-#| out.width: "85%"
-#| fig.align: "center"
-
-p3_A + p3_Sigma + plot_layout(ncol = 1) & theme(legend.position = "right")
 ```
@@
 
+## Simulation Settings: Supplementary Figures {#sec-sim-supplement}
+
+```{r fig settings 2x2}
+#| label: fig-settings-2x2
+#| fig-cap: "Heatmaps of the VAR(1) coefficient matrix and correlation matrix for the 2x2 structure."
+#| fig.height: 4
+#| fig.width: 6
+#| out.width: "55%"
+#| fig.align: "center"
+
+p1_A + p1_Sigma + plot_layout(ncol = 2) & theme(legend.position = "bottom")
+```
+
+```{r fig settings 50x2}
+#| label: fig-settings-50x2
+#| fig-cap: "Heatmaps of the VAR(1) coefficient matrix and correlation matrix for the 6×6 structure."
+#| fig.height: 12
+#| fig.width: 8
+#| out.width: "85%"
+#| fig.align: "center"
+
+p3_A + p3_Sigma + plot_layout(ncol = 1) & theme(legend.position = "right")
+```
+
 ## Algorithm: NOVELIST cross-validation for optimal threshold $\delta^*$ {#sec-novelist_cv}

### 2025-08-01 literature/paper_draft.qmd
refine 2nd sim setting

@@
 
-In this section we explore the effects of "sparse" and "dense" correlation matrix in our data generating process for series at the bottom level.
-We consider a large structure, having 2 groups of 50 bottom series.
-The cross-series within-group correlations are strong, and the between-group correlations will be weaker for the "dense" case, and zero for the "sparse" case.
-The dense and sparse correlation matrices for the DGP of bottom series are illustrated in @fig-Sigma-50x2.
-Note that these 2 cases have the same VAR(1) structure, only the correlation matrix is different.
+In our second simulation study, we design a data‐generating process that contrasts “dense” and “sparse” correlation regimes among bottom‐level series, reflecting settings one might encounter in practice.
+Specifically, we construct two groups of fifty series each, with strong within-group dependencies throughout and either modest between‐group correlations (the dense scenario) or complete independence (the sparse scenario).
+These correlation matrices are depicted in Figure @fig-Sigma-50x2.
+Both scenarios share the same VAR(1) coefficient structure as in our previous simulations; only the innovation covariance changes.
+Such a setup mirrors real‐world contexts where, for example, sales within a product line may exhibit strong co‐movements, while those in a separate line operate nearly independently.
 
-However, rather than aggregating series within group to form 2 level-1 series as we did earlier, we combine them in groups of size 10 to obtain 10 level-1 series.
-These 10 series will be splitted into 3 groups, having 4, 3, and 4 series respectively, and are aggregated into 3 level-2 series.
-Finally, they are added to obtain the top level.
-The intention behind this unusual aggregation path is to create a more complex structure for higher levels, in which the level-2 series will correlate with the bottom series in the first or second group, or both.
+Rather than collapsing each group into a single aggregate as in previous practice, we impose a more intricate, multi‐level aggregation path to stress‐test reconciliation methods.
+We first form ten intermediate series by summing bottom‐level series in consecutive blocks of ten.
+These ten series are then assigned to three second‐level aggregates--groups of four, three, and four--and finally summed to produce a single top‐level series.
+This asymmetric hierarchy creates overlapping correlation patterns: some level‐2 series share bottom‐level groups, while others draw from both.
+<!-- emulating practical scenarios such as regional sales aggregations that span multiple product categories or overlapping territories. -->
 
-Results are presented in @fig-sim-results-dsp, with the graphs in the first column being the "dense" case and the ones in second column being "sparse" case.
-We consider two time series length--54 and 304 observations with the last 4 being the test set--presenting at the first and second row respectively.
-As shown in the plots, MinT with either Shrinkage (mint_shr) or NOVELIST estimator (mint_n) return significant improvements over the base forecasts, especially in the 1-step-ahead horizon.
-However, the difference between their accuracy is hardly noticeable, with mint_n slightly outperform at the 1-step-ahead and mint_shr takes over in further forecasting horizons.
-mint_sample performs worse than base ARIMA models in short time series length, only to improve drastically in longer length.
-It is due to the sample covariance matrix gets more accurate as sample size gets larger.
+Figure @fig-sim-results-dsp presents out‐of‐sample mean squared error improvements over the base forecasts for each reconciliation strategy under both dense and sparse settings, with two panel lengths--54 and 304 observations--reserving the last four points for testing.
+In both short and long samples, MinT using either the shrinkage estimator (mint_shr) or the NOVELIST estimator (mint_n) delivers pronounced gains over incoherent ARIMA forecasts, particularly at the one‐step horizon where cross‐series correlations most directly inform the forecast adjustments.
+Although the two MinT variants perform almost indistinguishably overall, mint_n edges out mint_shr in the immediate horizon, whereas mint_shr slightly outperforms for longer horizons.
+By contrast, MinT with the raw sample covariance (mint_sample) suffers in small‐sample settings; as expected, its performance improves dramatically with 304 data points, since the sample covariance becomes more reliable with larger $n$.
+This highlights the practical necessity of regularized estimators in high‐dimensional, low‐sample contexts, a situation common in real applications where histories are short relative to the number of series.
 
-These simulations
+We also experimented with alternative DGPs: varying group sizes, aggregation paths, and VAR(1) structure.
+However, to our surprise, most designs failed to distinguish mint_shr from mint_n in any substantive way.
+Their nearly identical performance under these synthetic scenarios suggests that our current simulation may not unveil the full advantages of the thresholding estimators.
+This finding motivates our turn to empirical data, where latent structural features and regime shifts, which we will discuss in the next section, may reveal performance differences.
 

### 2025-08-01 literature/paper_draft.qmd
minor change

@@
     # labs(x = element_blank(), y = element_blank()) +
-    theme_minimal()
+    theme_bw()
 }
@@
 
+In our second simulation study, we design a data‐generating process that contrasts “dense” and “sparse” correlation regimes among bottom‐level series, reflecting settings one might encounter in practice.
+Specifically, we construct two groups of fifty series each, with strong within-group dependencies throughout and either modest between‐group correlations (the dense scenario) or complete independence (the sparse scenario).
+These correlation matrices are depicted in Figure @fig-Sigma-50x2.
+Both scenarios share the same VAR(1) coefficient structure as in our previous simulations; only the innovation covariance changes.
+Such a setup mirrors real‐world contexts where, for example, sales within a product line may exhibit strong co‐movements, while those in a separate line operate nearly independently.
+
+Rather than collapsing each group into a single aggregate as in previous practice, we impose a more intricate, multi‐level aggregation path to stress‐test reconciliation methods.
+We first form ten intermediate series by summing bottom‐level series in consecutive blocks of ten.
+These ten series are then assigned to three second‐level aggregates--groups of four, three, and four--and finally summed to produce a single top‐level series.
+This asymmetric hierarchy creates overlapping correlation patterns: some level‐2 series share bottom‐level groups, while others draw from both.
+<!-- emulating practical scenarios such as regional sales aggregations that span multiple product categories or overlapping territories. -->
+
 ```{r}
@@
 #| fig-cap: "Relative improvement of the MSE of reconciled forecasts over the base forecasts for the 50x2 hierarchical structure with dense and sparse DGP's correlation matrix, for 1 to 4 steps ahead forecasts, with 2 time series lengths (T = 54 and T = 304)."
-#| fig.height: 10
-#| fig.width: 10
+#| fig.height: 7
+#| fig.width: 8
 #| out.width: "90%"
@@
 ```
-
-In our second simulation study, we design a data‐generating process that contrasts “dense” and “sparse” correlation regimes among bottom‐level series, reflecting settings one might encounter in practice.
-Specifically, we construct two groups of fifty series each, with strong within-group dependencies throughout and either modest between‐group correlations (the dense scenario) or complete independence (the sparse scenario).
-These correlation matrices are depicted in Figure @fig-Sigma-50x2.
-Both scenarios share the same VAR(1) coefficient structure as in our previous simulations; only the innovation covariance changes.
-Such a setup mirrors real‐world contexts where, for example, sales within a product line may exhibit strong co‐movements, while those in a separate line operate nearly independently.
-
-Rather than collapsing each group into a single aggregate as in previous practice, we impose a more intricate, multi‐level aggregation path to stress‐test reconciliation methods.
-We first form ten intermediate series by summing bottom‐level series in consecutive blocks of ten.
-These ten series are then assigned to three second‐level aggregates--groups of four, three, and four--and finally summed to produce a single top‐level series.
-This asymmetric hierarchy creates overlapping correlation patterns: some level‐2 series share bottom‐level groups, while others draw from both.
-<!-- emulating practical scenarios such as regional sales aggregations that span multiple product categories or overlapping territories. -->
 

### 2025-08-01 literature/paper_draft.qmd
use 50x2 with weird agg path for 1st sim section as well

@@
 
-## Exploring Effects of Hierarchy's Size
-
-Using the data generating process described above, we consider three different 2-level hierarchical structures, with the bottom-level series $n_b$ being 2 groups of 2 (4 series), 6 groups of 6 (36 series), and 2 groups of 50 (100 series), respectively.
-The first structure is exactly the same as the one in @fig-hierarchy, with 2 series at the level 1.
-The second structure has 36 bottom series and are aggregated in groups of size 6 to form 6 level-1 series, which are finally aggregated to form the top level.
-The third structure is larger, having 100 bottom series aggregated in groups of size 50 to form 2 level-1 series, then aggregated to get the top level.
-
-Because of this structure, the VAR(1) coefficients matrix would have entries between groups being zeros.
-To save space, we will only illustrate the settings of the second structure (6 groups of 6) in @fig-settings-6x6, and the illustrations of the first and third structure is attached in @sec-sim-supplement Appendix.
-
 ```{r load results}
@@
 
-sim3_1 <- readRDS("./literature/lib/S100-2-1_T50_M200_run1.rds")
-sim3_2 <- readRDS("./literature/lib/S100-2-1_T100_M200_run1.rds")
+# sim3_1 <- readRDS("./literature/lib/S100-2-1_T50_M200_run1.rds")
+# sim3_2 <- readRDS("./literature/lib/S100-2-1_T100_M200_run1.rds")
 
-MSE1_1 <- transform_sim_MSE(sim1_1$MSE, F)
-MSE1_2 <- transform_sim_MSE(sim1_2$MSE, F)
+sim_dense_50 <- readRDS("./literature/lib/S100-10-3-1_T50_M200_dense.rds")
+sim_dense_300 <- readRDS("./literature/lib/S100-10-3-1_T300_M200_dense.rds")
 
-MSE2_1 <- transform_sim_MSE(sim2_1$MSE, F)
-MSE2_2 <- transform_sim_MSE(sim2_2$MSE, F)
+sim_sparse_50 <- readRDS("./literature/lib/S100-10-3-1_T50_M500_sparsebwgrp.rds")
+sim_sparse_300 <- readRDS("./literature/lib/S100-10-3-1_T300_M200_sparsebwgrp.rds")
+
+sim3_1 <- sim_dense_50 # using same with next section simulation
+sim3_2 <- sim_dense_300
+```
+
+## Exploring Effects of Hierarchy's Size
+
+In this setting, we aim to assess the performance of MinT with NOVELIST when the size of the hierarchy becomes larger.
+Using the data generating process described above, we consider three different 2-level hierarchical structures, with the bottom-level series $n_b$ being 2 groups of 2 (4 series), 6 groups of 6 (36 series), and 2 groups of 50 (100 series), respectively.
+The first structure is exactly the same as the one in @fig-hierarchy, with 2 series at the level 1.
+The second structure has 36 bottom series and are aggregated in groups of size 6 to form 6 level-1 series, which are finally aggregated to form the top level.
+
+The third structure is larger, we construct two groups of fifty series each for the bottom level.
+However, rather than collapsing each group into a single aggregate as in previous hierarchies, we impose a more intricate, multi‐level aggregation path to stress‐test reconciliation methods.
+We first form ten intermediate series by summing bottom‐level series in consecutive blocks of ten.
+These ten series are then assigned to three second‐level aggregates--groups of four, three, and four--and finally summed to produce a single top‐level series.
+This asymmetric hierarchy creates overlapping correlation patterns: some level‐2 series share bottom‐level groups, while others draw from both, emulating practical scenarios such as regional sales aggregations that span multiple product categories or overlapping territories.
+
+The VAR(1) coefficients matrices are block diagonal and would have entries between groups being zeros.
+To save space, we will only illustrate the settings of the second structure (6 groups of 6) in @fig-settings-6x6, and the illustrations of the first and third structure is attached in @sec-sim-supplement Appendix.
+
+```{r generate plots for A and Sigma}
+MSE1_1 <- transform_sim_MSE(sim1_1$MSE, F) |> 
+  filter(.model != "ols")
+MSE1_2 <- transform_sim_MSE(sim1_2$MSE, F) |> 
+  filter(.model != "ols")
+
+MSE2_1 <- transform_sim_MSE(sim2_1$MSE, F) |> 
+  filter(! .model %in% c("mint_n_hstep", "ols"))
+MSE2_2 <- transform_sim_MSE(sim2_2$MSE, F) |> 
+  filter(! .model %in% c("mint_n_hstep", "ols"))
 
 MSE3_1 <- transform_sim_MSE(sim3_1$MSE, F)
-MSE3_2 <- transform_sim_MSE(sim3_2$MSE, F)
-```
+MSE3_2 <- transform_sim_MSE(sim3_2$MSE, F) |> 
+  filter(.model != "ols")
 
-```{r generate plots for A and Sigma}
 p1_A <- plot_heatmap(sim1_1$A, T) + 
@@
 In our second simulation study, we design a data‐generating process that contrasts “dense” and “sparse” correlation regimes among bottom‐level series, reflecting settings one might encounter in practice.
-Specifically, we construct two groups of fifty series each, with strong within-group dependencies throughout and either modest between‐group correlations (the dense scenario) or complete independence (the sparse scenario).
+We consider the same hierarchical structure of two large groups of 50 bottom series as above.
+Specifically, the two groups would have strong within-group dependencies throughout and either modest between‐group correlations (the dense scenario) or complete independence (the sparse scenario).
 These correlation matrices are depicted in Figure @fig-Sigma-50x2.
@@
 
-Rather than collapsing each group into a single aggregate as in previous practice, we impose a more intricate, multi‐level aggregation path to stress‐test reconciliation methods.
+<!-- Rather than collapsing each group into a single aggregate as in previous practice, we impose a more intricate, multi‐level aggregation path to stress‐test reconciliation methods.
 We first form ten intermediate series by summing bottom‐level series in consecutive blocks of ten.
@@
 This asymmetric hierarchy creates overlapping correlation patterns: some level‐2 series share bottom‐level groups, while others draw from both.
+-->
 <!-- emulating practical scenarios such as regional sales aggregations that span multiple product categories or overlapping territories. -->
-
-```{r}
-sim_dense_50 <- readRDS("./literature/lib/S100-10-3-1_T50_M200_dense.rds")
-sim_dense_300 <- readRDS("./literature/lib/S100-10-3-1_T300_M200_dense.rds")
-
-sim_sparse_50 <- readRDS("./literature/lib/S100-10-3-1_T50_M500_sparsebwgrp.rds")
-sim_sparse_300 <- readRDS("./literature/lib/S100-10-3-1_T300_M200_sparsebwgrp.rds")
-```
 

### 2025-08-01 literature/paper_draft.qmd
finish draft for 1st sim section

@@
 
-In this setting, we aim to assess the performance of MinT with NOVELIST when the size of the hierarchy becomes larger. As shown in @fig-sim-results-1, ...
+As shown in @fig-sim-results-1, the larger, more complex the hierarchy is, the better reconciliation strategies perform across 1- to 4-step-ahead forecasting horizons.
+The length of training set gives better-fitted models, but does not affect the relative improvements of the methods over the base forecasts, except MinT with sample covariance (mint_sample).
+This is evidence that these high-dimensional covariance estimator--shrinkage and NOVELIST--successfully handle "$n \ll p$" data.
+However, MinT with either estimators rarely differ.
+
+NOVELIST has the flexibility to minimise the effects of weak correlations and retain the stronger ones, as opposed to shrinkage.
+Thus, this prompts us to give NOVELIST a scenario where the covariance matrix has entries that is noisy while they are, in fact, zeros.
+We simulate this via a sparse covariance structure of the data generating process for bottom series, which will be discussed next.
 

### 2025-08-01 literature/paper_draft.qmd
refine 1st sim section

--- 2025-08-01 literature/paper_draft.qmd
@@
 
-In this setting, we aim to assess the performance of MinT with NOVELIST when the size of the hierarchy becomes larger.
-Using the data generating process described above, we consider three different 2-level hierarchical structures, with the bottom-level series $n_b$ being 2 groups of 2 (4 series), 6 groups of 6 (36 series), and 2 groups of 50 (100 series), respectively.
-The first structure is exactly the same as the one in @fig-hierarchy, with 2 series at the level 1.
-The second structure has 36 bottom series and are aggregated in groups of size 6 to form 6 level-1 series, which are finally aggregated to form the top level.
+In our first set of experiments, we examine how MinT combined with the NOVELIST estimator scales as the hierarchy expands.
+We generate synthetic data from the same VAR(1) framework described earlier, but vary the number of bottom‐level series, $n_b$, across three different structures: the smallest with two groups of two ($n_b = 4$), an intermediate case with six groups of six ($n_b = 36$), and a much larger configuration with two clusters of fifty ($n_b = 100$).
+In the 4‐series hierarchy, each pair collapses into a single level‐1 series, which then sums to the top.
+In the 36‐series case, each block of six forms a level‐1 aggregate, and those six aggregates form the national total.
 
-The third structure is larger, we construct two groups of fifty series each for the bottom level.
-However, rather than collapsing each group into a single aggregate as in previous hierarchies, we impose a more intricate, multi‐level aggregation path to stress‐test reconciliation methods.
-We first form ten intermediate series by summing bottom‐level series in consecutive blocks of ten.
-These ten series are then assigned to three second‐level aggregates--groups of four, three, and four--and finally summed to produce a single top‐level series.
+The 100‐series design employs a deliberately intricate aggregation path to to stress‐test reconciliation methods.
+We first sum the one hundred bottom series into ten intermediate series by grouping them in contiguous blocks of ten.
+These ten series are then organised into three level-2 aggregates—four, three, and four series, respectively—before finally summing to a single top node.
 This asymmetric hierarchy creates overlapping correlation patterns: some level‐2 series share bottom‐level groups, while others draw from both, emulating practical scenarios such as regional sales aggregations that span multiple product categories or overlapping territories.
 
-The VAR(1) coefficients matrices are block diagonal and would have entries between groups being zeros.
-To save space, we will only illustrate the settings of the second structure (6 groups of 6) in @fig-settings-6x6, and the illustrations of the first and third structure is attached in @sec-sim-supplement Appendix.
+Across all designs, the VAR(1) coefficient matrices remain block-diagonal, ensuring that series within a block interact but blocks remain orthogonal to each other.
+To save space, we illustrate only the six-by-six's VAR(1) and correlation configuration in @fig-settings-6x6; the two-by-two and the one-hundred-series structures appear in Appendix @sec-sim-supplement.
 
@@
 
-For each series, $T = 54$ and $304$ observations are generated.
-The first 50 and 300 observations are used for training, and the last 4 observations are used for testing.
-The training data is used to compute the best fitted ARIMA models by minimising the AICc criterion, in which we use the automatic algorithm from Fabletools R package [@O-Hara-Wild2024-we].
-We refer to them as base models, and their base forecasts are then reconciled using the MinT with different covariance estimators.
-These include using the unbiased sample covariance matrix - mint_sample, the shrinkage estimator - mint_shr, and the NOVELIST estimator - mint_n.
-The Monte Carlo simulation is repeated $M = 200$ times, in which the parameters for data generating process is fixed.
+For each hierarchy, we simulate two panel lengths, $T=54$ and $T=304$, reserving the final four observations as an out-of-sample test set.
+In each Monte Carlo replication ($M=500$), we fit univariate ARIMA models (base models) to the training observations using an automatic AICc minimization algorithm from the fabletools package [@O-Hara-Wild2024-we], generating incoherent 1–4-step base forecasts.
+We then reconcile these forecasts under three covariance estimators: the raw sample covariance (mint_sample), the shrinkage estimator (mint_shr), and the NOVELIST estimator (mint_n).
 
@@
 
-In this setting, we aim to assess the performance of MinT with NOVELIST when the size of the hierarchy becomes larger. As shown in @fig-sim-results-1, ...
+Figure @fig-sim-results-1 illustrates that as hierarchy size and complexity increase, the relative improvements in MSE over the base forecasts from reconciliation amplify, across all 1- to 1-step horizons.
+While extending the training window (from 50 to 300 in‐sample observations) uniformly improves the accuracy of base ARIMA forecasts, it does not alter the percentage improvements of mint_shr or mint_n.
+This is evidence that both high-dimensional estimators robustly handle the "large $p \gg T$” regime.
+By contrast, mint_sample achieves significant accuracy when more data make the sample covariance more reliable.
 
+Despite their theoretical differences, mint_shr and mint_n exhibit nearly identical performance in these canonical settings.
+Yet NOVELIST’s thresholding ability to minimise effects of weak correlations and retain strong ones suggests it may excel in contexts where correlation matrix has small, noisy entries.
+This motivates our subsequent exploration of sparse covariance structures, in which many off-diagonals are truly zero.
 <!--

### 2025-08-01 literature/paper_draft.qmd
refining whole sim section

@@
 
-- Talk about the history and evolution of forecasting hierarchical time series, starting from the early heuristic methods to the modern statistical approaches.
+<!-- - Talk about the history and evolution of forecasting hierarchical time series, starting from the early heuristic methods to the modern statistical approaches.
 
@@
   - Bayesian, Machine learning
-  - Probabilistic methods
+  - Probabilistic methods -->
 
@@
 
-To overcome these issues, forecast reconciliation was introduced by @Hyndman2011-jv, and later developed by @Hyndman2016-ic, @Wickramasuriya2019-xq, and ...
-to achieve coherency in point forecasts and enhance accuracy.
+To overcome these issues, forecast reconciliation was introduced by @Hyndman2011-jv, and later developed by @Hyndman2016-ic, @Wickramasuriya2019-xq, and others to achieve coherency in point forecasts and enhance accuracy.
 Forecast reconciliation projects a collection of independent base forecasts into a set of coherent forecasts that respect the linear constraints defining a hierarchical or grouped time-series system.
@@
 
-- Discuss the interests in MinT and how it has become a standard method for forecast reconciliation.
+<!-- - Discuss the interests in MinT and how it has become a standard method for forecast reconciliation.
 
@@
 
-- Talk about the paper outline and what will be covered in the following sections.
+- Talk about the paper outline and what will be covered in the following sections. -->
 
@@
 
+For all hierarchies in our experiments, we simulate two panel lengths, $T=54$ and $T=304$, reserving the final four observations as an out-of-sample test set.
+In each Monte Carlo replication ($M=500$), we fit univariate ARIMA models (base models) to the training observations using an automatic AICc minimization algorithm from the fabletools package [@O-Hara-Wild2024-we], generating incoherent 1–4-step base forecasts.
+We then reconcile these forecasts under three covariance estimators: the raw sample covariance (mint_sample), the shrinkage estimator (mint_shr), and the NOVELIST estimator (mint_n).
+
 ```{r load results}
@@
 
-The 100‐series design employs a deliberately intricate aggregation path to to stress‐test reconciliation methods.
+The 100‐series design employs a deliberately intricate aggregation path to stress‐test reconciliation methods.
 We first sum the one hundred bottom series into ten intermediate series by grouping them in contiguous blocks of ten.
@@
 
-Across all designs, the VAR(1) coefficient matrices remain block-diagonal, ensuring that series within a block interact but blocks remain orthogonal to each other.
 To save space, we illustrate only the six-by-six's VAR(1) and correlation configuration in @fig-settings-6x6; the two-by-two and the one-hundred-series structures appear in Appendix @sec-sim-supplement.
@@
 p3_A <- plot_heatmap(sim3_1$A) +
-  ggtitle("50x2 VAR(1) Coefficients Matrix")
+  ggtitle("2x50 VAR(1) Coefficients Matrix")
 
 p3_Sigma <- plot_heatmap(cov2cor(sim3_1$Sigma), T) +
-  ggtitle("50x2 Correlation Matrix")
+  ggtitle("2x50 Correlation Matrix")
 ```
@@
 #| label: fig-settings-6x6
-#| fig-cap: "Heatmaps of the VAR(1) coefficient matrix and correlation matrix for the 50x2 structure."
+#| fig-cap: "Heatmaps of the VAR(1) coefficient matrix and correlation matrix for the 2x50 structure."
 #| fig.height: 4
@@
 ```
-
-For each hierarchy, we simulate two panel lengths, $T=54$ and $T=304$, reserving the final four observations as an out-of-sample test set.
-In each Monte Carlo replication ($M=500$), we fit univariate ARIMA models (base models) to the training observations using an automatic AICc minimization algorithm from the fabletools package [@O-Hara-Wild2024-we], generating incoherent 1–4-step base forecasts.
-We then reconcile these forecasts under three covariance estimators: the raw sample covariance (mint_sample), the shrinkage estimator (mint_shr), and the NOVELIST estimator (mint_n).
 
@@
 #| label: fig-sim-results-1
-#| fig-cap: "Relative improvement of the MSE of reconciled forecasts over the base forecasts for the 2x2, 6x6, and 50x2 hierarchical structures, for 1 to 4 steps ahead forecasts, with 2 time series lengths (T = 54 and T = 304)."
+#| fig-cap: "Relative improvement of the MSE of reconciled forecasts over the base forecasts for the 2x2, 6x6, and 2x50 hierarchical structures, for 1 to 4 steps ahead forecasts, with 2 time series lengths (T = 54 and T = 304)."
 #| fig.width: 8
@@
   plot_relative_improvement() +
-    labs(title = "50x2 Hierarchy")
+    labs(title = "2x50 Hierarchy")
 
@@
 
-Figure @fig-sim-results-1 illustrates that as hierarchy size and complexity increase, the relative improvements in MSE over the base forecasts from reconciliation amplify, across all 1- to 1-step horizons.
-While extending the training window (from 50 to 300 in‐sample observations) uniformly improves the accuracy of base ARIMA forecasts, it does not alter the percentage improvements of mint_shr or mint_n.
+Figure @fig-sim-results-1 illustrates that as hierarchy size and complexity increase, the relative improvements in MSE over the base forecasts from reconciliation amplify, across all 1- to 4-step horizons.
+While extending the training window (from 50 to 300 in‐sample observations) uniformly improves the accuracy of base ARIMA forecasts, it does not alter the relative improvements of mint_shr or mint_n.
 This is evidence that both high-dimensional estimators robustly handle the "large $p \gg T$” regime.
@@
 
-- 50x2 with usual path and weird path
+- 2x50 with usual path and weird path
 
@@
 Specifically, the two groups would have strong within-group dependencies throughout and either modest between‐group correlations (the dense scenario) or complete independence (the sparse scenario).
-These correlation matrices are depicted in Figure @fig-Sigma-50x2.
+These correlation matrices are depicted in Figure @fig-Sigma-2x50.
 Both scenarios share the same VAR(1) coefficient structure as in our previous simulations; only the innovation covariance changes.
@@
 ```{r}
-#| label: fig-Sigma-50x2
+#| label: fig-Sigma-2x50
 #| fig-cap: "Heatmaps of the dense and sparse correlation matrix of the data generating process."
@@
 #| label: fig-sim-results-dsp
-#| fig-cap: "Relative improvement of the MSE of reconciled forecasts over the base forecasts for the 50x2 hierarchical structure with dense and sparse DGP's correlation matrix, for 1 to 4 steps ahead forecasts, with 2 time series lengths (T = 54 and T = 304)."
+#| fig-cap: "Relative improvement of the MSE of reconciled forecasts over the base forecasts for the 2x50 hierarchical structure with dense and sparse DGP's correlation matrix, for 1 to 4 steps ahead forecasts, with 2 time series lengths (T = 54 and T = 304)."
 #| fig.height: 7
@@
 
-We also experimented with alternative DGPs: varying group sizes, aggregation paths, and VAR(1) structure.
-However, to our surprise, most designs failed to distinguish mint_shr from mint_n in any substantive way.
+Additional designs (varying block sizes, aggregation paths, correlation configurations) also failed to separate NOVELIST from Shrinkage.
 Their nearly identical performance under these synthetic scenarios suggests that our current simulation may not unveil the full advantages of the thresholding estimators.
@@
 
-- A table/figure to illustrate the hierarchy
+<!-- - A table/figure to illustrate the hierarchy -->
 
@@
 
-```{r fig settings 50x2}
-#| label: fig-settings-50x2
+```{r fig settings 2x50}
+#| label: fig-settings-2x50
 #| fig-cap: "Heatmaps of the VAR(1) coefficient matrix and correlation matrix for the 6×6 structure."

### 2025-08-01 literature/paper_draft.qmd
expand on introduction section, includeing literature review

@@
 
-To overcome these issues, forecast reconciliation was introduced by @Hyndman2011-jv, and later developed by @Hyndman2016-ic, @Wickramasuriya2019-xq, and others to achieve coherency in point forecasts and enhance accuracy.
+To overcome these issues, forecast reconciliation was introduced by @Hyndman2011-jv, and later developed by @Van_Erven2015-nx, @Hyndman2016-ic, @Ben-Taieb2019-yx, @Wickramasuriya2019-xq, and others to achieve coherency in point forecasts and enhance accuracy.
 Forecast reconciliation projects a collection of independent base forecasts into a set of coherent forecasts that respect the linear constraints defining a hierarchical or grouped time-series system.
+
+Among the modern reconciliation strategies, the Min Trace (MinT) approach developed by @Wickramasuriya2019-xq is widely used and perform significantly well under right conditions.
+Despite its properties, MinT relies on a good high-dimensional covariance estimator, which is positive-definite.
+However, not many researchers have explored this issue in the current literature, except @Carrara2025-rz, who introduced a new estimator for MinT--Double Shrinkage estimator.
+The current stage of this research piece is fairly early, leaving many potential gaps still not yet to explore.
+As a result, a study of alternative high-dimensional covariance estimators for MinT is timely and desirable.
+
+The remainder of the paper is organised as follows.
+@sec-pre-theory provides the basic theoretical framework for hierarchical time series and forecasting, and Min Trace approach, introducing notations, terminologies, and motivations for alternative estimators.
+@sec-cov-est walks through the main covariance estimators this paper explores, and argues their strengths and weaknesses.
+@sec-simulation covers the simulation design and currently explores the performance of NOVELIST on MinT.
+@sec-empirical shows a real-world application of MinT with NOVELIST, which produces results that did not occured in the simulation settings, suggesting further inspection and analysis.
 

### 2025-08-01 literature/paper_draft.qmd
timeline section

@@
 
+<!-- # Supplementary -->
+
 \pagebreak
 
-# Timeline & Milestones
-
-<!--
-# Discussion and Conclusion
-
--   Evaluate how covariance estimation impacts the MinT reconciliation.
-
--   Advantages and limitations of the NOVELIST estimator (with different ways of choosing threshold parameter).
-
--   Practical considerations: Computational efficiency, robustness, and ease of implementation.
-
--   Limitations of the research
--->
+# Timeline & Milestones {#sec-timeline}
+
+```{r timeline}
+#| echo: false
+#| message: false
+#| warning: false
+library(tibble)
+library(knitr)
+library(kableExtra)
+
+tibble(
+  Period = c(
+    "March",
+    "April", 
+    "May", 
+    "June",
+    "July",
+    "August", 
+    "September",
+    "October"
+  ),
+  Task = c(
+    "Initial setup and understanding the research scope. Preliminary knowledge and Literature review",
+
+    "Develop ReconCov, an R package implementing the Min Trace approach with the shrinkage and NOVELIST estimators, and rolling-window cross-validation process for choosing NOVELIST threshold parameter.
+Exploring simulation framework.",
+
+    "Code functions for producing simulation settings in ReconCov package. Initial simulation for MinT in hierarchical time series. Writing research proposal and presenting research project.",
+
+    "Intensive simulation on multiple different settings, in order to assess the performance and behaviours of NOVELIST under different conditions, and to explore cases where NOVELIST performs different than Shrinkage.",
+
+    "Empirical analysis on forecasting Australian Domestic Tourism data. Explore the real-world data generating process which can be used to refine the current simulation design",
+
+    "Implement the Tourism DGP configuration into the current simulation settings.
+Implementing POET and PC-adjusted NOVELIST estimators.
+Assessing whether prominent principal components structure can be exploited to improve the out-of-sample forecast performance.",
+
+    "Exploring other high-dimensional covariance estimators. More simulations on all explored covariance estimators. Final,  full-fledged assessment on all implemented estimators",
+
+    "Final writing, presenation, submission"
+  ),
+  Deliverable = c(
+    "Initial literature review + paper draft",
+    "ReconCov package: Main codes for estimators + simulation framework + cross-validation",
+    "NOVELIST initial simulation results + research proposal + presentation",
+    "More simulation results on NOVELIST",
+    "Empirical results on forecasting Tourism data",
+    "New simulation design, POET & PC-adjusted NOVELIST simulation results",
+    "Other estimators simulation results and full-fledged covariance estimators assessment",
+    "Presentation and thesis manuscript"
+  )
+) %>%
+  kable(
+    format = "latex",
+    booktabs = TRUE,
+    longtable = TRUE,
+    caption = "Project Timeline and Key Milestones",
+    escape = TRUE
+  ) %>%
+  column_spec(1, width = "1.9cm") %>%
+  column_spec(2, width = "8cm") %>%
+  column_spec(3, width = "4.8cm") %>%
+  kable_styling(latex_options = "hold_position", font_size = 10)
+```
 

### 2025-08-01 literature/paper_draft.qmd
mention reconcov :)

@@
 
+All data generation, covariance estimation, and reconciliation routines were implemented in the ReconCov R package, a package developed solely for this research, and is available under an open‐source license on GitHub [@ReconCov].
+
 ```{r load results}
@@
 
+The cross-validation algorithm for NOVELIST is available in the ReconCov package [@ReconCov].
+
 ```{=latex}

### 2025-08-01 literature/paper_draft.qmd
mention h-step CV on NOVELIST

@@
 On the other hand, the optimal threshold $\delta^*$ does not have a closed-form solution, and is typically obtained by executing a rolling-window cross-validation procedure.
-The idea is to find the threshold $\hat{\delta^*}$, with the corresponding $\hat{\lambda}^*$ and $\hat{\boldsymbol{R}}^{N}_{1}(\hat{\delta^*}, \hat{\lambda}^*)$, that minimises the average out-of-sample reconciled forecast errors.
+The idea is to find the threshold $\hat{\delta^*}$, with the corresponding $\hat{\lambda}^*$ and $\hat{\boldsymbol{R}}^{N}_{1}(\hat{\delta^*}, \hat{\lambda}^*)$, that minimises the average out-of-sample 1-step-ahead reconciled forecast mean squared error over all windows.
 The formal algorithm is given in the @sec-novelist_cv Appendix.
@@
 
+We also tested out minimising 1- to h-step-ahead MSE in the cross-validation procedure.
+Surprisingly, it returns almost the same best threshold parameter is in the 1-step-ahead case above.
 Note that when $\delta \in \bigl[ \text{max}_{i \neq j}|\hat{r}_{1,ij}|, \; 1 \bigr]$, the NOVELIST estimator collapses to the shrinkage estimator, and when $\delta = 0$, it becomes the sample covariance matrix.

### 2025-08-01 literature/paper_draft.qmd
final touch for Milestone 4

@@
 title: "Enhancing Forecast Reconciliation: A Study of Alternative Covariance Estimators"
-author: Vincent Su
+author: 
+  - Vincent Su
+  - Shanika Wickramasuriya (supv.)
+  - George Athanasopoulos (supv.)
 format: 
@@
 
-# Theoretical Framework
+# Theoretical Framework {#sec-pre-theory}
 
@@
 
-# Covariance Estimation Approaches
+# Covariance Estimation Approaches {#sec-cov-est}
 
@@
 
-# Simulation
+# Simulation Design {#sec-simulation}
 
@@
 
-# Forecasting Australian Domestics Tourism
+# Forecasting Australian Domestics Tourism {#sec-empirical}
 

### 2025-08-18 phd_application/phd_propossal.Rmd
draft for phd proposal (done basckground + motiv)

---
title: "Decision-Optimal Probabilistic Forecast Reconciliation for Expected Shortfall in Hierarchical Time Series"
author: 
  - Vincent Su
format: 
  pdf:
    include-in-header:
      text: |
        <!-- \usepackage{amsmath} -->
        <!-- \usepackage{mathspec} -->
        \usepackage{algorithm}
        \usepackage{algpseudocode}
        \usepackage[colorlinks=true, linkcolor=blue, citecolor=blue, urlcolor=blue]{hyperref}
        \usepackage{array}
    mathspec: true
    number-sections: true
# crossref:
#   custom:
#     - kind: float
#       key: alg
#       reference-prefix: Algorithm
#       caption-prefix: Algorithm
#       # for PDF output only, tell LaTeX to use the `algorithm` env:
#       latex-env: algorithm
execute:
  cache: false
  echo: false
  warning: false
bibliography: D:/Github/Recon_Honours_Thesis/references.bib
# bibliographystyle: apa
csl: D:/Github/Recon_Honours_Thesis/apa.csl
editor_options: 
  chunk_output_type: console
---

<!-- # Abstract {-} -->

# Background

<!--
- Data is hierarchical
- Decision making requires forecasts at different levels of hierarchy
- Base forecasts are incoherent and accuracy suffers
- Reconciliation methods are used to adjust forecasts to be coherent
- Point forecasts methods are developed
- Probabilistic forecasts are sometimes the quantity of interest, especially in risk management
-->

Many operational datasets are naturally hierarchical. For Starbucks, their total sales is the sum of all countries they are operating in, and each national sales is the sum for all cities, down to many outlets, forming an aggregation constraint. Forecasting such hierarchical time series arises in many decision-making contexts ranging from supply chains and electricity generation planning, to macroeconomic and tourism analysis, where stakeholders need forecasts at several aggregation levels to allocate resources and manage risk.

In practice, when base forecasts are produced independently for all series, they typically fail to satisfy the aggregation constraints. Such forecasts are *incoherent*, which can undermine downstream decisions that require internal consistency. Forecast reconciliation addresses this by adjusting the base forecasts so the final set is *coherent* with the hierarchy. Classical approaches such as OLS [@Hyndman2011-jv], WLS [@Hyndman2016-ic], and MinT [@Wickramasuriya2019-xq] implement this as linear mappings (often projections) so that the adjusted forecasts lie in the coherent subspace.

# Motivation

<!--
- Recent works mainly focus on point forecasts
- Existing point forecast reconciliation can be extended to probabilistic forecasts
- Point forecast reconciliation methods are optimised for scoring rules such as MSE
- Question: How can probabilistic reconciliation be optimised with respect to different loss functions targeted at the
downstream decision? 
- Value-at-risk (VaR) is common risk measure, however, it ignores the tail risk and not coherent risk measure
- Expected Shortfall (ES) is coherent risk measure and accounts for tail risk, and has recently superseded the Value at Risk as the preferred risk measure in banking regulation.
- This work investigates the optimal probabilistic forecast reconciliation for Expected Shortfall.
-->

In the current literature, most reconciliation methodology has been developed for point forecasts, and its benefits for accuracy and coordination across organisation's departments or "silos" are well documented. However, in many applications the quantity of interest is probabilistic forecasts, not just a specific value. Decision makers care about the full distribution of future outcomes to quantify uncertainty, especially in risk management, where tail risks matter (e.g., extreme demand in newsvendor settings or energy generation shortfalls). This motivates moving from point reconciliation to probabilistic reconciliation, so that coherent forecasts also carry coherent uncertainty.

Existing reconciliation methods for point forecasts can be generalised to reconcile probabilistic forecasts. However, these approaches focus on point forecasts and are optimised for scoring rules such as mean squared error (MSE). There is a gap in the literature on how *optimal* this approach is with respect to the downstream decision. As the current expansion of point reconciliation methods no longer guarantee optimality for probabilistic forecasts, such as quantiles or expected shortfall (ES), a more principled approach is needed. This leaves the question: *"How can probabilistic reconciliation be optimised with respect to different loss functions targeted at the downstream decision?"*

Among the risk measures, quantiles, also known as Value-at-Risk (VaR), is a common choice, but it has limitations. VaR ignores the severity of losses beyond that threshold and is not a coherent risk measure, since it does not satisfy the subadditivity property [@Yamai2005-me]. This can lead to inconsistencies in risk assessments across different levels of the hierarchy.

Meanwhile, Expected Shortfall (ES) is a coherent risk measure that accounts for the severity of losses in the tail of the distribution [@Yamai2005-me]. It has recently been adopted as the preferred risk measure in banking regulation since the Basel III regulatory framework [@Basel-iii]. It has ability to capture tail risks more effectively than VaR. ES is also applicable beyond finance, such as quantifying expected economic loss when electricity demand exceeds available capacity. Therefore, we aim to investigate and develop the optimal probabilistic forecast reconciliation under the Expected Shortfall criterion.

<!-- footnote for subadditivity property? -->

# Methodology

## Preliminaries

- Hierarchical time series (HTS) is a collection of time series that are organised in a hierarchy.
- Coherent subspace and reconciliation definition
- Reconciliation to probabilistic forecasts

## Optimisation Framework

- ES definition
- ES is not elicitable, but is jointly elicitable with the quantile function.
- the quantile and ES jointly minimise in expectation the Fissler-Ziegel class of loss functions
- utilise the Rockafellar-Uryasev joint characterisation of the quantile and ES
- thus jointly reconcile the quantile and ES using the following bilevel optimisation problem
- ...

# Applications 
- Finance: ES is prefered risk measure, hierarchical structure from composite indices and etfs, also accounting for varying weights. Data is publicly available.
- Energy: ES can be a measure of economic loss accrued if insufficient generation capacity, hierarchical structure from geographical aggregation of service zones, states, and national level. Electricity load data are made available by the Australian Energy Market Operator

# Contributions

- Methodological: cutting-edge techniques from optimisation, econometrics and statistics will be combined to develop new reconciliation methods, new methods will be applied to important problems in energy, retail, and finance thus demonstrating that the research can inform decisions that lower real-world economic costs.

- Economic, environmental and social benefits: the research will improve the accuracy of probabilistic forecasts for decision making in energy, retail, and finance. This will lead to more efficient allocation of resources, reduced economic costs, and improved risk management.

### 2025-08-19 phd_application/phd_propossal.Rmd
Update phd_propossal.Rmd

@@
 
-Existing reconciliation methods for point forecasts can be generalised to reconcile probabilistic forecasts.
-However, these approaches focus on point forecasts and are optimised for scoring rules such as mean squared error (MSE).
+Existing reconciliation methods for point forecasts has recently been generalised to reconcile probabilistic forecasts.
+However, these approaches originally focus on point forecasts and are optimised for scoring rules such as mean squared error (MSE).
 There is a gap in the literature on how *optimal* this approach is with respect to the downstream decision.
-As the current expansion of point reconciliation methods no longer guarantee optimality for probabilistic forecasts, such as quantiles or expected shortfall (ES), a more principled approach is needed.
+The current expansion of point reconciliation methods no longer guarantee optimality for probabilistic forecasts, such as quantiles or expected shortfall (ES).
 This leaves the question: *"How can probabilistic reconciliation be optimised with respect to different loss functions targeted at the downstream decision?"*
 
+Although @Panagiotelis2023-fs introduced an score optimisation for reconciliation weights using the energy score or variogram score, they are scoring rules for full multivariate predictive distributions and do not directly address the tail functionals.
 Among the risk measures, quantiles, also known as Value-at-Risk (VaR), is a common choice, but it has limitations.

### 2025-08-20 phd_application/phd_propossal.Rmd
prelim writing

@@
 ## Preliminaries
-
+<!-- 
 - Hierarchical time series (HTS) is a collection of time series that are organised in a hierarchy.
 - Coherent subspace and reconciliation definition
-- Reconciliation to probabilistic forecasts
+- Reconciliation to probabilistic forecasts\
+-->
+
+Let $\boldsymbol{y}_t \in \mathbb{R}^K$ denote a vector of $K$ time series variables from the hierarchy, at time $t$.
+These variables are subject to linear aggregation constraints represented by a matrix $\boldsymbol C$, such that $\boldsymbol C \boldsymbol{y}_t = 0$.
+For example, in a retail setting, the total sales at a warehouse ($y_1$) and two stores ($y_2$ and $y_3$) might satisfy the constraint $y_1 = y_2 + y_3$, leading to $\boldsymbol C = (1, -1, -1)$.
+
+<!-- 
+\tilde{\boldsymbol{y}}_t
+\hat{\boldsymbol{y}}_t
+-->
+
+Base forecasts $\hat{\boldsymbol{y}}_t$, which we generate independently for each series, are typically *incoherent* (i.e.
+$\boldsymbol C \hat{\boldsymbol{y}}_t \neq 0$), while reconciled forecasts $\tilde{\boldsymbol{y}}_t$ are coherent by construction (i.e.
+$\boldsymbol C \hat{\boldsymbol{y}}_t = 0$).
+We obtain coherence via a linear mapping $\tilde{\boldsymbol{y}}_t = \boldsymbol P \hat{\boldsymbol{y}}_t$, where $\boldsymbol P$ maps the full *incoherent* space $\mathbb{R}^K$ to the linear *coherent* subspace $\mathbb{S} \subset \mathbb{R}^K$, where aggregation constraints hold.
+This mapping $\boldsymbol P$ can be a projection operator $\boldsymbol S (\boldsymbol S' \boldsymbol S)^{-1} \boldsymbol S'$ [@Hyndman2011-jv], where $\boldsymbol S$ is a matrix designed such that $\boldsymbol S \boldsymbol C = 0$.
+The mapping $\boldsymbol P$ is not unique, and selecting an optimal $\boldsymbol P$ is the core design problem.
+In this project, we will focus solely on linear reconciliation.
+
+For probabilistic forecasts, let $\hat F_t$ be a multivariate base forecast on $\mathbb{R}^K$.
+Define the reconciled probabilistic forecast $\tilde F_t$ on $S$ so that probabilities are preserved under the reconciliation map (intuitively, events in the coherent subspace under $\tilde F_t$ correspond to pre-images under $\hat F_t$).
+This extends point reconciliation to the distributional setting while respecting the aggregation constraints.&#x20;
 
 ## Optimisation Framework
-
+<!-- 
 - ES definition
@@
 - ...
+-->
 
 # Applications 
+<!-- 
 - Finance: ES is prefered risk measure, hierarchical structure from composite indices and etfs, also accounting for varying weights. Data is publicly available.
@@
 Electricity load data are made available by the Australian Energy Market Operator
+-->
 
 # Contributions
-
+<!-- 
 - Methodological: cutting-edge techniques from optimisation, econometrics and statistics will be combined to develop new reconciliation methods, new methods will be applied to important problems in energy, retail, and finance thus demonstrating that the research can inform decisions that lower real-world economic costs.
@@
 This will lead to more efficient allocation of resources, reduced economic costs, and improved risk management.
+-->

### 2025-08-20 phd_application/phd_propossal.Rmd
initial draft

@@
 ---
-title: "Decision-Optimal Probabilistic Forecast Reconciliation for Expected Shortfall in Hierarchical Time Series"
+title: "\\LARGE Decision-Optimal Probabilistic Forecast Reconciliation for Expected Shortfall in Hierarchical Time Series"
 author: 
@@
     number-sections: true
-# crossref:
-#   custom:
-#     - kind: float
-#       key: alg
-#       reference-prefix: Algorithm
-#       caption-prefix: Algorithm
-#       # for PDF output only, tell LaTeX to use the `algorithm` env:
-#       latex-env: algorithm
+link-citations: true
 execute:
@@
 For Starbucks, their total sales is the sum of all countries they are operating in, and each national sales is the sum for all cities, down to many outlets, forming an aggregation constraint.
-Forecasting such hierarchical time series arises in many decision-making contexts ranging from supply chains and electricity generation planning, to macroeconomic and tourism analysis, where stakeholders need forecasts at several aggregation levels to allocate resources and manage risk.
+Forecasting such hierarchical time series arises in many decision-making contexts ranging from supply chains [@Seaman2022-bb; @Angam2025-od] and energy planning [@Di-Modica2021-ad], to macroeconomic [@El-Gemayel2022-hx; @Li2019-vv] and tourism analysis [@Athanasopoulos2009-lp], where stakeholders need forecasts at several aggregation levels to allocate resources and manage risk.
 
@@
 Forecast reconciliation addresses this by adjusting the base forecasts so the final set is *coherent* with the hierarchy.
-Classical approaches such as OLS [@Hyndman2011-jv], WLS [@Hyndman2016-ic], and MinT [@Wickramasuriya2019-xq] implement this as linear mappings (often projections) so that the adjusted forecasts lie in the coherent subspace.
+Approaches such as OLS [@Hyndman2011-jv], WLS [@Hyndman2016-ic], and MinT [@Wickramasuriya2019-xq] implement this as linear mappings (often projections) so that the adjusted forecasts lie in the coherent subspace; see @Athanasopoulos2024-as for a comprehensive review.
 
@@
 
-## Preliminaries
+### Preliminaries {-}
 <!-- 
@@
 Let $\boldsymbol{y}_t \in \mathbb{R}^K$ denote a vector of $K$ time series variables from the hierarchy, at time $t$.
-These variables are subject to linear aggregation constraints represented by a matrix $\boldsymbol C$, such that $\boldsymbol C \boldsymbol{y}_t = 0$.
+For simplicity, we suppress the subscript $t$ and denote it as $\boldsymbol{y}$.
+These variables are subject to linear aggregation constraints represented by a matrix $\boldsymbol C$, such that $\boldsymbol C \boldsymbol{y} = 0$.
 For example, in a retail setting, the total sales at a warehouse ($y_1$) and two stores ($y_2$ and $y_3$) might satisfy the constraint $y_1 = y_2 + y_3$, leading to $\boldsymbol C = (1, -1, -1)$.
@@
 <!-- 
-\tilde{\boldsymbol{y}}_t
-\hat{\boldsymbol{y}}_t
+\tilde{\boldsymbol{y}}
+\hat{\boldsymbol{y}}
 -->
 
-Base forecasts $\hat{\boldsymbol{y}}_t$, which we generate independently for each series, are typically *incoherent* (i.e.
-$\boldsymbol C \hat{\boldsymbol{y}}_t \neq 0$), while reconciled forecasts $\tilde{\boldsymbol{y}}_t$ are coherent by construction (i.e.
-$\boldsymbol C \hat{\boldsymbol{y}}_t = 0$).
-We obtain coherence via a linear mapping $\tilde{\boldsymbol{y}}_t = \boldsymbol P \hat{\boldsymbol{y}}_t$, where $\boldsymbol P$ maps the full *incoherent* space $\mathbb{R}^K$ to the linear *coherent* subspace $\mathbb{S} \subset \mathbb{R}^K$, where aggregation constraints hold.
+Base forecasts $\hat{\boldsymbol{y}}$, which we generate independently for each series, are typically *incoherent* (i.e.
+$\boldsymbol C \hat{\boldsymbol{y}} \neq 0$), while reconciled forecasts $\tilde{\boldsymbol{y}}$ are coherent by construction (i.e.
+$\boldsymbol C \hat{\boldsymbol{y}} = 0$).
+We obtain coherence via a linear mapping $\tilde{\boldsymbol{y}} = \boldsymbol P \hat{\boldsymbol{y}}$, where $\boldsymbol P$ maps the full *incoherent* space $\mathbb{R}^K$ to the linear *coherent* subspace $\mathbb{S} \subset \mathbb{R}^K$, where aggregation constraints hold.
 This mapping $\boldsymbol P$ can be a projection operator $\boldsymbol S (\boldsymbol S' \boldsymbol S)^{-1} \boldsymbol S'$ [@Hyndman2011-jv], where $\boldsymbol S$ is a matrix designed such that $\boldsymbol S \boldsymbol C = 0$.
@@
 
-For probabilistic forecasts, let $\hat F_t$ be a multivariate base forecast on $\mathbb{R}^K$.
-Define the reconciled probabilistic forecast $\tilde F_t$ on $S$ so that probabilities are preserved under the reconciliation map (intuitively, events in the coherent subspace under $\tilde F_t$ correspond to pre-images under $\hat F_t$).
-This extends point reconciliation to the distributional setting while respecting the aggregation constraints.&#x20;
+<!-- 
+For probabilistic forecasts, let $\hat F$ be a multivariate base probabilistic forecast on $\mathbb{R}^K$.
+Define the reconciled probabilistic forecast $\tilde F$ on $\mathbb{S}$ so that probabilities are preserved under the reconciliation map (intuitively, events in the coherent subspace under $\tilde F$ correspond to pre-images under $\hat F$).
+This extends point reconciliation to the distributional setting while respecting the aggregation constraints.
+-->
 
-## Optimisation Framework
+\color{red}
+Talk about the probabilistic reconciliation and how it is extended from point reconciliation.
+\color{black}
+
+### Optimisation Framework {-}
 <!-- 
@@
 -->
+
+For a random variable $Y$, the Expected Shortfall (ES) at level $\alpha \in (0, 1)$ is defined as $\mathrm{ES}_\alpha(Y) = E[Y \mid Y < \mathrm{VaR}_\alpha(Y)]$, where $\mathrm{VaR}_\alpha(Y)$ is the quantile $q$ at level $\alpha$.
+The ES alone is not elicitable, meaning there is no scoring rule or loss function that can uniquely identify the ES as the optimal optimiser.
+However, @Fissler2015-uu showed that the quantile function and ES are jointly elicitable, minimising the Fissler-Ziegel class of loss functions in expectation.
+One of which takes the form:
+
+$$
+L_\alpha^\mathrm{FZ}(y, q, e) = (\mathbb{I}(y \leq q) - \alpha)(q-y) 
++ \frac{1}{\alpha} \mathrm{exp}(e) \mathbb{I}(y \leq q)(q-y) + \mathrm{exp}(e)(e-q-1),
+$$
+
+where $y$ is the realisation of $Y$, $q$ is the quantile, and $e$ is the ES.
+This loss function is non-convex, thus greatly complicating the optimisation problem.
+Alternatively, we propose to use the Rockafellar-Uryasev joint characterisation of the quantile and ES [@Rockafellar2000-el]:
+$\widetilde{\mathrm{ES}}_k(\boldsymbol{P})=\max_{q}\, E_{\tilde{\boldsymbol{y}}}\!\left[ H_{\alpha}\!\big(\tilde y_k(\boldsymbol{P}), q\big) \right]$ 
+and 
+$\tilde q_k(\boldsymbol{P})=\arg\max_{q}\, E_{\tilde{\boldsymbol{y}}}\!\left[ H_{\alpha}\!\big(\tilde y_k(\boldsymbol{P}), q\big) \right]$,
+where
+$H_{\alpha}(y, q) = q - \frac{1}{\alpha}\max\{0, y - q\}$.
+Thus, we jointly reconcile the quantile and ES using the following bilevel optimisation problem:
+
+\vspace{-0.8cm}
+
+\begin{align}
+\min_{\boldsymbol{P}}\;&
+\sum_{k=1}^{K} E_{\boldsymbol y}\!\left[
+  L^{\mathrm{FZ}}_{\alpha}\!\big(y_k,\, \tilde q_k(\boldsymbol{P}),\, \widetilde{\mathrm{ES}}_k(\boldsymbol{P})\big)
+\right]
+\\
+\text{s.t. }&
+\tilde q_k(\boldsymbol{P})=\arg\max_{q}\, E_{\tilde{\boldsymbol{y}}}\!\left[ H_{\alpha}\!\big(\tilde y_k(\boldsymbol{P}), q\big) \right],
+\notag\\
+&
+\widetilde{\mathrm{ES}}_k(\boldsymbol{P})=\max_{q}\, E_{\tilde{\boldsymbol{y}}}\!\left[ H_{\alpha}\!\big(\tilde y_k(\boldsymbol{P}), q\big) \right],
+\qquad \forall\,k=1,\ldots,K.
+\notag
+\end{align}
+
+\vspace{-0.4cm}
+
+This ensures that the lower-level problem is still convex, however the upper-level objective function with the Fissler-Ziegler loss is non-convex.
+Irrespective of this non-convexity, smoothing techniques combined with the implicit function and envelope theorems can be used to devise gradient-based methods for locally minimising.
 
@@
 -->
+
+\pagebreak
+# References {-}

### 2025-08-24 phd_application/phd_propossal.Rmd
update with Tas advice

@@
 
-<!-- # Abstract {-} -->
+<!--
+========== Tas Advice ==========
+- Start with ES
+  - Talk about VaR and then ES
+  - Take portfolio as example
+  - Walk to hierarchical time series
+  - Motivation is bringing ES and HTS together
+
+- Quantiles are point forecasts
+
+-->
 
@@
 <!-- 
+========== Old version ==========
+
 - Data is hierarchical
@@
 - Probabilistic forecasts are sometimes the quantity of interest, especially in risk management
--->
 
@@
 Approaches such as OLS [@Hyndman2011-jv], WLS [@Hyndman2016-ic], and MinT [@Wickramasuriya2019-xq] implement this as linear mappings (often projections) so that the adjusted forecasts lie in the coherent subspace; see @Athanasopoulos2024-as for a comprehensive review.
+-->
+
+<!-- 
+- Talk about VaR and then ES
+- Take portfolio as example
+- Link to hierarchical time series with previous example
+- Mention HTS arises in many contexts
+- Brief about coherence and reconciliation
+-->
+
+Risk managers and regulators oftenly use quantitative risk measures to quantify and manage potential losses in uncertain environments.
+Two familiar choices are Value-at-Risk (VaR) and Expected Shortfall (ES).
+While VaR is conceptually straightforward, it only measures the percentiles of loss distributions and ignores the severity of losses beyond that threshold [@Yamai2005-me].
+ES, on the other hand, captures the average loss in the tail beyond the VaR threshold.
+This makes ES a more reliable metric for assessing and managing "tail risks", and it has been adopted as the preferred regulatory metric since the Basel III framework [@Basel-iii].
+
+These tail risks are, however, not managed in isolation.
+Consider a bank's trading book, which consists of multiple portfolios across different desks, where each desk manages various asset classes, and then individual securities.
+This is an example of a hierarchical structure.
+Risk managers need to assess and forecast the risk quantities at each aggregation level, from the overall trading book down to individual securities Ignoring this naturally-formed hierarchical structure can lead to poor forecasts of uncertainty, resulting in misallocation of capital and suboptimal risk management strategies.
+
+Forecasting such hierarchical time series also arises in many decision-making contexts.
+From supply chains [@Seaman2022-bb; @Angam2025-od] and energy planning [@Di-Modica2021-ad], to macroeconomic [@El-Gemayel2022-hx; @Li2019-vv] and tourism analysis [@Athanasopoulos2009-lp], stakeholders need forecasts at several aggregation levels to allocate resources and manage risk.
+These hierarchical structures naturally form aggregation constraints.
+
+In practice, when forecasts are produced for all series, they typically violate these aggregation constraints observed in the data.
+Such forecasts are *incoherent*, which often happens when forecasts at lower-level series (e.g.
+individual securities) do not sum up to the aggregated series (e.g.
+overall trading book).
+This can undermine downstream decisions that require internal consistency.
+Forecast reconciliation, a post-processing step, addresses this by adjusting the so-called base forecasts so the final adjusted set is *coherent* with the aggregation structure.
+Approaches such as OLS [@Hyndman2011-jv], WLS [@Hyndman2016-ic], and MinT [@Wickramasuriya2019-xq] implement this as linear mappings so that the adjusted forecasts lie in the coherent subspace; see @Athanasopoulos2024-as for a comprehensive review.
+
+An additional incentive for the rapid adoption of reconciliation methods is that they improve forecast accuracy.
+For example, the International Monetary Fund reports improvements in liquidity forecasts of $3 million a day in a medium-sized economy, generating economic benefits via more efficient monetary policy operations [@El-Gemayel2022-hx].
+This advantage, combined with the need for coherent forecasts, motivates extending reconciliation methods to probabilistic forecasts, especially for tail risk measures such as ES.
 
@@
 <!--
+========== Old version ==========
+
 - Recent works mainly focus on point forecasts
@@
 - This work investigates the optimal probabilistic forecast reconciliation for Expected Shortfall.
--->
 
@@
 Although @Panagiotelis2023-fs introduced an score optimisation for reconciliation weights using the energy score or variogram score, they are scoring rules for full multivariate predictive distributions and do not directly address the tail functionals.
+
 Among the risk measures, quantiles, also known as Value-at-Risk (VaR), is a common choice, but it has limitations.
@@
 Therefore, we aim to investigate and develop the optimal probabilistic forecast reconciliation under the Expected Shortfall criterion.
+-->
+
+In the current literature, most reconciliation methodology has been developed for point forecasts, and its benefits for accuracy and coordination across organisation's departments or "silos" are well documented.
+However, in many applications the quantity of interest is a probabilistic forecast, not just a typical mean or median value.
+Modern risk management and operations need more than a single number--they need quantities from the full predictive distribution that capture uncertainty and tail behaviour, because many decisions are tail-sensitive (e.g., extreme demand in newsvendor settings or energy generation shortfalls).
+
+Existing reconciliation methods for point forecasts has recently been generalised to reconcile probabilistic forecasts with downstream decisions based on these.
+However, these approaches originally focus on point forecasts and are optimised for scoring functions such as mean squared error (MSE).
+There is a gap in the literature on how *optimal* this approach when considering alternative quantities of interest, such as quantiles or expected shortfall (ES).
+This leaves the question: *"How can we construct probabilistic forecast reconciliation that are explicitly optimised for ES (and, where useful, jointly for VaR and ES)?"*
+
+Although @Panagiotelis2023-fs introduced an score optimisation for reconciliation weights using proper scoring rules, energy score and variogram score, for full multivariate predictive distributions, they do not directly address the tail functionals.
+Other than that, there is a lack of research on optimising reconciliation for specific risk measures such as VaR and ES.
+
+- briefly expand on limitaions of VaR: not coherent risk measure, not subadditive, on top of ignoring tail risk
+- focus on ES
+- however ES is not elicitable, but is jointly elicitable with the quantile function
+- Bring the work of Fissler and Ziegel (2016) and Rockafellar and Uryasev (2000) into the reconciliation framework
 

### 2025-08-25 phd_application/phd_propossal.Rmd
finish up motivation

@@
 
-Risk managers and regulators oftenly use quantitative risk measures to quantify and manage potential losses in uncertain environments.
-Two familiar choices are Value-at-Risk (VaR) and Expected Shortfall (ES).
+<!-- 
 While VaR is conceptually straightforward, it only measures the percentiles of loss distributions and ignores the severity of losses beyond that threshold [@Yamai2005-me].
@@
 This makes ES a more reliable metric for assessing and managing "tail risks", and it has been adopted as the preferred regulatory metric since the Basel III framework [@Basel-iii].
+-->
 
+Risk managers and regulators oftenly use quantitative risk measures to quantify and manage potential losses in uncertain environments.
+Two familiar choices are Value-at-Risk (VaR) and Expected Shortfall (ES; also known as Conditional VaR).
 These tail risks are, however, not managed in isolation.
@@
 This is an example of a hierarchical structure.
-Risk managers need to assess and forecast the risk quantities at each aggregation level, from the overall trading book down to individual securities Ignoring this naturally-formed hierarchical structure can lead to poor forecasts of uncertainty, resulting in misallocation of capital and suboptimal risk management strategies.
+Risk managers need to assess and forecast the risk quantities at each aggregation level, from the overall trading book down to individual securities.
+Ignoring this naturally-formed hierarchical structure can lead to poor forecasts of uncertainty, resulting in misallocation of capital and suboptimal risk management strategies.
 
@@
 Forecast reconciliation, a post-processing step, addresses this by adjusting the so-called base forecasts so the final adjusted set is *coherent* with the aggregation structure.
-Approaches such as OLS [@Hyndman2011-jv], WLS [@Hyndman2016-ic], and MinT [@Wickramasuriya2019-xq] implement this as linear mappings so that the adjusted forecasts lie in the coherent subspace; see @Athanasopoulos2024-as for a comprehensive review.
+Approaches such as OLS [@Hyndman2011-jv], WLS [@Hyndman2016-ic], and MinT [@Wickramasuriya2019-xq] are commonly used in practice; see @Athanasopoulos2024-as for a comprehensive review.
 
@@
 For example, the International Monetary Fund reports improvements in liquidity forecasts of $3 million a day in a medium-sized economy, generating economic benefits via more efficient monetary policy operations [@El-Gemayel2022-hx].
-This advantage, combined with the need for coherent forecasts, motivates extending reconciliation methods to probabilistic forecasts, especially for tail risk measures such as ES.
+This context motivates methods that deliver coherent, multi-level forecasts whose uncertainty aligns with the tail-risk metrics managers actually use.
 
@@
 However, these approaches originally focus on point forecasts and are optimised for scoring functions such as mean squared error (MSE).
-There is a gap in the literature on how *optimal* this approach when considering alternative quantities of interest, such as quantiles or expected shortfall (ES).
-This leaves the question: *"How can we construct probabilistic forecast reconciliation that are explicitly optimised for ES (and, where useful, jointly for VaR and ES)?"*
+There is a gap in the literature on how *optimal* this approach when considering alternative quantities of interest, such as Value-at-Risk or Expected Shortfall.
+This leaves the question: *"How can we construct probabilistic forecast reconciliation that are explicitly optimised for tail-risk measures?"*
 
-Although @Panagiotelis2023-fs introduced an score optimisation for reconciliation weights using proper scoring rules, energy score and variogram score, for full multivariate predictive distributions, they do not directly address the tail functionals.
+Although @Panagiotelis2023-fs introduced an score optimisation for reconciliation weights using proper scoring rules (energy score and variogram score) for full multivariate predictive distributions, they do not directly address the tail functionals.
 Other than that, there is a lack of research on optimising reconciliation for specific risk measures such as VaR and ES.
 
-- briefly expand on limitaions of VaR: not coherent risk measure, not subadditive, on top of ignoring tail risk
-- focus on ES
-- however ES is not elicitable, but is jointly elicitable with the quantile function
-- Bring the work of Fissler and Ziegel (2016) and Rockafellar and Uryasev (2000) into the reconciliation framework
+Both VaR and ES are popular risk measures, but they have different properties.
+While VaR is conceptually straightforward, it ignores the severity of losses beyond the VaR threshold and is not a coherent risk measure, since it does not satisfy the subadditivity property [@Yamai2005-me].
+VaR can mislead risk assessments across different levels of the hierarchy, and may underestimate the true risk exposure under market stress.
+ES, on the other hand, is a coherent metric that captures the average loss in the tail beyond the VaR threshold.
+This makes ES a more reliable metric for assessing and managing "tail risks", and it has been adopted as the preferred regulatory metric since the Basel III framework [@Basel-iii].
+
+Even with these advantages, ES has its own challenges.
+Notably, ES is not elicitable on its own, meaning there is no scoring rule or loss function that can uniquely identify the ES as the optimal optimiser.
+However, works from @Fissler2015-uu and @Rockafellar2000-el show that this is not impossible when ES is considered jointly with VaR.
+This joint elicitability opens the door to optimising reconciliation methods specifically for ES.
+As a result, the biggest objective of this project is to bring these recent theoretical advances into the reconciliation framework, to develop methods that produce coherent probabilistic forecasts optimised for ES.
 

### 2025-08-25 phd_application/phd_propossal.Rmd
finish up background adn motivation

@@
 author: 
-  - Vincent Su
+  - \Large Vincent Su
 format: 
@@
         \usepackage{array}
+
+        % adjust spacing
+        \usepackage{titling}
+        \pretitle{\begin{center}\vspace{-5em}}
+        \posttitle{\end{center}\vspace{-1.5em}} 
+        \preauthor{\begin{center}}
+        \postauthor{\end{center}\vspace{-4em}} 
     mathspec: true
@@
 execute:
-  cache: false
+  cache: true
   echo: false
@@
 
-Risk managers and regulators oftenly use quantitative risk measures to quantify and manage potential losses in uncertain environments.
-Two familiar choices are Value-at-Risk (VaR) and Expected Shortfall (ES; also known as Conditional VaR).
+Risk managers and regulators use quantitative risk measures to quantify and manage potential losses in uncertain environments.
+Two familiar choices are Value-at-Risk (VaR)--a loss threshold exceeded with some probability--and Expected Shortfall (ES; also known as Conditional VaR)--the average loss beyond that threshold.
+
 These tail risks are, however, not managed in isolation.
 Consider a bank's trading book, which consists of multiple portfolios across different desks, where each desk manages various asset classes, and then individual securities.
-This is an example of a hierarchical structure.
-Risk managers need to assess and forecast the risk quantities at each aggregation level, from the overall trading book down to individual securities.
-Ignoring this naturally-formed hierarchical structure can lead to poor forecasts of uncertainty, resulting in misallocation of capital and suboptimal risk management strategies.
+This is an example of a hierarchical structure, with multiple *aggregation levels*.
+Risk managers need to assess and forecast the risk quantities at several levels, from the overall trading book down to individual securities.
+Ignoring this naturally-formed hierarchical structure can lead to inconsistent numbers and poor forecasts, resulting in misallocation of capital and suboptimal risk management strategies.
 
@@
 From supply chains [@Seaman2022-bb; @Angam2025-od] and energy planning [@Di-Modica2021-ad], to macroeconomic [@El-Gemayel2022-hx; @Li2019-vv] and tourism analysis [@Athanasopoulos2009-lp], stakeholders need forecasts at several aggregation levels to allocate resources and manage risk.
-These hierarchical structures naturally form aggregation constraints.
+These hierarchical structures naturally form *aggregation constraints* (child levels must sum up to parents).
 
-In practice, when forecasts are produced for all series, they typically violate these aggregation constraints observed in the data.
-Such forecasts are *incoherent*, which often happens when forecasts at lower-level series (e.g.
-individual securities) do not sum up to the aggregated series (e.g.
-overall trading book).
+In practice, when forecasts are produced for all series (often called *base forecasts*), they typically violate these aggregation constraints observed in the data; such forecasts are *incoherent* (e.g., security-level numbers don’t sum to the overall trading book).
 This can undermine downstream decisions that require internal consistency.
 Forecast reconciliation, a post-processing step, addresses this by adjusting the so-called base forecasts so the final adjusted set is *coherent* with the aggregation structure.
-Approaches such as OLS [@Hyndman2011-jv], WLS [@Hyndman2016-ic], and MinT [@Wickramasuriya2019-xq] are commonly used in practice; see @Athanasopoulos2024-as for a comprehensive review.
+Approaches such as OLS [@Hyndman2011-jv], WLS [@Hyndman2016-ic], MinT [@Wickramasuriya2019-xq], and more are developed and commonly used in practice.
+An additional incentive for the rapid adoption of reconciliation methods is that they improve forecast accuracy [@Athanasopoulos2024-as], with example of the International Monetary Fund reporting improvements in their liquidity forecasts, generating economic benefits via more efficient monetary policy operations [@El-Gemayel2022-hx].
 
-An additional incentive for the rapid adoption of reconciliation methods is that they improve forecast accuracy.
-For example, the International Monetary Fund reports improvements in liquidity forecasts of $3 million a day in a medium-sized economy, generating economic benefits via more efficient monetary policy operations [@El-Gemayel2022-hx].
+\color{red}
+Importantly, ES behaves well under aggregation (satisfying subadditivity\footnote{\textbf{Subadditivity.} A risk measure $\rho$ on loss variables is subadditive if $\rho(X+Y)\le \rho(X)+\rho(Y)$ for any losses $X,Y$.
+It encodes diversification: the risk of a combined portfolio should not exceed the sum of stand-alone risks.
+Expected Shortfall satisfies subadditivity for all $\alpha\in(0,1)$, whereas VaR can violate it.
+Thus, the portfolio ES cannot exceed the sums of desk ESs.}), making it a natural target when risks are rolled up through a hierarchy.
 This context motivates methods that deliver coherent, multi-level forecasts whose uncertainty aligns with the tail-risk metrics managers actually use.
+
+(should i remove this?)
+\color{black}
 
@@
 
+<!--
+=========== 1st version =========
 In the current literature, most reconciliation methodology has been developed for point forecasts, and its benefits for accuracy and coordination across organisation's departments or "silos" are well documented.
-However, in many applications the quantity of interest is a probabilistic forecast, not just a typical mean or median value.
-Modern risk management and operations need more than a single number--they need quantities from the full predictive distribution that capture uncertainty and tail behaviour, because many decisions are tail-sensitive (e.g., extreme demand in newsvendor settings or energy generation shortfalls).
+However, modern risk management and operations need more than a typical mean or median value--they need quantities from the full predictive distribution that capture uncertainty and tail behaviour, because many decisions are tail-sensitive (e.g., extreme demand in newsvendor settings or energy generation shortfalls).
 
@@
 As a result, the biggest objective of this project is to bring these recent theoretical advances into the reconciliation framework, to develop methods that produce coherent probabilistic forecasts optimised for ES.
+-->
+
+Modern risk management and operations need more than a typical mean or median value--they need quantities from the full predictive distribution that capture uncertainty and tail behaviour, because many decisions are tail-sensitive (e.g., extreme demand in retail settings or energy generation shortfalls).
+While VaR is conceptually straightforward, it ignores the severity of losses beyond the VaR threshold and is not a coherent risk measure, since it does not satisfy the subadditivity property [@Yamai2005-me].
+VaR can mislead risk assessments across different levels of the hierarchy, and may underestimate the true risk exposure under market stress.
+ES, on the other hand, is a coherent metric that captures the average loss in the tail beyond the VaR threshold.
+This makes ES a more reliable metric for assessing and managing "tail risks", and it has been adopted as the preferred regulatory metric since the Basel III framework [@Basel-iii].
+If ES is the metric that risk managers actually use, reconciliation methods should guarantee the coherence and optimality with respect to ES.
+
+Most reconciliation methodology has been developed for point forecasts and are optimised for point scoring functions such as mean squared error (MSE).
+Generalising these methods to reconcile probabilistic forecasts does not guarantee optimality for tail-risk measures such as VaR or ES.
+Although @Panagiotelis2023-fs introduced an score optimisation for reconciliation weights using proper scoring rules (energy score and variogram score) for full multivariate predictive distributions, they do not directly address the tail functionals.
+This leaves the question: *"How can we construct probabilistic forecast reconciliation that are explicitly optimised for tail-risk measures such as ES?"*
+
+A practical challenge is that ES is not elicitable on its own, meaning there is no scoring rule or loss function that can uniquely identify ES as the optimiser.
+However, ES is jointly elicitable with VaR [@Fissler2015-uu], and is amenable to convex optimisation via an auxiliary VaR variable [@Rockafellar2000-el].
+This opens the door to train reconciliation specifically for ES.
+
+As a result, the objective of this project is to bring these theoretical advances into the reconciliation framework and develop methods that produce coherent probabilistic forecasts optimised for ES.
+Our goal is a reliable, consistent, and decision-aligned risk-forecasting framework that helps national and international organisations improve ES backtesting, stabilise capital, and allocate resources more effectively across hierarchical levels.
 
@@
 We obtain coherence via a linear mapping $\tilde{\boldsymbol{y}} = \boldsymbol P \hat{\boldsymbol{y}}$, where $\boldsymbol P$ maps the full *incoherent* space $\mathbb{R}^K$ to the linear *coherent* subspace $\mathbb{S} \subset \mathbb{R}^K$, where aggregation constraints hold.
+
+<!-- 
 This mapping $\boldsymbol P$ can be a projection operator $\boldsymbol S (\boldsymbol S' \boldsymbol S)^{-1} \boldsymbol S'$ [@Hyndman2011-jv], where $\boldsymbol S$ is a matrix designed such that $\boldsymbol S \boldsymbol C = 0$.
@@
 
-<!-- 
 For probabilistic forecasts, let $\hat F$ be a multivariate base probabilistic forecast on $\mathbb{R}^K$.
@@
 This extends point reconciliation to the distributional setting while respecting the aggregation constraints.
--->
 
@@
 \color{black}
+-->
 

### 2025-08-26 phd_application/phd_propossal.Rmd
update background from Monday feedback

@@
 These tail risks are, however, not managed in isolation.
-Consider a bank's trading book, which consists of multiple portfolios across different desks, where each desk manages various asset classes, and then individual securities.
+Consider a hedge fund's investment portfolio returns, which consists returns of different asset classes, and further contains returns of various individual securities.
 This is an example of a hierarchical structure, with multiple *aggregation levels*.
-Risk managers need to assess and forecast the risk quantities at several levels, from the overall trading book down to individual securities.
-Ignoring this naturally-formed hierarchical structure can lead to inconsistent numbers and poor forecasts, resulting in misallocation of capital and suboptimal risk management strategies.
+Risk managers need to assess and forecast the risk quantities at several levels, from the overall portfolio, down to individual securities.
+Ignoring this naturally-formed hierarchical structure can lead to misaligned forecasts, for example, the forecasted returns of individual securities are low while the overall portfolio return is high.
+This results in misallocation of capital and suboptimal risk management strategies.
 
-Forecasting such hierarchical time series also arises in many decision-making contexts.
-From supply chains [@Seaman2022-bb; @Angam2025-od] and energy planning [@Di-Modica2021-ad], to macroeconomic [@El-Gemayel2022-hx; @Li2019-vv] and tourism analysis [@Athanasopoulos2009-lp], stakeholders need forecasts at several aggregation levels to allocate resources and manage risk.
+Forecasting such hierarchical structure also arises in many other decision-making contexts, from supply chains [@Seaman2022-bb; @Angam2025-od] and energy planning [@Di-Modica2021-ad], to macroeconomic [@El-Gemayel2022-hx; @Li2019-vv] and tourism analysis [@Athanasopoulos2009-lp].
+Stakeholders in these settings need forecasts at several aggregation levels to allocate resources and manage risk.
 These hierarchical structures naturally form *aggregation constraints* (child levels must sum up to parents).
+Taken together, when we have a collection of time series organised in a hierarchy subject to aggregation constraints, we refer to this as *hierarchical time series*.
 
-In practice, when forecasts are produced for all series (often called *base forecasts*), they typically violate these aggregation constraints observed in the data; such forecasts are *incoherent* (e.g., security-level numbers don’t sum to the overall trading book).
+In practice, when forecasts are produced for all series (often called *base forecasts*), they typically violate the aggregation constraints observed in the data; such forecasts are *incoherent* (e.g., security-level numbers don’t sum to the overall trading book).
 This can undermine downstream decisions that require internal consistency.
-Forecast reconciliation, a post-processing step, addresses this by adjusting the so-called base forecasts so the final adjusted set is *coherent* with the aggregation structure.
+Forecast reconciliation, a post-processing step, addresses this by adjusting the base forecasts so the final adjusted set of forecasts is *coherent* with the aggregation structure.
 Approaches such as OLS [@Hyndman2011-jv], WLS [@Hyndman2016-ic], MinT [@Wickramasuriya2019-xq], and more are developed and commonly used in practice.
-An additional incentive for the rapid adoption of reconciliation methods is that they improve forecast accuracy [@Athanasopoulos2024-as], with example of the International Monetary Fund reporting improvements in their liquidity forecasts, generating economic benefits via more efficient monetary policy operations [@El-Gemayel2022-hx].
+An additional incentive for the rapid adoption of reconciliation methods is that they improve forecast accuracy [@Athanasopoulos2024-as].
+For instance, the International Monetary Fund reported improvements in their liquidity forecasts, generating economic benefits via more efficient monetary policy operations [@El-Gemayel2022-hx].
 

### 2025-08-27 phd_application/phd_propossal.Rmd
update bkgroud and motiv

@@
 These tail risks are, however, not managed in isolation.
-Consider a hedge fund's investment portfolio returns, which consists returns of different asset classes, and further contains returns of various individual securities.
+Consider a hedge fund's investment portfolio returns, which compose of returns of different asset classes, and each asset class further contain returns of various individual securities.
 This is an example of a hierarchical structure, with multiple *aggregation levels*.
-Risk managers need to assess and forecast the risk quantities at several levels, from the overall portfolio, down to individual securities.
-Ignoring this naturally-formed hierarchical structure can lead to misaligned forecasts, for example, the forecasted returns of individual securities are low while the overall portfolio return is high.
+Risk managers need to assess and forecast the returns or risk quantities at several levels, from the overall portfolio, down to individual securities.
+Ignoring this naturally-formed hierarchical structure can lead to misaligned forecasts, for example, the predicted returns of individual securities are low while the overall portfolio returns are high.
 This results in misallocation of capital and suboptimal risk management strategies.
@@
 These hierarchical structures naturally form *aggregation constraints* (child levels must sum up to parents).
-Taken together, when we have a collection of time series organised in a hierarchy subject to aggregation constraints, we refer to this as *hierarchical time series*.
+Taken together, when we have a collection of time series organised in a hierarchy subject to the aggregation constraints, we refer to this as *hierarchical time series*.
 
-In practice, when forecasts are produced for all series (often called *base forecasts*), they typically violate the aggregation constraints observed in the data; such forecasts are *incoherent* (e.g., security-level numbers don’t sum to the overall trading book).
+In practice, when forecasts are produced for all series (often called *base forecasts*), they typically violate the aggregation constraints observed in the data; such forecasts are *incoherent*.
 This can undermine downstream decisions that require internal consistency.
@@
 \color{red}
-Importantly, ES behaves well under aggregation (satisfying subadditivity\footnote{\textbf{Subadditivity.} A risk measure $\rho$ on loss variables is subadditive if $\rho(X+Y)\le \rho(X)+\rho(Y)$ for any losses $X,Y$.
-It encodes diversification: the risk of a combined portfolio should not exceed the sum of stand-alone risks.
-Expected Shortfall satisfies subadditivity for all $\alpha\in(0,1)$, whereas VaR can violate it.
-Thus, the portfolio ES cannot exceed the sums of desk ESs.}), making it a natural target when risks are rolled up through a hierarchy.
+Importantly, ES behaves well under aggregation (satisfying subadditivity), making it a natural target when risks are rolled up through a hierarchy.
 This context motivates methods that deliver coherent, multi-level forecasts whose uncertainty aligns with the tail-risk metrics managers actually use.
@@
 Modern risk management and operations need more than a typical mean or median value--they need quantities from the full predictive distribution that capture uncertainty and tail behaviour, because many decisions are tail-sensitive (e.g., extreme demand in retail settings or energy generation shortfalls).
-While VaR is conceptually straightforward, it ignores the severity of losses beyond the VaR threshold and is not a coherent risk measure, since it does not satisfy the subadditivity property [@Yamai2005-me].
-VaR can mislead risk assessments across different levels of the hierarchy, and may underestimate the true risk exposure under market stress.
-ES, on the other hand, is a coherent metric that captures the average loss in the tail beyond the VaR threshold.
-This makes ES a more reliable metric for assessing and managing "tail risks", and it has been adopted as the preferred regulatory metric since the Basel III framework [@Basel-iii].
+While VaR is conceptually straightforward, it ignores the severity of losses beyond the VaR threshold and does not satisfy the subadditivity property\footnote{\textbf{Subadditivity.} A risk measure $\rho$ on loss variables is subadditive if $\rho(X+Y)\le \rho(X)+\rho(Y)$ for any losses $X,Y$.
+Expected Shortfall satisfies subadditivity for all $\alpha\in(0,1)$, whereas VaR can violate it.
+Thus, the portfolio ES cannot exceed the sums of individual security ESs.} [@Yamai2005-me].
+Violating subadditivity is a critical problem, because it means that the risk measure of a combined portfolio can exceed the sum of stand-alone risks, which contradicts the principle of diversification.
+VaR therefore can mislead risk assessments across different levels of the hierarchy, and may underestimate the true risk exposure under market stress.
+ES, on the other hand, is a consistent metric that captures the average loss in the tail beyond the VaR threshold.
+This makes ES a more reliable measure for assessing and managing "tail risks", and it has been adopted as the preferred regulatory metric since the Basel III framework [@Basel-iii].
 If ES is the metric that risk managers actually use, reconciliation methods should guarantee the coherence and optimality with respect to ES.
@@
 Generalising these methods to reconcile probabilistic forecasts does not guarantee optimality for tail-risk measures such as VaR or ES.
-Although @Panagiotelis2023-fs introduced an score optimisation for reconciliation weights using proper scoring rules (energy score and variogram score) for full multivariate predictive distributions, they do not directly address the tail functionals.
+Although @Panagiotelis2023-fs introduced score optimisation for reconciliation weights using proper scoring rules (energy score and variogram score) for full multivariate predictive distributions, they do not directly address the tail functionals.
 This leaves the question: *"How can we construct probabilistic forecast reconciliation that are explicitly optimised for tail-risk measures such as ES?"*
+
+<!-- \color{red}
+**AP: I have thought a bit about how we discuss the point v probabilistic issue.
+Ultimately, a VaR or ES forecast is a point forecast.
+However, a) any point forecast can be derived from a distributional forecast and b) scoring rules give us a way to obtain the optimal VaR or ES if we know the underlying probabilistic distribution.
+These two facts taken together mean we can naturally adapt the score optimisation framework of Panagiotelis et al (2023) to obtain optimal ES forecasts while also guaranteeing that these forecast come from an underlying probabilitistic forecast that is coherent.**
+\color{black} -->
 
@@
 
-As a result, the objective of this project is to bring these theoretical advances into the reconciliation framework and develop methods that produce coherent probabilistic forecasts optimised for ES.
-Our goal is a reliable, consistent, and decision-aligned risk-forecasting framework that helps national and international organisations improve ES backtesting, stabilise capital, and allocate resources more effectively across hierarchical levels.
+As a result, the objective of my thesis is to bring these theoretical advances into the reconciliation framework and develop methods that produce coherent probabilistic forecasts optimised for ES.
+The goal is a reliable, consistent, and decision-aligned risk-forecasting framework that helps national and international organisations improve ES backtesting, stabilise capital, and allocate resources more effectively across hierarchical levels.
 

### 2025-08-27 phd_application/phd_propossal.Rmd
update methodology sec

@@
 Risk managers need to assess and forecast the returns or risk quantities at several levels, from the overall portfolio, down to individual securities.
-Ignoring this naturally-formed hierarchical structure can lead to misaligned forecasts, for example, the predicted returns of individual securities are low while the overall portfolio returns are high.
+Ignoring this naturally formed hierarchical structure can lead to misaligned forecasts, for example, the predicted returns of individual securities are low while the overall portfolio returns are high.
 This results in misallocation of capital and suboptimal risk management strategies.
@@
 
-\color{red}
+<!-- \color{red}
 Importantly, ES behaves well under aggregation (satisfying subadditivity), making it a natural target when risks are rolled up through a hierarchy.
@@
 (should i remove this?)
-\color{black}
+\color{black} -->
 
@@
 This makes ES a more reliable measure for assessing and managing "tail risks", and it has been adopted as the preferred regulatory metric since the Basel III framework [@Basel-iii].
-If ES is the metric that risk managers actually use, reconciliation methods should guarantee the coherence and optimality with respect to ES.
+Although ES is the metric that risk managers actually use, there is a lack of research on forecasting ES within the reconciliation framework.
 
+\color{red}
 Most reconciliation methodology has been developed for point forecasts and are optimised for point scoring functions such as mean squared error (MSE).
@@
 This leaves the question: *"How can we construct probabilistic forecast reconciliation that are explicitly optimised for tail-risk measures such as ES?"*
+\color{black}
 
-<!-- \color{red}
+<!-- 
 **AP: I have thought a bit about how we discuss the point v probabilistic issue.
@@
 These two facts taken together mean we can naturally adapt the score optimisation framework of Panagiotelis et al (2023) to obtain optimal ES forecasts while also guaranteeing that these forecast come from an underlying probabilitistic forecast that is coherent.**
-\color{black} -->
+-->
 
@@
 
-As a result, the objective of my thesis is to bring these theoretical advances into the reconciliation framework and develop methods that produce coherent probabilistic forecasts optimised for ES.
+As a result, the objective of this thesis is to bring these theoretical advances into the reconciliation framework and develop methods that produce coherent probabilistic forecasts optimised for ES.
 The goal is a reliable, consistent, and decision-aligned risk-forecasting framework that helps national and international organisations improve ES backtesting, stabilise capital, and allocate resources more effectively across hierarchical levels.
@@
 
+<!-- 
+=========== OLD version
+
 Let $\boldsymbol{y}_t \in \mathbb{R}^K$ denote a vector of $K$ time series variables from the hierarchy, at time $t$.
@@
 
-<!-- 
 \tilde{\boldsymbol{y}}
 \hat{\boldsymbol{y}}
--->
 
-Base forecasts $\hat{\boldsymbol{y}}$, which we generate independently for each series, are typically *incoherent* (i.e.
+Base forecasts $\hat{\boldsymbol{y}}$, which we generate for each series, are typically *incoherent* (i.e.
 $\boldsymbol C \hat{\boldsymbol{y}} \neq 0$), while reconciled forecasts $\tilde{\boldsymbol{y}}$ are coherent by construction (i.e.
@@
 
-<!-- 
 This mapping $\boldsymbol P$ can be a projection operator $\boldsymbol S (\boldsymbol S' \boldsymbol S)^{-1} \boldsymbol S'$ [@Hyndman2011-jv], where $\boldsymbol S$ is a matrix designed such that $\boldsymbol S \boldsymbol C = 0$.
@@
 
+Let $\boldsymbol{y}_t \in \mathbb{R}^K$ denote a vector of $K$ time series variables from the hierarchy, at time $t$.
+For simplicity, we suppress the subscript $t$ and denote it as $\boldsymbol{y}$.
+These variables are subject to linear aggregation constraints.
+Base forecasts $\hat{\boldsymbol{y}}$, which we generate for each series, are typically *incoherent*, while reconciled forecasts $\tilde{\boldsymbol{y}}$ are coherent by construction.
+We obtain coherence via a linear mapping $\tilde{\boldsymbol{y}} = \boldsymbol P \hat{\boldsymbol{y}}$, where $\boldsymbol P$ maps the full *incoherent* space $\mathbb{R}^K$ to the linear *coherent* subspace $\mathbb{S} \subset \mathbb{R}^K$, where aggregation constraints hold.
+Ultimately, selecting an optimal $\boldsymbol P$ is the core design problem.
+In this project, we will focus solely on linear reconciliation.
+
+For a random variable $Y$, the Expected Shortfall (ES) at level $\alpha \in (0, 1)$ is defined as $\mathrm{ES}_\alpha(Y) = E[Y \mid Y < \mathrm{VaR}_\alpha(Y)]$, where $\mathrm{VaR}_\alpha(Y)$ is the quantile $q$ at level $\alpha$.
+Thus, we can define $\widetilde{\mathrm{ES}}_k(\boldsymbol{P}) = \mathrm{ES}_\alpha(\tilde y_k(\boldsymbol{P}))$ as the ES of the $k$-th series after reconciliation, where $\tilde y_k(\boldsymbol{P})$ is the $k$-th element of the reconciled forecasts $\tilde{\boldsymbol{y}}(\boldsymbol{P}) = \boldsymbol P \hat{\boldsymbol{y}}$; and similarly define $\tilde q_k(\boldsymbol{P}) = \mathrm{VaR}_\alpha(\tilde y_k(\boldsymbol{P}))$ as the quantile of the $k$-th series after reconciliation.
+
 ### Optimisation Framework {-}
@@
 
-For a random variable $Y$, the Expected Shortfall (ES) at level $\alpha \in (0, 1)$ is defined as $\mathrm{ES}_\alpha(Y) = E[Y \mid Y < \mathrm{VaR}_\alpha(Y)]$, where $\mathrm{VaR}_\alpha(Y)$ is the quantile $q$ at level $\alpha$.
-The ES alone is not elicitable, meaning there is no scoring rule or loss function that can uniquely identify the ES as the optimal optimiser.
-However, @Fissler2015-uu showed that the quantile function and ES are jointly elicitable, minimising the Fissler-Ziegel class of loss functions in expectation.
-One of which takes the form:
+@Fissler2015-uu showed that the quantile function and ES are jointly elicitable, minimising the Fissler-Ziegel class of loss functions in expectation. One of which takes the form:
 

### 2025-08-28 phd_application/phd_propossal.Rmd
finalise

@@
 ---
-title: "\\LARGE Decision-Optimal Probabilistic Forecast Reconciliation for Expected Shortfall in Hierarchical Time Series"
+title: "\\LARGE Decision-Optimal Forecast Reconciliation for Expected Shortfall in Hierarchical Time Series"
 author: 
@@
         \usepackage{titling}
-        \pretitle{\begin{center}\vspace{-5em}}
+        \pretitle{\begin{center}\vspace{-3em}\bfseries}
         \posttitle{\end{center}\vspace{-1.5em}} 
@@
 
-Forecasting such hierarchical structure also arises in many other decision-making contexts, from supply chains [@Seaman2022-bb; @Angam2025-od] and energy planning [@Di-Modica2021-ad], to macroeconomic [@El-Gemayel2022-hx; @Li2019-vv] and tourism analysis [@Athanasopoulos2009-lp].
+Forecasting data from such hierarchical structures also arises in many other decision-making contexts, from supply chains [@Seaman2022-bb; @Angam2025-od] and energy planning [@Di-Modica2021-ad], to macroeconomics [@El-Gemayel2022-hx; @Li2019-vv] and tourism analysis [@Athanasopoulos2009-lp].
 Stakeholders in these settings need forecasts at several aggregation levels to allocate resources and manage risk.
 These hierarchical structures naturally form *aggregation constraints* (child levels must sum up to parents).
-Taken together, when we have a collection of time series organised in a hierarchy subject to the aggregation constraints, we refer to this as *hierarchical time series*.
+Taken together, a collection of time series organised in a hierarchy and subject to aggregation constraints is refered as *hierarchical time series*.
 
@@
 This can undermine downstream decisions that require internal consistency.
-Forecast reconciliation, a post-processing step, addresses this by adjusting the base forecasts so the final adjusted set of forecasts is *coherent* with the aggregation structure.
-Approaches such as OLS [@Hyndman2011-jv], WLS [@Hyndman2016-ic], MinT [@Wickramasuriya2019-xq], and more are developed and commonly used in practice.
-An additional incentive for the rapid adoption of reconciliation methods is that they improve forecast accuracy [@Athanasopoulos2024-as].
+Forecast reconciliation, a post-processing step, addresses this by adjusting the base forecasts to the final adjusted set of forecasts that are *coherent* with the aggregation structure.
+Reconciliation approaches such as OLS [@Hyndman2011-jv], WLS [@Hyndman2016-ic], MinT [@Wickramasuriya2019-xq], and more are developed and commonly used in practice.
+As well as guaranteeing coherence, reconciliation methods have been shown to improve forecast accuracy [@Athanasopoulos2024-as].
 For instance, the International Monetary Fund reported improvements in their liquidity forecasts, generating economic benefits via more efficient monetary policy operations [@El-Gemayel2022-hx].
@@
 
-Modern risk management and operations need more than a typical mean or median value--they need quantities from the full predictive distribution that capture uncertainty and tail behaviour, because many decisions are tail-sensitive (e.g., extreme demand in retail settings or energy generation shortfalls).
+Modern risk management and operations need more than a typical mean or median forecast.
+As many decisions are tail-sensitive (e.g., extreme demand in retail settings or energy generation shortfalls), they need quantities that capture uncertainty and tail behaviour.
 While VaR is conceptually straightforward, it ignores the severity of losses beyond the VaR threshold and does not satisfy the subadditivity property\footnote{\textbf{Subadditivity.} A risk measure $\rho$ on loss variables is subadditive if $\rho(X+Y)\le \rho(X)+\rho(Y)$ for any losses $X,Y$.
@@
 
-\color{red}
-Most reconciliation methodology has been developed for point forecasts and are optimised for point scoring functions such as mean squared error (MSE).
+Most reconciliation methodology has been developed for mean forecasts and are optimised for point scoring functions such as mean squared error (MSE).
 Generalising these methods to reconcile probabilistic forecasts does not guarantee optimality for tail-risk measures such as VaR or ES.
-Although @Panagiotelis2023-fs introduced score optimisation for reconciliation weights using proper scoring rules (energy score and variogram score) for full multivariate predictive distributions, they do not directly address the tail functionals.
-This leaves the question: *"How can we construct probabilistic forecast reconciliation that are explicitly optimised for tail-risk measures such as ES?"*
-\color{black}
+Recent works from @Panagiotelis2023-fs that introduced score optimisation for reconciliation weights using proper scoring rules (energy score and variogram score) has shred lights on reconciling full multivariate predictive distributions.
+Building on this, I aim to construct methods that directly address the tail functionals.
+The challenge will be to *"construct forecast reconciliation methods that are explicitly optimised for tail-risk measures such as ES."*
 
@@
 
-As a result, the objective of this thesis is to bring these theoretical advances into the reconciliation framework and develop methods that produce coherent probabilistic forecasts optimised for ES.
+As a result, the objective my PhD research project is to bring these theoretical advances into the reconciliation framework and develop methods that produce coherent probabilistic forecasts optimised for ES.
 The goal is a reliable, consistent, and decision-aligned risk-forecasting framework that helps national and international organisations improve ES backtesting, stabilise capital, and allocate resources more effectively across hierarchical levels.
@@
 
-For a random variable $Y$, the Expected Shortfall (ES) at level $\alpha \in (0, 1)$ is defined as $\mathrm{ES}_\alpha(Y) = E[Y \mid Y < \mathrm{VaR}_\alpha(Y)]$, where $\mathrm{VaR}_\alpha(Y)$ is the quantile $q$ at level $\alpha$.
-Thus, we can define $\widetilde{\mathrm{ES}}_k(\boldsymbol{P}) = \mathrm{ES}_\alpha(\tilde y_k(\boldsymbol{P}))$ as the ES of the $k$-th series after reconciliation, where $\tilde y_k(\boldsymbol{P})$ is the $k$-th element of the reconciled forecasts $\tilde{\boldsymbol{y}}(\boldsymbol{P}) = \boldsymbol P \hat{\boldsymbol{y}}$; and similarly define $\tilde q_k(\boldsymbol{P}) = \mathrm{VaR}_\alpha(\tilde y_k(\boldsymbol{P}))$ as the quantile of the $k$-th series after reconciliation.
+For a random variable $Y$, the Expected Shortfall (ES) at level $\alpha \in (0, 1)$ is defined as $\mathrm{ES}^{(\alpha)}(Y) = E[Y \mid Y < \mathrm{VaR}^{(\alpha)}(Y)]$, where $\mathrm{VaR}^{(\alpha)}(Y)$ is the quantile $q$ at level $\alpha$.
+Hereafter $\alpha$ is dropped for convenience.
+Thus, we can define $\widetilde{\mathrm{ES}}_k(\boldsymbol{P}) = \mathrm{ES}(\tilde y_k(\boldsymbol{P}))$ as the ES of the $k$-th series after reconciliation, where $\tilde y_k(\boldsymbol{P})$ is the $k$-th element of the reconciled forecasts $\tilde{\boldsymbol{y}}(\boldsymbol{P}) = \boldsymbol P \hat{\boldsymbol{y}}$; and similarly define $\tilde q_k(\boldsymbol{P}) = \mathrm{VaR}(\tilde y_k(\boldsymbol{P}))$ as the quantile of the $k$-th series after reconciliation.
 
@@
 $$
-L_\alpha^\mathrm{FZ}(y, q, e) = (\mathbb{I}(y \leq q) - \alpha)(q-y) 
+L^{(\alpha)}(y, q, e) = (\mathbb{I}(y \leq q) - \alpha)(q-y) 
 + \frac{1}{\alpha} \mathrm{exp}(e) \mathbb{I}(y \leq q)(q-y) + \mathrm{exp}(e)(e-q-1),
@@
 Alternatively, we propose to use the Rockafellar-Uryasev joint characterisation of the quantile and ES [@Rockafellar2000-el]:
-$\widetilde{\mathrm{ES}}_k(\boldsymbol{P})=\max_{q}\, E_{\tilde{\boldsymbol{y}}}\!\left[ H_{\alpha}\!\big(\tilde y_k(\boldsymbol{P}), q\big) \right]$ 
+$\widetilde{\mathrm{ES}}_k(\boldsymbol{P})=\max_{q}\, E_{\tilde{\boldsymbol{y}}}\!\left[ H\!\big(\tilde y_k(\boldsymbol{P}), q\big) \right]$ 
 and 
-$\tilde q_k(\boldsymbol{P})=\arg\max_{q}\, E_{\tilde{\boldsymbol{y}}}\!\left[ H_{\alpha}\!\big(\tilde y_k(\boldsymbol{P}), q\big) \right]$,
+$\tilde q_k(\boldsymbol{P})=\arg\max_{q}\, E_{\tilde{\boldsymbol{y}}}\!\left[ H\!\big(\tilde y_k(\boldsymbol{P}), q\big) \right]$,
 where
-$H_{\alpha}(y, q) = q - \frac{1}{\alpha}\max\{0, y - q\}$.
+$H(y, q) = H^{(\alpha)}(y, q) = q - \frac{1}{\alpha}\max\{0, y - q\}$.
 Thus, we jointly reconcile the quantile and ES using the following bilevel optimisation problem:
@@
 \sum_{k=1}^{K} E_{\boldsymbol y}\!\left[
-  L^{\mathrm{FZ}}_{\alpha}\!\big(y_k,\, \tilde q_k(\boldsymbol{P}),\, \widetilde{\mathrm{ES}}_k(\boldsymbol{P})\big)
+  L^{(\alpha)}\!\big(y_k,\, \tilde q_k(\boldsymbol{P}),\, \widetilde{\mathrm{ES}}_k(\boldsymbol{P})\big)
 \right]
@@
 \text{s.t. }&
-\tilde q_k(\boldsymbol{P})=\arg\max_{q}\, E_{\tilde{\boldsymbol{y}}}\!\left[ H_{\alpha}\!\big(\tilde y_k(\boldsymbol{P}), q\big) \right],
+\tilde q_k(\boldsymbol{P})=\arg\max_{q}\, E_{\tilde{\boldsymbol{y}}}\!\left[ H\!\big(\tilde y_k(\boldsymbol{P}), q\big) \right],
 \notag\\
 &
-\widetilde{\mathrm{ES}}_k(\boldsymbol{P})=\max_{q}\, E_{\tilde{\boldsymbol{y}}}\!\left[ H_{\alpha}\!\big(\tilde y_k(\boldsymbol{P}), q\big) \right],
+\widetilde{\mathrm{ES}}_k(\boldsymbol{P})=\max_{q}\, E_{\tilde{\boldsymbol{y}}}\!\left[ H\!\big(\tilde y_k(\boldsymbol{P}), q\big) \right],
 \qquad \forall\,k=1,\ldots,K.
@@
 
+<!-- 
 # Applications
-<!-- 
+
 - Finance: ES is prefered risk measure, hierarchical structure from composite indices and etfs, also accounting for varying weights. Data is publicly available.
@@
 Electricity load data are made available by the Australian Energy Market Operator
--->
 
 # Contributions
-<!-- 
+
 - Methodological: cutting-edge techniques from optimisation, econometrics and statistics will be combined to develop new reconciliation methods, new methods will be applied to important problems in energy, retail, and finance thus demonstrating that the research can inform decisions that lower real-world economic costs.

### 2025-09-22 literature/paper_draft.qmd
Update paper_draft.qmd

@@
 
+<!--
+Para 1
+- Introduce hierarchical time series and its importance in various applications.
+-->
+
 In time series forecasting, aggregation occurs in a variety of settings.
@@
 Starbucks operates in many countries, and each country has multiple cities where they have outlets.
-The sales data is structured hierarchically: the top level is the total sales across all countries, followed by national sales for each country, then city sales for each city within a country, and finally outlet sales for each outlet in a city.
+The sales data is *structured hierarchically*: the top level is the total sales across all countries, followed by national sales for each country, and then individual sales for each outlet in a city.
 As a result, there are over 40,000 individual outlet sales to forecast, plus additional series at higher levels of aggregation such as city and country.
@@
 
-This hierarchical structure is not unique to the Starbucks sales data; it can be found in many other domains, such as national tourism, electricity demand, or Gross Domestic Product (GDP).
+Forecasting data from such hierarchical structures also arises in many other decision-making contexts, from supply chains [@Seaman2022-bb; @Angam2025-od] and energy planning [@Di-Modica2021-ad], to macroeconomics [@El-Gemayel2022-hx; @Li2019-vv] and tourism analysis [@Athanasopoulos2009-lp].
+Stakeholders in these settings need forecasts at several aggregation levels to allocate resources and manage risk.
 The impact of methods for forecasting hierarchical time series has not been limited to academia, with industry also showing a strong interest.
@@
   - Probabilistic methods -->
+
+In practice, when forecasts are produced for all series (often called *base forecasts*), they typically violate the aggregation constraints observed in the data; such forecasts are *incoherent*.
+This can undermine downstream decisions that require internal consistency.
 
@@
 
-The hierarchical structure can be represented as a tree, as shown in @fig-hierarchy.
+The *hierarchical structure* can be represented as a tree, as shown in @fig-hierarchy.
 The top level of the tree represents the total value of all series, while the lower levels represent the series at different levels of disaggregation.
-When there are attributes of interest that are crossed, such as the Starbucks drinks sales at any aggregation level (brand-wise, national, city, or outlet) is also considered by kinds of drinks (e.g., coffees, refreshers), the structure is described as a grouped time series.
+These hierarchical structures naturally form *aggregation constraints* (child levels must sum up to parents).
+When there are attributes of interest that are crossed, such as the Starbucks drinks sales at any aggregation level (brand-wise, national, or outlet) is also considered by kinds of drinks (e.g., coffees, refreshers), the structure is described as a grouped time series.
 As illustrated in @fig-grouped, the aggregation or disaggregation paths are not unique.
@@
 
-For simplicity, we refer to both of these structures as hierarchical time series, we will distinguish between them if and when it is necessary.
-All hierarchical structures can be represented using matrix algebra:
+For simplicity, we refer to both of these structures as hierarchical structure, we will distinguish between them if and when it is necessary.
+Taken together, a collection of time series organised in a hierarchy and subject to aggregation constraints is refered as *hierarchical time series*.
+All hierarchical time series can be represented using matrix algebra:
 
@@
 
-Assume we produce $h$-step-ahead base forecasts $\hat{\boldsymbol{b}}_{t+h|t}$ for the bottom-level series, obtained by any prediction methods. Then pre-multiplying them by $\boldsymbol{S}$ we get:
+Assume we produce $h$-step-ahead base forecasts $\hat{\boldsymbol{b}}_{t+h|t}$ for the bottom-level series, obtained by any prediction methods, pre-multiplying them by $\boldsymbol{S}$ we get:
 
@@
 
-We refer to $\tilde{\boldsymbol{y}}_{t+h|t}$ as coherent forecasts, as they respect the aggregation structure.
+We refer to $\tilde{\boldsymbol{y}}_{t+h|t}$ as *coherent forecasts*, as they respect the aggregation constraints.
 We also refer to this way of obtaining coherent forecasts by summing the bottom-level forecasts as the bottom-up approach.
@@
 
-The NOVELIST (NOVEL Integration of the Sample and Thresholded Covariance) estimator, proposed by @Huang2019-ua, is currently the main focus of this research project.
-It introduces a way to control the target matrix's sparsity, retaining strong correlations while discarding weak, noisy effects.
+The NOVELIST (NOVEL Integration of the Sample and Thresholded Covariance) estimator, proposed by @Huang2019-ua, introduces a way to control the target matrix's sparsity, retaining strong correlations while discarding weak, noisy effects.
 NOVELIST offers more flexibility than the shrinkage estimator, which is useful when we believe that only a few variables are truly correlated.
@@
 
-## POET
+<!-- ## POET
 
@@
 where $\hat{\gamma}_k$ and $\hat{\boldsymbol{\xi}}_k$ are the $k$th largest eigenvalue and the corresponding eigenvector of the sample covariance matrix, respectively, and $T(.)$ is the thresholding function, which can be either soft-thresholding, hard-thresholding, or others.
-
-## PC-adjusted NOVELIST
+-->
+
+<!-- ## PC-adjusted NOVELIST
 
@@
 Alternatively, we can implement the algorithm of @Higham2002-te that computes the nearest positive definite matrix to a given matrix.
+-->
+
+## PC-adjusted Estimator
+
+This method takes the latent factors directly into its construction, and is appealing when there are common drivers in the time series within the hierarchy, as we saw in the Australian tourism example.
+It starts by decomposing the correlation matrix $\hat{\boldsymbol{R}}_1$ into a prominent principle components part (low-rank) and a orthogonal complement part $\hat{\boldsymbol{C}}^K_{1}$ (the correlation matrix after removing the first $K$ principal components).
+Then we can apply either shrinkage or NOVELIST estimator to $\hat{\boldsymbol{R}}_{1, K}$:
+
+$$
+\hat{\boldsymbol{R}}^{g, K}_{1} = \sum_{k = 1}^K \hat{\gamma}_k \hat{\boldsymbol{\xi}}_k \hat{\boldsymbol{\xi}}_k' + g(\hat{\boldsymbol{C}}^K_{1})
+$$
+
+where $g(.)$ is either the shrinkage or NOVELIST estimator, $\hat{\gamma}_k$ and $\hat{\boldsymbol{\xi}}_k$ are the $k$-th largest eigenvalue and the corresponding eigenvector of the sample covariance matrix, respectively.
+Similar to the NOVELIST estimator, $\hat{\boldsymbol{R}}^{*, K}_{1}$ is not guaranteed to be positive definite.
+
+## Scaled Variance
+
+Apply shrinkage or NOVELIST to the correlation matrix of 1-step-ahead base forecast errors, then scale it back to covariance matrix of h-step-ahead base forecast errors using the h-step-ahead standard deviations:
+
+$$
+\hat{\boldsymbol{W}}^{sv}_{h} = \boldsymbol{D}^{1/2}_h g(\hat{\boldsymbol{R}}_1) \boldsymbol{D}^{1/2}_h,
+$$
+
+where $\boldsymbol{D}_h = \text{diag}(\hat{\sigma}_{1,h}, \hat{\sigma}_{2,h}, \ldots, \hat{\sigma}_{n_b,h})$, and $\hat{\sigma}_{i,h}$ is the standard deviation of the $i$-th series' h-step-ahead base forecast errors.
+The asterisk $*$ indicates either shrinkage or NOVELIST estimator.
+
+## Constructing from h-step-ahead residuals
+
+Construct the sample covariance matrix directly from h-step-ahead base forecast errors, then apply either shrinkage or NOVELIST estimator:
+
+$$
+\hat{\boldsymbol{W}}^{g}_{h} = g(\hat{\boldsymbol{W}}_h)
+$$
+
+where $\hat{\boldsymbol{R}}_h$ is the sample correlation matrix of h-step-ahead base forecast errors, and $\circ$ denotes the Hadamard product (element-wise product).
+The asterisk $*$ indicates either shrinkage or NOVELIST estimator.
 
@@
 Each state is further sub-divided into tourism zones, which in turn break down into 77 regions.
+A complete illustration of this geographic hierarchy appears in Appendix @sec-tourism-supplement.
 Intersecting this geographic hierarchy is a second dimension--travel motive--partitioning tourism flows into four categories: holiday, business, visiting friends and relatives, and other.
 Altogether, this yields a grouped system of 560 series, from the most disaggregated regional-purpose cells up to the full national aggregate.
+@tbl-tourism-structure depicts this structure.
 

### 2025-09-22 literature/paper_draft.qmd
add tables to illstrate tourism structure

@@
 library(patchwork)
-library(ReconCov)
 library(dplyr)
@@
 library(lubridate)
-
-path <- "D:/Github/Recon_Honours_Thesis"
+library(knitr)
+library(kableExtra)
+library(stringr)
+
+library(ReconCov)
+
+path <- "D:/Github/Recon_Honours_Thesis/"
 knitr::opts_knit$set(root.dir = path)
@@
 
-<!-- - A table/figure to illustrate the hierarchy -->
+```{r}
+#| tbl-cap: "Hierarchical and grouped structure of Australian domestic tourism flows"
+#| label: tbl-tourism-structure
+
+num_series_geo <- c(1, 7, 27, 77)
+table_structure <- tibble(
+  `Geographical division` = c("Australia", "States", "Zones", "Regions"),
+  `Number of series per geographical division` = num_series_geo,
+  `Number of series per purpose` = 4*num_series_geo,
+  `Total number of series` = 5*num_series_geo
+)
+# add total per row
+table_structure <- rbind(
+  table_structure,
+  c("Total", colSums(table_structure[,-1]))
+)
+
+# width margin smaller (90%)
+table_structure |> 
+  knitr::kable(
+    align = "lccc",
+    booktabs = TRUE
+  )
+```
 
@@
 library(stringr)
-
 desired_order <- c(
@@
   ) |>
-  filter(.model != "mint_sample") %>%
+  filter(.model != "mint_sample") |>
   # Reorder the levels according to the desired order
@@
     )
-  ) %>% 
+  ) |> 
   ggplot(aes(x = h, y = pct_change, color = .model)) +
@@
 
-visnights_full %>% 
+visnights_full |> 
   filter(
@@
     !is_aggregated(State)
-  ) %>% 
+  ) |> 
   model(
     STL(Nights)
-  ) %>% 
-  components() %>% 
+  ) |> 
+  components() |> 
   pivot_longer(
@@
     values_to = "value"
-  ) %>%
+  ) |>
   filter(
@@
     year(Month) >= 2000
-  ) %>%
-  as_tibble() %>%
+  ) |>
+  as_tibble() |>
   mutate(component = recode(component,                     # give them nicer names
                             Nights = "Observed Nights",
-                            trend  = "STL Decomposed Trend")) %>% 
+                            trend  = "STL Decomposed Trend")) |> 
   ggplot(aes(Month, value, color = as.character(State))) +
@@
     )
-  ) %>%
+  ) |>
   ggplot(aes(x = h, y = pct_change, color = .model)) +
@@
   )
-) %>%
+) |>
   kable(
@@
     escape = TRUE
-  ) %>%
-  column_spec(1, width = "1.9cm") %>%
-  column_spec(2, width = "8cm") %>%
-  column_spec(3, width = "4.8cm") %>%
+  ) |>
+  column_spec(1, width = "1.9cm") |>
+  column_spec(2, width = "8cm") |>
+  column_spec(3, width = "4.8cm") |>
   kable_styling(latex_options = "hold_position", font_size = 10)
@@
 
+## Appendix: Australian Domestic Tourism Geographical Hierarchy {#sec-tourism-supplement}
+
+```{r}
+#| tbl-cap: "Geographical divisions of Australia."
+#| label: tbl-tourism-geo
+#| warning: false
+#| message: false
+
+source(paste0(path, "R/table_geo.R"))
+table_geo_labeled <- create_table_geo()
+table_geo_labeled[1,1] <- "Total"
+
+df <- table_geo_labeled %>% 
+  select(Series, Name, Label)
+
+n  <- ceiling(nrow(df) / 2)
+L  <- df[1:n, ]
+R  <- df[(n+1):nrow(df), ]
+
+# pad the shorter side
+if (nrow(R) < nrow(L)) {
+  R <- bind_rows(R, tibble(Series = rep(NA_integer_, n - nrow(R)),
+                           Name = "", Label = ""))
+}
+
+wide <- bind_cols(
+  setNames(L, c("Series","Name","Label")),
+  setNames(R, c("Series","Name","Label"))
+)
+
+kable(
+  wide,
+  align     = c("r","l","l","r","l","l"),
+  col.names = c("Series","Name","Label","Series","Name","Label")
+) |> 
+  kable_styling(latex_options = "hold_position", font_size = 8)
+```
+
 \pagebreak

### 2025-09-22 R/table_geo.R

@@ -1,0 +1,186 @@
+library(dplyr)
+library(tidyr)
+library(stringr)
+
+path <- here::here()
+
+# 0) Prepping --------------------------------
+# State 
+state_map <- list(
+  NSW = c(
+    "Blue Mountains", "Central Coast", "Central NSW",
+    "Hunter", "New England North West", "North Coast NSW", "Outback NSW",
+    "Riverina", "Snowy Mountains", "South Coast", "Sydney",
+    "Capital Country", "The Murray",
+    "Canberra" # ACT is included in NSW for simplicity
+  ),
+  VIC = c(
+    "Ballarat", "Bendigo Loddon", "Central Murray", "Geelong and the Bellarine",
+    "Gippsland", "Great Ocean Road", "High Country", "Macedon", "Mallee",
+    "Melbourne", "Melbourne East", "Murray East", "Peninsula", "Phillip Island",
+    "Spa Country", "Upper Yarra", "Western Grampians", "Wimmera",
+    "Central Highlands", "Lakes", "Goulburn"
+  ),
+  QLD = c(
+    "Brisbane", "Bundaberg", "Capricorn", "Fraser Coast", "Gladstone",
+    "Gold Coast", "Mackay", "Outback Queensland", "Southern Queensland Country",
+    "Sunshine Coast", "Townsville", "Tropical North Queensland",
+    "Whitsundays"
+  ),
+  SA = c(
+    "Adelaide", "Adelaide Hills", "Barossa", "Clare Valley", "Eyre Peninsula",
+    "Fleurieu Peninsula", "Flinders Ranges and Outback", "Kangaroo Island",
+    "Limestone Coast", "Murray River, Lakes and Coorong", "Riverland",
+    "Yorke Peninsula"
+  ),
+  WA = c(
+    "Australia's Coral Coast", "Australia's Golden Outback",
+    "Australia's North West", "Australia's South West", "Destination Perth"
+  ),
+  TAS = c(
+    "East Coast", "Hobart and the South",
+    "Launceston and the North", "North West", "West Coast"
+  ),
+  NT = c(
+    "Alice Springs", "Barkly", "Darwin", "Katherine Daly",
+    "Litchfield Kakadu Arnhem", "Lasseter", "MacDonnell"
+  )
+)
+region_state_lookup <- tibble(
+  State  = rep(names(state_map), lengths(state_map)),
+  Region = unlist(state_map)
+)
+# Zone 
[…]

### 2025-09-23 literature/archived.qmd
revise tourism result discussion

--- 2025-09-22 literature/paper_draft.qmd
@@
----
-title: "Enhancing Forecast Reconciliation: A Study of Alternative Covariance Estimators"
-author: 
-  - Vincent Su
-  - Shanika Wickramasuriya (supv.)
-  - George Athanasopoulos (supv.)
-format: 
-  pdf:
-    include-in-header:
-      text: |
-        <!-- \usepackage{amsmath} -->
-        <!-- \usepackage{mathspec} -->
-        \usepackage{algorithm}
-        \usepackage{algpseudocode}
-        \usepackage[colorlinks=true, linkcolor=blue, citecolor=blue, urlcolor=blue]{hyperref}
-        \usepackage{array}
-    mathspec: true
-    number-sections: true
-# crossref:
-#   custom:
-#     - kind: float
-#       key: alg
-#       reference-prefix: Algorithm
-#       caption-prefix: Algorithm
-#       # for PDF output only, tell LaTeX to use the `algorithm` env:
-#       latex-env: algorithm
-execute:
-  cache: true
-  echo: false
-  warning: false
-bibliography: D:/Github/Recon_Honours_Thesis/references.bib
-# bibliographystyle: apa
-csl: D:/Github/Recon_Honours_Thesis/apa.csl
-editor_options: 
-  chunk_output_type: console
----
-
-<!-- # Notation {-} -->
-
-<!-- -   Scalar $y_t$ -->
-
-<!-- -   Vector $\boldsymbol{y}_t$ -->
-
-<!-- -   Matrix $\boldsymbol{S}$ -->
-
-<!-- -   Covariance matrix of in-sample 1-step-ahead base forecast errors $\hat{\boldsymbol{W}}_1$ -->
-
-<!-- -   Its shrinkage estimator with diagonal target $\hat{\boldsymbol{W}}^{shr}_{1, D}$ -->
-
-<!-- -   Its NOVELIST estimator $\hat{\boldsymbol{W}}^{shr}_{1, thr}$ -->
-
-```{r library}
-library(ggplot2)
-library(patchwork)
-library(dplyr)
-library(tidyr)
-library(fable)
-library(fabletools)
-library(feasts)
-library(tsibble)
-library(lubridate)
-library(knitr)
-library(kableExtra)
-library(stringr)
-
-library(ReconCov)
-
-path <- "D:/Github/Recon_Honours_Thesis/"
-knitr::opts_knit$set(root.dir = path)
-# setwd(path)
-```
-
-# Abstract {-}
-
-<!-- [This is pasted from the Project Description]{style="color:red;"} -->
-
-A collection of time series connected via a set of linear constraints is known as hierarchical time series.
-Forecasting these series without respecting the hierarchical nature of the data can lead to incoherent forecasts across aggregation levels and lower accuracy.
-To mitigate this issue, various forecast reconciliation approaches have been proposed in the literature, where the individual forecasts are adjusted to satisfy the aggregation constraints.
-Among these, **MinT** (Minimum Trace) is widely used, however, it requires a good estimate of the covariance matrix of the base forecast errors.
-The current practice is to use the shrinkage estimator (often shrinking toward a diagonal matrix), but it lacks flexibility and might not fully utilise the prominent latent structure presented.
-In this project, we aim to assess the forecasting performance of MinT when different covariance estimators are used, namely NOVELIST (NOVEL Integration of the Sample and Thresholded Covariance), POET (Principal Orthogonal complEment Thresholding), and others.
-
-# Introduction
-
-<!-- ############  Intro  ############ -->
-
-<!--
-Para 1
-- Introduce hierarchical time series and its importance in various applications.
--->
-
-In time series forecasting, aggregation occurs in a variety of settings.
-While a formal definition of hierarchical time series can be found in @sec-hierarchy, we can think of Starbucks sales data as an illustrative example.
-Starbucks operates in many countries, and each country has multiple cities where they have outlets.
-The sales data is *structured hierarchically*: the top level is the total sales across all countries, followed by national sales for each country, and then individual sales for each outlet in a city.
-As a result, there are over 40,000 individual outlet sales to forecast, plus additional series at higher levels of aggregation such as city and country.
-The hierarchy can be even more complex if we consider the sales of different kinds of drinks (e.g., coffees, teas, refreshers) at each aggregation level.
-
-Forecasting data from such hierarchical structures also arises in many other decision-making contexts, from supply chains [@Seaman2022-bb; @Angam2025-od] and energy planning [@Di-Modica2021-ad], to macroeconomics [@El-Gemayel2022-hx; @Li2019-vv] and tourism analysis [@Athanasopoulos2009-lp].
-Stakeholders in these settings need forecasts at several aggregation levels to allocate resources and manage risk.
-The impact of methods for forecasting hierarchical time series has not been limited to academia, with industry also showing a strong interest.
-Many companies and organisations have adopted these methods in practice, including Amazon, the International Monetary Fund, IBM, SAP, and more [@Athanasopoulos2024-as].
-
-<!-- ##########  Literature Review  ########## -->
-
-<!-- - Talk about the history and evolution of forecasting hierarchical time series, starting from the early heuristic methods to the modern statistical approaches.
-
-  - Single level methods
-  - Optimal combination methods (OLS, WLS)
-  - MinT
-  - Bayesian, Machine learning
-  - Probabilistic methods -->
-
-In practice, when forecasts are produced for all series (often called *base forecasts*), they typically violate the aggregation constraints observed in the data; such forecasts are *incoherent*.
-This can undermine downstream decisions that require internal consistency.
-
-Traditionally, forecasting these hierarchical time series has been done using single-level methods, such as bottom-up, top-down, and middle-out approaches.
-Bottom-up methods involve generating forecasts for the bottom-level series and aggregating them to higher levels.
-Top-down methods start with forecasts for the only top-level series and disaggregate them down.
-Middle-out methods combine both approaches by forecasting middle-level series and then aggregating or disaggregating as needed.
-Despite their simplicity, these methods only anchor forecasts to a single level, implying a large loss of information on the hierarchy's inherent correlation structure.
-Additionally, the most disaggregated series often are very noisy or even intermittent, and the higher-level data might be smoother due to the aggregation.
-Furthermore, as we saw from the Starbucks example considering the sales of different kinds of drinks at each aggregation level -- formally defined as grouped structure in @sec-hierarchy -- the disaggregation becomes more complex since the disaggregation paths are not unique.
-Consequently, these single-level methods tend to give poor results across other levels of the hierarchy.
-
-To overcome these issues, forecast reconciliation was introduced by @Hyndman2011-jv, and later developed by @Van_Erven2015-nx, @Hyndman2016-ic, @Ben-Taieb2019-yx, @Wickramasuriya2019-xq, and others to achieve coherency in point forecasts and enhance accuracy.
-Forecast reconciliation projects a collection of independent base forecasts into a set of coherent forecasts that respect the linear constraints defining a hierarchical or grouped time-series system.
-
-Among the modern reconciliation strategies, the Min Trace (MinT) approach developed by @Wickramasuriya2019-xq is widely used and perform significantly well under right conditions.
-Despite its properties, MinT relies on a good high-dimensional covariance estimator, which is positive-definite.
-However, not many researchers have explored this issue in the current literature, except @Carrara2025-rz, who introduced a new estimator for MinT--Double Shrinkage estimator.
-The current stage of this research piece is fairly early, leaving many potential gaps still not yet to explore.
-As a result, a study of alternative high-dimensional covariance estimators for MinT is timely and desirable.
-
-The remainder of the paper is organised as follows.
-@sec-pre-theory provides the basic theoretical framework for hierarchical time series and forecasting, and Min Trace approach, introducing notations, terminologies, and motivations for alternative estimators.
-@sec-cov-est walks through the main covariance estimators this paper explores, and argues their strengths and weaknesses.
-@sec-simulation covers the simulation design and currently explores the performance of NOVELIST on MinT.
-@sec-empirical shows a real-world application of MinT with NOVELIST, which produces results that did not occured in the simulation settings, suggesting further inspection and analysis.
-
-<!-- Include more details on the gaps for each of these methods, such as bias, high variance? -->
-
-<!-- - Discuss the interests in MinT and how it has become a standard method for forecast reconciliation.
-
-- Discuss the MinT's reliance on a good estimate of the covariance matrix of base forecast errors and other gaps
-
-  - Empirical evidence of MinT under perform
-  - Comparison with other methods
-
-- Explain why this paper focus on exploring alternative covariance estimators for MinT. And is there any paper talk about this.
-
-  - Is better estimate of W_h really lead to better performance?
-
-- Talk about the paper outline and what will be covered in the following sections. -->
-
-<!--
--   **Problem Statement:**
-
-    The sample covariance matrix, although natural, suffers in high-dimensional settings.
-Especially when the number of series $p$ is huge and larger than the time dimension $T$, the sample covariance matrix is non-positive definite (rank T if p\>T).
-
-    The shrinkage estimators come in to tackle this issue.
-The shrinkage estimator with diagonal target (often shrinking toward a diagonal matrix) is proven to produce a guaranteed PD matrix (Schäfer & Strimmer, 2005).
-However, as it shrinks the covariance matrix toward a diagonal one, it does not have flexibility and might neglect the prominent structure presented in the covariance matrix.
-
-    An alternative approach is to perform shrinkage of the sample covariance towards its thresholded version, instead of a diagonal matrix.
-This is the NOVELIST (NOVEL Integration of the Sample and Thresholded covariance estimators) method proposed by Huang & Fryzlewicz (2019).
-They introduced thresholding functions applied only to off-diagonal elements, allowing for more flexibility in the estimation.
-
-    [... can include more estimators ...]{style="color:red;"}
-
--   **Research Aim:**
-
-    This paper assesses the reconciled forecasting performance of MinT approach using various covariance estimators, with a focus on the NOVELIST estimator.
-
--   **Paper Outline:**
-
-    The paper is structured as follows:
-
-    -   A literature review of forecast reconciliation and covariance estimation.
-    -   A description of the methodology, including the NOVELIST estimator and its principal-component-adjusted variant.
-    -   An experimental design using both synthetic and real hierarchical time series.
-    -   Empirical results and discussion.
-    -   Conclusions and suggestions for future work.
--->
-
-<!--
-# Literature Review
-
-## Forecast Reconciliation in Hierarchical and Grouped Time Series
-
-Forecast reconciliation converts a collection of independent base forecasts into a set of coherent forecasts that respect the linear constraints defining a hierarchical or grouped time-series system.
-Early work focused on heuristic single-level strategies, including bottom-up, top-down, and middle-out (...), each of which exploits only part of the information in the hierarchy and can induce bias or high variance.
-
--   Cite the single level approach
--   Athanasopoulos et al., 2024
-
-Hyndman et al.
-(2011) first showed that all single-level methods can be written as $\tilde{y} = SG\hat{y}$, where $S$ is the summing matrix and $G$ is a matrix that maps base forecasts $\hat{y}$ to into the bottom level.
-Treating reconciliation as a GLS regression problem, Hyndman et al.
-(2011) found that it yields a solution for $G$, but the required covariance matrix of reconcilation error is not identifiable in practice (Wickramasuriya et al., 2019).
-
--   (Talk more about how others transform it to OLS, WLS,..)
--   Di Fonzo and Marini (2011)
--   Athanasopoulos et al. (2009)
-
-Wickramasuriya, Athanasopoulos & Hyndman (2019) reframed the problem by taking an optimisation approach rather than the regression.
-They formulated the problem as minimising the variances of all reconciled forecasts, which happens to be equivalent to minimising the trace of the covariance matrix (sum of the diagonal elements).
-This is known as the Minimum Trace (MinT) reconciliation method.
-The MinT solution is given by $G_h = (S'W_h^{-1}S)^{-1}S'W_h^{-1}$, and $W_h$ is the covariance matrix of the h-step-ahead base forecast errors.
-
-The MinT approach is an algebraical generalisation of the GLS, and the OLS and WLS methods are special cases of MinT when $W_h$ is a diagonal or identity matrix, respectively.
-However, the MinT solution hinges on a reliable estimate of the h-step-ahead base forecast error covariance $W_h$.
-In high-dimensional setting, the usual sample covariance matrix is unstable, thus we need alternative covariance estimators.
-
--   Structural Scaling, based only on the struc- ture of the hierarchy (Athanasopoulos et al., 2017)
--   Shrinkage estimators (Schäfer & Strimmer, 2005; Ledoit & Wolf, 2004)
-
-Empirical evaluations have demonstrated that MinT with an appropriate covariance estimate often outperforms earlier methods in both simulation and real data studies.
-
-Wickramasuriya et al.
-(2019) adopt the diagonal-target Ledoit–Wolf approach and show analytically that any scalar multiple $k_h$ from $W_h = k_hW_1$ suffices for point-forecast reconciliation.
-However, simulation and empirical evidence suggest that ignoring cross-series error correlation can leave accuracy untapped, especially in highly-coupled systems (energy, retail SKUs).
-Their subsequent work imposes non-negativity constraints via quadratic programming citeturn0file2 and studies bias-robust variants (MinT-U).
-
-Empirical findings to date
-
-Athanasopoulos et al.
-(2024) synthesise the accumulated evidence: MinT dominates single-level baselines in most studies; WLS or diagonal-shrinkage MinT is the current industry default; but gains plateau when off-diagonals are weak citeturn0file9.
-Only a handful of papers explore richer covariance structures (e.g., block or factor shrinkage), and none benchmark them systematically against NOVELIST-type estimators.
-
-## Covariance Estimation in High Dimensions
-
--   Limitations of the sample covariance matrix.
--   Estimators used by Wickramasuriya et al. (2019).
--   Shrinkage estimators:
-    -   Diagonal shrinkage (e.g., Schäfer & Strimmer, Ledoit & Wolf).
-    -   NOVELIST estimator and its Cross-validation & PC-adjusted variant.
-    -   ...
-
-## Relevance to Forecast Reconciliation
-
--   Discuss how covariance estimation affects MinT performance.
--   Identify research gaps.
--->
-
-# Theoretical Framework {#sec-pre-theory}
-
-## Hierarchical tructure {#sec-hierarchy}
-
-The *hierarchical structure* can be represented as a tree, as shown in @fig-hierarchy.
-The top level of the tree represents the total value of all series, while the lower levels represent the series at different levels of disaggregation.
-These hierarchical structures naturally form *aggregation constraints* (child levels must sum up to parents).
-When there are attributes of interest that are crossed, such as the Starbucks drinks sales at any aggregation level (brand-wise, national, or outlet) is also considered by kinds of drinks (e.g., coffees, refreshers), the structure is described as a grouped time series.
-As illustrated in @fig-grouped, the aggregation or disaggregation paths are not unique.
-
-![A 2-level hierarchical tree structure](D:/Github/Recon_Honours_Thesis/research_proposal/figs/hierarchical_structure.png){#fig-hierarchy width="40%"}
-
-![A 2-level grouped structure, which can be considered as the union of two hierarchical trees with common top and bottom level series](D:/Github/Recon_Honours_Thesis/research_proposal/figs/grouped_structure.png){#fig-grouped width="90%"}
-
-For simplicity, we refer to both of these structures as hierarchical structure, we will distinguish between them if and when it is necessary.
-Taken together, a collection of time series organised in a hierarchy and subject to aggregation constraints is refered as *hierarchical time series*.
-All hierarchical time series can be represented using matrix algebra:
-
-$$
-\boldsymbol{y}_t = \boldsymbol{S} \boldsymbol{b}_t,
-$$
-
-where $\boldsymbol{S}$ is a summing matrix of order $n \times n_b$ which aggregates the bottom-level series $\boldsymbol{b}_t \in \mathbb{R}^{n_b}$ to the series at aggregation levels above.
-The vector $\boldsymbol{y}_t \in \mathbb{R}^n$ contains all observations at time $t$.
-The summing matrix $\boldsymbol{S}$ for the tree structure in @fig-hierarchy is:
-
-$$
-\boldsymbol{S} = 
-\left[
-\begin{array}{cccc}
-1 & 1 & 1 & 1 \\
-1 & 1 & 0 & 0 \\
-0 & 0 & 1 & 1 \\
-\multicolumn{4}{c}{\boldsymbol{I_4}}
-\end{array}
-\right].
-$$
-
-Assume we produce $h$-step-ahead base forecasts $\hat{\boldsymbol{b}}_{t+h|t}$ for the bottom-level series, obtained by any prediction methods, pre-multiplying them by $\boldsymbol{S}$ we get:
-
-$$
-\tilde{\boldsymbol{y}}_{t+h|t} = \boldsymbol{S} \hat{\boldsymbol{b}}_{t+h|t} \, .
-$$ {#eq-sb}
-
-We refer to $\tilde{\boldsymbol{y}}_{t+h|t}$ as *coherent forecasts*, as they respect the aggregation constraints.
-We also refer to this way of obtaining coherent forecasts by summing the bottom-level forecasts as the bottom-up approach.
-However, generating forecasts this way is anchored only to prediction models at a single level, and will not be utilising the inherent information from other levels.
-This drawback applies to the top-down and middle-out approaches.
-For example, the bottom-level data can be very noisy or even intermittent, and the higher-level data might be smoother due to the aggregation.
-
-Another issue with expressing reconciled methods as in @eq-sb is that it restricts the reconciliation to only single-level approaches.
-Thus, @Hyndman2011-jv suggested a generalised expression for all existing methods, which also provides a framework for new methods to be developed:
-
-$$
-\tilde{\boldsymbol{y}}_{t+h|t} = \boldsymbol{S} \boldsymbol{G} \hat{\boldsymbol{y}}_{t+h|t} \; ,
-$$ {#eq-sgy}
-
-for a suitable $n_b \times n$ matrix $\boldsymbol{G}$.
-$\boldsymbol{G}$ maps the base forecasts of all levels $\hat{\boldsymbol{y}}_{t+h|t}$ down into the bottom level, which is then aggregated to the higher levels by $\boldsymbol{S}$.
-The choice of $\boldsymbol{G}$ determines the composition of reconciled forecasts $\tilde{\boldsymbol{y}}_{t+h|t}$, and modern reconciliation methods are developed to estimate $\boldsymbol{G}$.
-
-<!-- including the OLS and WLS developed by @Hyndman2011-jv and @Hyndman2016-ic, respectively. A more detailed explanation of these methods can be found in the @sec-ols-wls Appendix. -->
-
-## The Minimum Trace (MinT) Reconciliation
-
-@Wickramasuriya2019-xq framed the problem as minimising the variances of all reconciled forecast errors $\text{Var}[y_{t+h} - \tilde{y}_{t+h|t}] = \boldsymbol{S} \boldsymbol{G} \boldsymbol{W}_h \boldsymbol{G}' \boldsymbol{S}'$, where $\boldsymbol{W}_h = \mathbb{E}( \hat{\boldsymbol{e}}_{t+h|t} \; \hat{\boldsymbol{e}}_{t+h|t}' )$ is the positive definite covariance matrix of the $h$-step-ahead base forecast errors.
-They showed that this is equivalent to minimising the trace of the reconciled forecast error covariance matrix (sum of the diagonal elements - the variances).
-The Minimum Trace (MinT) solution is given by
-
-$$
-\boldsymbol{G} = (\boldsymbol{S}' \boldsymbol{W}_h^{-1} \boldsymbol{S})^{-1}
-\boldsymbol{S}' \boldsymbol{W}_h^{-1} .
-$$
-
-@Wickramasuriya2019-xq also showed that MinT is an algebraic generalisation of the GLS, and the OLS and WLS methods are special cases of MinT when $\boldsymbol{W}_h$ is an identity matrix $I_{n_b}$ and a diagonal matrix $\text{diag}( \boldsymbol{W}_h )$, respectively.
-In this paper, we place our main focus on the MinT method.
-
-The MinT solution hinges on a reliable, positive-definite estimate of $\boldsymbol{W}_h$, which is challenging to estimate in high-dimensional setting.
-The sample covariance matrix is unstable and non-positive-definite when the number of series $n$ is huge and larger than the time dimension $T$.
-To tackle this issue, the original paper @Wickramasuriya2019-xq adopted the diagonal-target shrinkage estimator from @Schafer2005-yw, given by
-
-$$
-\hat{\boldsymbol{W}}^{shr}_{1} = \lambda_D \hat{\boldsymbol{W}}_{1, D} + (1 - \lambda_D) \hat{\boldsymbol{W}}_1 \, ,
-$$
-
-where $\hat{\boldsymbol{W}}_{1, D}$ is a diagonal matrix comprising the diagonal entries  $\text{diag}( \hat{\boldsymbol{W}}_1 )$.
-We refer to any $\lambda \in [0,1]$ as the shrinkage intensity parameter, the subscript specifies which estimator it belongs to.
-This approach shrinks the covariance matrix $\hat{\boldsymbol{W}}_1$ towards its diagonal matrix, meaning the off-diagonal elements are shrunk towards zero while the diagonal ones remain unchanged.
-
-@Schafer2005-yw also proposed an estimate of the optimal shrinkage intensity parameter $\lambda_D$:
-
-$$
-\hat{\lambda}_D = \frac{\sum_{i \neq j} \widehat{Var}(\hat{r}_{ij})} 
-{\sum_{i \neq j} \hat{r}_{ij}^2} \, ,
-$$
-
-where $\hat{r}_{ij}$ is the $ij$th element of $\hat{\boldsymbol{R}}_1$, the 1-step-ahead sample correlation matrix (obtained from $\hat{\boldsymbol{W}}_1$).
-The optimal estimate is obtained by minimising $MSE(\hat{\boldsymbol{W}}_1) = Bias(\hat{\boldsymbol{W}}_1)^2 + Var(\hat{\boldsymbol{W}}_1)$.
-More specifically, we trade the unbiasedness of the sample covariance matrix for a lower variance.
-
-However, the hierarchical time series data often exhibit a prominent principal components structure, which is not fully taken advantage.
-Taking an example of the Australian domestic overnight trips data set [@tourism-au], where the national trips are disaggregated into states and territories, and further into regions.
-We then fit ETS models to all series, using the algorithm from Fabletools R package [@O-Hara-Wild2024-we], and compute the one-step-ahead in-sample base forecast error covariance matrix $\hat{\boldsymbol{W}}_1$.
-The twenty largest eigenvalues of the covariance matrix are plotted in @fig-eigen.
-We can see that the point of inflexion occurs at the component with 5th largest eigenvalue, indicating a prominent principal components structure.
-
-![Twenty largest eigenvalues of one-step-ahead in-sample base forecast error covariance, Australian domestic overnight trips](D:/Github/Recon_Honours_Thesis/research_proposal/figs/eigen_tourism.png){#fig-eigen width="70%"}
-
-Additionally, the shrinkage estimator shrinks all off-diagonal elements towards zeros with equal weights $\lambda_D$.
-We might prefer to better preserve strong signals, and largely reduce the effects of small, noisy correlations.
-In the next sections, we will explore several options that take these two issues into account.
-
-# Covariance Estimation Approaches {#sec-cov-est}
-
-## NOVELIST Estimator
-
-The NOVELIST (NOVEL Integration of the Sample and Thresholded Covariance) estimator, proposed by @Huang2019-ua, introduces a way to control the target matrix's sparsity, retaining strong correlations while discarding weak, noisy effects.
-NOVELIST offers more flexibility than the shrinkage estimator, which is useful when we believe that only a few variables are truly correlated.
-However, it does not guarantee to be positive definite.
-
-The method is based on the idea of soft-thresholding the sample covariance matrix, then performing shrinkage towards this thresholded version.
-This introduces an extra parameter, the threshold $\delta$, which is used to control the amount of soft-thresholding.
-The NOVELIST estimator is given by:
-
-$$
-\hat{\boldsymbol{W}}^{N}_{1} = \lambda_{\delta} \hat{\boldsymbol{W}}_{1, \delta} + (1 - \lambda_{\delta}) \hat{\boldsymbol{W}}_1,
-$$ {#eq-novelist-cov}
-
-where $\hat{\boldsymbol{W}}_{1, \delta}$ is the thresholded version of $\hat{\boldsymbol{W}}_{1}$. By convenient setting, we can rewrite it in terms of correlation:
-
-$$
-\hat{\boldsymbol{R}}^{N}_{1} = \lambda_{\delta} \hat{\boldsymbol{R}}_{1,\delta} + (1 - \lambda_{\delta}) \hat{\boldsymbol{R}}_1,
-$$ {#eq-novelist-cor}
-
-In this setting, $\hat{\boldsymbol{R}}_{1,\delta}$ is the thresholded correlation matrix, where each element is regularised by:
-
-$$
-\hat{r}_{1,ij}^\delta = \text{sign}(\hat{r}_{1,ij}) \, 
-\text{max}(|\hat{r}_{1,ij}| - \delta, \; 0),
-$$ {#eq-soft-thr}
-
-where $\delta \in [0,1]$ is the threshold parameter.
-For a given threshold $\delta$, @Huang2019-ua derived an analytical expression for the optimal shrinkage intensity parameter $\lambda(\delta)$ using Ledoit-Wolf's lemma [@Ledoit2003-qv], following similar logic to @Schafer2005-yw.
-It can be computed as:
-
-<!-- - \textcolor{red}{@Huang2019-ua mentioned Ledoit-Wolf's lemma} -->
-
-$$
-\hat{\lambda}(\delta) = \frac{
-  \sum_{i \neq j} \widehat{Var}(\hat{r}_{1,ij}) \; \boldsymbol{1}(|\hat{r}_{1,ij}| \leq \delta)
-} {
-  \sum_{i \neq j} (\hat{r}_{1,ij} - \hat{r}_{1,ij}^\delta)^2
-},
-$$ {#eq-lambda-thr}
-
-where $\boldsymbol{1}(.)$ is the indicator function.
-
-On the other hand, the optimal threshold $\delta^*$ does not have a closed-form solution, and is typically obtained by executing a rolling-window cross-validation procedure.
-The idea is to find the threshold $\hat{\delta^*}$, with the corresponding $\hat{\lambda}^*$ and $\hat{\boldsymbol{R}}^{N}_{1}(\hat{\delta^*}, \hat{\lambda}^*)$, that minimises the average out-of-sample 1-step-ahead reconciled forecast mean squared error over all windows.
-The formal algorithm is given in the @sec-novelist_cv Appendix.
-Although it is not required to fit forecasting models multiple times, the cross-validation procedure is still computationally expensive as it computes the NOVELIST estimator and perform reconciliation for each threshold value.
-
-We also tested out minimising 1- to h-step-ahead MSE in the cross-validation procedure.
-Surprisingly, it returns almost the same best threshold parameter is in the 1-step-ahead case above.
-Note that when $\delta \in \bigl[ \text{max}_{i \neq j}|\hat{r}_{1,ij}|, \; 1 \bigr]$, the NOVELIST estimator collapses to the shrinkage estimator, and when $\delta = 0$, it becomes the sample covariance matrix.
-
-<!-- ## POET
-
-The POET (Principal Orthogonal complEment Thresholding) estimator, proposed by @Fan2013-jz, is another "sparse" + "non-sparse" covariance estimator.
-It takes the latent factors directly into its construction, and is appealing when there are common drivers in the time series within the hierarchy, as we saw in the Australian tourism example.
-
-The POET method starts by decomposing the correlation matrix $\hat{\boldsymbol{R}}_1$ into a prominent principle components part (low-rank) and a orthogonal complement part $\hat{\boldsymbol{R}}_{1, K}$ (the correlation matrix after removing the first $K$ principal components).
-Then it applies thresholding to $\hat{\boldsymbol{R}}_{1, K}$.
-The POET estimator is given by:
-
-$$
-\hat{\boldsymbol{R}}^{K}_{1} = \sum_{k = 1}^K \hat{\gamma}_k \hat{\boldsymbol{\xi}}_k \hat{\boldsymbol{\xi}}_k' + T(\hat{\boldsymbol{R}}_{1, K})
-$$
-
-where $\hat{\gamma}_k$ and $\hat{\boldsymbol{\xi}}_k$ are the $k$th largest eigenvalue and the corresponding eigenvector of the sample covariance matrix, respectively, and $T(.)$ is the thresholding function, which can be either soft-thresholding, hard-thresholding, or others.
--->
-
-<!-- ## PC-adjusted NOVELIST
-
-This approach is best of both worlds, leveraging the strengths of both NOVELIST and POET.
-The PC-adjusted (Principal-Component-adjusted) NOVELIST overcomes the shortcomings of the current shrinkage estimator, taking prominent PCs into account while also offers extra flexibility.
-The idea is to apply the NOVELIST estimator to the orthogonal complement part $\hat{\boldsymbol{R}}_{1, K}$, and then add the principal components part back.
-The PC-adjusted NOVELIST estimator is formulated as:
-
-$$
-\hat{\boldsymbol{R}}^{N, K}_{1} = \sum_{k = 1}^K \hat{\gamma}_k \hat{\boldsymbol{\xi}}_k \hat{\boldsymbol{\xi}}_k' + \hat{\boldsymbol{R}}^{N}_{1, K}
-\; ,
-$$
-
-where $\hat{\boldsymbol{R}}^{N}_{1, K}$ is the NOVELIST estimator applied to the orthogonal complement part $\hat{\boldsymbol{R}}_{1, K}$.
-Similar to the NOVELIST estimator, $\hat{\boldsymbol{R}}^{N, K}_{1}$ is not guaranteed to be positive definite.
-
-Methods to ensure positive definiteness of the NOVELIST estimator (and its PC-adjusted variant) will be explored and studied in the project.
-@Huang2019-ua proposed to diagonalise the NOVELIST estimator and replace any eigenvalues that fall under a certain small positive threshold by the value of that threshold.
-Alternatively, we can implement the algorithm of @Higham2002-te that computes the nearest positive definite matrix to a given matrix.
--->
-
-## PC-adjusted Estimator
-
-This method takes the latent factors directly into its construction, and is appealing when there are common drivers in the time series within the hierarchy, as we saw in the Australian tourism example.
-It starts by decomposing the correlation matrix $\hat{\boldsymbol{R}}_1$ into a prominent principle components part (low-rank) and a orthogonal complement part $\hat{\boldsymbol{C}}^K_{1}$ (the correlation matrix after removing the first $K$ principal components).
-Then we can apply either shrinkage or NOVELIST estimator to $\hat{\boldsymbol{R}}_{1, K}$:
-
-$$
-\hat{\boldsymbol{R}}^{g, K}_{1} = \sum_{k = 1}^K \hat{\gamma}_k \hat{\boldsymbol{\xi}}_k \hat{\boldsymbol{\xi}}_k' + g(\hat{\boldsymbol{C}}^K_{1})
-$$
-
-where $g(.)$ is either the shrinkage or NOVELIST estimator, $\hat{\gamma}_k$ and $\hat{\boldsymbol{\xi}}_k$ are the $k$-th largest eigenvalue and the corresponding eigenvector of the sample covariance matrix, respectively.
-Similar to the NOVELIST estimator, $\hat{\boldsymbol{R}}^{*, K}_{1}$ is not guaranteed to be positive definite.
-
-## Scaled Variance
-
-Apply shrinkage or NOVELIST to the correlation matrix of 1-step-ahead base forecast errors, then scale it back to covariance matrix of h-step-ahead base forecast errors using the h-step-ahead standard deviations:
-
-$$
-\hat{\boldsymbol{W}}^{sv}_{h} = \boldsymbol{D}^{1/2}_h g(\hat{\boldsymbol{R}}_1) \boldsymbol{D}^{1/2}_h,
-$$
-
-where $\boldsymbol{D}_h = \text{diag}(\hat{\sigma}_{1,h}, \hat{\sigma}_{2,h}, \ldots, \hat{\sigma}_{n_b,h})$, and $\hat{\sigma}_{i,h}$ is the standard deviation of the $i$-th series' h-step-ahead base forecast errors.
-The asterisk $*$ indicates either shrinkage or NOVELIST estimator.
-
-## Constructing from h-step-ahead residuals
-
-Construct the sample covariance matrix directly from h-step-ahead base forecast errors, then apply either shrinkage or NOVELIST estimator:
-
-$$
-\hat{\boldsymbol{W}}^{g}_{h} = g(\hat{\boldsymbol{W}}_h)
-$$
-
-where $\hat{\boldsymbol{R}}_h$ is the sample correlation matrix of h-step-ahead base forecast errors, and $\circ$ denotes the Hadamard product (element-wise product).
-The asterisk $*$ indicates either shrinkage or NOVELIST estimator.
-
-# Simulation Design {#sec-simulation}
-
-<!-- -   Talk about description of the hierarchical time series data set (e.g., economic, financial, or synthetic data). -->
-<!-- -   Characteristics such as dimensionality, frequency, and hierarchical structure. -->
-
-<!-- **Experimental Design** -->
-
-<!-- -   Design different cases of simulation studies and real-data experiments. -->
-<!-- -   Metrics: Forecast accuracy (e.g., RMSE, MAE), reconciliation error reduction, and matrix stability... -->
-<!-- -   Visualisation... -->
-
-The general design of data generating process for bottom-level series is a stationary VAR(1) process, with the following structure:
-
-$$
-\boldsymbol{b}_t = \boldsymbol{A} \boldsymbol{b}_{t-1} + \boldsymbol{\epsilon}_t,
-$$
-
-where $\boldsymbol{A}$ is a $n_b \times n_b$ block diagonal matrix of autoregressive coefficients $\boldsymbol{A} = diag(\boldsymbol{A}_1, \ldots, \boldsymbol{A}_m)$, with each $\boldsymbol{A}_i$ being a $n_{b,i} \times n_{b,i}$ matrix.
-The block diagonal structure ensures that the time series are grouped into $m$ groups, with each group having its own autoregressive coefficients.
-This aim to simulate the interdependencies between the time series within each group, where reconciliation will be expected to better performed than the usual base forecasts.
-
-The model is added with a Gaussian innovation process $\boldsymbol{\epsilon}_t$, with covariance matrix $\boldsymbol{\Sigma}$.
-The covariance matrix $\boldsymbol{\Sigma}$ is generated specifically using the Algorithm 1 in @Hardin2013-wu:
-
-1.
-A compound symmetric correlation matrix is used for each block of size $n_{b,i}$ in $\boldsymbol{A}_i$, where the entries $\rho_i$ for each block $i$ are sampled from a uniform distribution between 0 and 1.
-They are baseline correlations within group.
-
-2. A constant correlation, which is smaller than $\text{min} \{\rho_1, \rho_2, \dots, \rho_m \}$, is imposed on the entries between different blocks. It serves as baseline correlations between group.
-
-3. The entry-wise random noise is added on top of the entire correlation matrix.
-
-4. The covariance matrix $\boldsymbol{\Sigma}$ is then constructed by uniform sampling of standard deviations, in a range of $[\sqrt{2}, \sqrt{6}]$, for all $n_b$ series.
-
-We will randomly flip the signs of the covariance elements, which will create a more realistic structure in the innovation process.
-This can be done by pre- and post-multiplying $\boldsymbol{\Sigma}$ by a random diagonal matrix $\boldsymbol{V}$ with diagonal entries sampled from $\{-1, 1\}$, yielding $\boldsymbol{\Sigma}^* = \boldsymbol{V} \boldsymbol{\Sigma} \boldsymbol{V}$.
-
-For all hierarchies in our experiments, we simulate two panel lengths, $T=54$ and $T=304$, reserving the final four observations as an out-of-sample test set.
-In each Monte Carlo replication ($M=500$), we fit univariate ARIMA models (base models) to the training observations using an automatic AICc minimization algorithm from the fabletools package [@O-Hara-Wild2024-we], generating incoherent 1–4-step base forecasts.
-We then reconcile these forecasts under three covariance estimators: the raw sample covariance (mint_sample), the shrinkage estimator (mint_shr), and the NOVELIST estimator (mint_n).
-
-All data generation, covariance estimation, and reconciliation routines were implemented in the ReconCov R package, a package developed solely for this research, and is available under an open‐source license on GitHub [@ReconCov].
-
-```{r load results}
-sim1_1 <- readRDS("./literature/lib/S4-2-1_T50_M500.rds")
-sim1_2 <- readRDS("./literature/lib/S4-2-1_T300_M500.rds")
-
-sim2_1 <- readRDS("./literature/lib/S36-6-1_T50_M500.rds")
-sim2_2 <- readRDS("./literature/lib/S36-6-1_T300_M500.rds")
-
-# sim3_1 <- readRDS("./literature/lib/S100-2-1_T50_M200_run1.rds")
-# sim3_2 <- readRDS("./literature/lib/S100-2-1_T100_M200_run1.rds")
-
-sim_dense_50 <- readRDS("./literature/lib/S100-10-3-1_T50_M200_dense.rds")
-sim_dense_300 <- readRDS("./literature/lib/S100-10-3-1_T300_M200_dense.rds")
-
-sim_sparse_50 <- readRDS("./literature/lib/S100-10-3-1_T50_M500_sparsebwgrp.rds")
-sim_sparse_300 <- readRDS("./literature/lib/S100-10-3-1_T300_M200_sparsebwgrp.rds")
-
-sim3_1 <- sim_dense_50 # using same with next section simulation
-sim3_2 <- sim_dense_300
-```
-
-## Exploring Effects of Hierarchy's Size
-
-In our first set of experiments, we examine how MinT combined with the NOVELIST estimator scales as the hierarchy expands.
-We generate synthetic data from the same VAR(1) framework described earlier, but vary the number of bottom‐level series, $n_b$, across three different structures: the smallest with two groups of two ($n_b = 4$), an intermediate case with six groups of six ($n_b = 36$), and a much larger configuration with two clusters of fifty ($n_b = 100$).
-In the 4‐series hierarchy, each pair collapses into a single level‐1 series, which then sums to the top.
-In the 36‐series case, each block of six forms a level‐1 aggregate, and those six aggregates form the national total.
-
-The 100‐series design employs a deliberately intricate aggregation path to stress‐test reconciliation methods.
-We first sum the one hundred bottom series into ten intermediate series by grouping them in contiguous blocks of ten.
-These ten series are then organised into three level-2 aggregates—four, three, and four series, respectively—before finally summing to a single top node.
-This asymmetric hierarchy creates overlapping correlation patterns: some level‐2 series share bottom‐level groups, while others draw from both, emulating practical scenarios such as regional sales aggregations that span multiple product categories or overlapping territories.
-
-To save space, we illustrate only the six-by-six's VAR(1) and correlation configuration in @fig-settings-6x6; the two-by-two and the one-hundred-series structures appear in Appendix @sec-sim-supplement.
-
-```{r generate plots for A and Sigma}
-MSE1_1 <- transform_sim_MSE(sim1_1$MSE, F) |> 
-  filter(.model != "ols")
-MSE1_2 <- transform_sim_MSE(sim1_2$MSE, F) |> 
-  filter(.model != "ols")
-
-MSE2_1 <- transform_sim_MSE(sim2_1$MSE, F) |> 
-  filter(! .model %in% c("mint_n_hstep", "ols"))
-MSE2_2 <- transform_sim_MSE(sim2_2$MSE, F) |> 
-  filter(! .model %in% c("mint_n_hstep", "ols"))
-
-MSE3_1 <- transform_sim_MSE(sim3_1$MSE, F)
-MSE3_2 <- transform_sim_MSE(sim3_2$MSE, F) |> 
-  filter(.model != "ols")
-
-p1_A <- plot_heatmap(sim1_1$A, T) + 
-  ggtitle("2x2 VAR(1) Coefficients Matrix")
-
-p1_Sigma <- plot_heatmap(cov2cor(sim1_1$Sigma), T) + 
-  ggtitle("2x2 Correlation Matrix")
-
-p2_A <- plot_heatmap(sim2_1$A) +
-  ggtitle("VAR(1) Coefficients Matrix") + theme_void()
-
-p2_Sigma <- plot_heatmap(cov2cor(sim2_1$Sigma), T) +
-  ggtitle("Correlation Matrix") + theme_void()
-
-p3_A <- plot_heatmap(sim3_1$A) +
-  ggtitle("2x50 VAR(1) Coefficients Matrix")
-
-p3_Sigma <- plot_heatmap(cov2cor(sim3_1$Sigma), T) +
-  ggtitle("2x50 Correlation Matrix")
-```
-
-```{r fig settings 6x6}
-#| label: fig-settings-6x6
-#| fig-cap: "Heatmaps of the VAR(1) coefficient matrix and correlation matrix for the 2x50 structure."
-#| fig.height: 4
-#| fig.width: 6
-#| out.width: "85%"
-#| fig.align: "center"
-
-p2_A + p2_Sigma + plot_layout(ncol = 2) & theme(legend.position = "bottom")
-```
-
-```{r}
-plot_relative_improvement <- function(MSE_data) {
-  MSE_data |>
-    filter(
-      .model != "mint_true",
-      h <= 4
-    ) |>
-    group_by(series, h) |>
-    mutate(
-      base_MSE = MSE[.model == "base"]
-    ) |>
-    ungroup() |>
-    group_by(.model, h) |>
-    summarise(
-      MSE = mean(MSE),
-      base_MSE = mean(base_MSE),
-      pct_change = (MSE - base_MSE) / base_MSE * 100
-    ) |>
-    ggplot(aes(x = h, y = pct_change, color = .model)) +
-    geom_line() +
-    labs(
-      x = "Forecast Horizon (h)",
-      y = "Percentage Change (%)"
-    ) +
-    # labs(x = element_blank(), y = element_blank()) +
-    theme_bw()
-}
-```
-
-```{r fig-sim-results-1}
-#| label: fig-sim-results-1
-#| fig-cap: "Relative improvement of the MSE of reconciled forecasts over the base forecasts for the 2x2, 6x6, and 2x50 hierarchical structures, for 1 to 4 steps ahead forecasts, with 2 time series lengths (T = 54 and T = 304)."
-#| fig.width: 8
-#| fig.height: 6
-#| out.width: "100%"
-#| fig.align: "center"
-
-p_MSE1_1 <- MSE1_1 |> 
-  plot_relative_improvement() +
-    labs(title = "2x2 Hierarchy", subtitle = "T = 54")
-p_MSE2_1 <- MSE2_1 |> 
-  plot_relative_improvement() +
-    labs(title = "6x6 Hierarchy")
-p_MSE3_1 <- MSE3_1 |> 
-  plot_relative_improvement() +
-    labs(title = "2x50 Hierarchy")
-
-p_MSE1_2 <- MSE1_2 |> 
-  plot_relative_improvement() +
-    labs(subtitle = "T = 304")
-p_MSE2_2 <- MSE2_2 |> 
-  plot_relative_improvement()
-p_MSE3_2 <- MSE3_2 |> 
-  plot_relative_improvement()
-
-p_MSE1_1 + p_MSE2_1 + p_MSE3_1 + p_MSE1_2 + p_MSE2_2 + p_MSE3_2 +
-  plot_layout(ncol = 3, guides = "collect", axis_titles = "collect") &
-  theme(legend.position = "right")
-```
-
-Figure @fig-sim-results-1 illustrates that as hierarchy size and complexity increase, the relative improvements in MSE over the base forecasts from reconciliation amplify, across all 1- to 4-step horizons.
-While extending the training window (from 50 to 300 in‐sample observations) uniformly improves the accuracy of base ARIMA forecasts, it does not alter the relative improvements of mint_shr or mint_n.
-This is evidence that both high-dimensional estimators robustly handle the "large $p \gg T$” regime.
-By contrast, mint_sample achieves significant accuracy when more data make the sample covariance more reliable.
-
-Despite their theoretical differences, mint_shr and mint_n exhibit nearly identical performance in these canonical settings.
-Yet NOVELIST’s thresholding ability to minimise effects of weak correlations and retain strong ones suggests it may excel in contexts where correlation matrix has small, noisy entries.
-This motivates our subsequent exploration of sparse covariance structures, in which many off-diagonals are truly zero.
-<!--
-## Exploring Effects of Hierarchy's Aggregation
-
-- 2x50 with usual path and weird path
-
-- Finding: Complex aggregation MinT_n performs well
--->
-
-## Exploring the sparsity of the DGP covariance matrix
-
-In our second simulation study, we design a data‐generating process that contrasts “dense” and “sparse” correlation regimes among bottom‐level series, reflecting settings one might encounter in practice.
-We consider the same hierarchical structure of two large groups of 50 bottom series as above.
-Specifically, the two groups would have strong within-group dependencies throughout and either modest between‐group correlations (the dense scenario) or complete independence (the sparse scenario).
-These correlation matrices are depicted in Figure @fig-Sigma-2x50.
-Both scenarios share the same VAR(1) coefficient structure as in our previous simulations; only the innovation covariance changes.
-Such a setup mirrors real‐world contexts where, for example, sales within a product line may exhibit strong co‐movements, while those in a separate line operate nearly independently.
-
-<!-- Rather than collapsing each group into a single aggregate as in previous practice, we impose a more intricate, multi‐level aggregation path to stress‐test reconciliation methods.
-We first form ten intermediate series by summing bottom‐level series in consecutive blocks of ten.
-These ten series are then assigned to three second‐level aggregates--groups of four, three, and four--and finally summed to produce a single top‐level series.
-This asymmetric hierarchy creates overlapping correlation patterns: some level‐2 series share bottom‐level groups, while others draw from both.
--->
-<!-- emulating practical scenarios such as regional sales aggregations that span multiple product categories or overlapping territories. -->
-
-```{r}
-#| label: fig-Sigma-2x50
-#| fig-cap: "Heatmaps of the dense and sparse correlation matrix of the data generating process."
-#| fig.height: 12
-#| fig.width: 8
-#| out.width: "100%"
-#| fig.align: "center"
-p_dense_Sigma <- sim_dense_50$Sigma |>
-  cov2cor() |> 
-  plot_heatmap(T) +
-    labs(title = "Dense")
-
-p_sparse_Sigma <- sim_sparse_50$Sigma |>
-  cov2cor() |> 
-  plot_heatmap(T) +
-    labs(title = "Sparse")
-
-p_dense_Sigma + p_sparse_Sigma + 
-  plot_layout(ncol = 1, guides = "collect", axis_titles = "collect") & 
-  theme(legend.position = "right")
-```
-
-```{r}
-#| label: fig-sim-results-dsp
-#| fig-cap: "Relative improvement of the MSE of reconciled forecasts over the base forecasts for the 2x50 hierarchical structure with dense and sparse DGP's correlation matrix, for 1 to 4 steps ahead forecasts, with 2 time series lengths (T = 54 and T = 304)."
-#| fig.height: 7
-#| fig.width: 8
-#| out.width: "90%"
-#| fig.align: "center"
-p_dense_50 <- sim_dense_50$MSE |>
-  transform_sim_MSE(F) |> 
-  plot_relative_improvement() +
-    labs(title = "Dense DGP Covariance Matrix", subtitle = "T = 54")
-p_sparse_50 <- sim_sparse_50$MSE |>
-  transform_sim_MSE(F) |> 
-  plot_relative_improvement() +
-    labs(title = "Dense DGP Covariance Matrix")
-
-p_dense_300 <- sim_dense_300$MSE |>
-  transform_sim_MSE(F) |> 
-  filter(.model != "ols") |> 
-  plot_relative_improvement() +
-    labs(subtitle = "T = 304")
-p_sparse_300 <- sim_sparse_300$MSE |>
-  transform_sim_MSE(F) |> 
-  plot_relative_improvement()
-
-p_dense_50 + p_sparse_50 + p_dense_300 + p_sparse_300 +
-  plot_layout(ncol = 2, guides = "collect", axis_titles = "collect") &
-  theme(legend.position = "right")
-```
-
-Figure @fig-sim-results-dsp presents out‐of‐sample mean squared error improvements over the base forecasts for each reconciliation strategy under both dense and sparse settings, with two panel lengths--54 and 304 observations--reserving the last four points for testing.
-In both short and long samples, MinT using either the shrinkage estimator (mint_shr) or the NOVELIST estimator (mint_n) delivers pronounced gains over incoherent ARIMA forecasts, particularly at the one‐step horizon where cross‐series correlations most directly inform the forecast adjustments.
-Although the two MinT variants perform almost indistinguishably overall, mint_n edges out mint_shr in the immediate horizon, whereas mint_shr slightly outperforms for longer horizons.
-By contrast, MinT with the raw sample covariance (mint_sample) suffers in small‐sample settings; as expected, its performance improves dramatically with 304 data points, since the sample covariance becomes more reliable with larger $n$.
-This highlights the practical necessity of regularized estimators in high‐dimensional, low‐sample contexts, a situation common in real applications where histories are short relative to the number of series.
-
-Additional designs (varying block sizes, aggregation paths, correlation configurations) also failed to separate NOVELIST from Shrinkage.
-Their nearly identical performance under these synthetic scenarios suggests that our current simulation may not unveil the full advantages of the thresholding estimators.
-This finding motivates our turn to empirical data, where latent structural features and regime shifts, which we will discuss in the next section, may reveal performance differences.
-
-<!--
-## Exploring Grouped Structure
-
-- small vs big
-
-- Finding: NOVELIST performs well in both cases
--->
-
-# Forecasting Australian Domestics Tourism {#sec-empirical}
-
-Domestic tourism flows in Australia exhibit a natural hierarchical and grouped structure, driven both by geography and by purpose of travel.
-At the top of this hierarchy lies the national total, which splits into the seven states and territories.
-Each state is further sub-divided into tourism zones, which in turn break down into 77 regions.
-A complete illustration of this geographic hierarchy appears in Appendix @sec-tourism-supplement.
-Intersecting this geographic hierarchy is a second dimension--travel motive--partitioning tourism flows into four categories: holiday, business, visiting friends and relatives, and other.
-Altogether, this yields a grouped system of 560 series, from the most disaggregated regional-purpose cells up to the full national aggregate.
-@tbl-tourism-structure depicts this structure.
-
-```{r}
-#| tbl-cap: "Hierarchical and grouped structure of Australian domestic tourism flows"
-#| label: tbl-tourism-structure
-
-num_series_geo <- c(1, 7, 27, 77)
-table_structure <- tibble(
-  `Geographical division` = c("Australia", "States", "Zones", "Regions"),
-  `Number of series per geographical division` = num_series_geo,
-  `Number of series per purpose` = 4*num_series_geo,
-  `Total number of series` = 5*num_series_geo
-)
-# add total per row
-table_structure <- rbind(
-  table_structure,
-  c("Total", colSums(table_structure[,-1]))
-)
-
-# width margin smaller (90%)
-table_structure |> 
-  knitr::kable(
-    align = "lccc",
-    booktabs = TRUE
-  )
-```
-
-We quantify tourism demand via "visitor nights", the total number of nights spent by Australians away from home.
-The data is collected via the National Visitor Survey, managed by Tourism Research Australia, using computer assisted telephone interviews from nearly 120,000 Australian residents aged 15 years and over [@tourism-au].
+# Evaluating ex- and include 2016 onwards on Tourism
 
@@
 ```
-
-The data are monthly time series spanning from January 1998 to December 2019, resulting in 264 observations per series, producing a canonical "$n \ll p$" setting which is ideal for evaluating reconciliation approaches that rely on high-dimensional covariance estimation.
-The extreme dimensionality over sample size mirrors many contemporary business problems, for instance, Starbucks drink sales.
-Tourism demand is also economically vital yet highly volatile, with geographical and purpose‑specific patterns create a realistic stress‑test for reconciliation algorithms.
-Consequently, this panel offers both a rich policy case study and a stringent statistical laboratory for comparing reconciliation strategies that exploit cross-series information to improve forecasts when historical data are scarce.
-
-<!-- This dataset is particularly well‑suited to evaluating forecast‑reconciliation methods for three reasons.
-First, tourism demand is economically important and notoriously volatile; the heterogeneous regional and purpose‑specific patterns create a realistic stress‑test for reconciliation algorithms.
-Second, the clear structural shift towards stronger growth after 2016—visible in most aggregates—necessitates techniques capable of handling possible breaks or time‑varying error structures, making the exercise practically relevant.
-Third, the extreme dimensionality relative to sample size mirrors many contemporary business problems (e.g., SKU‑level retail sales), allowing insights to generalise beyond tourism.
-Using this panel therefore provides both a compelling policy context and a stringent statistical laboratory for comparing alternative reconciliation strategies, especially those that rely on regularised covariance estimation or exploit cross‑sectional information to stabilise forecasts when historical data are scant.
--->
-
-@Wickramasuriya2019-xq also argued that modelling spatial autocorrelations would be challenging as in the case of a large collection of time series.
-Reconciliation approaches have the advantage to implicitly model this spatial autocorreltion structure, especially when the MinT method is used.
-
-To assess forecasting performance, we adopt a rolling-window cross-validation scheme.
-Beginning with the first 96 monthly observations (January 1998-December 2005) as the initial training set, we obtain the best-fitted ARIMA model for each of the 560 series via the automatic algorithm in the fabletools package, by minimising AICc [@O-Hara-Wild2024-we].
-The 1- to 12-step-ahead base forecasts are then generated by these ARIMA models.
-These incoherent base forecasts are reconciled via OLS, MinT with shrinkage (mint_shr), and MinT with the NOVELIST estimator (mint_n).
-We then roll the training window forward by one month and refit all models, rebuild reconciliations, and produce another batch of 1- to 12-step-ahead forecasts, repeating until December 2018.
-In total, this results in 156 out-of-sample windows and an equal number of forecast sets for each series.
-
-Note that the number of series is larger than the number of observations (560 compared to 96), hence the sample covariance matrix is not positive definite and will not be considered.
 
@@
 <!-- perhaps via robust trend adjustment or state-space extensions that jointly model coherence and temporal dynamics. -->
-
-<!-- # Supplementary -->
-
-\pagebreak
-
-# Timeline & Milestones {#sec-timeline}
-
-```{r timeline}
-#| echo: false
-#| message: false
-#| warning: false
-library(tibble)
-library(knitr)
-library(kableExtra)
-
-tibble(
-  Period = c(
-    "March",
-    "April", 
-    "May", 
-    "June",
-    "July",
-    "August", 
-    "September",
-    "October"
-  ),
-  Task = c(
-    "Initial setup and understanding the research scope. Preliminary knowledge and Literature review",
-
-    "Develop ReconCov, an R package implementing the Min Trace approach with the shrinkage and NOVELIST estimators, and rolling-window cross-validation process for choosing NOVELIST threshold parameter.
-Exploring simulation framework.",
-
-    "Code functions for producing simulation settings in ReconCov package. Initial simulation for MinT in hierarchical time series. Writing research proposal and presenting research project.",
-
-    "Intensive simulation on multiple different settings, in order to assess the performance and behaviours of NOVELIST under different conditions, and to explore cases where NOVELIST performs different than Shrinkage.",
-
-    "Empirical analysis on forecasting Australian Domestic Tourism data. Explore the real-world data generating process which can be used to refine the current simulation design",
-
-    "Implement the Tourism DGP configuration into the current simulation settings.
-Implementing POET and PC-adjusted NOVELIST estimators.
-Assessing whether prominent principal components structure can be exploited to improve the out-of-sample forecast performance.",
-
-    "Exploring other high-dimensional covariance estimators. More simulations on all explored covariance estimators. Final,  full-fledged assessment on all implemented estimators",
-
-    "Final writing, presenation, submission"
-  ),
-  Deliverable = c(
-    "Initial literature review + paper draft",
-    "ReconCov package: Main codes for estimators + simulation framework + cross-validation",
-    "NOVELIST initial simulation results + research proposal + presentation",
-    "More simulation results on NOVELIST",
-    "Empirical results on forecasting Tourism data",
-    "New simulation design, POET & PC-adjusted NOVELIST simulation results",
-    "Other estimators simulation results and full-fledged covariance estimators assessment",
-    "Presentation and thesis manuscript"
-  )
-) |>
-  kable(
-    format = "latex",
-    booktabs = TRUE,
-    longtable = TRUE,
-    caption = "Project Timeline and Key Milestones",
-    escape = TRUE
-  ) |>
-  column_spec(1, width = "1.9cm") |>
-  column_spec(2, width = "8cm") |>
-  column_spec(3, width = "4.8cm") |>
-  kable_styling(latex_options = "hold_position", font_size = 10)
-```
-
-\pagebreak
-
-# Appendix {#sec-appendix}
-
-## Simulation Settings: Supplementary Figures {#sec-sim-supplement}
-
-```{r fig settings 2x2}
-#| label: fig-settings-2x2
-#| fig-cap: "Heatmaps of the VAR(1) coefficient matrix and correlation matrix for the 2x2 structure."
-#| fig.height: 4
-#| fig.width: 6
-#| out.width: "55%"
-#| fig.align: "center"
-
-p1_A + p1_Sigma + plot_layout(ncol = 2) & theme(legend.position = "bottom")
-```
-
-```{r fig settings 2x50}
-#| label: fig-settings-2x50
-#| fig-cap: "Heatmaps of the VAR(1) coefficient matrix and correlation matrix for the 6×6 structure."
-#| fig.height: 12
-#| fig.width: 8
-#| out.width: "85%"
-#| fig.align: "center"
-
-p3_A + p3_Sigma + plot_layout(ncol = 1) & theme(legend.position = "right")
-```
-
-\pagebreak
-
-## Algorithm: NOVELIST cross-validation for optimal threshold $\delta^*$ {#sec-novelist_cv}
-
-The cross-validation algorithm for NOVELIST is available in the ReconCov package [@ReconCov].
-
-```{=latex}
-\begin{algorithm}
-\caption{Cross-validation procedure}
-\begin{algorithmic}[1]
-
-\State \textbf{Input:} Observations and fitted values $\boldsymbol{y}_t, \hat{\boldsymbol{y}}_t \in \mathbb{R}^n$ for $t = 1,\dots,T$, set of threshold candidates $\Delta$, window size $v$.
-
-\State $\hat{\boldsymbol{e}}_t = \boldsymbol{y}_t - \hat{\boldsymbol{y}}_t$ for $t = 1,\dots,T$
-
-\For{$i = v:T-1$}
-    \State $j = i - v +1$
-    \State $\hat{\boldsymbol{W}}_j = \frac{1}{v} \sum_{t=j}^{i} \hat{\boldsymbol{e}}_{t} \hat{\boldsymbol{e}}_{t}'$
-    \State $\hat{\boldsymbol{D}}_j = \text{diag}(\hat{\boldsymbol{W}}_j)$
-    \State $\hat{\boldsymbol{R}}_j = \hat{\boldsymbol{D}}_j^{-1/2} \hat{\boldsymbol{W}}_j \hat{\boldsymbol{D}}_j^{-1/2}$
-    \For{$\delta \in \Delta$}
-        \State Compute thresholded correlation $\hat{\boldsymbol{R}}_{j,\delta}$ using Equation 5
-        \State Compute $\hat{\lambda}_{j,\delta}$ using Equation 6
-        \State Compute $\hat{\boldsymbol{R}}^{N}_{j,\delta}$ using Equation 4
-        \State $\hat{\boldsymbol{W}}^{N}_{j,\delta} = \hat{\boldsymbol{D}}_j^{1/2} \hat{\boldsymbol{R}}^{N}_{j,\delta} \hat{\boldsymbol{D}}_j^{1/2}$
-        \State $\boldsymbol{G} = (\boldsymbol{S}' \hat{\boldsymbol{W}}^{N^{-1}}_{j,\delta} \boldsymbol{S})^{-1} \boldsymbol{S}' \hat{\boldsymbol{W}}^{N^{-1}}_{j,\delta}$
-        \State Reconciled forecasts $\tilde{\boldsymbol{y}}_{i+1 | \delta} = \boldsymbol{S} \boldsymbol{G} \hat{\boldsymbol{y}}_{i+1}$
-        \State $\tilde{\boldsymbol{e}}_{i+1 | \delta} = \boldsymbol{y}_{i+1} - \tilde{\boldsymbol{y}}_{i+1 | \delta}$
-    \EndFor
-\EndFor
-
-\State $\text{MSE}_{\delta} = \frac{1}{T-v} \sum_{i=v}^{T-1} (\tilde{\boldsymbol{e}}_{i+1 | \delta})^2$ for each $\delta \in \Delta$
-\State $\hat{\delta}^* = \arg\min_{\delta \in \Delta} \text{MSE}_{\delta}$
-\State Compute $\hat{\lambda}^*$ on all training data using $\hat{\delta}^*$
-\State Compute $\hat{\boldsymbol{R}}_1^*$ using $\hat{\delta}^*$ and $\hat{\lambda}^*$ on all training data, using Equation 3
-
-\State \textbf{Output:} Estimate of optimal $\hat{\delta}^*$
-
-\end{algorithmic}
-\end{algorithm}
-```
-
-## Appendix: Australian Domestic Tourism Geographical Hierarchy {#sec-tourism-supplement}
-
-```{r}
-#| tbl-cap: "Geographical divisions of Australia."
-#| label: tbl-tourism-geo
-#| warning: false
-#| message: false
-
-source(paste0(path, "R/table_geo.R"))
-table_geo_labeled <- create_table_geo()
-table_geo_labeled[1,1] <- "Total"
-
-df <- table_geo_labeled %>% 
-  select(Series, Name, Label)
-
-n  <- ceiling(nrow(df) / 2)
-L  <- df[1:n, ]
-R  <- df[(n+1):nrow(df), ]
-
-# pad the shorter side
-if (nrow(R) < nrow(L)) {
-  R <- bind_rows(R, tibble(Series = rep(NA_integer_, n - nrow(R)),
-                           Name = "", Label = ""))
-}
-
-wide <- bind_cols(
-  setNames(L, c("Series","Name","Label")),
-  setNames(R, c("Series","Name","Label"))
-)
-
-kable(
-  wide,
-  align     = c("r","l","l","r","l","l"),
-  col.names = c("Series","Name","Label","Series","Name","Label")
-) |> 
-  kable_styling(latex_options = "hold_position", font_size = 8)
-```
-
-\pagebreak
-
-# References

### 2025-09-23 literature/paper_draft.qmd
revise tourism result discussion

@@
 #       latex-env: algorithm
+link-citations: true
 execute:
@@
 library(stringr)
+library(forcats)
 
@@
 
-# Abstract {-}
+# Abstract {.unnumbered}
 
@@
 <!-- -   Talk about description of the hierarchical time series data set (e.g., economic, financial, or synthetic data). -->
+
 <!-- -   Characteristics such as dimensionality, frequency, and hierarchical structure. -->
@@
 <!-- -   Design different cases of simulation studies and real-data experiments. -->
+
 <!-- -   Metrics: Forecast accuracy (e.g., RMSE, MAE), reconciliation error reduction, and matrix stability... -->
+
 <!-- -   Visualisation... -->
@@
 -->
+
 <!-- emulating practical scenarios such as regional sales aggregations that span multiple product categories or overlapping territories. -->
@@
 
+```{r}
+tourism_plots <- readRDS("literature/tourism_plots.rds")
+```
+
 Domestic tourism flows in Australia exhibit a natural hierarchical and grouped structure, driven both by geography and by purpose of travel.
@@
 
-```{r}
-library(stringr)
-desired_order <- c(
-  "Australia", "Australia_by_Purpose",
-  "State",     "State_by_Purpose",
-  "Zone",      "Zone_by_Purpose",
-  "Region",    "Region_by_Purpose"
-)
-```
-
-```{r fig tourism results}
-#| label: fig-tourism-results
-#| fig-cap: "Relative improvement of the MSE of reconciled forecasts over the base forecasts for the Australian domestic tourism flows"
-#| fig.width: 8
-#| fig.height: 10
-#| out.width: "100%"
-
-tourism_MSE <- readRDS("./literature/lib/tourism_MSE.rds")
-
-# % improv plot by level
-tourism_MSE |>
-  group_by(series, h) |>
-  mutate(
-    base_MSE = MSE[.model == "base"]
-  ) |>
-  ungroup() |>
-  group_by(.model, level, h) |>
-  summarise(
-    MSE = mean(MSE),
-    base_MSE = mean(base_MSE),
-    pct_change = (MSE - base_MSE) / base_MSE * 100
-  ) |>
-  filter(.model != "mint_sample") |>
-  # Reorder the levels according to the desired order
-  mutate(
-    level = factor(
-      level,
-      levels = desired_order,
-      labels = str_replace_all(desired_order, "_", " ")  # replace "_" with " "
-    )
-  ) |> 
-  ggplot(aes(x = h, y = pct_change, color = .model)) +
-    geom_line() +
-    labs(x = "Horizon", y = "% improvements") +
-    facet_wrap(~ level, ncol=2, scales = "free_y") +
-    theme_bw() +
-    scale_x_continuous(limits = c(1, 12), expand = c(0, 0)) +
-    theme(legend.position = "bottom", legend.title = element_blank())
-```
-
-The data are monthly time series spanning from January 1998 to December 2019, resulting in 264 observations per series, producing a canonical "$n \ll p$" setting which is ideal for evaluating reconciliation approaches that rely on high-dimensional covariance estimation.
+The data are monthly time series spanning from January 1998 to December 2016, resulting in 228 observations per series, producing a canonical "$n \ll p$" setting which is ideal for evaluating reconciliation approaches that rely on high-dimensional covariance estimation.
 The extreme dimensionality over sample size mirrors many contemporary business problems, for instance, Starbucks drink sales.
 Tourism demand is also economically vital yet highly volatile, with geographical and purpose‑specific patterns create a realistic stress‑test for reconciliation algorithms.
-Consequently, this panel offers both a rich policy case study and a stringent statistical laboratory for comparing reconciliation strategies that exploit cross-series information to improve forecasts when historical data are scarce.
+
+<!-- Consequently, this panel offers both a rich policy case study and a stringent statistical laboratory for comparing reconciliation strategies that exploit cross-series information to improve forecasts when historical data are scarce.
+-->
 
@@
 
-@Wickramasuriya2019-xq also argued that modelling spatial autocorrelations would be challenging as in the case of a large collection of time series.
-Reconciliation approaches have the advantage to implicitly model this spatial autocorreltion structure, especially when the MinT method is used.
+@Wickramasuriya2019-xq also argued that modelling spatial autocorrelations directly from the start would be challenging as in this case of a large collection of time series.
+Post-processing reconciliation approaches have the advantage to implicitly model this spatial autocorrelation structure, especially true for MinT.
 
 To assess forecasting performance, we adopt a rolling-window cross-validation scheme.
-Beginning with the first 96 monthly observations (January 1998-December 2005) as the initial training set, we obtain the best-fitted ARIMA model for each of the 560 series via the automatic algorithm in the fabletools package, by minimising AICc [@O-Hara-Wild2024-we].
-The 1- to 12-step-ahead base forecasts are then generated by these ARIMA models.
-These incoherent base forecasts are reconciled via OLS, MinT with shrinkage (mint_shr), and MinT with the NOVELIST estimator (mint_n).
-We then roll the training window forward by one month and refit all models, rebuild reconciliations, and produce another batch of 1- to 12-step-ahead forecasts, repeating until December 2018.
-In total, this results in 156 out-of-sample windows and an equal number of forecast sets for each series.
+Beginning with the first 120 monthly observations (January 1998-December 2005) as the initial training set, we obtain the best-fitted ARIMA model for each of the 560 series via the automatic algorithm by minimising AICc from @Hyndman2008-rf.
+The 1- to 12-step-ahead base forecasts are then generated by these ARIMA models, and then reconciled using multiple approaches.
+We then roll the training window forward by one month and refit all models, rebuild reconciliations, and produce another batch of 1- to 12-step-ahead forecasts, repeating until the training set reaches December 2015.
+In total, this results in 133 out-of-sample windows.
+
+<!-- 
+Should MinT-sample be included in the paper?
 
 Note that the number of series is larger than the number of observations (560 compared to 96), hence the sample covariance matrix is not positive definite and will not be considered.
-
-The results are presented in @fig-tourism-results, which shows the relative improvement of the MSE of reconciled forecasts over the base forecasts for the Australian domestic tourism flows.
-The results are grouped by levels of aggregation.
-MinT variants show improvement over the base ARIMA forecasts for middle to lower aggregation levels.
-We hardly differentiate the point accuracy performance between mint_shr and mint_n in these level, except at the most disaggregated, where mint_shr slightly edges out mint_n.
-
-```{r fig STL tourism}
-#| label: fig-STL-tourism
-#| fig-cap: "Monthly domestic overnight tourism nights in Australia by state, with STL decomposition to extract trend component"
+-->
+
+```{r}
+#| label: fig-tourism-MSE-line
+#| fig.height: 6
 #| fig.width: 8
-#| fig.height: 6
-#| out.width: "100%"
-#| fig.align: "center"
-
-visnights_full <- readRDS("literature/lib/visnights_full.rds")
-
-visnights_full |> 
-  filter(
-    is_aggregated(Purpose) & is_aggregated(Zone) & is_aggregated(Region),
-    !is_aggregated(State)
-  ) |> 
-  model(
-    STL(Nights)
-  ) |> 
-  components() |> 
-  pivot_longer(
-    cols = c(Nights, trend, season_year, remainder, season_adjust),
-    names_to = "component",
-    values_to = "value"
-  ) |>
-  filter(
-    component == c("Nights", "trend"),
-    year(Month) >= 2000
-  ) |>
-  as_tibble() |>
-  mutate(component = recode(component,                     # give them nicer names
-                            Nights = "Observed Nights",
-                            trend  = "STL Decomposed Trend")) |> 
-  ggplot(aes(Month, value, color = as.character(State))) +
-    geom_line() +
-    facet_wrap(~ component, ncol=1, scales = "free_y") +
-    labs(
-      x = "Month",
-      y = "Visitor Nights (in thousands)",
-      color = "State"
+#| fig-cap: "Percentage relative improvement in the mean squared error (MSE) of different reconciled forecasts over the base forecasts for the Australian domestic tourism data, for 1 to 12 steps ahead forecasts.
+The positive entries indicate an decrease in MSE."
+tourism_plots$MSE_line + 
+  labs(title = NULL, subtitle = NULL) + 
+  theme(
+    legend.position = "bottom", 
+    legend.direction = "vertical", 
+    legend.title.position = "left"
   ) + 
-    theme_bw()
-```
-
-Surprisingly, however, MinT underperforms OLS at higher aggregation level and even the base forecasts at country-wise.
-This contradicts with the canonical results from @Wickramasuriya2019-xq, prompting further investigation.
-Applying STL decomposition by @Bandara2022-og to state-aggregated observations reveals a structural change around 2016: four large states (New South Wales, Victoria, Queensland, Western Australia) develop a modest upward trend in visitor nights, as shown in @fig-STL-tourism.
-The trend is likely due to the increasing popularity of tourism in these places, as well as the impact of various factors such as economic growth, population growth, and changes in consumer preferences.
+  scale_x_continuous(breaks = c(1:12), minor_breaks = NULL)
+```
 
 ```{r}
-#| label: fig-tourism-results-ex2016
-#| fig-cap: "Relative improvement of the MSE of reconciled forecasts over the base forecasts for the Australian domestic tourism flows, excluding the period after 2016"
+#| label: fig-tourism-MSE-heat
+#| fig.height: 5
+#| fig.width: 7
+#| out.width: "90%"
+#| fig-cap: "Heatmap of relative improvement in the mean squared error (MSE) of different reconciled forecasts over the base forecasts for the Australian domestic tourism data, for 1 to 12 steps ahead forecasts.
+The values are scaled to the range of 0 to 100 for better visualisation, with darker colors indicating greater improvement and best performance is noted by a star."
+
+tourism_plots$MSE_heat + 
+  labs(title = NULL, subtitle = NULL) + 
+  theme(legend.title = element_blank(), legend.position = "bottom") + 
+  scale_x_continuous(breaks = c(1:12), minor_breaks = NULL)
+```
+
+@fig-tourism-MSE-line and @fig-tourism-MSE-heat show that reconciliation is beneficial for point forecasts: across horizons $h=1,2,\dots,12$, all reconciled methods improve MSE over incoherent base ARIMA on average.
+Among vanilla MinT variants, the shrinkage estimator (*MinT-S*) marginally outperforms NOVELIST (*MinT-N*) at most horizons, though the gap is small.
+Adjusting for a single dominant factor via PC decomposition tightens performance further: both *MinT-S(PC1)* and *MinT-N(PC1)* beat their unadjusted counterparts, with *MinT-N(PC1)* narrowly ahead.
+Adding more than one PC brings no additional benefit, likely due to injecting estimation noise from weaker components.
+Variants that modify the multi-step covariance, either via scaled-variance or direct h-step residual covariances, underperform standard MinT, suggesting that extra estimation at horizon $h>1$ is not rewarded in this setting.
+
+```{r}
+#| label: fig-tourism-probscores
+#| fig.height: 12
+#| fig.width: 12
+#| fig-cap: "Percentage relative improvement in the Winkler score at 80% and 95% nominal coverage, CRPS, and Energy score of multiple reconciled forecasts over the base forecasts for the Australian domestic tourism data, for 1-step-ahead forecasts.
+The positive entries indicate an decrease in the probabilistic scores."
+
+# a 2x2 showing 4 plots showing the 1-step-ahead % improv in Winkler score 80, 95, CRPS, and Energy
+
+tourism_plots$W80 + labs(title = "Winkler score 80%", subtitle = NULL) + 
+  tourism_plots$W95 + labs(title = "Winkler score 95%", subtitle = NULL) +
+  tourism_plots$CRPS + labs(title = "CRPS", subtitle = NULL) +
+  tourism_plots$Energy + labs(title = "Energy score", subtitle = NULL) +
+  plot_layout(ncol = 2, guides = "collect", axis_titles = "collect") & 
+  theme(legend.position = "bottom", legend.title = element_blank())
+
+```
+
+Turning to probabilistic forecasts (1-step-ahead forecasts), @fig-tourism-probscores shows that *MinT-N* consistently outperforms *MinT-S* across CRPS and both Winkler scores, and PC1-adjustment again yields the best improvements.
+Results are aligned across probabilistic scoring rules, but the 95% Winkler reveals a notable pattern, as all MinT variants lose to the base while *OLS* performs best.
+
+\color{red}
+This points to tail-calibration sensitivity at high nominal coverage.
+The reconciliation that helps central accuracy can degrade extreme-tail interval performance when uncertainty is misallocated across the hierarchy.
+\color{black}
+
+In the multivariate evaluation, the Energy score places *OLS* close to the PC1-adjusted MinT methods, indicating that simple coherent projection might be able to capture a substantial share of cross-series dependence benefits even without sophisticated covariance regularisation.
+
+The summary radar graph in @fig-tourism-radar consolidates these findings: among the selected MinT models by probabilistic criteria, *MinT-N(PC1)* clearly leads across CRPS, W80, W95, and Energy, extending the improvements seen in the single-metric panels.
+
+Taken together, the evidence supports the use of reconciliation for both point and probabilistic forecasts in this high-dimensional setting.
+For point forecasts, MinT with shrinkage is a solid default; for probabilistic forecasts, NOVELIST performs more reliably.
+PC adjustment with a single dominant factor consistently enhances performance and should be considered when a dominant latent factor is evident.
+More complex adjustments, such as multiple PCs or horizon-specific covariances, add estimation noise without clear benefits and can be omitted for parsimony.
+
+```{r}
+#| label: fig-tourism-radar
+#| fig.height: 8
 #| fig.width: 8
-#| fig.height: 10
-#| out.width: "100%"
-
-tourism_MSE_ex2016 <- readRDS("literature/lib/tourism_MSE_ex2016.rds")
-
-# % improv plot by level
-tourism_MSE_ex2016 |>
-  group_by(series, h) |>
-  mutate(
-    base_MSE = MSE[.model == "base"]
-  ) |>
-  ungroup() |>
-  group_by(.model, level, h) |>
-  summarise(
-    MSE = mean(MSE),
-    base_MSE = mean(base_MSE),
-    pct_change = (MSE - base_MSE) / base_MSE * 100
-  ) |>
-  filter(.model != "mint_sample") |>
-  # Reorder the levels according to the desired order
-  mutate(
-    level = factor(
-      level,
-      levels = desired_order,
-      labels = str_replace_all(desired_order, "_", " ")  # replace "_" with " "
-    )
-  ) |>
-  ggplot(aes(x = h, y = pct_change, color = .model)) +
-    geom_line() +
-    labs(x = "Horizon", y = "% improvements") +
-    facet_wrap(~ level, ncol=2, scales = "free_y") +
-    theme_bw() +
-    scale_x_continuous(limits = c(1, 12), expand = c(0, 0)) +
-    theme(legend.position = "bottom", legend.title = element_blank())
-```
-
-Neither ARIMA nor reconciliation strategies capture this shift, degrading accuracy at the most aggregated level.
-From the results in @fig-tourism-results-ex2016, excluding the post-2016 period (i.e., restricting the final training window to end on December 2015) restores reconciliation’s performance: both MinT methods now clearly outperform OLS and the base forecasts at all levels.
-Additionally, we notice that mint_n increasingly outperforms mint_shr at higher levels (Australia, State, and Australia by Purposes levels).
-This is an advantage of NOVELIST estimator.
-At higher levels most series are combinations of many bottom cells, their cross-series base forecast error correlations are strong and genuinely useful.
-NOVELIST keeps these strong links at those higher levels, while its thresholding feature minimises the effects of weak, noisy ones.
-
-The empirical analysis on real, hierarchical Australian domestic tourism shreds lights on the strengths of the NOVELIST estimator--advantages that were not fully apparent in previous simulation studies.
-Going forward, we plan to embed this structure into our data-generating processes and to pinpoint contexts where NOVELIST’s thresholding yields gains beyond those of Shrinkage.
-Moreover, our findings underscore the need to accommodate structural changes--the post-2016 Australian domestic tourism demand shifts we encountered.
-
-<!-- perhaps via robust trend adjustment or state-space extensions that jointly model coherence and temporal dynamics. -->
+#| out.width: "90%"
+#| fig-cap: "Radar plot of relative improvements in probabilistic scores (Winkler 80, Winkler 95, CRPS, Energy) over the base forecasts.
+The scores are scaled to a range of 0 to 100, with larger values indicating better performance.
+Only the top 5 models are shown."
+
+# a radar plot of 5 top models based on probabilistic scores (Winkler 80, 95, CRPS, Energy) scaled to 0-100 
+par(mfrow=c(1,1), mar=c(0,0,0,0))
+tourism_plots$radar
+```
 
@@
 
+<!-- 
 \pagebreak
@@
 ```
+-->
 

### 2025-10-05 literature/archived.qmd
BIG COMMIT: revised for milestone 5/6
all the graphs are in images PNG. Need to insert code for final submission

@@
 <!-- perhaps via robust trend adjustment or state-space extensions that jointly model coherence and temporal dynamics. -->
+
+<!-- ## POET
+
+The POET (Principal Orthogonal complEment Thresholding) estimator, proposed by @Fan2013-jz, is another "sparse" + "non-sparse" covariance estimator.
+It takes the latent factors directly into its construction, and is appealing when there are common drivers in the time series within the hierarchy, as we saw in the Australian tourism example.
+
+The POET method starts by decomposing the correlation matrix $\hat{\boldsymbol{R}}_1$ into a prominent principle components part (low-rank) and a orthogonal complement part $\hat{\boldsymbol{R}}_{1, K}$ (the correlation matrix after removing the first $K$ principal components).
+Then it applies thresholding to $\hat{\boldsymbol{R}}_{1, K}$.
+The POET estimator is given by:
+
+$$
+\hat{\boldsymbol{R}}^{K}_{1} = \sum_{k = 1}^K \hat{\gamma}_k \hat{\boldsymbol{\xi}}_k \hat{\boldsymbol{\xi}}_k' + T(\hat{\boldsymbol{R}}_{1, K})
+$$
+
+where $\hat{\gamma}_k$ and $\hat{\boldsymbol{\xi}}_k$ are the $k$th largest eigenvalue and the corresponding eigenvector of the sample covariance matrix, respectively, and $T(.)$ is the thresholding function, which can be either soft-thresholding, hard-thresholding, or others.
+-->
+
+<!-- ## PC-adjusted NOVELIST
+
+This approach is best of both worlds, leveraging the strengths of both NOVELIST and POET.
+The PC-adjusted (Principal-Component-adjusted) NOVELIST overcomes the shortcomings of the current shrinkage estimator, taking prominent PCs into account while also offers extra flexibility.
+The idea is to apply the NOVELIST estimator to the orthogonal complement part $\hat{\boldsymbol{R}}_{1, K}$, and then add the principal components part back.
+The PC-adjusted NOVELIST estimator is formulated as:
+
+$$
+\hat{\boldsymbol{R}}^{N, K}_{1} = \sum_{k = 1}^K \hat{\gamma}_k \hat{\boldsymbol{\xi}}_k \hat{\boldsymbol{\xi}}_k' + \hat{\boldsymbol{R}}^{N}_{1, K}
+\; ,
+$$
+
+where $\hat{\boldsymbol{R}}^{N}_{1, K}$ is the NOVELIST estimator applied to the orthogonal complement part $\hat{\boldsymbol{R}}_{1, K}$.
+Similar to the NOVELIST estimator, $\hat{\boldsymbol{R}}^{N, K}_{1}$ is not guaranteed to be positive definite.
+
+Methods to ensure positive definiteness of the NOVELIST estimator (and its PC-adjusted variant) will be explored and studied in the project.
+@Huang2019-ua proposed to diagonalise the NOVELIST estimator and replace any eigenvalues that fall under a certain small positive threshold by the value of that threshold.
+Alternatively, we can implement the algorithm of @Higham2002-te that computes the nearest positive definite matrix to a given matrix.
+-->
+
+<!--  
+```{r}
+plot_relative_improvement <- function(MSE_data) {
+  MSE_data |>
+    filter(
+      .model != "mint_true",
+      h <= 4
+    ) |>
+    group_by(series, h) |>
+    mutate(
+      base_MSE = MSE[.model == "base"]
+    ) |>
+    ungroup() |>
+    group_by(.model, h) |>
+    summarise(
+      MSE = mean(MSE),
+      base_MSE = mean(base_MSE),
+      pct_change = (MSE - base_MSE) / base_MSE * 100
+    ) |>
+    ggplot(aes(x = h, y = pct_change, color = .model)) +
+    geom_line() +
+    labs(
+      x = "Forecast Horizon (h)",
+      y = "Percentage Change (%)"
+    ) +
+    # labs(x = element_blank(), y = element_blank()) +
+    theme_bw()
+}
+```
+
+```{r fig-sim-results-1}
+#| label: fig-sim-results-1
+#| fig-cap: "Relative improvement of the MSE of reconciled forecasts over the base forecasts for the 2x2, 6x6, and 2x50 hierarchical structures, for 1 to 4 steps ahead forecasts, with 2 time series lengths (T = 54 and T = 304)."
+#| fig.width: 8
+#| fig.height: 6
+#| out.width: "100%"
+#| fig.align: "center"
+
+p_MSE1_1 <- MSE1_1 |> 
+  plot_relative_improvement() +
+    labs(title = "2x2 Hierarchy", subtitle = "T = 54")
+p_MSE2_1 <- MSE2_1 |> 
+  plot_relative_improvement() +
+    labs(title = "6x6 Hierarchy")
+p_MSE3_1 <- MSE3_1 |> 
+  plot_relative_improvement() +
+    labs(title = "2x50 Hierarchy")
+
+p_MSE1_2 <- MSE1_2 |> 
+  plot_relative_improvement() +
+    labs(subtitle = "T = 304")
+p_MSE2_2 <- MSE2_2 |> 
+  plot_relative_improvement()
+p_MSE3_2 <- MSE3_2 |> 
+  plot_relative_improvement()
+
+p_MSE1_1 + p_MSE2_1 + p_MSE3_1 + p_MSE1_2 + p_MSE2_2 + p_MSE3_2 +
+  plot_layout(ncol = 3, guides = "collect", axis_titles = "collect") &
+  theme(legend.position = "right")
+```
+
+-->
+
+Figure @fig-sim-results-1 illustrates that as hierarchy size and complexity increase, the relative improvements in MSE over the base forecasts from reconciliation amplify, across all 1- to 4-step horizons.
+While extending the training window (from 50 to 300 in‐sample observations) uniformly improves the accuracy of base ARIMA forecasts, it does not alter the relative improvements of mint_shr or mint_n.
+This is evidence that both high-dimensional estimators robustly handle the "large $p \gg T$” regime.
+By contrast, mint_sample achieves significant accuracy when more data make the sample covariance more reliable.
+
+Despite their theoretical differences, mint_shr and mint_n exhibit nearly identical performance in these canonical settings.
+Yet NOVELIST’s thresholding ability to minimise effects of weak correlations and retain strong ones suggests it may excel in contexts where correlation matrix has small, noisy entries.
+This motivates our subsequent exploration of sparse covariance structures, in which many off-diagonals are truly zero.

### 2025-10-05 literature/paper_draft.qmd
BIG COMMIT: revised for milestone 5/6
all the graphs are in images PNG. Need to insert code for final submission

@@
         \usepackage{array}
+        \usepackage{float}
     mathspec: true
@@
 
-path <- "D:/Github/Recon_Honours_Thesis/"
+path <- "D:/Github/Recon_Honours_Thesis"
 knitr::opts_knit$set(root.dir = path)
@@
 The current practice is to use the shrinkage estimator (often shrinking toward a diagonal matrix), but it lacks flexibility and might not fully utilise the prominent latent structure presented.
-In this project, we aim to assess the forecasting performance of MinT when different covariance estimators are used, namely NOVELIST (NOVEL Integration of the Sample and Thresholded Covariance), POET (Principal Orthogonal complEment Thresholding), and others.
+In this project, we aim to assess the forecasting performance of MinT when different covariance estimators are used, namely NOVELIST (NOVEL Integration of the Sample and Thresholded Covariance), PC-adjusted estimators (taking the latent factors directly into its construction), and others.
 
@@
 
-## Hierarchical tructure {#sec-hierarchy}
+## Hierarchical structure {#sec-hierarchy}
+
+\color{red}
+to be revised with probabilistic framework
+\color{black}
 
@@
 
-![A 2-level hierarchical tree structure](D:/Github/Recon_Honours_Thesis/research_proposal/figs/hierarchical_structure.png){#fig-hierarchy width="40%"}
-
-![A 2-level grouped structure, which can be considered as the union of two hierarchical trees with common top and bottom level series](D:/Github/Recon_Honours_Thesis/research_proposal/figs/grouped_structure.png){#fig-grouped width="90%"}
+![A 2-level hierarchical tree structure](D:/Github/Recon_Honours_Thesis/figs/hierarchical_structure.png){#fig-hierarchy width="40%"}
+
+![A 2-level grouped structure, which can be considered as the union of two hierarchical trees with common top and bottom level series](D:/Github/Recon_Honours_Thesis/figs/grouped_structure.png){#fig-grouped width="90%"}
 
@@
 
-The MinT solution hinges on a reliable, positive-definite estimate of $\boldsymbol{W}_h$, which is challenging to estimate in high-dimensional setting.
+The MinT solution hinges on a reliable, positive-definite estimate of $\boldsymbol{W}_h$, which is often not available and challenging to estimate in high-dimensional setting.
 The sample covariance matrix is unstable and non-positive-definite when the number of series $n$ is huge and larger than the time dimension $T$.
-To tackle this issue, the original paper @Wickramasuriya2019-xq adopted the diagonal-target shrinkage estimator from @Schafer2005-yw, given by
-
-$$
-\hat{\boldsymbol{W}}^{shr}_{1} = \lambda_D \hat{\boldsymbol{W}}_{1, D} + (1 - \lambda_D) \hat{\boldsymbol{W}}_1 \, ,
+To tackle this issue, the original paper @Wickramasuriya2019-xq assumed a proportionality relationship between $\hat{\boldsymbol{W}}^g_h = k_h g( \hat{\boldsymbol{W}}_1 )$, where $\hat{\boldsymbol{W}}_1$ is the covariance matrix of the in-sample 1-step-ahead base residuals (to approximate $\boldsymbol{W}_1$), $k_h$ is a positive scaling constant (which will be cancelled out in point-forecast reconciliation), and $g(.)$ is a covariance estimator that produces a positive-definite matrix.
+They suggested using the shrinkage estimator with diagonal target from @Schafer2005-yw, given by:
+
+$$
+\hat{\boldsymbol{W}}^{S}_{1} = \lambda_D \hat{\boldsymbol{W}}_{1, D} + (1 - \lambda_D) \hat{\boldsymbol{W}}_1 \, ,
 $$
@@
 
-However, the hierarchical time series data often exhibit a prominent principal components structure, which is not fully taken advantage.
+Despite its relative simplicity and guaranteed positive definiteness, MinT with shrinkage estimator has three main shortcomings.
+
+### Problem 1: Uniform shrinkage {.unnumbered}
+
+The shrinkage estimator shrinks all off-diagonal elements towards zeros with equal weights $\lambda_D$.
+We might prefer to better preserve strong signals, and largely reduce the effects of small, noisy correlations.
+This is especially true in high-dimensional settings, where only a few series to be truly correlated; or due to the aggregation effects, bottom-level series will correlate more strongly with their parents.
+The shrinkage estimator lacks this flexibility.
+
+### Problem 2: Latent factors {.unnumbered}
+
+The hierarchical time series data often exhibit a prominent principal components structure, which is not fully taken advantage.
 Taking an example of the Australian domestic overnight trips data set [@tourism-au], where the national trips are disaggregated into states and territories, and further into regions.
-We then fit ETS models to all series, using the algorithm from Fabletools R package [@O-Hara-Wild2024-we], and compute the one-step-ahead in-sample base forecast error covariance matrix $\hat{\boldsymbol{W}}_1$.
+We then fit ETS models to all series, using the algorithm from Fabletools R package [@O-Hara-Wild2024-we], and compute the one-step-ahead in-sample base forecast residual covariance matrix $\hat{\boldsymbol{W}}_1$.
 The twenty largest eigenvalues of the covariance matrix are plotted in @fig-eigen.
@@
 
-![Twenty largest eigenvalues of one-step-ahead in-sample base forecast error covariance, Australian domestic overnight trips](D:/Github/Recon_Honours_Thesis/research_proposal/figs/eigen_tourism.png){#fig-eigen width="70%"}
-
-Additionally, the shrinkage estimator shrinks all off-diagonal elements towards zeros with equal weights $\lambda_D$.
-We might prefer to better preserve strong signals, and largely reduce the effects of small, noisy correlations.
-In the next sections, we will explore several options that take these two issues into account.
+![Twenty largest eigenvalues of one-step-ahead in-sample base forecast error covariance, Australian domestic overnight trips](D:/Github/Recon_Honours_Thesis/figs/eigen_tourism.png){#fig-eigen width="70%"}
+
+### Problem 3: Proportional scaling for h-step-ahead {.unnumbered}
+
+The proportionality relationship $\hat{\boldsymbol{W}}^g_h = k_h g( \hat{\boldsymbol{W}}_1 )$ might not hold in practice.
+The covariance structure of h-step-ahead base forecast errors can be different from that of 1-step-ahead ones.
+
+In the next sections, we will explore several options that tackle these issues one by one.
 
@@
 NOVELIST offers more flexibility than the shrinkage estimator, which is useful when we believe that only a few variables are truly correlated.
-However, it does not guarantee to be positive definite.
 
@@
 
-In this setting, $\hat{\boldsymbol{R}}_{1,\delta}$ is the thresholded correlation matrix, where each element is regularised by:
+where, $\hat{\boldsymbol{R}}_{1,\delta}$ is the thresholded correlation matrix, where each element is regularised by:
 
@@
 where $\delta \in [0,1]$ is the threshold parameter.
-For a given threshold $\delta$, @Huang2019-ua derived an analytical expression for the optimal shrinkage intensity parameter $\lambda(\delta)$ using Ledoit-Wolf's lemma [@Ledoit2003-qv], following similar logic to @Schafer2005-yw.
+For a given threshold $\delta$, @Huang2019-ua derived an analytical expression for the optimal shrinkage intensity parameter $\lambda(\delta)$ using Ledoit-Wolf's lemma \[@Ledoit2003-qv\], following similar logic to @Schafer2005-yw.
 It can be computed as:
@@
 
-On the other hand, the optimal threshold $\delta^*$ does not have a closed-form solution, and is typically obtained by executing a rolling-window cross-validation procedure.
-The idea is to find the threshold $\hat{\delta^*}$, with the corresponding $\hat{\lambda}^*$ and $\hat{\boldsymbol{R}}^{N}_{1}(\hat{\delta^*}, \hat{\lambda}^*)$, that minimises the average out-of-sample 1-step-ahead reconciled forecast mean squared error over all windows.
-The formal algorithm is given in the @sec-novelist_cv Appendix.
-Although it is not required to fit forecasting models multiple times, the cross-validation procedure is still computationally expensive as it computes the NOVELIST estimator and perform reconciliation for each threshold value.
-
-We also tested out minimising 1- to h-step-ahead MSE in the cross-validation procedure.
+On the other hand, the optimal threshold $\hat \delta$ does not have a closed-form solution, and is typically obtained by executing a rolling-window cross-validation procedure.
+The idea is to find the threshold $\delta^*$, with the corresponding $\lambda^*$ and $\hat{\boldsymbol{R}}^{N}_{1}(\delta^*, \lambda^*)$, that minimises the average out-of-sample 1-step-ahead reconciled forecast mean squared error over all windows.
+The formal algorithm is given in @sec-novelist_cv.
+
+We also tested out minimising 1- to h-step-ahead overall MSE in the cross-validation procedure.
 Surprisingly, it returns almost the same best threshold parameter is in the 1-step-ahead case above.
 Note that when $\delta \in \bigl[ \text{max}_{i \neq j}|\hat{r}_{1,ij}|, \; 1 \bigr]$, the NOVELIST estimator collapses to the shrinkage estimator, and when $\delta = 0$, it becomes the sample covariance matrix.
-
-<!-- ## POET
-
-The POET (Principal Orthogonal complEment Thresholding) estimator, proposed by @Fan2013-jz, is another "sparse" + "non-sparse" covariance estimator.
-It takes the latent factors directly into its construction, and is appealing when there are common drivers in the time series within the hierarchy, as we saw in the Australian tourism example.
-
-The POET method starts by decomposing the correlation matrix $\hat{\boldsymbol{R}}_1$ into a prominent principle components part (low-rank) and a orthogonal complement part $\hat{\boldsymbol{R}}_{1, K}$ (the correlation matrix after removing the first $K$ principal components).
-Then it applies thresholding to $\hat{\boldsymbol{R}}_{1, K}$.
-The POET estimator is given by:
-
-$$
-\hat{\boldsymbol{R}}^{K}_{1} = \sum_{k = 1}^K \hat{\gamma}_k \hat{\boldsymbol{\xi}}_k \hat{\boldsymbol{\xi}}_k' + T(\hat{\boldsymbol{R}}_{1, K})
-$$
-
-where $\hat{\gamma}_k$ and $\hat{\boldsymbol{\xi}}_k$ are the $k$th largest eigenvalue and the corresponding eigenvector of the sample covariance matrix, respectively, and $T(.)$ is the thresholding function, which can be either soft-thresholding, hard-thresholding, or others.
--->
-
-<!-- ## PC-adjusted NOVELIST
-
-This approach is best of both worlds, leveraging the strengths of both NOVELIST and POET.
-The PC-adjusted (Principal-Component-adjusted) NOVELIST overcomes the shortcomings of the current shrinkage estimator, taking prominent PCs into account while also offers extra flexibility.
-The idea is to apply the NOVELIST estimator to the orthogonal complement part $\hat{\boldsymbol{R}}_{1, K}$, and then add the principal components part back.
-The PC-adjusted NOVELIST estimator is formulated as:
-
-$$
-\hat{\boldsymbol{R}}^{N, K}_{1} = \sum_{k = 1}^K \hat{\gamma}_k \hat{\boldsymbol{\xi}}_k \hat{\boldsymbol{\xi}}_k' + \hat{\boldsymbol{R}}^{N}_{1, K}
-\; ,
-$$
-
-where $\hat{\boldsymbol{R}}^{N}_{1, K}$ is the NOVELIST estimator applied to the orthogonal complement part $\hat{\boldsymbol{R}}_{1, K}$.
-Similar to the NOVELIST estimator, $\hat{\boldsymbol{R}}^{N, K}_{1}$ is not guaranteed to be positive definite.
-
-Methods to ensure positive definiteness of the NOVELIST estimator (and its PC-adjusted variant) will be explored and studied in the project.
-@Huang2019-ua proposed to diagonalise the NOVELIST estimator and replace any eigenvalues that fall under a certain small positive threshold by the value of that threshold.
-Alternatively, we can implement the algorithm of @Higham2002-te that computes the nearest positive definite matrix to a given matrix.
--->
+An additional concern is that the estimator does not guarantee to be positive definite, but we can use @Higham2002-te algorithm to compute the nearest positive definite matrix if needed.
+
+### NOVELIST cross-validation algorithm {#sec-novelist_cv}
+
+It is only required to fit the base models once on the whole training data $\{\boldsymbol{y}_t\}_{t=1}^T$, and obtain the in-sample fitted values $\{\hat{\boldsymbol{y}}_t\}_{t=1}^T$.
+
+```{=latex}
+\begin{algorithm}[H]
+\caption{Cross-validation procedure}
+\begin{algorithmic}[1]
+
+\State \textbf{Input:} Observations and fitted values $\boldsymbol{y}_t, \hat{\boldsymbol{y}}_t \in \mathbb{R}^n$ for $t = 1,\dots,T$, set of threshold candidates $\Delta$, window size $v$.
+
+\State $\hat{\boldsymbol{e}}_t = \boldsymbol{y}_t - \hat{\boldsymbol{y}}_t$ for $t = 1,\dots,T$
+
+\For{$i = v:T-1$}
+    \State $j = i - v +1$
+    \State $\hat{\boldsymbol{W}}_j = \frac{1}{v} \sum_{t=j}^{i} \hat{\boldsymbol{e}}_{t} \hat{\boldsymbol{e}}_{t}'$
+    \State $\hat{\boldsymbol{D}}_j = \text{diag}(\hat{\boldsymbol{W}}_j)$
+    \State $\hat{\boldsymbol{R}}_j = \hat{\boldsymbol{D}}_j^{-1/2} \hat{\boldsymbol{W}}_j \hat{\boldsymbol{D}}_j^{-1/2}$
+    \For{$\delta \in \Delta$}
+        \State Compute thresholded correlation $\hat{\boldsymbol{R}}_{j,\delta}$ using Equation 5
+        \State Compute $\hat{\lambda}_{j,\delta}$ using Equation 6
+        \State Compute $\hat{\boldsymbol{R}}^{N}_{j,\delta}$ using Equation 4
+        \State $\hat{\boldsymbol{W}}^{N}_{j,\delta} = \hat{\boldsymbol{D}}_j^{1/2} \hat{\boldsymbol{R}}^{N}_{j,\delta} \hat{\boldsymbol{D}}_j^{1/2}$
+        \State $\boldsymbol{G} = (\boldsymbol{S}' \hat{\boldsymbol{W}}^{N^{-1}}_{j,\delta} \boldsymbol{S})^{-1} \boldsymbol{S}' \hat{\boldsymbol{W}}^{N^{-1}}_{j,\delta}$
+        \State Reconciled forecasts $\tilde{\boldsymbol{y}}_{i+1 | \delta} = \boldsymbol{S} \boldsymbol{G} \hat{\boldsymbol{y}}_{i+1}$
+        \State $\tilde{\boldsymbol{e}}_{i+1 | \delta} = \boldsymbol{y}_{i+1} - \tilde{\boldsymbol{y}}_{i+1 | \delta}$
+    \EndFor
+\EndFor
+
+\State $\text{MSE}_{\delta} = \frac{1}{T-v} \sum_{i=v}^{T-1} (\tilde{\boldsymbol{e}}_{i+1 | \delta})^2$ for each $\delta \in \Delta$
+\State $\hat{\delta}^* = \arg\min_{\delta \in \Delta} \text{MSE}_{\delta}$
+\State Compute $\hat{\lambda}^*$ on all training data using $\hat{\delta}^*$
+\State Compute $\hat{\boldsymbol{R}}_1^*$ using $\hat{\delta}^*$ and $\hat{\lambda}^*$ on all training data, using Equation 3
+
+\State \textbf{Output:} Estimate of optimal $\hat{\delta}^*$
+
+\end{algorithmic}
+\end{algorithm}
+```
+
+The cross-validation algorithm for NOVELIST is available in the ReconCov package [@ReconCov].
 
@@
 
-This method takes the latent factors directly into its construction, and is appealing when there are common drivers in the time series within the hierarchy, as we saw in the Australian tourism example.
-It starts by decomposing the correlation matrix $\hat{\boldsymbol{R}}_1$ into a prominent principle components part (low-rank) and a orthogonal complement part $\hat{\boldsymbol{C}}^K_{1}$ (the correlation matrix after removing the first $K$ principal components).
-Then we can apply either shrinkage or NOVELIST estimator to $\hat{\boldsymbol{R}}_{1, K}$:
-
-$$
-\hat{\boldsymbol{R}}^{g, K}_{1} = \sum_{k = 1}^K \hat{\gamma}_k \hat{\boldsymbol{\xi}}_k \hat{\boldsymbol{\xi}}_k' + g(\hat{\boldsymbol{C}}^K_{1})
+To utilise the latent factors structure for better shrinkage, the PC-adjusted method takes the latent factors directly into its construction, and is appealing when there are common drivers in the time series within the hierarchy, as we saw in the Australian tourism example.
+It starts by decomposing the covariance matrix $\hat{\boldsymbol{W}}_1$ into a prominent principle components part (low-rank) and a orthogonal complement part $\hat{\boldsymbol{W}}^K_{1}$ (the correlation matrix after removing the first $K$ principal components).
+Then we can apply either shrinkage or NOVELIST estimator to $\hat{\boldsymbol{W}}^K_{1}$:
+
+$$
+\hat{\boldsymbol{R}}^{g, K}_{1} = \sum_{k = 1}^K \hat{\gamma}_k \hat{\boldsymbol{\xi}}_k \hat{\boldsymbol{\xi}}_k' + g(\hat{\boldsymbol{W}}^K_{1})
 $$
@@
 
-Apply shrinkage or NOVELIST to the correlation matrix of 1-step-ahead base forecast errors, then scale it back to covariance matrix of h-step-ahead base forecast errors using the h-step-ahead standard deviations:
-
-$$
-\hat{\boldsymbol{W}}^{sv}_{h} = \boldsymbol{D}^{1/2}_h g(\hat{\boldsymbol{R}}_1) \boldsymbol{D}^{1/2}_h,
-$$
-
-where $\boldsymbol{D}_h = \text{diag}(\hat{\sigma}_{1,h}, \hat{\sigma}_{2,h}, \ldots, \hat{\sigma}_{n_b,h})$, and $\hat{\sigma}_{i,h}$ is the standard deviation of the $i$-th series' h-step-ahead base forecast errors.
-The asterisk $*$ indicates either shrinkage or NOVELIST estimator.
+To address the potential issue of the proportionality relationship $\hat{\boldsymbol{W}}^g_h = k_h g( \hat{\boldsymbol{W}}_1 )$ not holding in practice, we can relax the assumption by allowing the variances to scale differently.
+This is done by scaling the $1$-step-ahead correlation matrix back to $h$-step-ahead covariance matrix using the h-step-ahead standard deviations.
+The scaled variance estimator is given by:
+
+$$
+\hat{\boldsymbol{W}}^{g, sv}_{h} = \boldsymbol{D}^{1/2}_h g(\hat{\boldsymbol{R}}_1) \boldsymbol{D}^{1/2}_h,
+$$
+
+where $\boldsymbol{D}_h = \text{diag}(\hat{\sigma}^2_{1,h}, \ldots, \hat{\sigma}^2_{n,h})$, and $\hat{\sigma}^2_{i,h}$ is the variance of the $i$-th series' h-step-ahead base forecast errors.
+Similarly, $g(.)$ is either the shrinkage or NOVELIST estimator.
+
+If NOVELIST is used to produce $h$-step-ahead reconciled forecasts, the cross-validation procedure is slightly modified to evaluate the out-of-sample reconciled forecast MSE using $\hat{\boldsymbol{W}}^{N, sv}_{h}$ instead of $\hat{\boldsymbol{W}}^{N}_{1}$.
 
@@
 
-Construct the sample covariance matrix directly from h-step-ahead base forecast errors, then apply either shrinkage or NOVELIST estimator:
+Another approach is not to rely on the assumption of proportionality at all, and directly estimate the covariance matrix from the h-step-ahead base forecast errors:
 
@@
 
-where $\hat{\boldsymbol{R}}_h$ is the sample correlation matrix of h-step-ahead base forecast errors, and $\circ$ denotes the Hadamard product (element-wise product).
-The asterisk $*$ indicates either shrinkage or NOVELIST estimator.
-
-# Simulation Design {#sec-simulation}
-
-<!-- -   Talk about description of the hierarchical time series data set (e.g., economic, financial, or synthetic data). -->
-
-<!-- -   Characteristics such as dimensionality, frequency, and hierarchical structure. -->
-
-<!-- **Experimental Design** -->
-
-<!-- -   Design different cases of simulation studies and real-data experiments. -->
-
-<!-- -   Metrics: Forecast accuracy (e.g., RMSE, MAE), reconciliation error reduction, and matrix stability... -->
-
-<!-- -   Visualisation... -->
+where $\hat{\boldsymbol{W}}_h$ is the covariance matrix of in-sample h-step-ahead base forecast residuals, and $g(.)$ is either the shrinkage or NOVELIST estimator.
+Similar to the scaled variance approach, if NOVELIST is used, the cross-validation procedure is modified to compute $\hat{\boldsymbol{W}}^{N}_{h}$ and evaluate the out-of-sample reconciled forecast MSE using it.
+
+## Summary of MinT with Covariance Estimators {-}
+
+| Covariance estimators used              | Abbreviation       | 
+|-----------------------------------------|--------------------|
+| Shrinkage                               | MinT-S             |
+| NOVELIST                                | MinT-N             |
+| PC-adjusted Shrinkage with *k* PCs      | MinT-S(PC*k*)      |
+| PC-adjusted NOVELIST with *k* PCs       | MinT-N(PC*k*)      |
+| Scaled Variance Shrinkage               | MinT-S(sv)         |
+| Scaled Variance NOVELIST                | MinT-N(sv)         |
+| Constructed from h-step-ahead Shrinkage | MinT-S(hcov)       |
+| Constructed from h-step-ahead NOVELIST  | MinT-N(hcov)       |
+
+# Simulation {#sec-simulation}
+
+## General Design Framework {-}
 
@@
 For all hierarchies in our experiments, we simulate two panel lengths, $T=54$ and $T=304$, reserving the final four observations as an out-of-sample test set.
-In each Monte Carlo replication ($M=500$), we fit univariate ARIMA models (base models) to the training observations using an automatic AICc minimization algorithm from the fabletools package [@O-Hara-Wild2024-we], generating incoherent 1–4-step base forecasts.
+In each Monte Carlo replication ($M=500$), we fit univariate ARIMA models (base models) to the training observations using an automatic AICc minimization algorithm from @Hyndman2008-rf, implemented in the *fabletools* package [@O-Hara-Wild2024-we], generating incoherent 1–4-step base forecasts.
 We then reconcile these forecasts under three covariance estimators: the raw sample covariance (mint_sample), the shrinkage estimator (mint_shr), and the NOVELIST estimator (mint_n).
 
-All data generation, covariance estimation, and reconciliation routines were implemented in the ReconCov R package, a package developed solely for this research, and is available under an open‐source license on GitHub [@ReconCov].
+All data generation, covariance estimation, and reconciliation routines were implemented in the ReconCov R package and is available under an open‐source license on GitHub [@ReconCov].
 
 ```{r load results}
+path_sim <- "D:/Github/Recon_Honours_Analysis/sim/thesis_sim/"
+
 sim1_1 <- readRDS("./literature/lib/S4-2-1_T50_M500.rds")
@@
 
-# sim3_1 <- readRDS("./literature/lib/S100-2-1_T50_M200_run1.rds")
-# sim3_2 <- readRDS("./literature/lib/S100-2-1_T100_M200_run1.rds")
-
 sim_dense_50 <- readRDS("./literature/lib/S100-10-3-1_T50_M200_dense.rds")
@@
 
+## Scoring Metric {-}
+
+\color{red}
+to be included...
+\color{black}
+
 ## Exploring Effects of Hierarchy's Size
 
-In our first set of experiments, we examine how MinT combined with the NOVELIST estimator scales as the hierarchy expands.
-We generate synthetic data from the same VAR(1) framework described earlier, but vary the number of bottom‐level series, $n_b$, across three different structures: the smallest with two groups of two ($n_b = 4$), an intermediate case with six groups of six ($n_b = 36$), and a much larger configuration with two clusters of fifty ($n_b = 100$).
-In the 4‐series hierarchy, each pair collapses into a single level‐1 series, which then sums to the top.
+In our first set of experiments, we examine how MinT combined with the different estimators perform as the hierarchy expands.
+We generate synthetic data from the same VAR(1) framework described earlier, but vary the number of bottom‐level series, $n_b$, across two structures: a small structure with six groups of six bottom series ($n_b = 6$x$6 = 36$), and a much larger configuration with two groups of fifty ($n_b = 2$x$50 = 100$).
+
 In the 36‐series case, each block of six forms a level‐1 aggregate, and those six aggregates form the national total.
-
 The 100‐series design employs a deliberately intricate aggregation path to stress‐test reconciliation methods.
@@
 This asymmetric hierarchy creates overlapping correlation patterns: some level‐2 series share bottom‐level groups, while others draw from both, emulating practical scenarios such as regional sales aggregations that span multiple product categories or overlapping territories.
-
-To save space, we illustrate only the six-by-six's VAR(1) and correlation configuration in @fig-settings-6x6; the two-by-two and the one-hundred-series structures appear in Appendix @sec-sim-supplement.
+The aggregation paths for both structures are illustrated in @fig-sim-structures.
+
+![Aggregation structures used in the simulation experiments: 6 groups of 6 (left) and 2 groups of 50 (right)](D:/Github/Recon_Honours_Thesis/figs/sim_structures.png){#fig-sim-structures}
+
+The VAR(1) and correlation configurations for the 6 by 6 case and 2 by 50 case are illustrated in @fig-settings-6x6 and @fig-settings-2x50, respectively.
+The block diagonal structure of the VAR(1) coefficient matrices $\boldsymbol{A}$ reflects the grouping of series, and the correlation matrices $\boldsymbol{\Sigma}$ show higher correlations among series within the same group.
 
@@
 
-```{r fig settings 6x6}
+```{r settings 6x6}
 #| label: fig-settings-6x6
-#| fig-cap: "Heatmaps of the VAR(1) coefficient matrix and correlation matrix for the 2x50 structure."
+#| fig-cap: "Heatmaps of the VAR(1) coefficient matrix and correlation matrix for the 6 by 6 structure."
 #| fig.height: 4
@@
 
-```{r}
-plot_relative_improvement <- function(MSE_data) {
-  MSE_data |>
-    filter(
-      .model != "mint_true",
-      h <= 4
-    ) |>
-    group_by(series, h) |>
-    mutate(
-      base_MSE = MSE[.model == "base"]
-    ) |>
-    ungroup() |>
-    group_by(.model, h) |>
-    summarise(
-      MSE = mean(MSE),
-      base_MSE = mean(base_MSE),
-      pct_change = (MSE - base_MSE) / base_MSE * 100
-    ) |>
-    ggplot(aes(x = h, y = pct_change, color = .model)) +
-    geom_line() +
-    labs(
-      x = "Forecast Horizon (h)",
-      y = "Percentage Change (%)"
-    ) +
-    # labs(x = element_blank(), y = element_blank()) +
-    theme_bw()
-}
-```
-
-```{r fig-sim-results-1}
-#| label: fig-sim-results-1
-#| fig-cap: "Relative improvement of the MSE of reconciled forecasts over the base forecasts for the 2x2, 6x6, and 2x50 hierarchical structures, for 1 to 4 steps ahead forecasts, with 2 time series lengths (T = 54 and T = 304)."
+```{r settings 2x50}
+#| label: fig-settings-2x50
+#| fig-cap: "Heatmaps of the VAR(1) coefficient matrix and correlation matrix for the 2 by 50 structure."
+#| fig.height: 12
 #| fig.width: 8
-#| fig.height: 6
 #| out.width: "100%"
@@
 
-p_MSE1_1 <- MSE1_1 |> 
-  plot_relative_improvement() +
-    labs(title = "2x2 Hierarchy", subtitle = "T = 54")
-p_MSE2_1 <- MSE2_1 |> 
-  plot_relative_improvement() +
-    labs(title = "6x6 Hierarchy")
-p_MSE3_1 <- MSE3_1 |> 
-  plot_relative_improvement() +
-    labs(title = "2x50 Hierarchy")
-
-p_MSE1_2 <- MSE1_2 |> 
-  plot_relative_improvement() +
-    labs(subtitle = "T = 304")
-p_MSE2_2 <- MSE2_2 |> 
-  plot_relative_improvement()
-p_MSE3_2 <- MSE3_2 |> 
-  plot_relative_improvement()
-
-p_MSE1_1 + p_MSE2_1 + p_MSE3_1 + p_MSE1_2 + p_MSE2_2 + p_MSE3_2 +
-  plot_layout(ncol = 3, guides = "collect", axis_titles = "collect") &
-  theme(legend.position = "right")
-```
-
-Figure @fig-sim-results-1 illustrates that as hierarchy size and complexity increase, the relative improvements in MSE over the base forecasts from reconciliation amplify, across all 1- to 4-step horizons.
-While extending the training window (from 50 to 300 in‐sample observations) uniformly improves the accuracy of base ARIMA forecasts, it does not alter the relative improvements of mint_shr or mint_n.
-This is evidence that both high-dimensional estimators robustly handle the "large $p \gg T$” regime.
-By contrast, mint_sample achieves significant accuracy when more data make the sample covariance more reliable.
-
-Despite their theoretical differences, mint_shr and mint_n exhibit nearly identical performance in these canonical settings.
-Yet NOVELIST’s thresholding ability to minimise effects of weak correlations and retain strong ones suggests it may excel in contexts where correlation matrix has small, noisy entries.
-This motivates our subsequent exploration of sparse covariance structures, in which many off-diagonals are truly zero.
-<!--
-## Exploring Effects of Hierarchy's Aggregation
-
-- 2x50 with usual path and weird path
-
-- Finding: Complex aggregation MinT_n performs well
--->
+p3_A + p3_Sigma + plot_layout(ncol = 1) & theme(legend.position = "right")
+```
+
+![Percentage relative improvement in MSE of reconciled forecasts over the base forecasts in the 6 by 6 case (top row) and the 2 by 50 case (bottom row), T=50 (left column) and T=300 (right column), for 1- to 4-step-ahead forecasts.
+The positive (negative) entries indicate a decrease (increase) in MSE relative to base.](D:/Github/Recon_Honours_Thesis/figs/sim_MSE.png){#fig-sim-results-1}
+
+Figure @fig-sim-results-1 illustrates the relative improvements in mean squared error (MSE) of reconciled forecasts over the incoherent base forecasts, across two structures and time series lengths.
+The MinT with shrinkage (*MinT-S*) and its variants are colored in mint green, while MinT with NOVELIST (*MinT-N*) are in purple.
+The concrete lines represent the vanilla *MinT-S* and *MinT-N*; the dashed lines with dot points denote the PC-adjusted variants (e.g.
+*MinT-S(PC1), *MinT-N(PC2)*); and the dotted or dashed-dotted lines indicate the scaled variance and h-step-ahead residuals versions (e.g.
+*MinT-S(SV), MinT-N(hcov)*).
+
+The first key observation is that methods with shrinkage slightly outperform those with NOVELIST across all scenarios, despite the differences being small.
+Second, the PC-adjusted variants (using one and two principal components) do not yield improvements over the vanilla versions.
+This is expected since the synthetic data generating process does not simulate from strong latent factors.
+Third, the scaled variance and h-step-ahead residuals approaches do not enhance performance as the forecast horizon increases, suggesting that the proportionality assumption may not be severely violated in this VAR(1) setup.
+Lastly, as the hierarchy expands from 36 to 100 series, forecasts from NOVELIST methods become worse relative to base forecasts when we move to $2$-step-ahead and beyond.
+When looking into each MCMC replication, we find that there are instances where the NOVELIST estimator collapses to the shrinkage estimator due to a large optimal threshold $\hat{\delta}^*$ being selected in the cross-validation step.
+
+![Percentage relative improvement in Energy score in both hierarchies and time dimensions, for 1-step-ahead forecasts.
+The positive entries indicate a decrease in Energy score relative to base.](D:/Github/Recon_Honours_Thesis/figs/sim_energy.png){#fig-sim-results-2}
+
+Moving on to probabilistic forecasts, we evaluate the performance of reconciliation methods using the energy score [@Gneiting2011-km], a proper scoring rule for multivariate distributions.
+Figure @fig-sim-results-2 presents the percentage relative improvement in energy score for 1-step-ahead forecasts.
+Across both hierarchies and time dimensions, the MinT methods consistently outperform the base forecasts.
+Meanwhile, the differences among *MinT-S* and *MinT-N* variants are insignificant, except for the 6 by 6 case with 50 observations, where shrinkage has a slight edge.
+The PC-adjusted variants again degrade performance as we add more principal components.
+The scaled variance and h-step-ahead residuals approaches are not available since we only evaluate 1-step-ahead covariance estimates.
 
@@
 
-In our second simulation study, we design a data‐generating process that contrasts “dense” and “sparse” correlation regimes among bottom‐level series, reflecting settings one might encounter in practice.
-We consider the same hierarchical structure of two large groups of 50 bottom series as above.
-Specifically, the two groups would have strong within-group dependencies throughout and either modest between‐group correlations (the dense scenario) or complete independence (the sparse scenario).
-These correlation matrices are depicted in Figure @fig-Sigma-2x50.
-Both scenarios share the same VAR(1) coefficient structure as in our previous simulations; only the innovation covariance changes.
-Such a setup mirrors real‐world contexts where, for example, sales within a product line may exhibit strong co‐movements, while those in a separate line operate nearly independently.
-
-<!-- Rather than collapsing each group into a single aggregate as in previous practice, we impose a more intricate, multi‐level aggregation path to stress‐test reconciliation methods.
-We first form ten intermediate series by summing bottom‐level series in consecutive blocks of ten.
-These ten series are then assigned to three second‐level aggregates--groups of four, three, and four--and finally summed to produce a single top‐level series.
-This asymmetric hierarchy creates overlapping correlation patterns: some level‐2 series share bottom‐level groups, while others draw from both.
--->
-
-<!-- emulating practical scenarios such as regional sales aggregations that span multiple product categories or overlapping territories. -->
+\color{red}
+to be revised for final draft
+\color{black}
 
@@
 
+In our second simulation study, we design a data‐generating process that contrasts “dense” and “sparse” correlation regimes among bottom‐level series, reflecting settings one might encounter in practice.
+We consider the same hierarchical structure of two large groups of 50 bottom series as above.
+Specifically, the two groups would have strong within-group dependencies throughout and either modest between‐group correlations (the dense scenario) or complete independence (the sparse scenario).
+These correlation matrices are depicted in Figure @fig-Sigma-2x50.
+Both scenarios share the same VAR(1) coefficient structure as in our previous simulations; only the innovation covariance changes.
+Such a setup mirrors real‐world contexts where, for example, sales within a product line may exhibit strong co‐movements, while those in a separate line operate nearly independently.
+
+<!-- Rather than collapsing each group into a single aggregate as in previous practice, we impose a more intricate, multi‐level aggregation path to stress‐test reconciliation methods.
+We first form ten intermediate series by summing bottom‐level series in consecutive blocks of ten.
+These ten series are then assigned to three second‐level aggregates--groups of four, three, and four--and finally summed to produce a single top‐level series.
+This asymmetric hierarchy creates overlapping correlation patterns: some level‐2 series share bottom‐level groups, while others draw from both.
+-->
+
+<!-- emulating practical scenarios such as regional sales aggregations that span multiple product categories or overlapping territories. -->
+
 Figure @fig-sim-results-dsp presents out‐of‐sample mean squared error improvements over the base forecasts for each reconciliation strategy under both dense and sparse settings, with two panel lengths--54 and 304 observations--reserving the last four points for testing.
@@
 
-<!--
-## Exploring Grouped Structure
-
-- small vs big
-
-- Finding: NOVELIST performs well in both cases
--->
-
 # Forecasting Australian Domestics Tourism {#sec-empirical}
@@
 ```{r}
-tourism_plots <- readRDS("literature/tourism_plots.rds")
+tourism_plots <- readRDS("D:/Github/Recon_Honours_Thesis/literature/tourism_plots.rds")
 ```
@@
 To assess forecasting performance, we adopt a rolling-window cross-validation scheme.
-Beginning with the first 120 monthly observations (January 1998-December 2005) as the initial training set, we obtain the best-fitted ARIMA model for each of the 560 series via the automatic algorithm by minimising AICc from @Hyndman2008-rf.
+Beginning with the first 120 monthly observations (January 1998-December 2005) as the initial training set, we obtain the best-fitted ARIMA model for each of the 560 series via the automatic algorithm by minimising AICc from @Hyndman2008-rf, implemented in the *fabletools* package [@O-Hara-Wild2024-we].
 The 1- to 12-step-ahead base forecasts are then generated by these ARIMA models, and then reconciled using multiple approaches.
 We then roll the training window forward by one month and refit all models, rebuild reconciliations, and produce another batch of 1- to 12-step-ahead forecasts, repeating until the training set reaches December 2015.
-In total, this results in 133 out-of-sample windows.
-
+In total, this results in 97 out-of-sample windows.
+
+```{=html}
 <!-- 
@@
 -->
+```
 
@@
 
+<!-- 
 ```{r}
@@
 ```
+-->
+
+![Percentage relative improvement in the Winkler score at 80% and 95% nominal coverage, CRPS, and Energy score of multiple reconciled forecasts over the base forecasts for the Australian domestic tourism data, for 1-step-ahead forecasts.
+The positive (negative) entries indicate a decrease (increase) in the probabilistic scores relative to base.](D:/Github/Recon_Honours_Thesis/figs/tourism_probscores.png){#fig-tourism-probscores}
 
@@
 Results are aligned across probabilistic scoring rules, but the 95% Winkler reveals a notable pattern, as all MinT variants lose to the base while *OLS* performs best.
-
-\color{red}
-This points to tail-calibration sensitivity at high nominal coverage.
-The reconciliation that helps central accuracy can degrade extreme-tail interval performance when uncertainty is misallocated across the hierarchy.
-\color{black}
-
 In the multivariate evaluation, the Energy score places *OLS* close to the PC1-adjusted MinT methods, indicating that simple coherent projection might be able to capture a substantial share of cross-series dependence benefits even without sophisticated covariance regularisation.
+
+![Percentage relative improvement in Winkler score at 95% nominal coverage by level (left), and empirical coverage of 95% prediction intervals by level (right)](D:/Github/Recon_Honours_Thesis/figs/tourism_coverage_95.png){#fig-tourism-coverage-95-level}
+
+To further dissect the 95% Winkler score results, @fig-tourism-coverage-95-level breaks down performance by hierarchical level, and examines empirical coverage of the 95% prediction intervals.
+The left panel shows that *MinT-S* underperforms the base at all levels, especially in the higher aggregated levels.
+This is due to the overly shrunk variances from the shrinkage estimator, leading to narrow prediction intervals and thus high Winkler penalties when observations fall outside.
+The *MinT-N* method has relatively good coverage and improve the Winkler score at bottom levels, but still underperforms at higher levels.
+The PC-adjusted variants seem to strike a better balance, improving overall coverage and relative Winkler scores over the base.
 
@@
 
-## Simulation Settings: Supplementary Figures {#sec-sim-supplement}
-
-```{r fig settings 2x2}
-#| label: fig-settings-2x2
-#| fig-cap: "Heatmaps of the VAR(1) coefficient matrix and correlation matrix for the 2x2 structure."
-#| fig.height: 4
-#| fig.width: 6
-#| out.width: "55%"
-#| fig.align: "center"
-
-p1_A + p1_Sigma + plot_layout(ncol = 2) & theme(legend.position = "bottom")
-```
-
-```{r fig settings 2x50}
-#| label: fig-settings-2x50
-#| fig-cap: "Heatmaps of the VAR(1) coefficient matrix and correlation matrix for the 6×6 structure."
-#| fig.height: 12
-#| fig.width: 8
-#| out.width: "85%"
-#| fig.align: "center"
-
-p3_A + p3_Sigma + plot_layout(ncol = 1) & theme(legend.position = "right")
-```
-
-\pagebreak
-
-## Algorithm: NOVELIST cross-validation for optimal threshold $\delta^*$ {#sec-novelist_cv}
-
-The cross-validation algorithm for NOVELIST is available in the ReconCov package [@ReconCov].
-
-```{=latex}
-\begin{algorithm}
-\caption{Cross-validation procedure}
-\begin{algorithmic}[1]
-
-\State \textbf{Input:} Observations and fitted values $\boldsymbol{y}_t, \hat{\boldsymbol{y}}_t \in \mathbb{R}^n$ for $t = 1,\dots,T$, set of threshold candidates $\Delta$, window size $v$.
-
-\State $\hat{\boldsymbol{e}}_t = \boldsymbol{y}_t - \hat{\boldsymbol{y}}_t$ for $t = 1,\dots,T$
-
-\For{$i = v:T-1$}
-    \State $j = i - v +1$
-    \State $\hat{\boldsymbol{W}}_j = \frac{1}{v} \sum_{t=j}^{i} \hat{\boldsymbol{e}}_{t} \hat{\boldsymbol{e}}_{t}'$
-    \State $\hat{\boldsymbol{D}}_j = \text{diag}(\hat{\boldsymbol{W}}_j)$
-    \State $\hat{\boldsymbol{R}}_j = \hat{\boldsymbol{D}}_j^{-1/2} \hat{\boldsymbol{W}}_j \hat{\boldsymbol{D}}_j^{-1/2}$
-    \For{$\delta \in \Delta$}
-        \State Compute thresholded correlation $\hat{\boldsymbol{R}}_{j,\delta}$ using Equation 5
-        \State Compute $\hat{\lambda}_{j,\delta}$ using Equation 6
-        \State Compute $\hat{\boldsymbol{R}}^{N}_{j,\delta}$ using Equation 4
-        \State $\hat{\boldsymbol{W}}^{N}_{j,\delta} = \hat{\boldsymbol{D}}_j^{1/2} \hat{\boldsymbol{R}}^{N}_{j,\delta} \hat{\boldsymbol{D}}_j^{1/2}$
-        \State $\boldsymbol{G} = (\boldsymbol{S}' \hat{\boldsymbol{W}}^{N^{-1}}_{j,\delta} \boldsymbol{S})^{-1} \boldsymbol{S}' \hat{\boldsymbol{W}}^{N^{-1}}_{j,\delta}$
-        \State Reconciled forecasts $\tilde{\boldsymbol{y}}_{i+1 | \delta} = \boldsymbol{S} \boldsymbol{G} \hat{\boldsymbol{y}}_{i+1}$
-        \State $\tilde{\boldsymbol{e}}_{i+1 | \delta} = \boldsymbol{y}_{i+1} - \tilde{\boldsymbol{y}}_{i+1 | \delta}$
-    \EndFor
-\EndFor
-
-\State $\text{MSE}_{\delta} = \frac{1}{T-v} \sum_{i=v}^{T-1} (\tilde{\boldsymbol{e}}_{i+1 | \delta})^2$ for each $\delta \in \Delta$
-\State $\hat{\delta}^* = \arg\min_{\delta \in \Delta} \text{MSE}_{\delta}$
-\State Compute $\hat{\lambda}^*$ on all training data using $\hat{\delta}^*$
-\State Compute $\hat{\boldsymbol{R}}_1^*$ using $\hat{\delta}^*$ and $\hat{\lambda}^*$ on all training data, using Equation 3
-
-\State \textbf{Output:} Estimate of optimal $\hat{\delta}^*$
-
-\end{algorithmic}
-\end{algorithm}
-```
-
 ## Appendix: Australian Domestic Tourism Geographical Hierarchy {#sec-tourism-supplement}

### 2025-10-05 literature/temp.qmd
BIG COMMIT: revised for milestone 5/6
all the graphs are in images PNG. Need to insert code for final submission

--- 2025-09-23 literature/paper_draft.qmd
@@
 ---
-title: "Enhancing Forecast Reconciliation: A Study of Alternative Covariance Estimators"
-author: 
-  - Vincent Su
-  - Shanika Wickramasuriya (supv.)
-  - George Athanasopoulos (supv.)
+title: ""
 format: 
@@
     number-sections: true
-# crossref:
-#   custom:
-#     - kind: float
-#       key: alg
-#       reference-prefix: Algorithm
-#       caption-prefix: Algorithm
-#       # for PDF output only, tell LaTeX to use the `algorithm` env:
-#       latex-env: algorithm
 link-citations: true
 execute:
-  cache: true
+  cache: false
   echo: false
@@
 ---
-
-<!-- # Notation {-} -->
-
-<!-- -   Scalar $y_t$ -->
-
-<!-- -   Vector $\boldsymbol{y}_t$ -->
-
-<!-- -   Matrix $\boldsymbol{S}$ -->
-
-<!-- -   Covariance matrix of in-sample 1-step-ahead base forecast errors $\hat{\boldsymbol{W}}_1$ -->
-
-<!-- -   Its shrinkage estimator with diagonal target $\hat{\boldsymbol{W}}^{shr}_{1, D}$ -->
-
-<!-- -   Its NOVELIST estimator $\hat{\boldsymbol{W}}^{shr}_{1, thr}$ -->
 
@@
 ```
-
-# Abstract {.unnumbered}
-
-<!-- [This is pasted from the Project Description]{style="color:red;"} -->
-
-A collection of time series connected via a set of linear constraints is known as hierarchical time series.
-Forecasting these series without respecting the hierarchical nature of the data can lead to incoherent forecasts across aggregation levels and lower accuracy.
-To mitigate this issue, various forecast reconciliation approaches have been proposed in the literature, where the individual forecasts are adjusted to satisfy the aggregation constraints.
-Among these, **MinT** (Minimum Trace) is widely used, however, it requires a good estimate of the covariance matrix of the base forecast errors.
-The current practice is to use the shrinkage estimator (often shrinking toward a diagonal matrix), but it lacks flexibility and might not fully utilise the prominent latent structure presented.
-In this project, we aim to assess the forecasting performance of MinT when different covariance estimators are used, namely NOVELIST (NOVEL Integration of the Sample and Thresholded Covariance), POET (Principal Orthogonal complEment Thresholding), and others.
-
-# Introduction
-
-<!-- ############  Intro  ############ -->
-
-<!--
-Para 1
-- Introduce hierarchical time series and its importance in various applications.
--->
-
-In time series forecasting, aggregation occurs in a variety of settings.
-While a formal definition of hierarchical time series can be found in @sec-hierarchy, we can think of Starbucks sales data as an illustrative example.
-Starbucks operates in many countries, and each country has multiple cities where they have outlets.
-The sales data is *structured hierarchically*: the top level is the total sales across all countries, followed by national sales for each country, and then individual sales for each outlet in a city.
-As a result, there are over 40,000 individual outlet sales to forecast, plus additional series at higher levels of aggregation such as city and country.
-The hierarchy can be even more complex if we consider the sales of different kinds of drinks (e.g., coffees, teas, refreshers) at each aggregation level.
-
-Forecasting data from such hierarchical structures also arises in many other decision-making contexts, from supply chains [@Seaman2022-bb; @Angam2025-od] and energy planning [@Di-Modica2021-ad], to macroeconomics [@El-Gemayel2022-hx; @Li2019-vv] and tourism analysis [@Athanasopoulos2009-lp].
-Stakeholders in these settings need forecasts at several aggregation levels to allocate resources and manage risk.
-The impact of methods for forecasting hierarchical time series has not been limited to academia, with industry also showing a strong interest.
-Many companies and organisations have adopted these methods in practice, including Amazon, the International Monetary Fund, IBM, SAP, and more [@Athanasopoulos2024-as].
-
-<!-- ##########  Literature Review  ########## -->
-
-<!-- - Talk about the history and evolution of forecasting hierarchical time series, starting from the early heuristic methods to the modern statistical approaches.
-
-  - Single level methods
-  - Optimal combination methods (OLS, WLS)
-  - MinT
-  - Bayesian, Machine learning
-  - Probabilistic methods -->
-
-In practice, when forecasts are produced for all series (often called *base forecasts*), they typically violate the aggregation constraints observed in the data; such forecasts are *incoherent*.
-This can undermine downstream decisions that require internal consistency.
-
-Traditionally, forecasting these hierarchical time series has been done using single-level methods, such as bottom-up, top-down, and middle-out approaches.
-Bottom-up methods involve generating forecasts for the bottom-level series and aggregating them to higher levels.
-Top-down methods start with forecasts for the only top-level series and disaggregate them down.
-Middle-out methods combine both approaches by forecasting middle-level series and then aggregating or disaggregating as needed.
-Despite their simplicity, these methods only anchor forecasts to a single level, implying a large loss of information on the hierarchy's inherent correlation structure.
-Additionally, the most disaggregated series often are very noisy or even intermittent, and the higher-level data might be smoother due to the aggregation.
-Furthermore, as we saw from the Starbucks example considering the sales of different kinds of drinks at each aggregation level -- formally defined as grouped structure in @sec-hierarchy -- the disaggregation becomes more complex since the disaggregation paths are not unique.
-Consequently, these single-level methods tend to give poor results across other levels of the hierarchy.
-
-To overcome these issues, forecast reconciliation was introduced by @Hyndman2011-jv, and later developed by @Van_Erven2015-nx, @Hyndman2016-ic, @Ben-Taieb2019-yx, @Wickramasuriya2019-xq, and others to achieve coherency in point forecasts and enhance accuracy.
-Forecast reconciliation projects a collection of independent base forecasts into a set of coherent forecasts that respect the linear constraints defining a hierarchical or grouped time-series system.
-
-Among the modern reconciliation strategies, the Min Trace (MinT) approach developed by @Wickramasuriya2019-xq is widely used and perform significantly well under right conditions.
-Despite its properties, MinT relies on a good high-dimensional covariance estimator, which is positive-definite.
-However, not many researchers have explored this issue in the current literature, except @Carrara2025-rz, who introduced a new estimator for MinT--Double Shrinkage estimator.
-The current stage of this research piece is fairly early, leaving many potential gaps still not yet to explore.
-As a result, a study of alternative high-dimensional covariance estimators for MinT is timely and desirable.
-
-The remainder of the paper is organised as follows.
-@sec-pre-theory provides the basic theoretical framework for hierarchical time series and forecasting, and Min Trace approach, introducing notations, terminologies, and motivations for alternative estimators.
-@sec-cov-est walks through the main covariance estimators this paper explores, and argues their strengths and weaknesses.
-@sec-simulation covers the simulation design and currently explores the performance of NOVELIST on MinT.
-@sec-empirical shows a real-world application of MinT with NOVELIST, which produces results that did not occured in the simulation settings, suggesting further inspection and analysis.
-
-<!-- Include more details on the gaps for each of these methods, such as bias, high variance? -->
-
-<!-- - Discuss the interests in MinT and how it has become a standard method for forecast reconciliation.
-
-- Discuss the MinT's reliance on a good estimate of the covariance matrix of base forecast errors and other gaps
-
-  - Empirical evidence of MinT under perform
-  - Comparison with other methods
-
-- Explain why this paper focus on exploring alternative covariance estimators for MinT. And is there any paper talk about this.
-
-  - Is better estimate of W_h really lead to better performance?
-
-- Talk about the paper outline and what will be covered in the following sections. -->
-
-<!--
--   **Problem Statement:**
-
-    The sample covariance matrix, although natural, suffers in high-dimensional settings.
-Especially when the number of series $p$ is huge and larger than the time dimension $T$, the sample covariance matrix is non-positive definite (rank T if p\>T).
-
-    The shrinkage estimators come in to tackle this issue.
-The shrinkage estimator with diagonal target (often shrinking toward a diagonal matrix) is proven to produce a guaranteed PD matrix (Schäfer & Strimmer, 2005).
-However, as it shrinks the covariance matrix toward a diagonal one, it does not have flexibility and might neglect the prominent structure presented in the covariance matrix.
-
-    An alternative approach is to perform shrinkage of the sample covariance towards its thresholded version, instead of a diagonal matrix.
-This is the NOVELIST (NOVEL Integration of the Sample and Thresholded covariance estimators) method proposed by Huang & Fryzlewicz (2019).
-They introduced thresholding functions applied only to off-diagonal elements, allowing for more flexibility in the estimation.
-
-    [... can include more estimators ...]{style="color:red;"}
-
--   **Research Aim:**
-
-    This paper assesses the reconciled forecasting performance of MinT approach using various covariance estimators, with a focus on the NOVELIST estimator.
-
--   **Paper Outline:**
-
-    The paper is structured as follows:
-
-    -   A literature review of forecast reconciliation and covariance estimation.
-    -   A description of the methodology, including the NOVELIST estimator and its principal-component-adjusted variant.
-    -   An experimental design using both synthetic and real hierarchical time series.
-    -   Empirical results and discussion.
-    -   Conclusions and suggestions for future work.
--->
-
-<!--
-# Literature Review
-
-## Forecast Reconciliation in Hierarchical and Grouped Time Series
-
-Forecast reconciliation converts a collection of independent base forecasts into a set of coherent forecasts that respect the linear constraints defining a hierarchical or grouped time-series system.
-Early work focused on heuristic single-level strategies, including bottom-up, top-down, and middle-out (...), each of which exploits only part of the information in the hierarchy and can induce bias or high variance.
-
--   Cite the single level approach
--   Athanasopoulos et al., 2024
-
-Hyndman et al.
-(2011) first showed that all single-level methods can be written as $\tilde{y} = SG\hat{y}$, where $S$ is the summing matrix and $G$ is a matrix that maps base forecasts $\hat{y}$ to into the bottom level.
-Treating reconciliation as a GLS regression problem, Hyndman et al.
-(2011) found that it yields a solution for $G$, but the required covariance matrix of reconcilation error is not identifiable in practice (Wickramasuriya et al., 2019).
-
--   (Talk more about how others transform it to OLS, WLS,..)
--   Di Fonzo and Marini (2011)
--   Athanasopoulos et al. (2009)
-
-Wickramasuriya, Athanasopoulos & Hyndman (2019) reframed the problem by taking an optimisation approach rather than the regression.
-They formulated the problem as minimising the variances of all reconciled forecasts, which happens to be equivalent to minimising the trace of the covariance matrix (sum of the diagonal elements).
-This is known as the Minimum Trace (MinT) reconciliation method.
-The MinT solution is given by $G_h = (S'W_h^{-1}S)^{-1}S'W_h^{-1}$, and $W_h$ is the covariance matrix of the h-step-ahead base forecast errors.
-
-The MinT approach is an algebraical generalisation of the GLS, and the OLS and WLS methods are special cases of MinT when $W_h$ is a diagonal or identity matrix, respectively.
-However, the MinT solution hinges on a reliable estimate of the h-step-ahead base forecast error covariance $W_h$.
-In high-dimensional setting, the usual sample covariance matrix is unstable, thus we need alternative covariance estimators.
-
--   Structural Scaling, based only on the struc- ture of the hierarchy (Athanasopoulos et al., 2017)
--   Shrinkage estimators (Schäfer & Strimmer, 2005; Ledoit & Wolf, 2004)
-
-Empirical evaluations have demonstrated that MinT with an appropriate covariance estimate often outperforms earlier methods in both simulation and real data studies.
-
-Wickramasuriya et al.
-(2019) adopt the diagonal-target Ledoit–Wolf approach and show analytically that any scalar multiple $k_h$ from $W_h = k_hW_1$ suffices for point-forecast reconciliation.
-However, simulation and empirical evidence suggest that ignoring cross-series error correlation can leave accuracy untapped, especially in highly-coupled systems (energy, retail SKUs).
-Their subsequent work imposes non-negativity constraints via quadratic programming citeturn0file2 and studies bias-robust variants (MinT-U).
-
-Empirical findings to date
-
-Athanasopoulos et al.
-(2024) synthesise the accumulated evidence: MinT dominates single-level baselines in most studies; WLS or diagonal-shrinkage MinT is the current industry default; but gains plateau when off-diagonals are weak citeturn0file9.
-Only a handful of papers explore richer covariance structures (e.g., block or factor shrinkage), and none benchmark them systematically against NOVELIST-type estimators.
-
-## Covariance Estimation in High Dimensions
-
--   Limitations of the sample covariance matrix.
--   Estimators used by Wickramasuriya et al. (2019).
--   Shrinkage estimators:
-    -   Diagonal shrinkage (e.g., Schäfer & Strimmer, Ledoit & Wolf).
-    -   NOVELIST estimator and its Cross-validation & PC-adjusted variant.
-    -   ...
-
-## Relevance to Forecast Reconciliation
-
--   Discuss how covariance estimation affects MinT performance.
--   Identify research gaps.
--->
-
-# Theoretical Framework {#sec-pre-theory}
-
-## Hierarchical tructure {#sec-hierarchy}
-
-The *hierarchical structure* can be represented as a tree, as shown in @fig-hierarchy.
-The top level of the tree represents the total value of all series, while the lower levels represent the series at different levels of disaggregation.
-These hierarchical structures naturally form *aggregation constraints* (child levels must sum up to parents).
-When there are attributes of interest that are crossed, such as the Starbucks drinks sales at any aggregation level (brand-wise, national, or outlet) is also considered by kinds of drinks (e.g., coffees, refreshers), the structure is described as a grouped time series.
-As illustrated in @fig-grouped, the aggregation or disaggregation paths are not unique.
-
-![A 2-level hierarchical tree structure](D:/Github/Recon_Honours_Thesis/research_proposal/figs/hierarchical_structure.png){#fig-hierarchy width="40%"}
-
-![A 2-level grouped structure, which can be considered as the union of two hierarchical trees with common top and bottom level series](D:/Github/Recon_Honours_Thesis/research_proposal/figs/grouped_structure.png){#fig-grouped width="90%"}
-
-For simplicity, we refer to both of these structures as hierarchical structure, we will distinguish between them if and when it is necessary.
-Taken together, a collection of time series organised in a hierarchy and subject to aggregation constraints is refered as *hierarchical time series*.
-All hierarchical time series can be represented using matrix algebra:
-
-$$
-\boldsymbol{y}_t = \boldsymbol{S} \boldsymbol{b}_t,
-$$
-
-where $\boldsymbol{S}$ is a summing matrix of order $n \times n_b$ which aggregates the bottom-level series $\boldsymbol{b}_t \in \mathbb{R}^{n_b}$ to the series at aggregation levels above.
-The vector $\boldsymbol{y}_t \in \mathbb{R}^n$ contains all observations at time $t$.
-The summing matrix $\boldsymbol{S}$ for the tree structure in @fig-hierarchy is:
-
-$$
-\boldsymbol{S} = 
-\left[
-\begin{array}{cccc}
-1 & 1 & 1 & 1 \\
-1 & 1 & 0 & 0 \\
-0 & 0 & 1 & 1 \\
-\multicolumn{4}{c}{\boldsymbol{I_4}}
-\end{array}
-\right].
-$$
-
-Assume we produce $h$-step-ahead base forecasts $\hat{\boldsymbol{b}}_{t+h|t}$ for the bottom-level series, obtained by any prediction methods, pre-multiplying them by $\boldsymbol{S}$ we get:
-
-$$
-\tilde{\boldsymbol{y}}_{t+h|t} = \boldsymbol{S} \hat{\boldsymbol{b}}_{t+h|t} \, .
-$$ {#eq-sb}
-
-We refer to $\tilde{\boldsymbol{y}}_{t+h|t}$ as *coherent forecasts*, as they respect the aggregation constraints.
-We also refer to this way of obtaining coherent forecasts by summing the bottom-level forecasts as the bottom-up approach.
-However, generating forecasts this way is anchored only to prediction models at a single level, and will not be utilising the inherent information from other levels.
-This drawback applies to the top-down and middle-out approaches.
-For example, the bottom-level data can be very noisy or even intermittent, and the higher-level data might be smoother due to the aggregation.
-
-Another issue with expressing reconciled methods as in @eq-sb is that it restricts the reconciliation to only single-level approaches.
-Thus, @Hyndman2011-jv suggested a generalised expression for all existing methods, which also provides a framework for new methods to be developed:
-
-$$
-\tilde{\boldsymbol{y}}_{t+h|t} = \boldsymbol{S} \boldsymbol{G} \hat{\boldsymbol{y}}_{t+h|t} \; ,
-$$ {#eq-sgy}
-
-for a suitable $n_b \times n$ matrix $\boldsymbol{G}$.
-$\boldsymbol{G}$ maps the base forecasts of all levels $\hat{\boldsymbol{y}}_{t+h|t}$ down into the bottom level, which is then aggregated to the higher levels by $\boldsymbol{S}$.
-The choice of $\boldsymbol{G}$ determines the composition of reconciled forecasts $\tilde{\boldsymbol{y}}_{t+h|t}$, and modern reconciliation methods are developed to estimate $\boldsymbol{G}$.
-
-<!-- including the OLS and WLS developed by @Hyndman2011-jv and @Hyndman2016-ic, respectively. A more detailed explanation of these methods can be found in the @sec-ols-wls Appendix. -->
-
-## The Minimum Trace (MinT) Reconciliation
-
-@Wickramasuriya2019-xq framed the problem as minimising the variances of all reconciled forecast errors $\text{Var}[y_{t+h} - \tilde{y}_{t+h|t}] = \boldsymbol{S} \boldsymbol{G} \boldsymbol{W}_h \boldsymbol{G}' \boldsymbol{S}'$, where $\boldsymbol{W}_h = \mathbb{E}( \hat{\boldsymbol{e}}_{t+h|t} \; \hat{\boldsymbol{e}}_{t+h|t}' )$ is the positive definite covariance matrix of the $h$-step-ahead base forecast errors.
-They showed that this is equivalent to minimising the trace of the reconciled forecast error covariance matrix (sum of the diagonal elements - the variances).
-The Minimum Trace (MinT) solution is given by
-
-$$
-\boldsymbol{G} = (\boldsymbol{S}' \boldsymbol{W}_h^{-1} \boldsymbol{S})^{-1}
-\boldsymbol{S}' \boldsymbol{W}_h^{-1} .
-$$
-
-@Wickramasuriya2019-xq also showed that MinT is an algebraic generalisation of the GLS, and the OLS and WLS methods are special cases of MinT when $\boldsymbol{W}_h$ is an identity matrix $I_{n_b}$ and a diagonal matrix $\text{diag}( \boldsymbol{W}_h )$, respectively.
-In this paper, we place our main focus on the MinT method.
-
-The MinT solution hinges on a reliable, positive-definite estimate of $\boldsymbol{W}_h$, which is challenging to estimate in high-dimensional setting.
-The sample covariance matrix is unstable and non-positive-definite when the number of series $n$ is huge and larger than the time dimension $T$.
-To tackle this issue, the original paper @Wickramasuriya2019-xq adopted the diagonal-target shrinkage estimator from @Schafer2005-yw, given by
-
-$$
-\hat{\boldsymbol{W}}^{shr}_{1} = \lambda_D \hat{\boldsymbol{W}}_{1, D} + (1 - \lambda_D) \hat{\boldsymbol{W}}_1 \, ,
-$$
-
-where $\hat{\boldsymbol{W}}_{1, D}$ is a diagonal matrix comprising the diagonal entries $\text{diag}( \hat{\boldsymbol{W}}_1 )$.
-We refer to any $\lambda \in [0,1]$ as the shrinkage intensity parameter, the subscript specifies which estimator it belongs to.
-This approach shrinks the covariance matrix $\hat{\boldsymbol{W}}_1$ towards its diagonal matrix, meaning the off-diagonal elements are shrunk towards zero while the diagonal ones remain unchanged.
-
-@Schafer2005-yw also proposed an estimate of the optimal shrinkage intensity parameter $\lambda_D$:
-
-$$
-\hat{\lambda}_D = \frac{\sum_{i \neq j} \widehat{Var}(\hat{r}_{ij})} 
-{\sum_{i \neq j} \hat{r}_{ij}^2} \, ,
-$$
-
-where $\hat{r}_{ij}$ is the $ij$th element of $\hat{\boldsymbol{R}}_1$, the 1-step-ahead sample correlation matrix (obtained from $\hat{\boldsymbol{W}}_1$).
-The optimal estimate is obtained by minimising $MSE(\hat{\boldsymbol{W}}_1) = Bias(\hat{\boldsymbol{W}}_1)^2 + Var(\hat{\boldsymbol{W}}_1)$.
-More specifically, we trade the unbiasedness of the sample covariance matrix for a lower variance.
-
-However, the hierarchical time series data often exhibit a prominent principal components structure, which is not fully taken advantage.
-Taking an example of the Australian domestic overnight trips data set [@tourism-au], where the national trips are disaggregated into states and territories, and further into regions.
-We then fit ETS models to all series, using the algorithm from Fabletools R package [@O-Hara-Wild2024-we], and compute the one-step-ahead in-sample base forecast error covariance matrix $\hat{\boldsymbol{W}}_1$.
-The twenty largest eigenvalues of the covariance matrix are plotted in @fig-eigen.
-We can see that the point of inflexion occurs at the component with 5th largest eigenvalue, indicating a prominent principal components structure.
-
-![Twenty largest eigenvalues of one-step-ahead in-sample base forecast error covariance, Australian domestic overnight trips](D:/Github/Recon_Honours_Thesis/research_proposal/figs/eigen_tourism.png){#fig-eigen width="70%"}
-
-Additionally, the shrinkage estimator shrinks all off-diagonal elements towards zeros with equal weights $\lambda_D$.
-We might prefer to better preserve strong signals, and largely reduce the effects of small, noisy correlations.
-In the next sections, we will explore several options that take these two issues into account.
-
-# Covariance Estimation Approaches {#sec-cov-est}
-
-## NOVELIST Estimator
-
-The NOVELIST (NOVEL Integration of the Sample and Thresholded Covariance) estimator, proposed by @Huang2019-ua, introduces a way to control the target matrix's sparsity, retaining strong correlations while discarding weak, noisy effects.
-NOVELIST offers more flexibility than the shrinkage estimator, which is useful when we believe that only a few variables are truly correlated.
-However, it does not guarantee to be positive definite.
-
-The method is based on the idea of soft-thresholding the sample covariance matrix, then performing shrinkage towards this thresholded version.
-This introduces an extra parameter, the threshold $\delta$, which is used to control the amount of soft-thresholding.
-The NOVELIST estimator is given by:
-
-$$
-\hat{\boldsymbol{W}}^{N}_{1} = \lambda_{\delta} \hat{\boldsymbol{W}}_{1, \delta} + (1 - \lambda_{\delta}) \hat{\boldsymbol{W}}_1,
-$$ {#eq-novelist-cov}
-
-where $\hat{\boldsymbol{W}}_{1, \delta}$ is the thresholded version of $\hat{\boldsymbol{W}}_{1}$. By convenient setting, we can rewrite it in terms of correlation:
-
-$$
-\hat{\boldsymbol{R}}^{N}_{1} = \lambda_{\delta} \hat{\boldsymbol{R}}_{1,\delta} + (1 - \lambda_{\delta}) \hat{\boldsymbol{R}}_1,
-$$ {#eq-novelist-cor}
-
-In this setting, $\hat{\boldsymbol{R}}_{1,\delta}$ is the thresholded correlation matrix, where each element is regularised by:
-
-$$
-\hat{r}_{1,ij}^\delta = \text{sign}(\hat{r}_{1,ij}) \, 
-\text{max}(|\hat{r}_{1,ij}| - \delta, \; 0),
-$$ {#eq-soft-thr}
-
-where $\delta \in [0,1]$ is the threshold parameter.
-For a given threshold $\delta$, @Huang2019-ua derived an analytical expression for the optimal shrinkage intensity parameter $\lambda(\delta)$ using Ledoit-Wolf's lemma [@Ledoit2003-qv], following similar logic to @Schafer2005-yw.
-It can be computed as:
-
-<!-- - \textcolor{red}{@Huang2019-ua mentioned Ledoit-Wolf's lemma} -->
-
-$$
-\hat{\lambda}(\delta) = \frac{
-  \sum_{i \neq j} \widehat{Var}(\hat{r}_{1,ij}) \; \boldsymbol{1}(|\hat{r}_{1,ij}| \leq \delta)
-} {
-  \sum_{i \neq j} (\hat{r}_{1,ij} - \hat{r}_{1,ij}^\delta)^2
-},
-$$ {#eq-lambda-thr}
-
-where $\boldsymbol{1}(.)$ is the indicator function.
-
-On the other hand, the optimal threshold $\delta^*$ does not have a closed-form solution, and is typically obtained by executing a rolling-window cross-validation procedure.
-The idea is to find the threshold $\hat{\delta^*}$, with the corresponding $\hat{\lambda}^*$ and $\hat{\boldsymbol{R}}^{N}_{1}(\hat{\delta^*}, \hat{\lambda}^*)$, that minimises the average out-of-sample 1-step-ahead reconciled forecast mean squared error over all windows.
-The formal algorithm is given in the @sec-novelist_cv Appendix.
-Although it is not required to fit forecasting models multiple times, the cross-validation procedure is still computationally expensive as it computes the NOVELIST estimator and perform reconciliation for each threshold value.
-
-We also tested out minimising 1- to h-step-ahead MSE in the cross-validation procedure.
-Surprisingly, it returns almost the same best threshold parameter is in the 1-step-ahead case above.
-Note that when $\delta \in \bigl[ \text{max}_{i \neq j}|\hat{r}_{1,ij}|, \; 1 \bigr]$, the NOVELIST estimator collapses to the shrinkage estimator, and when $\delta = 0$, it becomes the sample covariance matrix.
-
-<!-- ## POET
-
-The POET (Principal Orthogonal complEment Thresholding) estimator, proposed by @Fan2013-jz, is another "sparse" + "non-sparse" covariance estimator.
-It takes the latent factors directly into its construction, and is appealing when there are common drivers in the time series within the hierarchy, as we saw in the Australian tourism example.
-
-The POET method starts by decomposing the correlation matrix $\hat{\boldsymbol{R}}_1$ into a prominent principle components part (low-rank) and a orthogonal complement part $\hat{\boldsymbol{R}}_{1, K}$ (the correlation matrix after removing the first $K$ principal components).
-Then it applies thresholding to $\hat{\boldsymbol{R}}_{1, K}$.
-The POET estimator is given by:
-
-$$
-\hat{\boldsymbol{R}}^{K}_{1} = \sum_{k = 1}^K \hat{\gamma}_k \hat{\boldsymbol{\xi}}_k \hat{\boldsymbol{\xi}}_k' + T(\hat{\boldsymbol{R}}_{1, K})
-$$
-
-where $\hat{\gamma}_k$ and $\hat{\boldsymbol{\xi}}_k$ are the $k$th largest eigenvalue and the corresponding eigenvector of the sample covariance matrix, respectively, and $T(.)$ is the thresholding function, which can be either soft-thresholding, hard-thresholding, or others.
--->
-
-<!-- ## PC-adjusted NOVELIST
-
-This approach is best of both worlds, leveraging the strengths of both NOVELIST and POET.
-The PC-adjusted (Principal-Component-adjusted) NOVELIST overcomes the shortcomings of the current shrinkage estimator, taking prominent PCs into account while also offers extra flexibility.
-The idea is to apply the NOVELIST estimator to the orthogonal complement part $\hat{\boldsymbol{R}}_{1, K}$, and then add the principal components part back.
-The PC-adjusted NOVELIST estimator is formulated as:
-
-$$
-\hat{\boldsymbol{R}}^{N, K}_{1} = \sum_{k = 1}^K \hat{\gamma}_k \hat{\boldsymbol{\xi}}_k \hat{\boldsymbol{\xi}}_k' + \hat{\boldsymbol{R}}^{N}_{1, K}
-\; ,
-$$
-
-where $\hat{\boldsymbol{R}}^{N}_{1, K}$ is the NOVELIST estimator applied to the orthogonal complement part $\hat{\boldsymbol{R}}_{1, K}$.
-Similar to the NOVELIST estimator, $\hat{\boldsymbol{R}}^{N, K}_{1}$ is not guaranteed to be positive definite.
-
-Methods to ensure positive definiteness of the NOVELIST estimator (and its PC-adjusted variant) will be explored and studied in the project.
-@Huang2019-ua proposed to diagonalise the NOVELIST estimator and replace any eigenvalues that fall under a certain small positive threshold by the value of that threshold.
-Alternatively, we can implement the algorithm of @Higham2002-te that computes the nearest positive definite matrix to a given matrix.
--->
-
-## PC-adjusted Estimator
-
-This method takes the latent factors directly into its construction, and is appealing when there are common drivers in the time series within the hierarchy, as we saw in the Australian tourism example.
-It starts by decomposing the correlation matrix $\hat{\boldsymbol{R}}_1$ into a prominent principle components part (low-rank) and a orthogonal complement part $\hat{\boldsymbol{C}}^K_{1}$ (the correlation matrix after removing the first $K$ principal components).
-Then we can apply either shrinkage or NOVELIST estimator to $\hat{\boldsymbol{R}}_{1, K}$:
-
-$$
-\hat{\boldsymbol{R}}^{g, K}_{1} = \sum_{k = 1}^K \hat{\gamma}_k \hat{\boldsymbol{\xi}}_k \hat{\boldsymbol{\xi}}_k' + g(\hat{\boldsymbol{C}}^K_{1})
-$$
-
-where $g(.)$ is either the shrinkage or NOVELIST estimator, $\hat{\gamma}_k$ and $\hat{\boldsymbol{\xi}}_k$ are the $k$-th largest eigenvalue and the corresponding eigenvector of the sample covariance matrix, respectively.
-Similar to the NOVELIST estimator, $\hat{\boldsymbol{R}}^{*, K}_{1}$ is not guaranteed to be positive definite.
-
-## Scaled Variance
-
-Apply shrinkage or NOVELIST to the correlation matrix of 1-step-ahead base forecast errors, then scale it back to covariance matrix of h-step-ahead base forecast errors using the h-step-ahead standard deviations:
-
-$$
-\hat{\boldsymbol{W}}^{sv}_{h} = \boldsymbol{D}^{1/2}_h g(\hat{\boldsymbol{R}}_1) \boldsymbol{D}^{1/2}_h,
-$$
-
-where $\boldsymbol{D}_h = \text{diag}(\hat{\sigma}_{1,h}, \hat{\sigma}_{2,h}, \ldots, \hat{\sigma}_{n_b,h})$, and $\hat{\sigma}_{i,h}$ is the standard deviation of the $i$-th series' h-step-ahead base forecast errors.
-The asterisk $*$ indicates either shrinkage or NOVELIST estimator.
-
-## Constructing from h-step-ahead residuals
-
-Construct the sample covariance matrix directly from h-step-ahead base forecast errors, then apply either shrinkage or NOVELIST estimator:
-
-$$
-\hat{\boldsymbol{W}}^{g}_{h} = g(\hat{\boldsymbol{W}}_h)
-$$
-
-where $\hat{\boldsymbol{R}}_h$ is the sample correlation matrix of h-step-ahead base forecast errors, and $\circ$ denotes the Hadamard product (element-wise product).
-The asterisk $*$ indicates either shrinkage or NOVELIST estimator.
-
-# Simulation Design {#sec-simulation}
-
-<!-- -   Talk about description of the hierarchical time series data set (e.g., economic, financial, or synthetic data). -->
-
-<!-- -   Characteristics such as dimensionality, frequency, and hierarchical structure. -->
-
-<!-- **Experimental Design** -->
-
-<!-- -   Design different cases of simulation studies and real-data experiments. -->
-
-<!-- -   Metrics: Forecast accuracy (e.g., RMSE, MAE), reconciliation error reduction, and matrix stability... -->
-
-<!-- -   Visualisation... -->
-
-The general design of data generating process for bottom-level series is a stationary VAR(1) process, with the following structure:
-
-$$
-\boldsymbol{b}_t = \boldsymbol{A} \boldsymbol{b}_{t-1} + \boldsymbol{\epsilon}_t,
-$$
-
-where $\boldsymbol{A}$ is a $n_b \times n_b$ block diagonal matrix of autoregressive coefficients $\boldsymbol{A} = diag(\boldsymbol{A}_1, \ldots, \boldsymbol{A}_m)$, with each $\boldsymbol{A}_i$ being a $n_{b,i} \times n_{b,i}$ matrix.
-The block diagonal structure ensures that the time series are grouped into $m$ groups, with each group having its own autoregressive coefficients.
-This aim to simulate the interdependencies between the time series within each group, where reconciliation will be expected to better performed than the usual base forecasts.
-
-The model is added with a Gaussian innovation process $\boldsymbol{\epsilon}_t$, with covariance matrix $\boldsymbol{\Sigma}$.
-The covariance matrix $\boldsymbol{\Sigma}$ is generated specifically using the Algorithm 1 in @Hardin2013-wu:
-
-1.
-A compound symmetric correlation matrix is used for each block of size $n_{b,i}$ in $\boldsymbol{A}_i$, where the entries $\rho_i$ for each block $i$ are sampled from a uniform distribution between 0 and 1.
-They are baseline correlations within group.
-
-2.  A constant correlation, which is smaller than $\text{min} \{\rho_1, \rho_2, \dots, \rho_m \}$, is imposed on the entries between different blocks. It serves as baseline correlations between group.
-
-3.  The entry-wise random noise is added on top of the entire correlation matrix.
-
-4.  The covariance matrix $\boldsymbol{\Sigma}$ is then constructed by uniform sampling of standard deviations, in a range of $[\sqrt{2}, \sqrt{6}]$, for all $n_b$ series.
-
-We will randomly flip the signs of the covariance elements, which will create a more realistic structure in the innovation process.
-This can be done by pre- and post-multiplying $\boldsymbol{\Sigma}$ by a random diagonal matrix $\boldsymbol{V}$ with diagonal entries sampled from $\{-1, 1\}$, yielding $\boldsymbol{\Sigma}^* = \boldsymbol{V} \boldsymbol{\Sigma} \boldsymbol{V}$.
-
-For all hierarchies in our experiments, we simulate two panel lengths, $T=54$ and $T=304$, reserving the final four observations as an out-of-sample test set.
-In each Monte Carlo replication ($M=500$), we fit univariate ARIMA models (base models) to the training observations using an automatic AICc minimization algorithm from the fabletools package [@O-Hara-Wild2024-we], generating incoherent 1–4-step base forecasts.
-We then reconcile these forecasts under three covariance estimators: the raw sample covariance (mint_sample), the shrinkage estimator (mint_shr), and the NOVELIST estimator (mint_n).
-
-All data generation, covariance estimation, and reconciliation routines were implemented in the ReconCov R package, a package developed solely for this research, and is available under an open‐source license on GitHub [@ReconCov].
-
-```{r load results}
-sim1_1 <- readRDS("./literature/lib/S4-2-1_T50_M500.rds")
-sim1_2 <- readRDS("./literature/lib/S4-2-1_T300_M500.rds")
-
-sim2_1 <- readRDS("./literature/lib/S36-6-1_T50_M500.rds")
-sim2_2 <- readRDS("./literature/lib/S36-6-1_T300_M500.rds")
-
-# sim3_1 <- readRDS("./literature/lib/S100-2-1_T50_M200_run1.rds")
-# sim3_2 <- readRDS("./literature/lib/S100-2-1_T100_M200_run1.rds")
-
-sim_dense_50 <- readRDS("./literature/lib/S100-10-3-1_T50_M200_dense.rds")
-sim_dense_300 <- readRDS("./literature/lib/S100-10-3-1_T300_M200_dense.rds")
-
-sim_sparse_50 <- readRDS("./literature/lib/S100-10-3-1_T50_M500_sparsebwgrp.rds")
-sim_sparse_300 <- readRDS("./literature/lib/S100-10-3-1_T300_M200_sparsebwgrp.rds")
-
-sim3_1 <- sim_dense_50 # using same with next section simulation
-sim3_2 <- sim_dense_300
-```
-
-## Exploring Effects of Hierarchy's Size
-
-In our first set of experiments, we examine how MinT combined with the NOVELIST estimator scales as the hierarchy expands.
-We generate synthetic data from the same VAR(1) framework described earlier, but vary the number of bottom‐level series, $n_b$, across three different structures: the smallest with two groups of two ($n_b = 4$), an intermediate case with six groups of six ($n_b = 36$), and a much larger configuration with two clusters of fifty ($n_b = 100$).
-In the 4‐series hierarchy, each pair collapses into a single level‐1 series, which then sums to the top.
-In the 36‐series case, each block of six forms a level‐1 aggregate, and those six aggregates form the national total.
-
-The 100‐series design employs a deliberately intricate aggregation path to stress‐test reconciliation methods.
-We first sum the one hundred bottom series into ten intermediate series by grouping them in contiguous blocks of ten.
-These ten series are then organised into three level-2 aggregates—four, three, and four series, respectively—before finally summing to a single top node.
-This asymmetric hierarchy creates overlapping correlation patterns: some level‐2 series share bottom‐level groups, while others draw from both, emulating practical scenarios such as regional sales aggregations that span multiple product categories or overlapping territories.
-
-To save space, we illustrate only the six-by-six's VAR(1) and correlation configuration in @fig-settings-6x6; the two-by-two and the one-hundred-series structures appear in Appendix @sec-sim-supplement.
-
-```{r generate plots for A and Sigma}
-MSE1_1 <- transform_sim_MSE(sim1_1$MSE, F) |> 
-  filter(.model != "ols")
-MSE1_2 <- transform_sim_MSE(sim1_2$MSE, F) |> 
-  filter(.model != "ols")
-
-MSE2_1 <- transform_sim_MSE(sim2_1$MSE, F) |> 
-  filter(! .model %in% c("mint_n_hstep", "ols"))
-MSE2_2 <- transform_sim_MSE(sim2_2$MSE, F) |> 
-  filter(! .model %in% c("mint_n_hstep", "ols"))
-
-MSE3_1 <- transform_sim_MSE(sim3_1$MSE, F)
-MSE3_2 <- transform_sim_MSE(sim3_2$MSE, F) |> 
-  filter(.model != "ols")
-
-p1_A <- plot_heatmap(sim1_1$A, T) + 
-  ggtitle("2x2 VAR(1) Coefficients Matrix")
-
-p1_Sigma <- plot_heatmap(cov2cor(sim1_1$Sigma), T) + 
-  ggtitle("2x2 Correlation Matrix")
-
-p2_A <- plot_heatmap(sim2_1$A) +
-  ggtitle("VAR(1) Coefficients Matrix") + theme_void()
-
-p2_Sigma <- plot_heatmap(cov2cor(sim2_1$Sigma), T) +
-  ggtitle("Correlation Matrix") + theme_void()
-
-p3_A <- plot_heatmap(sim3_1$A) +
-  ggtitle("2x50 VAR(1) Coefficients Matrix")
-
-p3_Sigma <- plot_heatmap(cov2cor(sim3_1$Sigma), T) +
-  ggtitle("2x50 Correlation Matrix")
-```
-
-```{r fig settings 6x6}
-#| label: fig-settings-6x6
-#| fig-cap: "Heatmaps of the VAR(1) coefficient matrix and correlation matrix for the 2x50 structure."
-#| fig.height: 4
-#| fig.width: 6
-#| out.width: "85%"
-#| fig.align: "center"
-
-p2_A + p2_Sigma + plot_layout(ncol = 2) & theme(legend.position = "bottom")
-```
-
-```{r}
-plot_relative_improvement <- function(MSE_data) {
-  MSE_data |>
-    filter(
-      .model != "mint_true",
-      h <= 4
-    ) |>
-    group_by(series, h) |>
-    mutate(
-      base_MSE = MSE[.model == "base"]
-    ) |>
-    ungroup() |>
-    group_by(.model, h) |>
-    summarise(
-      MSE = mean(MSE),
-      base_MSE = mean(base_MSE),
-      pct_change = (MSE - base_MSE) / base_MSE * 100
-    ) |>
-    ggplot(aes(x = h, y = pct_change, color = .model)) +
-    geom_line() +
-    labs(
-      x = "Forecast Horizon (h)",
-      y = "Percentage Change (%)"
-    ) +
-    # labs(x = element_blank(), y = element_blank()) +
-    theme_bw()
-}
-```
-
-```{r fig-sim-results-1}
-#| label: fig-sim-results-1
-#| fig-cap: "Relative improvement of the MSE of reconciled forecasts over the base forecasts for the 2x2, 6x6, and 2x50 hierarchical structures, for 1 to 4 steps ahead forecasts, with 2 time series lengths (T = 54 and T = 304)."
-#| fig.width: 8
-#| fig.height: 6
-#| out.width: "100%"
-#| fig.align: "center"
-
-p_MSE1_1 <- MSE1_1 |> 
-  plot_relative_improvement() +
-    labs(title = "2x2 Hierarchy", subtitle = "T = 54")
-p_MSE2_1 <- MSE2_1 |> 
-  plot_relative_improvement() +
-    labs(title = "6x6 Hierarchy")
-p_MSE3_1 <- MSE3_1 |> 
-  plot_relative_improvement() +
-    labs(title = "2x50 Hierarchy")
-
-p_MSE1_2 <- MSE1_2 |> 
-  plot_relative_improvement() +
-    labs(subtitle = "T = 304")
-p_MSE2_2 <- MSE2_2 |> 
-  plot_relative_improvement()
-p_MSE3_2 <- MSE3_2 |> 
-  plot_relative_improvement()
-
-p_MSE1_1 + p_MSE2_1 + p_MSE3_1 + p_MSE1_2 + p_MSE2_2 + p_MSE3_2 +
-  plot_layout(ncol = 3, guides = "collect", axis_titles = "collect") &
-  theme(legend.position = "right")
-```
-
-Figure @fig-sim-results-1 illustrates that as hierarchy size and complexity increase, the relative improvements in MSE over the base forecasts from reconciliation amplify, across all 1- to 4-step horizons.
-While extending the training window (from 50 to 300 in‐sample observations) uniformly improves the accuracy of base ARIMA forecasts, it does not alter the relative improvements of mint_shr or mint_n.
-This is evidence that both high-dimensional estimators robustly handle the "large $p \gg T$” regime.
-By contrast, mint_sample achieves significant accuracy when more data make the sample covariance more reliable.
-
-Despite their theoretical differences, mint_shr and mint_n exhibit nearly identical performance in these canonical settings.
-Yet NOVELIST’s thresholding ability to minimise effects of weak correlations and retain strong ones suggests it may excel in contexts where correlation matrix has small, noisy entries.
-This motivates our subsequent exploration of sparse covariance structures, in which many off-diagonals are truly zero.
-<!--
-## Exploring Effects of Hierarchy's Aggregation
-
-- 2x50 with usual path and weird path
-
-- Finding: Complex aggregation MinT_n performs well
--->
-
-## Exploring the sparsity of the DGP covariance matrix
-
-In our second simulation study, we design a data‐generating process that contrasts “dense” and “sparse” correlation regimes among bottom‐level series, reflecting settings one might encounter in practice.
-We consider the same hierarchical structure of two large groups of 50 bottom series as above.
-Specifically, the two groups would have strong within-group dependencies throughout and either modest between‐group correlations (the dense scenario) or complete independence (the sparse scenario).
-These correlation matrices are depicted in Figure @fig-Sigma-2x50.
-Both scenarios share the same VAR(1) coefficient structure as in our previous simulations; only the innovation covariance changes.
-Such a setup mirrors real‐world contexts where, for example, sales within a product line may exhibit strong co‐movements, while those in a separate line operate nearly independently.
-
-<!-- Rather than collapsing each group into a single aggregate as in previous practice, we impose a more intricate, multi‐level aggregation path to stress‐test reconciliation methods.
-We first form ten intermediate series by summing bottom‐level series in consecutive blocks of ten.
-These ten series are then assigned to three second‐level aggregates--groups of four, three, and four--and finally summed to produce a single top‐level series.
-This asymmetric hierarchy creates overlapping correlation patterns: some level‐2 series share bottom‐level groups, while others draw from both.
--->
-
-<!-- emulating practical scenarios such as regional sales aggregations that span multiple product categories or overlapping territories. -->
-
-```{r}
-#| label: fig-Sigma-2x50
-#| fig-cap: "Heatmaps of the dense and sparse correlation matrix of the data generating process."
-#| fig.height: 12
-#| fig.width: 8
-#| out.width: "100%"
-#| fig.align: "center"
-p_dense_Sigma <- sim_dense_50$Sigma |>
-  cov2cor() |> 
-  plot_heatmap(T) +
-    labs(title = "Dense")
-
-p_sparse_Sigma <- sim_sparse_50$Sigma |>
-  cov2cor() |> 
-  plot_heatmap(T) +
-    labs(title = "Sparse")
-
-p_dense_Sigma + p_sparse_Sigma + 
-  plot_layout(ncol = 1, guides = "collect", axis_titles = "collect") & 
-  theme(legend.position = "right")
-```
-
-```{r}
-#| label: fig-sim-results-dsp
-#| fig-cap: "Relative improvement of the MSE of reconciled forecasts over the base forecasts for the 2x50 hierarchical structure with dense and sparse DGP's correlation matrix, for 1 to 4 steps ahead forecasts, with 2 time series lengths (T = 54 and T = 304)."
-#| fig.height: 7
-#| fig.width: 8
-#| out.width: "90%"
-#| fig.align: "center"
-p_dense_50 <- sim_dense_50$MSE |>
-  transform_sim_MSE(F) |> 
-  plot_relative_improvement() +
-    labs(title = "Dense DGP Covariance Matrix", subtitle = "T = 54")
-p_sparse_50 <- sim_sparse_50$MSE |>
-  transform_sim_MSE(F) |> 
-  plot_relative_improvement() +
-    labs(title = "Dense DGP Covariance Matrix")
-
-p_dense_300 <- sim_dense_300$MSE |>
-  transform_sim_MSE(F) |> 
-  filter(.model != "ols") |> 
-  plot_relative_improvement() +
-    labs(subtitle = "T = 304")
-p_sparse_300 <- sim_sparse_300$MSE |>
-  transform_sim_MSE(F) |> 
-  plot_relative_improvement()
-
-p_dense_50 + p_sparse_50 + p_dense_300 + p_sparse_300 +
-  plot_layout(ncol = 2, guides = "collect", axis_titles = "collect") &
-  theme(legend.position = "right")
-```
-
-Figure @fig-sim-results-dsp presents out‐of‐sample mean squared error improvements over the base forecasts for each reconciliation strategy under both dense and sparse settings, with two panel lengths--54 and 304 observations--reserving the last four points for testing.
-In both short and long samples, MinT using either the shrinkage estimator (mint_shr) or the NOVELIST estimator (mint_n) delivers pronounced gains over incoherent ARIMA forecasts, particularly at the one‐step horizon where cross‐series correlations most directly inform the forecast adjustments.
-Although the two MinT variants perform almost indistinguishably overall, mint_n edges out mint_shr in the immediate horizon, whereas mint_shr slightly outperforms for longer horizons.
-By contrast, MinT with the raw sample covariance (mint_sample) suffers in small‐sample settings; as expected, its performance improves dramatically with 304 data points, since the sample covariance becomes more reliable with larger $n$.
-This highlights the practical necessity of regularized estimators in high‐dimensional, low‐sample contexts, a situation common in real applications where histories are short relative to the number of series.
-
-Additional designs (varying block sizes, aggregation paths, correlation configurations) also failed to separate NOVELIST from Shrinkage.
-Their nearly identical performance under these synthetic scenarios suggests that our current simulation may not unveil the full advantages of the thresholding estimators.
-This finding motivates our turn to empirical data, where latent structural features and regime shifts, which we will discuss in the next section, may reveal performance differences.
-
-<!--
-## Exploring Grouped Structure
-
-- small vs big
-
-- Finding: NOVELIST performs well in both cases
--->
 
@@
 Results are aligned across probabilistic scoring rules, but the 95% Winkler reveals a notable pattern, as all MinT variants lose to the base while *OLS* performs best.
+In the multivariate evaluation, the Energy score places *OLS* close to the PC1-adjusted MinT methods, indicating that simple coherent projection might be able to capture a substantial share of cross-series dependence benefits even without sophisticated covariance regularisation.
 
 \color{red}
-This points to tail-calibration sensitivity at high nominal coverage.
-The reconciliation that helps central accuracy can degrade extreme-tail interval performance when uncertainty is misallocated across the hierarchy.
+Discuss the 95% empirical coverage 
 \color{black}
-
-In the multivariate evaluation, the Energy score places *OLS* close to the PC1-adjusted MinT methods, indicating that simple coherent projection might be able to capture a substantial share of cross-series dependence benefits even without sophisticated covariance regularisation.
 
@@
 
-<!-- # Supplementary -->
-
-<!-- 
-\pagebreak
-
-# Timeline & Milestones {#sec-timeline}
-
-```{r timeline}
-#| echo: false
-#| message: false
-#| warning: false
-library(tibble)
-library(knitr)
-library(kableExtra)
-
-tibble(
-  Period = c(
-    "March",
-    "April", 
-    "May", 
-    "June",
-    "July",
-    "August", 
-    "September",
-    "October"
-  ),
-  Task = c(
-    "Initial setup and understanding the research scope. Preliminary knowledge and Literature review",
-
-    "Develop ReconCov, an R package implementing the Min Trace approach with the shrinkage and NOVELIST estimators, and rolling-window cross-validation process for choosing NOVELIST threshold parameter.
-Exploring simulation framework.",
-
-    "Code functions for producing simulation settings in ReconCov package. Initial simulation for MinT in hierarchical time series. Writing research proposal and presenting research project.",
-
-    "Intensive simulation on multiple different settings, in order to assess the performance and behaviours of NOVELIST under different conditions, and to explore cases where NOVELIST performs different than Shrinkage.",
-
-    "Empirical analysis on forecasting Australian Domestic Tourism data. Explore the real-world data generating process which can be used to refine the current simulation design",
-
-    "Implement the Tourism DGP configuration into the current simulation settings.
-Implementing POET and PC-adjusted NOVELIST estimators.
-Assessing whether prominent principal components structure can be exploited to improve the out-of-sample forecast performance.",
-
-    "Exploring other high-dimensional covariance estimators. More simulations on all explored covariance estimators. Final,  full-fledged assessment on all implemented estimators",
-
-    "Final writing, presenation, submission"
-  ),
-  Deliverable = c(
-    "Initial literature review + paper draft",
-    "ReconCov package: Main codes for estimators + simulation framework + cross-validation",
-    "NOVELIST initial simulation results + research proposal + presentation",
-    "More simulation results on NOVELIST",
-    "Empirical results on forecasting Tourism data",
-    "New simulation design, POET & PC-adjusted NOVELIST simulation results",
-    "Other estimators simulation results and full-fledged covariance estimators assessment",
-    "Presentation and thesis manuscript"
-  )
-) |>
-  kable(
-    format = "latex",
-    booktabs = TRUE,
-    longtable = TRUE,
-    caption = "Project Timeline and Key Milestones",
-    escape = TRUE
-  ) |>
-  column_spec(1, width = "1.9cm") |>
-  column_spec(2, width = "8cm") |>
-  column_spec(3, width = "4.8cm") |>
-  kable_styling(latex_options = "hold_position", font_size = 10)
-```
--->
-
 \pagebreak
@@
 # Appendix {#sec-appendix}
-
-## Simulation Settings: Supplementary Figures {#sec-sim-supplement}
-
-```{r fig settings 2x2}
-#| label: fig-settings-2x2
-#| fig-cap: "Heatmaps of the VAR(1) coefficient matrix and correlation matrix for the 2x2 structure."
-#| fig.height: 4
-#| fig.width: 6
-#| out.width: "55%"
-#| fig.align: "center"
-
-p1_A + p1_Sigma + plot_layout(ncol = 2) & theme(legend.position = "bottom")
-```
-
-```{r fig settings 2x50}
-#| label: fig-settings-2x50
-#| fig-cap: "Heatmaps of the VAR(1) coefficient matrix and correlation matrix for the 6×6 structure."
-#| fig.height: 12
-#| fig.width: 8
-#| out.width: "85%"
-#| fig.align: "center"
-
-p3_A + p3_Sigma + plot_layout(ncol = 1) & theme(legend.position = "right")
-```
-
-\pagebreak
-
-## Algorithm: NOVELIST cross-validation for optimal threshold $\delta^*$ {#sec-novelist_cv}
-
-The cross-validation algorithm for NOVELIST is available in the ReconCov package [@ReconCov].
-
-```{=latex}
-\begin{algorithm}
-\caption{Cross-validation procedure}
-\begin{algorithmic}[1]
-
-\State \textbf{Input:} Observations and fitted values $\boldsymbol{y}_t, \hat{\boldsymbol{y}}_t \in \mathbb{R}^n$ for $t = 1,\dots,T$, set of threshold candidates $\Delta$, window size $v$.
-
-\State $\hat{\boldsymbol{e}}_t = \boldsymbol{y}_t - \hat{\boldsymbol{y}}_t$ for $t = 1,\dots,T$
-
-\For{$i = v:T-1$}
-    \State $j = i - v +1$
-    \State $\hat{\boldsymbol{W}}_j = \frac{1}{v} \sum_{t=j}^{i} \hat{\boldsymbol{e}}_{t} \hat{\boldsymbol{e}}_{t}'$
-    \State $\hat{\boldsymbol{D}}_j = \text{diag}(\hat{\boldsymbol{W}}_j)$
-    \State $\hat{\boldsymbol{R}}_j = \hat{\boldsymbol{D}}_j^{-1/2} \hat{\boldsymbol{W}}_j \hat{\boldsymbol{D}}_j^{-1/2}$
-    \For{$\delta \in \Delta$}
-        \State Compute thresholded correlation $\hat{\boldsymbol{R}}_{j,\delta}$ using Equation 5
-        \State Compute $\hat{\lambda}_{j,\delta}$ using Equation 6
-        \State Compute $\hat{\boldsymbol{R}}^{N}_{j,\delta}$ using Equation 4
-        \State $\hat{\boldsymbol{W}}^{N}_{j,\delta} = \hat{\boldsymbol{D}}_j^{1/2} \hat{\boldsymbol{R}}^{N}_{j,\delta} \hat{\boldsymbol{D}}_j^{1/2}$
-        \State $\boldsymbol{G} = (\boldsymbol{S}' \hat{\boldsymbol{W}}^{N^{-1}}_{j,\delta} \boldsymbol{S})^{-1} \boldsymbol{S}' \hat{\boldsymbol{W}}^{N^{-1}}_{j,\delta}$
-        \State Reconciled forecasts $\tilde{\boldsymbol{y}}_{i+1 | \delta} = \boldsymbol{S} \boldsymbol{G} \hat{\boldsymbol{y}}_{i+1}$
-        \State $\tilde{\boldsymbol{e}}_{i+1 | \delta} = \boldsymbol{y}_{i+1} - \tilde{\boldsymbol{y}}_{i+1 | \delta}$
-    \EndFor
-\EndFor
-
-\State $\text{MSE}_{\delta} = \frac{1}{T-v} \sum_{i=v}^{T-1} (\tilde{\boldsymbol{e}}_{i+1 | \delta})^2$ for each $\delta \in \Delta$
-\State $\hat{\delta}^* = \arg\min_{\delta \in \Delta} \text{MSE}_{\delta}$
-\State Compute $\hat{\lambda}^*$ on all training data using $\hat{\delta}^*$
-\State Compute $\hat{\boldsymbol{R}}_1^*$ using $\hat{\delta}^*$ and $\hat{\lambda}^*$ on all training data, using Equation 3
-
-\State \textbf{Output:} Estimate of optimal $\hat{\delta}^*$
-
-\end{algorithmic}
-\end{algorithm}
-```
 
@@
 ```
-
-\pagebreak
-
-# References

### 2025-10-07 literature/archived.qmd
move old to archived

@@
+<!--
+-   **Problem Statement:**
+
+    The sample covariance matrix, although natural, suffers in high-dimensional settings.
+Especially when the number of series $p$ is huge and larger than the time dimension $T$, the sample covariance matrix is non-positive definite (rank T if p\>T).
+
+    The shrinkage estimators come in to tackle this issue.
+The shrinkage estimator with diagonal target (often shrinking toward a diagonal matrix) is proven to produce a guaranteed PD matrix (Schäfer & Strimmer, 2005).
+However, as it shrinks the covariance matrix toward a diagonal one, it does not have flexibility and might neglect the prominent structure presented in the covariance matrix.
+
+    An alternative approach is to perform shrinkage of the sample covariance towards its thresholded version, instead of a diagonal matrix.
+This is the NOVELIST (NOVEL Integration of the Sample and Thresholded covariance estimators) method proposed by Huang & Fryzlewicz (2019).
+They introduced thresholding functions applied only to off-diagonal elements, allowing for more flexibility in the estimation.
+
+    [... can include more estimators ...]{style="color:red;"}
+
+-   **Research Aim:**
+
+    This paper assesses the reconciled forecasting performance of MinT approach using various covariance estimators, with a focus on the NOVELIST estimator.
+
+-   **Paper Outline:**
+
+    The paper is structured as follows:
+
+    -   A literature review of forecast reconciliation and covariance estimation.
+    -   A description of the methodology, including the NOVELIST estimator and its principal-component-adjusted variant.
+    -   An experimental design using both synthetic and real hierarchical time series.
+    -   Empirical results and discussion.
+    -   Conclusions and suggestions for future work.
+-->
+
+<!--
+# Literature Review
+
+## Forecast Reconciliation in Hierarchical and Grouped Time Series
+
+Forecast reconciliation converts a collection of independent base forecasts into a set of coherent forecasts that respect the linear constraints defining a hierarchical or grouped time-series system.
+Early work focused on heuristic single-level strategies, including bottom-up, top-down, and middle-out (...), each of which exploits only part of the information in the hierarchy and can induce bias or high variance.
+
+-   Cite the single level approach
+-   Athanasopoulos et al., 2024
+
+Hyndman et al.
+(2011) first showed that all single-level methods can be written as $\tilde{y} = SG\hat{y}$, where $S$ is the summing matrix and $G$ is a matrix that maps base forecasts $\hat{y}$ to into the bottom level.
+Treating reconciliation as a GLS regression problem, Hyndman et al.
+(2011) found that it yields a solution for $G$, but the required covariance matrix of reconcilation error is not identifiable in practice (Wickramasuriya et al., 2019).
+
+-   (Talk more about how others transform it to OLS, WLS,..)
+-   Di Fonzo and Marini (2011)
+-   Athanasopoulos et al. (2009)
+
+Wickramasuriya, Athanasopoulos & Hyndman (2019) reframed the problem by taking an optimisation approach rather than the regression.
+They formulated the problem as minimising the variances of all reconciled forecasts, which happens to be equivalent to minimising the trace of the covariance matrix (sum of the diagonal elements).
+This is known as the Minimum Trace (MinT) reconciliation method.
+The MinT solution is given by $G_h = (S'W_h^{-1}S)^{-1}S'W_h^{-1}$, and $W_h$ is the covariance matrix of the h-step-ahead base forecast errors.
+
+The MinT approach is an algebraical generalisation of the GLS, and the OLS and WLS methods are special cases of MinT when $W_h$ is a diagonal or identity matrix, respectively.
+However, the MinT solution hinges on a reliable estimate of the h-step-ahead base forecast error covariance $W_h$.
+In high-dimensional setting, the usual sample covariance matrix is unstable, thus we need alternative covariance estimators.
+
+-   Structural Scaling, based only on the struc- ture of the hierarchy (Athanasopoulos et al., 2017)
+-   Shrinkage estimators (Schäfer & Strimmer, 2005; Ledoit & Wolf, 2004)
+
+Empirical evaluations have demonstrated that MinT with an appropriate covariance estimate often outperforms earlier methods in both simulation and real data studies.
+
+Wickramasuriya et al.
+(2019) adopt the diagonal-target Ledoit–Wolf approach and show analytically that any scalar multiple $k_h$ from $W_h = k_hW_1$ suffices for point-forecast reconciliation.
+However, simulation and empirical evidence suggest that ignoring cross-series error correlation can leave accuracy untapped, especially in highly-coupled systems (energy, retail SKUs).
+Their subsequent work imposes non-negativity constraints via quadratic programming citeturn0file2 and studies bias-robust variants (MinT-U).
+
+Empirical findings to date
+
+Athanasopoulos et al.
+(2024) synthesise the accumulated evidence: MinT dominates single-level baselines in most studies; WLS or diagonal-shrinkage MinT is the current industry default; but gains plateau when off-diagonals are weak citeturn0file9.
+Only a handful of papers explore richer covariance structures (e.g., block or factor shrinkage), and none benchmark them systematically against NOVELIST-type estimators.
+
+## Covariance Estimation in High Dimensions
+
+-   Limitations of the sample covariance matrix.
+-   Estimators used by Wickramasuriya et al. (2019).
+-   Shrinkage estimators:
+    -   Diagonal shrinkage (e.g., Schäfer & Strimmer, Ledoit & Wolf).
+    -   NOVELIST estimator and its Cross-validation & PC-adjusted variant.
+    -   ...
+
+## Relevance to Forecast Reconciliation
+
+-   Discuss how covariance estimation affects MinT performance.
+-   Identify research gaps.
+-->
+
 # Evaluating ex- and include 2016 onwards on Tourism

### 2025-10-07 literature/paper_draft.qmd
move old to archived

@@
 
-<!--
--   **Problem Statement:**
-
-    The sample covariance matrix, although natural, suffers in high-dimensional settings.
-Especially when the number of series $p$ is huge and larger than the time dimension $T$, the sample covariance matrix is non-positive definite (rank T if p\>T).
-
-    The shrinkage estimators come in to tackle this issue.
-The shrinkage estimator with diagonal target (often shrinking toward a diagonal matrix) is proven to produce a guaranteed PD matrix (Schäfer & Strimmer, 2005).
-However, as it shrinks the covariance matrix toward a diagonal one, it does not have flexibility and might neglect the prominent structure presented in the covariance matrix.
-
-    An alternative approach is to perform shrinkage of the sample covariance towards its thresholded version, instead of a diagonal matrix.
-This is the NOVELIST (NOVEL Integration of the Sample and Thresholded covariance estimators) method proposed by Huang & Fryzlewicz (2019).
-They introduced thresholding functions applied only to off-diagonal elements, allowing for more flexibility in the estimation.
-
-    [... can include more estimators ...]{style="color:red;"}
-
--   **Research Aim:**
-
-    This paper assesses the reconciled forecasting performance of MinT approach using various covariance estimators, with a focus on the NOVELIST estimator.
-
--   **Paper Outline:**
-
-    The paper is structured as follows:
-
-    -   A literature review of forecast reconciliation and covariance estimation.
-    -   A description of the methodology, including the NOVELIST estimator and its principal-component-adjusted variant.
-    -   An experimental design using both synthetic and real hierarchical time series.
-    -   Empirical results and discussion.
-    -   Conclusions and suggestions for future work.
--->
-
-<!--
-# Literature Review
-
-## Forecast Reconciliation in Hierarchical and Grouped Time Series
-
-Forecast reconciliation converts a collection of independent base forecasts into a set of coherent forecasts that respect the linear constraints defining a hierarchical or grouped time-series system.
-Early work focused on heuristic single-level strategies, including bottom-up, top-down, and middle-out (...), each of which exploits only part of the information in the hierarchy and can induce bias or high variance.
-
--   Cite the single level approach
--   Athanasopoulos et al., 2024
-
-Hyndman et al.
-(2011) first showed that all single-level methods can be written as $\tilde{y} = SG\hat{y}$, where $S$ is the summing matrix and $G$ is a matrix that maps base forecasts $\hat{y}$ to into the bottom level.
-Treating reconciliation as a GLS regression problem, Hyndman et al.
-(2011) found that it yields a solution for $G$, but the required covariance matrix of reconcilation error is not identifiable in practice (Wickramasuriya et al., 2019).
-
--   (Talk more about how others transform it to OLS, WLS,..)
--   Di Fonzo and Marini (2011)
--   Athanasopoulos et al. (2009)
-
-Wickramasuriya, Athanasopoulos & Hyndman (2019) reframed the problem by taking an optimisation approach rather than the regression.
-They formulated the problem as minimising the variances of all reconciled forecasts, which happens to be equivalent to minimising the trace of the covariance matrix (sum of the diagonal elements).
-This is known as the Minimum Trace (MinT) reconciliation method.
-The MinT solution is given by $G_h = (S'W_h^{-1}S)^{-1}S'W_h^{-1}$, and $W_h$ is the covariance matrix of the h-step-ahead base forecast errors.
-
-The MinT approach is an algebraical generalisation of the GLS, and the OLS and WLS methods are special cases of MinT when $W_h$ is a diagonal or identity matrix, respectively.
-However, the MinT solution hinges on a reliable estimate of the h-step-ahead base forecast error covariance $W_h$.
-In high-dimensional setting, the usual sample covariance matrix is unstable, thus we need alternative covariance estimators.
-
--   Structural Scaling, based only on the struc- ture of the hierarchy (Athanasopoulos et al., 2017)
--   Shrinkage estimators (Schäfer & Strimmer, 2005; Ledoit & Wolf, 2004)
-
-Empirical evaluations have demonstrated that MinT with an appropriate covariance estimate often outperforms earlier methods in both simulation and real data studies.
-
-Wickramasuriya et al.
-(2019) adopt the diagonal-target Ledoit–Wolf approach and show analytically that any scalar multiple $k_h$ from $W_h = k_hW_1$ suffices for point-forecast reconciliation.
-However, simulation and empirical evidence suggest that ignoring cross-series error correlation can leave accuracy untapped, especially in highly-coupled systems (energy, retail SKUs).
-Their subsequent work imposes non-negativity constraints via quadratic programming citeturn0file2 and studies bias-robust variants (MinT-U).
-
-Empirical findings to date
-
-Athanasopoulos et al.
-(2024) synthesise the accumulated evidence: MinT dominates single-level baselines in most studies; WLS or diagonal-shrinkage MinT is the current industry default; but gains plateau when off-diagonals are weak citeturn0file9.
-Only a handful of papers explore richer covariance structures (e.g., block or factor shrinkage), and none benchmark them systematically against NOVELIST-type estimators.
-
-## Covariance Estimation in High Dimensions
-
--   Limitations of the sample covariance matrix.
--   Estimators used by Wickramasuriya et al. (2019).
--   Shrinkage estimators:
-    -   Diagonal shrinkage (e.g., Schäfer & Strimmer, Ledoit & Wolf).
-    -   NOVELIST estimator and its Cross-validation & PC-adjusted variant.
-    -   ...
-
-## Relevance to Forecast Reconciliation
-
--   Discuss how covariance estimation affects MinT performance.
--   Identify research gaps.
--->
-
 # Theoretical Framework {#sec-pre-theory}
@@
 | NOVELIST                                | MinT-N             |
-| PC-adjusted Shrinkage with *k* PCs      | MinT-S(PC*k*)      |
-| PC-adjusted NOVELIST with *k* PCs       | MinT-N(PC*k*)      |
+| PC-adjusted Shrinkage with *K* PCs      | MinT-S(PC*K*)      |
+| PC-adjusted NOVELIST with *K* PCs       | MinT-N(PC*K*)      |
 | Scaled Variance Shrinkage               | MinT-S(sv)         |

### 2025-10-07 literature/paper_draft.qmd
iterate theorectical section

@@
 
-## Hierarchical structure {#sec-hierarchy}
+## Hierarchical Time Series {#sec-hierarchy}
 
@@
 to be revised with probabilistic framework
+
+- start with hierarchical time series
+  - mention hier and grouped example (but stick to the general notation)
+- talk about S, containing A
+  - example of S
+- coherent subspace
+- coherent point forecasts
+  - projection
+- coherent probabilistic forecasts
+  - definition
+  - image
+
+*Hierarchical time series* are multivariate time series $\boldsymbol{y}_t \in \mathbb{R}^n$ organised in a structure where the series adheres to some *constraints*.
+For example, @fig-hierarchy illustrates a simple 2-level hierarchical structure with one top-level series, two middle-level series, and four bottom-level series.
+Here, the constraints imply that the four bottom-level series (or the most disaggregated series) must sum up to the two middle-level series, which in turn sum up to the top-level series (or the most aggregated series).
+
+The bottom-level series are denoted as $\boldsymbol{b}_t \in \mathbb{R}^{n_b}$. Thus, the hierarchical time series can be represented as:
+
+$$
+\boldsymbol{y}_t = \boldsymbol{S} \boldsymbol{b}_t,
+$$
+
+where $\boldsymbol{S} \in \mathbb{R}^{n \times n_b}$ is a summing matrix that aggregates the bottom-level series to all-level series.
+The summing matrix $\boldsymbol{S}$ for the tree structure in @fig-hierarchy is:
+
+$$
+\boldsymbol{S} = 
+\left[
+\begin{array}{cccc}
+1 & 1 & 1 & 1 \\
+1 & 1 & 0 & 0 \\
+0 & 0 & 1 & 1 \\
+\multicolumn{4}{c}{\boldsymbol{I_4}}
+\end{array}
+\right].
+$$
+
+The matrix $\boldsymbol{S}$ encodes the aggregation constraints presented in the structure.
+Hence, the columns of $\boldsymbol{S}$ span a linear subspace.
+Any observation $\boldsymbol{y}_t$ that lies inside this subspace is called *coherent*, while those outside are *incoherent*.
+We refer to the subspace spanned by $\boldsymbol{S}$ as the *coherent subspace* $\mathfrak{s} \in \mathbb{R}^{n_b}$.
+
+This setting is not restricted to hierarchical (nested) structures.
+When there are attributes of interest that are crossed, such as the company sales at any aggregation level (company-wise, city-wise, or outlet-wise) is also considered by kinds of products, the structure is described as a *grouped structure*.
+As illustrated in @fig-grouped, the aggregation or disaggregation paths are not unique.
+These constraints formed by the grouped structure can also be represented using a summing matrix $\boldsymbol{S}$.
+For simplicity, we refer to both of these structures as hierarchical structure, we will distinguish between them if and when it is necessary.
+
+When we produce forecasts for each individual series, referred to as *base forecasts* $\hat{\boldsymbol{y}}_{t+h|t}$, they often do not respect the aggregation constraints, and thus are incoherent.
+Coherency can be achieved by linearly projecting the base forecasts onto the coherent subspace $\mathfrak{s}$ using a projection matrix $\boldsymbol{P}$: $\tilde{\boldsymbol{y}}_{t+h|t} = \boldsymbol{P} \hat{\boldsymbol{y}}_{t+h|t}$, where $\tilde{\boldsymbol{y}}_{t+h|t}$ are the *reconciled forecasts*.
+
+- Probabilistic reconciliation
+
+- Definition from @Panagiotelis2023-fs?
+
+Many existing reconciliation methods including the OLS [@Hyndman2011-jv], WLS [@Hyndman2016-ic], and MinT [@Wickramasuriya2019-xq] express the projection matrix as $\boldsymbol{P} = \boldsymbol{S} \boldsymbol{G}$, for a suitable $n_b \times n$ matrix $\boldsymbol{G}$.
+The idea is to map the base forecasts of all levels $\hat{\boldsymbol{y}}_{t+h|t}$ down into the bottom level, which is then aggregated to the higher levels by $\boldsymbol{S}$.
+The choice of $\boldsymbol{G}$ determines the composition of reconciled forecasts $\tilde{\boldsymbol{y}}_{t+h|t}$, and modern reconciliation methods are developed to estimate $\boldsymbol{G}$.
+
 \color{black}
 
+<!-- 
 The *hierarchical structure* can be represented as a tree, as shown in @fig-hierarchy.
@@
 <!-- including the OLS and WLS developed by @Hyndman2011-jv and @Hyndman2016-ic, respectively. A more detailed explanation of these methods can be found in the @sec-ols-wls Appendix. -->
+ -->
 

### 2025-10-08 literature/paper_draft.qmd
write scoring rules

@@
         \usepackage{float}
+        \usepackage{xcolor}
     mathspec: true
@@
 
+\color{black}
+
 *Hierarchical time series* are multivariate time series $\boldsymbol{y}_t \in \mathbb{R}^n$ organised in a structure where the series adheres to some *constraints*.
@@
 Here, the constraints imply that the four bottom-level series (or the most disaggregated series) must sum up to the two middle-level series, which in turn sum up to the top-level series (or the most aggregated series).
+
+![A 2-level hierarchical tree structure](D:/Github/Recon_Honours_Thesis/figs/hierarchical_structure.png){#fig-hierarchy width="40%"}
 
@@
 The choice of $\boldsymbol{G}$ determines the composition of reconciled forecasts $\tilde{\boldsymbol{y}}_{t+h|t}$, and modern reconciliation methods are developed to estimate $\boldsymbol{G}$.
-
-\color{black}
 
@@
 
-## Scoring Metric {-}
-
-\color{red}
-to be included...
-\color{black}
+## Probabilistic Scoring Rules {-}
+
+We use the mean squared error (MSE) to evaluate the accuracy of point forecasts.
+Meanwhile, to assess the quality of the probabilistic forecasts, it is common to use scoring rules.
+A scoring rule is a function $S(., .)$ taking a predictive distribution as its first argument and a realisation as its second argument, then returns a numerical score.
+We follow a convention that lower scores are better.
+A scoring rule is said to be *proper* if $\mathbb{E}_Q[S(Q, y)] \leq \mathbb{E}_Q[S(P, y)]$ for all $P$, where $P$ is a predictive distribution produced by forecasting model, $Q$ is the true distribution of the realisation $y$, and $\mathbb{E}_Q$ is the expectation with respect to $Q$.
+Hence, the expected score is minimised when the forecast distribution matches the true distribution.
+
+We employ Winkler score and continuous ranked probability score as our univariate scoring rules, and energy score as the multivariate scoring rule. All three are proper scoring rules.
+
+\textcolor{blue}{\textit{Winkler score (WS).}} If the 100$(1-\alpha)$\% prediction interval is $[l, u]$ (the $\alpha / 2$ and $1 - $\alpha / 2$ quantiles), then the Winkler score is defined as:
+
+$$
+WS_{\alpha}(l, u; y) = (u - l) + \frac{2}{\alpha} (l - y) \boldsymbol{1}(y < l) + \frac{2}{\alpha} (y - u) \boldsymbol{1}(y > u),
+$$
+
+where $y$ is the observed value, and $\boldsymbol{1}(.)$ is the indicator function.
+The Winkler score rewards narrow intervals that contain the observation, and penalises intervals that do not contain the observation.
+
+\textcolor{blue}{\textit{Continuous ranked probability score (CRPS).}} The CRPS is defined as the squared difference between the predictive cumulative distribution function (CDF) $P$ and the empirical CDF of the observation $y$:
+
+$$
+CRPS(P, y) = \int_{-\infty}^{\infty} (P(x) - \boldsymbol{1}(x \geq y))^2 dx.
+$$
+
+When the predictive distribution is Gaussian with mean $\mu$ and standard deviation $\sigma$, the CRPS has a closed-form expression:
+
+$$
+CRPS(P, y) = \sigma \left[ z \left( 2 \Phi(z) -1 \right) + 2 \phi(z) - \frac{1}{\sqrt{\pi}} \right],
+$$
+
+where $z = \frac{y - \mu}{\sigma}$, and $\Phi(.)$ and $\phi(.)$ are the CDF and probability density function (PDF) of a standard normal distribution, respectively.
+
+\textcolor{blue}{\textit{Energy score (ES).}} The energy score is a multivariate generalisation of the CRPS. It is defined as:
+
+$$
+ES(P, \boldsymbol{y}) = \mathbb{E}_P ||\boldsymbol{X} - \boldsymbol{y}||^{\beta} - \frac{1}{2} \mathbb{E}_P ||\boldsymbol{X} - \boldsymbol{X}'||^{\beta},
+$$
+
+where $\boldsymbol{X}$ and $\boldsymbol{X}'$ are independent random vectors with distribution $P$, $\boldsymbol{y}$ is the observed vector, $||.||$ is the Euclidean norm, and $\beta \in (0, 2]$.
+We set $\beta = 1$ following common convention.
+
+Since the closed-form expression of the ES may not available, we approximate it using Monte Carlo samples $\{\boldsymbol{x}_1, \ldots, \boldsymbol{x}_M\}$ drawn from $P$:
+
+$$
+\widehat{ES}(P, \boldsymbol{y}) = \frac{1}{M} \sum_{m=1}^M ||\boldsymbol{x}_m - \boldsymbol{y}|| - \frac{1}{2M(M-1)} \sum_{m=1}^M ||\boldsymbol{x}_m - \boldsymbol{x}^*_m||,
+$$
+
+where $\boldsymbol{x}^*_m$ is a randomly selected sample from $\{\boldsymbol{x}_1, \ldots, \boldsymbol{x}_M\} \setminus \{\boldsymbol{x}_m\}$.
+In our experiments, we use $M=10000$ samples to approximate the ES.
 

### 2025-10-09 literature/archived.qmd
rearragne

@@
 This motivates our subsequent exploration of sparse covariance structures, in which many off-diagonals are truly zero.
+
+<!-- 
+\pagebreak
+
+# Timeline & Milestones {#sec-timeline}
+
+```{r timeline}
+#| echo: false
+#| message: false
+#| warning: false
+library(tibble)
+library(knitr)
+library(kableExtra)
+
+tibble(
+  Period = c(
+    "March",
+    "April", 
+    "May", 
+    "June",
+    "July",
+    "August", 
+    "September",
+    "October"
+  ),
+  Task = c(
+    "Initial setup and understanding the research scope. Preliminary knowledge and Literature review",
+
+    "Develop ReconCov, an R package implementing the Min Trace approach with the shrinkage and NOVELIST estimators, and rolling-window cross-validation process for choosing NOVELIST threshold parameter.
+Exploring simulation framework.",
+
+    "Code functions for producing simulation settings in ReconCov package. Initial simulation for MinT in hierarchical time series. Writing research proposal and presenting research project.",
+
+    "Intensive simulation on multiple different settings, in order to assess the performance and behaviours of NOVELIST under different conditions, and to explore cases where NOVELIST performs different than Shrinkage.",
+
+    "Empirical analysis on forecasting Australian Domestic Tourism data. Explore the real-world data generating process which can be used to refine the current simulation design",
+
+    "Implement the Tourism DGP configuration into the current simulation settings.
+Implementing POET and PC-adjusted NOVELIST estimators.
+Assessing whether prominent principal components structure can be exploited to improve the out-of-sample forecast performance.",
+
+    "Exploring other high-dimensional covariance estimators. More simulations on all explored covariance estimators. Final,  full-fledged assessment on all implemented estimators",
+
+    "Final writing, presenation, submission"
+  ),
+  Deliverable = c(
+    "Initial literature review + paper draft",
+    "ReconCov package: Main codes for estimators + simulation framework + cross-validation",
+    "NOVELIST initial simulation results + research proposal + presentation",
+    "More simulation results on NOVELIST",
+    "Empirical results on forecasting Tourism data",
+    "New simulation design, POET & PC-adjusted NOVELIST simulation results",
+    "Other estimators simulation results and full-fledged covariance estimators assessment",
+    "Presentation and thesis manuscript"
+  )
+) |>
+  kable(
+    format = "latex",
+    booktabs = TRUE,
+    longtable = TRUE,
+    caption = "Project Timeline and Key Milestones",
+    escape = TRUE
+  ) |>
+  column_spec(1, width = "1.9cm") |>
+  column_spec(2, width = "8cm") |>
+  column_spec(3, width = "4.8cm") |>
+  kable_styling(latex_options = "hold_position", font_size = 10)
+```
+-->

### 2025-10-09 literature/paper_draft.qmd
rearragne

@@
 
-# Simulation {#sec-simulation}
-
-## General Design Framework {-}
-
-The general design of data generating process for bottom-level series is a stationary VAR(1) process, with the following structure:
-
-$$
-\boldsymbol{b}_t = \boldsymbol{A} \boldsymbol{b}_{t-1} + \boldsymbol{\epsilon}_t,
-$$
-
-where $\boldsymbol{A}$ is a $n_b \times n_b$ block diagonal matrix of autoregressive coefficients $\boldsymbol{A} = diag(\boldsymbol{A}_1, \ldots, \boldsymbol{A}_m)$, with each $\boldsymbol{A}_i$ being a $n_{b,i} \times n_{b,i}$ matrix.
-The block diagonal structure ensures that the time series are grouped into $m$ groups, with each group having its own autoregressive coefficients.
-This aim to simulate the interdependencies between the time series within each group, where reconciliation will be expected to better performed than the usual base forecasts.
-
-The model is added with a Gaussian innovation process $\boldsymbol{\epsilon}_t$, with covariance matrix $\boldsymbol{\Sigma}$.
-The covariance matrix $\boldsymbol{\Sigma}$ is generated specifically using the Algorithm 1 in @Hardin2013-wu:
-
-1.
-A compound symmetric correlation matrix is used for each block of size $n_{b,i}$ in $\boldsymbol{A}_i$, where the entries $\rho_i$ for each block $i$ are sampled from a uniform distribution between 0 and 1.
-They are baseline correlations within group.
-
-2.  A constant correlation, which is smaller than $\text{min} \{\rho_1, \rho_2, \dots, \rho_m \}$, is imposed on the entries between different blocks. It serves as baseline correlations between group.
-
-3.  The entry-wise random noise is added on top of the entire correlation matrix.
-
-4.  The covariance matrix $\boldsymbol{\Sigma}$ is then constructed by uniform sampling of standard deviations, in a range of $[\sqrt{2}, \sqrt{6}]$, for all $n_b$ series.
-
-We will randomly flip the signs of the covariance elements, which will create a more realistic structure in the innovation process.
-This can be done by pre- and post-multiplying $\boldsymbol{\Sigma}$ by a random diagonal matrix $\boldsymbol{V}$ with diagonal entries sampled from $\{-1, 1\}$, yielding $\boldsymbol{\Sigma}^* = \boldsymbol{V} \boldsymbol{\Sigma} \boldsymbol{V}$.
-
-For all hierarchies in our experiments, we simulate two panel lengths, $T=54$ and $T=304$, reserving the final four observations as an out-of-sample test set.
-In each Monte Carlo replication ($M=500$), we fit univariate ARIMA models (base models) to the training observations using an automatic AICc minimization algorithm from @Hyndman2008-rf, implemented in the *fabletools* package [@O-Hara-Wild2024-we], generating incoherent 1–4-step base forecasts.
-We then reconcile these forecasts under three covariance estimators: the raw sample covariance (mint_sample), the shrinkage estimator (mint_shr), and the NOVELIST estimator (mint_n).
-
-All data generation, covariance estimation, and reconciliation routines were implemented in the ReconCov R package and is available under an open‐source license on GitHub [@ReconCov].
-
-```{r load results}
-path_sim <- "D:/Github/Recon_Honours_Analysis/sim/thesis_sim/"
-
-sim1_1 <- readRDS("./literature/lib/S4-2-1_T50_M500.rds")
-sim1_2 <- readRDS("./literature/lib/S4-2-1_T300_M500.rds")
-
-sim2_1 <- readRDS("./literature/lib/S36-6-1_T50_M500.rds")
-sim2_2 <- readRDS("./literature/lib/S36-6-1_T300_M500.rds")
-
-sim_dense_50 <- readRDS("./literature/lib/S100-10-3-1_T50_M200_dense.rds")
-sim_dense_300 <- readRDS("./literature/lib/S100-10-3-1_T300_M200_dense.rds")
-
-sim_sparse_50 <- readRDS("./literature/lib/S100-10-3-1_T50_M500_sparsebwgrp.rds")
-sim_sparse_300 <- readRDS("./literature/lib/S100-10-3-1_T300_M200_sparsebwgrp.rds")
-
-sim3_1 <- sim_dense_50 # using same with next section simulation
-sim3_2 <- sim_dense_300
-```
-
-## Probabilistic Scoring Rules {-}
+# Probabilistic Scoring Rules {#sec-scores}
 
@@
 In our experiments, we use $M=10000$ samples to approximate the ES.
+
+# Simulation {#sec-simulation}
+
+## General Design Framework {-}
+
+The general design of data generating process for bottom-level series is a stationary VAR(1) process, with the following structure:
+
+$$
+\boldsymbol{b}_t = \boldsymbol{A} \boldsymbol{b}_{t-1} + \boldsymbol{\epsilon}_t,
+$$
+
+where $\boldsymbol{A}$ is a $n_b \times n_b$ block diagonal matrix of autoregressive coefficients $\boldsymbol{A} = diag(\boldsymbol{A}_1, \ldots, \boldsymbol{A}_m)$, with each $\boldsymbol{A}_i$ being a $n_{b,i} \times n_{b,i}$ matrix.
+The block diagonal structure ensures that the time series are grouped into $m$ groups, with each group having its own autoregressive coefficients.
+This aim to simulate the interdependencies between the time series within each group, where reconciliation will be expected to better performed than the usual base forecasts.
+
+The model is added with a Gaussian innovation process $\boldsymbol{\epsilon}_t$, with covariance matrix $\boldsymbol{\Sigma}$.
+The covariance matrix $\boldsymbol{\Sigma}$ is generated specifically using the Algorithm 1 in @Hardin2013-wu:
+
+1.
+A compound symmetric correlation matrix is used for each block of size $n_{b,i}$ in $\boldsymbol{A}_i$, where the entries $\rho_i$ for each block $i$ are sampled from a uniform distribution between 0 and 1.
+They are baseline correlations within group.
+
+2.  A constant correlation, which is smaller than $\text{min} \{\rho_1, \rho_2, \dots, \rho_m \}$, is imposed on the entries between different blocks. It serves as baseline correlations between group.
+
+3.  The entry-wise random noise is added on top of the entire correlation matrix.
+
+4.  The covariance matrix $\boldsymbol{\Sigma}$ is then constructed by uniform sampling of standard deviations, in a range of $[\sqrt{2}, \sqrt{6}]$, for all $n_b$ series.
+
+We will randomly flip the signs of the covariance elements, which will create a more realistic structure in the innovation process.
+This can be done by pre- and post-multiplying $\boldsymbol{\Sigma}$ by a random diagonal matrix $\boldsymbol{V}$ with diagonal entries sampled from $\{-1, 1\}$, yielding $\boldsymbol{\Sigma}^* = \boldsymbol{V} \boldsymbol{\Sigma} \boldsymbol{V}$.
+
+For all hierarchies in our experiments, we simulate two panel lengths, $T=54$ and $T=304$, reserving the final four observations as an out-of-sample test set.
+In each Monte Carlo replication ($M=500$), we fit univariate ARIMA models (base models) to the training observations using an automatic AICc minimization algorithm from @Hyndman2008-rf, implemented in the *fabletools* package [@O-Hara-Wild2024-we], generating incoherent 1–4-step base forecasts.
+We then reconcile these forecasts under three covariance estimators: the raw sample covariance (mint_sample), the shrinkage estimator (mint_shr), and the NOVELIST estimator (mint_n).
+
+All data generation, covariance estimation, and reconciliation routines were implemented in the ReconCov R package and is available under an open‐source license on GitHub [@ReconCov].
+
+```{r load results}
+path_sim <- "D:/Github/Recon_Honours_Analysis/sim/thesis_sim/"
+
+sim1_1 <- readRDS("./literature/lib/S4-2-1_T50_M500.rds")
+sim1_2 <- readRDS("./literature/lib/S4-2-1_T300_M500.rds")
+
+sim2_1 <- readRDS("./literature/lib/S36-6-1_T50_M500.rds")
+sim2_2 <- readRDS("./literature/lib/S36-6-1_T300_M500.rds")
+
+sim_dense_50 <- readRDS("./literature/lib/S100-10-3-1_T50_M200_dense.rds")
+sim_dense_300 <- readRDS("./literature/lib/S100-10-3-1_T300_M200_dense.rds")
+
+sim_sparse_50 <- readRDS("./literature/lib/S100-10-3-1_T50_M500_sparsebwgrp.rds")
+sim_sparse_300 <- readRDS("./literature/lib/S100-10-3-1_T300_M200_sparsebwgrp.rds")
+
+sim3_1 <- sim_dense_50 # using same with next section simulation
+sim3_2 <- sim_dense_300
+```
 
@@
 
-<!-- 
-\pagebreak
-
-# Timeline & Milestones {#sec-timeline}
-
-```{r timeline}
-#| echo: false
-#| message: false
-#| warning: false
-library(tibble)
-library(knitr)
-library(kableExtra)
-
-tibble(
-  Period = c(
-    "March",
-    "April", 
-    "May", 
-    "June",
-    "July",
-    "August", 
-    "September",
-    "October"
-  ),
-  Task = c(
-    "Initial setup and understanding the research scope. Preliminary knowledge and Literature review",
-
-    "Develop ReconCov, an R package implementing the Min Trace approach with the shrinkage and NOVELIST estimators, and rolling-window cross-validation process for choosing NOVELIST threshold parameter.
-Exploring simulation framework.",
-
-    "Code functions for producing simulation settings in ReconCov package. Initial simulation for MinT in hierarchical time series. Writing research proposal and presenting research project.",
-
-    "Intensive simulation on multiple different settings, in order to assess the performance and behaviours of NOVELIST under different conditions, and to explore cases where NOVELIST performs different than Shrinkage.",
-
-    "Empirical analysis on forecasting Australian Domestic Tourism data. Explore the real-world data generating process which can be used to refine the current simulation design",
-
-    "Implement the Tourism DGP configuration into the current simulation settings.
-Implementing POET and PC-adjusted NOVELIST estimators.
-Assessing whether prominent principal components structure can be exploited to improve the out-of-sample forecast performance.",
-
-    "Exploring other high-dimensional covariance estimators. More simulations on all explored covariance estimators. Final,  full-fledged assessment on all implemented estimators",
-
-    "Final writing, presenation, submission"
-  ),
-  Deliverable = c(
-    "Initial literature review + paper draft",
-    "ReconCov package: Main codes for estimators + simulation framework + cross-validation",
-    "NOVELIST initial simulation results + research proposal + presentation",
-    "More simulation results on NOVELIST",
-    "Empirical results on forecasting Tourism data",
-    "New simulation design, POET & PC-adjusted NOVELIST simulation results",
-    "Other estimators simulation results and full-fledged covariance estimators assessment",
-    "Presentation and thesis manuscript"
-  )
-) |>
-  kable(
-    format = "latex",
-    booktabs = TRUE,
-    longtable = TRUE,
-    caption = "Project Timeline and Key Milestones",
-    escape = TRUE
-  ) |>
-  column_spec(1, width = "1.9cm") |>
-  column_spec(2, width = "8cm") |>
-  column_spec(3, width = "4.8cm") |>
-  kable_styling(latex_options = "hold_position", font_size = 10)
-```
--->
-
 \pagebreak

### 2025-10-09 literature/paper_draft.qmd
refine tourism section

@@
 
-```{=html}
 <!-- 
@@
 -->
-```
 
@@
 
-@fig-tourism-MSE-line and @fig-tourism-MSE-heat show that reconciliation is beneficial for point forecasts: across horizons $h=1,2,\dots,12$, all reconciled methods improve MSE over incoherent base ARIMA on average.
+<!-- @fig-tourism-MSE-line and @fig-tourism-MSE-heat show that reconciliation is beneficial for point forecasts: across horizons $h=1,2,\dots,12$, all reconciled methods improve MSE over incoherent base ARIMA on average.
 Among vanilla MinT variants, the shrinkage estimator (*MinT-S*) marginally outperforms NOVELIST (*MinT-N*) at most horizons, though the gap is small.
@@
 Variants that modify the multi-step covariance, either via scaled-variance or direct h-step residual covariances, underperform standard MinT, suggesting that extra estimation at horizon $h>1$ is not rewarded in this setting.
+-->
+
+@fig-tourism-MSE-line show a main difference compared to the one of simulation results.
+Adjusting for a single dominant factor via PC decomposition (dashed line with dot points) tightens performance further: both *MinT-S(PC1)* and *MinT-N(PC1)* beat their unadjusted counterparts.
+Other than that, similar patterns are observed: among vanilla MinT variants, the shrinkage estimator (*MinT-S*) slightly outperforms NOVELIST (*MinT-N*) at most horizons; adding more than one PC brings no additional benefit, likely due to injecting estimation noise from weaker components; variants that modify the multi-step covariance, either via scaled-variance or direct h-step residual covariances, underperform standard MinT, suggesting that extra estimation at horizon $h>1$ is not rewarded in this empirical analysis.
+
+@fig-tourism-MSE-heat provides a complementary heatmap view, scaling each method's MSE improvement to a 0-100 range for better visual discrimination.
+Here, *MinT-N(PC1)* and *MinT-S(PC1)* emerge as the top performers across most horizons.
+The heatmap underscores the consistent gains from PC adjustment and highlights the diminishing returns from more complex covariance treatments.
 
@@
 
-Turning to probabilistic forecasts (1-step-ahead forecasts), @fig-tourism-probscores shows that *MinT-N* consistently outperforms *MinT-S* across CRPS and both Winkler scores, and PC1-adjustment again yields the best improvements.
+<!-- Turning to probabilistic forecasts (1-step-ahead forecasts), @fig-tourism-probscores shows that *MinT-N* consistently outperforms *MinT-S* across CRPS and both Winkler scores, and PC1-adjustment again yields the best improvements.
 Results are aligned across probabilistic scoring rules, but the 95% Winkler reveals a notable pattern, as all MinT variants lose to the base while *OLS* performs best.
 In the multivariate evaluation, the Energy score places *OLS* close to the PC1-adjusted MinT methods, indicating that simple coherent projection might be able to capture a substantial share of cross-series dependence benefits even without sophisticated covariance regularisation.
+-->
+
+Turning to probabilistic forecasts (1-step-ahead forecasts), @fig-tourism-probscores shows that *MinT-N* (purple bars) consistently outperforms *MinT-S* (green bars) across univariate and multivariate scores.
+The PC1-adjusted variants again yield improvements over their vanilla counterparts, with *MinT-N(PC1)* leading overall.
+In the multivariate evaluation, the Energy score places *OLS* close to the PC1-adjusted MinT methods.
+One surprising finding is that all MinT variants underperform the base forecasts when moving to the 95% Winkler score, while *OLS* performs best.
 
@@
 
-To further dissect the 95% Winkler score results, @fig-tourism-coverage-95-level breaks down performance by hierarchical level, and examines empirical coverage of the 95% prediction intervals.
-The left panel shows that *MinT-S* underperforms the base at all levels, especially in the higher aggregated levels.
-This is due to the overly shrunk variances from the shrinkage estimator, leading to narrow prediction intervals and thus high Winkler penalties when observations fall outside.
+To dissect the 95% Winkler score results, @fig-tourism-coverage-95-level breaks down performance by hierarchical level, and examines empirical coverage of the 95% prediction intervals.
+The left panel shows that *MinT-S* underperforms the base at all levels, especially in higher aggregated levels.
+From our inspection, this is due to the overly shrunk variances from the shrinkage estimator, leading to narrow prediction intervals and thus high Winkler penalties when observations fall outside.
 The *MinT-N* method has relatively good coverage and improve the Winkler score at bottom levels, but still underperforms at higher levels.
 The PC-adjusted variants seem to strike a better balance, improving overall coverage and relative Winkler scores over the base.
-
-The summary radar graph in @fig-tourism-radar consolidates these findings: among the selected MinT models by probabilistic criteria, *MinT-N(PC1)* clearly leads across CRPS, W80, W95, and Energy, extending the improvements seen in the single-metric panels.
-
-Taken together, the evidence supports the use of reconciliation for both point and probabilistic forecasts in this high-dimensional setting.
-For point forecasts, MinT with shrinkage is a solid default; for probabilistic forecasts, NOVELIST performs more reliably.
-PC adjustment with a single dominant factor consistently enhances performance and should be considered when a dominant latent factor is evident.
-More complex adjustments, such as multiple PCs or horizon-specific covariances, add estimation noise without clear benefits and can be omitted for parsimony.
 
@@
 The scores are scaled to a range of 0 to 100, with larger values indicating better performance.
+The outermost polygon represents the best possible score (100) and the innermost polygon represents the worst possible score (0).
 Only the top 5 models are shown."
@@
 ```
+
+The summary radar graph in @fig-tourism-radar consolidates these findings: among the selected MinT models by probabilistic criteria, *MinT-N(PC1)* (the yellow polygon) clearly leads across CRPS, W80, W95, and Energy, extending the improvements seen in the single-metric panels.
+
+Taken together, the evidence supports the use of reconciliation for both point and probabilistic forecasts in this high-dimensional setting.
+For point forecasts, MinT with shrinkage is a solid default; for probabilistic forecasts, NOVELIST performs more reliably.
+PC adjustment with a single dominant factor consistently enhances performance in both point and probabilistic forecast and should be considered when a dominant latent factor is evident.
+More complex adjustments, such as multiple PCs or horizon-specific covariances, add estimation noise without clear benefits and can be omitted for parsimony.
 

### 2025-10-09 literature/paper_draft.qmd
slight fix + reorganise figs

@@
 $$
-\hat{\boldsymbol{R}}^{g, K}_{1} = \sum_{k = 1}^K \hat{\gamma}_k \hat{\boldsymbol{\xi}}_k \hat{\boldsymbol{\xi}}_k' + g(\hat{\boldsymbol{W}}^K_{1})
+\hat{\boldsymbol{W}}^{g, K}_{1} = \sum_{k = 1}^K \hat{\gamma}_k \hat{\boldsymbol{\xi}}_k \hat{\boldsymbol{\xi}}_k' + g(\hat{\boldsymbol{W}}^K_{1})
 $$
@@
 where $g(.)$ is either the shrinkage or NOVELIST estimator, $\hat{\gamma}_k$ and $\hat{\boldsymbol{\xi}}_k$ are the $k$-th largest eigenvalue and the corresponding eigenvector of the sample covariance matrix, respectively.
-Similar to the NOVELIST estimator, $\hat{\boldsymbol{R}}^{*, K}_{1}$ is not guaranteed to be positive definite.
-
+Similar to the NOVELIST estimator, its PC-adjusted variant $\hat{\boldsymbol{W}}^{N, K}_{1}$ requires a cross-validation procedure and adjustment to obtain positive definiteness.
 ## Scaled Variance
@@
 
-Another approach is not to rely on the assumption of proportionality at all, and directly estimate the covariance matrix from the h-step-ahead base forecast errors:
+Another alternative is not to rely on the assumption of proportionality at all, and directly estimate the covariance matrix from the h-step-ahead base forecast errors:
 
@@
 The concrete lines represent the vanilla *MinT-S* and *MinT-N*; the dashed lines with dot points denote the PC-adjusted variants (e.g.
-*MinT-S(PC1), *MinT-N(PC2)*); and the dotted or dashed-dotted lines indicate the scaled variance and h-step-ahead residuals versions (e.g.
-*MinT-S(SV), MinT-N(hcov)*).
+*MinT-S(PC1)*, *MinT-N(PC2)*); and the dotted or dashed-dotted lines indicate the scaled variance and h-step-ahead residuals versions (e.g.
+*MinT-S(SV)*, *MinT-N(hcov)*).
 

### 2025-10-13 literature/paper_draft.qmd
expand on rolling window tourism with fig

@@
 
-To assess forecasting performance, we adopt a rolling-window cross-validation scheme.
+![Rolling-window cross-validation scheme for evaluating forecasting performance in Australia tourism data](D:/Github/Recon_Honours_Thesis/figs/tourism_cv.png){#fig-tourism-cv width="70%"}
+
+To assess forecasting performance between models, we adopt a rolling-window cross-validation scheme.
 Beginning with the first 120 monthly observations (January 1998-December 2005) as the initial training set, we obtain the best-fitted ARIMA model for each of the 560 series via the automatic algorithm by minimising AICc from @Hyndman2008-rf, implemented in the *fabletools* package [@O-Hara-Wild2024-we].
 The 1- to 12-step-ahead base forecasts are then generated by these ARIMA models, and then reconciled using multiple approaches.
+To estimate the NOVELIST and its variants, we would have an extra cross-validation prodecure within this training window, as described in @sec-novelist_cv.
 We then roll the training window forward by one month and refit all models, rebuild reconciliations, and produce another batch of 1- to 12-step-ahead forecasts, repeating until the training set reaches December 2015.
 In total, this results in 97 out-of-sample windows.
+The entire procedure is illustrated in @fig-tourism-cv.
 

### 2025-10-13 literature/archived.qmd
expand on theorectical section, projection and link to MinT

@@
+# Old literature review (archived)
+
 <!--
@@
 -   Identify research gaps.
+-->
+
+# Old preliminary section
+
+<!-- 
+The *hierarchical structure* can be represented as a tree, as shown in @fig-hierarchy.
+The top level of the tree represents the total value of all series, while the lower levels represent the series at different levels of disaggregation.
+These hierarchical structures naturally form *aggregation constraints* (child levels must sum up to parents).
+When there are attributes of interest that are crossed, such as the Starbucks drinks sales at any aggregation level (brand-wise, national, or outlet) is also considered by kinds of drinks (e.g., coffees, refreshers), the structure is described as a grouped time series.
+As illustrated in @fig-grouped, the aggregation or disaggregation paths are not unique.
+
+![A 2-level hierarchical tree structure](D:/Github/Recon_Honours_Thesis/figs/hierarchical_structure.png){#fig-hierarchy width="40%"}
+
+![A 2-level grouped structure, which can be considered as the union of two hierarchical trees with common top and bottom level series](D:/Github/Recon_Honours_Thesis/figs/grouped_structure.png){#fig-grouped width="90%"}
+
+For simplicity, we refer to both of these structures as hierarchical structure, we will distinguish between them if and when it is necessary.
+Taken together, a collection of time series organised in a hierarchy and subject to aggregation constraints is refered as *hierarchical time series*.
+All hierarchical time series can be represented using matrix algebra:
+
+$$
+\boldsymbol{y}_t = \boldsymbol{S} \boldsymbol{b}_t,
+$$
+
+where $\boldsymbol{S}$ is a summing matrix of order $n \times n_b$ which aggregates the bottom-level series $\boldsymbol{b}_t \in \mathbb{R}^{n_b}$ to the series at aggregation levels above.
+The vector $\boldsymbol{y}_t \in \mathbb{R}^n$ contains all observations at time $t$.
+The summing matrix $\boldsymbol{S}$ for the tree structure in @fig-hierarchy is:
+
+$$
+\boldsymbol{S} = 
+\left[
+\begin{array}{cccc}
+1 & 1 & 1 & 1 \\
+1 & 1 & 0 & 0 \\
+0 & 0 & 1 & 1 \\
+\multicolumn{4}{c}{\boldsymbol{I_4}}
+\end{array}
+\right].
+$$
+
+Assume we produce $h$-step-ahead base forecasts $\hat{\boldsymbol{b}}_{t+h|t}$ for the bottom-level series, obtained by any prediction methods, pre-multiplying them by $\boldsymbol{S}$ we get:
+
+$$
+\tilde{\boldsymbol{y}}_{t+h|t} = \boldsymbol{S} \hat{\boldsymbol{b}}_{t+h|t} \, .
+$$ {#eq-sb}
+
+We refer to $\tilde{\boldsymbol{y}}_{t+h|t}$ as *coherent forecasts*, as they respect the aggregation constraints.
+We also refer to this way of obtaining coherent forecasts by summing the bottom-level forecasts as the bottom-up approach.
+However, generating forecasts this way is anchored only to prediction models at a single level, and will not be utilising the inherent information from other levels.
+This drawback applies to the top-down and middle-out approaches.
+For example, the bottom-level data can be very noisy or even intermittent, and the higher-level data might be smoother due to the aggregation.
+
+Another issue with expressing reconciled methods as in @eq-sb is that it restricts the reconciliation to only single-level approaches.
+Thus, @Hyndman2011-jv suggested a generalised expression for all existing methods, which also provides a framework for new methods to be developed:
+
+$$
+\tilde{\boldsymbol{y}}_{t+h|t} = \boldsymbol{S} \boldsymbol{G} \hat{\boldsymbol{y}}_{t+h|t} \; ,
+$$ {#eq-sgy}
+
+for a suitable $n_b \times n$ matrix $\boldsymbol{G}$.
+$\boldsymbol{G}$ maps the base forecasts of all levels $\hat{\boldsymbol{y}}_{t+h|t}$ down into the bottom level, which is then aggregated to the higher levels by $\boldsymbol{S}$.
+The choice of $\boldsymbol{G}$ determines the composition of reconciled forecasts $\tilde{\boldsymbol{y}}_{t+h|t}$, and modern reconciliation methods are developed to estimate $\boldsymbol{G}$.
+
+including the OLS and WLS developed by @Hyndman2011-jv and @Hyndman2016-ic, respectively. A more detailed explanation of these methods can be found in the @sec-ols-wls Appendix.
+
 -->

### 2025-10-13 literature/paper_draft.qmd
expand on theorectical section, projection and link to MinT

@@
 *Hierarchical time series* are multivariate time series $\boldsymbol{y}_t \in \mathbb{R}^n$ organised in a structure where the series adheres to some *constraints*.
-For example, @fig-hierarchy illustrates a simple 2-level hierarchical structure with one top-level series, two middle-level series, and four bottom-level series.
-Here, the constraints imply that the four bottom-level series (or the most disaggregated series) must sum up to the two middle-level series, which in turn sum up to the top-level series (or the most aggregated series).
+For example, @fig-hierarchy illustrates a simple 2-level hierarchical structure with one top-level series $y_{Tot,t}$, two middle-level series $(y_{A,t}, y_{B,t})'$, and four bottom-level series $(y_{A1,t}, y_{A2,t}, y_{B1,t}, y_{B2,t})'$.
+Here, the *aggregation constraints* imply that $y_{Tot,t} = y_{A,t} + y_{B,t}$, $y_{A,t} = y_{A1,t} + y_{A2,t}$, and $y_{B,t} = y_{B1,t} + y_{B2,t}$.
 
-![A 2-level hierarchical tree structure](D:/Github/Recon_Honours_Thesis/figs/hierarchical_structure.png){#fig-hierarchy width="40%"}
+![A 2-level hierarchical tree structure](../figs/hierarchical_structure.png){#fig-hierarchy width="40%"}
 
-The bottom-level series are denoted as $\boldsymbol{b}_t \in \mathbb{R}^{n_b}$. Thus, the hierarchical time series can be represented as:
+The bottom-level (or most disaggregated) series are denoted as $\boldsymbol{b}_t \in \mathbb{R}^{n_b}$. Thus, the hierarchical time series can be represented as:
 
@@
 
-where $\boldsymbol{S} \in \mathbb{R}^{n \times n_b}$ is a summing matrix that aggregates the bottom-level series to all-level series.
+where $\boldsymbol{S} \in \mathbb{R}^{n \times n_b}$ is a summing matrix that aggregates the bottom-level to all-level series.
 The summing matrix $\boldsymbol{S}$ for the tree structure in @fig-hierarchy is:
@@
 
+![A 2-level grouped structure, which can be considered as the union of two hierarchical trees with common top and bottom level series](../figs/grouped_structure.png){#fig-grouped width="90%"}
+
 This setting is not restricted to hierarchical (nested) structures.
@@
 
+<!-- There are other forms of structures, such as temporal hierarchies or non-linear constraints, but they are beyond the scope of this paper.  -->
+
 When we produce forecasts for each individual series, referred to as *base forecasts* $\hat{\boldsymbol{y}}_{t+h|t}$, they often do not respect the aggregation constraints, and thus are incoherent.
 Coherency can be achieved by linearly projecting the base forecasts onto the coherent subspace $\mathfrak{s}$ using a projection matrix $\boldsymbol{P}$: $\tilde{\boldsymbol{y}}_{t+h|t} = \boldsymbol{P} \hat{\boldsymbol{y}}_{t+h|t}$, where $\tilde{\boldsymbol{y}}_{t+h|t}$ are the *reconciled forecasts*.
+A schematic illustration of this projection is depicted in @fig-geometry.
 
-- Probabilistic reconciliation
+![Geometry of probabilistic forecast reconciliation.
+The eliptical base forecast distribution is projected orthogonally onto the coherent subspace (purple line), resulting in the reconciled forecast distribution (red).
+The projection is defined by the projection matrix $\boldsymbol{P}$, and MinT allows oblique projections.
+Note that this figure is schematic since most applications are high-dimensional.](../figs/geometry_orthogonal.png){#fig-geometry width="60%"}
 
-- Definition from @Panagiotelis2023-fs?
+Many existing reconciliation methods including the OLS [@Hyndman2011-jv], WLS [@Hyndman2016-ic], and MinT [@Wickramasuriya2019-xq] express the projection matrix as $\boldsymbol{P} = \boldsymbol{S} \boldsymbol{G}$, for a suitable $n_b \times n$ mapping matrix $\boldsymbol{G}$.
+The idea is to map the base forecasts of all levels $\hat{\boldsymbol{y}}_{t+h|t}$ down into the bottom level, which is then aggregated to the higher levels by $\boldsymbol{S}$.
+Since the projection matrix $\boldsymbol{P}$ is idempotent and symmetric, $\boldsymbol{G}$ must satisfy the condition $\boldsymbol{S} \boldsymbol{G} \boldsymbol{S} = \boldsymbol{S}$.
+Thus, we generally have the mapping matrix $\boldsymbol{G} = (\boldsymbol{S}' \boldsymbol{M}^{-1} \boldsymbol{S})^{-1} \boldsymbol{S}' \boldsymbol{M}^{-1}$, for some positive definite matrix $\boldsymbol{M}$.
 
-Many existing reconciliation methods including the OLS [@Hyndman2011-jv], WLS [@Hyndman2016-ic], and MinT [@Wickramasuriya2019-xq] express the projection matrix as $\boldsymbol{P} = \boldsymbol{S} \boldsymbol{G}$, for a suitable $n_b \times n$ matrix $\boldsymbol{G}$.
-The idea is to map the base forecasts of all levels $\hat{\boldsymbol{y}}_{t+h|t}$ down into the bottom level, which is then aggregated to the higher levels by $\boldsymbol{S}$.
-The choice of $\boldsymbol{G}$ determines the composition of reconciled forecasts $\tilde{\boldsymbol{y}}_{t+h|t}$, and modern reconciliation methods are developed to estimate $\boldsymbol{G}$.
-
-<!-- 
-The *hierarchical structure* can be represented as a tree, as shown in @fig-hierarchy.
-The top level of the tree represents the total value of all series, while the lower levels represent the series at different levels of disaggregation.
-These hierarchical structures naturally form *aggregation constraints* (child levels must sum up to parents).
-When there are attributes of interest that are crossed, such as the Starbucks drinks sales at any aggregation level (brand-wise, national, or outlet) is also considered by kinds of drinks (e.g., coffees, refreshers), the structure is described as a grouped time series.
-As illustrated in @fig-grouped, the aggregation or disaggregation paths are not unique.
-
-![A 2-level hierarchical tree structure](D:/Github/Recon_Honours_Thesis/figs/hierarchical_structure.png){#fig-hierarchy width="40%"}
-
-![A 2-level grouped structure, which can be considered as the union of two hierarchical trees with common top and bottom level series](D:/Github/Recon_Honours_Thesis/figs/grouped_structure.png){#fig-grouped width="90%"}
-
-For simplicity, we refer to both of these structures as hierarchical structure, we will distinguish between them if and when it is necessary.
-Taken together, a collection of time series organised in a hierarchy and subject to aggregation constraints is refered as *hierarchical time series*.
-All hierarchical time series can be represented using matrix algebra:
-
-$$
-\boldsymbol{y}_t = \boldsymbol{S} \boldsymbol{b}_t,
-$$
-
-where $\boldsymbol{S}$ is a summing matrix of order $n \times n_b$ which aggregates the bottom-level series $\boldsymbol{b}_t \in \mathbb{R}^{n_b}$ to the series at aggregation levels above.
-The vector $\boldsymbol{y}_t \in \mathbb{R}^n$ contains all observations at time $t$.
-The summing matrix $\boldsymbol{S}$ for the tree structure in @fig-hierarchy is:
-
-$$
-\boldsymbol{S} = 
-\left[
-\begin{array}{cccc}
-1 & 1 & 1 & 1 \\
-1 & 1 & 0 & 0 \\
-0 & 0 & 1 & 1 \\
-\multicolumn{4}{c}{\boldsymbol{I_4}}
-\end{array}
-\right].
-$$
-
-Assume we produce $h$-step-ahead base forecasts $\hat{\boldsymbol{b}}_{t+h|t}$ for the bottom-level series, obtained by any prediction methods, pre-multiplying them by $\boldsymbol{S}$ we get:
-
-$$
-\tilde{\boldsymbol{y}}_{t+h|t} = \boldsymbol{S} \hat{\boldsymbol{b}}_{t+h|t} \, .
-$$ {#eq-sb}
-
-We refer to $\tilde{\boldsymbol{y}}_{t+h|t}$ as *coherent forecasts*, as they respect the aggregation constraints.
-We also refer to this way of obtaining coherent forecasts by summing the bottom-level forecasts as the bottom-up approach.
-However, generating forecasts this way is anchored only to prediction models at a single level, and will not be utilising the inherent information from other levels.
-This drawback applies to the top-down and middle-out approaches.
-For example, the bottom-level data can be very noisy or even intermittent, and the higher-level data might be smoother due to the aggregation.
-
-Another issue with expressing reconciled methods as in @eq-sb is that it restricts the reconciliation to only single-level approaches.
-Thus, @Hyndman2011-jv suggested a generalised expression for all existing methods, which also provides a framework for new methods to be developed:
-
-$$
-\tilde{\boldsymbol{y}}_{t+h|t} = \boldsymbol{S} \boldsymbol{G} \hat{\boldsymbol{y}}_{t+h|t} \; ,
-$$ {#eq-sgy}
-
-for a suitable $n_b \times n$ matrix $\boldsymbol{G}$.
-$\boldsymbol{G}$ maps the base forecasts of all levels $\hat{\boldsymbol{y}}_{t+h|t}$ down into the bottom level, which is then aggregated to the higher levels by $\boldsymbol{S}$.
-The choice of $\boldsymbol{G}$ determines the composition of reconciled forecasts $\tilde{\boldsymbol{y}}_{t+h|t}$, and modern reconciliation methods are developed to estimate $\boldsymbol{G}$.
-
-<!-- including the OLS and WLS developed by @Hyndman2011-jv and @Hyndman2016-ic, respectively. A more detailed explanation of these methods can be found in the @sec-ols-wls Appendix. -->
- -->
+When setting $\boldsymbol{M} = \boldsymbol{I}_n$, the identity matrix, we obtain the OLS method, which also corresponds to an orthogonal projection onto the coherent subspace.
+Research have been done to introduce $\boldsymbol{M}$ that better utilise the inherent information from the observed data, resulting in oblique projections to deliver better-performing reconciled forecasts.
 
@@
 
-@Wickramasuriya2019-xq framed the problem as minimising the variances of all reconciled forecast errors $\text{Var}[y_{t+h} - \tilde{y}_{t+h|t}] = \boldsymbol{S} \boldsymbol{G} \boldsymbol{W}_h \boldsymbol{G}' \boldsymbol{S}'$, where $\boldsymbol{W}_h = \mathbb{E}( \hat{\boldsymbol{e}}_{t+h|t} \; \hat{\boldsymbol{e}}_{t+h|t}' )$ is the positive definite covariance matrix of the $h$-step-ahead base forecast errors.
+- we link to MinT saying that when M is the covariance matrix of base forecast errors, minimise the trace of the reconciled forecast error covariance matrix
+
+@Wickramasuriya2019-xq showed that by setting $\boldsymbol{M} = \boldsymbol{W}_h = \mathbb{E}( \hat{\boldsymbol{e}}_{t+h|t} \; \hat{\boldsymbol{e}}_{t+h|t}' )$, the covariance matrix of the $h$-step-ahead base forecast errors $\hat{\boldsymbol{e}}_{t+h|t} = \boldsymbol{y}_{t+h} - \hat{\boldsymbol{y}}_{t+h|t}$, we essentially minimise the total variance of the reconciled forecast errors across all series.
+In which, it is minimising the trace of $\text{Var}[y_{t+h} - \tilde{y}_{t+h|t}] = \boldsymbol{S} \boldsymbol{G} \boldsymbol{W}_h \boldsymbol{G}' \boldsymbol{S}'$.
+This method is called Minimum Trace (MinT) reconciliation.
+The matrix $\boldsymbol{G}$ is thus given by:
+
+$$
+\boldsymbol{G} = (\boldsymbol{S}' \boldsymbol{W}_h^{-1} \boldsymbol{S})^{-1}
+\boldsymbol{S}' \boldsymbol{W}_h^{-1} .
+$$
+
+<!-- @Wickramasuriya2019-xq framed the problem as minimising the variances of all reconciled forecast errors $\text{Var}[y_{t+h} - \tilde{y}_{t+h|t}] = \boldsymbol{S} \boldsymbol{G} \boldsymbol{W}_h \boldsymbol{G}' \boldsymbol{S}'$, where $\boldsymbol{W}_h = \mathbb{E}( \hat{\boldsymbol{e}}_{t+h|t} \; \hat{\boldsymbol{e}}_{t+h|t}' )$ is the positive definite covariance matrix of the $h$-step-ahead base forecast errors.
 They showed that this is equivalent to minimising the trace of the reconciled forecast error covariance matrix (sum of the diagonal elements - the variances).
@@
 In this paper, we place our main focus on the MinT method.
+-->
 

### 2025-10-13 literature/archived.qmd
expand on MinT, gaussian framework, minor changes to shrinkage notation

@@
 
+-->
+
+<!-- @Wickramasuriya2019-xq framed the problem as minimising the variances of all reconciled forecast errors $\text{Var}[y_{t+h} - \tilde{y}_{t+h|t}] = \boldsymbol{S} \boldsymbol{G} \boldsymbol{W}_h \boldsymbol{G}' \boldsymbol{S}'$, where $\boldsymbol{W}_h = \mathbb{E}( \hat{\boldsymbol{e}}_{t+h|t} \; \hat{\boldsymbol{e}}_{t+h|t}' )$ is the positive definite covariance matrix of the $h$-step-ahead base forecast errors.
+They showed that this is equivalent to minimising the trace of the reconciled forecast error covariance matrix (sum of the diagonal elements - the variances).
+The Minimum Trace (MinT) solution is given by
+
+$$
+\boldsymbol{G} = (\boldsymbol{S}' \boldsymbol{W}_h^{-1} \boldsymbol{S})^{-1}
+\boldsymbol{S}' \boldsymbol{W}_h^{-1} .
+$$
+
+@Wickramasuriya2019-xq also showed that MinT is an algebraic generalisation of the GLS, and the OLS and WLS methods are special cases of MinT when $\boldsymbol{W}_h$ is an identity matrix $I_{n_b}$ and a diagonal matrix $\text{diag}( \boldsymbol{W}_h )$, respectively.
+In this paper, we place our main focus on the MinT method.
 -->

### 2025-10-13 literature/paper_draft.qmd
expand on MinT, gaussian framework, minor changes to shrinkage notation

@@
 
-\color{red}
-to be revised with probabilistic framework
-
-- start with hierarchical time series
-  - mention hier and grouped example (but stick to the general notation)
-- talk about S, containing A
-  - example of S
-- coherent subspace
-- coherent point forecasts
-  - projection
-- coherent probabilistic forecasts
-  - definition
-  - image
-
-\color{black}
-
 *Hierarchical time series* are multivariate time series $\boldsymbol{y}_t \in \mathbb{R}^n$ organised in a structure where the series adheres to some *constraints*.
@@
 Since the projection matrix $\boldsymbol{P}$ is idempotent and symmetric, $\boldsymbol{G}$ must satisfy the condition $\boldsymbol{S} \boldsymbol{G} \boldsymbol{S} = \boldsymbol{S}$.
-Thus, we generally have the mapping matrix $\boldsymbol{G} = (\boldsymbol{S}' \boldsymbol{M}^{-1} \boldsymbol{S})^{-1} \boldsymbol{S}' \boldsymbol{M}^{-1}$, for some positive definite matrix $\boldsymbol{M}$.
+Thus, we generally have the mapping matrix $\boldsymbol{G} = (\boldsymbol{S}' \boldsymbol{M}^{-1} \boldsymbol{S})^{-1} \boldsymbol{S}' \boldsymbol{M}^{-1}$, for some positive definite matrix $\boldsymbol{M}$ [@Gamakumara2020-ml].
 
-When setting $\boldsymbol{M} = \boldsymbol{I}_n$, the identity matrix, we obtain the OLS method, which also corresponds to an orthogonal projection onto the coherent subspace.
-Research have been done to introduce $\boldsymbol{M}$ that better utilise the inherent information from the observed data, resulting in oblique projections to deliver better-performing reconciled forecasts.
+When setting $\boldsymbol{M} = \boldsymbol{I}_n$, the identity matrix, we obtain the OLS method, which also corresponds to an orthogonal projection onto the coherent subspace (similar to @fig-geometry).
+Research have been done to introduce $\boldsymbol{M}$ that better utilise the inherent information from the observed data, to construct oblique projections and deliver better-performing forecasts.
 
@@
 
-- we link to MinT saying that when M is the covariance matrix of base forecast errors, minimise the trace of the reconciled forecast error covariance matrix
-
 @Wickramasuriya2019-xq showed that by setting $\boldsymbol{M} = \boldsymbol{W}_h = \mathbb{E}( \hat{\boldsymbol{e}}_{t+h|t} \; \hat{\boldsymbol{e}}_{t+h|t}' )$, the covariance matrix of the $h$-step-ahead base forecast errors $\hat{\boldsymbol{e}}_{t+h|t} = \boldsymbol{y}_{t+h} - \hat{\boldsymbol{y}}_{t+h|t}$, we essentially minimise the total variance of the reconciled forecast errors across all series.
-In which, it is minimising the trace of $\text{Var}[y_{t+h} - \tilde{y}_{t+h|t}] = \boldsymbol{S} \boldsymbol{G} \boldsymbol{W}_h \boldsymbol{G}' \boldsymbol{S}'$.
+In which, it is minimising the trace of $\text{Var}[y_{t+h} - \tilde{y}_{t+h|t}] = \boldsymbol{S} \boldsymbol{G}_h \boldsymbol{W}_h \boldsymbol{G}_h' \boldsymbol{S}'$.
 This method is called Minimum Trace (MinT) reconciliation.
-The matrix $\boldsymbol{G}$ is thus given by:
+The matrix $\boldsymbol{G}_h$ is thus given by:
 
 $$
-\boldsymbol{G} = (\boldsymbol{S}' \boldsymbol{W}_h^{-1} \boldsymbol{S})^{-1}
+\boldsymbol{G}_h = (\boldsymbol{S}' \boldsymbol{W}_h^{-1} \boldsymbol{S})^{-1}
 \boldsymbol{S}' \boldsymbol{W}_h^{-1} .
@@
 
-<!-- @Wickramasuriya2019-xq framed the problem as minimising the variances of all reconciled forecast errors $\text{Var}[y_{t+h} - \tilde{y}_{t+h|t}] = \boldsymbol{S} \boldsymbol{G} \boldsymbol{W}_h \boldsymbol{G}' \boldsymbol{S}'$, where $\boldsymbol{W}_h = \mathbb{E}( \hat{\boldsymbol{e}}_{t+h|t} \; \hat{\boldsymbol{e}}_{t+h|t}' )$ is the positive definite covariance matrix of the $h$-step-ahead base forecast errors.
-They showed that this is equivalent to minimising the trace of the reconciled forecast error covariance matrix (sum of the diagonal elements - the variances).
-The Minimum Trace (MinT) solution is given by
+This is initially derived under point forecast reconciliation setting.
+However, @Wickramasuriya2024-mb showed that under the Gaussian base forecast distribution assumption, MinT also yields a Gaussian reconciled forecast distribution.
+Let $h$-step-ahead base forecast distribution be $\hat{\boldsymbol{y}}_{t+h|t} \sim \mathcal{N}(\hat{\boldsymbol{y}}_{t+h|t}, \boldsymbol{W}_h)$, then the reconciled forecast distribution is given by $\tilde{\boldsymbol{y}}_{t+h|t} \sim \mathcal{N}(\tilde{\boldsymbol{y}}_{t+h|t}, \boldsymbol{S} \boldsymbol{G}_h \boldsymbol{W}_h \boldsymbol{G}_h' \boldsymbol{S}')$.
+Additionally, MinT also minimises the logarithmic score of the reconciled predictive distribution among all projection matrices.
+
+For probabilistic forecasts, this paper focuses on evaluating the performance of MinT with different covariance estimators under the Gaussian framework.
+Even so, it is worth to mention that the methods can be extended to non-Gaussian settings by bootstrapping [@Gamakumara2020-ml; @Panagiotelis2023-fs].
+
+## Shrinkage Estimator for MinT
+
+The MinT solution hinges on a reliable, positive-definite estimate of $\boldsymbol{W}_h$, which comes in both the mapping matrix $\boldsymbol{G}_h$ and the reconciled forecast variance $\boldsymbol{S} \boldsymbol{G}_h \boldsymbol{W}_h \boldsymbol{G}_h' \boldsymbol{S}'$.
+
+However, the covariance matrix $\boldsymbol{W}_h$ is often not available in theoretical derivation, and is challenging to estimate in high-dimensional setting where the number of series $n$ is larger than the time dimension $T$.
+To tackle this issue, the original paper @Wickramasuriya2019-xq assumed a proportionality relationship between $\hat{\boldsymbol{W}}^g_h = k_h g( \hat{\boldsymbol{W}}_1 )$, where $\hat{\boldsymbol{W}}_1$ is the covariance matrix of the in-sample 1-step-ahead base residuals (to approximate $\boldsymbol{W}_1$) and $k_h$ is a positive scaling constant (which will be cancelled out in point-forecast reconciliation).
+The function $g(.)$ is a covariance estimator that produces a positive-definite matrix, the main focus of this paper.
+
+The authors suggested using the shrinkage estimator with diagonal target from @Schafer2005-yw, given by:
 
 $$
-\boldsymbol{G} = (\boldsymbol{S}' \boldsymbol{W}_h^{-1} \boldsymbol{S})^{-1}
-\boldsymbol{S}' \boldsymbol{W}_h^{-1} .
+\hat{\boldsymbol{W}}^{S}_{1} = \lambda_S \, \text{diag}( \hat{\boldsymbol{W}}_1 ) + (1 - \lambda_S) \hat{\boldsymbol{W}}_1 \, ,
 $$
 
-@Wickramasuriya2019-xq also showed that MinT is an algebraic generalisation of the GLS, and the OLS and WLS methods are special cases of MinT when $\boldsymbol{W}_h$ is an identity matrix $I_{n_b}$ and a diagonal matrix $\text{diag}( \boldsymbol{W}_h )$, respectively.
-In this paper, we place our main focus on the MinT method.
--->
+where $\text{diag}( \hat{\boldsymbol{W}}_1 )$ comprises only the diagonal elements of $\hat{\boldsymbol{W}}_1$.
+We refer to any $\lambda_S \in [0,1]$ as the shrinkage intensity parameter, the subscript specifies the shrinkage estimator it belongs to.
+This approach shrinks the covariance matrix $\hat{\boldsymbol{W}}_1$ towards its diagonal matrix, meaning the off-diagonal elements are shrunk towards zero while the diagonal ones remain unchanged.
 
-The MinT solution hinges on a reliable, positive-definite estimate of $\boldsymbol{W}_h$, which is often not available and challenging to estimate in high-dimensional setting.
-The sample covariance matrix is unstable and non-positive-definite when the number of series $n$ is huge and larger than the time dimension $T$.
-To tackle this issue, the original paper @Wickramasuriya2019-xq assumed a proportionality relationship between $\hat{\boldsymbol{W}}^g_h = k_h g( \hat{\boldsymbol{W}}_1 )$, where $\hat{\boldsymbol{W}}_1$ is the covariance matrix of the in-sample 1-step-ahead base residuals (to approximate $\boldsymbol{W}_1$), $k_h$ is a positive scaling constant (which will be cancelled out in point-forecast reconciliation), and $g(.)$ is a covariance estimator that produces a positive-definite matrix.
-They suggested using the shrinkage estimator with diagonal target from @Schafer2005-yw, given by:
+@Schafer2005-yw also proposed an estimate of the optimal shrinkage intensity parameter $\lambda_S$:
 
 $$
-\hat{\boldsymbol{W}}^{S}_{1} = \lambda_D \hat{\boldsymbol{W}}_{1, D} + (1 - \lambda_D) \hat{\boldsymbol{W}}_1 \, ,
-$$
-
-where $\hat{\boldsymbol{W}}_{1, D}$ is a diagonal matrix comprising the diagonal entries $\text{diag}( \hat{\boldsymbol{W}}_1 )$.
-We refer to any $\lambda \in [0,1]$ as the shrinkage intensity parameter, the subscript specifies which estimator it belongs to.
-This approach shrinks the covariance matrix $\hat{\boldsymbol{W}}_1$ towards its diagonal matrix, meaning the off-diagonal elements are shrunk towards zero while the diagonal ones remain unchanged.
-
-@Schafer2005-yw also proposed an estimate of the optimal shrinkage intensity parameter $\lambda_D$:
-
-$$
-\hat{\lambda}_D = \frac{\sum_{i \neq j} \widehat{Var}(\hat{r}_{ij})} 
+\hat{\lambda}_S = \frac{\sum_{i \neq j} \widehat{Var}(\hat{r}_{ij})} 
 {\sum_{i \neq j} \hat{r}_{ij}^2} \, ,
@@
 
-The shrinkage estimator shrinks all off-diagonal elements towards zeros with equal weights $\lambda_D$.
+The shrinkage estimator shrinks all off-diagonal elements towards zeros with equal weights $\lambda_S$.
 We might prefer to better preserve strong signals, and largely reduce the effects of small, noisy correlations.
@@
 ## Summary of MinT with Covariance Estimators {-}
+
+The covariance estimators explored in this paper are summarised in the table below. The abbreviations will be used in the following sections.
 

### 2025-10-13 literature/paper_draft.qmd
expand of evaluation and scoring rules, use subscript i for univar score

@@
 
-## Constructing from h-step-ahead residuals
+## Constructing from h-step-ahead Residuals
 
@@
 
-# Probabilistic Scoring Rules {#sec-scores}
+# Evaluation of Point and Probabilistic Forecasts {#sec-scores}
 
-We use the mean squared error (MSE) to evaluate the accuracy of point forecasts.
-Meanwhile, to assess the quality of the probabilistic forecasts, it is common to use scoring rules.
+This section briefly introduces the scoring rules used to evaluate the point and probabilistic forecast accuracy in @sec-simulation and @sec-empirical.
+
+For point forecasts, we use the mean squared error (MSE) to evaluate the accuracy of different reconciliation methods: $MSE = \frac{1}{n} \sum_{i=1}^n (y_{i, t+h} - \tilde{y}_{i,t+h|t})^2$, where $y_{i, t+h}$ is the realised value of series $i$ at time $t+h$, and $\tilde{y}_{i,t+h|t}$ is the reconciled point forecast.
+
+Meanwhile, to assess the quality of the probabilistic forecasts, it is common to use proper scoring rules.
 A scoring rule is a function $S(., .)$ taking a predictive distribution as its first argument and a realisation as its second argument, then returns a numerical score.
 We follow a convention that lower scores are better.
-A scoring rule is said to be *proper* if $\mathbb{E}_Q[S(Q, y)] \leq \mathbb{E}_Q[S(P, y)]$ for all $P$, where $P$ is a predictive distribution produced by forecasting model, $Q$ is the true distribution of the realisation $y$, and $\mathbb{E}_Q$ is the expectation with respect to $Q$.
+A scoring rule is said to be *proper* if $\mathbb{E}_Q[S(Q, y)] \leq \mathbb{E}_Q[S(F, y)]$ for all $F$, where $F$ is a predictive distribution produced by forecasting model, $Q$ is the true distribution of the realisation $y$, and $\mathbb{E}_Q$ is the expectation with respect to $Q$.
 Hence, the expected score is minimised when the forecast distribution matches the true distribution.
 
-We employ Winkler score and continuous ranked probability score as our univariate scoring rules, and energy score as the multivariate scoring rule. All three are proper scoring rules.
+We employ Winkler score and continuous ranked probability score as our univariate scoring rules, and energy score as the multivariate scoring rule.
+All three are proper scoring rules.
+In this paper, we only evaluate $1$-step-ahead probabilistic forecasts, and thus we drop the subscript $t$ and $h$ for simplicity.
 
-\textcolor{blue}{\textit{Winkler score (WS).}} If the 100$(1-\alpha)$\% prediction interval is $[l, u]$ (the $\alpha / 2$ and $1 - $\alpha / 2$ quantiles), then the Winkler score is defined as:
+\textcolor{blue}{\textit{Winkler score (WS).}} If the 100$(1-\alpha)$\% prediction interval of $i$-th series is $[l_i, u_i]$ (the $\alpha / 2$ and $1 - \alpha / 2$ quantiles), then the Winkler score is defined as:
 
 $$
-WS_{\alpha}(l, u; y) = (u - l) + \frac{2}{\alpha} (l - y) \boldsymbol{1}(y < l) + \frac{2}{\alpha} (y - u) \boldsymbol{1}(y > u),
+WS_{\alpha}(l_i, u_i; y_i) = (u_i - l_i) + \frac{2}{\alpha} (l_i - y_i) \boldsymbol{1}(y_i < l_i) + \frac{2}{\alpha} (y_i - u_i) \boldsymbol{1}(y_i > u_i),
 $$
 
-where $y$ is the observed value, and $\boldsymbol{1}(.)$ is the indicator function.
+where $y_i$ is the observed value of $i$-th series, and $\boldsymbol{1}(.)$ is the indicator function.
 The Winkler score rewards narrow intervals that contain the observation, and penalises intervals that do not contain the observation.
 
-\textcolor{blue}{\textit{Continuous ranked probability score (CRPS).}} The CRPS is defined as the squared difference between the predictive cumulative distribution function (CDF) $P$ and the empirical CDF of the observation $y$:
+\textcolor{blue}{\textit{Continuous ranked probability score (CRPS).}} The CRPS is defined as the squared difference between the predictive cumulative distribution function (CDF) $F_i$ and the empirical CDF of the observation $y_i$ of series $i$:
 
 $$
-CRPS(P, y) = \int_{-\infty}^{\infty} (P(x) - \boldsymbol{1}(x \geq y))^2 dx.
+CRPS(F_i, y_i) = \int_{-\infty}^{\infty} (F_i(x) - \boldsymbol{1}(x \geq y_i))^2 dx.
 $$
 
-When the predictive distribution is Gaussian with mean $\mu$ and standard deviation $\sigma$, the CRPS has a closed-form expression:
+When the predictive distribution is Gaussian with mean $\mu_i$ and standard deviation $\sigma_i$, the CRPS has a closed-form expression:
 
 $$
-CRPS(P, y) = \sigma \left[ z \left( 2 \Phi(z) -1 \right) + 2 \phi(z) - \frac{1}{\sqrt{\pi}} \right],
+CRPS(F_i, y_i) = \sigma_i \left[ z_i \left( 2 \Phi(z_i) -1 \right) + 2 \phi(z_i) - \frac{1}{\sqrt{\pi}} \right],
 $$
 
-where $z = \frac{y - \mu}{\sigma}$, and $\Phi(.)$ and $\phi(.)$ are the CDF and probability density function (PDF) of a standard normal distribution, respectively.
+where $z_i = \frac{y_i - \mu_i}{\sigma_i}$, and $\Phi(.)$ and $\phi(.)$ are the CDF and probability density function (PDF) of a standard normal distribution, respectively.
 
@@
 $$
-ES(P, \boldsymbol{y}) = \mathbb{E}_P ||\boldsymbol{X} - \boldsymbol{y}||^{\beta} - \frac{1}{2} \mathbb{E}_P ||\boldsymbol{X} - \boldsymbol{X}'||^{\beta},
+ES(F, \boldsymbol{y}) = \mathbb{E}_F ||\boldsymbol{X} - \boldsymbol{y}||^{\beta} - \frac{1}{2} \mathbb{E}_F ||\boldsymbol{X} - \boldsymbol{X}'||^{\beta},
 $$
 
-where $\boldsymbol{X}$ and $\boldsymbol{X}'$ are independent random vectors with distribution $P$, $\boldsymbol{y}$ is the observed vector, $||.||$ is the Euclidean norm, and $\beta \in (0, 2]$.
+where $\boldsymbol{X}$ and $\boldsymbol{X}'$ are independent random vectors with multivariate distribution $F$, $\boldsymbol{y}$ is the observed vector, $||.||$ is the Euclidean norm, and $\beta \in (0, 2]$.
 We set $\beta = 1$ following common convention.
@@
 $$
-\widehat{ES}(P, \boldsymbol{y}) = \frac{1}{M} \sum_{m=1}^M ||\boldsymbol{x}_m - \boldsymbol{y}|| - \frac{1}{2M(M-1)} \sum_{m=1}^M ||\boldsymbol{x}_m - \boldsymbol{x}^*_m||,
+\widehat{ES}(F, \boldsymbol{y}) = \frac{1}{M} \sum_{m=1}^M ||\boldsymbol{x}_m - \boldsymbol{y}|| - \frac{1}{2M(M-1)} \sum_{m=1}^M ||\boldsymbol{x}_m - \boldsymbol{x}^*_m||,
 $$

### 2025-10-13 thesis/paper_draft.Rmd
rearrange folder

--- 2025-04-28 literature/paper_draft.Rmd
@@
 
-The model is added with a Gaussian innovation process $\bm{e}_t$, with covariance matrix $\Sigma$. The covariance matrix $\Sigma$ is generated specifically in the following way:
+The model is added with a Gaussian innovation process $\boldsymbol{\epsilon}_t$, with covariance matrix $\Sigma$. The covariance matrix $\Sigma$ is generated specifically in the following way:
 
-1. A compound symmetric correlation matrix is used for each block of size $p_i$ in $\bm{A}_i$, where the coefficients are sampled from a uniform distribution.
+1. A compound symmetric correlation matrix is used for each block of size $n_{b,i}$ in $\boldsymbol{A}_i$, where the coefficients are sampled from a uniform distribution between 0 and 1.
 
-2. The correlations between different blocks are imposed using the Algorithm 1 in Hardin, Garcia & Golan (2013).
+2. The correlations between different blocks are imposed using the Algorithm 1 in @Hardin2013-wu.
 
-3. The covariance matrix $\Sigma$ is then constructed by uniform sampling of standard deviations for all $p$ series.
+3. The covariance matrix $\Sigma$ is then constructed by uniform sampling of standard deviations, in a range of $[\sqrt{2}, \sqrt{6}]$, for all $n_b$ series.
 
 We have an option to randomly flip the signs of the covariance elements, which will create a more realistic structure in the innovation process.
-This is also to simulate the real-world scenario where the observed covariance matrix is not necessarily positive definite.
+This can be done by pre- and post-multiplying $\Sigma$ by a random diagonal matrix $\boldsymbol{V}$ with entries sampled from $\{-1, 1\}$, yielding $\Sigma^* = \boldsymbol{V} \Sigma \boldsymbol{V}$.
+
+For each series, $T = 116$ or $316$ observations are generated.
+The first 100 or 300 observations are used for training, and the last 16 observations are used for testing.
+The remaining data is used to compute the best fitted ARIMA models by minimising the AICc criterion, in which we use the automatic algorithm from Fabletools R package [@fabletools-fa].
+We refer to them as base models, and their base forecasts are then reconciled using the MinT with different covariance estimators.
+These include using the unbiased sample covariance matrix - MinT(Sample), the shrinkage estimator - MinT(Shrink), and the NOVELIST estimator - MinT(N).
+The Monte Carlo simulation is repeated $M = 10000$ times, in which the parameters for data generating process is fixed.
 

### 2025-10-13 thesis/paper_draft.qmd
rearrange folder

--- 2025-10-13 literature/paper_draft.qmd
@@
 
-sim1_1 <- readRDS("./literature/lib/S4-2-1_T50_M500.rds")
-sim1_2 <- readRDS("./literature/lib/S4-2-1_T300_M500.rds")
+sim1_1 <- readRDS("./thesis/lib/S4-2-1_T50_M500.rds")
+sim1_2 <- readRDS("./thesis/lib/S4-2-1_T300_M500.rds")
 
-sim2_1 <- readRDS("./literature/lib/S36-6-1_T50_M500.rds")
-sim2_2 <- readRDS("./literature/lib/S36-6-1_T300_M500.rds")
+sim2_1 <- readRDS("./thesis/lib/S36-6-1_T50_M500.rds")
+sim2_2 <- readRDS("./thesis/lib/S36-6-1_T300_M500.rds")
 
-sim_dense_50 <- readRDS("./literature/lib/S100-10-3-1_T50_M200_dense.rds")
-sim_dense_300 <- readRDS("./literature/lib/S100-10-3-1_T300_M200_dense.rds")
+sim_dense_50 <- readRDS("./thesis/lib/S100-10-3-1_T50_M200_dense.rds")
+sim_dense_300 <- readRDS("./thesis/lib/S100-10-3-1_T300_M200_dense.rds")
 
-sim_sparse_50 <- readRDS("./literature/lib/S100-10-3-1_T50_M500_sparsebwgrp.rds")
-sim_sparse_300 <- readRDS("./literature/lib/S100-10-3-1_T300_M200_sparsebwgrp.rds")
+sim_sparse_50 <- readRDS("./thesis/lib/S100-10-3-1_T50_M500_sparsebwgrp.rds")
+sim_sparse_300 <- readRDS("./thesis/lib/S100-10-3-1_T300_M200_sparsebwgrp.rds")
 
@@
 ```{r}
-tourism_plots <- readRDS("D:/Github/Recon_Honours_Thesis/literature/tourism_plots.rds")
+tourism_plots <- readRDS("D:/Github/Recon_Honours_Thesis/thesis/tourism_plots.rds")
 ```

### 2025-10-13 R/plot_helpers.R
copy tourism results and plotting helpers to thesis github

@@ -1,0 +1,142 @@
+# ---------- 1) Encoder: adds family, variant, labels, and aesthetics ----------
+encode_recon_models <- function(df, model_col = ".model", drop_large_pc = FALSE) {
+  df %>%
+    mutate(
+      method_raw = .data[[model_col]],
+      # Families
+      family = case_when(
+        str_starts(method_raw, "base") ~ "Base",
+        str_starts(method_raw, "ols") ~ "OLS",
+        str_starts(method_raw, "mint_shr") ~ "MinT-S",
+        str_starts(method_raw, "mint_n") ~ "MinT-N",
+        TRUE ~ "Other"
+      ),
+      # Variants
+      variant_type = case_when(
+        str_detect(method_raw, "_pc_") ~ "pc",
+        str_ends(method_raw, "_sv") ~ "sv",
+        str_ends(method_raw, "_hcov") ~ "hcov",
+        TRUE ~ "vanilla"
+      ),
+      pc_k = if_else(str_detect(method_raw, "_pc_"),
+                     as.integer(str_extract(method_raw, "(?<=_pc_K)\\d+")),
+                     NA_integer_),
+      # model labels (clean legend text)
+      model_label = case_when(
+        method_raw == "base" ~ "Base",
+        method_raw == "ols" ~ "OLS",
+        method_raw == "mint_shr" ~ "MinT-S",
+        method_raw == "mint_n" ~ "MinT-N",
+        # shrinkage of OLS and Base (for probabilistic forecasts)
+        method_raw == "ols_shr" ~ "OLS-S",
+        method_raw == "base_shr" ~ "Base-S",
+        # PC-adjusted (explicit PC1/PC2; general fallback for other K)
+        str_detect(method_raw, "mint_shr_pc_K1") ~ "MinT-S(PC1)",
+        str_detect(method_raw, "mint_shr_pc_K2") ~ "MinT-S(PC2)",
+        str_detect(method_raw, "mint_n_pc_K1")   ~ "MinT-N(PC1)",
+        str_detect(method_raw, "mint_n_pc_K2")   ~ "MinT-N(PC2)",
+        str_detect(method_raw, "_pc_K") ~ paste0(if_else(str_detect(method_raw, "mint_shr"), "MinT-S", "MinT-N"),
+                                                 "(PC", pc_k, ")"),
+        # sv / hcov
[…]

### 2025-10-13 R/table_geo.R

@@ -1,0 +1,186 @@
+library(dplyr)
+library(tidyr)
+library(stringr)
+
+path <- here::here()
+
+# 0) Prepping --------------------------------
+# State 
+state_map <- list(
+  NSW = c(
+    "Blue Mountains", "Central Coast", "Central NSW",
+    "Hunter", "New England North West", "North Coast NSW", "Outback NSW",
+    "Riverina", "Snowy Mountains", "South Coast", "Sydney",
+    "Capital Country", "The Murray",
+    "Canberra" # ACT is included in NSW for simplicity
+  ),
+  VIC = c(
+    "Ballarat", "Bendigo Loddon", "Central Murray", "Geelong and the Bellarine",
+    "Gippsland", "Great Ocean Road", "High Country", "Macedon", "Mallee",
+    "Melbourne", "Melbourne East", "Murray East", "Peninsula", "Phillip Island",
+    "Spa Country", "Upper Yarra", "Western Grampians", "Wimmera",
+    "Central Highlands", "Lakes", "Goulburn"
+  ),
+  QLD = c(
+    "Brisbane", "Bundaberg", "Capricorn", "Fraser Coast", "Gladstone",
+    "Gold Coast", "Mackay", "Outback Queensland", "Southern Queensland Country",
+    "Sunshine Coast", "Townsville", "Tropical North Queensland",
+    "Whitsundays"
+  ),
+  SA = c(
+    "Adelaide", "Adelaide Hills", "Barossa", "Clare Valley", "Eyre Peninsula",
+    "Fleurieu Peninsula", "Flinders Ranges and Outback", "Kangaroo Island",
+    "Limestone Coast", "Murray River, Lakes and Coorong", "Riverland",
+    "Yorke Peninsula"
+  ),
+  WA = c(
+    "Australia's Coral Coast", "Australia's Golden Outback",
+    "Australia's North West", "Australia's South West", "Destination Perth"
+  ),
+  TAS = c(
+    "East Coast", "Hobart and the South",
+    "Launceston and the North", "North West", "West Coast"
+  ),
+  NT = c(
+    "Alice Springs", "Barkly", "Darwin", "Katherine Daly",
+    "Litchfield Kakadu Arnhem", "Lasseter", "MacDonnell"
+  )
+)
+region_state_lookup <- tibble(
+  State  = rep(names(state_map), lengths(state_map)),
+  Region = unlist(state_map)
+)
+# Zone 
[…]

### 2025-10-13 thesis/paper_draft.qmd
graphs for point and probs results tourism

@@
 
+# source all R scripts in thesis/R folder
+R_files <- list.files("D:/Github/Recon_Honours_Thesis/R", pattern = "*.R", full.names = TRUE)
+invisible(lapply(R_files, source))
+
 path <- "D:/Github/Recon_Honours_Thesis"
@@
 ```{r}
+plot_relative_improvement <- function(MSE_data) {
+  MSE_data |>
+    filter(
+      .model != "mint_true",
+      h <= 4
+    ) |>
+    group_by(series, h) |>
+    mutate(
+      base_MSE = MSE[.model == "base"]
+    ) |>
+    ungroup() |>
+    group_by(.model, h) |>
+    summarise(
+      MSE = mean(MSE),
+      base_MSE = mean(base_MSE),
+      pct_change = (MSE - base_MSE) / base_MSE * 100
+    ) |>
+    ggplot(aes(x = h, y = pct_change, color = .model)) +
+    geom_line() +
+    labs(
+      x = "Forecast Horizon (h)",
+      y = "Percentage Change (%)"
+    ) +
+    # labs(x = element_blank(), y = element_blank()) +
+    theme_bw()
+}
+```
+
+```{r}
 #| label: fig-Sigma-2x50
@@
 
-```{r}
+```{r summary structure tourism}
 #| tbl-cap: "Hierarchical and grouped structure of Australian domestic tourism flows"
@@
 
-![Rolling-window cross-validation scheme for evaluating forecasting performance in Australia tourism data](D:/Github/Recon_Honours_Thesis/figs/tourism_cv.png){#fig-tourism-cv width="70%"}
+![Rolling-window cross-validation scheme for evaluating forecasting performance in Australia tourism data](D:/Github/Recon_Honours_Thesis/figs/tourism_cv.png){#fig-tourism-cv width="50%"}
 
@@
 
-```{r}
+```{r tourism-MSE-line}
 #| label: fig-tourism-MSE-line
@@
 The positive entries indicate an decrease in MSE."
-tourism_plots$MSE_line + 
-  labs(title = NULL, subtitle = NULL) + 
+
+tourism_MSE_results <- readRDS(paste0(path, "/thesis/lib/tourism_MSE_results.rds"))
+
+tourism_MSE_results |> 
+  mutate(shape_lab = ifelse(shape_lab %in% c("PC1","PC2"), shape_lab, "")) |>
+  ggplot(aes(x = h, y = pct_change, color = family, linetype = variant_type, shape = shape_lab)) +
+  geom_line(linewidth = 0.8) +
+  geom_point(alpha = 0.85, size = 2.6, na.rm = TRUE) +
+  recon_scales +
+  labs(x = "Horizon", y = "% improvements") +
+  # scale_y_continuous(limits = c(-10, 0.2)) +
+  theme_minimal() + 
   theme(
@@
 
-tourism_plots$MSE_heat + 
-  labs(title = NULL, subtitle = NULL) + 
+tourism_MSE_results |> 
+  # best models per h
+  filter(.model != "base") |> 
+  group_by(h) |>
+  mutate(
+    is_best = pct_change == max(pct_change),
+    # rescale to (0 to -1)
+    scaled_pct_change = scales::rescale(pct_change, to = c(0, 100))
+  ) |>
+  filter(.model != "base") |> 
+  ggplot(aes(
+    x = h, y = fct_rev(model_label), 
+    # fill = pct_change
+    fill = scaled_pct_change
+  )) +
+    geom_tile() +
+    geom_text(aes(label = ifelse(is_best, "*", "")), color = "black", size = 8) +
+    scale_fill_gradient2(high = "#ca2b33ff", mid = "#f7ffadff", midpoint = 0) +
+    labs(x = "Horizon", y = "Model") +
+    theme_minimal() + 
     theme(legend.title = element_blank(), legend.position = "bottom") + 
@@
 
-<!-- 
-```{r}
+```{r tourism-probscores}
 #| label: fig-tourism-probscores
-#| fig.height: 12
-#| fig.width: 12
+#| fig.height: 10
+#| fig.width: 10
 #| fig-cap: "Percentage relative improvement in the Winkler score at 80% and 95% nominal coverage, CRPS, and Energy score of multiple reconciled forecasts over the base forecasts for the Australian domestic tourism data, for 1-step-ahead forecasts.
-The positive entries indicate an decrease in the probabilistic scores."
+The positive (negative) entries indicate a decrease (increase) in the probabilistic scores relative to base."
 
@@
 
-tourism_plots$W80 + labs(title = "Winkler score 80%", subtitle = NULL) + 
-  tourism_plots$W95 + labs(title = "Winkler score 95%", subtitle = NULL) +
-  tourism_plots$CRPS + labs(title = "CRPS", subtitle = NULL) +
-  tourism_plots$Energy + labs(title = "Energy score", subtitle = NULL) +
+tourism_uniscores <- readRDS(paste0(path, "./thesis/lib/tourism_uniscore_results.rds"))
+tourism_energy <- readRDS(paste0(path, "./thesis/lib/tourism_energyscore_results.rds"))
+
+plot_uniscore <- function(score_label, title) {
+  tourism_uniscores %>%
+    filter(rule == score_label) %>%
+    group_by(family, model_label, pc_k, iter) %>%
+    summarise(
+      score = mean(score),
+      base_score = mean(base_score),
+      pct_change = (base_score - score) / base_score * 100,
+      .groups = "drop_last"
+    ) %>%
+    ungroup() %>%
+    group_by(family, model_label, pc_k) %>%
+    summarise(
+      pct_change_avg = pmin(mean(pct_change), 10),   # cap at 10
+      capped = as.integer(pct_change_avg >= 10),
+      .groups = "drop"
+    ) %>%
+    filter(model_label != "Base", model_label != "OLS-S") %>%
+    mutate(pc_k = ifelse(is.na(pc_k), "0", as.character(pc_k))) %>%
+    # ggplot
+    ggplot(aes(x = model_label, y = pct_change_avg,
+              fill = family, alpha = pc_k)) +
+      geom_col(color = "black", size = 0.5) +
+      # geom_text(aes(label = ifelse(capped == 1, "> 10%", "")), vjust = -0.5) +
+      recon_fill_scale +
+      scale_alpha_manual(values = c("0" = 1, "1" = 0.6, "2" = 0.3), guide = "none") +
+      labs(
+        x = " ", y = " ",
+        title = title
+      ) +
+      theme_minimal(base_size = 12) +
+      theme(axis.text.x = element_text(angle = 25, hjust = 0.6)) +
+      guides(fill = "none")
+}
+
+plot_W80 <- plot_uniscore("W80", "Winkler score 80%")
+plot_W95 <- plot_uniscore("W95", "Winkler score 95%")
+plot_CRPS <- plot_uniscore("CRPS", "CRPS")
+
+plot_energy <- tourism_energy |> 
+  group_by(family, model_label, pc_k) |> 
+  summarise(
+    pct_change_avg = mean(pct_change_per_series)
+  ) |>
+  mutate(pc_k = ifelse(is.na(pc_k), "0", as.character(pc_k))) |> 
+  filter(! grepl("Ba|OLS-S", model_label)) |>
+  ggplot(aes(x = model_label, y = pct_change_avg, fill = family, alpha = pc_k)) +
+    geom_col(color = "black", size = 0.5) +
+    recon_fill_scale +
+    scale_alpha_manual(values = c("0" = 1, "1" = 0.6, "2" = 0.3), guide = "none") +
+    labs(x = " ", y = " ", title = "Energy Score") +
+    theme_minimal() +
+    theme(axis.text.x = element_text(angle = 25, hjust = 0.6))
+
+p <- plot_W80 + plot_W95 + plot_CRPS + plot_energy +
   plot_layout(ncol = 2, guides = "collect", axis_titles = "collect") & 
@@
 
-```
--->
-
-![Percentage relative improvement in the Winkler score at 80% and 95% nominal coverage, CRPS, and Energy score of multiple reconciled forecasts over the base forecasts for the Australian domestic tourism data, for 1-step-ahead forecasts.
-The positive (negative) entries indicate a decrease (increase) in the probabilistic scores relative to base.](D:/Github/Recon_Honours_Thesis/figs/tourism_probscores.png){#fig-tourism-probscores}
+cowplot::ggdraw(p) +
+  cowplot::draw_label("% improvements", angle = 90, x = 0.02, y = 0.5, size = 16)
+
+# tourism_plots$W80 + labs(title = "Winkler score 80%", subtitle = NULL) + 
+#   tourism_plots$W95 + labs(title = "Winkler score 95%", subtitle = NULL) +
+#   tourism_plots$CRPS + labs(title = "CRPS", subtitle = NULL) +
+#   tourism_plots$Energy + labs(title = "Energy score", subtitle = NULL) +
+#   plot_layout(ncol = 2, guides = "collect", axis_titles = "collect") & 
+#   theme(legend.position = "bottom", legend.title = element_blank())
+```
+
+<!-- ![Percentage relative improvement in the Winkler score at 80% and 95% nominal coverage, CRPS, and Energy score of multiple reconciled forecasts over the base forecasts for the Australian domestic tourism data, for 1-step-ahead forecasts.
+The positive (negative) entries indicate a decrease (increase) in the probabilistic scores relative to base.](D:/Github/Recon_Honours_Thesis/figs/tourism_probscores.png){#fig-tourism-probscores} -->
 
@@
 
-source(paste0(path, "R/table_geo.R"))
 table_geo_labeled <- create_table_geo()

### 2025-10-13 thesis/R/table_geo.R

@@ -1,186 +1,0 @@
-library(dplyr)
-library(tidyr)
-library(stringr)
-
-path <- here::here()
-
-# 0) Prepping --------------------------------
-# State 
-state_map <- list(
-  NSW = c(
-    "Blue Mountains", "Central Coast", "Central NSW",
-    "Hunter", "New England North West", "North Coast NSW", "Outback NSW",
-    "Riverina", "Snowy Mountains", "South Coast", "Sydney",
-    "Capital Country", "The Murray",
-    "Canberra" # ACT is included in NSW for simplicity
-  ),
-  VIC = c(
-    "Ballarat", "Bendigo Loddon", "Central Murray", "Geelong and the Bellarine",
-    "Gippsland", "Great Ocean Road", "High Country", "Macedon", "Mallee",
-    "Melbourne", "Melbourne East", "Murray East", "Peninsula", "Phillip Island",
-    "Spa Country", "Upper Yarra", "Western Grampians", "Wimmera",
-    "Central Highlands", "Lakes", "Goulburn"
-  ),
-  QLD = c(
-    "Brisbane", "Bundaberg", "Capricorn", "Fraser Coast", "Gladstone",
-    "Gold Coast", "Mackay", "Outback Queensland", "Southern Queensland Country",
-    "Sunshine Coast", "Townsville", "Tropical North Queensland",
-    "Whitsundays"
-  ),
-  SA = c(
-    "Adelaide", "Adelaide Hills", "Barossa", "Clare Valley", "Eyre Peninsula",
-    "Fleurieu Peninsula", "Flinders Ranges and Outback", "Kangaroo Island",
-    "Limestone Coast", "Murray River, Lakes and Coorong", "Riverland",
-    "Yorke Peninsula"
-  ),
-  WA = c(
-    "Australia's Coral Coast", "Australia's Golden Outback",
-    "Australia's North West", "Australia's South West", "Destination Perth"
-  ),
-  TAS = c(
-    "East Coast", "Hobart and the South",
-    "Launceston and the North", "North West", "West Coast"
-  ),
-  NT = c(
-    "Alice Springs", "Barkly", "Darwin", "Katherine Daly",
-    "Litchfield Kakadu Arnhem", "Lasseter", "MacDonnell"
-  )
-)
-region_state_lookup <- tibble(
-  State  = rep(names(state_map), lengths(state_map)),
-  Region = unlist(state_map)
-)
-# Zone 
[…]

### 2025-10-13 thesis/archived.qmd
shorten simulation section, refining

--- 2025-10-13 literature/archived.qmd
@@
 In this paper, we place our main focus on the MinT method.
+-->
+
+# Sparse correlation simulation discussion
+
+<!-- ## Exploring the sparsity of the DGP covariance matrix
+
+\color{red}
+to be revised for final draft
+\color{black}
+
+```{r}
+plot_relative_improvement <- function(MSE_data) {
+  MSE_data |>
+    filter(
+      .model != "mint_true",
+      h <= 4
+    ) |>
+    group_by(series, h) |>
+    mutate(
+      base_MSE = MSE[.model == "base"]
+    ) |>
+    ungroup() |>
+    group_by(.model, h) |>
+    summarise(
+      MSE = mean(MSE),
+      base_MSE = mean(base_MSE),
+      pct_change = (MSE - base_MSE) / base_MSE * 100
+    ) |>
+    ggplot(aes(x = h, y = pct_change, color = .model)) +
+    geom_line() +
+    labs(
+      x = "Forecast Horizon (h)",
+      y = "Percentage Change (%)"
+    ) +
+    # labs(x = element_blank(), y = element_blank()) +
+    theme_bw()
+}
+```
+
+```{r}
+# #| label: fig-Sigma-2x50
+# #| fig-cap: "Heatmaps of the dense and sparse correlation matrix of the data generating process."
+# #| fig.height: 12
+# #| fig.width: 8
+# #| out.width: "100%"
+# #| fig.align: "center"
+p_dense_Sigma <- sim_dense_50$Sigma |>
+  cov2cor() |> 
+  plot_heatmap(T) +
+    labs(title = "Dense")
+
+p_sparse_Sigma <- sim_sparse_50$Sigma |>
+  cov2cor() |> 
+  plot_heatmap(T) +
+    labs(title = "Sparse")
+
+p_dense_Sigma + p_sparse_Sigma + 
+  plot_layout(ncol = 1, guides = "collect", axis_titles = "collect") & 
+  theme(legend.position = "right")
+```
+
+```{r}
+#| label: fig-sim-results-dsp
+#| fig-cap: "Relative improvement of the MSE of reconciled forecasts over the base forecasts for the 2x50 hierarchical structure with dense and sparse DGP's correlation matrix, for 1 to 4 steps ahead forecasts, with 2 time series lengths (T = 54 and T = 304)."
+#| fig.height: 7
+#| fig.width: 8
+#| out.width: "90%"
+#| fig.align: "center"
+p_dense_50 <- sim_dense_50$MSE |>
+  transform_sim_MSE(F) |> 
+  plot_relative_improvement() +
+    labs(title = "Dense DGP Covariance Matrix", subtitle = "T = 54")
+p_sparse_50 <- sim_sparse_50$MSE |>
+  transform_sim_MSE(F) |> 
+  plot_relative_improvement() +
+    labs(title = "Dense DGP Covariance Matrix")
+
+p_dense_300 <- sim_dense_300$MSE |>
+  transform_sim_MSE(F) |> 
+  filter(.model != "ols") |> 
+  plot_relative_improvement() +
+    labs(subtitle = "T = 304")
+p_sparse_300 <- sim_sparse_300$MSE |>
+  transform_sim_MSE(F) |> 
+  plot_relative_improvement()
+
+p_dense_50 + p_sparse_50 + p_dense_300 + p_sparse_300 +
+  plot_layout(ncol = 2, guides = "collect", axis_titles = "collect") &
+  theme(legend.position = "right")
+```
+
+In our second simulation study, we design a data‐generating process that contrasts “dense” and “sparse” correlation regimes among bottom‐level series, reflecting settings one might encounter in practice.
+We consider the same hierarchical structure of two large groups of 50 bottom series as above.
+Specifically, the two groups would have strong within-group dependencies throughout and either modest between‐group correlations (the dense scenario) or complete independence (the sparse scenario).
+These correlation matrices are depicted in Figure @fig-Sigma-2x50.
+Both scenarios share the same VAR(1) coefficient structure as in our previous simulations; only the innovation covariance changes.
+Such a setup mirrors real‐world contexts where, for example, sales within a product line may exhibit strong co‐movements, while those in a separate line operate nearly independently.
+
+Rather than collapsing each group into a single aggregate as in previous practice, we impose a more intricate, multi‐level aggregation path to stress‐test reconciliation methods.
+We first form ten intermediate series by summing bottom‐level series in consecutive blocks of ten.
+These ten series are then assigned to three second‐level aggregates--groups of four, three, and four--and finally summed to produce a single top‐level series.
+This asymmetric hierarchy creates overlapping correlation patterns: some level‐2 series share bottom‐level groups, while others draw from both.
+
+emulating practical scenarios such as regional sales aggregations that span multiple product categories or overlapping territories.
+
+Figure @fig-sim-results-dsp presents out‐of‐sample mean squared error improvements over the base forecasts for each reconciliation strategy under both dense and sparse settings, with two panel lengths--54 and 304 observations--reserving the last four points for testing.
+In both short and long samples, MinT using either the shrinkage estimator (mint_shr) or the NOVELIST estimator (mint_n) delivers pronounced gains over incoherent ARIMA forecasts, particularly at the one‐step horizon where cross‐series correlations most directly inform the forecast adjustments.
+Although the two MinT variants perform almost indistinguishably overall, mint_n edges out mint_shr in the immediate horizon, whereas mint_shr slightly outperforms for longer horizons.
+By contrast, MinT with the raw sample covariance (mint_sample) suffers in small‐sample settings; as expected, its performance improves dramatically with 304 data points, since the sample covariance becomes more reliable with larger $n$.
+This highlights the practical necessity of regularized estimators in high‐dimensional, low‐sample contexts, a situation common in real applications where histories are short relative to the number of series.
+
+Additional designs (varying block sizes, aggregation paths, correlation configurations) also failed to separate NOVELIST from Shrinkage.
+Their nearly identical performance under these synthetic scenarios suggests that our current simulation may not unveil the full advantages of the thresholding estimators.
+This finding motivates our turn to empirical data, where latent structural features and regime shifts, which we will discuss in the next section, may reveal performance differences.
 -->

### 2025-10-13 thesis/paper_draft.qmd
shorten simulation section, refining

@@
 
-## General Design Framework {-}
+## General Design
 
@@
 The VAR(1) and correlation configurations for the 6 by 6 case and 2 by 50 case are illustrated in @fig-settings-6x6 and @fig-settings-2x50, respectively.
-The block diagonal structure of the VAR(1) coefficient matrices $\boldsymbol{A}$ reflects the grouping of series, and the correlation matrices $\boldsymbol{\Sigma}$ show higher correlations among series within the same group.
+The block diagonal structure of the VAR(1) coefficient matrices $\boldsymbol{A}$ reflects the grouping of series, and the correlation matrices- show higher correlations among series within the same group.
 
@@
 p1_Sigma <- plot_heatmap(cov2cor(sim1_1$Sigma), T) + 
-  ggtitle("2x2 Correlation Matrix")
+  ggtitle("2x2 Correlation Matrix of Innovation process")
 
@@
 p2_Sigma <- plot_heatmap(cov2cor(sim2_1$Sigma), T) +
-  ggtitle("Correlation Matrix") + theme_void()
+  ggtitle("Correlation Matrix of Innovation process") + theme_void()
 
 p3_A <- plot_heatmap(sim3_1$A) +
-  ggtitle("2x50 VAR(1) Coefficients Matrix")
+  ggtitle("VAR(1) Coefficients Matrix")
 
 p3_Sigma <- plot_heatmap(cov2cor(sim3_1$Sigma), T) +
-  ggtitle("2x50 Correlation Matrix")
+  ggtitle("Correlation Matrix of Innovation process")
 ```
@@
 #| label: fig-settings-6x6
-#| fig-cap: "Heatmaps of the VAR(1) coefficient matrix and correlation matrix for the 6 by 6 structure."
-#| fig.height: 4
-#| fig.width: 6
-#| out.width: "85%"
+#| fig-cap: "The VAR(1) coefficients matrix (left) and correlation matrix of the innovation process (right) for the 6 groups of 6 structure."
+#| fig.height: 5
+#| fig.width: 7
+#| out.width: "75%"
 #| fig.align: "center"
@@
 #| label: fig-settings-2x50
-#| fig-cap: "Heatmaps of the VAR(1) coefficient matrix and correlation matrix for the 2 by 50 structure."
+#| fig-cap: "The VAR(1) coefficients matrix (top) and correlation matrix of the innovation process (bottom) for the 2 groups of 50 structure."
 #| fig.height: 12
@@
 This is expected since the synthetic data generating process does not simulate from strong latent factors.
-Third, the scaled variance and h-step-ahead residuals approaches do not enhance performance as the forecast horizon increases, suggesting that the proportionality assumption may not be severely violated in this VAR(1) setup.
-Lastly, as the hierarchy expands from 36 to 100 series, forecasts from NOVELIST methods become worse relative to base forecasts when we move to $2$-step-ahead and beyond.
-When looking into each MCMC replication, we find that there are instances where the NOVELIST estimator collapses to the shrinkage estimator due to a large optimal threshold $\hat{\delta}^*$ being selected in the cross-validation step.
+Lastly, the scaled variance and h-step-ahead residuals approaches do not enhance performance as the forecast horizon increases, suggesting that the proportionality assumption may not be severely violated in this VAR(1) setup.
+
+<!-- Lastly, as the hierarchy expands from 36 to 100 series, forecasts from NOVELIST methods become worse relative to base forecasts when we move to $2$-step-ahead and beyond.  -->
+
+When looking into each MCMC replication, we find that there are instances where the NOVELIST estimator collapses to the shrinkage estimator due to a large optimal threshold $\hat{\delta}$ being selected in the cross-validation step, resulting in a diagonal shrinkage target.
 
@@
 
-Moving on to probabilistic forecasts, we evaluate the performance of reconciliation methods using the energy score [@Gneiting2011-km], a proper scoring rule for multivariate distributions.
+Moving on to probabilistic forecasts, we evaluate the performance of reconciliation methods using the energy score, a proper scoring rule for multivariate predictive distributions.
 Figure @fig-sim-results-2 presents the percentage relative improvement in energy score for 1-step-ahead forecasts.
@@
 
-## Exploring the sparsity of the DGP covariance matrix
-
-\color{red}
-to be revised for final draft
-\color{black}
-
-```{r}
-plot_relative_improvement <- function(MSE_data) {
-  MSE_data |>
-    filter(
-      .model != "mint_true",
-      h <= 4
-    ) |>
-    group_by(series, h) |>
-    mutate(
-      base_MSE = MSE[.model == "base"]
-    ) |>
-    ungroup() |>
-    group_by(.model, h) |>
-    summarise(
-      MSE = mean(MSE),
-      base_MSE = mean(base_MSE),
-      pct_change = (MSE - base_MSE) / base_MSE * 100
-    ) |>
-    ggplot(aes(x = h, y = pct_change, color = .model)) +
-    geom_line() +
-    labs(
-      x = "Forecast Horizon (h)",
-      y = "Percentage Change (%)"
-    ) +
-    # labs(x = element_blank(), y = element_blank()) +
-    theme_bw()
-}
-```
+Other scoring rules, such as the Winkler score and CRPS, yield similar conclusions.
+
+## Other Data Generating Processes
+
+In attempts to differentiate the performance of NOVELIST from the shrinkage estimator, we also simulate from a sparse covariance matrix for a 2 groups of 50 scenario, as illustrated in @fig-Sigma-2x50.
+The sparse covariance matrix is obtained by randomly choosing 40% of the bottom series and setting their correlations with all other series to zero, resulting in a grid-like sparse structure.
+The VAR(1) coefficient matrix remains the same as in the dense case.
+The idea is to allow NOVELIST to exploit the sparsity in the covariance structure, since it can control the sparsity of the shrinkage target via the thresholding parameter $\delta$.
+However, no profound insights can be drawn from the results.
 
@@
 #| label: fig-Sigma-2x50
-#| fig-cap: "Heatmaps of the dense and sparse correlation matrix of the data generating process."
-#| fig.height: 12
+#| fig-cap: "Sparse correlation matrix of the innovation process for 2 groups of 50 structure."
+#| fig.height: 6
 #| fig.width: 8
-#| out.width: "100%"
+#| out.width: "80%"
 #| fig.align: "center"
-p_dense_Sigma <- sim_dense_50$Sigma |>
+
+sim_sparse_50 <- readRDS(paste0(path, "/thesis/lib/params_S100_10_3_1_sparse46.rds"))
+sim_sparse_50$Sigma |>
   cov2cor() |> 
-  plot_heatmap(T) +
-    labs(title = "Dense")
-
-p_sparse_Sigma <- sim_sparse_50$Sigma |>
-  cov2cor() |> 
-  plot_heatmap(T) +
-    labs(title = "Sparse")
-
-p_dense_Sigma + p_sparse_Sigma + 
-  plot_layout(ncol = 1, guides = "collect", axis_titles = "collect") & 
-  theme(legend.position = "right")
-```
-
-```{r}
-#| label: fig-sim-results-dsp
-#| fig-cap: "Relative improvement of the MSE of reconciled forecasts over the base forecasts for the 2x50 hierarchical structure with dense and sparse DGP's correlation matrix, for 1 to 4 steps ahead forecasts, with 2 time series lengths (T = 54 and T = 304)."
-#| fig.height: 7
-#| fig.width: 8
-#| out.width: "90%"
-#| fig.align: "center"
-p_dense_50 <- sim_dense_50$MSE |>
-  transform_sim_MSE(F) |> 
-  plot_relative_improvement() +
-    labs(title = "Dense DGP Covariance Matrix", subtitle = "T = 54")
-p_sparse_50 <- sim_sparse_50$MSE |>
-  transform_sim_MSE(F) |> 
-  plot_relative_improvement() +
-    labs(title = "Dense DGP Covariance Matrix")
-
-p_dense_300 <- sim_dense_300$MSE |>
-  transform_sim_MSE(F) |> 
-  filter(.model != "ols") |> 
-  plot_relative_improvement() +
-    labs(subtitle = "T = 304")
-p_sparse_300 <- sim_sparse_300$MSE |>
-  transform_sim_MSE(F) |> 
-  plot_relative_improvement()
-
-p_dense_50 + p_sparse_50 + p_dense_300 + p_sparse_300 +
-  plot_layout(ncol = 2, guides = "collect", axis_titles = "collect") &
-  theme(legend.position = "right")
-```
-
-In our second simulation study, we design a data‐generating process that contrasts “dense” and “sparse” correlation regimes among bottom‐level series, reflecting settings one might encounter in practice.
-We consider the same hierarchical structure of two large groups of 50 bottom series as above.
-Specifically, the two groups would have strong within-group dependencies throughout and either modest between‐group correlations (the dense scenario) or complete independence (the sparse scenario).
-These correlation matrices are depicted in Figure @fig-Sigma-2x50.
-Both scenarios share the same VAR(1) coefficient structure as in our previous simulations; only the innovation covariance changes.
-Such a setup mirrors real‐world contexts where, for example, sales within a product line may exhibit strong co‐movements, while those in a separate line operate nearly independently.
-
-<!-- Rather than collapsing each group into a single aggregate as in previous practice, we impose a more intricate, multi‐level aggregation path to stress‐test reconciliation methods.
-We first form ten intermediate series by summing bottom‐level series in consecutive blocks of ten.
-These ten series are then assigned to three second‐level aggregates--groups of four, three, and four--and finally summed to produce a single top‐level series.
-This asymmetric hierarchy creates overlapping correlation patterns: some level‐2 series share bottom‐level groups, while others draw from both.
--->
-
-<!-- emulating practical scenarios such as regional sales aggregations that span multiple product categories or overlapping territories. -->
-
-Figure @fig-sim-results-dsp presents out‐of‐sample mean squared error improvements over the base forecasts for each reconciliation strategy under both dense and sparse settings, with two panel lengths--54 and 304 observations--reserving the last four points for testing.
-In both short and long samples, MinT using either the shrinkage estimator (mint_shr) or the NOVELIST estimator (mint_n) delivers pronounced gains over incoherent ARIMA forecasts, particularly at the one‐step horizon where cross‐series correlations most directly inform the forecast adjustments.
-Although the two MinT variants perform almost indistinguishably overall, mint_n edges out mint_shr in the immediate horizon, whereas mint_shr slightly outperforms for longer horizons.
-By contrast, MinT with the raw sample covariance (mint_sample) suffers in small‐sample settings; as expected, its performance improves dramatically with 304 data points, since the sample covariance becomes more reliable with larger $n$.
-This highlights the practical necessity of regularized estimators in high‐dimensional, low‐sample contexts, a situation common in real applications where histories are short relative to the number of series.
-
-Additional designs (varying block sizes, aggregation paths, correlation configurations) also failed to separate NOVELIST from Shrinkage.
+  plot_heatmap(T)
+```
+
+Additional designs (varying block sizes, grouped structure, aggregation paths, correlation configurations) also failed to separate NOVELIST from Shrinkage.
 Their nearly identical performance under these synthetic scenarios suggests that our current simulation may not unveil the full advantages of the thresholding estimators.
-This finding motivates our turn to empirical data, where latent structural features and regime shifts, which we will discuss in the next section, may reveal performance differences.
+Nevertheless, we have not explored settings where PC variants or using h-step-ahead residuals approaches would have an edge.
+
+These findings motivates our turn to empirical data in the next section, where latent structural features, regime shifts, and noisy, intermittent series will reveal performance differences.
 
@@
 
+## Appendix: Simulation Supplementary {#sec-sim-supplement}
+
 ## Appendix: Australian Domestic Tourism Geographical Hierarchy {#sec-tourism-supplement}

### 2025-10-13 thesis/paper_draft.qmd
changed radar plot to include Base

@@
 library(forcats)
+library(tibble)
+library(fmsb)
 
@@
       base_score = mean(base_score),
-      pct_change = (base_score - score) / base_score * 100,
+      pct_change = (base_score - score) / base_score * 100, # positive indicate decrease in score
       .groups = "drop_last"
@@
 #| label: fig-tourism-radar
-#| fig.height: 8
-#| fig.width: 8
-#| out.width: "90%"
-#| fig-cap: "Radar plot of relative improvements in probabilistic scores (Winkler 80, Winkler 95, CRPS, Energy) over the base forecasts.
+#| fig.height: 7
+#| fig.width: 7
+#| out.width: "70%"
+#| fig-cap: "Radar plot of relative improvements in probabilistic scores (Winkler score at 80% and 95% intervals, CRPS, and Energy) over the base forecasts.
 The scores are scaled to a range of 0 to 100, with larger values indicating better performance.
 The outermost polygon represents the best possible score (100) and the innermost polygon represents the worst possible score (0).
-Only the top 5 models are shown."
+Only the top 4 MinT approaches are shown (with the base model)."
 
 # a radar plot of 5 top models based on probabilistic scores (Winkler 80, 95, CRPS, Energy) scaled to 0-100 
-par(mfrow=c(1,1), mar=c(0,0,0,0))
-tourism_plots$radar
-```
-
-The summary radar graph in @fig-tourism-radar consolidates these findings: among the selected MinT models by probabilistic criteria, *MinT-N(PC1)* (the yellow polygon) clearly leads across CRPS, W80, W95, and Energy, extending the improvements seen in the single-metric panels.
+par(mfrow=c(1,1), mar=c(1,2,1,2))
+plot_radar(tourism_uniscores, tourism_energy)
+```
+
+The summary radar graph in @fig-tourism-radar consolidates these findings: among the selected MinT models by probabilistic criteria, *MinT-N(PC1)* (the yellow polygon) clearly leads across CRPS, Winkler score at 80% and 95% intervals, and Energy, extending the improvements seen in the single-metric panels.
+Except for the Winkler score at 95% interval, all MinT variants outperform the base forecasts.
 

### 2025-10-13 R/plot_functions.R

@@ -1,0 +1,80 @@
+#' Plot radar chart comparing methods across multiple metrics
+#' (FOR THESIS USE ONLY)
+#' 
+#' @param tourism_uniscores Data frame of univariate scores
+#' @param tourism_energy Data frame of energy scores
+#' 
+#' @returns radar chart for tourism data
+plot_radar <- function(tourism_uniscores, tourism_energy) {
+
+  radar_df <- tourism_uniscores |> 
+    filter(.model %in% c("base", "mint_shr", "mint_n", "mint_shr_pc_K1", "mint_n_pc_K1")) |>
+    # average across series per iteration & compute % improv per iter
+    group_by(model_label, rule, iter) |>
+    summarise(
+      score = mean(score),
+      base_score = mean(base_score),
+      pct_change = (base_score - score) / base_score * 100
+    ) |>
+    ungroup() |>
+    # average across iter
+    group_by(model_label, rule) |>
+    summarise(
+      pct_change_avg = mean(pct_change)
+    ) |>
+    ungroup() |>
+    # filter(model_label != "Base") |>
+    pivot_wider(names_from = rule, values_from = pct_change_avg) |>
+    column_to_rownames("model_label")
+
+  radar_df <- cbind(
+    radar_df,
+    Energy = tourism_energy |> 
+      filter(.model %in% c("base", "mint_shr", "mint_n", "mint_shr_pc_K1", "mint_n_pc_K1")) |> 
+      group_by(model_label) |> 
+      summarise(
+        mean_score = mean(score),
+        pct_change_avg = mean(pct_change_per_series)
+      ) |> 
+      pull(pct_change_avg)
+  )
+
+  # rename the cols (scores)
+  colnames(radar_df) <- c("CRPS", "Winkler\n80%", "Winkler\n95%", "Energy")
+
+
+  {# --- 2) Add required max/min rows for fmsb::radarchart ---
+  maxs <- apply(radar_df, 2, max, na.rm = TRUE)
+  mins <- apply(radar_df, 2, min, na.rm = TRUE)
+  mins <- pmin(mins, 0) # ensure min is at most 0
+
+  # pad the scale a bit so polygons don’t touch edges
+  pad  <- 0.01 * (maxs - mins)
+  scale_top <- maxs + pad
+  scale_bot <- mins - pad
+
[…]

### 2025-10-14 thesis/paper_draft.qmd
expanding on intro & literature review
revise each paragraph idea, hts, recon, mint and shortcomings, current research (only DS), modern cov ests

@@
 
-<!-- # Notation {-} -->
-
-<!-- -   Scalar $y_t$ -->
-
-<!-- -   Vector $\boldsymbol{y}_t$ -->
-
-<!-- -   Matrix $\boldsymbol{S}$ -->
-
-<!-- -   Covariance matrix of in-sample 1-step-ahead base forecast errors $\hat{\boldsymbol{W}}_1$ -->
-
-<!-- -   Its shrinkage estimator with diagonal target $\hat{\boldsymbol{W}}^{shr}_{1, D}$ -->
-
-<!-- -   Its NOVELIST estimator $\hat{\boldsymbol{W}}^{shr}_{1, thr}$ -->
-
 ```{r library}
@@
 
-<!-- ############  Intro  ############ -->
-
 <!--
-Para 1
-- Introduce hierarchical time series and its importance in various applications.
+
+############  Paragraphs  ############
+
+1. Introduce hierarchical time series and its importance in various applications.
+2. Discuss the challenges in forecasting hierarchical time series, and current methods used. Introduce MinT.
+3. Highlight the reliance of MinT on a good estimate of the covariance matrix of base forecast errors, and the problems with current estimation methods.
+4. Note the Double Shrinkage estimator paper, but still early stage.
+5. Mention advances in high-dimensional covariance estimation that can be potentially useful.
+6. State the purpose and contribution of this paper.
+
 -->
@@
 In time series forecasting, aggregation occurs in a variety of settings.
-While a formal definition of hierarchical time series can be found in @sec-hierarchy, we can think of Starbucks sales data as an illustrative example.
-Starbucks operates in many countries, and each country has multiple cities where they have outlets.
-The sales data is *structured hierarchically*: the top level is the total sales across all countries, followed by national sales for each country, and then individual sales for each outlet in a city.
-As a result, there are over 40,000 individual outlet sales to forecast, plus additional series at higher levels of aggregation such as city and country.
+For example, Starbucks operates in many countries, and each country has multiple cities where they have outlets.
+The sales data is structured *hierarchically*: the top level is the total sales across all countries, which are the sum of sales in each country, which in turn are the sum of sales in each city, and finally the sum of sales from each outlet in the city.
+As a result, there are over 50,000 sales series across all *aggregation levels*, and decision makers need forecasts at each level to manage inventory and plan marketing strategies effectively.
 The hierarchy can be even more complex if we consider the sales of different kinds of drinks (e.g., coffees, teas, refreshers) at each aggregation level.
-
-Forecasting data from such hierarchical structures also arises in many other decision-making contexts, from supply chains [@Seaman2022-bb; @Angam2025-od] and energy planning [@Di-Modica2021-ad], to macroeconomics [@El-Gemayel2022-hx; @Li2019-vv] and tourism analysis [@Athanasopoulos2009-lp].
+Such hierarchical structures also arises in many other decision-making contexts, from supply chains [@Seaman2022-bb; @Angam2025-od] and energy planning [@Di-Modica2021-ad], to macroeconomics [@El-Gemayel2022-hx; @Li2019-vv] and tourism analysis [@Athanasopoulos2009-lp].
 Stakeholders in these settings need forecasts at several aggregation levels to allocate resources and manage risk.
-The impact of methods for forecasting hierarchical time series has not been limited to academia, with industry also showing a strong interest.
-Many companies and organisations have adopted these methods in practice, including Amazon, the International Monetary Fund, IBM, SAP, and more [@Athanasopoulos2024-as].
-
-<!-- ##########  Literature Review  ########## -->
-
-<!-- - Talk about the history and evolution of forecasting hierarchical time series, starting from the early heuristic methods to the modern statistical approaches.
-
-  - Single level methods
-  - Optimal combination methods (OLS, WLS)
-  - MinT
-  - Bayesian, Machine learning
-  - Probabilistic methods -->
-
-In practice, when forecasts are produced for all series (often called *base forecasts*), they typically violate the aggregation constraints observed in the data; such forecasts are *incoherent*.
-This can undermine downstream decisions that require internal consistency.
-
-Traditionally, forecasting these hierarchical time series has been done using single-level methods, such as bottom-up, top-down, and middle-out approaches.
-Bottom-up methods involve generating forecasts for the bottom-level series and aggregating them to higher levels.
-Top-down methods start with forecasts for the only top-level series and disaggregate them down.
-Middle-out methods combine both approaches by forecasting middle-level series and then aggregating or disaggregating as needed.
-Despite their simplicity, these methods only anchor forecasts to a single level, implying a large loss of information on the hierarchy's inherent correlation structure.
-Additionally, the most disaggregated series often are very noisy or even intermittent, and the higher-level data might be smoother due to the aggregation.
-Furthermore, as we saw from the Starbucks example considering the sales of different kinds of drinks at each aggregation level -- formally defined as grouped structure in @sec-hierarchy -- the disaggregation becomes more complex since the disaggregation paths are not unique.
-Consequently, these single-level methods tend to give poor results across other levels of the hierarchy.
-
-To overcome these issues, forecast reconciliation was introduced by @Hyndman2011-jv, and later developed by @Van_Erven2015-nx, @Hyndman2016-ic, @Ben-Taieb2019-yx, @Wickramasuriya2019-xq, and others to achieve coherency in point forecasts and enhance accuracy.
-Forecast reconciliation projects a collection of independent base forecasts into a set of coherent forecasts that respect the linear constraints defining a hierarchical or grouped time-series system.
-
-Among the modern reconciliation strategies, the Min Trace (MinT) approach developed by @Wickramasuriya2019-xq is widely used and perform significantly well under right conditions.
-Despite its properties, MinT relies on a good high-dimensional covariance estimator, which is positive-definite.
-However, not many researchers have explored this issue in the current literature, except @Carrara2025-rz, who introduced a new estimator for MinT--Double Shrinkage estimator.
-The current stage of this research piece is fairly early, leaving many potential gaps still not yet to explore.
-As a result, a study of alternative high-dimensional covariance estimators for MinT is timely and desirable.
+
+In practice, when forecasts are produced for all series (often called *base forecasts*), they typically violate the aggregation constraints observed in the data (the sum of all countries' sales does not equal the total sales).
+Such forecasts are called *incoherent*.
+First, they undermine downstream decisions that require internal consistency, and second, our forecast performance suffers.
+To overcome these issues, forecast reconciliation was introduced.
+Forecast reconciliation, a post-processing step, utilises the information from the hierarchical structure and data to adjust the initially produced base forecasts, so that the resulting forecasts are *coherent* (i.e., respecting the aggregation constraints).
+It first introduced by @Hyndman2011-jv, and later developed by @Van_Erven2015-nx, @Hyndman2016-ic, @Ben-Taieb2019-yx, @Wickramasuriya2019-xq, @Wickramasuriya2020-uk, and others to enhance point forecast performance.
+Among the modern methods, the Min Trace (MinT) approach developed by @Wickramasuriya2019-xq provides good theoretical properties of minimising the total variance of the reconciled forecast errors, fast computational speed, and significant empirical improvements in many applications.
+It has become a standard method for forecast reconciliation, implemented in popular R and Python softwares/packages [@O-Hara-Wild2024-we; @Nixtla].
+
+A difficulty with MinT is in estimating the covariance matrix of the base forecast errors, especially for 2-step-ahead forecasts and beyond.
+This is a high-dimensional covariance estimation problem, where the number of series $n$ is often larger than the time dimension $T$.
+@Wickramasuriya2019-xq proposed using a shrinkage estimator with diagonal target from @Schafer2005-yw to estimate the 1-step-ahead covariance matrix, which is then scaled by a constant to approximate the h-step-ahead covariance matrix.
+While this approach is relatively simple and guarantees positive definiteness, it has three main shortcomings.
+First, it uniformly shrinks all off-diagonal elements towards zeros with equal weights, lacking flexibility to effectively preserve strong signals in the data.
+Second, in many real-world applications, hierarchical time series data often exhibit a prominent principal components structure, which is not fully utilised.
+Third, the proportionality relationship between the h-step-ahead and 1-step-ahead forecast error covariance matrices might not hold in practice.
+
+The theorectical advantage of MinT might not be achieved in finite sample settings if the covariance estimate is not close enough to the true covariance matrix.
+Despite its importance, not many researchers have explored this issue in the current literature, except @Carrara2025-rz, who proposed a Double Shrinkage estimator for MinT.
+The idea is to introduce a second shrinkage target, which incorporates conditional dependence information suggested by the hierarchical structure.
+This paper is still in its early stage, leaving many potential gaps still not yet to explore.
+
+In the current literature, there are many advancements in high-dimensional covariance estimation that can be potentially useful for MinT.
+Building upon the shrinkage estimator, @Huang2019-ua introduces a soft-thresholded target matrix instead of the diagonal target.
+Extending beyond linear shrinkage, @Ledoit2012-ga proposed a nonlinear shrinkage estimator that replaces each sample eigenvalue with a data-driven nonlinear transformation of itself, which is further developed by @Ledoit2020-xt.
+On the other hand, @Fan2013-jz introduced an estimator that directly preserves the latent factors structure, decomposing the covariance matrix into a low-rank component and a sparse component, and applying thresholding to the latter.
+There are also approaches related to thresholding, such as the adaptive thresholding estimator for sparse covariance matrices by @Cai2011-jf.
+The list of modern high-dimensional covariance estimators is long, presenting many opportunities to explore their potential in improving the performance of MinT.
+
+The purpose of this paper is to assess the forecasting performance of MinT when different covariance estimators are used. First, we show
 

### 2025-10-15 thesis/paper_draft.qmd
finished the contribution part in intro section, and write conclusion, limitation/future works

@@
 It first introduced by @Hyndman2011-jv, and later developed by @Van_Erven2015-nx, @Hyndman2016-ic, @Ben-Taieb2019-yx, @Wickramasuriya2019-xq, @Wickramasuriya2020-uk, and others to enhance point forecast performance.
+@Athanasopoulos2024-as provided a comprehensive review of the literature on forecast reconciliation.
 Among the modern methods, the Min Trace (MinT) approach developed by @Wickramasuriya2019-xq provides good theoretical properties of minimising the total variance of the reconciled forecast errors, fast computational speed, and significant empirical improvements in many applications.
@@
 The idea is to introduce a second shrinkage target, which incorporates conditional dependence information suggested by the hierarchical structure.
-This paper is still in its early stage, leaving many potential gaps still not yet to explore.
+This paper is still in its early stage, and does not fully address the shortcomings mentioned above.
 
@@
 
-The purpose of this paper is to assess the forecasting performance of MinT when different covariance estimators are used. First, we show
+The purpose of this paper is to explore the shortcomings of the shrinkage estimator and introduce alternative estimators that address these issues one by one.
+We evaluate the performance of MinT with these different covariance estimators in both point and probabilistic forecast reconciliation settings.
+In our empirical results from the complex, large-hierarchy dataset, we observe the following: (i) MinT with NOVELIST estimator with brings improvements over the shrinkage estimator under the probabilistic setting, as its predictive distributions have higher coverage rates; (ii) Covariance estimators that utilise the latent factors structure consistently produce better point and probabilistic forecasts than those that do not, when the data is evidently driven by such structure; (iii) Scaling the 1-step-ahead covariance matrix to approximate the h-step-ahead one gives significantly better performance than estimators that do not use this assumption, suggesting that the scaling assumption is not entirely incorrect.
+The estimators relaxing on the proportionality assumption that we propose can also be further extended to h-step-ahead probabilistic forecast reconciliation.
 
 The remainder of the paper is organised as follows.
-@sec-pre-theory provides the basic theoretical framework for hierarchical time series and forecasting, and Min Trace approach, introducing notations, terminologies, and motivations for alternative estimators.
+@sec-pre-theory provides the basic theoretical framework for hierarchical time series and forecasting, and Minimum Trace approach, introducing notations, terminologies, and motivations for alternative estimators.
 @sec-cov-est walks through the main covariance estimators this paper explores, and argues their strengths and weaknesses.
@@
 
-The cross-validation algorithm for NOVELIST is available in the ReconCov package [@ReconCov].
+<!-- The cross-validation algorithm for NOVELIST is available in the ReconCov package [@ReconCov]. -->
 
@@
 In each Monte Carlo replication ($M=500$), we fit univariate ARIMA models (base models) to the training observations using an automatic AICc minimization algorithm from @Hyndman2008-rf, implemented in the *fabletools* package [@O-Hara-Wild2024-we], generating incoherent 1–4-step base forecasts.
-We then reconcile these forecasts under three covariance estimators: the raw sample covariance (mint_sample), the shrinkage estimator (mint_shr), and the NOVELIST estimator (mint_n).
-
-All data generation, covariance estimation, and reconciliation routines were implemented in the ReconCov R package and is available under an open‐source license on GitHub [@ReconCov].
+We then reconcile these forecasts under different methods, and evaluate their point and probabilistic forecast accuracy on the test set.
 
@@
 
+# Conclusions and Future Work {#sec-conclusion}
+
+<!-- 
+- Aim to study the shortcomings of current MinT, a method attracting lots of attention in academia and practice.
+- Propose a NOVELIST, PC-adjusted variants, and multi-step covariance variants.
+- Empirical results on tourism data show that:
+  - MinT with shrinkage is a solid default for point forecasts. NOVELIST performs more reliably for probabilistic forecasts.
+  - PC adjustment with a single dominant factor consistently enhances performance in both point and probabilistic forecast and should be considered when a dominant latent factor is evident.
+  - Proportionality assumption still gives reasonably good results.
+  - More complex adjustments, such as multiple PCs or horizon-specific covariances, add estimation noise without clear benefits and can be omitted for parsimony.
+- Limitations & Future works:
+  - More complex simulation studies needed to tease out the differences between shrinkage and NOVELIST. No simulating from factor models.
+  - Not standardising the data for PC-variants methods, higher-level series can dominate the PCs.
+  - Explore other methods utilising cross-series information, eigenvalue/eigenvector structures,..
+  - Explore why MinT shrinkage fails for Winkler 95% score, especially at higher levels.
+  - h-step-ahead probabilistic forecast evaluation.
+-->
+
+This paper aims to address the limitations of the current Minimum Trace reconciliation method, a method that has garnered significant attention in both academic and practical forecasting contexts.
+We propose the NOVELIST estimator to address the lack of flexibility in the uniform shrinkage target of the traditional shrinkage estimator.
+Additionally, we introduce PC-adjusted variants of shrinkage and NOVELIST to take into account dominant latent factors, as well as multi-step covariance variants that relax the proportionality assumption.
+
+Empirical results on Australian domestic tourism data reveal several insights.
+For point forecasts, shrinkage (and its PC-adjusted version) remains a robust default choice, while NOVELIST demonstrates improved performance for probabilistic forecasts.
+Adjusting for a single dominant principal component consistently enhances performance across both point and probabilistic forecasts, suggesting its utility when such latent structures are present.
+More complex adjustments, such as incorporating multiple principal components or horizon-specific covariances, tend to introduce additional estimation noise without notable benefits.
+
+However, several limitations and avenues for future research remain.
+More intricate simulation studies are necessary to further highlight the differences between shrinkage and NOVELIST, particularly in scenarios involving factor models.
+The current implementation does not standardise data for PC-variant methods, which may lead to higher-level series largely influencing the principal components.
+Future work could also explore alternative methods that leverage cross-series information or eigenvalue/eigenvector structures.
+Additionally, understanding why MinT with shrinkage underperforms in the Winkler 95% score, especially at higher aggregation levels, needs further investigation.
+Finally, evaluating h-step-ahead probabilistic forecasts remains an open area for further research.
+
+All data generation, covariance estimation, and reconciliation routines were implemented in the ReconCov R package and is available under an open‐source license on GitHub [@ReconCov].
+
 \pagebreak

### 2025-10-15 thesis/paper_draft.qmd
refine abstract

@@
 To mitigate this issue, various forecast reconciliation approaches have been proposed in the literature, where the individual forecasts are adjusted to satisfy the aggregation constraints.
-Among these, **MinT** (Minimum Trace) is widely used, however, it requires a good estimate of the covariance matrix of the base forecast errors.
-The current practice is to use the shrinkage estimator (often shrinking toward a diagonal matrix), but it lacks flexibility and might not fully utilise the prominent latent structure presented.
-In this project, we aim to assess the forecasting performance of MinT when different covariance estimators are used, namely NOVELIST (NOVEL Integration of the Sample and Thresholded Covariance), PC-adjusted estimators (taking the latent factors directly into its construction), and others.
+Among these, MinT (Minimum Trace) is widely used, however, it requires a good estimate of the covariance matrix of the base forecast errors.
+The current practice is to use the shrinkage estimator (shrinking toward a diagonal matrix) for 1-step-ahead covariance estimation and proportionally scale it to approximate the h-step-ahead covariance matrix, but it leaves a question of whether this is appropriate for all real-world applications.
+We propose several alternative covariance estimators that address the shortcomings of the shrinkage estimator, including the NOVELIST estimator (which shrinks toward a soft-thresholded target), the PC-adjusted estimators (which utilises the latent factors structure), and estimators that relax the proportional scaling assumption.
+We evaluate the performance of MinT with these different covariance estimators in both point and probabilistic forecast reconciliation settings, and demonstrate their effectiveness and improvements over the shrinkage estimator in a complex, large-hierarchy dataset.
 

### 2025-10-15 thesis/paper_draft.qmd
enhance abstract, intro, and theorectical secs

--- 2025-10-15 thesis/paper_draft.qmd
@@
 A collection of time series connected via a set of linear constraints is known as hierarchical time series.
-Forecasting these series without respecting the hierarchical nature of the data can lead to incoherent forecasts across aggregation levels and lower accuracy.
-To mitigate this issue, various forecast reconciliation approaches have been proposed in the literature, where the individual forecasts are adjusted to satisfy the aggregation constraints.
-Among these, **MinT** (Minimum Trace) is widely used, however, it requires a good estimate of the covariance matrix of the base forecast errors.
-The current practice is to use the shrinkage estimator (often shrinking toward a diagonal matrix), but it lacks flexibility and might not fully utilise the prominent latent structure presented.
-In this project, we aim to assess the forecasting performance of MinT when different covariance estimators are used, namely NOVELIST (NOVEL Integration of the Sample and Thresholded Covariance), PC-adjusted estimators (taking the latent factors directly into its construction), and others.
+Forecasting these series without respecting the hierarchical nature of the data can lead to incoherent forecasts across aggregation levels and, in practice, reduced accuracy.
+Forecast reconciliation corrects this by adjusting base forecasts to satisfy such constraints.
+Among modern reconciliation methods, Minimum Trace (MinT) is widely used, however, it requires a good estimate of the forecast error covariance matrix.
+The current practice is to use the linear shrinkage toward a diagonal target to estimate 1-step-ahead covariance and proportionally scale it to approximate the h-step-ahead covariance matrix.
+This leaves a question of whether this method is appropriate for all real-world applications.
+We study alternative estimators that address the shortcomings of the current practice, including the NOVELIST estimator (shrinkage toward a soft-thresholded target), the PC-adjusted shrinkage (which utilises the latent factors structure), and horizon-specific estimators that relax proportional scaling.
+We evaluate MinT under these covariance estimates for both point and probabilistic reconciliation, and demonstrate their effectiveness and improvements over the shrinkage estimator in a complex, large-hierarchy dataset.
 
@@
 
-In practice, when forecasts are produced for all series (often called *base forecasts*), they typically violate the aggregation constraints observed in the data (the sum of all countries' sales does not equal the total sales).
+In practice, when forecasts are produced for all series (often called *base forecasts*), they typically violate the aggregation constraints observed in the data (e.g.
+the sum of all countries' sales does not equal the total sales).
 Such forecasts are called *incoherent*.
-First, they undermine downstream decisions that require internal consistency, and second, our forecast performance suffers.
-To overcome these issues, forecast reconciliation was introduced.
-Forecast reconciliation, a post-processing step, utilises the information from the hierarchical structure and data to adjust the initially produced base forecasts, so that the resulting forecasts are *coherent* (i.e., respecting the aggregation constraints).
+Incoherence undermines downstream decisions that require internal consistency and can degrade forecasting performance.
+To tackle this problem, forecast reconciliation was introduced.
+Forecast reconciliation, a post-processing step, utilises the information from the hierarchical structure and data to adjust the initially produced base forecasts, so that the resulting forecasts are *coherent* (i.e.
+respecting the aggregation constraints).
 It first introduced by @Hyndman2011-jv, and later developed by @Van_Erven2015-nx, @Hyndman2016-ic, @Ben-Taieb2019-yx, @Wickramasuriya2019-xq, @Wickramasuriya2020-uk, and others to enhance point forecast performance.
 @Athanasopoulos2024-as provided a comprehensive review of the literature on forecast reconciliation.
-Among the modern methods, the Min Trace (MinT) approach developed by @Wickramasuriya2019-xq provides good theoretical properties of minimising the total variance of the reconciled forecast errors, fast computational speed, and significant empirical improvements in many applications.
-It has become a standard method for forecast reconciliation, implemented in popular R and Python softwares/packages [@O-Hara-Wild2024-we; @Nixtla].
-
-A difficulty with MinT is in estimating the covariance matrix of the base forecast errors, especially for 2-step-ahead forecasts and beyond.
-This is a high-dimensional covariance estimation problem, where the number of series $n$ is often larger than the time dimension $T$.
-@Wickramasuriya2019-xq proposed using a shrinkage estimator with diagonal target from @Schafer2005-yw to estimate the 1-step-ahead covariance matrix, which is then scaled by a constant to approximate the h-step-ahead covariance matrix.
-While this approach is relatively simple and guarantees positive definiteness, it has three main shortcomings.
-First, it uniformly shrinks all off-diagonal elements towards zeros with equal weights, lacking flexibility to effectively preserve strong signals in the data.
-Second, in many real-world applications, hierarchical time series data often exhibit a prominent principal components structure, which is not fully utilised.
-Third, the proportionality relationship between the h-step-ahead and 1-step-ahead forecast error covariance matrices might not hold in practice.
-
-The theorectical advantage of MinT might not be achieved in finite sample settings if the covariance estimate is not close enough to the true covariance matrix.
-Despite its importance, not many researchers have explored this issue in the current literature, except @Carrara2025-rz, who proposed a Double Shrinkage estimator for MinT.
-The idea is to introduce a second shrinkage target, which incorporates conditional dependence information suggested by the hierarchical structure.
-This paper is still in its early stage, and does not fully address the shortcomings mentioned above.
-
-In the current literature, there are many advancements in high-dimensional covariance estimation that can be potentially useful for MinT.
-Building upon the shrinkage estimator, @Huang2019-ua introduces a soft-thresholded target matrix instead of the diagonal target.
-Extending beyond linear shrinkage, @Ledoit2012-ga proposed a nonlinear shrinkage estimator that replaces each sample eigenvalue with a data-driven nonlinear transformation of itself, which is further developed by @Ledoit2020-xt.
-On the other hand, @Fan2013-jz introduced an estimator that directly preserves the latent factors structure, decomposing the covariance matrix into a low-rank component and a sparse component, and applying thresholding to the latter.
-There are also approaches related to thresholding, such as the adaptive thresholding estimator for sparse covariance matrices by @Cai2011-jf.
-The list of modern high-dimensional covariance estimators is long, presenting many opportunities to explore their potential in improving the performance of MinT.
-
-The purpose of this paper is to explore the shortcomings of the shrinkage estimator and introduce alternative estimators that address these issues one by one.
-We evaluate the performance of MinT with these different covariance estimators in both point and probabilistic forecast reconciliation settings.
-In our empirical results from the complex, large-hierarchy dataset, we observe the following: (i) MinT with NOVELIST estimator with brings improvements over the shrinkage estimator under the probabilistic setting, as its predictive distributions have higher coverage rates; (ii) Covariance estimators that utilise the latent factors structure consistently produce better point and probabilistic forecasts than those that do not, when the data is evidently driven by such structure; (iii) Scaling the 1-step-ahead covariance matrix to approximate the h-step-ahead one gives significantly better performance than estimators that do not use this assumption, suggesting that the scaling assumption is not entirely incorrect.
-The estimators relaxing on the proportionality assumption that we propose can also be further extended to h-step-ahead probabilistic forecast reconciliation.
+Among the modern methods, the Min Trace (MinT) approach developed by @Wickramasuriya2019-xq is is widely used due to its strong theoretical properties for minimising total reconciled error variance, computational efficiency, and robust empirical performance.
+MinT is implemented in popular R and Python software ecosystems [@O-Hara-Wild2024-we; @Nixtla].
+
+A central difficulty for MinT is estimating the covariance matrix of base-forecast errors, particularly beyond one-step horizons.
+This is a high-dimensional estimation problem in which the number of series often exceeds the effective time dimension.
+A common practice, following @Wickramasuriya2019-xq, is to estimate the one-step covariance using linear shrinkage toward a diagonal target [@Schafer2005-yw], and to obtain multi-step covariances by proportional scaling of the one-step estimate.
+While convenient and guaranteed to produce a positive-definite estimator, this practice has three important shortcomings.
+First, the shrinkage is uniform across off-diagonals, applying a single penalty that may over-shrink genuine dependence and under-shrink noise.
+Second, many hierarchical data sets exhibit strong latent low-rank structure that is not explicitly exploited.
+Third, the proportionality relationship between the h-step-ahead and 1-step-ahead forecast error covariance matrices might not hold when error dynamics change with horizon.
+
+These limitations matter because the theoretical advantages of MinT depend on the quality of the covariance estimate available in finite samples.
+Despite its central role, there has been limited work on tailored covariance estimation for reconciliation.
+An exception is the recent double-shrinkage proposal in @Carrara2025-rz, which introduces an additional target designed to encode conditional dependence suggested by the hierarchy.
+This line of work still remains at an early stage and does not fully address the aforementioned limitations.
+
+Meanwhile, there has been substantial progress in high-dimensional covariance estimation that can be leveraged for MinT.
+Building on shrinkage, @Huang2019-ua proposed NOVELIST, which shrinks toward a soft-thresholded target rather than a diagonal one, improving flexibility to sparse targets.
+Extending beyond linear shrinkage, @Ledoit2012-ga (and further developed by @Ledoit2020-xt) introduced nonlinear shrinkage that replaces sample eigenvalues with data-driven nonlinear transformation of themselves.
+In a complementary direction, @Fan2013-jz introduce factor-based estimators that explicitly preserve latent low-rank structure while thresholding the idiosyncratic component.
+Related thresholding approaches, including hard thresholding [@Bickel2008-nh], generalised thresholding [@Rothman2009-yf], and adaptive thresholding [@Cai2011-jf], have also been developed.
+This growing toolkit provides multiple pathways to maximise the performance of MinT.
+
+This paper focuses on covariance estimation for MinT.
+We examine the shortcomings of the standard shrinkage plus scaling practice and introduce alternative estimators that address these issues individually and in combination.
+Specifically, we study NOVELIST as a more adaptive target-based shrinkage; principal-component-adjusted (PC-adjusted) estimators that exploit latent factor structure; and horizon-specific estimators that relax proportional scaling, including direct estimation from multi-step residuals.
+We evaluate MinT under these alternatives for both point and probabilistic reconciliation.
+In a large, complex real-world hierarchy, we find: first, NOVELIST improves probabilistic performance relative to linear shrinkage, including higher empirical coverage; second, PC-adjusted estimators consistently outperform other estimators when common components are strong, for both point and probabilistic metrics; third, proportional scaling of the one-step covariance remains a competitive baseline, indicating that the assumption is not universally invalid.
 
 The remainder of the paper is organised as follows.
-@sec-pre-theory provides the basic theoretical framework for hierarchical time series and forecasting, and Minimum Trace approach, introducing notations, terminologies, and motivations for alternative estimators.
-@sec-cov-est walks through the main covariance estimators this paper explores, and argues their strengths and weaknesses.
-@sec-simulation covers the simulation design and currently explores the performance of NOVELIST on MinT.
-@sec-empirical shows a real-world application of MinT with NOVELIST, which produces results that did not occured in the simulation settings, suggesting further inspection and analysis.
-
-<!-- Include more details on the gaps for each of these methods, such as bias, high variance? -->
-
-<!-- - Discuss the interests in MinT and how it has become a standard method for forecast reconciliation.
-
-- Discuss the MinT's reliance on a good estimate of the covariance matrix of base forecast errors and other gaps
-
-  - Empirical evidence of MinT under perform
-  - Comparison with other methods
-
-- Explain why this paper focus on exploring alternative covariance estimators for MinT. And is there any paper talk about this.
-
-  - Is better estimate of W_h really lead to better performance?
-
-- Talk about the paper outline and what will be covered in the following sections. -->
+@sec-pre-theory sets out the framework for hierarchical time series, forecast reconciliation, and MinT, and motivates the need for improved covariance estimation.
+@sec-cov-est presents the covariance estimators considered, outlining their strengths and limitations in the reconciliation context.
+@sec-simulation details the simulation design and examines the performance of NOVELIST compared to shrinkage.
+@sec-empirical dives into a real-world application with all proposed methods and highlights behaviours not observed in the simplified simulations.
 
@@
 
-*Hierarchical time series* are multivariate time series $\boldsymbol{y}_t \in \mathbb{R}^n$ organised in a structure where the series adheres to some *constraints*.
-For example, @fig-hierarchy illustrates a simple 2-level hierarchical structure with one top-level series $y_{Tot,t}$, two middle-level series $(y_{A,t}, y_{B,t})'$, and four bottom-level series $(y_{A1,t}, y_{A2,t}, y_{B1,t}, y_{B2,t})'$.
+*Hierarchical time series* are multivariate time series $\boldsymbol{y}_t \in \mathbb{R}^n$ organised in a structure where the series adheres to some constraints.
+@fig-hierarchy illustrates a simple two-level hierarchical structure with one top-level series $y_{Tot,t}$, disaggregating down to two level-1 series $(y_{A,t}, y_{B,t})'$, and to four bottom-level series $(y_{A1,t}, y_{A2,t}, y_{B1,t}, y_{B2,t})'$.
 Here, the *aggregation constraints* imply that $y_{Tot,t} = y_{A,t} + y_{B,t}$, $y_{A,t} = y_{A1,t} + y_{A2,t}$, and $y_{B,t} = y_{B1,t} + y_{B2,t}$.
@@
 
-The bottom-level (or most disaggregated) series are denoted as $\boldsymbol{b}_t \in \mathbb{R}^{n_b}$. Thus, the hierarchical time series can be represented as:
+The bottom-level (or most disaggregated) series are denoted as $\boldsymbol{b}_t \in \mathbb{R}^{n_b}$. Thus, the full vector of all series in the hierarchy can be represented as:
 
@@
 
-The matrix $\boldsymbol{S}$ encodes the aggregation constraints presented in the structure.
+The matrix $\boldsymbol{S}$ encodes the aggregation constraints implied by the structure.
 Hence, the columns of $\boldsymbol{S}$ span a linear subspace.
@@
 
-This setting is not restricted to hierarchical (nested) structures.
-When there are attributes of interest that are crossed, such as the company sales at any aggregation level (company-wise, city-wise, or outlet-wise) is also considered by kinds of products, the structure is described as a *grouped structure*.
-As illustrated in @fig-grouped, the aggregation or disaggregation paths are not unique.
-These constraints formed by the grouped structure can also be represented using a summing matrix $\boldsymbol{S}$.
-For simplicity, we refer to both of these structures as hierarchical structure, we will distinguish between them if and when it is necessary.
+This representation extends beyond hierarchical (nested) structures.
+When attributes of interest are crossed, such as the company sales at any aggregation level (company-wise, city-wise, or outlet-wise) is also considered by kinds of products, the structure is described as a *grouped structure*.
+In grouped systems, as illustrated in @fig-grouped, aggregation and disaggregation paths are not unique, but the linear constraints can still be written compactly through a summing matrix $\boldsymbol{S}$.
+For simplicity, we refer to both structures as hierarchical structure and distinguish between them when needed.
 
@@
 
-When we produce forecasts for each individual series, referred to as *base forecasts* $\hat{\boldsymbol{y}}_{t+h|t}$, they often do not respect the aggregation constraints, and thus are incoherent.
-Coherency can be achieved by linearly projecting the base forecasts onto the coherent subspace $\mathfrak{s}$ using a projection matrix $\boldsymbol{P}$: $\tilde{\boldsymbol{y}}_{t+h|t} = \boldsymbol{P} \hat{\boldsymbol{y}}_{t+h|t}$, where $\tilde{\boldsymbol{y}}_{t+h|t}$ are the *reconciled forecasts*.
-A schematic illustration of this projection is depicted in @fig-geometry.
+In practice, when we produce forecasts for each individual series, referred to as *base forecasts* $\hat{\boldsymbol{y}}_{t+h|t}$, they generally violate the aggregation constraints, and thus are incoherent.
+Coherency can be restored by linearly projecting the base forecasts onto the coherent subspace $\mathfrak{s}$ using a projection matrix $\boldsymbol{P}$: $\tilde{\boldsymbol{y}}_{t+h|t} = \boldsymbol{P} \hat{\boldsymbol{y}}_{t+h|t}$, where $\tilde{\boldsymbol{y}}_{t+h|t}$ are the *reconciled forecasts*.
 
 ![Geometry of probabilistic forecast reconciliation.
-The eliptical base forecast distribution is projected orthogonally onto the coherent subspace (purple line), resulting in the reconciled forecast distribution (red).
-The projection is defined by the projection matrix $\boldsymbol{P}$, and MinT allows oblique projections.
+The base forecast distribution is projected orthogonally onto the coherent subspace (purple line), resulting in the reconciled forecast distribution (red).
+The projection is defined by the projection matrix $\boldsymbol{P}$.
 Note that this figure is schematic since most applications are high-dimensional.](../figs/geometry_orthogonal.png){#fig-geometry width="60%"}
@@
 The idea is to map the base forecasts of all levels $\hat{\boldsymbol{y}}_{t+h|t}$ down into the bottom level, which is then aggregated to the higher levels by $\boldsymbol{S}$.
-Since the projection matrix $\boldsymbol{P}$ is idempotent and symmetric, $\boldsymbol{G}$ must satisfy the condition $\boldsymbol{S} \boldsymbol{G} \boldsymbol{S} = \boldsymbol{S}$.
-Thus, we generally have the mapping matrix $\boldsymbol{G} = (\boldsymbol{S}' \boldsymbol{M}^{-1} \boldsymbol{S})^{-1} \boldsymbol{S}' \boldsymbol{M}^{-1}$, for some positive definite matrix $\boldsymbol{M}$ [@Gamakumara2020-ml].
-
-When setting $\boldsymbol{M} = \boldsymbol{I}_n$, the identity matrix, we obtain the OLS method, which also corresponds to an orthogonal projection onto the coherent subspace (similar to @fig-geometry).
-Research have been done to introduce $\boldsymbol{M}$ that better utilise the inherent information from the observed data, to construct oblique projections and deliver better-performing forecasts.
+Since the projection matrix $\boldsymbol{P}$ is idempotent, $\boldsymbol{G}$ must satisfy the condition $\boldsymbol{S} \boldsymbol{G} \boldsymbol{S} = \boldsymbol{S}$.
+Within this class, a broad family of mapping matrix is given by: $\boldsymbol{G} = (\boldsymbol{S}' \boldsymbol{M}^{-1} \boldsymbol{S})^{-1} \boldsymbol{S}' \boldsymbol{M}^{-1}$, for some positive definite matrix $\boldsymbol{M} \in \mathbb{R}^{n \times n}$ [@Gamakumara2020-ml].
+
+When setting $\boldsymbol{M} = \boldsymbol{I}_n$, the identity matrix, this reduces to the OLS reconciliation, which corresponds to an orthogonal projection onto the coherent subspace (similar to @fig-geometry).
+A schematic illustration of this projection is depicted in @fig-geometry.
+Reconciliation takes the (possibly elliptical) base forecast distribution and projects it onto the coherent subspace, resulting in the reconciled forecast distribution.
+The choice of $\boldsymbol{M}$ determines the direction of the projection, which can yield oblique projection to deliver better-performing forecasts.
 
@@
 @Wickramasuriya2019-xq showed that by setting $\boldsymbol{M} = \boldsymbol{W}_h = \mathbb{E}( \hat{\boldsymbol{e}}_{t+h|t} \; \hat{\boldsymbol{e}}_{t+h|t}' )$, the covariance matrix of the $h$-step-ahead base forecast errors $\hat{\boldsymbol{e}}_{t+h|t} = \boldsymbol{y}_{t+h} - \hat{\boldsymbol{y}}_{t+h|t}$, we essentially minimise the total variance of the reconciled forecast errors across all series.
-In which, it is minimising the trace of $\text{Var}[y_{t+h} - \tilde{y}_{t+h|t}] = \boldsymbol{S} \boldsymbol{G}_h \boldsymbol{W}_h \boldsymbol{G}_h' \boldsymbol{S}'$.
-This method is called Minimum Trace (MinT) reconciliation.
+Equivalently, MinT is the unique linear-unbiased reconciler that minimises the trace of $\text{Var}[y_{t+h} - \tilde{y}_{t+h|t}] = \boldsymbol{S} \boldsymbol{G}_h \boldsymbol{W}_h \boldsymbol{G}_h' \boldsymbol{S}'$.
+This method is thus called Minimum Trace (MinT) reconciliation.
 The matrix $\boldsymbol{G}_h$ is thus given by:
@@
 \boldsymbol{G}_h = (\boldsymbol{S}' \boldsymbol{W}_h^{-1} \boldsymbol{S})^{-1}
-\boldsymbol{S}' \boldsymbol{W}_h^{-1} .
-$$
-
-This is initially derived under point forecast reconciliation setting.
-However, @Wickramasuriya2024-mb showed that under the Gaussian base forecast distribution assumption, MinT also yields a Gaussian reconciled forecast distribution.
-Let $h$-step-ahead base forecast distribution be $\hat{\boldsymbol{y}}_{t+h|t} \sim \mathcal{N}(\hat{\boldsymbol{y}}_{t+h|t}, \boldsymbol{W}_h)$, then the reconciled forecast distribution is given by $\tilde{\boldsymbol{y}}_{t+h|t} \sim \mathcal{N}(\tilde{\boldsymbol{y}}_{t+h|t}, \boldsymbol{S} \boldsymbol{G}_h \boldsymbol{W}_h \boldsymbol{G}_h' \boldsymbol{S}')$.
-Additionally, MinT also minimises the logarithmic score of the reconciled predictive distribution among all projection matrices.
-
-For probabilistic forecasts, this paper focuses on evaluating the performance of MinT with different covariance estimators under the Gaussian framework.
-Even so, it is worth to mention that the methods can be extended to non-Gaussian settings by bootstrapping [@Gamakumara2020-ml; @Panagiotelis2023-fs].
+\boldsymbol{S}' \boldsymbol{W}_h^{-1},
+$$
+
+provided that $\boldsymbol{W}_h$ is positive definite.
+Intuitively, $\boldsymbol{W}_h^{-1}$ down-weights the base forecasts of series with high uncertainty, and up-weights those with low uncertainty, when mapping to the bottom level.
+
+Although originally developed for point reconciliation, @Wickramasuriya2024-mb showed that MinT extends naturally to the probabilistic setting when base predictive distributions are Gaussian.
+If $h$-step-ahead base forecast distribution be $\hat{\boldsymbol{y}}_{t+h|t} \sim \mathcal{N}(\hat{\boldsymbol{y}}_{t+h|t}, \boldsymbol{W}_h)$, then reconciliation gives $\tilde{\boldsymbol{y}}_{t+h|t} = \boldsymbol{S} \boldsymbol{G}_h \hat{\boldsymbol{y}}_{t+h|t} \sim \mathcal{N}(\tilde{\boldsymbol{y}}_{t+h|t}, \boldsymbol{S} \boldsymbol{G}_h \boldsymbol{W}_h \boldsymbol{G}_h' \boldsymbol{S}')$.
+Within the class of linear-unbiased reconcilers, MinT also minimises the negative log score (equivalently, the Gaussian log predictive score) of the reconciled distribution.
+
+In this paper we adopt this Gaussian framework to compare covariance estimators for MinT in both point and probabilistic reconciliation.
+It is also worth to mention that the methods can be extended to non-Gaussian settings by bootstrapping [@Gamakumara2020-ml; @Panagiotelis2023-fs], which is beyond the scope of this paper.
 
@@
 
-The MinT solution hinges on a reliable, positive-definite estimate of $\boldsymbol{W}_h$, which comes in both the mapping matrix $\boldsymbol{G}_h$ and the reconciled forecast variance $\boldsymbol{S} \boldsymbol{G}_h \boldsymbol{W}_h \boldsymbol{G}_h' \boldsymbol{S}'$.
-
-However, the covariance matrix $\boldsymbol{W}_h$ is often not available in theoretical derivation, and is challenging to estimate in high-dimensional setting where the number of series $n$ is larger than the time dimension $T$.
-To tackle this issue, the original paper @Wickramasuriya2019-xq assumed a proportionality relationship between $\hat{\boldsymbol{W}}^g_h = k_h g( \hat{\boldsymbol{W}}_1 )$, where $\hat{\boldsymbol{W}}_1$ is the covariance matrix of the in-sample 1-step-ahead base residuals (to approximate $\boldsymbol{W}_1$) and $k_h$ is a positive scaling constant (which will be cancelled out in point-forecast reconciliation).
+The performance of MinT hinges on a reliable, positive-definite estimate of $\boldsymbol{W}_h$, which comes in both the mapping matrix $\boldsymbol{G}_h$ and the reconciled forecast variance $\boldsymbol{S} \boldsymbol{G}_h \boldsymbol{W}_h \boldsymbol{G}_h' \boldsymbol{S}'$.
+
+However, the covariance matrix $\boldsymbol{W}_h$ is often not available in closed-form, and is challenging to estimate in high-dimensional setting where the number of series $n$ is larger than the time dimension $T$.
+To tackle this issue, the original paper @Wickramasuriya2019-xq assumed a proportionality relationship between $\hat{\boldsymbol{W}}^g_h = k_h g( \hat{\boldsymbol{W}}_1 )$, where $\hat{\boldsymbol{W}}_1$ is the covariance matrix of the in-sample 1-step-ahead base residuals (to approximate $\boldsymbol{W}_1$) and $k_h > 0$ is a scaling constant (which will be algebraically cancelled out in point-forecast reconciliation).
 The function $g(.)$ is a covariance estimator that produces a positive-definite matrix, the main focus of this paper.
 
-The authors suggested using the shrinkage estimator with diagonal target from @Schafer2005-yw, given by:
+The recommended choice for $g(.)$ in the original work is the shrinkage estimator with diagonal target from @Schafer2005-yw:
 
@@
 where $\text{diag}( \hat{\boldsymbol{W}}_1 )$ comprises only the diagonal elements of $\hat{\boldsymbol{W}}_1$.
-We refer to any $\lambda_S \in [0,1]$ as the shrinkage intensity parameter, the subscript specifies the shrinkage estimator it belongs to.
+We refer to any $\lambda_S \in [0,1]$ as the shrinkage intensity of the shrinkage estimator.
 This approach shrinks the covariance matrix $\hat{\boldsymbol{W}}_1$ towards its diagonal matrix, meaning the off-diagonal elements are shrunk towards zero while the diagonal ones remain unchanged.
 
-@Schafer2005-yw also proposed an estimate of the optimal shrinkage intensity parameter $\lambda_S$:
+@Schafer2005-yw also provided an closed-form estimate of the optimal shrinkage intensity parameter $\lambda_S$:
 
@@
 
-where $\hat{r}_{ij}$ is the $ij$th element of $\hat{\boldsymbol{R}}_1$, the 1-step-ahead sample correlation matrix (obtained from $\hat{\boldsymbol{W}}_1$).
+where $\hat{r}_{ij}$ is the $i,j$-th element of $\hat{\boldsymbol{R}}_1$, the 1-step-ahead sample correlation matrix (obtained from $\hat{\boldsymbol{W}}_1$).
 The optimal estimate is obtained by minimising $MSE(\hat{\boldsymbol{W}}_1) = Bias(\hat{\boldsymbol{W}}_1)^2 + Var(\hat{\boldsymbol{W}}_1)$.
-More specifically, we trade the unbiasedness of the sample covariance matrix for a lower variance.
-
-Despite its relative simplicity and guaranteed positive definiteness, MinT with shrinkage estimator has three main shortcomings.
+More specifically, we trade a small bias for a substantial variance reduction, which is especially valuable in high dimension.
+
+Despite its simplicity and guaranteed positive-definiteness, MinT coupled with diagonal-target shrinkage presents three important limitations.
 
@@
 
-The shrinkage estimator shrinks all off-diagonal elements towards zeros with equal weights $\lambda_S$.
-We might prefer to better preserve strong signals, and largely reduce the effects of small, noisy correlations.
-This is especially true in high-dimensional settings, where only a few series to be truly correlated; or due to the aggregation effects, bottom-level series will correlate more strongly with their parents.
-The shrinkage estimator lacks this flexibility.
+Linear shrinkage shrinks all off-diagonal elements towards zeros with equal weights $\lambda_S$.
+The resulting penalty is global and non-adaptive: strong, genuinely systematic correlations are shrunk at the same rate as weak, noisy ones.
+In hierarchical and grouped systems, this can be problematic.
+Aggregation naturally induces stronger dependence among related nodes (for example, a region and a neighbouring region, or a region and its parent state), while many unrelated pairs exhibit near-zero correlations.
+A uniform penalty may over-shrink informative co-movements and under-shrink idiosyncratic noise, reducing reconciliation efficiency.
 
@@
 
-The hierarchical time series data often exhibit a prominent principal components structure, which is not fully taken advantage.
+Many real-world hierarchical data sets exhibit a prominent low-rank (factor) structure.
+
+In Australian domestic tourism, for example, national and state-level movements often load on a small number of common components, with residual dependence concentrated in local idiosyncrasies.
+A scree plot of the one-step residual covariance typically shows a marked elbow after a handful of principal components, indicating strong latent structure.
+Diagonal-target shrinkage is factor-unaware: it neither preserves the low-rank common subspace nor differentially shrinks the idiosyncratic remainder, and so can be inefficient when common factors dominate.
+
 Taking an example of the Australian domestic overnight trips data set [@tourism-au], where the national trips are disaggregated into states and territories, and further into regions.
-We then fit ETS models to all series, using the algorithm from Fabletools R package [@O-Hara-Wild2024-we], and compute the one-step-ahead in-sample base forecast residual covariance matrix $\hat{\boldsymbol{W}}_1$.
-The twenty largest eigenvalues of the covariance matrix are plotted in @fig-eigen.
-We can see that the point of inflexion occurs at the component with 5th largest eigenvalue, indicating a prominent principal components structure.
-
-![Twenty largest eigenvalues of one-step-ahead in-sample base forecast error covariance, Australian domestic overnight trips](D:/Github/Recon_Honours_Thesis/figs/eigen_tourism.png){#fig-eigen width="70%"}
+The tourism activities might be driven by a few common factors, such as economic conditions, fuel prices, or major events, which might be left in the residuals after fitting the base models.
+From @fig-eigen, a scree plot of the one-step residual covariance $\hat{\boldsymbol{W}}_1$ shows a marked elbow after a handful of principal components (largest eigenvalues), indicating strong latent structure.
+
+![Twenty largest eigenvalues of one-step-ahead in-sample base forecast error covariance, Australian domestic overnight trips.
+The point of inflection occurs around the fifth largest eigenvalue.](D:/Github/Recon_Honours_Thesis/figs/eigen_tourism.png){#fig-eigen width="70%"}
 
@@
 
-The proportionality relationship $\hat{\boldsymbol{W}}^g_h = k_h g( \hat{\boldsymbol{W}}_1 )$ might not hold in practice.
-The covariance structure of h-step-ahead base forecast errors can be different from that of 1-step-ahead ones.
-
-In the next sections, we will explore several options that tackle these issues one by one.
+The proportionality relationship $\hat{\boldsymbol{W}}^g_h = k_h g( \hat{\boldsymbol{W}}_1 )$ enforces horizon-invariant cross-series dependence of the forecast errors up to a scalar factor.
+In many applications, the shape of the error covariance might change with horizon due to evolving error dynamics or horizon-specific interactions among series.
+Proportional scaling may therefore misrepresent multi-step dependence.
+
+Additionally, while the scalar factor cancels in point reconciliation, it directly controls dispersion in probabilistic reconciliation and can lead to miscalibrated predictive distributions.
+However, we will not explore this issue in this paper, and leave it for future research.
+
+In the next sections, we investigate covariance estimators that address these issues.
+We consider a estimator that mitigate uniform shrinkage, factor-aware adjustments that preserve low-rank structure while shrinking idiosyncratic noise, and horizon-specific estimators that relax proportional scaling.
 

### 2025-10-15 thesis/paper_draft.qmd
enhance cov ests

@@
 
-In the next sections, we investigate covariance estimators that address these issues.
-We consider a estimator that mitigate uniform shrinkage, factor-aware adjustments that preserve low-rank structure while shrinking idiosyncratic noise, and horizon-specific estimators that relax proportional scaling.
+In the next sections, we introduce alternative covariance estimators that address these limitations individually and in combination.
 
@@
 
-The NOVELIST (NOVEL Integration of the Sample and Thresholded Covariance) estimator, proposed by @Huang2019-ua, introduces a way to control the target matrix's sparsity, retaining strong correlations while discarding weak, noisy effects.
-NOVELIST offers more flexibility than the shrinkage estimator, which is useful when we believe that only a few variables are truly correlated.
-
-The method is based on the idea of soft-thresholding the sample covariance matrix, then performing shrinkage towards this thresholded version.
-This introduces an extra parameter, the threshold $\delta$, which is used to control the amount of soft-thresholding.
-The NOVELIST estimator is given by:
+NOVELIST (NOVEL Integration of the Sample and Thresholded Covariance) estimator, proposed by @Huang2019-ua, introduces a parameter to control the sparsity level in the target toward which the sample matrix is shrunk.
+By soft-thresholding small correlations before shrinkage, NOVELIST preserves strong signals while attenuating weak, noisy ones, addressing the uniform, non-adaptive nature of diagonal-target shrinkage.
+
+The construction proceeds in two steps.
+First, apply elementwise soft-thresholding to the sample correlation matrix; second, shrink the sample correlation toward this thresholded target.
+Working in correlation space avoids rescaling issues and keeps diagonal entries at one.
+The method introduces an extra parameter, the threshold $\delta$, which is used to control the amount of soft-thresholding.
+The NOVELIST estimator for covariance matrix is given by:
 
@@
 
-where $\hat{\boldsymbol{W}}_{1, \delta}$ is the thresholded version of $\hat{\boldsymbol{W}}_{1}$. By convenient setting, we can rewrite it in terms of correlation:
+where $\hat{\boldsymbol{W}}_{1, \delta}$ is the thresholded version of $\hat{\boldsymbol{W}}_{1}$. By convenient setting as discussed above, we rewrite it in terms of correlation:
 
@@
 where $\delta \in [0,1]$ is the threshold parameter.
-For a given threshold $\delta$, @Huang2019-ua derived an analytical expression for the optimal shrinkage intensity parameter $\lambda(\delta)$ using Ledoit-Wolf's lemma \[@Ledoit2003-qv\], following similar logic to @Schafer2005-yw.
+For a given threshold $\delta$, @Huang2019-ua derived an analytical expression for the optimal shrinkage intensity parameter $\lambda(\delta)$, following similar logic to @Schafer2005-yw.
 It can be computed as:
@@
 
-On the other hand, the optimal threshold $\hat \delta$ does not have a closed-form solution, and is typically obtained by executing a rolling-window cross-validation procedure.
+On the other hand, the optimal threshold $\hat \delta$ does not have a closed-form solution, and is typically obtained by rolling-window cross-validation procedure.
 The idea is to find the threshold $\delta^*$, with the corresponding $\lambda^*$ and $\hat{\boldsymbol{R}}^{N}_{1}(\delta^*, \lambda^*)$, that minimises the average out-of-sample 1-step-ahead reconciled forecast mean squared error over all windows.
@@
 
-We also tested out minimising 1- to h-step-ahead overall MSE in the cross-validation procedure.
-Surprisingly, it returns almost the same best threshold parameter is in the 1-step-ahead case above.
 Note that when $\delta \in \bigl[ \text{max}_{i \neq j}|\hat{r}_{1,ij}|, \; 1 \bigr]$, the NOVELIST estimator collapses to the shrinkage estimator, and when $\delta = 0$, it becomes the sample covariance matrix.
@@
 
+*Remark.* Minimising a multi-step-ahead forecast error metric in the cross-validation procedure oftern yields a $\hat{\delta}^*$ that is close to the 1-step-ahead case.
+
 <!-- The cross-validation algorithm for NOVELIST is available in the ReconCov package [@ReconCov]. -->
@@
 
-To utilise the latent factors structure for better shrinkage, the PC-adjusted method takes the latent factors directly into its construction, and is appealing when there are common drivers in the time series within the hierarchy, as we saw in the Australian tourism example.
+To exploit latent common components explicitly, the PC-adjusted method takes the latent factors directly into its construction.
 It starts by decomposing the covariance matrix $\hat{\boldsymbol{W}}_1$ into a prominent principle components part (low-rank) and a orthogonal complement part $\hat{\boldsymbol{W}}^K_{1}$ (the correlation matrix after removing the first $K$ principal components).
@@
 where $g(.)$ is either the shrinkage or NOVELIST estimator, $\hat{\gamma}_k$ and $\hat{\boldsymbol{\xi}}_k$ are the $k$-th largest eigenvalue and the corresponding eigenvector of the sample covariance matrix, respectively.
-Similar to the NOVELIST estimator, its PC-adjusted variant $\hat{\boldsymbol{W}}^{N, K}_{1}$ requires a cross-validation procedure and adjustment to obtain positive definiteness.
+The number of principal components $K$ can be selected using information criteria.
+
+Similar to the NOVELIST estimator, its PC-adjusted variant $\hat{\boldsymbol{W}}^{N, K}_{1}$ requires a cross-validation procedure to select the threshold paramter and adjustment to obtain positive definiteness.
+
 ## Scaled Variance
 
-To address the potential issue of the proportionality relationship $\hat{\boldsymbol{W}}^g_h = k_h g( \hat{\boldsymbol{W}}_1 )$ not holding in practice, we can relax the assumption by allowing the variances to scale differently.
-This is done by scaling the $1$-step-ahead correlation matrix back to $h$-step-ahead covariance matrix using the h-step-ahead standard deviations.
+To address the potential issue of the proportionality relationship $\hat{\boldsymbol{W}}^g_h = k_h g( \hat{\boldsymbol{W}}_1 )$ not holding in practice.
+A simple relaxation retains the $1$-step-ahead correlation shape but allows the variances to scale differently with horizon.
 The scaled variance estimator is given by:
@@
 
-If NOVELIST is used to produce $h$-step-ahead reconciled forecasts, the cross-validation procedure is slightly modified to evaluate the out-of-sample reconciled forecast MSE using $\hat{\boldsymbol{W}}^{N, sv}_{h}$ instead of $\hat{\boldsymbol{W}}^{N}_{1}$.
+When NOVELIST is used to produce $h$-step-ahead reconciled forecasts, the cross-validation procedure is slightly modified to evaluate the out-of-sample reconciled forecast MSE using $\hat{\boldsymbol{W}}^{N, sv}_{h}$ instead of $\hat{\boldsymbol{W}}^{N}_{1}$.
 
@@
 
-Another alternative is not to rely on the assumption of proportionality at all, and directly estimate the covariance matrix from the h-step-ahead base forecast errors:
+Another alternative is to directly estimate the covariance matrix from the h-step-ahead base forecast errors, without assuming any proportionality relationship:
 
@@
 where $\hat{\boldsymbol{W}}_h$ is the covariance matrix of in-sample h-step-ahead base forecast residuals, and $g(.)$ is either the shrinkage or NOVELIST estimator.
+
 Similar to the scaled variance approach, if NOVELIST is used, the cross-validation procedure is modified to compute $\hat{\boldsymbol{W}}^{N}_{h}$ and evaluate the out-of-sample reconciled forecast MSE using it.
@@
 
-Taken together, the evidence supports the use of reconciliation for both point and probabilistic forecasts in this high-dimensional setting.
+<!-- Taken together, the evidence supports the use of reconciliation for both point and probabilistic forecasts in this high-dimensional setting.
 For point forecasts, MinT with shrinkage is a solid default; for probabilistic forecasts, NOVELIST performs more reliably.
@@
 More complex adjustments, such as multiple PCs or horizon-specific covariances, add estimation noise without clear benefits and can be omitted for parsimony.
+-->
 

### 2025-10-20 thesis/paper_draft.qmd
refine abstract & intro from George feedback

@@
 Among modern reconciliation methods, Minimum Trace (MinT) is widely used, however, it requires a good estimate of the forecast error covariance matrix.
-The current practice is to use the linear shrinkage toward a diagonal target to estimate 1-step-ahead covariance and proportionally scale it to approximate the h-step-ahead covariance matrix.
+The current practice is to use the linear shrinkage towards a diagonal target.
+Furthermore, the covariance estimate is based on 1-step-ahead residuals, then proportionally scale it to approximate the h-step-ahead covariance matrix.
 This leaves a question of whether this method is appropriate for all real-world applications.
-We study alternative estimators that address the shortcomings of the current practice, including the NOVELIST estimator (shrinkage toward a soft-thresholded target), the PC-adjusted shrinkage (which utilises the latent factors structure), and horizon-specific estimators that relax proportional scaling.
-We evaluate MinT under these covariance estimates for both point and probabilistic reconciliation, and demonstrate their effectiveness and improvements over the shrinkage estimator in a complex, large-hierarchy dataset.
+We study the shortcomings of current practice and propose alternative covariance estimators, including the NOVELIST estimator (shrinkage towards a soft-thresholded target), PC-adjusted shrinkage (which utilises latent factor structures), and horizon-specific estimators that relax proportional scaling.
+We evaluate MinT using these covariance estimates for both point and probabilistic reconciliation, and demonstrate their effectiveness and improvements over the shrinkage estimator in a complex, large-hierarchy dataset.
 
@@
 In time series forecasting, aggregation occurs in a variety of settings.
-For example, Starbucks operates in many countries, and each country has multiple cities where they have outlets.
-The sales data is structured *hierarchically*: the top level is the total sales across all countries, which are the sum of sales in each country, which in turn are the sum of sales in each city, and finally the sum of sales from each outlet in the city.
+For example, Starbucks Corporation operates in many countries, and each country has multiple cities where they have outlets.
+The sales data is structured *hierarchically*: the top level is the company's total sales, which disaggregates into sales by country, then sales by city within each country, and finally down to sales by individual outlet within each city.
 As a result, there are over 50,000 sales series across all *aggregation levels*, and decision makers need forecasts at each level to manage inventory and plan marketing strategies effectively.
-The hierarchy can be even more complex if we consider the sales of different kinds of drinks (e.g., coffees, teas, refreshers) at each aggregation level.
-Such hierarchical structures also arises in many other decision-making contexts, from supply chains [@Seaman2022-bb; @Angam2025-od] and energy planning [@Di-Modica2021-ad], to macroeconomics [@El-Gemayel2022-hx; @Li2019-vv] and tourism analysis [@Athanasopoulos2009-lp].
+The hierarchy can be even more complex if we consider the sales of different categories of products (e.g., beverage, food, etc.) and the sales of product within each category (e.g., latte, cappuccino, etc.) at each aggregation level.
+In this case, the structure is called a *grouped structure*, where the aggregation paths are not unique.
+Such structures also arises in many other decision-making contexts, from supply chains [@Seaman2022-bb; @Angam2025-od] and energy planning [@Di-Modica2021-ad], to macroeconomics [@El-Gemayel2022-hx; @Li2019-vv] and tourism analysis [@Athanasopoulos2009-lp].
 Stakeholders in these settings need forecasts at several aggregation levels to allocate resources and manage risk.
 
-In practice, when forecasts are produced for all series (often called *base forecasts*), they typically violate the aggregation constraints observed in the data (e.g.
-the sum of all countries' sales does not equal the total sales).
+In practice, when forecasts are produced for all series (often called *base forecasts*), they typically violate the aggregation constraints observed in the data (e.g., the sum of all countries' sales forecasts does not equal the total sales forecast).
 Such forecasts are called *incoherent*.
@@
 To tackle this problem, forecast reconciliation was introduced.
-Forecast reconciliation, a post-processing step, utilises the information from the hierarchical structure and data to adjust the initially produced base forecasts, so that the resulting forecasts are *coherent* (i.e.
-respecting the aggregation constraints).
-It first introduced by @Hyndman2011-jv, and later developed by @Van_Erven2015-nx, @Hyndman2016-ic, @Ben-Taieb2019-yx, @Wickramasuriya2019-xq, @Wickramasuriya2020-uk, and others to enhance point forecast performance.
-@Athanasopoulos2024-as provided a comprehensive review of the literature on forecast reconciliation.
-Among the modern methods, the Min Trace (MinT) approach developed by @Wickramasuriya2019-xq is is widely used due to its strong theoretical properties for minimising total reconciled error variance, computational efficiency, and robust empirical performance.
+Forecast reconciliation, a post-processing step, utilises the information from the hierarchical structure and data to adjust the initially produced base forecasts, so that the resulting *reconciled forecasts* are *coherent* (i.e., respecting the aggregation constraints).
+It was first introduced by @Hyndman2011-jv, and later developed by @Van_Erven2015-nx, @Hyndman2016-ic, @Ben-Taieb2019-yx, @Wickramasuriya2019-xq, @Wickramasuriya2020-uk, and others that focus point forecast reconciliation.
+Recognising the important of probabilistic forecasting, probabilistic forecast reconciliation was studied by @Shang2017-ux, @Jeon2019-wm, and @Ben-Taieb2021-bn, and later formally defined by @Panagiotelis2023-fs.
+@Athanasopoulos2024-as provide a comprehensive review of the literature on forecast reconciliation.
+Among the modern methods, the Min Trace (MinT) approach developed by @Wickramasuriya2019-xq is is widely used due to its strong theoretical properties for minimising total reconciled forecast error variance, computational efficiency, and robust empirical performance.
+MinT is later extended to probabilistic reconciliation by @Wickramasuriya2024-mb, showing that it also minimises the negative log score of the reconciled distribution under Gaussian assumptions.
 MinT is implemented in popular R and Python software ecosystems [@O-Hara-Wild2024-we; @Nixtla].
 
-A central difficulty for MinT is estimating the covariance matrix of base-forecast errors, particularly beyond one-step horizons.
-This is a high-dimensional estimation problem in which the number of series often exceeds the effective time dimension.
-A common practice, following @Wickramasuriya2019-xq, is to estimate the one-step covariance using linear shrinkage toward a diagonal target [@Schafer2005-yw], and to obtain multi-step covariances by proportional scaling of the one-step estimate.
-While convenient and guaranteed to produce a positive-definite estimator, this practice has three important shortcomings.
+A central difficulty for MinT is estimating the covariance matrix of base-forecast errors, particularly beyond one-step forecast horizons.
+This is a high-dimensional estimation problem in which the number of series often exceeds the time dimension.
+A common practice, following @Wickramasuriya2019-xq, is to estimate the 1-step-ahead covariance from the residuals, using linear shrinkage toward a diagonal target [@Schafer2005-yw].
+Then, proportionally scale this estimate to approximate the multi-step-ahead covariance matrix.
+While convenient and guaranteed to produce a positive-definite estimate, this practice has three important shortcomings.
 First, the shrinkage is uniform across off-diagonals, applying a single penalty that may over-shrink genuine dependence and under-shrink noise.
-Second, many hierarchical data sets exhibit strong latent low-rank structure that is not explicitly exploited.
+Second, many hierarchical data sets exhibit strong latent low-rank structures that can be explicitly exploited.
 Third, the proportionality relationship between the h-step-ahead and 1-step-ahead forecast error covariance matrices might not hold when error dynamics change with horizon.
@@
 We examine the shortcomings of the standard shrinkage plus scaling practice and introduce alternative estimators that address these issues individually and in combination.
-Specifically, we study NOVELIST as a more adaptive target-based shrinkage; principal-component-adjusted (PC-adjusted) estimators that exploit latent factor structure; and horizon-specific estimators that relax proportional scaling, including direct estimation from multi-step residuals.
+Specifically, we study NOVELIST as a more adaptive target-based shrinkage; principal-component-adjusted (PC-adjusted) estimators that exploit latent factor structures; and horizon-specific estimators that relax proportional scaling, including direct estimation from multi-step residuals.
 We evaluate MinT under these alternatives for both point and probabilistic reconciliation.
-In a large, complex real-world hierarchy, we find: first, NOVELIST improves probabilistic performance relative to linear shrinkage, including higher empirical coverage; second, PC-adjusted estimators consistently outperform other estimators when common components are strong, for both point and probabilistic metrics; third, proportional scaling of the one-step covariance remains a competitive baseline, indicating that the assumption is not universally invalid.
+In a large, complex real-world hierarchy, our findings reveal three main insights.
+First, NOVELIST improves probabilistic performance relative to linear shrinkage, including higher empirical coverage.
+Second, PC-adjusted estimators consistently outperform other estimators when common components are strong, for both point and probabilistic metrics.
+Third, proportional scaling of the one-step covariance remains a competitive baseline, indicating that the assumption may not be universally invalid.
 
@@
 @sec-cov-est presents the covariance estimators considered, outlining their strengths and limitations in the reconciliation context.
+@sec-scores defines the evaluation metrics for point and probabilistic reconciliation.
 @sec-simulation details the simulation design and examines the performance of NOVELIST compared to shrinkage.

### 2025-10-20 thesis/paper_draft.qmd
codes for plotting sim results (replacing the images)

@@
 
-path <- "D:/Github/Recon_Honours_Thesis"
+path <- here::here()
 knitr::opts_knit$set(root.dir = path)
@@
 ```{r load results}
-path_sim <- "D:/Github/Recon_Honours_Analysis/sim/thesis_sim/"
-
-sim1_1 <- readRDS("./thesis/lib/S4-2-1_T50_M500.rds")
-sim1_2 <- readRDS("./thesis/lib/S4-2-1_T300_M500.rds")
-
-sim2_1 <- readRDS("./thesis/lib/S36-6-1_T50_M500.rds")
-sim2_2 <- readRDS("./thesis/lib/S36-6-1_T300_M500.rds")
-
-sim_dense_50 <- readRDS("./thesis/lib/S100-10-3-1_T50_M200_dense.rds")
-sim_dense_300 <- readRDS("./thesis/lib/S100-10-3-1_T300_M200_dense.rds")
-
-sim_sparse_50 <- readRDS("./thesis/lib/S100-10-3-1_T50_M500_sparsebwgrp.rds")
-sim_sparse_300 <- readRDS("./thesis/lib/S100-10-3-1_T300_M200_sparsebwgrp.rds")
+path_lib <- paste0(path, "/thesis/lib/")
+
+sim1_1 <- readRDS(paste0(path_lib, "S4-2-1_T50_M500.rds"))
+sim1_2 <- readRDS(paste0(path_lib, "S4-2-1_T300_M500.rds"))
+
+sim2_1 <- readRDS(paste0(path_lib, "S36-6-1_T50_M500.rds"))
+sim2_2 <- readRDS(paste0(path_lib, "S36-6-1_T300_M500.rds"))
+
+sim_dense_50 <- readRDS(paste0(path_lib, "S100-10-3-1_T50_M200_dense.rds"))
+sim_dense_300 <- readRDS(paste0(path_lib, "S100-10-3-1_T300_M200_dense.rds"))
+
+sim_sparse_50 <- readRDS(paste0(path_lib, "S100-10-3-1_T50_M500_sparsebwgrp.rds"))
+sim_sparse_300 <- readRDS(paste0(path_lib, "S100-10-3-1_T300_M200_sparsebwgrp.rds"))
 
@@
 sim3_2 <- sim_dense_300
+
+sim_MSE <- readRDS(paste0(path_lib, "sim_MSE.rds"))
+sim_prob <- readRDS(paste0(path_lib, "sim_prob.rds"))
 ```
@@
 
-In our first set of experiments, we examine how MinT combined with the different estimators perform as the hierarchy expands.
+In our main experiments, we examine how MinT combined with the different estimators perform as the hierarchy expands.
 We generate synthetic data from the same VAR(1) framework described earlier, but vary the number of bottom‐level series, $n_b$, across two structures: a small structure with six groups of six bottom series ($n_b = 6$x$6 = 36$), and a much larger configuration with two groups of fifty ($n_b = 2$x$50 = 100$).
@@
 
-![Percentage relative improvement in MSE of reconciled forecasts over the base forecasts in the 6 by 6 case (top row) and the 2 by 50 case (bottom row), T=50 (left column) and T=300 (right column), for 1- to 4-step-ahead forecasts.
-The positive (negative) entries indicate a decrease (increase) in MSE relative to base.](D:/Github/Recon_Honours_Thesis/figs/sim_MSE.png){#fig-sim-results-1}
+<!-- ![Percentage relative improvement in MSE of reconciled forecasts over the base forecasts in the 6 by 6 case (top row) and the 2 by 50 case (bottom row), T=50 (left column) and T=300 (right column), for 1- to 4-step-ahead forecasts.
+The positive (negative) entries indicate a decrease (increase) in MSE relative to base.](D:/Github/Recon_Honours_Thesis/figs/sim_MSE.png){#fig-sim-results-1} -->
+
+```{r}
+#| label: fig-sim-results-1
+#| fig-cap: "Percentage relative improvement in MSE of reconciled forecasts over the base forecasts in the 6 by 6 case (top row) and the 2 by 50 case (bottom row), T=50 (left column) and T=300 (right column), for 1- to 4-step-ahead forecasts.
+The positive (negative) entries indicate a decrease (increase) in MSE relative to base.
+The negative improvements beyond 6% are capped."
+#| fig.height: 10
+#| fig.width: 10
+#| out.width: "100%"
+
+MSE_plots <- list()
+
+cap <- -6
+for (sim_name in names(sim_MSE)) {
+
+  plot_df <- sim_MSE[[sim_name]] |> 
+    # cap pct_change upper 10%
+    mutate(
+      capped = pct_change <= cap,
+      capped_pct_change = pmax(pct_change, cap)
+    ) |> 
+    mutate(shape_lab = ifelse(shape_lab %in% c("PC1","PC2"), shape_lab, ""))
+  p <- plot_df |>
+    ggplot(aes(
+      x = h, y = capped_pct_change,
+      color = family, shape = shape_lab, linetype = variant_type
+    )) +
+      geom_line(linewidth = 0.8) +
+      geom_point(alpha = 0.85, size = 2.6, na.rm = TRUE) +
+      recon_scales +
+      # # label actual values for capped points
+      # geom_text(
+      #   data = subset(plot_df, pct_change <= cap),
+      #   aes(label = round(pct_change, 1)),
+      #   vjust = -0.5, color = "black", size = 3
+      # ) +
+      theme_minimal() +
+      labs(x = "Forecast Horizon", y = "% improvements")
+  MSE_plots[[sim_name]] <- p
+}
+
+MSE_plots[[1]] + labs(title = "6 Groups of 6", subtitle = "T=50") +
+  MSE_plots[[2]] + labs(subtitle = "T=300") +
+  MSE_plots[[3]] + labs(title = "2 Groups of 50", subtitle = "T=50") +
+  MSE_plots[[4]] + labs(subtitle = "T=300") +
+  plot_layout(ncol = 2, guides = "collect", axis_titles = "collect") & 
+  theme(legend.position = "bottom", legend.direction = "vertical", legend.title.position = "left")
+```
 
@@
 
-![Percentage relative improvement in Energy score in both hierarchies and time dimensions, for 1-step-ahead forecasts.
-The positive entries indicate a decrease in Energy score relative to base.](D:/Github/Recon_Honours_Thesis/figs/sim_energy.png){#fig-sim-results-2}
+<!-- ![Percentage relative improvement in Energy score in both hierarchies and time dimensions, for 1-step-ahead forecasts.
+The positive entries indicate a decrease in Energy score relative to base.](D:/Github/Recon_Honours_Thesis/figs/sim_energy.png){#fig-sim-results-2} -->
+
+```{r}
+#| label: fig-sim-results-2
+#| fig-cap: "Percentage relative improvement in Energy score in both hierarchies and time dimensions, for 1-step-ahead forecasts.
+The positive entries indicate a decrease in Energy score relative to base."
+#| fig.height: 10
+#| fig.width: 10
+#| out.width: "100%"
+
+energy_plots <- list()
+
+for (sim_name in names(sim_prob)) {
+
+  p <- sim_prob[[sim_name]] |>
+    filter(rule == "Energy") |>
+    filter(model_label != "Base") |>
+    ggplot(aes(x = model_label, y = pct_change_avg, fill = family, alpha=pc_k)) +
+      geom_col(color = "black") +
+      scale_alpha_manual(values = c("0" = 1, "1" = 0.7, "2" = 0.4), guide = "none") +
+      recon_fill_scale +
+      labs(x = "Model", y = "% improvements") +
+      theme_minimal() +
+      theme(
+        axis.text.x = element_text(angle = 20, hjust = 0.5)
+      )
+
+  energy_plots[[sim_name]] <- p
+}
+
+energy_plots[[1]] + labs(title = "6 Groups of 6", subtitle = "T=50") +
+  energy_plots[[2]] + labs(subtitle = "T=300") +
+  energy_plots[[3]] + labs(title = "2 Groups of 50", subtitle = "T=50") +
+  energy_plots[[4]] + labs(subtitle = "T=300") +
+  plot_layout(ncol = 2, guides = "collect", axis_titles = "collect") & 
+  theme(legend.position = "bottom", legend.title = element_blank())
+
+```
 

### 2025-10-20 thesis/paper_draft.qmd
Refine Simulation section from george feedback

@@
 
+Table: Summary of covariance estimators for MinT reconciliation {#tbl-cov}
+
 | Covariance estimators used              | Abbreviation       | 
@@
 
-where $\boldsymbol{A}$ is a $n_b \times n_b$ block diagonal matrix of autoregressive coefficients $\boldsymbol{A} = diag(\boldsymbol{A}_1, \ldots, \boldsymbol{A}_m)$, with each $\boldsymbol{A}_i$ being a $n_{b,i} \times n_{b,i}$ matrix.
-The block diagonal structure ensures that the time series are grouped into $m$ groups, with each group having its own autoregressive coefficients.
-This aim to simulate the interdependencies between the time series within each group, where reconciliation will be expected to better performed than the usual base forecasts.
-
-The model is added with a Gaussian innovation process $\boldsymbol{\epsilon}_t$, with covariance matrix $\boldsymbol{\Sigma}$.
+where $\boldsymbol{A}$ is a $n_b \times n_b$ block diagonal matrix of autoregressive coefficients $\boldsymbol{A} = diag(\boldsymbol{A}_1, \ldots, \boldsymbol{A}_M)$, with each $\boldsymbol{A}_m$ being a $n_{b,m} \times n_{b,m}$ matrix.
+The block diagonal structure ensures that the time series are grouped into $M$ groups (blocks), with each group having its own autoregressive coefficients.
+This aims to simulate the interdependencies between the time series within each group, where reconciliation will be expected to better capture these compared to independent base forecasts.
+
+$\boldsymbol{\epsilon}_t$ is a Gaussian innovation process, with covariance matrix $\boldsymbol{\Sigma}$.
 The covariance matrix $\boldsymbol{\Sigma}$ is generated specifically using the Algorithm 1 in @Hardin2013-wu:
@@
 1.
-A compound symmetric correlation matrix is used for each block of size $n_{b,i}$ in $\boldsymbol{A}_i$, where the entries $\rho_i$ for each block $i$ are sampled from a uniform distribution between 0 and 1.
-They are baseline correlations within group.
-
-2.  A constant correlation, which is smaller than $\text{min} \{\rho_1, \rho_2, \dots, \rho_m \}$, is imposed on the entries between different blocks. It serves as baseline correlations between group.
-
-3.  The entry-wise random noise is added on top of the entire correlation matrix.
-
-4.  The covariance matrix $\boldsymbol{\Sigma}$ is then constructed by uniform sampling of standard deviations, in a range of $[\sqrt{2}, \sqrt{6}]$, for all $n_b$ series.
-
-We will randomly flip the signs of the covariance elements, which will create a more realistic structure in the innovation process.
-This can be done by pre- and post-multiplying $\boldsymbol{\Sigma}$ by a random diagonal matrix $\boldsymbol{V}$ with diagonal entries sampled from $\{-1, 1\}$, yielding $\boldsymbol{\Sigma}^* = \boldsymbol{V} \boldsymbol{\Sigma} \boldsymbol{V}$.
-
-For all hierarchies in our experiments, we simulate two panel lengths, $T=54$ and $T=304$, reserving the final four observations as an out-of-sample test set.
-In each Monte Carlo replication ($M=500$), we fit univariate ARIMA models (base models) to the training observations using an automatic AICc minimization algorithm from @Hyndman2008-rf, implemented in the *fabletools* package [@O-Hara-Wild2024-we], generating incoherent 1–4-step base forecasts.
-We then reconcile these forecasts under different methods, and evaluate their point and probabilistic forecast accuracy on the test set.
+A compound symmetric correlation matrix is used for each block of size $n_{b,m}$ in $\boldsymbol{A}_m$, where the correlation entries $\rho_j$ for each block $m$ are sampled from a uniform distribution between 0 and 1.
+They are within group baseline correlations.
+
+2.  A constant correlation, which is smaller than $\text{min} \{\rho_1, \rho_2, \dots, \rho_M \}$, is imposed on the entries between different blocks. It serves as between group baseline correlations.
+
+3.
+We select a constant $\varepsilon$ such that $0 \le \varepsilon < 1 - \text{max} \{\rho_1, \rho_2, \dots, \rho_M \}$ to control the noise intensity.
+We then generate $n_b$ unit vectors $\boldsymbol{u}_1, \dots, \boldsymbol{u}_{n_b}$ and obtain a entry-wise noise element as $\varepsilon \boldsymbol{u}_i' \boldsymbol{u}_j$ for the $i,j$-th entry.
+These entry-wise noise elements are added on top of the baseline correlation matrix (except the diagonal entries), creating a noisy positive correlation matrix $\boldsymbol{R}^+$.
+
+4.
+The covariance matrix $\boldsymbol{\Sigma}^+$ is then constructed by $\boldsymbol{\Sigma}^+ = \boldsymbol{D} \boldsymbol{R}^+ \boldsymbol{D}$, where $\boldsymbol{D} = diag(\sigma_1, \dots, \sigma_{n_b})$ is a diagonal matrix of standard deviations.
+Each $\sigma_i$ is uniformly sampled from the range of $[\sqrt{2}, \sqrt{6}]$, for all $n_b$ series.
+
+These steps result in a positive covariance matrix $\boldsymbol{\Sigma}^+$ where all the elements are positive.
+To allow a mixture of positive and negative covariance, we randomly flip the signs of them but need to maintain the positive definiteness.
+This can be done by pre- and post-multiplying $\boldsymbol{\Sigma}$ by a random diagonal matrix $\boldsymbol{V}$ with diagonal entries sampled from $\{-1, 1\}$ with equal probability.
+The final covariance matrix is given by $\boldsymbol{\Sigma} = \boldsymbol{V} \boldsymbol{\Sigma}^+ \boldsymbol{V}$.
+
+For all hierarchies in our experiments, we simulate two panel lengths, $T=54$ ("short") and $T=304$ ("long"), reserving the final four observations as an out-of-sample test set.
+We perform 500 Monte Carlo replications for each configuration.
+In each replication, we fit univariate ARIMA models to the training observations using the automatic AICc minimization algorithm of @Hyndman2008-rf, implemented in the *fabletools* package [@O-Hara-Wild2024-we], generating 1–4-step base forecasts.
+We then reconcile the base forecasts using different reconciliation methods, and evaluate their point and probabilistic forecast accuracy on the test set.
 
@@
 In our main experiments, we examine how MinT combined with the different estimators perform as the hierarchy expands.
-We generate synthetic data from the same VAR(1) framework described earlier, but vary the number of bottom‐level series, $n_b$, across two structures: a small structure with six groups of six bottom series ($n_b = 6$x$6 = 36$), and a much larger configuration with two groups of fifty ($n_b = 2$x$50 = 100$).
-
-In the 36‐series case, each block of six forms a level‐1 aggregate, and those six aggregates form the national total.
+We generate synthetic data from the VAR(1) framework described earlier, varying the number of bottom‐level series, $n_b$, across two structures: a "small" structure with six groups of six bottom-level series ($n_b = 6$x$6 = 36$), and a "large" configuration with two groups of fifty bottom-level series ($n_b = 2$x$50 = 100$).
+
+In the 36‐series case, each block of six forms a level‐1 aggregate, and those six aggregates form the total.
 The 100‐series design employs a deliberately intricate aggregation path to stress‐test reconciliation methods.
@@
 
-Figure @fig-sim-results-1 illustrates the relative improvements in mean squared error (MSE) of reconciled forecasts over the incoherent base forecasts, across two structures and time series lengths.
-The MinT with shrinkage (*MinT-S*) and its variants are colored in mint green, while MinT with NOVELIST (*MinT-N*) are in purple.
-The concrete lines represent the vanilla *MinT-S* and *MinT-N*; the dashed lines with dot points denote the PC-adjusted variants (e.g.
-*MinT-S(PC1)*, *MinT-N(PC2)*); and the dotted or dashed-dotted lines indicate the scaled variance and h-step-ahead residuals versions (e.g.
+Figure @fig-sim-results-1 illustrates the relative improvements in mean squared error (MSE) of reconciled forecasts over the incoherent base forecasts, across the two structures (small and large) and time series lengths (short and long).
+MinT with shrinkage (*MinT-S*) and its variants (as defined in @tbl-cov) are colored in mint green, while MinT with NOVELIST (*MinT-N*) and its variants (as defined in same table) are in purple.
+The solid lines represent vanilla *MinT-S* and *MinT-N*; the dashed lines with dot points denote the PC-adjusted variants (e.g.
+*MinT-S(PC1)*, *MinT-N(PC2)*); and the dotted or dashed-dotted lines indicate the scaled variance and h-step-ahead residual versions (e.g.
 *MinT-S(SV)*, *MinT-N(hcov)*).
 
-The first key observation is that methods with shrinkage slightly outperform those with NOVELIST across all scenarios, despite the differences being small.
+The first key observation is that methods with shrinkage slightly outperform those with NOVELIST across all scenarios, although the differences are small.
 Second, the PC-adjusted variants (using one and two principal components) do not yield improvements over the vanilla versions.
 This is expected since the synthetic data generating process does not simulate from strong latent factors.
-Lastly, the scaled variance and h-step-ahead residuals approaches do not enhance performance as the forecast horizon increases, suggesting that the proportionality assumption may not be severely violated in this VAR(1) setup.
+Lastly, the scaled variance and h-step-ahead residual approaches do not enhance performance as the forecast horizon increases, suggesting that the proportionality assumption may not be severely violated in this VAR(1) setup.
 
@@
 
-When looking into each MCMC replication, we find that there are instances where the NOVELIST estimator collapses to the shrinkage estimator due to a large optimal threshold $\hat{\delta}$ being selected in the cross-validation step, resulting in a diagonal shrinkage target.
+When looking into each Monte Carlo replication, we find that there are instances where the NOVELIST estimator collapses to the shrinkage estimator due to a large optimal threshold $\hat{\delta}$ being selected in the cross-validation step, resulting in a diagonal shrinkage target.
 
@@
 Moving on to probabilistic forecasts, we evaluate the performance of reconciliation methods using the energy score, a proper scoring rule for multivariate predictive distributions.
-Figure @fig-sim-results-2 presents the percentage relative improvement in energy score for 1-step-ahead forecasts.
+@fig-sim-results-2 presents the percentage relative improvement in energy score for 1-step-ahead forecasts.
 Across both hierarchies and time dimensions, the MinT methods consistently outperform the base forecasts.
-Meanwhile, the differences among *MinT-S* and *MinT-N* variants are insignificant, except for the 6 by 6 case with 50 observations, where shrinkage has a slight edge.
+Meanwhile, the differences among *MinT-S* and *MinT-N* variants are small, except for the 6 by 6 case with 50 observations, where shrinkage has a pronounced edge.
 The PC-adjusted variants again degrade performance as we add more principal components.
@@
 
-In attempts to differentiate the performance of NOVELIST from the shrinkage estimator, we also simulate from a sparse covariance matrix for a 2 groups of 50 scenario, as illustrated in @fig-Sigma-2x50.
+In attempts to differentiate the performance of NOVELIST from the shrinkage estimator, we also simulate from a sparse covariance matrix for the previous section's 2 groups of 50 bottom series setting, as illustrated in @fig-Sigma-2x50.
 The sparse covariance matrix is obtained by randomly choosing 40% of the bottom series and setting their correlations with all other series to zero, resulting in a grid-like sparse structure.
@@
 
-These findings motivates our turn to empirical data in the next section, where latent structural features, regime shifts, and noisy, intermittent series will reveal performance differences.
+These findings motivates our turn to empirical data in the next section, where latent structural features, regime shifts, and noisy, intermittent series may possibly reveal performance differences.
 

### 2025-10-21 thesis/paper_draft.qmd
Refine tourism section (and slight changes to others) from George feedback

@@
 @Athanasopoulos2024-as provide a comprehensive review of the literature on forecast reconciliation.
+
 Among the modern methods, the Min Trace (MinT) approach developed by @Wickramasuriya2019-xq is is widely used due to its strong theoretical properties for minimising total reconciled forecast error variance, computational efficiency, and robust empirical performance.
 MinT is later extended to probabilistic reconciliation by @Wickramasuriya2024-mb, showing that it also minimises the negative log score of the reconciled distribution under Gaussian assumptions.
+@Wickramasuriya2019-xq also argued that modelling spatial autocorrelations directly from the start would be challenging as in this case of a large collection of time series.
+Post-processing reconciliation has the advantage to implicitly model this spatial autocorrelation structure, especially true for MinT.
 MinT is implemented in popular R and Python software ecosystems [@O-Hara-Wild2024-we; @Nixtla].
@@
 
-## Exploring Effects of Hierarchy's Size
+## Exploring Effects of Hierarchy's Size {#sec-sim-hierarchy-size}
 
@@
 Figure @fig-sim-results-1 illustrates the relative improvements in mean squared error (MSE) of reconciled forecasts over the incoherent base forecasts, across the two structures (small and large) and time series lengths (short and long).
-MinT with shrinkage (*MinT-S*) and its variants (as defined in @tbl-cov) are colored in mint green, while MinT with NOVELIST (*MinT-N*) and its variants (as defined in same table) are in purple.
+We evaluate ten different MinT variants based on the covariance estimators discussed in @tbl-cov, with visual distinctions made through colors, line types, and point shapes.
+MinT with shrinkage (*MinT-S*) and its variants are colored in mint green, while MinT with NOVELIST (*MinT-N*) and its variants are in purple.
 The solid lines represent vanilla *MinT-S* and *MinT-N*; the dashed lines with dot points denote the PC-adjusted variants (e.g.
@@
 
-Domestic tourism flows in Australia exhibit a natural hierarchical and grouped structure, driven both by geography and by purpose of travel.
+<!-- briefly talk about the important of forecasting tourism when describing the data-->
+
+Forecasting domestic tourism flows is of great economic and policy significance, as policymakers and stakeholders rely on accurate forecasts to make informed decisions regarding resource allocation, infrastructure development, and marketing strategies.
+The domestic tourism flows in Australia exhibit a natural hierarchical and grouped structure, driven both by geography and by purpose of travel.
 At the top of this hierarchy lies the national total, which splits into the seven states and territories.
@@
 A complete illustration of this geographic hierarchy appears in Appendix @sec-tourism-supplement.
-Intersecting this geographic hierarchy is a second dimension--travel motive--partitioning tourism flows into four categories: holiday, business, visiting friends and relatives, and other.
-Altogether, this yields a grouped system of 560 series, from the most disaggregated regional-purpose cells up to the full national aggregate.
+Intersecting this geographic hierarchy is a second dimension: policymakers are also interested in travel motives.
+This partitions tourism flows into four categories: holiday, business, visiting friends and relatives, and other.
+Altogether, this yields a grouped structure of 560 series, from the most disaggregated regional-purpose cells up to the full national aggregate.
 @tbl-tourism-structure depicts this structure.
@@
 
-The data are monthly time series spanning from January 1998 to December 2016, resulting in 228 observations per series, producing a canonical "$n \ll p$" setting which is ideal for evaluating reconciliation approaches that rely on high-dimensional covariance estimation.
-The extreme dimensionality over sample size mirrors many contemporary business problems, for instance, Starbucks drink sales.
+The data are monthly time series spanning from January 1998 to December 2016, resulting in 228 observations per series.
+This gives a challenging high-dimensional forecasting problem with number of series (560) doubled the number of observations, which is ideal for evaluating reconciliation approaches that rely on high-dimensional covariance estimation.
+The extreme dimensionality over sample size mirrors many contemporary business problems, for instance, Starbucks Corporation sales.
 Tourism demand is also economically vital yet highly volatile, with geographical and purpose‑specific patterns create a realistic stress‑test for reconciliation algorithms.
@@
 
-@Wickramasuriya2019-xq also argued that modelling spatial autocorrelations directly from the start would be challenging as in this case of a large collection of time series.
-Post-processing reconciliation approaches have the advantage to implicitly model this spatial autocorrelation structure, especially true for MinT.
-
 ![Rolling-window cross-validation scheme for evaluating forecasting performance in Australia tourism data](D:/Github/Recon_Honours_Thesis/figs/tourism_cv.png){#fig-tourism-cv width="50%"}
@@
 The 1- to 12-step-ahead base forecasts are then generated by these ARIMA models, and then reconciled using multiple approaches.
-To estimate the NOVELIST and its variants, we would have an extra cross-validation prodecure within this training window, as described in @sec-novelist_cv.
-We then roll the training window forward by one month and refit all models, rebuild reconciliations, and produce another batch of 1- to 12-step-ahead forecasts, repeating until the training set reaches December 2015.
+To estimate the NOVELIST and its variants, we implement an extra cross-validation prodecure within this training window, as described in @sec-novelist_cv.
+We then roll the training window forward by one month and refit all models, re-estimate reconciliation variants, and produce another batch of 1- to 12-step-ahead forecasts, repeating until the training set reaches December 2015.
 In total, this results in 97 out-of-sample windows.
@@
 #| fig-cap: "Heatmap of relative improvement in the mean squared error (MSE) of different reconciled forecasts over the base forecasts for the Australian domestic tourism data, for 1 to 12 steps ahead forecasts.
-The values are scaled to the range of 0 to 100 for better visualisation, with darker colors indicating greater improvement and best performance is noted by a star."
+The values are scaled to the range of 0 to 100 for better visualisation, with darker colors indicating greater improvement and best performance is noted by a star.
+This re-scaling does not affect relative rankings among methods."
 
@@
 
-@fig-tourism-MSE-line show a main difference compared to the one of simulation results.
+<!--
+- describe the fig
+- say models in @tbl-cov
+- encoded by color, linetype, shape similar to simulation plot
+- quickly refresh the memory of what each means
+-->
+
+@fig-tourism-MSE-line reports percentage relative improvements in MSE of reconciled forecasts over the incoherent base ARIMA forecasts, across horizons from 1 to 12 months.
+We evaluate ten MinT variants (as defined in @tbl-cov), similar to the simulation's @fig-sim-results-1.
+The MinT with shrinkage (*MinT-S*) and its variants are colored in mint green, while MinT with NOVELIST (*MinT-N*) and its variants are in purple.
+Solid lines represent "vanilla" *MinT-S* and *MinT-N*; dashed lines with dot points denote the PC-adjusted variants (e.g.
+*MinT-S(PC1)*, *MinT-N(PC2)*); and dotted or dashed-dotted lines indicate the scaled variance and h-step-ahead residual versions (e.g.
+*MinT-S(SV)*, *MinT-N(hcov)*).
+
+Three patterns stand out.
+First, the vanilla *MinT-S* (solid green) slightly outperform *MinT-N* (solid purple) at most horizons, although the gap is small.
+Second, variants that alter the multi-step covariance, either via scaled variance or direct h-step residual covariances, underperform standard MinT, suggesting that extra estimation at horizon $h>1$ is not rewarded in this empirical analysis.
+Most importantly, incorporating a single dominant factor via PC-adjustment (dashed line with round points) delivers best performance among all methods.
+Both *MinT-S(PC1)* and *MinT-N(PC1)* consistently beat their unadjusted counterparts across horizons.
+However, adding more than one PC brings no additional benefit and can erode performance, likely due to injecting idiosyncratic noise from weaker components.
+
+<!-- @fig-tourism-MSE-line shows a main difference compared to the one of simulation results.
 Adjusting for a single dominant factor via PC decomposition (dashed line with dot points) tightens performance further: both *MinT-S(PC1)* and *MinT-N(PC1)* beat their unadjusted counterparts.
 Other than that, similar patterns are observed: among vanilla MinT variants, the shrinkage estimator (*MinT-S*) slightly outperforms NOVELIST (*MinT-N*) at most horizons; adding more than one PC brings no additional benefit, likely due to injecting estimation noise from weaker components; variants that modify the multi-step covariance, either via scaled-variance or direct h-step residual covariances, underperform standard MinT, suggesting that extra estimation at horizon $h>1$ is not rewarded in this empirical analysis.
-
-@fig-tourism-MSE-heat provides a complementary heatmap view, scaling each method's MSE improvement to a 0-100 range for better visual discrimination.
-Here, *MinT-N(PC1)* and *MinT-S(PC1)* emerge as the top performers across most horizons.
-The heatmap underscores the consistent gains from PC adjustment and highlights the diminishing returns from more complex covariance treatments.
+-->
+
+@fig-tourism-MSE-heat complements the line plot with a heatmap that standardises each method’s MSE improvement to a 0–100 scale to enhance visual discrimination across horizons.
+Darker shades indicate larger improvements.
+The heatmap highlights the pattern from the line plot: *MinT-N(PC1)* and *MinT-S(PC1)* exhibit the highest concentration of dark cells and are repeatedly marked as the best method across horizons (star markers).
+Overall, the results underscores the consistent gains from PC adjustment and highlights the diminishing returns from more complex covariance treatments (lighter cells for scaled variance and h-step residuals).
 
@@
 
-Turning to probabilistic forecasts (1-step-ahead forecasts), @fig-tourism-probscores shows that *MinT-N* (purple bars) consistently outperforms *MinT-S* (green bars) across univariate and multivariate scores.
+Turning to probabilistic forecasts (1-step-ahead forecasts), @fig-tourism-probscores shows that *MinT-N* variants (purple bars) consistently outperform *MinT-S* variants (green bars) across univariate and multivariate scores.
 The PC1-adjusted variants again yield improvements over their vanilla counterparts, with *MinT-N(PC1)* leading overall.
 In the multivariate evaluation, the Energy score places *OLS* close to the PC1-adjusted MinT methods.
-One surprising finding is that all MinT variants underperform the base forecasts when moving to the 95% Winkler score, while *OLS* performs best.
+One surprising finding is that all MinT variants produce inferior forecasts compared to the base forecasts when considering 95% Winkler score, while *OLS* performs best.
 
@@
 To dissect the 95% Winkler score results, @fig-tourism-coverage-95-level breaks down performance by hierarchical level, and examines empirical coverage of the 95% prediction intervals.
-The left panel shows that *MinT-S* underperforms the base at all levels, especially in higher aggregated levels.
-From our inspection, this is due to the overly shrunk variances from the shrinkage estimator, leading to narrow prediction intervals and thus high Winkler penalties when observations fall outside.
-The *MinT-N* method has relatively good coverage and improve the Winkler score at bottom levels, but still underperforms at higher levels.
+The left panel shows that *MinT-S* produces inferior forecasts compared to the base forecasts at all levels, especially in higher aggregated levels.
+From our inspection, this is due to the overly shrunk variances from the shrinkage estimator, leading to narrow prediction intervals.
+The right panel confirms this, as *MinT-S* has the lowest empirical coverage across all levels, indicating that its prediction intervals fail to capture the true observations adequately and resulting in poor Winkler scores.
+*MinT-N* has relatively good coverage and improved Winkler score at bottom levels, but still is inferior at higher levels.
 The PC-adjusted variants seem to strike a better balance, improving overall coverage and relative Winkler scores over the base.
@@
 The outermost polygon represents the best possible score (100) and the innermost polygon represents the worst possible score (0).
-Only the top 4 MinT approaches are shown (with the base model)."
+Only the top 4 MinT approaches are shown, together with the base forecasts."
 
@@
 
-The summary radar graph in @fig-tourism-radar consolidates these findings: among the selected MinT models by probabilistic criteria, *MinT-N(PC1)* (the yellow polygon) clearly leads across CRPS, Winkler score at 80% and 95% intervals, and Energy, extending the improvements seen in the single-metric panels.
+The summary radar graph in @fig-tourism-radar consolidates these findings.
+Among the selected MinT variants considering probabilistic criteria, *MinT-N(PC1)* (yellow polygon) clearly leads across CRPS, Winkler score at 80% and 95% intervals, and Energy score, extending the improvements seen in the single-metric panels.
 Except for the Winkler score at 95% interval, all MinT variants outperform the base forecasts.

### 2025-10-21 thesis/paper_draft.qmd
codes for by level plot in tourism

@@
 
-![Percentage relative improvement in Winkler score at 95% nominal coverage by level (left), and empirical coverage of 95% prediction intervals by level (right)](D:/Github/Recon_Honours_Thesis/figs/tourism_coverage_95.png){#fig-tourism-coverage-95-level}
+<!-- ![Percentage relative improvement in Winkler score at 95% nominal coverage by level (left), and empirical coverage of 95% prediction intervals by level (right)](D:/Github/Recon_Honours_Thesis/figs/tourism_coverage_95.png){#fig-tourism-coverage-95-level} -->
+
+```{r}
+#| label: fig-tourism-coverage-95-level
+#| fig-cap: "Percentage relative improvement in Winkler score at 95% nominal coverage by aggregation level (left), and empirical coverage of 95% prediction intervals by aggregation level (right)."
+#| fig.height: 12
+#| fig.width: 9
+#| out.width: "100%"
+
+tourism_coverage95 <- readRDS(paste0(path, "/thesis/lib/tourism_coverage95.rds"))
+tourism_W95_level <- readRDS(paste0(path, "/thesis/lib/tourism_W95_level.rds"))
+
+p_c95_lvl <- tourism_coverage95 |> 
+  group_by(level, .model) |> 
+  summarise(mean_coverage = mean(coverage)) |> 
+  encode_recon_models() |> 
+  ggplot(aes(x = model_label, y = mean_coverage, fill = family, alpha = pc_k)) +
+    geom_col(color = "black") +
+    scale_alpha_manual(values = c("0" = 1, "1" = 0.7, "2" = 0.4), guide = "none") +
+    labs(x = "", y = "Empirical Coverage",
+         title = "Empirical coverage of 95% prediction intervals by level") +
+    geom_hline(yintercept = 0.95, linetype = "dashed", color = "black") + # draw line for coverage
+    facet_wrap(~ level, ncol=2, scales = "free_y") +
+    recon_fill_scale +
+    theme_minimal() +
+    # rotate the x axis labels
+    theme(axis.text.x = element_text(angle = 20, hjust = 0.5))
+
+p_W95_lvl <- tourism_W95_level |> 
+  ggplot(aes(x = model_label, y = pct_change, fill = family, alpha = pc_k)) +
+    geom_col(color = "black") +
+    scale_alpha_manual(values = c("0" = 1, "1" = 0.7, "2" = 0.4), guide = "none") +
+    labs(x = "", y = "% improvements",
+         title = "Winkler score 95% by level") +
+    facet_wrap(~ level, ncol=2, scales = "free_y") +
+    recon_fill_scale  +
+    guides(fill = "none") +
+    theme_minimal() +
+    # rotate the x axis labels
+    theme(axis.text.x = element_text(angle = 20, hjust = 0.5))
+
+p_W95_lvl + p_c95_lvl +
+  plot_layout(ncol = 1, guides = "collect", axis_titles = "collect") &
+  theme(legend.position = "right", legend.direction = "vertical", legend.title.position = "top")
+```
 

### 2025-10-21 R/plot_helpers.R

@@ -126,7 +126,7 @@ recon_scales <- list(
 # recon_fill_scale <- scale_fill_manual(
 #   name = "Family",
 #   values = c(
-#     "Base"   = "#9490b7ff",  # grey
+#     "Base"   = "#7A7A7A",  # grey
 #     "OLS"    = "#d1d400ff",  # black
 #     "MinT-S" = "#00c086ff",  # teal (colorblind-friendly)
 #     "MinT-N" = "#9500c2ff"   # purple
@@ -135,8 +135,8 @@ recon_scales <- list(
 
 recon_fill_scale <- scale_fill_manual(
   name   = "Family",
-  limits = c("OLS","MinT-S","MinT-N"),
-  breaks = c("OLS","MinT-S","MinT-N"),
-  values = c("OLS"="#d1d400", "MinT-S"="#00c086", "MinT-N"="#9500c2"),
+  limits = c("Base", "OLS","MinT-S","MinT-N"),
+  breaks = c("Base", "OLS","MinT-S","MinT-N"),
+  values = c("Base" = "#7A7A7A", "OLS"="#d1d400", "MinT-S"="#00c086", "MinT-N"="#9500c2"),
   drop   = FALSE
 )

### 2025-10-21 thesis/paper_draft.qmd
organise old functions and re render

@@
 
-sim1_1 <- readRDS(paste0(path_lib, "S4-2-1_T50_M500.rds"))
-sim1_2 <- readRDS(paste0(path_lib, "S4-2-1_T300_M500.rds"))
-
 sim2_1 <- readRDS(paste0(path_lib, "S36-6-1_T50_M500.rds"))
@@
 
-sim_dense_50 <- readRDS(paste0(path_lib, "S100-10-3-1_T50_M200_dense.rds"))
-sim_dense_300 <- readRDS(paste0(path_lib, "S100-10-3-1_T300_M200_dense.rds"))
-
-sim_sparse_50 <- readRDS(paste0(path_lib, "S100-10-3-1_T50_M500_sparsebwgrp.rds"))
-sim_sparse_300 <- readRDS(paste0(path_lib, "S100-10-3-1_T300_M200_sparsebwgrp.rds"))
-
-sim3_1 <- sim_dense_50 # using same with next section simulation
-sim3_2 <- sim_dense_300
+sim3_1 <- readRDS(paste0(path_lib, "S100-10-3-1_T50_M200_dense.rds"))
+sim3_2 <- readRDS(paste0(path_lib, "S100-10-3-1_T300_M200_dense.rds"))
 
@@
 The block diagonal structure of the VAR(1) coefficient matrices $\boldsymbol{A}$ reflects the grouping of series, and the correlation matrices- show higher correlations among series within the same group.
-
-```{r generate plots for A and Sigma}
-MSE1_1 <- transform_sim_MSE(sim1_1$MSE, F) |> 
-  filter(.model != "ols")
-MSE1_2 <- transform_sim_MSE(sim1_2$MSE, F) |> 
-  filter(.model != "ols")
-
-MSE2_1 <- transform_sim_MSE(sim2_1$MSE, F) |> 
-  filter(! .model %in% c("mint_n_hstep", "ols"))
-MSE2_2 <- transform_sim_MSE(sim2_2$MSE, F) |> 
-  filter(! .model %in% c("mint_n_hstep", "ols"))
-
-MSE3_1 <- transform_sim_MSE(sim3_1$MSE, F)
-MSE3_2 <- transform_sim_MSE(sim3_2$MSE, F) |> 
-  filter(.model != "ols")
-
-p1_A <- plot_heatmap(sim1_1$A, T) + 
-  ggtitle("2x2 VAR(1) Coefficients Matrix")
-
-p1_Sigma <- plot_heatmap(cov2cor(sim1_1$Sigma), T) + 
-  ggtitle("2x2 Correlation Matrix of Innovation process")
-
-p2_A <- plot_heatmap(sim2_1$A) +
-  ggtitle("VAR(1) Coefficients Matrix") + theme_void()
-
-p2_Sigma <- plot_heatmap(cov2cor(sim2_1$Sigma), T) +
-  ggtitle("Correlation Matrix of Innovation process") + theme_void()
-
-p3_A <- plot_heatmap(sim3_1$A) +
-  ggtitle("VAR(1) Coefficients Matrix")
-
-p3_Sigma <- plot_heatmap(cov2cor(sim3_1$Sigma), T) +
-  ggtitle("Correlation Matrix of Innovation process")
-```
 
@@
 
+p2_A <- plot_heatmap(sim2_1$A) +
+  ggtitle("VAR(1) Coefficients Matrix") + theme_void()
+
+p2_Sigma <- plot_heatmap(cov2cor(sim2_1$Sigma), T) +
+  ggtitle("Correlation Matrix of Innovation process") + theme_void()
+
 p2_A + p2_Sigma + plot_layout(ncol = 2) & theme(legend.position = "bottom")
@@
 #| fig.align: "center"
+
+p3_A <- plot_heatmap(sim3_1$A) +
+  ggtitle("VAR(1) Coefficients Matrix")
+
+p3_Sigma <- plot_heatmap(cov2cor(sim3_1$Sigma), T) +
+  ggtitle("Correlation Matrix of Innovation process")
 
@@
 
-sim_sparse_50 <- readRDS(paste0(path, "/thesis/lib/params_S100_10_3_1_sparse46.rds"))
-sim_sparse_50$Sigma |>
+sim_params_sparse <- readRDS(paste0(path, "/thesis/lib/params_S100_10_3_1_sparse46.rds"))
+sim_params_sparse$Sigma |>
   cov2cor() |> 

### 2025-10-21 thesis/paper_draft.qmd
fixing the font size and spacing, minor fix to scale_fill_manual for bar plots

@@
   pdf:
+    fontsize: 11pt
+    linestretch: 1.15
     include-in-header:
@@
 
+\vspace{2cm}
+
 # Abstract {.unnumbered}
@@
 We evaluate MinT using these covariance estimates for both point and probabilistic reconciliation, and demonstrate their effectiveness and improvements over the shrinkage estimator in a complex, large-hierarchy dataset.
+
+\pagebreak
 
@@
     facet_wrap(~ level, ncol=2, scales = "free_y") +
-    recon_fill_scale +
+    recon_fill_scale_w_base +
     theme_minimal() +
@@
 
-This paper aims to address the limitations of the current Minimum Trace reconciliation method, a method that has garnered significant attention in both academic and practical forecasting contexts.
-We propose the NOVELIST estimator to address the lack of flexibility in the uniform shrinkage target of the traditional shrinkage estimator.
-Additionally, we introduce PC-adjusted variants of shrinkage and NOVELIST to take into account dominant latent factors, as well as multi-step covariance variants that relax the proportionality assumption.
-
-Empirical results on Australian domestic tourism data reveal several insights.
-For point forecasts, shrinkage (and its PC-adjusted version) remains a robust default choice, while NOVELIST demonstrates improved performance for probabilistic forecasts.
-Adjusting for a single dominant principal component consistently enhances performance across both point and probabilistic forecasts, suggesting its utility when such latent structures are present.
-More complex adjustments, such as incorporating multiple principal components or horizon-specific covariances, tend to introduce additional estimation noise without notable benefits.
-
-However, several limitations and avenues for future research remain.
-More intricate simulation studies are necessary to further highlight the differences between shrinkage and NOVELIST, particularly in scenarios involving factor models.
-The current implementation does not standardise data for PC-variant methods, which may lead to higher-level series largely influencing the principal components.
-Future work could also explore alternative methods that leverage cross-series information or eigenvalue/eigenvector structures.
-Additionally, understanding why MinT with shrinkage underperforms in the Winkler 95% score, especially at higher aggregation levels, needs further investigation.
-Finally, evaluating h-step-ahead probabilistic forecasts remains an open area for further research.
+This paper tackles a central bottleneck in Minimum Trace (MinT) reconciliation: obtaining reliable, positive-definite estimates of the base-forecast error covariance.
+The original practice, following @Wickramasuriya2019-xq, estimates the one-step covariance via diagonal-target shrinkage and obtains multi-step horizons by proportional scaling.
+We identified three limitations of this approach--uniform shrinkage, lack of factor awareness, and horizon-invariant dependence--and proposed alternatives that address them individually and in combination.
+Specifically, we adopt NOVELIST to introduce a more flexible shrinkage target; develop PC-adjusted variants of shrinkage and NOVELIST to preserve dominant latent components in the data; and consider multi-step covariance constructions that relax proportional scaling assumption.
+
+Empirical evidence from Australian domestic tourism yields several insights.
+For point reconciliation, shrinkage (and its PC-adjusted version) remains a robust default choice, while NOVELIST demonstrates improved performance for probabilistic forecasts.
+Across both settings, adjusting for a single dominant principal component consistently enhances performance, indicating its utility when such latent structures are present.
+More complex adjustments, such as incorporating multiple principal components or horizon-specific covariances, tend to introduce additional idiosyncratic noise without notable benefits.
+
+Several limitations and directions for future work remain.
+First, more intricate simulation designs are needed to further highlight the differences between shrinkage and NOVELIST, particularly under explicit factor models and varying sparsity/noise levels.
+Second, our current PC-adjusted implementations do not standardise series prior to factor extraction, which may lead to higher-level series largely influencing the principal components.
+Third, we can explore alternative methods that leverage cross-series information or eigenvalue/eigenvector structures.
+Fourth, the observed inferior performance of MinT with shrinkage under the Winkler 95% score at higher aggregation levels needs detailed investigation.
+Finally, evaluating genuinely $h$-step probabilistic reconciliation remains an open avenue, including how to best regularise multi-step covariances given limited effective samples.
 

### 2025-10-21 R/plot_helpers.R

@@ -135,6 +135,14 @@ recon_scales <- list(
 
 recon_fill_scale <- scale_fill_manual(
   name   = "Family",
+  limits = c("OLS","MinT-S","MinT-N"),
+  breaks = c("OLS","MinT-S","MinT-N"),
+  values = c("OLS"="#d1d400", "MinT-S"="#00c086", "MinT-N"="#9500c2"),
+  drop   = FALSE
+)
+
+recon_fill_scale_w_base <- scale_fill_manual(
+  name   = "Family",
   limits = c("Base", "OLS","MinT-S","MinT-N"),
   breaks = c("Base", "OLS","MinT-S","MinT-N"),
   values = c("Base" = "#7A7A7A", "OLS"="#d1d400", "MinT-S"="#00c086", "MinT-N"="#9500c2"),

### 2025-10-21 thesis/paper_draft.qmd
minor fixes to algorithm

@@
 \hat{\boldsymbol{W}}^{N}_{1} = \lambda_{\delta} \hat{\boldsymbol{W}}_{1, \delta} + (1 - \lambda_{\delta}) \hat{\boldsymbol{W}}_1,
-$$ {#eq-novelist-cov}
+$$
 
@@
     \For{$\delta \in \Delta$}
-        \State Compute thresholded correlation $\hat{\boldsymbol{R}}_{j,\delta}$ using Equation 5
-        \State Compute $\hat{\lambda}_{j,\delta}$ using Equation 6
-        \State Compute $\hat{\boldsymbol{R}}^{N}_{j,\delta}$ using Equation 4
+        \State Compute thresholded correlation $\hat{\boldsymbol{R}}_{j,\delta}$ using Equation 2
+        \State Compute $\hat{\lambda}_{j,\delta}$ using Equation 3
+        \State Compute $\hat{\boldsymbol{R}}^{N}_{j,\delta}$ using Equation 1
         \State $\hat{\boldsymbol{W}}^{N}_{j,\delta} = \hat{\boldsymbol{D}}_j^{1/2} \hat{\boldsymbol{R}}^{N}_{j,\delta} \hat{\boldsymbol{D}}_j^{1/2}$
@@
 \State $\text{MSE}_{\delta} = \frac{1}{T-v} \sum_{i=v}^{T-1} (\tilde{\boldsymbol{e}}_{i+1 | \delta})^2$ for each $\delta \in \Delta$
-\State $\hat{\delta}^* = \arg\min_{\delta \in \Delta} \text{MSE}_{\delta}$
-\State Compute $\hat{\lambda}^*$ on all training data using $\hat{\delta}^*$
-\State Compute $\hat{\boldsymbol{R}}_1^*$ using $\hat{\delta}^*$ and $\hat{\lambda}^*$ on all training data, using Equation 3
-
-\State \textbf{Output:} Estimate of optimal $\hat{\delta}^*$
+\State $\hat{\delta} = \arg\min_{\delta \in \Delta} \text{MSE}_{\delta}$
+
+\State \textbf{Output:} Estimate of optimal threshold parameter $\hat{\delta}$
 
@@
 
-*Remark.* Minimising a multi-step-ahead forecast error metric in the cross-validation procedure oftern yields a $\hat{\delta}^*$ that is close to the 1-step-ahead case.
+*Remark.* Minimising a multi-step-ahead forecast error metric in the cross-validation procedure oftern yields a $\hat{\delta}$ that is close to the 1-step-ahead case.
 

### 2025-10-22 thesis/paper_draft.qmd
minor change, change params files for simulations

@@
 
-\vspace{2cm}
+\vspace{1cm}
 
@@
 We evaluate MinT using these covariance estimates for both point and probabilistic reconciliation, and demonstrate their effectiveness and improvements over the shrinkage estimator in a complex, large-hierarchy dataset.
+
+\vspace{1cm}
+
+***Keywords:** Coherent forecasts, Hierarchical time series, Covariance estimation, Australian tourism, Aggregation.*
 
@@
 
-sim2_1 <- readRDS(paste0(path_lib, "S36-6-1_T50_M500.rds"))
-sim2_2 <- readRDS(paste0(path_lib, "S36-6-1_T300_M500.rds"))
-
-sim3_1 <- readRDS(paste0(path_lib, "S100-10-3-1_T50_M200_dense.rds"))
-sim3_2 <- readRDS(paste0(path_lib, "S100-10-3-1_T300_M200_dense.rds"))
+sim_params_S36 <- readRDS(paste0(path_lib, "params_S36_6_1.rds"))
+sim_params_S100 <- readRDS(paste0(path_lib, "params_S100_10_3_1_dense.rds"))
+sim_params_sparse <- readRDS(paste0(path_lib, "params_S100_10_3_1_sparse46.rds"))
 
@@
 
-p2_A <- plot_heatmap(sim2_1$A) +
+p36_A <- plot_heatmap(sim_params_S36$A) +
   ggtitle("VAR(1) Coefficients Matrix") + theme_void()
 
-p2_Sigma <- plot_heatmap(cov2cor(sim2_1$Sigma), T) +
+p36_Sigma <- plot_heatmap(cov2cor(sim_params_S36$Sigma), T) +
   ggtitle("Correlation Matrix of Innovation process") + theme_void()
 
-p2_A + p2_Sigma + plot_layout(ncol = 2) & theme(legend.position = "bottom")
+p36_A + p36_Sigma + plot_layout(ncol = 2) & theme(legend.position = "bottom")
 ```
@@
 
-p3_A <- plot_heatmap(sim3_1$A) +
+p100_A <- plot_heatmap(sim_params_S100$A) +
   ggtitle("VAR(1) Coefficients Matrix")
 
-p3_Sigma <- plot_heatmap(cov2cor(sim3_1$Sigma), T) +
+p100_Sigma <- plot_heatmap(cov2cor(sim_params_S100$Sigma), T) +
   ggtitle("Correlation Matrix of Innovation process")
 
-p3_A + p3_Sigma + plot_layout(ncol = 1) & theme(legend.position = "right")
+p100_A + p100_Sigma + plot_layout(ncol = 1) & theme(legend.position = "right")
 ```
@@
 
-sim_params_sparse <- readRDS(paste0(path, "/thesis/lib/params_S100_10_3_1_sparse46.rds"))
 sim_params_sparse$Sigma |>

### 2025-10-22 thesis/paper_draft.qmd
update title page (cover page)

@@
 ---
-title: "Enhancing Forecast Reconciliation: A Study of Alternative Covariance Estimators"
-author: 
-  - Vincent Su
-  - Shanika Wickramasuriya (supv.)
-  - George Athanasopoulos (supv.)
+# title: "\\LARGE Enhancing Forecast Reconciliation: A Study of Alternative Covariance Estimators"
+# author: 
+#   - \large Vincent Su
+#   - \large Shanika Wickramasuriya (supv.)
+#   - \large George Athanasopoulos (supv.)
 format:
@@
     linestretch: 1.15
+    classoption: [titlepage]
     include-in-header:
@@
         \usepackage{xcolor}
+
+        % Heading font sizes
+        \usepackage{sectsty}
+        \sectionfont{\fontsize{16}{30}\selectfont}
+        \subsectionfont{\fontsize{14}{16}\selectfont}
+        \subsubsectionfont{\fontsize{12}{14}\selectfont}
+
+        \usepackage{graphicx}
     mathspec: true
@@
 
+<!-- Write cover page my own -->
+
+::: {=latex}
+\begin{titlepage}
+
+\begin{center}
+\includegraphics[height=2cm,keepaspectratio]{D:/Github/Recon_Honours_Thesis/assets/monash-logo.png}
+\end{center}
+
+\centering
+
+{%
+\sffamily
+
+% --- move the whole block down a bit ---
+\vspace*{1.5cm}
+
+{\huge \bfseries Enhancing Forecast Reconciliation:\\
+A Study of\\
+Alternative Covariance Estimators\par}
+
+\vspace{2cm}
+
+{\Large \bfseries Vincent Su\par}
+{\large B.Com. (Hons), Monash University\par}
+
+\vspace{1cm}
+
+{\large Supervised by\par}
+{\Large Shanika Wickramasuriya\\
+and George Athanasopoulos\par}
+
+\vfill
+
+{\large A thesis submitted for the degree of \\
+Bachelor of Commerce (Honours)\\
+at Monash University\\
+Department of Econometrics and Business Statistics\par}
+
+\vspace{0.3cm}
+
+{\large 25 October 2025\par}
+
+}%
+\end{titlepage}
+:::
+
 ```{r library}
@@
 
-\vspace{1cm}
+\pagebreak
 
@@
 \pagebreak
+
+\clearpage
+\tableofcontents
+\clearpage
 

### 2025-10-22 thesis/paper_draft.qmd
reorganise paths (move lib folder outside), rerender

@@
 
-<!-- Write cover page my own -->
+<!-- COVER PAGE -->
 
@@
 \begin{center}
-\includegraphics[height=2cm,keepaspectratio]{D:/Github/Recon_Honours_Thesis/assets/monash-logo.png}
+\includegraphics[height=2cm,keepaspectratio]{D:/Github/Recon_Honours_Thesis/figs/monash-logo.png}
 \end{center}
@@
 ```{r library}
+library(dplyr)
+library(tidyr)
+library(tibble)
 library(ggplot2)
 library(patchwork)
-library(dplyr)
-library(tidyr)
 library(fable)
@@
 library(forcats)
-library(tibble)
 library(fmsb)
@@
 path <- here::here()
+path_lib <- paste0(path, "/lib/") # lib folder (sim, empi results)
 knitr::opts_knit$set(root.dir = path)
-# setwd(path)
 ```
@@
 # Abstract {.unnumbered}
-
-<!-- [This is pasted from the Project Description]{style="color:red;"} -->
 
@@
 Among modern reconciliation methods, Minimum Trace (MinT) is widely used, however, it requires a good estimate of the forecast error covariance matrix.
-The current practice is to use the linear shrinkage towards a diagonal target.
-Furthermore, the covariance estimate is based on 1-step-ahead residuals, then proportionally scale it to approximate the h-step-ahead covariance matrix.
+The current practice uses linear shrinkage towards a diagonal target.
+Furthermore, the covariance estimate is based on 1-step-ahead residuals, then proportionally scale it to approximate h-step-ahead covariance matrix.
 This leaves a question of whether this method is appropriate for all real-world applications.
@@
 ```{r load results}
-path_lib <- paste0(path, "/thesis/lib/")
-
 sim_params_S36 <- readRDS(paste0(path_lib, "params_S36_6_1.rds"))
@@
 
-```{r}
+<!-- ```{r}
 tourism_plots <- readRDS("D:/Github/Recon_Honours_Thesis/thesis/tourism_plots.rds")
-```
+``` -->
 
@@
 
-tourism_MSE_results <- readRDS(paste0(path, "/thesis/lib/tourism_MSE_results.rds"))
+tourism_MSE_results <- readRDS(paste0(path_lib, "tourism_MSE_results.rds"))
 
@@
 
-tourism_uniscores <- readRDS(paste0(path, "./thesis/lib/tourism_uniscore_results.rds"))
-tourism_energy <- readRDS(paste0(path, "./thesis/lib/tourism_energyscore_results.rds"))
+tourism_uniscores <- readRDS(paste0(path_lib, "tourism_uniscore_results.rds"))
+tourism_energy <- readRDS(paste0(path_lib, "tourism_energyscore_results.rds"))
 
@@
 
-tourism_coverage95 <- readRDS(paste0(path, "/thesis/lib/tourism_coverage95.rds"))
-tourism_W95_level <- readRDS(paste0(path, "/thesis/lib/tourism_W95_level.rds"))
+tourism_coverage95 <- readRDS(paste0(path_lib, "tourism_coverage95.rds"))
+tourism_W95_level <- readRDS(paste0(path_lib, "tourism_W95_level.rds"))
 

### 2025-10-22 R/table_geo.R

@@ -3,6 +3,7 @@ library(tidyr)
 library(stringr)
 
 path <- here::here()
+path_lib <- paste0(path,"/lib/")
 
 # 0) Prepping --------------------------------
 # State 
@@ -97,7 +98,7 @@ region_zone_lookup <- tibble(
   Region = unlist(zone_map)
 )
 
-level_map <- readRDS(paste0(path,"/thesis/lib/level_info.rds"))$map
+level_map <- readRDS(paste0(path_lib, "level_info.rds"))$map
 # only geo
 table_geo <- level_map[str_sub(level_map, -7, -1) != "Purpose"]
 table_geo <- tibble(

### 2025-10-22 thesis/paper_draft.qmd
refine hts/recon and mint section from Shanika feedback

@@
 
-```{r library}
+```{r}
+#| label: setup
 library(dplyr)
@@
 Forecast reconciliation, a post-processing step, utilises the information from the hierarchical structure and data to adjust the initially produced base forecasts, so that the resulting *reconciled forecasts* are *coherent* (i.e., respecting the aggregation constraints).
-It was first introduced by @Hyndman2011-jv, and later developed by @Van_Erven2015-nx, @Hyndman2016-ic, @Ben-Taieb2019-yx, @Wickramasuriya2019-xq, @Wickramasuriya2020-uk, and others that focus point forecast reconciliation.
+It was first introduced by @Hyndman2011-jv, and later developed by @Van_Erven2015-nx, @Hyndman2016-ic, @Ben-Taieb2019-yx, @Wickramasuriya2019-xq, @Wickramasuriya2020-uk, and others that focus point forecast reconciliation (more accurately, reconciling the means of predictive distributions).
 Recognising the important of probabilistic forecasting, probabilistic forecast reconciliation was studied by @Shang2017-ux, @Jeon2019-wm, and @Ben-Taieb2021-bn, and later formally defined by @Panagiotelis2023-fs.
@@
 
-## Hierarchical Time Series {#sec-hierarchy}
-
-*Hierarchical time series* are multivariate time series $\boldsymbol{y}_t \in \mathbb{R}^n$ organised in a structure where the series adheres to some constraints.
+## Hierarchical Time Series and Reconciliation {#sec-hierarchy}
+
+*Hierarchical time series* are multivariate time series $\boldsymbol{y}_t \in \mathbb{R}^n$ organised in a structure where the series adheres to linear constraints.
 @fig-hierarchy illustrates a simple two-level hierarchical structure with one top-level series $y_{Tot,t}$, disaggregating down to two level-1 series $(y_{A,t}, y_{B,t})'$, and to four bottom-level series $(y_{A1,t}, y_{A2,t}, y_{B1,t}, y_{B2,t})'$.
-Here, the *aggregation constraints* imply that $y_{Tot,t} = y_{A,t} + y_{B,t}$, $y_{A,t} = y_{A1,t} + y_{A2,t}$, and $y_{B,t} = y_{B1,t} + y_{B2,t}$.
+Here, the *aggregation constraints* imply that $y_{Tot,t} = y_{A,t} + y_{B,t}$, $y_{A,t} = y_{A1,t} + y_{A2,t}$, and $y_{B,t} = y_{B1,t} + y_{B2,t}$, for all time points $t$.
 
@@
 \end{array}
-\right].
-$$
-
+\right],
+$$
+
+where $\boldsymbol{I_4}$ is the $4 \times 4$ identity matrix.
 The matrix $\boldsymbol{S}$ encodes the aggregation constraints implied by the structure.
@@
 
-![A 2-level grouped structure, which can be considered as the union of two hierarchical trees with common top and bottom level series](../figs/grouped_structure.png){#fig-grouped width="90%"}
-
 This representation extends beyond hierarchical (nested) structures.
@@
 
+![A 2-level grouped structure, which can be considered as the union of two hierarchical trees with common top and bottom level series](../figs/grouped_structure.png){#fig-grouped width="90%"}
+
 <!-- There are other forms of structures, such as temporal hierarchies or non-linear constraints, but they are beyond the scope of this paper.  -->
 
-In practice, when we produce forecasts for each individual series, referred to as *base forecasts* $\hat{\boldsymbol{y}}_{t+h|t}$, they generally violate the aggregation constraints, and thus are incoherent.
-Coherency can be restored by linearly projecting the base forecasts onto the coherent subspace $\mathfrak{s}$ using a projection matrix $\boldsymbol{P}$: $\tilde{\boldsymbol{y}}_{t+h|t} = \boldsymbol{P} \hat{\boldsymbol{y}}_{t+h|t}$, where $\tilde{\boldsymbol{y}}_{t+h|t}$ are the *reconciled forecasts*.
+In practice, when we produce $h$-step-ahead forecasts for each individual series, referred to as *base forecasts* $\hat{\boldsymbol{y}}_{t+h|t}$, they generally violate the aggregation constraints, and thus are incoherent.
+Coherency can be restored by linearly projecting the base forecasts onto the coherent subspace $\mathfrak{s}$ using a projection matrix $\boldsymbol{P}$: $\tilde{\boldsymbol{y}}_{t+h|t} = \boldsymbol{P} \hat{\boldsymbol{y}}_{t+h|t}$, where $\tilde{\boldsymbol{y}}_{t+h|t}$ are $h$-step-ahead *reconciled forecasts*.
+
+Many existing reconciliation methods including the OLS from @Hyndman2011-jv, WLS from @Hyndman2016-ic, and MinT from @Wickramasuriya2019-xq express the projection matrix as $\boldsymbol{P} = \boldsymbol{S} \boldsymbol{G}$, for a suitable $n_b \times n$ mapping matrix $\boldsymbol{G}$.
+The role of this matrix is to map the base forecasts of all levels $\hat{\boldsymbol{y}}_{t+h|t}$ down into the bottom level, which is then aggregated to the higher levels by premultiplying with $\boldsymbol{S}$.
+Since the projection matrix $\boldsymbol{P}$ is idempotent, $\boldsymbol{G}$ must satisfy the condition $\boldsymbol{S} \boldsymbol{G} \boldsymbol{S} = \boldsymbol{S}$.
+Within this class, a broad family of mapping matrices is given by: $\boldsymbol{G} = (\boldsymbol{S}' \boldsymbol{M}^{-1} \boldsymbol{S})^{-1} \boldsymbol{S}' \boldsymbol{M}^{-1}$, for some positive definite matrix $\boldsymbol{M} \in \mathbb{R}^{n \times n}$ [@Gamakumara2020-ml].
+
+When setting $\boldsymbol{M} = \boldsymbol{I}_n$, the identity matrix, the $\boldsymbol{P}$ matrix reduces to the OLS reconciliation, which corresponds to an orthogonal projection onto $\mathfrak{s}$.
+For a general positive definite matrix $\boldsymbol{M}$, $\boldsymbol{P}$ is idempotent but not necessarily symmetric, resulting in an oblique projection.
+Thus, the choice of $\boldsymbol{M}$ determines the projection direction and, in turn, how disagreements among levels are resolved.
 
@@
 The projection is defined by the projection matrix $\boldsymbol{P}$.
-Note that this figure is schematic since most applications are high-dimensional.](../figs/geometry_orthogonal.png){#fig-geometry width="60%"}
-
-Many existing reconciliation methods including the OLS [@Hyndman2011-jv], WLS [@Hyndman2016-ic], and MinT [@Wickramasuriya2019-xq] express the projection matrix as $\boldsymbol{P} = \boldsymbol{S} \boldsymbol{G}$, for a suitable $n_b \times n$ mapping matrix $\boldsymbol{G}$.
-The idea is to map the base forecasts of all levels $\hat{\boldsymbol{y}}_{t+h|t}$ down into the bottom level, which is then aggregated to the higher levels by $\boldsymbol{S}$.
-Since the projection matrix $\boldsymbol{P}$ is idempotent, $\boldsymbol{G}$ must satisfy the condition $\boldsymbol{S} \boldsymbol{G} \boldsymbol{S} = \boldsymbol{S}$.
-Within this class, a broad family of mapping matrix is given by: $\boldsymbol{G} = (\boldsymbol{S}' \boldsymbol{M}^{-1} \boldsymbol{S})^{-1} \boldsymbol{S}' \boldsymbol{M}^{-1}$, for some positive definite matrix $\boldsymbol{M} \in \mathbb{R}^{n \times n}$ [@Gamakumara2020-ml].
-
-When setting $\boldsymbol{M} = \boldsymbol{I}_n$, the identity matrix, this reduces to the OLS reconciliation, which corresponds to an orthogonal projection onto the coherent subspace (similar to @fig-geometry).
-A schematic illustration of this projection is depicted in @fig-geometry.
+Note that this figure is a schematic since most applications are high-dimensional.](../figs/geometry_orthogonal.png){#fig-geometry width="60%"}
+
+<!-- connect point to probabilistic, at some point mention we refer point forecasts as forecasts of means -->
+
+This projection view extends seamlessly from point to probabilistic forecast reconciliation.
+Let $\hat{\boldsymbol{Y}}_{t+h|t}$ denote the random vector representing the $h$-step-ahead base forecasts.
+The random reconciled forecast vector is obtained similarly via projection: $\tilde{\boldsymbol{Y}}_{t+h|t} = \boldsymbol{P} \hat{\boldsymbol{Y}}_{t+h|t}$.
+Consequently, point forecasts (which we refer to as means of the predictive distributions) reconcile as $\mathbb{E}(\tilde{\boldsymbol{Y}}_{t+h|t}) = \mathbb{E}(\boldsymbol{P} \hat{\boldsymbol{Y}}_{t+h|t}) = \boldsymbol{P} \hat{\boldsymbol{y}}_{t+h|t} = \tilde{\boldsymbol{y}}_{t+h|t}$, and the covariance of the predictive distribution transforms linearly as $\text{Var}(\tilde{\boldsymbol{Y}}_{t+h|t}) = \text{Var}(\boldsymbol{P} \hat{\boldsymbol{Y}}_{t+h|t}) = \boldsymbol{P} \boldsymbol{W}_h \boldsymbol{P}'$, where $\boldsymbol{W}_h = \text{Var}(\hat{\boldsymbol{Y}}_{t+h|t})$.
+
+@fig-geometry schematically illustrates this projection.
+Reconciliation takes the (possibly elliptical) base forecast distribution and "pushes" the entire mass linearly onto the coherent subspace, resulting in the reconciled forecast distribution.
+
+<!-- If the base forecast distribution has mean $\mathbb{E}(\hat{\boldsymbol{Y}}_{t+h|t}) = \hat{\boldsymbol{y}}_{t+h|t}$ and covariance $\text{Var}(\hat{\boldsymbol{Y}}_{t+h|t}) = \boldsymbol{W}_h$, then the reconciled forecast distribution has mean $\mathbb{E}(\tilde{\boldsymbol{Y}}_{t+h|t}) = \mathbb{E}(\boldsymbol{P} \hat{\boldsymbol{Y}}_{t+h|t}) = \boldsymbol{P} \hat{\boldsymbol{y}}_{t+h|t} = \tilde{\boldsymbol{y}}_{t+h|t}$ and covariance $\text{Var}(\tilde{\boldsymbol{Y}}_{t+h|t}) = \text{Var}(\boldsymbol{P} \hat{\boldsymbol{Y}}_{t+h|t}) = \boldsymbol{P} \boldsymbol{W}_h \boldsymbol{P}'$.
+-->
+
+<!-- A schematic illustration of this projection is depicted in @fig-geometry.
 Reconciliation takes the (possibly elliptical) base forecast distribution and projects it onto the coherent subspace, resulting in the reconciled forecast distribution.
-The choice of $\boldsymbol{M}$ determines the direction of the projection, which can yield oblique projection to deliver better-performing forecasts.
-
-## The Minimum Trace (MinT) Reconciliation
+-->
+
+## The Minimum Trace Reconciliation
 
 @Wickramasuriya2019-xq showed that by setting $\boldsymbol{M} = \boldsymbol{W}_h = \mathbb{E}( \hat{\boldsymbol{e}}_{t+h|t} \; \hat{\boldsymbol{e}}_{t+h|t}' )$, the covariance matrix of the $h$-step-ahead base forecast errors $\hat{\boldsymbol{e}}_{t+h|t} = \boldsymbol{y}_{t+h} - \hat{\boldsymbol{y}}_{t+h|t}$, we essentially minimise the total variance of the reconciled forecast errors across all series.
-Equivalently, MinT is the unique linear-unbiased reconciler that minimises the trace of $\text{Var}[y_{t+h} - \tilde{y}_{t+h|t}] = \boldsymbol{S} \boldsymbol{G}_h \boldsymbol{W}_h \boldsymbol{G}_h' \boldsymbol{S}'$.
+Equivalently, MinT is the unique linear-unbiased reconciler that minimises the trace of $\text{Var}[\boldsymbol{y}_{t+h} - \tilde{\boldsymbol{y}}_{t+h|t}] = \boldsymbol{S} \boldsymbol{G}_h \boldsymbol{W}_h \boldsymbol{G}_h' \boldsymbol{S}'$.
 This method is thus called Minimum Trace (MinT) reconciliation.
@@
 provided that $\boldsymbol{W}_h$ is positive definite.
-Intuitively, $\boldsymbol{W}_h^{-1}$ down-weights the base forecasts of series with high uncertainty, and up-weights those with low uncertainty, when mapping to the bottom level.
+
+<!-- Intuitively, $\boldsymbol{W}_h^{-1}$ down-weights the base forecasts of series with high uncertainty, and up-weights those with low uncertainty, when mapping to the bottom level. -->
 
 Although originally developed for point reconciliation, @Wickramasuriya2024-mb showed that MinT extends naturally to the probabilistic setting when base predictive distributions are Gaussian.
-If $h$-step-ahead base forecast distribution be $\hat{\boldsymbol{y}}_{t+h|t} \sim \mathcal{N}(\hat{\boldsymbol{y}}_{t+h|t}, \boldsymbol{W}_h)$, then reconciliation gives $\tilde{\boldsymbol{y}}_{t+h|t} = \boldsymbol{S} \boldsymbol{G}_h \hat{\boldsymbol{y}}_{t+h|t} \sim \mathcal{N}(\tilde{\boldsymbol{y}}_{t+h|t}, \boldsymbol{S} \boldsymbol{G}_h \boldsymbol{W}_h \boldsymbol{G}_h' \boldsymbol{S}')$.
+If $h$-step-ahead base forecast distribution is $\mathcal{N}(\hat{\boldsymbol{y}}_{t+h|t}, \boldsymbol{W}_h)$, then reconciliation gives $\tilde{\boldsymbol{Y}}_{t+h|t} = \boldsymbol{S} \boldsymbol{G}_h \hat{\boldsymbol{Y}}_{t+h|t} \sim \mathcal{N}(\tilde{\boldsymbol{y}}_{t+h|t}, \boldsymbol{S} \boldsymbol{G}_h \boldsymbol{W}_h \boldsymbol{G}_h' \boldsymbol{S}')$.
 Within the class of linear-unbiased reconcilers, MinT also minimises the negative log score (equivalently, the Gaussian log predictive score) of the reconciled distribution.
 
-In this paper we adopt this Gaussian framework to compare covariance estimators for MinT in both point and probabilistic reconciliation.
+In this paper we adopt the Gaussian framework to compare the impacts of different covariance estimators for MinT in both point and probabilistic reconciliation.
 It is also worth to mention that the methods can be extended to non-Gaussian settings by bootstrapping [@Gamakumara2020-ml; @Panagiotelis2023-fs], which is beyond the scope of this paper.
@@
 
-```{r load results}
+```{r}
+#| label: load sim results
 sim_params_S36 <- readRDS(paste0(path_lib, "params_S36_6_1.rds"))
@@
 
-```{r settings 6x6}
+```{r}
 #| label: fig-settings-6x6
@@
 
-```{r settings 2x50}
+```{r}
 #| label: fig-settings-2x50

### 2025-10-23 thesis/paper_draft.qmd
refine shriankge and problem sections + change screeplot to corr (codes)

@@
 
-The performance of MinT hinges on a reliable, positive-definite estimate of $\boldsymbol{W}_h$, which comes in both the mapping matrix $\boldsymbol{G}_h$ and the reconciled forecast variance $\boldsymbol{S} \boldsymbol{G}_h \boldsymbol{W}_h \boldsymbol{G}_h' \boldsymbol{S}'$.
+The performance of MinT hinges on a reliable, positive-definite estimate of $\boldsymbol{W}_h$, which comes in both the mapping matrix $\boldsymbol{G}_h$ and the reconciled forecast error variance $\boldsymbol{S} \boldsymbol{G}_h \boldsymbol{W}_h \boldsymbol{G}_h' \boldsymbol{S}'$.
 
 However, the covariance matrix $\boldsymbol{W}_h$ is often not available in closed-form, and is challenging to estimate in high-dimensional setting where the number of series $n$ is larger than the time dimension $T$.
-To tackle this issue, the original paper @Wickramasuriya2019-xq assumed a proportionality relationship between $\hat{\boldsymbol{W}}^g_h = k_h g( \hat{\boldsymbol{W}}_1 )$, where $\hat{\boldsymbol{W}}_1$ is the covariance matrix of the in-sample 1-step-ahead base residuals (to approximate $\boldsymbol{W}_1$) and $k_h > 0$ is a scaling constant (which will be algebraically cancelled out in point-forecast reconciliation).
+To tackle this issue, the original paper by @Wickramasuriya2019-xq assumed a proportionality relationship $\hat{\boldsymbol{W}}^g_h = k_h g( \hat{\boldsymbol{W}}_1 )$, where $\hat{\boldsymbol{W}}_1$ is the sample covariance matrix of the in-sample 1-step-ahead base forecast errors (to approximate $\boldsymbol{W}_1$) and $k_h > 0$ is a scaling constant (which will be algebraically cancelled out in point-forecast reconciliation).
 The function $g(.)$ is a covariance estimator that produces a positive-definite matrix, the main focus of this paper.
@@
 
-where $\text{diag}( \hat{\boldsymbol{W}}_1 )$ comprises only the diagonal elements of $\hat{\boldsymbol{W}}_1$.
+where $\text{diag}( \hat{\boldsymbol{W}}_1 )$ is the diagonal matrix comprising the diagonal elements of $\hat{\boldsymbol{W}}_1$.
 We refer to any $\lambda_S \in [0,1]$ as the shrinkage intensity of the shrinkage estimator.
-This approach shrinks the covariance matrix $\hat{\boldsymbol{W}}_1$ towards its diagonal matrix, meaning the off-diagonal elements are shrunk towards zero while the diagonal ones remain unchanged.
-
-@Schafer2005-yw also provided an closed-form estimate of the optimal shrinkage intensity parameter $\lambda_S$:
+This approach shrinks the covariance matrix $\hat{\boldsymbol{W}}_1$ towards the diagonal target $\text{diag}( \hat{\boldsymbol{W}}_1 )$, meaning the off-diagonal elements are shrunk towards zero while the diagonal ones remain unchanged.
+
+@Schafer2005-yw also provided a closed-form estimate of the optimal shrinkage intensity parameter $\lambda_S$:
 
@@
 
-Linear shrinkage shrinks all off-diagonal elements towards zeros with equal weights $\lambda_S$.
+Linear shrinkage shrinks all off-diagonal elements of $\hat{\boldsymbol{W}}_1$ towards zeros with equal weights $1 - \lambda_S$.
 The resulting penalty is global and non-adaptive: strong, genuinely systematic correlations are shrunk at the same rate as weak, noisy ones.
@@
 Aggregation naturally induces stronger dependence among related nodes (for example, a region and a neighbouring region, or a region and its parent state), while many unrelated pairs exhibit near-zero correlations.
-A uniform penalty may over-shrink informative co-movements and under-shrink idiosyncratic noise, reducing reconciliation efficiency.
+A uniform penalty may over-shrink informative co-movements or under-shrink idiosyncratic noise, reducing reconciliation efficiency.
 
@@
 
-Many real-world hierarchical data sets exhibit a prominent low-rank (factor) structure.
-
-In Australian domestic tourism, for example, national and state-level movements often load on a small number of common components, with residual dependence concentrated in local idiosyncrasies.
-A scree plot of the one-step residual covariance typically shows a marked elbow after a handful of principal components, indicating strong latent structure.
+Many real-world hierarchical data sets may exhibit a prominent low-rank (factor) structure.
+A clear example is the Australian domestic overnight trips [@tourism-au], where the national trips are disaggregated into states and territories, and further into regions.
+The tourism activities might be driven by a few common factors, such as economic conditions, fuel prices, or major events, which might be left in the forecast errors after fitting the base models.
+@fig-eigen illustrates the largest eigenvalues from the one-step forecast error correlation matrix $\hat{\boldsymbol{R}}_1$, typically showing a marked elbow after a handful principal components (largest eigenvalues), indicating strong latent structure.
 Diagonal-target shrinkage is factor-unaware: it neither preserves the low-rank common subspace nor differentially shrinks the idiosyncratic remainder, and so can be inefficient when common factors dominate.
 
-Taking an example of the Australian domestic overnight trips data set [@tourism-au], where the national trips are disaggregated into states and territories, and further into regions.
-The tourism activities might be driven by a few common factors, such as economic conditions, fuel prices, or major events, which might be left in the residuals after fitting the base models.
-From @fig-eigen, a scree plot of the one-step residual covariance $\hat{\boldsymbol{W}}_1$ shows a marked elbow after a handful of principal components (largest eigenvalues), indicating strong latent structure.
-
-![Twenty largest eigenvalues of one-step-ahead in-sample base forecast error covariance, Australian domestic overnight trips.
-The point of inflection occurs around the fifth largest eigenvalue.](D:/Github/Recon_Honours_Thesis/figs/eigen_tourism.png){#fig-eigen width="70%"}
-
-### Problem 3: Proportional scaling for h-step-ahead {.unnumbered}
+<!-- ![Scree plot of the eigenvalues of the 1-step-ahead forecast error covariance matrix from the Australian domestic tourism dataset.
+The point of inflection occurs around the fifth largest eigenvalue.](D:/Github/Recon_Honours_Thesis/figs/eigen_tourism.png){#fig-eigen width="70%"} -->
+
+```{r}
+#| label: fig-eigen
+#| fig.cap: "Scree plot of the eigenvalues of the 1-step-ahead forecast error covariance matrix from the Australian domestic tourism dataset.
+The point of inflection occurs around the fourth largest eigenvalue."
+#| fig.width: 7
+#| fig.height: 4
+#| out.width: "70%"
+tourism_resid <- readRDS(paste0(path_lib, "tourism_resid.rds"))
+
+par(mfrow = c(1,1), mar = c(4,4,1,1))
+
+(eigen(cor(tourism_resid))$values[1:20]/1e6) %>%
+  plot(type = "b",
+       xlab = "Component",
+       ylab = "Eigenvalue ('000,000)"
+  )
+# add a vertical line at 5
+abline(v = 4, col = "purple", lty = 2, lwd = 2)
+```
+
+### Problem 3: Proportional scaling {.unnumbered}
 
 The proportionality relationship $\hat{\boldsymbol{W}}^g_h = k_h g( \hat{\boldsymbol{W}}_1 )$ enforces horizon-invariant cross-series dependence of the forecast errors up to a scalar factor.
-In many applications, the shape of the error covariance might change with horizon due to evolving error dynamics or horizon-specific interactions among series.
+In many applications, the structure of the error covariance might change with horizon due to evolving error dynamics or horizon-specific interactions among series.
 Proportional scaling may therefore misrepresent multi-step dependence.

### 2025-10-23 thesis/paper_draft.qmd
refine NOVELIST and algo cv + re render

@@
 
-NOVELIST (NOVEL Integration of the Sample and Thresholded Covariance) estimator, proposed by @Huang2019-ua, introduces a parameter to control the sparsity level in the target toward which the sample matrix is shrunk.
+To tackle the first limitation of uniform shrinkage, we consider the NOVELIST estimator from @Huang2019-ua.
+The NOVELIST (NOVEL Integration of the Sample and Thresholded Covariance) estimator introduces extra parameter, the threshold $\delta$, to control the sparsity level in the target toward which the sample covariance matrix is shrunk.
 By soft-thresholding small correlations before shrinkage, NOVELIST preserves strong signals while attenuating weak, noisy ones, addressing the uniform, non-adaptive nature of diagonal-target shrinkage.
@@
 First, apply elementwise soft-thresholding to the sample correlation matrix; second, shrink the sample correlation toward this thresholded target.
-Working in correlation space avoids rescaling issues and keeps diagonal entries at one.
-The method introduces an extra parameter, the threshold $\delta$, which is used to control the amount of soft-thresholding.
 The NOVELIST estimator for covariance matrix is given by:
@@
 
-where $\hat{\boldsymbol{W}}_{1, \delta}$ is the thresholded version of $\hat{\boldsymbol{W}}_{1}$. By convenient setting as discussed above, we rewrite it in terms of correlation:
+where $\hat{\boldsymbol{W}}_{1, \delta}$ is the thresholded covariance matrix of $\hat{\boldsymbol{W}}_{1}$.
+As working with correlations avoids rescaling issues and keeps diagonal entries at one, we rewrite it in terms of correlations:
 
@@
 
-where $\delta \in [0,1]$ is the threshold parameter.
+where $\delta \in [0,1]$.
 For a given threshold $\delta$, @Huang2019-ua derived an analytical expression for the optimal shrinkage intensity parameter $\lambda(\delta)$, following similar logic to @Schafer2005-yw.
@@
 
-On the other hand, the optimal threshold $\hat \delta$ does not have a closed-form solution, and is typically obtained by rolling-window cross-validation procedure.
-The idea is to find the threshold $\delta^*$, with the corresponding $\lambda^*$ and $\hat{\boldsymbol{R}}^{N}_{1}(\delta^*, \lambda^*)$, that minimises the average out-of-sample 1-step-ahead reconciled forecast mean squared error over all windows.
-The formal algorithm is given in @sec-novelist_cv.
+On the other hand, the optimal threshold $\hat \delta$ does not have a closed-form solution, and is obtained by rolling-window cross-validation procedure.
+The idea is to find the threshold $\delta^*$, with the corresponding $\lambda^*$ and $\hat{\boldsymbol{R}}^{N}_{1}(\delta^*, \lambda^*)$, that minimises the average out-of-sample 1-step-ahead mean squared reconciled forecast errors over all windows.
+The formal algorithm is given in Algorithm \ref{alg:novelist_cv}.
 
 Note that when $\delta \in \bigl[ \text{max}_{i \neq j}|\hat{r}_{1,ij}|, \; 1 \bigr]$, the NOVELIST estimator collapses to the shrinkage estimator, and when $\delta = 0$, it becomes the sample covariance matrix.
-An additional concern is that the estimator does not guarantee to be positive definite, but we can use @Higham2002-te algorithm to compute the nearest positive definite matrix if needed.
-
-### NOVELIST cross-validation algorithm {#sec-novelist_cv}
-
-It is only required to fit the base models once on the whole training data $\{\boldsymbol{y}_t\}_{t=1}^T$, and obtain the in-sample fitted values $\{\hat{\boldsymbol{y}}_t\}_{t=1}^T$.
+An additional concern is that the estimator is not guaranteed to be positive definite, but we can use @Higham2002-te algorithm to compute the nearest positive definite matrix if needed.
 
@@
 \begin{algorithm}[H]
-\caption{Cross-validation procedure}
+\caption{Cross-validation procedure}\label{alg:novelist_cv}
 \begin{algorithmic}[1]
 
-\State \textbf{Input:} Observations and fitted values $\boldsymbol{y}_t, \hat{\boldsymbol{y}}_t \in \mathbb{R}^n$ for $t = 1,\dots,T$, set of threshold candidates $\Delta$, window size $v$.
+\State \textbf{Input:} Observations and fitted values $\boldsymbol{y}_t, \hat{\boldsymbol{y}}_t \in \mathbb{R}^n$ for $t = 1,\dots,T$\footnotemark, set of threshold candidates $\Delta$, window size $v$.
 
@@
 ```
+
+\footnotetext{It is only required to fit the base models once on the whole training data $\{\boldsymbol{y}_t\}_{t=1}^T$, and obtain the in-sample fitted values $\{\hat{\boldsymbol{y}}_t\}_{t=1}^T$.}
+
+\vspace{-1em}
 
@@
 The 1- to 12-step-ahead base forecasts are then generated by these ARIMA models, and then reconciled using multiple approaches.
-To estimate the NOVELIST and its variants, we implement an extra cross-validation prodecure within this training window, as described in @sec-novelist_cv.
+To estimate the NOVELIST and its variants, we implement an extra cross-validation prodecure within this training window, as described in Algorithm \ref{alg:novelist_cv}.
 We then roll the training window forward by one month and refit all models, re-estimate reconciliation variants, and produce another batch of 1- to 12-step-ahead forecasts, repeating until the training set reaches December 2015.

### 2025-10-23 thesis/paper_draft.qmd
change algo, remove subscript

@@
     number-sections: true
-# crossref:
-#   custom:
-#     - kind: float
-#       key: alg
-#       reference-prefix: Algorithm
-#       caption-prefix: Algorithm
-#       # for PDF output only, tell LaTeX to use the `algorithm` env:
-#       latex-env: algorithm
 link-citations: true
@@
 
-### Problem 1: Uniform shrinkage {.unnumbered}
+### Problem 1: Uniform shrinkage {#prob1 .unnumbered}
 
@@
 
-To tackle the first limitation of uniform shrinkage, we consider the NOVELIST estimator from @Huang2019-ua.
+To tackle the Problem [1](#prob1) of uniform shrinkage, we consider the NOVELIST estimator from @Huang2019-ua.
 The NOVELIST (NOVEL Integration of the Sample and Thresholded Covariance) estimator introduces extra parameter, the threshold $\delta$, to control the sparsity level in the target toward which the sample covariance matrix is shrunk.
@@
 
-```{=latex}
+<!-- ```{=latex}
 \begin{algorithm}[H]
@@
 
-\State \textbf{Input:} Observations and fitted values $\boldsymbol{y}_t, \hat{\boldsymbol{y}}_t \in \mathbb{R}^n$ for $t = 1,\dots,T$\footnotemark, set of threshold candidates $\Delta$, window size $v$.
+\State \textbf{Input:} Observations and fitted values $\boldsymbol{y}_t, \hat{\boldsymbol{y}}_t \in \mathbb{R}^n$ for $t = 1,\dots,T$ \footnotemark, set of threshold candidates $\Delta$, window size $v$.
 
@@
 \State $\text{MSE}_{\delta} = \frac{1}{T-v} \sum_{i=v}^{T-1} (\tilde{\boldsymbol{e}}_{i+1 | \delta})^2$ for each $\delta \in \Delta$
+\State $\hat{\delta} = \arg\min_{\delta \in \Delta} \text{MSE}_{\delta}$
+
+\State \textbf{Output:} Estimate of optimal threshold parameter $\hat{\delta}$
+
+\end{algorithmic}
+\end{algorithm}
+``` -->
+
+```{=latex}
+\begin{algorithm}[H]
+\caption{Cross-validation procedure}\label{alg:novelist_cv}
+\begin{algorithmic}[1]
+
+\State \textbf{Input:} Observations and fitted values $\boldsymbol{y}_t, \hat{\boldsymbol{y}}_t \in \mathbb{R}^n$ for $t = 1,\dots,T$ \footnotemark, set of threshold candidates $\Delta$, window size $v$.
+
+\State $\hat{\boldsymbol{e}}_t = \boldsymbol{y}_t - \hat{\boldsymbol{y}}_t$ for $t = 1,\dots,T$
+
+\For{$i = v:T-1$}
+    \State $j = i - v +1$
+    \State $\hat{\boldsymbol{W}}_1 = \frac{1}{v} \sum_{t=j}^{i} \hat{\boldsymbol{e}}_{t} \hat{\boldsymbol{e}}_{t}'$
+    \State $\hat{\boldsymbol{D}} = \text{diag}(\hat{\boldsymbol{W}}_1)$
+    \State $\hat{\boldsymbol{R}}_1 = \hat{\boldsymbol{D}}^{-1/2} \hat{\boldsymbol{W}}_1 \hat{\boldsymbol{D}}^{-1/2}$
+    \For{$\delta \in \Delta$}
+        \State Compute thresholded correlation $\hat{\boldsymbol{R}}_{1,\delta}$ using Equation 2
+        \State Compute shrinkage parameter $\hat{\lambda}_{\delta}$ using Equation 3
+        \State Compute $\hat{\boldsymbol{R}}^{N}_{1,\delta}$ using Equation 1
+        \State $\hat{\boldsymbol{W}}^{N}_{1,\delta} = \hat{\boldsymbol{D}}^{1/2} \hat{\boldsymbol{R}}^{N}_{1,\delta} \hat{\boldsymbol{D}}^{1/2}$
+        \State $\boldsymbol{G} = (\boldsymbol{S}' \hat{\boldsymbol{W}}^{N^{-1}}_{1,\delta} \boldsymbol{S})^{-1} \boldsymbol{S}' \hat{\boldsymbol{W}}^{N^{-1}}_{1,\delta}$
+        \State Reconciled forecasts $\tilde{\boldsymbol{y}}_{i+1} = \boldsymbol{S} \boldsymbol{G} \hat{\boldsymbol{y}}_{i+1}$
+        \State $\tilde{\boldsymbol{e}}_{i+1, \delta} = \boldsymbol{y}_{i+1} - \tilde{\boldsymbol{y}}_{i+1}$
+    \EndFor
+\EndFor
+
+\State $\text{MSE}_{\delta} = \frac{1}{T-v} \sum_{i=v}^{T-1} (\tilde{\boldsymbol{e}}_{i+1, \delta})^2$ for each $\delta \in \Delta$
 \State $\hat{\delta} = \arg\min_{\delta \in \Delta} \text{MSE}_{\delta}$

### 2025-10-23 thesis/paper_draft.qmd
revert screeplot back to cov

@@
 The tourism activities might be driven by a few common factors, such as economic conditions, fuel prices, or major events, which might be left in the forecast errors after fitting the base models.
-@fig-eigen illustrates the largest eigenvalues from the one-step forecast error correlation matrix $\hat{\boldsymbol{R}}_1$, typically showing a marked elbow after a handful principal components (largest eigenvalues), indicating strong latent structure.
+@fig-eigen illustrates the largest eigenvalues from the one-step forecast error correlation matrix $\hat{\boldsymbol{R}}_1$, typically showing a marked elbow after a few principal components (largest eigenvalues), indicating strong latent structure.
 Diagonal-target shrinkage is factor-unaware: it neither preserves the low-rank common subspace nor differentially shrinks the idiosyncratic remainder, and so can be inefficient when common factors dominate.
@@
 #| fig.cap: "Scree plot of the eigenvalues of the 1-step-ahead forecast error covariance matrix from the Australian domestic tourism dataset.
-The point of inflection occurs around the fourth largest eigenvalue."
+The point of inflection occurs around the second largest eigenvalue."
 #| fig.width: 7
@@
 
-(eigen(cor(tourism_resid))$values[1:20]/1e6) %>%
+(eigen(cov(tourism_resid))$values[1:20]/1e6) %>%
   plot(type = "b",
@@
 # add a vertical line at 5
-abline(v = 4, col = "purple", lty = 2, lwd = 2)
+abline(v = 2, col = "purple", lty = 2, lwd = 2)
 ```

### 2025-10-23 thesis/paper_draft.qmd
refine other cov ests, consistent g(.) for cov only

@@
 $$
-\hat{\boldsymbol{W}}^{S}_{1} = \lambda_S \, \text{diag}( \hat{\boldsymbol{W}}_1 ) + (1 - \lambda_S) \hat{\boldsymbol{W}}_1 \, ,
-$$
-
-where $\text{diag}( \hat{\boldsymbol{W}}_1 )$ is the diagonal matrix comprising the diagonal elements of $\hat{\boldsymbol{W}}_1$.
+\hat{\boldsymbol{W}}^{S}_{1} = \lambda_S \, \boldsymbol{D}_1 + (1 - \lambda_S) \hat{\boldsymbol{W}}_1 \, ,
+$$
+
+where $\boldsymbol{D}_1 = \text{diag}( \hat{\boldsymbol{W}}_1 )$ is the diagonal matrix comprising the diagonal elements of $\hat{\boldsymbol{W}}_1$ (i.e., the sample variances).
 We refer to any $\lambda_S \in [0,1]$ as the shrinkage intensity of the shrinkage estimator.
@@
 
-### Problem 2: Latent factors {.unnumbered}
+### Problem 2: Latent factors {#prob2 .unnumbered}
 
@@
 
-### Problem 3: Proportional scaling {.unnumbered}
+### Problem 3: Proportional scaling {#prob3 .unnumbered}
 
@@
 
-To exploit latent common components explicitly, the PC-adjusted method takes the latent factors directly into its construction.
-It starts by decomposing the covariance matrix $\hat{\boldsymbol{W}}_1$ into a prominent principle components part (low-rank) and a orthogonal complement part $\hat{\boldsymbol{W}}^K_{1}$ (the correlation matrix after removing the first $K$ principal components).
+To address Problem [2](#prob2) and exploit latent common components explicitly, the PC-adjusted (Principal-Component-adjusted) method takes the latent factors directly into its construction.
+It starts by decomposing the covariance matrix $\hat{\boldsymbol{W}}_1$ into a prominent principle components part (low-rank) and a orthogonal complement part $\hat{\Upsilon}$ (the covariance matrix after removing the first $K$ principal components).
 Then we can apply either shrinkage or NOVELIST estimator to $\hat{\boldsymbol{W}}^K_{1}$:
@@
 $$
-\hat{\boldsymbol{W}}^{g, K}_{1} = \sum_{k = 1}^K \hat{\gamma}_k \hat{\boldsymbol{\xi}}_k \hat{\boldsymbol{\xi}}_k' + g(\hat{\boldsymbol{W}}^K_{1})
+\hat{\boldsymbol{W}}^{g, K}_{1} = \sum_{k = 1}^K \hat{\gamma}_k \hat{\boldsymbol{\xi}}_k \hat{\boldsymbol{\xi}}_k' + g(\hat{\Upsilon})
 $$
@@
 where $g(.)$ is either the shrinkage or NOVELIST estimator, $\hat{\gamma}_k$ and $\hat{\boldsymbol{\xi}}_k$ are the $k$-th largest eigenvalue and the corresponding eigenvector of the sample covariance matrix, respectively.
-The number of principal components $K$ can be selected using information criteria.
+
+<!-- The number of principal components $K$ can be selected using information criteria. -->
 
 Similar to the NOVELIST estimator, its PC-adjusted variant $\hat{\boldsymbol{W}}^{N, K}_{1}$ requires a cross-validation procedure to select the threshold paramter and adjustment to obtain positive definiteness.
+The reconciled forecasts for each cross-validation window are obtained using MinT with $\hat{\boldsymbol{W}}^{N, K}_{1}$.
 
@@
 
-To address the potential issue of the proportionality relationship $\hat{\boldsymbol{W}}^g_h = k_h g( \hat{\boldsymbol{W}}_1 )$ not holding in practice.
-A simple relaxation retains the $1$-step-ahead correlation shape but allows the variances to scale differently with horizon.
+To address the potential issue of Problem [3](#prob3) of proportionality relationship $\hat{\boldsymbol{W}}^g_h = k_h g( \hat{\boldsymbol{W}}_1 )$ not holding in practice.
+A simple relaxation retains the $1$-step-ahead correlation structure but allows the variances to scale differently with horizon.
 The scaled variance estimator is given by:
@@
 $$
-\hat{\boldsymbol{W}}^{g, sv}_{h} = \boldsymbol{D}^{1/2}_h g(\hat{\boldsymbol{R}}_1) \boldsymbol{D}^{1/2}_h,
-$$
-
-where $\boldsymbol{D}_h = \text{diag}(\hat{\sigma}^2_{1,h}, \ldots, \hat{\sigma}^2_{n,h})$, and $\hat{\sigma}^2_{i,h}$ is the variance of the $i$-th series' h-step-ahead base forecast errors.
-Similarly, $g(.)$ is either the shrinkage or NOVELIST estimator.
-
-When NOVELIST is used to produce $h$-step-ahead reconciled forecasts, the cross-validation procedure is slightly modified to evaluate the out-of-sample reconciled forecast MSE using $\hat{\boldsymbol{W}}^{N, sv}_{h}$ instead of $\hat{\boldsymbol{W}}^{N}_{1}$.
+\hat{\boldsymbol{W}}^{g, sv}_{h} = \boldsymbol{D}^{1/2}_h \hat{\boldsymbol{R}}^g_1 \boldsymbol{D}^{1/2}_h, \quad
+
+\hat{\boldsymbol{R}}^g_1 = D^{-1/2}_1 \, g(\hat{\boldsymbol{W}}_1) \, D^{-1/2}_1,
+$$
+
+where $\boldsymbol{D}_h$ is a diagonal matrix comprising $\hat{\sigma}^2_{1,h} , \ldots, \hat{\sigma}^2_{n,h}$ on its diagonal, and $\hat{\sigma}^2_{i,h}$ is $h$-step-ahead base forecast error variance of the $i$-th series.
+$\hat{\boldsymbol{R}}^g_1$ is the estimated one-step correlation matrix using either shrinkage or NOVELIST estimator.
+
+When NOVELIST is used to produce $h$-step-ahead reconciled forecasts, the cross-validation procedure is modified to evaluate the out-of-sample $h$-step-ahead reconciled forecasts for each window and estimate optimal threshold parameter (instead of evaluating $1$-step-ahead reconciled forecasts).
+The reconciled forecasts are obtained using MinT with $\hat{\boldsymbol{W}}^{N, sv}_{h}$.
 
@@
 
-Another alternative is to directly estimate the covariance matrix from the h-step-ahead base forecast errors, without assuming any proportionality relationship:
-
-$$
-\hat{\boldsymbol{W}}^{g}_{h} = g(\hat{\boldsymbol{W}}_h)
-$$
-
-where $\hat{\boldsymbol{W}}_h$ is the covariance matrix of in-sample h-step-ahead base forecast residuals, and $g(.)$ is either the shrinkage or NOVELIST estimator.
-
-Similar to the scaled variance approach, if NOVELIST is used, the cross-validation procedure is modified to compute $\hat{\boldsymbol{W}}^{N}_{h}$ and evaluate the out-of-sample reconciled forecast MSE using it.
+Another alternative is to directly estimate the covariance matrix from the $h$-step-ahead base forecast errors, without assuming any proportionality relationship:
+
+$$
+\hat{\boldsymbol{W}}^{g}_{h} = g(\hat{\boldsymbol{W}}_h),
+$$
+
+where $\hat{\boldsymbol{W}}_h$ is the covariance matrix of in-sample h-step-ahead base forecast errors, and $g(.)$ is either the shrinkage or NOVELIST estimator.
+
+Similarly, if NOVELIST is used, the cross-validation procedure is modified to evaluate the out-of-sample $h$-step-ahead reconciled forecasts for each window.
+The reconciled forecasts are obtained using MinT with $\hat{\boldsymbol{W}}^{N}_{h}$.
 

### 2025-10-23 thesis/paper_draft.tex
refine other cov ests, consistent g(.) for cov only

% Options for packages loaded elsewhere
% Options for packages loaded elsewhere
\PassOptionsToPackage{unicode}{hyperref}
\PassOptionsToPackage{hyphens}{url}
\PassOptionsToPackage{dvipsnames,svgnames,x11names}{xcolor}
%
\documentclass[
  11pt,
  letterpaper,
  DIV=11,
  numbers=noendperiod,
  titlepage]{scrartcl}
\usepackage{xcolor}
\usepackage{amsmath,amssymb}
\setcounter{secnumdepth}{5}
\usepackage{iftex}
\ifPDFTeX
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
  \usepackage{textcomp} % provide euro and other symbols
\else % if luatex or xetex
  \ifXeTeX
    \usepackage{mathspec} % this also loads fontspec
  \else
    \usepackage{unicode-math} % this also loads fontspec
  \fi
  \defaultfontfeatures{Scale=MatchLowercase}
  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
\fi
\usepackage{lmodern}
\ifPDFTeX\else
  % xetex/luatex font selection
\fi
% Use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\IfFileExists{microtype.sty}{% use microtype if available
  \usepackage[]{microtype}
  \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\usepackage{setspace}
\makeatletter
\@ifundefined{KOMAClassName}{% if non-KOMA class
  \IfFileExists{parskip.sty}{%
    \usepackage{parskip}
  }{% else
    \setlength{\parindent}{0pt}
    \setlength{\parskip}{6pt plus 2pt minus 1pt}}
}{% if KOMA class
  \KOMAoptions{parskip=half}}
\makeatother
% Make \paragraph and \subparagraph free-standing
\makeatletter
\ifx\paragraph\undefined\else
  \let\oldparagraph\paragraph
  \renewcommand{\paragraph}{
    \@ifstar
      \xxxParagraphStar
      \xxxParagraphNoStar
  }
  \newcommand{\xxxParagraphStar}[1]{\oldparagraph*{#1}\mbox{}}
  \newcommand{\xxxParagraphNoStar}[1]{\oldparagraph{#1}\mbox{}}
\fi
\ifx\subparagraph\undefined\else
  \let\oldsubparagraph\subparagraph
  \renewcommand{\subparagraph}{
    \@ifstar
      \xxxSubParagraphStar
      \xxxSubParagraphNoStar
  }
  \newcommand{\xxxSubParagraphStar}[1]{\oldsubparagraph*{#1}\mbox{}}
  \newcommand{\xxxSubParagraphNoStar}[1]{\oldsubparagraph{#1}\mbox{}}
\fi
\makeatother

\usepackage{longtable,booktabs,array}
\usepackage{calc} % for calculating minipage widths
% Correct order of tables after \paragraph or \subparagraph
\usepackage{etoolbox}
\makeatletter
\patchcmd\longtable{\par}{\if@noskipsec\mbox{}\fi\par}{}{}
\makeatother
% Allow footnotes in longtable head/foot
\IfFileExists{footnotehyper.sty}{\usepackage{footnotehyper}}{\usepackage{footnote}}
\makesavenoteenv{longtable}
\usepackage{graphicx}
\makeatletter
\newsavebox\pandoc@box
\newcommand*\pandocbounded[1]{% scales image to fit in text height/width
  \sbox\pandoc@box{#1}%
  \Gscale@div\@tempa{\textheight}{\dimexpr\ht\pandoc@box+\dp\pandoc@box\relax}%
  \Gscale@div\@tempb{\linewidth}{\wd\pandoc@box}%
  \ifdim\@tempb\p@<\@tempa\p@\let\@tempa\@tempb\fi% select the smaller of both
  \ifdim\@tempa\p@<\p@\scalebox{\@tempa}{\usebox\pandoc@box}%
  \else\usebox{\pandoc@box}%
  \fi%
}
% Set default figure placement to htbp
\def\fps@figure{htbp}
\makeatother

% definitions for citeproc citations
\NewDocumentCommand\citeproctext{}{}
\NewDocumentCommand\citeproc{mm}{%
  \begingroup\def\citeproctext{#2}\cite{#1}\endgroup}
\makeatletter
 % allow citations to break across lines
 \let\@cite@ofmt\@firstofone
 % avoid brackets around text for \cite:
 \def\@biblabel#1{}
 \def\@cite#1#2{{#1\if@tempswa , #2\fi}}
\makeatother
\newlength{\cslhangindent}
\setlength{\cslhangindent}{1.5em}
\newlength{\csllabelwidth}
\setlength{\csllabelwidth}{3em}
\newenvironment{CSLReferences}[2] % #1 hanging-indent, #2 entry-spacing
 {\begin{list}{}{%
  \setlength{\itemindent}{0pt}
  \setlength{\leftmargin}{0pt}
  \setlength{\parsep}{0pt}
  % turn on hanging indent if param 1 is 1
  \ifodd #1
   \setlength{\leftmargin}{\cslhangindent}
   \setlength{\itemindent}{-1\cslhangindent}
  \fi
  % set entry spacing
  \setlength{\itemsep}{#2\baselineskip}}}
 {\end{list}}
\usepackage{calc}
\newcommand{\CSLBlock}[1]{\hfill\break\parbox[t]{\linewidth}{\strut\ignorespaces#1\strut}}
\newcommand{\CSLLeftMargin}[1]{\parbox[t]{\csllabelwidth}{\strut#1\strut}}
\newcommand{\CSLRightInline}[1]{\parbox[t]{\linewidth - \csllabelwidth}{\strut#1\strut}}
\newcommand{\CSLIndent}[1]{\hspace{\cslhangindent}#1}

\setlength{\emergencystretch}{3em} % prevent overfull lines

\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}

\usepackage{algorithm}
\usepackage{algpseudocode}
\usepackage[colorlinks=true, linkcolor=blue, citecolor=blue, urlcolor=blue]{hyperref}
\usepackage{array}
\usepackage{float}
\usepackage{xcolor}

% Heading font sizes
\usepackage{sectsty}
\sectionfont{\fontsize{16}{30}\selectfont}
\subsectionfont{\fontsize{14}{16}\selectfont}
\subsubsectionfont{\fontsize{12}{14}\selectfont}

\usepackage{graphicx}
\KOMAoption{captions}{tableheading}
\makeatletter
\@ifpackageloaded{caption}{}{\usepackage{caption}}
\AtBeginDocument{%
\ifdefined\contentsname
  \renewcommand*\contentsname{Table of contents}
\else
  \newcommand\contentsname{Table of contents}
\fi
\ifdefined\listfigurename
  \renewcommand*\listfigurename{List of Figures}
\else
  \newcommand\listfigurename{List of Figures}
\fi
\ifdefined\listtablename
  \renewcommand*\listtablename{List of Tables}
\else
  \newcommand\listtablename{List of Tables}
\fi
\ifdefined\figurename
  \renewcommand*\figurename{Figure}
\else
  \newcommand\figurename{Figure}
\fi
\ifdefined\tablename
  \renewcommand*\tablename{Table}
\else
  \newcommand\tablename{Table}
\fi
}
\@ifpackageloaded{float}{}{\usepackage{float}}
\floatstyle{ruled}
\@ifundefined{c@chapter}{\newfloat{codelisting}{h}{lop}}{\newfloat{codelisting}{h}{lop}[chapter]}
\floatname{codelisting}{Listing}
\newcommand*\listoflistings{\listof{codelisting}{List of Listings}}
\makeatother
\makeatletter
\makeatother
\makeatletter
\@ifpackageloaded{caption}{}{\usepackage{caption}}
\@ifpackageloaded{subcaption}{}{\usepackage{subcaption}}
\makeatother
\usepackage{bookmark}
\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\urlstyle{same}
\hypersetup{
  colorlinks=true,
  linkcolor={blue},
  filecolor={Maroon},
  citecolor={Blue},
  urlcolor={Blue},
  pdfcreator={LaTeX via pandoc}}

\author{}
\date{}
\begin{document}

\setstretch{1.15}
\begin{titlepage}

\begin{center}
\includegraphics[height=2cm,keepaspectratio]{D:/Github/Recon_Honours_Thesis/figs/monash-logo.png}
\end{center}

\centering

{%
\sffamily

% --- move the whole block down a bit ---
\vspace*{1.5cm}

{\huge \bfseries Enhancing Forecast Reconciliation:\\
A Study of\\
Alternative Covariance Estimators\par}

\vspace{2cm}

{\Large \bfseries Vincent Su\par}
{\large B.Com. (Hons), Monash University\par}

\vspace{1cm}

{\large Supervised by\par}
{\Large Shanika Wickramasuriya\\
and George Athanasopoulos\par}

\vfill

{\large A thesis submitted for the degree of \\
Bachelor of Commerce (Honours)\\
at Monash University\\
Department of Econometrics and Business Statistics\par}

\vspace{0.3cm}

{\large 25 October 2025\par}

}%
\end{titlepage}

\pagebreak

\section*{Abstract}\label{abstract}
\addcontentsline{toc}{section}{Abstract}

A collection of time series connected via a set of linear constraints is
known as hierarchical time series. Forecasting these series without
respecting the hierarchical nature of the data can lead to incoherent
forecasts across aggregation levels and, in practice, reduced accuracy.
Forecast reconciliation corrects this by adjusting base forecasts to
satisfy such constraints. Among modern reconciliation methods, Minimum
Trace (MinT) is widely used, however, it requires a good estimate of the
forecast error covariance matrix. The current practice uses linear
shrinkage towards a diagonal target. Furthermore, the covariance
estimate is based on 1-step-ahead residuals, then proportionally scale
it to approximate h-step-ahead covariance matrix. This leaves a question
of whether this method is appropriate for all real-world applications.
We study the shortcomings of current practice and propose alternative
covariance estimators, including the NOVELIST estimator (shrinkage
towards a soft-thresholded target), PC-adjusted shrinkage (which
utilises latent factor structures), and horizon-specific estimators that
relax proportional scaling. We evaluate MinT using these covariance
estimates for both point and probabilistic reconciliation, and
demonstrate their effectiveness and improvements over the shrinkage
estimator in a complex, large-hierarchy dataset.

\vspace{1cm}

\emph{\textbf{Keywords:} Coherent forecasts, Hierarchical time series,
Covariance estimation, Australian tourism, Aggregation.}

\pagebreak

\clearpage
\tableofcontents
\clearpage

\section{Introduction}\label{introduction}

In time series forecasting, aggregation occurs in a variety of settings.
For example, Starbucks Corporation operates in many countries, and each
country has multiple cities where they have outlets. The sales data is
structured \emph{hierarchically}: the top level is the company's total
sales, which disaggregates into sales by country, then sales by city
within each country, and finally down to sales by individual outlet
within each city. As a result, there are over 50,000 sales series across
all \emph{aggregation levels}, and decision makers need forecasts at
each level to manage inventory and plan marketing strategies
effectively. The hierarchy can be even more complex if we consider the
sales of different categories of products (e.g., beverage, food, etc.)
and the sales of product within each category (e.g., latte, cappuccino,
etc.) at each aggregation level. In this case, the structure is called a
\emph{grouped structure}, where the aggregation paths are not unique.
Such structures also arises in many other decision-making contexts, from
supply chains (\citeproc{ref-Angam2025-od}{Angam et al., 2025};
\citeproc{ref-Seaman2022-bb}{Seaman \& Bowman, 2022}) and energy
planning (\citeproc{ref-Di-Modica2021-ad}{Di Modica et al., 2021}), to
macroeconomics (\citeproc{ref-El-Gemayel2022-hx}{El Gemayel et al.,
2022}; \citeproc{ref-Li2019-vv}{Li et al., 2019}) and tourism analysis
(\citeproc{ref-Athanasopoulos2009-lp}{Athanasopoulos et al., 2009}).
Stakeholders in these settings need forecasts at several aggregation
levels to allocate resources and manage risk.

In practice, when forecasts are produced for all series (often called
\emph{base forecasts}), they typically violate the aggregation
constraints observed in the data (e.g., the sum of all countries' sales
forecasts does not equal the total sales forecast). Such forecasts are
called \emph{incoherent}. Incoherence undermines downstream decisions
that require internal consistency and can degrade forecasting
performance. To tackle this problem, forecast reconciliation was
introduced. Forecast reconciliation, a post-processing step, utilises
the information from the hierarchical structure and data to adjust the
initially produced base forecasts, so that the resulting
\emph{reconciled forecasts} are \emph{coherent} (i.e., respecting the
aggregation constraints). It was first introduced by Hyndman et al.
(\citeproc{ref-Hyndman2011-jv}{2011}), and later developed by Erven \&
Cugliari (\citeproc{ref-Van_Erven2015-nx}{2015}), Hyndman et al.
(\citeproc{ref-Hyndman2016-ic}{2016}), Ben Taieb \& Koo
(\citeproc{ref-Ben-Taieb2019-yx}{2019}), Wickramasuriya et al.
(\citeproc{ref-Wickramasuriya2019-xq}{2019}), Wickramasuriya et al.
(\citeproc{ref-Wickramasuriya2020-uk}{2020}), and others that focus
point forecast reconciliation (more accurately, reconciling the means of
predictive distributions). Recognising the important of probabilistic
forecasting, probabilistic forecast reconciliation was studied by Shang
\& Hyndman (\citeproc{ref-Shang2017-ux}{2017}), Jeon et al.
(\citeproc{ref-Jeon2019-wm}{2019}), and Ben Taieb et al.
(\citeproc{ref-Ben-Taieb2021-bn}{2021}), and later formally defined by
Panagiotelis et al. (\citeproc{ref-Panagiotelis2023-fs}{2023}).
Athanasopoulos et al. (\citeproc{ref-Athanasopoulos2024-as}{2024})
provide a comprehensive review of the literature on forecast
reconciliation.

Among the modern methods, the Min Trace (MinT) approach developed by
Wickramasuriya et al. (\citeproc{ref-Wickramasuriya2019-xq}{2019}) is is
widely used due to its strong theoretical properties for minimising
total reconciled forecast error variance, computational efficiency, and
robust empirical performance. MinT is later extended to probabilistic
reconciliation by Wickramasuriya
(\citeproc{ref-Wickramasuriya2024-mb}{2024}), showing that it also
minimises the negative log score of the reconciled distribution under
Gaussian assumptions. Wickramasuriya et al.
(\citeproc{ref-Wickramasuriya2019-xq}{2019}) also argued that modelling
spatial autocorrelations directly from the start would be challenging as
in this case of a large collection of time series. Post-processing
reconciliation has the advantage to implicitly model this spatial
autocorrelation structure, especially true for MinT. MinT is implemented
in popular R and Python software ecosystems
(\citeproc{ref-Nixtla}{Nixtla, 2025};
\citeproc{ref-O-Hara-Wild2024-we}{O'Hara-Wild et al., 2024}).

A central difficulty for MinT is estimating the covariance matrix of
base-forecast errors, particularly beyond one-step forecast horizons.
This is a high-dimensional estimation problem in which the number of
series often exceeds the time dimension. A common practice, following
Wickramasuriya et al. (\citeproc{ref-Wickramasuriya2019-xq}{2019}), is
to estimate the 1-step-ahead covariance from the residuals, using linear
shrinkage toward a diagonal target
(\citeproc{ref-Schafer2005-yw}{Schäfer \& Strimmer, 2005}). Then,
proportionally scale this estimate to approximate the multi-step-ahead
covariance matrix. While convenient and guaranteed to produce a
positive-definite estimate, this practice has three important
shortcomings. First, the shrinkage is uniform across off-diagonals,
applying a single penalty that may over-shrink genuine dependence and
under-shrink noise. Second, many hierarchical data sets exhibit strong
latent low-rank structures that can be explicitly exploited. Third, the
proportionality relationship between the h-step-ahead and 1-step-ahead
forecast error covariance matrices might not hold when error dynamics
change with horizon.

These limitations matter because the theoretical advantages of MinT
depend on the quality of the covariance estimate available in finite
samples. Despite its central role, there has been limited work on
tailored covariance estimation for reconciliation. An exception is the
recent double-shrinkage proposal in Carrara et al.
(\citeproc{ref-Carrara2025-rz}{2025}), which introduces an additional
target designed to encode conditional dependence suggested by the
hierarchy. This line of work still remains at an early stage and does
not fully address the aforementioned limitations.

Meanwhile, there has been substantial progress in high-dimensional
covariance estimation that can be leveraged for MinT. Building on
shrinkage, Huang \& Fryzlewicz (\citeproc{ref-Huang2019-ua}{2019})
proposed NOVELIST, which shrinks toward a soft-thresholded target rather
than a diagonal one, improving flexibility to sparse targets. Extending
beyond linear shrinkage, Ledoit \& Wolf
(\citeproc{ref-Ledoit2012-ga}{2012}) (and further developed by Ledoit \&
Wolf (\citeproc{ref-Ledoit2020-xt}{2020})) introduced nonlinear
shrinkage that replaces sample eigenvalues with data-driven nonlinear
transformation of themselves. In a complementary direction, Fan et al.
(\citeproc{ref-Fan2013-jz}{2013}) introduce factor-based estimators that
explicitly preserve latent low-rank structure while thresholding the
idiosyncratic component. Related thresholding approaches, including hard
thresholding (\citeproc{ref-Bickel2008-nh}{Bickel \& Levina, 2008}),
generalised thresholding (\citeproc{ref-Rothman2009-yf}{Rothman et al.,
2009}), and adaptive thresholding (\citeproc{ref-Cai2011-jf}{Cai \& Liu,
2011}), have also been developed. This growing toolkit provides multiple
pathways to maximise the performance of MinT.

This paper focuses on covariance estimation for MinT. We examine the
shortcomings of the standard shrinkage plus scaling practice and
introduce alternative estimators that address these issues individually
and in combination. Specifically, we study NOVELIST as a more adaptive
target-based shrinkage; principal-component-adjusted (PC-adjusted)
estimators that exploit latent factor structures; and horizon-specific
estimators that relax proportional scaling, including direct estimation
from multi-step residuals. We evaluate MinT under these alternatives for
both point and probabilistic reconciliation. In a large, complex
real-world hierarchy, our findings reveal three main insights. First,
NOVELIST improves probabilistic performance relative to linear
shrinkage, including higher empirical coverage. Second, PC-adjusted
estimators consistently outperform other estimators when common
components are strong, for both point and probabilistic metrics. Third,
proportional scaling of the one-step covariance remains a competitive
baseline, indicating that the assumption may not be universally invalid.

The remainder of the paper is organised as follows.
Section~\ref{sec-pre-theory} sets out the framework for hierarchical
time series, forecast reconciliation, and MinT, and motivates the need
for improved covariance estimation. Section~\ref{sec-cov-est} presents
the covariance estimators considered, outlining their strengths and
limitations in the reconciliation context. Section~\ref{sec-scores}
defines the evaluation metrics for point and probabilistic
reconciliation. Section~\ref{sec-simulation} details the simulation
design and examines the performance of NOVELIST compared to shrinkage.
Section~\ref{sec-empirical} dives into a real-world application with all
proposed methods and highlights behaviours not observed in the
simplified simulations.

\section{Theoretical Framework}\label{sec-pre-theory}

\subsection{Hierarchical Time Series and
Reconciliation}\label{sec-hierarchy}

\emph{Hierarchical time series} are multivariate time series
\(\boldsymbol{y}_t \in \mathbb{R}^n\) organised in a structure where the
series adheres to linear constraints. Figure~\ref{fig-hierarchy}
illustrates a simple two-level hierarchical structure with one top-level
series \(y_{Tot,t}\), disaggregating down to two level-1 series
\((y_{A,t}, y_{B,t})'\), and to four bottom-level series
\((y_{A1,t}, y_{A2,t}, y_{B1,t}, y_{B2,t})'\). Here, the
\emph{aggregation constraints} imply that
\(y_{Tot,t} = y_{A,t} + y_{B,t}\), \(y_{A,t} = y_{A1,t} + y_{A2,t}\),
and \(y_{B,t} = y_{B1,t} + y_{B2,t}\), for all time points \(t\).

\begin{figure}

\centering{

\includegraphics[width=0.4\linewidth,height=\textheight,keepaspectratio]{../figs/hierarchical_structure.png}

}

\caption{\label{fig-hierarchy}A 2-level hierarchical tree structure}

\end{figure}%

The bottom-level (or most disaggregated) series are denoted as
\(\boldsymbol{b}_t \in \mathbb{R}^{n_b}\). Thus, the full vector of all
series in the hierarchy can be represented as:

\[
\boldsymbol{y}_t = \boldsymbol{S} \boldsymbol{b}_t,
\]

where \(\boldsymbol{S} \in \mathbb{R}^{n \times n_b}\) is a summing
matrix that aggregates the bottom-level to all-level series. The summing
matrix \(\boldsymbol{S}\) for the tree structure in
Figure~\ref{fig-hierarchy} is:

\[
\boldsymbol{S} = 
\left[
\begin{array}{cccc}
1 & 1 & 1 & 1 \\
1 & 1 & 0 & 0 \\
0 & 0 & 1 & 1 \\
\multicolumn{4}{c}{\boldsymbol{I_4}}
\end{array}
\right],
\]

where \(\boldsymbol{I_4}\) is the \(4 \times 4\) identity matrix. The
matrix \(\boldsymbol{S}\) encodes the aggregation constraints implied by
the structure. Hence, the columns of \(\boldsymbol{S}\) span a linear
subspace. Any observation \(\boldsymbol{y}_t\) that lies inside this
subspace is called \emph{coherent}, while those outside are
\emph{incoherent}. We refer to the subspace spanned by
\(\boldsymbol{S}\) as the \emph{coherent subspace}
\(\mathfrak{s} \in \mathbb{R}^{n_b}\).

This representation extends beyond hierarchical (nested) structures.
When attributes of interest are crossed, such as the company sales at
any aggregation level (company-wise, city-wise, or outlet-wise) is also
considered by kinds of products, the structure is described as a
\emph{grouped structure}. In grouped systems, as illustrated in
Figure~\ref{fig-grouped}, aggregation and disaggregation paths are not
unique, but the linear constraints can still be written compactly
through a summing matrix \(\boldsymbol{S}\). For simplicity, we refer to
both structures as hierarchical structure and distinguish between them
when needed.

\begin{figure}

\centering{

\includegraphics[width=0.9\linewidth,height=\textheight,keepaspectratio]{../figs/grouped_structure.png}

}

\caption{\label{fig-grouped}A 2-level grouped structure, which can be
considered as the union of two hierarchical trees with common top and
bottom level series}

\end{figure}%

In practice, when we produce \(h\)-step-ahead forecasts for each
individual series, referred to as \emph{base forecasts}
\(\hat{\boldsymbol{y}}_{t+h|t}\), they generally violate the aggregation
constraints, and thus are incoherent. Coherency can be restored by
linearly projecting the base forecasts onto the coherent subspace
\(\mathfrak{s}\) using a projection matrix \(\boldsymbol{P}\):
\(\tilde{\boldsymbol{y}}_{t+h|t} = \boldsymbol{P} \hat{\boldsymbol{y}}_{t+h|t}\),
where \(\tilde{\boldsymbol{y}}_{t+h|t}\) are \(h\)-step-ahead
\emph{reconciled forecasts}.

Many existing reconciliation methods including the OLS from Hyndman et
al. (\citeproc{ref-Hyndman2011-jv}{2011}), WLS from Hyndman et al.
(\citeproc{ref-Hyndman2016-ic}{2016}), and MinT from Wickramasuriya et
al. (\citeproc{ref-Wickramasuriya2019-xq}{2019}) express the projection
matrix as \(\boldsymbol{P} = \boldsymbol{S} \boldsymbol{G}\), for a
suitable \(n_b \times n\) mapping matrix \(\boldsymbol{G}\). The role of
this matrix is to map the base forecasts of all levels
\(\hat{\boldsymbol{y}}_{t+h|t}\) down into the bottom level, which is
then aggregated to the higher levels by premultiplying with
\(\boldsymbol{S}\). Since the projection matrix \(\boldsymbol{P}\) is
idempotent, \(\boldsymbol{G}\) must satisfy the condition
\(\boldsymbol{S} \boldsymbol{G} \boldsymbol{S} = \boldsymbol{S}\).
Within this class, a broad family of mapping matrices is given by:
\(\boldsymbol{G} = (\boldsymbol{S}' \boldsymbol{M}^{-1} \boldsymbol{S})^{-1} \boldsymbol{S}' \boldsymbol{M}^{-1}\),
for some positive definite matrix
\(\boldsymbol{M} \in \mathbb{R}^{n \times n}\)
(\citeproc{ref-Gamakumara2020-ml}{Gamakumara, 2020}).

When setting \(\boldsymbol{M} = \boldsymbol{I}_n\), the identity matrix,
the \(\boldsymbol{P}\) matrix reduces to the OLS reconciliation, which
corresponds to an orthogonal projection onto \(\mathfrak{s}\). For a
general positive definite matrix \(\boldsymbol{M}\), \(\boldsymbol{P}\)
is idempotent but not necessarily symmetric, resulting in an oblique
projection. Thus, the choice of \(\boldsymbol{M}\) determines the
projection direction and, in turn, how disagreements among levels are
resolved.

\begin{figure}

\centering{

\includegraphics[width=0.6\linewidth,height=\textheight,keepaspectratio]{../figs/geometry_orthogonal.png}

}

\caption{\label{fig-geometry}Geometry of probabilistic forecast
reconciliation. The base forecast distribution is projected orthogonally
onto the coherent subspace (purple line), resulting in the reconciled
forecast distribution (red). The projection is defined by the projection
matrix \(\boldsymbol{P}\). Note that this figure is a schematic since
most applications are high-dimensional.}

\end{figure}%

This projection view extends seamlessly from point to probabilistic
forecast reconciliation. Let \(\hat{\boldsymbol{Y}}_{t+h|t}\) denote the
random vector representing the \(h\)-step-ahead base forecasts. The
random reconciled forecast vector is obtained similarly via projection:
\(\tilde{\boldsymbol{Y}}_{t+h|t} = \boldsymbol{P} \hat{\boldsymbol{Y}}_{t+h|t}\).
Consequently, point forecasts (which we refer to as means of the
predictive distributions) reconcile as
\(\mathbb{E}(\tilde{\boldsymbol{Y}}_{t+h|t}) = \mathbb{E}(\boldsymbol{P} \hat{\boldsymbol{Y}}_{t+h|t}) = \boldsymbol{P} \hat{\boldsymbol{y}}_{t+h|t} = \tilde{\boldsymbol{y}}_{t+h|t}\),
and the covariance of the predictive distribution transforms linearly as
\(\text{Var}(\tilde{\boldsymbol{Y}}_{t+h|t}) = \text{Var}(\boldsymbol{P} \hat{\boldsymbol{Y}}_{t+h|t}) = \boldsymbol{P} \boldsymbol{W}_h \boldsymbol{P}'\),
where \(\boldsymbol{W}_h = \text{Var}(\hat{\boldsymbol{Y}}_{t+h|t})\).

Figure~\ref{fig-geometry} schematically illustrates this projection.
Reconciliation takes the (possibly elliptical) base forecast
distribution and ``pushes'' the entire mass linearly onto the coherent
subspace, resulting in the reconciled forecast distribution.

\subsection{The Minimum Trace
Reconciliation}\label{the-minimum-trace-reconciliation}

Wickramasuriya et al. (\citeproc{ref-Wickramasuriya2019-xq}{2019})
showed that by setting
\(\boldsymbol{M} = \boldsymbol{W}_h = \mathbb{E}( \hat{\boldsymbol{e}}_{t+h|t} \; \hat{\boldsymbol{e}}_{t+h|t}' )\),
the covariance matrix of the \(h\)-step-ahead base forecast errors
\(\hat{\boldsymbol{e}}_{t+h|t} = \boldsymbol{y}_{t+h} - \hat{\boldsymbol{y}}_{t+h|t}\),
we essentially minimise the total variance of the reconciled forecast
errors across all series. Equivalently, MinT is the unique
linear-unbiased reconciler that minimises the trace of
\(\text{Var}[\boldsymbol{y}_{t+h} - \tilde{\boldsymbol{y}}_{t+h|t}] = \boldsymbol{S} \boldsymbol{G}_h \boldsymbol{W}_h \boldsymbol{G}_h' \boldsymbol{S}'\).
This method is thus called Minimum Trace (MinT) reconciliation. The
matrix \(\boldsymbol{G}_h\) is thus given by:

\[
\boldsymbol{G}_h = (\boldsymbol{S}' \boldsymbol{W}_h^{-1} \boldsymbol{S})^{-1}
\boldsymbol{S}' \boldsymbol{W}_h^{-1},
\]

provided that \(\boldsymbol{W}_h\) is positive definite.

Although originally developed for point reconciliation, Wickramasuriya
(\citeproc{ref-Wickramasuriya2024-mb}{2024}) showed that MinT extends
naturally to the probabilistic setting when base predictive
distributions are Gaussian. If \(h\)-step-ahead base forecast
distribution is
\(\mathcal{N}(\hat{\boldsymbol{y}}_{t+h|t}, \boldsymbol{W}_h)\), then
reconciliation gives
\(\tilde{\boldsymbol{Y}}_{t+h|t} = \boldsymbol{S} \boldsymbol{G}_h \hat{\boldsymbol{Y}}_{t+h|t} \sim \mathcal{N}(\tilde{\boldsymbol{y}}_{t+h|t}, \boldsymbol{S} \boldsymbol{G}_h \boldsymbol{W}_h \boldsymbol{G}_h' \boldsymbol{S}')\).
Within the class of linear-unbiased reconcilers, MinT also minimises the
negative log score (equivalently, the Gaussian log predictive score) of
the reconciled distribution.

In this paper we adopt the Gaussian framework to compare the impacts of
different covariance estimators for MinT in both point and probabilistic
reconciliation. It is also worth to mention that the methods can be
extended to non-Gaussian settings by bootstrapping
(\citeproc{ref-Gamakumara2020-ml}{Gamakumara, 2020};
\citeproc{ref-Panagiotelis2023-fs}{Panagiotelis et al., 2023}), which is
beyond the scope of this paper.

\subsection{Shrinkage Estimator for
MinT}\label{shrinkage-estimator-for-mint}

The performance of MinT hinges on a reliable, positive-definite estimate
of \(\boldsymbol{W}_h\), which comes in both the mapping matrix
\(\boldsymbol{G}_h\) and the reconciled forecast error variance
\(\boldsymbol{S} \boldsymbol{G}_h \boldsymbol{W}_h \boldsymbol{G}_h' \boldsymbol{S}'\).

However, the covariance matrix \(\boldsymbol{W}_h\) is often not
available in closed-form, and is challenging to estimate in
high-dimensional setting where the number of series \(n\) is larger than
the time dimension \(T\). To tackle this issue, the original paper by
Wickramasuriya et al. (\citeproc{ref-Wickramasuriya2019-xq}{2019})
assumed a proportionality relationship
\(\hat{\boldsymbol{W}}^g_h = k_h g( \hat{\boldsymbol{W}}_1 )\), where
\(\hat{\boldsymbol{W}}_1\) is the sample covariance matrix of the
in-sample 1-step-ahead base forecast errors (to approximate
\(\boldsymbol{W}_1\)) and \(k_h > 0\) is a scaling constant (which will
be algebraically cancelled out in point-forecast reconciliation). The
function \(g(.)\) is a covariance estimator that produces a
positive-definite matrix, the main focus of this paper.

The recommended choice for \(g(.)\) in the original work is the
shrinkage estimator with diagonal target from Schäfer \& Strimmer
(\citeproc{ref-Schafer2005-yw}{2005}):

\[
\hat{\boldsymbol{W}}^{S}_{1} = \lambda_S \, \boldsymbol{D}_1 + (1 - \lambda_S) \hat{\boldsymbol{W}}_1 \, ,
\]

where \(\boldsymbol{D}_1 = \text{diag}( \hat{\boldsymbol{W}}_1 )\) is
the diagonal matrix comprising the diagonal elements of
\(\hat{\boldsymbol{W}}_1\) (i.e., the sample variances). We refer to any
\(\lambda_S \in [0,1]\) as the shrinkage intensity of the shrinkage
estimator. This approach shrinks the covariance matrix
\(\hat{\boldsymbol{W}}_1\) towards the diagonal target
\(\text{diag}( \hat{\boldsymbol{W}}_1 )\), meaning the off-diagonal
elements are shrunk towards zero while the diagonal ones remain
unchanged.

Schäfer \& Strimmer (\citeproc{ref-Schafer2005-yw}{2005}) also provided
a closed-form estimate of the optimal shrinkage intensity parameter
\(\lambda_S\):

\[
\hat{\lambda}_S = \frac{\sum_{i \neq j} \widehat{Var}(\hat{r}_{ij})} 
{\sum_{i \neq j} \hat{r}_{ij}^2} \, ,
\]

where \(\hat{r}_{ij}\) is the \(i,j\)-th element of
\(\hat{\boldsymbol{R}}_1\), the 1-step-ahead sample correlation matrix
(obtained from \(\hat{\boldsymbol{W}}_1\)). The optimal estimate is
obtained by minimising
\(MSE(\hat{\boldsymbol{W}}_1) = Bias(\hat{\boldsymbol{W}}_1)^2 + Var(\hat{\boldsymbol{W}}_1)\).
More specifically, we trade a small bias for a substantial variance
reduction, which is especially valuable in high dimension.

Despite its simplicity and guaranteed positive-definiteness, MinT
coupled with diagonal-target shrinkage presents three important
limitations.

\subsubsection*{Problem 1: Uniform shrinkage}\label{prob1}
\addcontentsline{toc}{subsubsection}{Problem 1: Uniform shrinkage}

Linear shrinkage shrinks all off-diagonal elements of
\(\hat{\boldsymbol{W}}_1\) towards zeros with equal weights
\(1 - \lambda_S\). The resulting penalty is global and non-adaptive:
strong, genuinely systematic correlations are shrunk at the same rate as
weak, noisy ones. In hierarchical and grouped systems, this can be
problematic. Aggregation naturally induces stronger dependence among
related nodes (for example, a region and a neighbouring region, or a
region and its parent state), while many unrelated pairs exhibit
near-zero correlations. A uniform penalty may over-shrink informative
co-movements or under-shrink idiosyncratic noise, reducing
reconciliation efficiency.

\subsubsection*{Problem 2: Latent factors}\label{prob2}
\addcontentsline{toc}{subsubsection}{Problem 2: Latent factors}

Many real-world hierarchical data sets may exhibit a prominent low-rank
(factor) structure. A clear example is the Australian domestic overnight
trips (\citeproc{ref-tourism-au}{\emph{Tourism Research Australia},
2024}), where the national trips are disaggregated into states and
territories, and further into regions. The tourism activities might be
driven by a few common factors, such as economic conditions, fuel
prices, or major events, which might be left in the forecast errors
after fitting the base models. Figure~\ref{fig-eigen} illustrates the
largest eigenvalues from the one-step forecast error correlation matrix
\(\hat{\boldsymbol{R}}_1\), typically showing a marked elbow after a few
principal components (largest eigenvalues), indicating strong latent
structure. Diagonal-target shrinkage is factor-unaware: it neither
preserves the low-rank common subspace nor differentially shrinks the
idiosyncratic remainder, and so can be inefficient when common factors
dominate.

\begin{figure}

\centering{

\includegraphics[width=0.7\linewidth,height=\textheight,keepaspectratio]{paper_draft_files/figure-pdf/fig-eigen-1.pdf}

}

\caption{\label{fig-eigen}Scree plot of the eigenvalues of the
1-step-ahead forecast error covariance matrix from the Australian
domestic tourism dataset. The point of inflection occurs around the
second largest eigenvalue.}

\end{figure}%

\subsubsection*{Problem 3: Proportional scaling}\label{prob3}
\addcontentsline{toc}{subsubsection}{Problem 3: Proportional scaling}

The proportionality relationship
\(\hat{\boldsymbol{W}}^g_h = k_h g( \hat{\boldsymbol{W}}_1 )\) enforces
horizon-invariant cross-series dependence of the forecast errors up to a
scalar factor. In many applications, the structure of the error
covariance might change with horizon due to evolving error dynamics or
horizon-specific interactions among series. Proportional scaling may
therefore misrepresent multi-step dependence.

Additionally, while the scalar factor cancels in point reconciliation,
it directly controls dispersion in probabilistic reconciliation and can
lead to miscalibrated predictive distributions. However, we will not
explore this issue in this paper, and leave it for future research.

In the next sections, we introduce alternative covariance estimators
that address these limitations individually and in combination.

\section{Covariance Estimation Approaches}\label{sec-cov-est}

\subsection{NOVELIST Estimator}\label{novelist-estimator}

To tackle the Problem \hyperref[prob1]{1} of uniform shrinkage, we
consider the NOVELIST estimator from Huang \& Fryzlewicz
(\citeproc{ref-Huang2019-ua}{2019}). The NOVELIST (NOVEL Integration of
the Sample and Thresholded Covariance) estimator introduces extra
parameter, the threshold \(\delta\), to control the sparsity level in
the target toward which the sample covariance matrix is shrunk. By
soft-thresholding small correlations before shrinkage, NOVELIST
preserves strong signals while attenuating weak, noisy ones, addressing
the uniform, non-adaptive nature of diagonal-target shrinkage.

The construction proceeds in two steps. First, apply elementwise
soft-thresholding to the sample correlation matrix; second, shrink the
sample correlation toward this thresholded target. The NOVELIST
estimator for covariance matrix is given by:

\[
\hat{\boldsymbol{W}}^{N}_{1} = \lambda_{\delta} \hat{\boldsymbol{W}}_{1, \delta} + (1 - \lambda_{\delta}) \hat{\boldsymbol{W}}_1,
\]

where \(\hat{\boldsymbol{W}}_{1, \delta}\) is the thresholded covariance
matrix of \(\hat{\boldsymbol{W}}_{1}\). As working with correlations
avoids rescaling issues and keeps diagonal entries at one, we rewrite it
in terms of correlations:

\begin{equation}\phantomsection\label{eq-novelist-cor}{
\hat{\boldsymbol{R}}^{N}_{1} = \lambda_{\delta} \hat{\boldsymbol{R}}_{1,\delta} + (1 - \lambda_{\delta}) \hat{\boldsymbol{R}}_1,
}\end{equation}

where, \(\hat{\boldsymbol{R}}_{1,\delta}\) is the thresholded
correlation matrix, where each element is regularised by:

\begin{equation}\phantomsection\label{eq-soft-thr}{
\hat{r}_{1,ij}^\delta = \text{sign}(\hat{r}_{1,ij}) \, 
\text{max}(|\hat{r}_{1,ij}| - \delta, \; 0),
}\end{equation}

where \(\delta \in [0,1]\). For a given threshold \(\delta\), Huang \&
Fryzlewicz (\citeproc{ref-Huang2019-ua}{2019}) derived an analytical
expression for the optimal shrinkage intensity parameter
\(\lambda(\delta)\), following similar logic to Schäfer \& Strimmer
(\citeproc{ref-Schafer2005-yw}{2005}). It can be computed as:

\begin{equation}\phantomsection\label{eq-lambda-thr}{
\hat{\lambda}(\delta) = \frac{
  \sum_{i \neq j} \widehat{Var}(\hat{r}_{1,ij}) \; \boldsymbol{1}(|\hat{r}_{1,ij}| \leq \delta)
} {
  \sum_{i \neq j} (\hat{r}_{1,ij} - \hat{r}_{1,ij}^\delta)^2
},
}\end{equation}

where \(\boldsymbol{1}(.)\) is the indicator function.

On the other hand, the optimal threshold \(\hat \delta\) does not have a
closed-form solution, and is obtained by rolling-window cross-validation
procedure. The idea is to find the threshold \(\delta^*\), with the
corresponding \(\lambda^*\) and
\(\hat{\boldsymbol{R}}^{N}_{1}(\delta^*, \lambda^*)\), that minimises
the average out-of-sample 1-step-ahead mean squared reconciled forecast
errors over all windows. The formal algorithm is given in Algorithm
\ref{alg:novelist_cv}.

Note that when
\(\delta \in \bigl[ \text{max}_{i \neq j}|\hat{r}_{1,ij}|, \; 1 \bigr]\),
the NOVELIST estimator collapses to the shrinkage estimator, and when
\(\delta = 0\), it becomes the sample covariance matrix. An additional
concern is that the estimator is not guaranteed to be positive definite,
but we can use Higham (\citeproc{ref-Higham2002-te}{2002}) algorithm to
compute the nearest positive definite matrix if needed.

\begin{algorithm}[H]
\caption{Cross-validation procedure}\label{alg:novelist_cv}
\begin{algorithmic}[1]

\State \textbf{Input:} Observations and fitted values $\boldsymbol{y}_t, \hat{\boldsymbol{y}}_t \in \mathbb{R}^n$ for $t = 1,\dots,T$ \footnotemark, set of threshold candidates $\Delta$, window size $v$.

\State $\hat{\boldsymbol{e}}_t = \boldsymbol{y}_t - \hat{\boldsymbol{y}}_t$ for $t = 1,\dots,T$

\For{$i = v:T-1$}
    \State $j = i - v +1$
    \State $\hat{\boldsymbol{W}}_1 = \frac{1}{v} \sum_{t=j}^{i} \hat{\boldsymbol{e}}_{t} \hat{\boldsymbol{e}}_{t}'$
    \State $\hat{\boldsymbol{D}} = \text{diag}(\hat{\boldsymbol{W}}_1)$
    \State $\hat{\boldsymbol{R}}_1 = \hat{\boldsymbol{D}}^{-1/2} \hat{\boldsymbol{W}}_1 \hat{\boldsymbol{D}}^{-1/2}$
    \For{$\delta \in \Delta$}
        \State Compute thresholded correlation $\hat{\boldsymbol{R}}_{1,\delta}$ using Equation 2
        \State Compute shrinkage parameter $\hat{\lambda}_{\delta}$ using Equation 3
        \State Compute $\hat{\boldsymbol{R}}^{N}_{1,\delta}$ using Equation 1
        \State $\hat{\boldsymbol{W}}^{N}_{1,\delta} = \hat{\boldsymbol{D}}^{1/2} \hat{\boldsymbol{R}}^{N}_{1,\delta} \hat{\boldsymbol{D}}^{1/2}$
        \State $\boldsymbol{G} = (\boldsymbol{S}' \hat{\boldsymbol{W}}^{N^{-1}}_{1,\delta} \boldsymbol{S})^{-1} \boldsymbol{S}' \hat{\boldsymbol{W}}^{N^{-1}}_{1,\delta}$
        \State Reconciled forecasts $\tilde{\boldsymbol{y}}_{i+1} = \boldsymbol{S} \boldsymbol{G} \hat{\boldsymbol{y}}_{i+1}$
        \State $\tilde{\boldsymbol{e}}_{i+1, \delta} = \boldsymbol{y}_{i+1} - \tilde{\boldsymbol{y}}_{i+1}$
    \EndFor
\EndFor

\State $\text{MSE}_{\delta} = \frac{1}{T-v} \sum_{i=v}^{T-1} (\tilde{\boldsymbol{e}}_{i+1, \delta})^2$ for each $\delta \in \Delta$
\State $\hat{\delta} = \arg\min_{\delta \in \Delta} \text{MSE}_{\delta}$

\State \textbf{Output:} Estimate of optimal threshold parameter $\hat{\delta}$

\end{algorithmic}
\end{algorithm}

\footnotetext{It is only required to fit the base models once on the whole training data $\{\boldsymbol{y}_t\}_{t=1}^T$, and obtain the in-sample fitted values $\{\hat{\boldsymbol{y}}_t\}_{t=1}^T$.}

\vspace{-1em}

\emph{Remark.} Minimising a multi-step-ahead forecast error metric in
the cross-validation procedure oftern yields a \(\hat{\delta}\) that is
close to the 1-step-ahead case.

\subsection{PC-adjusted Estimator}\label{pc-adjusted-estimator}

To address Problem \hyperref[prob2]{2} and exploit latent common
components explicitly, the PC-adjusted (Principal-Component-adjusted)
method takes the latent factors directly into its construction. It
starts by decomposing the covariance matrix \(\hat{\boldsymbol{W}}_1\)
into a prominent principle components part (low-rank) and a orthogonal
complement part \(\hat{\Upsilon}\) (the covariance matrix after removing
the first \(K\) principal components). Then we can apply either
shrinkage or NOVELIST estimator to \(\hat{\boldsymbol{W}}^K_{1}\):

\[
\hat{\boldsymbol{W}}^{g, K}_{1} = \sum_{k = 1}^K \hat{\gamma}_k \hat{\boldsymbol{\xi}}_k \hat{\boldsymbol{\xi}}_k' + g(\hat{\Upsilon})
\]

where \(g(.)\) is either the shrinkage or NOVELIST estimator,
\(\hat{\gamma}_k\) and \(\hat{\boldsymbol{\xi}}_k\) are the \(k\)-th
largest eigenvalue and the corresponding eigenvector of the sample
covariance matrix, respectively.

Similar to the NOVELIST estimator, its PC-adjusted variant
\(\hat{\boldsymbol{W}}^{N, K}_{1}\) requires a cross-validation
procedure to select the threshold paramter and adjustment to obtain
positive definiteness. The reconciled forecasts for each
cross-validation window are obtained using MinT with
\(\hat{\boldsymbol{W}}^{N, K}_{1}\).

\subsection{Scaled Variance}\label{scaled-variance}

To address the potential issue of Problem \hyperref[prob3]{3} of
proportionality relationship
\(\hat{\boldsymbol{W}}^g_h = k_h g( \hat{\boldsymbol{W}}_1 )\) not
holding in practice. A simple relaxation retains the \(1\)-step-ahead
correlation structure but allows the variances to scale differently with
horizon. The scaled variance estimator is given by:

\$\$ \hat{\boldsymbol{W}}\^{}\{g, sv\}\_\{h\} =
\boldsymbol{D}\^{}\{1/2\}\_h \hat{\boldsymbol{R}}\^{}g\_1
\boldsymbol{D}\^{}\{1/2\}\_h, \quad

\hat{\boldsymbol{R}}\^{}g\_1 = D\^{}\{-1/2\}\_1 ,
g(\hat{\boldsymbol{W}}\_1) , D\^{}\{-1/2\}\_1, \$\$

where \(\boldsymbol{D}_h\) is a diagonal matrix comprising
\(\hat{\sigma}^2_{1,h} , \ldots, \hat{\sigma}^2_{n,h}\) on its diagonal,
and \(\hat{\sigma}^2_{i,h}\) is \(h\)-step-ahead base forecast error
variance of the \(i\)-th series. \(\hat{\boldsymbol{R}}^g_1\) is the
estimated one-step correlation matrix using either shrinkage or NOVELIST
estimator.

When NOVELIST is used to produce \(h\)-step-ahead reconciled forecasts,
the cross-validation procedure is modified to evaluate the out-of-sample
\(h\)-step-ahead reconciled forecasts for each window and estimate
optimal threshold parameter (instead of evaluating \(1\)-step-ahead
reconciled forecasts). The reconciled forecasts are obtained using MinT
with \(\hat{\boldsymbol{W}}^{N, sv}_{h}\).

\subsection{Constructing from h-step-ahead
Residuals}\label{constructing-from-h-step-ahead-residuals}

Another alternative is to directly estimate the covariance matrix from
the \(h\)-step-ahead base forecast errors, without assuming any
proportionality relationship:

\[
\hat{\boldsymbol{W}}^{g}_{h} = g(\hat{\boldsymbol{W}}_h),
\]

where \(\hat{\boldsymbol{W}}_h\) is the covariance matrix of in-sample
h-step-ahead base forecast errors, and \(g(.)\) is either the shrinkage
or NOVELIST estimator.

Similarly, if NOVELIST is used, the cross-validation procedure is
modified to evaluate the out-of-sample \(h\)-step-ahead reconciled
forecasts for each window. The reconciled forecasts are obtained using
MinT with \(\hat{\boldsymbol{W}}^{N}_{h}\).

\subsection*{Summary of MinT with Covariance
Estimators}\label{summary-of-mint-with-covariance-estimators}
\addcontentsline{toc}{subsection}{Summary of MinT with Covariance
Estimators}

The covariance estimators explored in this paper are summarised in the
table below. The abbreviations will be used in the following sections.

\begin{longtable}[]{@{}ll@{}}
\caption{Summary of covariance estimators for MinT
reconciliation}\label{tbl-cov}\tabularnewline
\toprule\noalign{}
Covariance estimators used & Abbreviation \\
\midrule\noalign{}
\endfirsthead
\toprule\noalign{}
Covariance estimators used & Abbreviation \\
\midrule\noalign{}
\endhead
\bottomrule\noalign{}
\endlastfoot
Shrinkage & MinT-S \\
NOVELIST & MinT-N \\
PC-adjusted Shrinkage with \emph{K} PCs & MinT-S(PC\emph{K}) \\
PC-adjusted NOVELIST with \emph{K} PCs & MinT-N(PC\emph{K}) \\
Scaled Variance Shrinkage & MinT-S(sv) \\
Scaled Variance NOVELIST & MinT-N(sv) \\
Constructed from h-step-ahead Shrinkage & MinT-S(hcov) \\
Constructed from h-step-ahead NOVELIST & MinT-N(hcov) \\
\end{longtable}

\section{Evaluation of Point and Probabilistic
Forecasts}\label{sec-scores}

This section briefly introduces the scoring rules used to evaluate the
point and probabilistic forecast accuracy in
Section~\ref{sec-simulation} and Section~\ref{sec-empirical}.

For point forecasts, we use the mean squared error (MSE) to evaluate the
accuracy of different reconciliation methods:
\(MSE = \frac{1}{n} \sum_{i=1}^n (y_{i, t+h} - \tilde{y}_{i,t+h|t})^2\),
where \(y_{i, t+h}\) is the realised value of series \(i\) at time
\(t+h\), and \(\tilde{y}_{i,t+h|t}\) is the reconciled point forecast.

Meanwhile, to assess the quality of the probabilistic forecasts, it is
common to use proper scoring rules. A scoring rule is a function
\(S(., .)\) taking a predictive distribution as its first argument and a
realisation as its second argument, then returns a numerical score. We
follow a convention that lower scores are better. A scoring rule is said
to be \emph{proper} if
\(\mathbb{E}_Q[S(Q, y)] \leq \mathbb{E}_Q[S(F, y)]\) for all \(F\),
where \(F\) is a predictive distribution produced by forecasting model,
\(Q\) is the true distribution of the realisation \(y\), and
\(\mathbb{E}_Q\) is the expectation with respect to \(Q\). Hence, the
expected score is minimised when the forecast distribution matches the
true distribution.

We employ Winkler score and continuous ranked probability score as our
univariate scoring rules, and energy score as the multivariate scoring
rule. All three are proper scoring rules. In this paper, we only
evaluate \(1\)-step-ahead probabilistic forecasts, and thus we drop the
subscript \(t\) and \(h\) for simplicity.

\textcolor{blue}{\textit{Winkler score (WS).}} If the
100\((1-\alpha)\)\% prediction interval of \(i\)-th series is
\([l_i, u_i]\) (the \(\alpha / 2\) and \(1 - \alpha / 2\) quantiles),
then the Winkler score is defined as:

\[
WS_{\alpha}(l_i, u_i; y_i) = (u_i - l_i) + \frac{2}{\alpha} (l_i - y_i) \boldsymbol{1}(y_i < l_i) + \frac{2}{\alpha} (y_i - u_i) \boldsymbol{1}(y_i > u_i),
\]

where \(y_i\) is the observed value of \(i\)-th series, and
\(\boldsymbol{1}(.)\) is the indicator function. The Winkler score
rewards narrow intervals that contain the observation, and penalises
intervals that do not contain the observation.

\textcolor{blue}{\textit{Continuous ranked probability score (CRPS).}}
The CRPS is defined as the squared difference between the predictive
cumulative distribution function (CDF) \(F_i\) and the empirical CDF of
the observation \(y_i\) of series \(i\):

\[
CRPS(F_i, y_i) = \int_{-\infty}^{\infty} (F_i(x) - \boldsymbol{1}(x \geq y_i))^2 dx.
\]

When the predictive distribution is Gaussian with mean \(\mu_i\) and
standard deviation \(\sigma_i\), the CRPS has a closed-form expression:

\[
CRPS(F_i, y_i) = \sigma_i \left[ z_i \left( 2 \Phi(z_i) -1 \right) + 2 \phi(z_i) - \frac{1}{\sqrt{\pi}} \right],
\]

where \(z_i = \frac{y_i - \mu_i}{\sigma_i}\), and \(\Phi(.)\) and
\(\phi(.)\) are the CDF and probability density function (PDF) of a
standard normal distribution, respectively.

\textcolor{blue}{\textit{Energy score (ES).}} The energy score is a
multivariate generalisation of the CRPS. It is defined as:

\[
ES(F, \boldsymbol{y}) = \mathbb{E}_F ||\boldsymbol{X} - \boldsymbol{y}||^{\beta} - \frac{1}{2} \mathbb{E}_F ||\boldsymbol{X} - \boldsymbol{X}'||^{\beta},
\]

where \(\boldsymbol{X}\) and \(\boldsymbol{X}'\) are independent random
vectors with multivariate distribution \(F\), \(\boldsymbol{y}\) is the
observed vector, \(||.||\) is the Euclidean norm, and
\(\beta \in (0, 2]\). We set \(\beta = 1\) following common convention.

Since the closed-form expression of the ES may not available, we
approximate it using Monte Carlo samples
\(\{\boldsymbol{x}_1, \ldots, \boldsymbol{x}_M\}\) drawn from \(P\):

\[
\widehat{ES}(F, \boldsymbol{y}) = \frac{1}{M} \sum_{m=1}^M ||\boldsymbol{x}_m - \boldsymbol{y}|| - \frac{1}{2M(M-1)} \sum_{m=1}^M ||\boldsymbol{x}_m - \boldsymbol{x}^*_m||,
\]

where \(\boldsymbol{x}^*_m\) is a randomly selected sample from
\(\{\boldsymbol{x}_1, \ldots, \boldsymbol{x}_M\} \setminus \{\boldsymbol{x}_m\}\).
In our experiments, we use \(M=10000\) samples to approximate the ES.

\section{Simulation}\label{sec-simulation}

\subsection{General Design}\label{general-design}

The general design of data generating process for bottom-level series is
a stationary VAR(1) process, with the following structure:

\[
\boldsymbol{b}_t = \boldsymbol{A} \boldsymbol{b}_{t-1} + \boldsymbol{\epsilon}_t,
\]

where \(\boldsymbol{A}\) is a \(n_b \times n_b\) block diagonal matrix
of autoregressive coefficients
\(\boldsymbol{A} = diag(\boldsymbol{A}_1, \ldots, \boldsymbol{A}_M)\),
with each \(\boldsymbol{A}_m\) being a \(n_{b,m} \times n_{b,m}\)
matrix. The block diagonal structure ensures that the time series are
grouped into \(M\) groups (blocks), with each group having its own
autoregressive coefficients. This aims to simulate the interdependencies
between the time series within each group, where reconciliation will be
expected to better capture these compared to independent base forecasts.

\(\boldsymbol{\epsilon}_t\) is a Gaussian innovation process, with
covariance matrix \(\boldsymbol{\Sigma}\). The covariance matrix
\(\boldsymbol{\Sigma}\) is generated specifically using the Algorithm 1
in Hardin et al. (\citeproc{ref-Hardin2013-wu}{2013}):

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\item
  A compound symmetric correlation matrix is used for each block of size
  \(n_{b,m}\) in \(\boldsymbol{A}_m\), where the correlation entries
  \(\rho_j\) for each block \(m\) are sampled from a uniform
  distribution between 0 and 1. They are within group baseline
  correlations.
\item
  A constant correlation, which is smaller than
  \(\text{min} \{\rho_1, \rho_2, \dots, \rho_M \}\), is imposed on the
  entries between different blocks. It serves as between group baseline
  correlations.
\item
  We select a constant \(\varepsilon\) such that
  \(0 \le \varepsilon < 1 - \text{max} \{\rho_1, \rho_2, \dots, \rho_M \}\)
  to control the noise intensity. We then generate \(n_b\) unit vectors
  \(\boldsymbol{u}_1, \dots, \boldsymbol{u}_{n_b}\) and obtain a
  entry-wise noise element as
  \(\varepsilon \boldsymbol{u}_i' \boldsymbol{u}_j\) for the \(i,j\)-th
  entry. These entry-wise noise elements are added on top of the
  baseline correlation matrix (except the diagonal entries), creating a
  noisy positive correlation matrix \(\boldsymbol{R}^+\).
\item
  The covariance matrix \(\boldsymbol{\Sigma}^+\) is then constructed by
  \(\boldsymbol{\Sigma}^+ = \boldsymbol{D} \boldsymbol{R}^+ \boldsymbol{D}\),
  where \(\boldsymbol{D} = diag(\sigma_1, \dots, \sigma_{n_b})\) is a
  diagonal matrix of standard deviations. Each \(\sigma_i\) is uniformly
  sampled from the range of \([\sqrt{2}, \sqrt{6}]\), for all \(n_b\)
  series.
\end{enumerate}

These steps result in a positive covariance matrix
\(\boldsymbol{\Sigma}^+\) where all the elements are positive. To allow
a mixture of positive and negative covariance, we randomly flip the
signs of them but need to maintain the positive definiteness. This can
be done by pre- and post-multiplying \(\boldsymbol{\Sigma}\) by a random
diagonal matrix \(\boldsymbol{V}\) with diagonal entries sampled from
\(\{-1, 1\}\) with equal probability. The final covariance matrix is
given by
\(\boldsymbol{\Sigma} = \boldsymbol{V} \boldsymbol{\Sigma}^+ \boldsymbol{V}\).

For all hierarchies in our experiments, we simulate two panel lengths,
\(T=54\) (``short'') and \(T=304\) (``long''), reserving the final four
observations as an out-of-sample test set. We perform 500 Monte Carlo
replications for each configuration. In each replication, we fit
univariate ARIMA models to the training observations using the automatic
AICc minimization algorithm of Hyndman \& Khandakar
(\citeproc{ref-Hyndman2008-rf}{2008}), implemented in the
\emph{fabletools} package (\citeproc{ref-O-Hara-Wild2024-we}{O'Hara-Wild
et al., 2024}), generating 1--4-step base forecasts. We then reconcile
the base forecasts using different reconciliation methods, and evaluate
their point and probabilistic forecast accuracy on the test set.

\subsection{Exploring Effects of Hierarchy's
Size}\label{sec-sim-hierarchy-size}

In our main experiments, we examine how MinT combined with the different
estimators perform as the hierarchy expands. We generate synthetic data
from the VAR(1) framework described earlier, varying the number of
bottom‐level series, \(n_b\), across two structures: a ``small''
structure with six groups of six bottom-level series
(\(n_b = 6\)x\(6 = 36\)), and a ``large'' configuration with two groups
of fifty bottom-level series (\(n_b = 2\)x\(50 = 100\)).

In the 36‐series case, each block of six forms a level‐1 aggregate, and
those six aggregates form the total. The 100‐series design employs a
deliberately intricate aggregation path to stress‐test reconciliation
methods. We first sum the one hundred bottom series into ten
intermediate series by grouping them in contiguous blocks of ten. These
ten series are then organised into three level-2 aggregates---four,
three, and four series, respectively---before finally summing to a
single top node. This asymmetric hierarchy creates overlapping
correlation patterns: some level‐2 series share bottom‐level groups,
while others draw from both, emulating practical scenarios such as
regional sales aggregations that span multiple product categories or
overlapping territories. The aggregation paths for both structures are
illustrated in Figure~\ref{fig-sim-structures}.

\begin{figure}

\centering{

\pandocbounded{\includegraphics[keepaspectratio]{D:/Github/Recon_Honours_Thesis/figs/sim_structures.png}}

}

\caption{\label{fig-sim-structures}Aggregation structures used in the
simulation experiments: 6 groups of 6 (left) and 2 groups of 50 (right)}

\end{figure}%

The VAR(1) and correlation configurations for the 6 by 6 case and 2 by
50 case are illustrated in Figure~\ref{fig-settings-6x6} and
Figure~\ref{fig-settings-2x50}, respectively. The block diagonal
structure of the VAR(1) coefficient matrices \(\boldsymbol{A}\) reflects
the grouping of series, and the correlation matrices- show higher
correlations among series within the same group.

\begin{figure}

\centering{

\includegraphics[width=0.75\linewidth,height=\textheight,keepaspectratio]{paper_draft_files/figure-pdf/fig-settings-6x6-1.pdf}

}

\caption{\label{fig-settings-6x6}The VAR(1) coefficients matrix (left)
and correlation matrix of the innovation process (right) for the 6
groups of 6 structure.}

\end{figure}%

\begin{figure}

\centering{

\includegraphics[width=1\linewidth,height=\textheight,keepaspectratio]{paper_draft_files/figure-pdf/fig-settings-2x50-1.pdf}

}

\caption{\label{fig-settings-2x50}The VAR(1) coefficients matrix (top)
and correlation matrix of the innovation process (bottom) for the 2
groups of 50 structure.}

\end{figure}%

\begin{figure}

\centering{

\includegraphics[width=1\linewidth,height=\textheight,keepaspectratio]{paper_draft_files/figure-pdf/fig-sim-results-1-1.pdf}

}

\caption{\label{fig-sim-results-1}Percentage relative improvement in MSE
of reconciled forecasts over the base forecasts in the 6 by 6 case (top
row) and the 2 by 50 case (bottom row), T=50 (left column) and T=300
(right column), for 1- to 4-step-ahead forecasts. The positive
(negative) entries indicate a decrease (increase) in MSE relative to
base. The negative improvements beyond 6\% are capped.}

\end{figure}%

Figure Figure~\ref{fig-sim-results-1} illustrates the relative
improvements in mean squared error (MSE) of reconciled forecasts over
the incoherent base forecasts, across the two structures (small and
large) and time series lengths (short and long). We evaluate ten
different MinT variants based on the covariance estimators discussed in
Table~\ref{tbl-cov}, with visual distinctions made through colors, line
types, and point shapes. MinT with shrinkage (\emph{MinT-S}) and its
variants are colored in mint green, while MinT with NOVELIST
(\emph{MinT-N}) and its variants are in purple. The solid lines
represent vanilla \emph{MinT-S} and \emph{MinT-N}; the dashed lines with
dot points denote the PC-adjusted variants (e.g.~\emph{MinT-S(PC1)},
\emph{MinT-N(PC2)}); and the dotted or dashed-dotted lines indicate the
scaled variance and h-step-ahead residual versions
(e.g.~\emph{MinT-S(SV)}, \emph{MinT-N(hcov)}).

The first key observation is that methods with shrinkage slightly
outperform those with NOVELIST across all scenarios, although the
differences are small. Second, the PC-adjusted variants (using one and
two principal components) do not yield improvements over the vanilla
versions. This is expected since the synthetic data generating process
does not simulate from strong latent factors. Lastly, the scaled
variance and h-step-ahead residual approaches do not enhance performance
as the forecast horizon increases, suggesting that the proportionality
assumption may not be severely violated in this VAR(1) setup.

When looking into each Monte Carlo replication, we find that there are
instances where the NOVELIST estimator collapses to the shrinkage
estimator due to a large optimal threshold \(\hat{\delta}\) being
selected in the cross-validation step, resulting in a diagonal shrinkage
target.

\begin{figure}

\centering{

\includegraphics[width=1\linewidth,height=\textheight,keepaspectratio]{paper_draft_files/figure-pdf/fig-sim-results-2-1.pdf}

}

\caption{\label{fig-sim-results-2}Percentage relative improvement in
Energy score in both hierarchies and time dimensions, for 1-step-ahead
forecasts. The positive entries indicate a decrease in Energy score
relative to base.}

\end{figure}%

Moving on to probabilistic forecasts, we evaluate the performance of
reconciliation methods using the energy score, a proper scoring rule for
multivariate predictive distributions. Figure~\ref{fig-sim-results-2}
presents the percentage relative improvement in energy score for
1-step-ahead forecasts. Across both hierarchies and time dimensions, the
MinT methods consistently outperform the base forecasts. Meanwhile, the
differences among \emph{MinT-S} and \emph{MinT-N} variants are small,
except for the 6 by 6 case with 50 observations, where shrinkage has a
pronounced edge. The PC-adjusted variants again degrade performance as
we add more principal components. The scaled variance and h-step-ahead
residuals approaches are not available since we only evaluate
1-step-ahead covariance estimates.

Other scoring rules, such as the Winkler score and CRPS, yield similar
conclusions.

\subsection{Other Data Generating
Processes}\label{other-data-generating-processes}

In attempts to differentiate the performance of NOVELIST from the
shrinkage estimator, we also simulate from a sparse covariance matrix
for the previous section's 2 groups of 50 bottom series setting, as
illustrated in Figure~\ref{fig-Sigma-2x50}. The sparse covariance matrix
is obtained by randomly choosing 40\% of the bottom series and setting
their correlations with all other series to zero, resulting in a
grid-like sparse structure. The VAR(1) coefficient matrix remains the
same as in the dense case. The idea is to allow NOVELIST to exploit the
sparsity in the covariance structure, since it can control the sparsity
of the shrinkage target via the thresholding parameter \(\delta\).
However, no profound insights can be drawn from the results.

\begin{figure}

\centering{

\includegraphics[width=0.8\linewidth,height=\textheight,keepaspectratio]{paper_draft_files/figure-pdf/fig-Sigma-2x50-1.pdf}

}

\caption{\label{fig-Sigma-2x50}Sparse correlation matrix of the
innovation process for 2 groups of 50 structure.}

\end{figure}%

Additional designs (varying block sizes, grouped structure, aggregation
paths, correlation configurations) also failed to separate NOVELIST from
Shrinkage. Their nearly identical performance under these synthetic
scenarios suggests that our current simulation may not unveil the full
advantages of the thresholding estimators. Nevertheless, we have not
explored settings where PC variants or using h-step-ahead residuals
approaches would have an edge.

These findings motivates our turn to empirical data in the next section,
where latent structural features, regime shifts, and noisy, intermittent
series may possibly reveal performance differences.

\section{Forecasting Australian Domestics Tourism}\label{sec-empirical}

Forecasting domestic tourism flows is of great economic and policy
significance, as policymakers and stakeholders rely on accurate
forecasts to make informed decisions regarding resource allocation,
infrastructure development, and marketing strategies. The domestic
tourism flows in Australia exhibit a natural hierarchical and grouped
structure, driven both by geography and by purpose of travel. At the top
of this hierarchy lies the national total, which splits into the seven
states and territories. Each state is further sub-divided into tourism
zones, which in turn break down into 77 regions. A complete illustration
of this geographic hierarchy appears in Appendix
Section~\ref{sec-tourism-supplement}. Intersecting this geographic
hierarchy is a second dimension: policymakers are also interested in
travel motives. This partitions tourism flows into four categories:
holiday, business, visiting friends and relatives, and other.
Altogether, this yields a grouped structure of 560 series, from the most
disaggregated regional-purpose cells up to the full national aggregate.
Table~\ref{tbl-tourism-structure} depicts this structure.

\begin{longtable}[]{@{}
  >{\raggedright\arraybackslash}p{(\linewidth - 6\tabcolsep) * \real{0.1833}}
  >{\centering\arraybackslash}p{(\linewidth - 6\tabcolsep) * \real{0.3667}}
  >{\centering\arraybackslash}p{(\linewidth - 6\tabcolsep) * \real{0.2500}}
  >{\centering\arraybackslash}p{(\linewidth - 6\tabcolsep) * \real{0.2000}}@{}}

\caption{\label{tbl-tourism-structure}Hierarchical and grouped structure
of Australian domestic tourism flows}

\tabularnewline

\toprule\noalign{}
\begin{minipage}[b]{\linewidth}\raggedright
Geographical division
\end{minipage} & \begin{minipage}[b]{\linewidth}\centering
Number of series per geographical division
\end{minipage} & \begin{minipage}[b]{\linewidth}\centering
Number of series per purpose
\end{minipage} & \begin{minipage}[b]{\linewidth}\centering
Total number of series
\end{minipage} \\
\midrule\noalign{}
\endhead
\bottomrule\noalign{}
\endlastfoot
Australia & 1 & 4 & 5 \\
States & 7 & 28 & 35 \\
Zones & 27 & 108 & 135 \\
Regions & 77 & 308 & 385 \\
Total & 112 & 448 & 560 \\

\end{longtable}

We quantify tourism demand via ``visitor nights'', the total number of
nights spent by Australians away from home. The data is collected via
the National Visitor Survey, managed by Tourism Research Australia,
using computer assisted telephone interviews from nearly 120,000
Australian residents aged 15 years and over
(\citeproc{ref-tourism-au}{\emph{Tourism Research Australia}, 2024}).

The data are monthly time series spanning from January 1998 to December
2016, resulting in 228 observations per series. This gives a challenging
high-dimensional forecasting problem with number of series (560) doubled
the number of observations, which is ideal for evaluating reconciliation
approaches that rely on high-dimensional covariance estimation. The
extreme dimensionality over sample size mirrors many contemporary
business problems, for instance, Starbucks Corporation sales. Tourism
demand is also economically vital yet highly volatile, with geographical
and purpose‑specific patterns create a realistic stress‑test for
reconciliation algorithms.

\begin{figure}

\centering{

\includegraphics[width=0.5\linewidth,height=\textheight,keepaspectratio]{D:/Github/Recon_Honours_Thesis/figs/tourism_cv.png}

}

\caption{\label{fig-tourism-cv}Rolling-window cross-validation scheme
for evaluating forecasting performance in Australia tourism data}

\end{figure}%

To assess forecasting performance between models, we adopt a
rolling-window cross-validation scheme. Beginning with the first 120
monthly observations (January 1998-December 2005) as the initial
training set, we obtain the best-fitted ARIMA model for each of the 560
series via the automatic algorithm by minimising AICc from Hyndman \&
Khandakar (\citeproc{ref-Hyndman2008-rf}{2008}), implemented in the
\emph{fabletools} package (\citeproc{ref-O-Hara-Wild2024-we}{O'Hara-Wild
et al., 2024}). The 1- to 12-step-ahead base forecasts are then
generated by these ARIMA models, and then reconciled using multiple
approaches. To estimate the NOVELIST and its variants, we implement an
extra cross-validation prodecure within this training window, as
described in Algorithm \ref{alg:novelist_cv}. We then roll the training
window forward by one month and refit all models, re-estimate
reconciliation variants, and produce another batch of 1- to
12-step-ahead forecasts, repeating until the training set reaches
December 2015. In total, this results in 97 out-of-sample windows. The
entire procedure is illustrated in Figure~\ref{fig-tourism-cv}.

\begin{figure}

\centering{

\pandocbounded{\includegraphics[keepaspectratio]{paper_draft_files/figure-pdf/fig-tourism-MSE-line-1.pdf}}

}

\caption{\label{fig-tourism-MSE-line}Percentage relative improvement in
the mean squared error (MSE) of different reconciled forecasts over the
base forecasts for the Australian domestic tourism data, for 1 to 12
steps ahead forecasts. The positive entries indicate an decrease in
MSE.}

\end{figure}%

\begin{figure}

\centering{

\includegraphics[width=0.9\linewidth,height=\textheight,keepaspectratio]{paper_draft_files/figure-pdf/fig-tourism-MSE-heat-1.pdf}

}

\caption{\label{fig-tourism-MSE-heat}Heatmap of relative improvement in
the mean squared error (MSE) of different reconciled forecasts over the
base forecasts for the Australian domestic tourism data, for 1 to 12
steps ahead forecasts. The values are scaled to the range of 0 to 100
for better visualisation, with darker colors indicating greater
improvement and best performance is noted by a star. This re-scaling
does not affect relative rankings among methods.}

\end{figure}%

Figure~\ref{fig-tourism-MSE-line} reports percentage relative
improvements in MSE of reconciled forecasts over the incoherent base
ARIMA forecasts, across horizons from 1 to 12 months. We evaluate ten
MinT variants (as defined in Table~\ref{tbl-cov}), similar to the
simulation's Figure~\ref{fig-sim-results-1}. The MinT with shrinkage
(\emph{MinT-S}) and its variants are colored in mint green, while MinT
with NOVELIST (\emph{MinT-N}) and its variants are in purple. Solid
lines represent ``vanilla'' \emph{MinT-S} and \emph{MinT-N}; dashed
lines with dot points denote the PC-adjusted variants
(e.g.~\emph{MinT-S(PC1)}, \emph{MinT-N(PC2)}); and dotted or
dashed-dotted lines indicate the scaled variance and h-step-ahead
residual versions (e.g.~\emph{MinT-S(SV)}, \emph{MinT-N(hcov)}).

Three patterns stand out. First, the vanilla \emph{MinT-S} (solid green)
slightly outperform \emph{MinT-N} (solid purple) at most horizons,
although the gap is small. Second, variants that alter the multi-step
covariance, either via scaled variance or direct h-step residual
covariances, underperform standard MinT, suggesting that extra
estimation at horizon \(h>1\) is not rewarded in this empirical
analysis. Most importantly, incorporating a single dominant factor via
PC-adjustment (dashed line with round points) delivers best performance
among all methods. Both \emph{MinT-S(PC1)} and \emph{MinT-N(PC1)}
consistently beat their unadjusted counterparts across horizons.
However, adding more than one PC brings no additional benefit and can
erode performance, likely due to injecting idiosyncratic noise from
weaker components.

Figure~\ref{fig-tourism-MSE-heat} complements the line plot with a
heatmap that standardises each method's MSE improvement to a 0--100
scale to enhance visual discrimination across horizons. Darker shades
indicate larger improvements. The heatmap highlights the pattern from
the line plot: \emph{MinT-N(PC1)} and \emph{MinT-S(PC1)} exhibit the
highest concentration of dark cells and are repeatedly marked as the
best method across horizons (star markers). Overall, the results
underscores the consistent gains from PC adjustment and highlights the
diminishing returns from more complex covariance treatments (lighter
cells for scaled variance and h-step residuals).

\begin{figure}

\centering{

\pandocbounded{\includegraphics[keepaspectratio]{paper_draft_files/figure-pdf/fig-tourism-probscores-1.pdf}}

}

\caption{\label{fig-tourism-probscores}Percentage relative improvement
in the Winkler score at 80\% and 95\% nominal coverage, CRPS, and Energy
score of multiple reconciled forecasts over the base forecasts for the
Australian domestic tourism data, for 1-step-ahead forecasts. The
positive (negative) entries indicate a decrease (increase) in the
probabilistic scores relative to base.}

\end{figure}%

Turning to probabilistic forecasts (1-step-ahead forecasts),
Figure~\ref{fig-tourism-probscores} shows that \emph{MinT-N} variants
(purple bars) consistently outperform \emph{MinT-S} variants (green
bars) across univariate and multivariate scores. The PC1-adjusted
variants again yield improvements over their vanilla counterparts, with
\emph{MinT-N(PC1)} leading overall. In the multivariate evaluation, the
Energy score places \emph{OLS} close to the PC1-adjusted MinT methods.
One surprising finding is that all MinT variants produce inferior
forecasts compared to the base forecasts when considering 95\% Winkler
score, while \emph{OLS} performs best.

\begin{figure}

\centering{

\includegraphics[width=1\linewidth,height=\textheight,keepaspectratio]{paper_draft_files/figure-pdf/fig-tourism-coverage-95-level-1.pdf}

}

\caption{\label{fig-tourism-coverage-95-level}Percentage relative
improvement in Winkler score at 95\% nominal coverage by aggregation
level (left), and empirical coverage of 95\% prediction intervals by
aggregation level (right).}

\end{figure}%

To dissect the 95\% Winkler score results,
Figure~\ref{fig-tourism-coverage-95-level} breaks down performance by
hierarchical level, and examines empirical coverage of the 95\%
prediction intervals. The left panel shows that \emph{MinT-S} produces
inferior forecasts compared to the base forecasts at all levels,
especially in higher aggregated levels. From our inspection, this is due
to the overly shrunk variances from the shrinkage estimator, leading to
narrow prediction intervals. The right panel confirms this, as
\emph{MinT-S} has the lowest empirical coverage across all levels,
indicating that its prediction intervals fail to capture the true
observations adequately and resulting in poor Winkler scores.
\emph{MinT-N} has relatively good coverage and improved Winkler score at
bottom levels, but still is inferior at higher levels. The PC-adjusted
variants seem to strike a better balance, improving overall coverage and
relative Winkler scores over the base.

\begin{figure}

\centering{

\includegraphics[width=0.7\linewidth,height=\textheight,keepaspectratio]{paper_draft_files/figure-pdf/fig-tourism-radar-1.pdf}

}

\caption{\label{fig-tourism-radar}Radar plot of relative improvements in
probabilistic scores (Winkler score at 80\% and 95\% intervals, CRPS,
and Energy) over the base forecasts. The scores are scaled to a range of
0 to 100, with larger values indicating better performance. The
outermost polygon represents the best possible score (100) and the
innermost polygon represents the worst possible score (0). Only the top
4 MinT approaches are shown, together with the base forecasts.}

\end{figure}%

The summary radar graph in Figure~\ref{fig-tourism-radar} consolidates
these findings. Among the selected MinT variants considering
probabilistic criteria, \emph{MinT-N(PC1)} (yellow polygon) clearly
leads across CRPS, Winkler score at 80\% and 95\% intervals, and Energy
score, extending the improvements seen in the single-metric panels.
Except for the Winkler score at 95\% interval, all MinT variants
outperform the base forecasts.

\section{Conclusions and Future Work}\label{sec-conclusion}

This paper tackles a central bottleneck in Minimum Trace (MinT)
reconciliation: obtaining reliable, positive-definite estimates of the
base-forecast error covariance. The original practice, following
Wickramasuriya et al. (\citeproc{ref-Wickramasuriya2019-xq}{2019}),
estimates the one-step covariance via diagonal-target shrinkage and
obtains multi-step horizons by proportional scaling. We identified three
limitations of this approach--uniform shrinkage, lack of factor
awareness, and horizon-invariant dependence--and proposed alternatives
that address them individually and in combination. Specifically, we
adopt NOVELIST to introduce a more flexible shrinkage target; develop
PC-adjusted variants of shrinkage and NOVELIST to preserve dominant
latent components in the data; and consider multi-step covariance
constructions that relax proportional scaling assumption.

Empirical evidence from Australian domestic tourism yields several
insights. For point reconciliation, shrinkage (and its PC-adjusted
version) remains a robust default choice, while NOVELIST demonstrates
improved performance for probabilistic forecasts. Across both settings,
adjusting for a single dominant principal component consistently
enhances performance, indicating its utility when such latent structures
are present. More complex adjustments, such as incorporating multiple
principal components or horizon-specific covariances, tend to introduce
additional idiosyncratic noise without notable benefits.

Several limitations and directions for future work remain. First, more
intricate simulation designs are needed to further highlight the
differences between shrinkage and NOVELIST, particularly under explicit
factor models and varying sparsity/noise levels. Second, our current
PC-adjusted implementations do not standardise series prior to factor
extraction, which may lead to higher-level series largely influencing
the principal components. Third, we can explore alternative methods that
leverage cross-series information or eigenvalue/eigenvector structures.
Fourth, the observed inferior performance of MinT with shrinkage under
the Winkler 95\% score at higher aggregation levels needs detailed
investigation. Finally, evaluating genuinely \(h\)-step probabilistic
reconciliation remains an open avenue, including how to best regularise
multi-step covariances given limited effective samples.

All data generation, covariance estimation, and reconciliation routines
were implemented in the ReconCov R package and is available under an
open‐source license on GitHub (\citeproc{ref-ReconCov}{Su, 2025}).

\pagebreak

\section{Appendix}\label{sec-appendix}

\subsection{Appendix: Simulation
Supplementary}\label{sec-sim-supplement}

\subsection{Appendix: Australian Domestic Tourism Geographical
Hierarchy}\label{sec-tourism-supplement}

\begingroup\fontsize{8}{10}\selectfont

\begin{longtable}[t]{rllrll}

\caption{\label{tbl-tourism-geo}Geographical divisions of Australia.}

\tabularnewline

\toprule
Series & Name & Label & Series & Name & Label\\
\midrule
1 & Australia & Total & 57 & Bundaberg & CAA\\
2 & NSW & A & 58 & Capricorn & CAB\\
3 & NT & B & 59 & Fraser Coast & CAC\\
4 & QLD & C & 60 & Gladstone & CAD\\
5 & SA & D & 61 & Mackay & CAE\\
\addlinespace
6 & TAS & E & 62 & Southern Queensland Country & CAF\\
7 & VIC & F & 63 & Outback Queensland & CBA\\
8 & WA & G & 64 & Brisbane & CCA\\
9 & ACT & AA & 65 & Gold Coast & CCB\\
10 & Metro NSW & AB & 66 & Sunshine Coast & CCC\\
\addlinespace
11 & Nth Coast NSW & AC & 67 & Townsville & CDA\\
12 & Nth NSW & AD & 68 & Tropical North Queensland & CDB\\
13 & Sth Coast NSW & AE & 69 & Whitsundays & CDC\\
14 & Sth NSW & AF & 70 & Clare Valley & DAA\\
15 & Central NT & BA & 71 & Flinders Ranges and Outback & DAB\\
\addlinespace
16 & Nth Coast NT & BB & 72 & Murray River, Lakes and Coorong & DAC\\
17 & Central Coast QLD & CA & 73 & Riverland & DAD\\
18 & Inland QLD & CB & 74 & Adelaide & DBA\\
19 & Metro QLD & CC & 75 & Adelaide Hills & DBB\\
20 & Nth Coast QLD & CD & 76 & Barossa & DBC\\
\addlinespace
21 & Inland SA & DA & 77 & Fleurieu Peninsula & DCA\\
22 & Metro SA & DB & 78 & Kangaroo Island & DCB\\
23 & Sth Coast SA & DC & 79 & Limestone Coast & DCC\\
24 & West Coast SA & DD & 80 & Eyre Peninsula & DDA\\
25 & Nth East TAS & EA & 81 & Yorke Peninsula & DDB\\
\addlinespace
26 & Nth West TAS & EB & 82 & East Coast & EAA\\
27 & Sth TAS & EC & 83 & Launceston and the North & EAB\\
28 & East Coast VIC & FA & 84 & North West & EBA\\
29 & Metro VIC & FB & 85 & West Coast & EBB\\
30 & Nth East VIC & FC & 86 & Hobart and the South & ECA\\
\addlinespace
31 & Nth West VIC & FD & 87 & Gippsland & FAA\\
32 & West Coast VIC & FE & 88 & Lakes & FAB\\
33 & Nth WA & GA & 89 & Phillip Island & FAC\\
34 & Sth WA & GB & 90 & Geelong and the Bellarine & FBA\\
35 & West Coast WA & GC & 91 & Melbourne & FBB\\
\addlinespace
36 & Canberra & AAA & 92 & Peninsula & FBC\\
37 & Central Coast & ABA & 93 & Central Murray & FCA\\
38 & Sydney & ABB & 94 & Goulburn & FCB\\
39 & Hunter & ACA & 95 & High Country & FCC\\
40 & North Coast NSW & ACB & 96 & Melbourne East & FCD\\
\addlinespace
41 & Blue Mountains & ADA & 97 & Murray East & FCE\\
42 & Central NSW & ADB & 98 & Upper Yarra & FCF\\
43 & New England North West & ADC & 99 & Ballarat & FDA\\
44 & Outback NSW & ADD & 100 & Bendigo Loddon & FDB\\
45 & South Coast & AEA & 101 & Central Highlands & FDC\\
\addlinespace
46 & Capital Country & AFA & 102 & Macedon & FDD\\
47 & Riverina & AFB & 103 & Mallee & FDE\\
48 & Snowy Mountains & AFC & 104 & Spa Country & FDF\\
49 & The Murray & AFD & 105 & Western Grampians & FDG\\
50 & Alice Springs & BAA & 106 & Wimmera & FDH\\
\addlinespace
51 & Barkly & BAB & 107 & Great Ocean Road & FEA\\
52 & Lasseter & BAC & 108 & Australia's North West & GAA\\
53 & MacDonnell & BAD & 109 & Australia's Golden Outback & GBA\\
54 & Darwin & BBA & 110 & Australia's Coral Coast & GCA\\
55 & Katherine Daly & BBB & 111 & Australia's South West & GCB\\
\addlinespace
56 & Litchfield Kakadu Arnhem & BBC & 112 & Destination Perth & GCC\\
\bottomrule

\end{longtable}

\endgroup{}

\pagebreak

\section*{References}\label{references}
\addcontentsline{toc}{section}{References}

\phantomsection\label{refs}
\begin{CSLReferences}{1}{0}
\bibitem[\citeproctext]{ref-Angam2025-od}
Angam, B., Beretta, A., De Poorter, E., Duvinage, M., \& Peralta, D.
(2025). Forecast reconciliation for vaccine supply chain optimization.
In \emph{Communications in computer and information science} (pp.
101--118). Springer Nature Switzerland.
\url{https://doi.org/10.1007/978-3-031-74650-5/_6}

\bibitem[\citeproctext]{ref-Athanasopoulos2009-lp}
Athanasopoulos, G., Ahmed, R. A., \& Hyndman, R. J. (2009). Hierarchical
forecasts for australian domestic tourism. \emph{International Journal
of Forecasting}, \emph{25}(1), 146--166.
\url{https://doi.org/10.1016/j.ijforecast.2008.07.004}

\bibitem[\citeproctext]{ref-Athanasopoulos2024-as}
Athanasopoulos, G., Hyndman, R. J., Kourentzes, N., \& Panagiotelis, A.
(2024). \emph{Forecast reconciliation: A review}. \emph{40}(2),
430--456.
\url{https://www.sciencedirect.com/science/article/pii/S0169207023001097}

\bibitem[\citeproctext]{ref-Ben-Taieb2019-yx}
Ben Taieb, S., \& Koo, B. (2019). Regularized regression for
hierarchical forecasting without unbiasedness conditions.
\emph{Proceedings of the 25th ACM SIGKDD International Conference on
Knowledge Discovery \& Data Mining}.
\url{https://doi.org/10.1145/3292500.3330976}

\bibitem[\citeproctext]{ref-Ben-Taieb2021-bn}
Ben Taieb, S., Taylor, J. W., \& Hyndman, R. J. (2021). Hierarchical
probabilistic forecasting of electricity demand with smart meter data.
\emph{Journal of the American Statistical Association}, \emph{116}(533),
27--43. \url{https://doi.org/10.1080/01621459.2020.1736081}

\bibitem[\citeproctext]{ref-Bickel2008-nh}
Bickel, P. J., \& Levina, E. (2008). Covariance regularization by
thresholding. \emph{Annals of Statistics}, \emph{36}(6), 2577--2604.
\url{https://doi.org/10.1214/08-aos600}

\bibitem[\citeproctext]{ref-Cai2011-jf}
Cai, T., \& Liu, W. (2011). Adaptive thresholding for sparse covariance
matrix estimation. \emph{Journal of the American Statistical
Association}, \emph{106}(494), 672--684.
\url{https://doi.org/10.1198/jasa.2011.tm10560}

\bibitem[\citeproctext]{ref-Carrara2025-rz}
Carrara, C., Zambon, L., Azzimonti, D., \& Corani, G. (2025). A novel
shrinkage estimator of the covariance matrix for hierarchical time
series. In \emph{Italian statistical society series on advances in
statistics} (pp. 140--145). Springer Nature Switzerland.
\url{https://doi.org/10.1007/978-3-031-96736-8/_24}

\bibitem[\citeproctext]{ref-Di-Modica2021-ad}
Di Modica, C., Pinson, P., \& Ben Taieb, S. (2021). Online forecast
reconciliation in wind power prediction. \emph{Electric Power Systems
Research}, \emph{190}(106637), 106637.
\url{https://doi.org/10.1016/j.epsr.2020.106637}

\bibitem[\citeproctext]{ref-El-Gemayel2022-hx}
El Gemayel, J., Lafarguette, R., Itd, K. M., et al. (2022). \emph{United
arab emirates: Technical assistance reportliquidity management and
forecasting}.

\bibitem[\citeproctext]{ref-Van_Erven2015-nx}
Erven, T. van, \& Cugliari, J. (2015). Game-theoretically optimal
reconciliation of contemporaneous hierarchical time series forecasts. In
\emph{Modeling and stochastic learning for forecasting in high
dimensions} (pp. 297--317). Springer International Publishing.
\url{https://doi.org/10.1007/978-3-319-18732-7/_15}

\bibitem[\citeproctext]{ref-Fan2013-jz}
Fan, J., Liao, Y., \& Mincheva, M. (2013). Large covariance estimation
by thresholding principal orthogonal complements. \emph{Journal of the
Royal Statistical Society. Series B, Statistical Methodology},
\emph{75}(4), 603--680. \url{https://doi.org/10.1111/rssb.12016}

\bibitem[\citeproctext]{ref-Gamakumara2020-ml}
Gamakumara, P. (2020). \emph{Probabilistic forecast reconciliation:
Theory and applications} {[}PhD thesis, Monash University{]}.
\url{https://doi.org/10.26180/5e4ca9d0c4b9d}

\bibitem[\citeproctext]{ref-Hardin2013-wu}
Hardin, J., Garcia, S. R., \& Golan, D. (2013). A method for generating
realistic correlation matrices. \emph{The Annals of Applied Statistics},
\emph{7}(3), 1733--1762. \url{https://www.jstor.org/stable/23566492}

\bibitem[\citeproctext]{ref-Higham2002-te}
Higham, N. (2002). Computing the nearest correlation matrix---a problem
from finance. \emph{Ima Journal of Numerical Analysis}, \emph{22},
329--343. \url{https://doi.org/10.1093/IMANUM/22.3.329}

\bibitem[\citeproctext]{ref-Huang2019-ua}
Huang, N., \& Fryzlewicz, P. (2019). {NOVELIST} estimator of large
correlation and covariance matrices and their inverses. \emph{Test
(Madrid, Spain)}, \emph{28}(3), 694--727.
\url{https://doi.org/10.1007/s11749-018-0592-4}

\bibitem[\citeproctext]{ref-Hyndman2011-jv}
Hyndman, R. J., Ahmed, R. A., Athanasopoulos, G., \& Shang, H. L.
(2011). Optimal combination forecasts for hierarchical time series.
\emph{Computational Statistics \& Data Analysis}, \emph{55}(9),
2579--2589. \url{https://doi.org/10.1016/j.csda.2011.03.006}

\bibitem[\citeproctext]{ref-Hyndman2008-rf}
Hyndman, R. J., \& Khandakar, Y. (2008). Automatic time series
forecasting: The forecast package for {R}. \emph{Journal of Statistical
Software}, \emph{27}, 1--22. \url{https://doi.org/10.18637/JSS.V027.I03}

\bibitem[\citeproctext]{ref-Hyndman2016-ic}
Hyndman, R. J., Lee, A. J., \& Wang, E. (2016). Fast computation of
reconciled forecasts for hierarchical and grouped time series.
\emph{Computational Statistics \& Data Analysis}, \emph{97}, 16--32.
\url{https://doi.org/10.1016/j.csda.2015.11.007}

\bibitem[\citeproctext]{ref-Jeon2019-wm}
Jeon, J., Panagiotelis, A., \& Petropoulos, F. (2019). Probabilistic
forecast reconciliation with applications to wind power and electric
load. \emph{European Journal of Operational Research}, \emph{279}(2),
364--379. \url{https://doi.org/10.1016/j.ejor.2019.05.020}

\bibitem[\citeproctext]{ref-Ledoit2012-ga}
Ledoit, O., \& Wolf, M. (2012). Nonlinear shrinkage estimation of
large-dimensional covariance matrices. \emph{Annals of Statistics}.
\url{https://doi.org/10.1214/12-AOS989}

\bibitem[\citeproctext]{ref-Ledoit2020-xt}
Ledoit, O., \& Wolf, M. (2020). Analytical nonlinear shrinkage of
large-dimensional covariance matrices. \emph{Annals of Statistics},
\emph{48}(5), 3043--3065. \url{https://doi.org/10.1214/19-AOS1921}

\bibitem[\citeproctext]{ref-Li2019-vv}
Li, H., Li, H., Lu, Y., \& Panagiotelis, A. (2019). A forecast
reconciliation approach to cause-of-death mortality modeling.
\emph{Insurance, Mathematics \& Economics}, \emph{86}, 122--133.
\url{https://doi.org/10.1016/j.insmatheco.2019.02.011}

\bibitem[\citeproctext]{ref-Nixtla}
Nixtla. (2025). \emph{Time series forecasting software}.
\url{https://www.nixtla.io/}

\bibitem[\citeproctext]{ref-O-Hara-Wild2024-we}
O'Hara-Wild, M., Hyndman, R. J., \& Wang, E. (2024). \emph{Fabletools
{R} package} (Version v0.5.0). \url{https://fabletools.tidyverts.org/}

\bibitem[\citeproctext]{ref-Panagiotelis2023-fs}
Panagiotelis, A., Gamakumara, P., Athanasopoulos, G., \& Hyndman, R. J.
(2023). Probabilistic forecast reconciliation: Properties, evaluation
and score optimisation. \emph{European Journal of Operational Research},
\emph{306}(2), 693--706.
\url{https://doi.org/10.1016/j.ejor.2022.07.040}

\bibitem[\citeproctext]{ref-Rothman2009-yf}
Rothman, A. J., Levina, E., \& Zhu, J. (2009). Generalized thresholding
of large covariance matrices. \emph{Journal of the American Statistical
Association}, \emph{104}(485), 177--186.
\url{https://doi.org/10.1198/jasa.2009.0101}

\bibitem[\citeproctext]{ref-Schafer2005-yw}
Schäfer, J., \& Strimmer, K. (2005). A shrinkage approach to large-scale
covariance matrix estimation and implications for functional genomics.
\emph{Statistical Applications in Genetics and Molecular Biology},
\emph{4}(1), Article32. \url{https://doi.org/10.2202/1544-6115.1175}

\bibitem[\citeproctext]{ref-Seaman2022-bb}
Seaman, B., \& Bowman, J. (2022). Applicability of the {M5} to
forecasting at walmart. \emph{International Journal of Forecasting},
\emph{38}(4), 1468--1472.
\url{https://doi.org/10.1016/j.ijforecast.2021.06.002}

\bibitem[\citeproctext]{ref-Shang2017-ux}
Shang, H. L., \& Hyndman, R. J. (2017). Grouped functional time series
forecasting: An application to age-specific mortality rates.
\emph{Journal of Computational and Graphical Statistics: A Joint
Publication of American Statistical Association, Institute of
Mathematical Statistics, Interface Foundation of North America},
\emph{26}(2), 330--343.
\url{https://doi.org/10.1080/10618600.2016.1237877}

\bibitem[\citeproctext]{ref-ReconCov}
Su, V. (2025). \emph{{ReconCov} {R} package} (Version beta).
\url{https://github.com/lordtahdus/ReconCov}

\bibitem[\citeproctext]{ref-tourism-au}
\emph{Tourism research australia}. (2024).
\url{https://www.tra.gov.au/}.

\bibitem[\citeproctext]{ref-Wickramasuriya2024-mb}
Wickramasuriya, S. L. (2024). Probabilistic forecast reconciliation
under the gaussian framework. \emph{Journal of Business \& Economic
Statistics: A Publication of the American Statistical Association},
\emph{42}(1), 272--285.
\url{https://doi.org/10.1080/07350015.2023.2181176}

\bibitem[\citeproctext]{ref-Wickramasuriya2019-xq}
Wickramasuriya, S. L., Athanasopoulos, G., \& Hyndman, R. J. (2019).
Optimal forecast reconciliation for hierarchical and grouped time series
through trace minimization. \emph{Journal of the American Statistical
Association}, \emph{114}(526), 804--819.
\url{https://doi.org/10.1080/01621459.2018.1448825}

\bibitem[\citeproctext]{ref-Wickramasuriya2020-uk}
Wickramasuriya, S. L., Turlach, B. A., \& Hyndman, R. J. (2020). Optimal
non-negative forecast reconciliation. \emph{Statistics and Computing},
\emph{30}(5), 1167--1182.
\url{https://doi.org/10.1007/s11222-020-09930-0}

\end{CSLReferences}

\end{document}

### 2025-10-23 thesis/paper_draft.qmd
bug fixes
fixes after meeting with shanika

@@
 The current practice uses linear shrinkage towards a diagonal target.
-Furthermore, the covariance estimate is based on 1-step-ahead residuals, then proportionally scale it to approximate h-step-ahead covariance matrix.
+Furthermore, the covariance estimate is based on in-sample 1-step-ahead forecast errors, then proportionally scale it to approximate $h$-step-ahead covariance matrix.
 This leaves a question of whether this method is appropriate for all real-world applications.
-We study the shortcomings of current practice and propose alternative covariance estimators, including the NOVELIST estimator (shrinkage towards a soft-thresholded target), PC-adjusted shrinkage (which utilises latent factor structures), and horizon-specific estimators that relax proportional scaling.
-We evaluate MinT using these covariance estimates for both point and probabilistic reconciliation, and demonstrate their effectiveness and improvements over the shrinkage estimator in a complex, large-hierarchy dataset.
+We study the shortcomings of current practice and assess the impacts of alternative covariance estimators on MinT, including the NOVELIST estimator (shrinkage towards a soft-thresholded target), PC-adjusted shrinkage (which utilises latent factor structures), and horizon-specific estimators that relax proportional scaling.
+We evaluate the accuracy of MinT using these covariance estimates for both point and probabilistic reconciled forecasts, and demonstrate their effectiveness and improvements over the shrinkage estimator in a complex, large-hierarchical dataset.
 
@@
 Second, many hierarchical data sets exhibit strong latent low-rank structures that can be explicitly exploited.
-Third, the proportionality relationship between the h-step-ahead and 1-step-ahead forecast error covariance matrices might not hold when error dynamics change with horizon.
+Third, the proportionality relationship between the $h$-step-ahead and 1-step-ahead forecast error covariance matrices might not hold when error dynamics change with horizon.
 
@@
 Since the projection matrix $\boldsymbol{P}$ is idempotent, $\boldsymbol{G}$ must satisfy the condition $\boldsymbol{S} \boldsymbol{G} \boldsymbol{S} = \boldsymbol{S}$.
-Within this class, a broad family of mapping matrices is given by: $\boldsymbol{G} = (\boldsymbol{S}' \boldsymbol{M}^{-1} \boldsymbol{S})^{-1} \boldsymbol{S}' \boldsymbol{M}^{-1}$, for some positive definite matrix $\boldsymbol{M} \in \mathbb{R}^{n \times n}$ [@Gamakumara2020-ml].
+Within this class, a broad family of mapping matrices is given by: $\boldsymbol{G} = (\boldsymbol{S}' \boldsymbol{M}^{-1} \boldsymbol{S})^{-1} \boldsymbol{S}' \boldsymbol{M}^{-1}$, for some positive definite matrix $\boldsymbol{M} \in \mathbb{R}^{n \times n}$ [@Panagiotelis2021-ci].
 
@@
 This projection view extends seamlessly from point to probabilistic forecast reconciliation.
-Let $\hat{\boldsymbol{Y}}_{t+h|t}$ denote the random vector representing the $h$-step-ahead base forecasts.
+Given an $h$-step-ahead base forecast density with mean $\hat{\boldsymbol{y}}_{t+h|t}$ and covariance $\boldsymbol{W}_h$, via the linear projection with $\boldsymbol{P}$, the resulting reconciled forecast density has mean $\tilde{\boldsymbol{y}}_{t+h|t} = \boldsymbol{P} \hat{\boldsymbol{y}}_{t+h|t}$ and covariance $\boldsymbol{P} \boldsymbol{W}_h \boldsymbol{P}'$.
+@fig-geometry schematically illustrates this projection.
+Reconciliation takes the (possibly elliptical) base forecast distribution and "pushes" the entire mass linearly onto the coherent subspace, resulting in the reconciled forecast distribution.
+
+<!-- Let $\hat{\boldsymbol{Y}}_{t+h|t}$ denote the random vector representing the $h$-step-ahead base forecasts.
 The random reconciled forecast vector is obtained similarly via projection: $\tilde{\boldsymbol{Y}}_{t+h|t} = \boldsymbol{P} \hat{\boldsymbol{Y}}_{t+h|t}$.
 Consequently, point forecasts (which we refer to as means of the predictive distributions) reconcile as $\mathbb{E}(\tilde{\boldsymbol{Y}}_{t+h|t}) = \mathbb{E}(\boldsymbol{P} \hat{\boldsymbol{Y}}_{t+h|t}) = \boldsymbol{P} \hat{\boldsymbol{y}}_{t+h|t} = \tilde{\boldsymbol{y}}_{t+h|t}$, and the covariance of the predictive distribution transforms linearly as $\text{Var}(\tilde{\boldsymbol{Y}}_{t+h|t}) = \text{Var}(\boldsymbol{P} \hat{\boldsymbol{Y}}_{t+h|t}) = \boldsymbol{P} \boldsymbol{W}_h \boldsymbol{P}'$, where $\boldsymbol{W}_h = \text{Var}(\hat{\boldsymbol{Y}}_{t+h|t})$.
-
-@fig-geometry schematically illustrates this projection.
-Reconciliation takes the (possibly elliptical) base forecast distribution and "pushes" the entire mass linearly onto the coherent subspace, resulting in the reconciled forecast distribution.
+-->
 
@@
 Although originally developed for point reconciliation, @Wickramasuriya2024-mb showed that MinT extends naturally to the probabilistic setting when base predictive distributions are Gaussian.
-If $h$-step-ahead base forecast distribution is $\mathcal{N}(\hat{\boldsymbol{y}}_{t+h|t}, \boldsymbol{W}_h)$, then reconciliation gives $\tilde{\boldsymbol{Y}}_{t+h|t} = \boldsymbol{S} \boldsymbol{G}_h \hat{\boldsymbol{Y}}_{t+h|t} \sim \mathcal{N}(\tilde{\boldsymbol{y}}_{t+h|t}, \boldsymbol{S} \boldsymbol{G}_h \boldsymbol{W}_h \boldsymbol{G}_h' \boldsymbol{S}')$.
+If $h$-step-ahead base forecast distribution is $\mathcal{N}(\hat{\boldsymbol{y}}_{t+h|t}, \boldsymbol{W}_h)$, then reconciliation gives a reconciled forecast distribution of $\mathcal{N}(\tilde{\boldsymbol{y}}_{t+h|t}, \boldsymbol{S} \boldsymbol{G}_h \boldsymbol{W}_h \boldsymbol{G}_h' \boldsymbol{S}')$.
 Within the class of linear-unbiased reconcilers, MinT also minimises the negative log score (equivalently, the Gaussian log predictive score) of the reconciled distribution.
@@
 The tourism activities might be driven by a few common factors, such as economic conditions, fuel prices, or major events, which might be left in the forecast errors after fitting the base models.
-@fig-eigen illustrates the largest eigenvalues from the one-step forecast error correlation matrix $\hat{\boldsymbol{R}}_1$, typically showing a marked elbow after a few principal components (largest eigenvalues), indicating strong latent structure.
-Diagonal-target shrinkage is factor-unaware: it neither preserves the low-rank common subspace nor differentially shrinks the idiosyncratic remainder, and so can be inefficient when common factors dominate.
 
@@
 ```
+
+@fig-eigen illustrates the largest eigenvalues from the one-step forecast error correlation matrix $\hat{\boldsymbol{R}}_1$, typically showing a marked elbow after a few principal components (largest eigenvalues), indicating strong latent structure.
+Diagonal-target shrinkage is factor-unaware: it neither preserves the low-rank common subspace nor differentially shrinks the idiosyncratic remainder, and so can be inefficient when common factors dominate.
 
@@
 $$
-\hat{\boldsymbol{W}}^{g, sv}_{h} = \boldsymbol{D}^{1/2}_h \hat{\boldsymbol{R}}^g_1 \boldsymbol{D}^{1/2}_h, \quad
-
-\hat{\boldsymbol{R}}^g_1 = D^{-1/2}_1 \, g(\hat{\boldsymbol{W}}_1) \, D^{-1/2}_1,
+\hat{\boldsymbol{W}}^{g, sv}_{h} = \boldsymbol{D}^{1/2}_h \hat{\boldsymbol{R}}^{g}_1 \boldsymbol{D}^{1/2}_h, \quad
+\hat{\boldsymbol{R}}^{g}_1 = D^{-1/2}_1 \, g(\hat{\boldsymbol{W}}_1) \, D^{-1/2}_1,
 $$
@@
 where $\boldsymbol{D}_h$ is a diagonal matrix comprising $\hat{\sigma}^2_{1,h} , \ldots, \hat{\sigma}^2_{n,h}$ on its diagonal, and $\hat{\sigma}^2_{i,h}$ is $h$-step-ahead base forecast error variance of the $i$-th series.
-$\hat{\boldsymbol{R}}^g_1$ is the estimated one-step correlation matrix using either shrinkage or NOVELIST estimator.
+$\hat{\boldsymbol{R}}^{g}_1$ is the estimated one-step correlation matrix, obtained from either shrinkage or NOVELIST covariance estimate.
 
@@
 
-## Constructing from h-step-ahead Residuals
+## Constructing from $h$-step-ahead Residuals
 
@@
 
-where $\hat{\boldsymbol{W}}_h$ is the covariance matrix of in-sample h-step-ahead base forecast errors, and $g(.)$ is either the shrinkage or NOVELIST estimator.
+where $\hat{\boldsymbol{W}}_h$ is the covariance matrix of in-sample $h$-step-ahead base forecast errors, and $g(.)$ is either the shrinkage or NOVELIST estimator.
 
@@
 | Scaled Variance NOVELIST                | MinT-N(sv)         |
-| Constructed from h-step-ahead Shrinkage | MinT-S(hcov)       |
-| Constructed from h-step-ahead NOVELIST  | MinT-N(hcov)       |
+| Constructed from $h$-step-ahead Shrinkage | MinT-S(hcov)       |
+| Constructed from $h$-step-ahead NOVELIST  | MinT-N(hcov)       |
 
@@
 The solid lines represent vanilla *MinT-S* and *MinT-N*; the dashed lines with dot points denote the PC-adjusted variants (e.g.
-*MinT-S(PC1)*, *MinT-N(PC2)*); and the dotted or dashed-dotted lines indicate the scaled variance and h-step-ahead residual versions (e.g.
+*MinT-S(PC1)*, *MinT-N(PC2)*); and the dotted or dashed-dotted lines indicate the scaled variance and $h$-step-ahead residual versions (e.g.
 *MinT-S(SV)*, *MinT-N(hcov)*).
@@
 This is expected since the synthetic data generating process does not simulate from strong latent factors.
-Lastly, the scaled variance and h-step-ahead residual approaches do not enhance performance as the forecast horizon increases, suggesting that the proportionality assumption may not be severely violated in this VAR(1) setup.
+Lastly, the scaled variance and $h$-step-ahead residual approaches do not enhance performance as the forecast horizon increases, suggesting that the proportionality assumption may not be severely violated in this VAR(1) setup.
 
@@
 The PC-adjusted variants again degrade performance as we add more principal components.
-The scaled variance and h-step-ahead residuals approaches are not available since we only evaluate 1-step-ahead covariance estimates.
+The scaled variance and $h$-step-ahead residuals approaches are not available since we only evaluate 1-step-ahead covariance estimates.
 
@@
 Their nearly identical performance under these synthetic scenarios suggests that our current simulation may not unveil the full advantages of the thresholding estimators.
-Nevertheless, we have not explored settings where PC variants or using h-step-ahead residuals approaches would have an edge.
+Nevertheless, we have not explored settings where PC variants or using $h$-step-ahead residuals approaches would have an edge.
 
@@
 Adding more than one PC brings no additional benefit, likely due to injecting estimation noise from weaker components.
-Variants that modify the multi-step covariance, either via scaled-variance or direct h-step residual covariances, underperform standard MinT, suggesting that extra estimation at horizon $h>1$ is not rewarded in this setting.
+Variants that modify the multi-step covariance, either via scaled-variance or direct $h$-step residual covariances, underperform standard MinT, suggesting that extra estimation at horizon $h>1$ is not rewarded in this setting.
 -->
@@
 Solid lines represent "vanilla" *MinT-S* and *MinT-N*; dashed lines with dot points denote the PC-adjusted variants (e.g.
-*MinT-S(PC1)*, *MinT-N(PC2)*); and dotted or dashed-dotted lines indicate the scaled variance and h-step-ahead residual versions (e.g.
+*MinT-S(PC1)*, *MinT-N(PC2)*); and dotted or dashed-dotted lines indicate the scaled variance and $h$-step-ahead residual versions (e.g.
 *MinT-S(SV)*, *MinT-N(hcov)*).
@@
 First, the vanilla *MinT-S* (solid green) slightly outperform *MinT-N* (solid purple) at most horizons, although the gap is small.
-Second, variants that alter the multi-step covariance, either via scaled variance or direct h-step residual covariances, underperform standard MinT, suggesting that extra estimation at horizon $h>1$ is not rewarded in this empirical analysis.
+Second, variants that alter the multi-step covariance, either via scaled variance or direct $h$-step residual covariances, underperform standard MinT, suggesting that extra estimation at horizon $h>1$ is not rewarded in this empirical analysis.
 Most importantly, incorporating a single dominant factor via PC-adjustment (dashed line with round points) delivers best performance among all methods.
@@
 Adjusting for a single dominant factor via PC decomposition (dashed line with dot points) tightens performance further: both *MinT-S(PC1)* and *MinT-N(PC1)* beat their unadjusted counterparts.
-Other than that, similar patterns are observed: among vanilla MinT variants, the shrinkage estimator (*MinT-S*) slightly outperforms NOVELIST (*MinT-N*) at most horizons; adding more than one PC brings no additional benefit, likely due to injecting estimation noise from weaker components; variants that modify the multi-step covariance, either via scaled-variance or direct h-step residual covariances, underperform standard MinT, suggesting that extra estimation at horizon $h>1$ is not rewarded in this empirical analysis.
+Other than that, similar patterns are observed: among vanilla MinT variants, the shrinkage estimator (*MinT-S*) slightly outperforms NOVELIST (*MinT-N*) at most horizons; adding more than one PC brings no additional benefit, likely due to injecting estimation noise from weaker components; variants that modify the multi-step covariance, either via scaled-variance or direct $h$-step residual covariances, underperform standard MinT, suggesting that extra estimation at horizon $h>1$ is not rewarded in this empirical analysis.
 -->
@@
 The heatmap highlights the pattern from the line plot: *MinT-N(PC1)* and *MinT-S(PC1)* exhibit the highest concentration of dark cells and are repeatedly marked as the best method across horizons (star markers).
-Overall, the results underscores the consistent gains from PC adjustment and highlights the diminishing returns from more complex covariance treatments (lighter cells for scaled variance and h-step residuals).
+Overall, the results underscores the consistent gains from PC adjustment and highlights the diminishing returns from more complex covariance treatments (lighter cells for scaled variance and $h$-step residuals).
 
@@
   - Explore why MinT shrinkage fails for Winkler 95% score, especially at higher levels.
-  - h-step-ahead probabilistic forecast evaluation.
+  - $h$-step-ahead probabilistic forecast evaluation.
 -->
@@
 The original practice, following @Wickramasuriya2019-xq, estimates the one-step covariance via diagonal-target shrinkage and obtains multi-step horizons by proportional scaling.
-We identified three limitations of this approach--uniform shrinkage, lack of factor awareness, and horizon-invariant dependence--and proposed alternatives that address them individually and in combination.
+We identified three limitations of this approach: (1) uniform shrinkage, (2) lack of factor awareness, and (3) horizon-invariant dependence.
+We thus explored alternatives that address them individually and in combination.
 Specifically, we adopt NOVELIST to introduce a more flexible shrinkage target; develop PC-adjusted variants of shrinkage and NOVELIST to preserve dominant latent components in the data; and consider multi-step covariance constructions that relax proportional scaling assumption.

### 2025-10-23 thesis/paper_draft.qmd
minor refine scoring rules section

@@
 
-This section briefly introduces the scoring rules used to evaluate the point and probabilistic forecast accuracy in @sec-simulation and @sec-empirical.
-
-For point forecasts, we use the mean squared error (MSE) to evaluate the accuracy of different reconciliation methods: $MSE = \frac{1}{n} \sum_{i=1}^n (y_{i, t+h} - \tilde{y}_{i,t+h|t})^2$, where $y_{i, t+h}$ is the realised value of series $i$ at time $t+h$, and $\tilde{y}_{i,t+h|t}$ is the reconciled point forecast.
+This section summarises the scoring rules used to evaluate the point and probabilistic forecast accuracy in @sec-simulation and @sec-empirical.
+
+For point forecasts, we use the mean squared error to evaluate the accuracy of different reconciliation methods: $MSE = \frac{1}{n} \sum_{i=1}^n (y_{i, t+h} - \tilde{y}_{i,t+h|t})^2$, where $y_{i, t+h}$ is the realised value of series $i$ at time $t+h$, and $\tilde{y}_{i,t+h|t}$ is the reconciled point forecast.
 
@@
 
-where $y_i$ is the observed value of $i$-th series, and $\boldsymbol{1}(.)$ is the indicator function.
-The Winkler score rewards narrow intervals that contain the observation, and penalises intervals that do not contain the observation.
+where $y_i$ is the observed value of $i$-th series. The Winkler score rewards narrow intervals that contain the observation, and penalises intervals that do not contain the observation.
 
@@
 
-where $z_i = \frac{y_i - \mu_i}{\sigma_i}$, and $\Phi(.)$ and $\phi(.)$ are the CDF and probability density function (PDF) of a standard normal distribution, respectively.
+where $z_i = \frac{y_i - \mu_i}{\sigma_i}$, and $\Phi(.)$ and $\phi(.)$ are the CDF and PDF (probability density function) of a standard normal distribution, respectively.
 
@@
 
-where $\boldsymbol{X}$ and $\boldsymbol{X}'$ are independent random vectors with multivariate distribution $F$, $\boldsymbol{y}$ is the observed vector, $||.||$ is the Euclidean norm, and $\beta \in (0, 2]$.
+where $\boldsymbol{X}$ and $\boldsymbol{X}'$ are independent random vectors having multivariate distribution $F$, $\boldsymbol{y}$ is the observed vector, $||.||$ is the Euclidean norm, and $\beta \in (0, 2]$.
 We set $\beta = 1$ following common convention.
 
-Since the closed-form expression of the ES may not available, we approximate it using Monte Carlo samples $\{\boldsymbol{x}_1, \ldots, \boldsymbol{x}_M\}$ drawn from $P$:
+Since the closed-form expression of the ES is not available, we approximate it using Monte Carlo samples $\{\boldsymbol{x}_1, \ldots, \boldsymbol{x}_M\}$ drawn from $F$:
 

### 2025-10-23 thesis/paper_draft.qmd
add acknowledgement section

@@
 {\large Supervised by\par}
-{\Large Shanika Wickramasuriya\\
-and George Athanasopoulos\par}
+{\Large Dr. Shanika Wickramasuriya\\
+and Prof. George Athanasopoulos\par}
 
@@
 ***Keywords:** Coherent forecasts, Hierarchical time series, Covariance estimation, Australian tourism, Aggregation.*
+
+\pagebreak
+
+# Acknowledgements {-}
+
+I would like to express my sincere gratitude to my supervisors, Prof.
+George Athanasopoulos and Dr.
+Shanika Wickramasuriya, for their invaluable guidance, support, and encouragement throughout my Honours research journey.
+Their directions and insights have been instrumental not only in shaping this thesis but also in my personal and academic growth, letting me understand what it means to be a researcher.
+Despite being busy with their own work, they always made time to assist me whenever I knocked on their doors.
+I also thank for the laughs and light-hearted moments during our weekly meetings, which made the process enjoyable.
+
+I wish to thank Assoc.
+Prof.
+Natalia Bailey for being a wonderful Honours coordinator, always ready to help and provide constructive comments after each milestone, and for organising the Honours seminar series throughout the year.
+She also provided great advice when I was considering pursuing Honours initially, and I would say I made the right choice.
+My gratitude should also go to Prof.
+Jiti Gao for his advice on my decision to pursue Honours.
+
+I am grateful to my friends and peers, and to the professors and lecturers for their insightful classes throughout my Honours studies.
+I would also like to thank the Monash University Department of Econometrics and Business Statistics for providing a supportive and stimulating academic environment.
 

### 2025-10-23 thesis/paper_draft.qmd
re arrange heatmap in tourism

@@
 
-```{r tourism-MSE-line}
+```{r}
 #| label: fig-tourism-MSE-line
@@
 
+<!-- @fig-tourism-MSE-line and @fig-tourism-MSE-heat show that reconciliation is beneficial for point forecasts: across horizons $h=1,2,\dots,12$, all reconciled methods improve MSE over incoherent base ARIMA on average.
+Among vanilla MinT variants, the shrinkage estimator (*MinT-S*) marginally outperforms NOVELIST (*MinT-N*) at most horizons, though the gap is small.
+Adjusting for a single dominant factor via PC decomposition tightens performance further: both *MinT-S(PC1)* and *MinT-N(PC1)* beat their unadjusted counterparts, with *MinT-N(PC1)* narrowly ahead.
+Adding more than one PC brings no additional benefit, likely due to injecting estimation noise from weaker components.
+Variants that modify the multi-step covariance, either via scaled-variance or direct $h$-step residual covariances, underperform standard MinT, suggesting that extra estimation at horizon $h>1$ is not rewarded in this setting.
+-->
+
+<!--
+- describe the fig
+- say models in @tbl-cov
+- encoded by color, linetype, shape similar to simulation plot
+- quickly refresh the memory of what each means
+-->
+
+@fig-tourism-MSE-line reports percentage relative improvements in MSE of reconciled forecasts over the incoherent base ARIMA forecasts, across horizons from 1 to 12 months.
+We evaluate ten MinT variants (as defined in @tbl-cov), similar to the simulation's @fig-sim-results-1.
+The MinT with shrinkage (*MinT-S*) and its variants are colored in mint green, while MinT with NOVELIST (*MinT-N*) and its variants are in purple.
+Solid lines represent "vanilla" *MinT-S* and *MinT-N*; dashed lines with dot points denote the PC-adjusted variants (e.g.
+*MinT-S(PC1)*, *MinT-N(PC2)*); and dotted or dashed-dotted lines indicate the scaled variance and $h$-step-ahead residual versions (e.g.
+*MinT-S(SV)*, *MinT-N(hcov)*).
+
+Three patterns stand out.
+First, the vanilla *MinT-S* (solid green) slightly outperform *MinT-N* (solid purple) at most horizons, although the gap is small.
+Second, variants that alter the multi-step covariance, either via scaled variance or direct $h$-step residual covariances, underperform standard MinT, suggesting that extra estimation at horizon $h>1$ is not rewarded in this empirical analysis.
+Most importantly, incorporating a single dominant factor via PC-adjustment (dashed line with round points) delivers best performance among all methods.
+Both *MinT-S(PC1)* and *MinT-N(PC1)* consistently beat their unadjusted counterparts across horizons.
+However, adding more than one PC brings no additional benefit and can erode performance, likely due to injecting idiosyncratic noise from weaker components.
+
 ```{r}
@@
 #| fig.width: 7
-#| out.width: "90%"
+#| out.width: "85%"
 #| fig-cap: "Heatmap of relative improvement in the mean squared error (MSE) of different reconciled forecasts over the base forecasts for the Australian domestic tourism data, for 1 to 12 steps ahead forecasts.
@@
 ```
-
-<!-- @fig-tourism-MSE-line and @fig-tourism-MSE-heat show that reconciliation is beneficial for point forecasts: across horizons $h=1,2,\dots,12$, all reconciled methods improve MSE over incoherent base ARIMA on average.
-Among vanilla MinT variants, the shrinkage estimator (*MinT-S*) marginally outperforms NOVELIST (*MinT-N*) at most horizons, though the gap is small.
-Adjusting for a single dominant factor via PC decomposition tightens performance further: both *MinT-S(PC1)* and *MinT-N(PC1)* beat their unadjusted counterparts, with *MinT-N(PC1)* narrowly ahead.
-Adding more than one PC brings no additional benefit, likely due to injecting estimation noise from weaker components.
-Variants that modify the multi-step covariance, either via scaled-variance or direct $h$-step residual covariances, underperform standard MinT, suggesting that extra estimation at horizon $h>1$ is not rewarded in this setting.
--->
-
-<!--
-- describe the fig
-- say models in @tbl-cov
-- encoded by color, linetype, shape similar to simulation plot
-- quickly refresh the memory of what each means
--->
-
-@fig-tourism-MSE-line reports percentage relative improvements in MSE of reconciled forecasts over the incoherent base ARIMA forecasts, across horizons from 1 to 12 months.
-We evaluate ten MinT variants (as defined in @tbl-cov), similar to the simulation's @fig-sim-results-1.
-The MinT with shrinkage (*MinT-S*) and its variants are colored in mint green, while MinT with NOVELIST (*MinT-N*) and its variants are in purple.
-Solid lines represent "vanilla" *MinT-S* and *MinT-N*; dashed lines with dot points denote the PC-adjusted variants (e.g.
-*MinT-S(PC1)*, *MinT-N(PC2)*); and dotted or dashed-dotted lines indicate the scaled variance and $h$-step-ahead residual versions (e.g.
-*MinT-S(SV)*, *MinT-N(hcov)*).
-
-Three patterns stand out.
-First, the vanilla *MinT-S* (solid green) slightly outperform *MinT-N* (solid purple) at most horizons, although the gap is small.
-Second, variants that alter the multi-step covariance, either via scaled variance or direct $h$-step residual covariances, underperform standard MinT, suggesting that extra estimation at horizon $h>1$ is not rewarded in this empirical analysis.
-Most importantly, incorporating a single dominant factor via PC-adjustment (dashed line with round points) delivers best performance among all methods.
-Both *MinT-S(PC1)* and *MinT-N(PC1)* consistently beat their unadjusted counterparts across horizons.
-However, adding more than one PC brings no additional benefit and can erode performance, likely due to injecting idiosyncratic noise from weaker components.
 
@@
 
+\pagebreak
+
 # Conclusions and Future Work {#sec-conclusion}

### 2025-10-23 thesis/paper_draft.qmd
update appendix, include codes for sim supplemetary, reproducibility

@@
 <!-- ![Percentage relative improvement in Energy score in both hierarchies and time dimensions, for 1-step-ahead forecasts.
-The positive entries indicate a decrease in Energy score relative to base.](D:/Github/Recon_Honours_Thesis/figs/sim_energy.png){#fig-sim-results-2} -->
-
-```{r}
-#| label: fig-sim-results-2
+The positive entries indicate a decrease in Energy score relative to base.](D:/Github/Recon_Honours_Thesis/figs/sim_energy.png){#fig-sim-energy} -->
+
+```{r}
+#| label: fig-sim-energy
 #| fig-cap: "Percentage relative improvement in Energy score in both hierarchies and time dimensions, for 1-step-ahead forecasts.
@@
   theme(legend.position = "bottom", legend.title = element_blank())
-
 ```
@@
 Moving on to probabilistic forecasts, we evaluate the performance of reconciliation methods using the energy score, a proper scoring rule for multivariate predictive distributions.
-@fig-sim-results-2 presents the percentage relative improvement in energy score for 1-step-ahead forecasts.
+@fig-sim-energy presents the percentage relative improvement in energy score for 1-step-ahead forecasts.
 Across both hierarchies and time dimensions, the MinT methods consistently outperform the base forecasts.
@@
 
-Other scoring rules, such as the Winkler score and CRPS, yield similar conclusions.
-
-## Other Data Generating Processes
+Other scoring rules, such as the Winkler score and CRPS, yield similar conclusions. More detailed results are provided in Appendix @sec-sim-supplement.
+
+## Other Data Generating Processes {#sec-sim-other}
 
@@
 The idea is to allow NOVELIST to exploit the sparsity in the covariance structure, since it can control the sparsity of the shrinkage target via the thresholding parameter $\delta$.
-However, no profound insights can be drawn from the results.
+However, no profound insights can be drawn from the results (provided in Appendix @sec-sim-supplement).
 
@@
 
-```{r summary structure tourism}
+```{r}
+#| label: tbl-tourism-structure
 #| tbl-cap: "Hierarchical and grouped structure of Australian domestic tourism flows"
-#| label: tbl-tourism-structure
 
@@
 All data generation, covariance estimation, and reconciliation routines were implemented in the ReconCov R package and is available under an open‐source license on GitHub [@ReconCov].
+Regarding reproducibility, see Appendix @sec-reproducibility.
 
@@
 
+\vspace{-0.5em}
+
 ## Appendix: Simulation Supplementary {#sec-sim-supplement}
 
-## Appendix: Australian Domestic Tourism Geographical Hierarchy {#sec-tourism-supplement}
+\vspace{-0.5em}
+
+**Supplementary figures for simulation study in @sec-sim-hierarchy-size.**
+
+```{r}
+#| label: fig-appendix-sim-winkler
+#| fig-cap: "Percentage relative improvement in the Winkler score at 80% (left column) and 95% (right column) nominal coverage of different 1-step-ahead reconciled forecasts over the base forecasts for both hierarchical structures and time series lengths.
+The positive entries indicate a decrease in Winkler score relative to base."
+#| fig.height: 10
+#| fig.width: 8.5
+#| out.width: "83%"
+#| fig-pos: "H"
+sim_plots <- list()
+
+for (sim_name in names(sim_prob)) {
+  for (rule_name in c("W80", "W95")) {
+    p <- sim_prob[[sim_name]] |>
+      filter(rule == rule_name) |>
+      filter(model_label != "Base") |>
+      ggplot(aes(x = model_label, y = pct_change_avg, fill = family, alpha=pc_k)) +
+        geom_col(color = "black") +
+        scale_alpha_manual(values = c("0" = 1, "1" = 0.7, "2" = 0.4), guide = "none") +
+        recon_fill_scale +
+        labs(x = NULL, y = "% improvements") +
+        theme_minimal() +
+        theme(
+          axis.text.x = element_text(angle = 20, hjust = 0.5)
+        )
+    # remove x axis labels for all but last row
+    if (sim_name != "S100_T300") p <- p + theme(axis.text.x = element_blank())
+    name <- paste0(sim_name, "_", rule_name)
+    sim_plots[[name]] <- p
+  }
+}
+
+sim_plots[[1]] + labs(title = "Winkler score 80%", subtitle = "6 Groups of 6, T=50") +
+  sim_plots[[2]] + labs(title = "Winkler score 95%", subtitle = "6 Groups of 6, T=50") +
+  sim_plots[[3]] + labs(subtitle = "6 Groups of 6, T=300") +
+  sim_plots[[4]] + labs(subtitle = "6 Groups of 6, T=300") +
+  sim_plots[[5]] + labs(subtitle = "2 Groups of 50, T=50") +
+  sim_plots[[6]] + labs(subtitle = "2 Groups of 50, T=50") +
+  sim_plots[[7]] + labs(subtitle = "2 Groups of 50, T=300") +
+  sim_plots[[8]] + labs(subtitle = "2 Groups of 50, T=300") +
+  plot_layout(ncol = 2, guides = "collect", axis_titles = "collect") & 
+  theme(legend.position = "right", legend.title = element_blank())
+```
+
+To complement the main simulation probabilistic forecast results (using Energy score) in @sec-sim-hierarchy-size, @fig-appendix-sim-winkler presents percentage relative improvements in Winkler scores at 80% and 95% nominal coverage for 1-step-ahead forecasts across both hierarchical structures (small and large) and time series lengths (short and long).
+The patterns largely mirror those observed with the Energy score in @fig-sim-energy.
+
+\vspace{3em}
+
+**Supplementary figure for simulation study with sparse covariance structure in @sec-sim-other.**
+
+```{r}
+#| label: fig-appendix-sim-sparse-MSE
+#| fig-cap: "Percentage relative improvement in MSE of 1-to-12-step reconciled forecasts for sparse covariance (of the innovation process) setting, for the large 2-groups-of-50 hierarchy and both time series lengths.
+The positive entries indicate a decrease in MSE."
+#| fig.height: 6
+#| fig.width: 10
+#| out.width: "100%"
+#| fig-pos: "H"
+sim_sparse_MSE <- readRDS(paste0(path_lib, "sim_sparse_MSE.rds"))
+
+sparse_MSE_plots <- list()
+
+cap <- -6
+for (sim_name in names(sim_sparse_MSE)) {
+  plot_df <- sim_sparse_MSE[[sim_name]] |> 
+    mutate(
+      capped = pct_change <= cap,
+      capped_pct_change = pmax(pct_change, cap)
+    ) |> 
+    mutate(shape_lab = ifelse(shape_lab %in% c("PC1","PC2"), shape_lab, ""))
+  p <- plot_df |>
+    ggplot(aes(
+      x = h, y = capped_pct_change,
+      color = family, shape = shape_lab, linetype = variant_type
+    )) +
+      geom_line(linewidth = 0.8) +
+      geom_point(alpha = 0.85, size = 2.6, na.rm = TRUE) +
+      recon_scales +
+      theme_minimal() +
+      labs(x = "Forecast Horizon", y = "% improvements")
+  sparse_MSE_plots[[sim_name]] <- p
+}
+
+sparse_MSE_plots[[1]] + labs(subtitle = "T=50") +
+  sparse_MSE_plots[[2]] + labs(subtitle = "T=300") +
+  plot_layout(ncol = 2, guides = "collect", axis_titles = "collect") & 
+  theme(legend.position = "bottom", legend.direction = "vertical", legend.title.position = "left")
+```
+
+```{r}
+#| label: fig-appendix-sim-sparse-energy
+#| fig-cap: "Percentage relative improvement in Energy score of 1-step reconciled forecasts for sparse covariance (of the innovation process) setting, for the large 2-groups-of-50 hierarchy and both time series lengths.
+The positive entries indicate a decrease in Energy score."
+#| fig.height: 6
+#| fig.width: 10
+#| out.width: "90%"
+#| fig-pos: "H"
+sim_sparse_prob <- readRDS(paste0(path_lib, "sim_sparse_prob.rds"))
+sparse_energy_plots <- list()
+
+for (sim_name in names(sim_sparse_prob)) {
+  p <- sim_sparse_prob[[sim_name]] |>
+    filter(rule == "Energy") |>
+    filter(model_label != "Base") |>
+    ggplot(aes(x = model_label, y = pct_change_avg, fill = family, alpha=pc_k)) +
+      geom_col(color = "black") +
+      scale_alpha_manual(values = c("0" = 1, "1" = 0.7, "2" = 0.4), guide = "none") +
+      recon_fill_scale +
+      labs(x = "Model", y = "% improvements") +
+      theme_minimal() +
+      theme(
+        axis.text.x = element_text(angle = 20, hjust = 0.5)
+      )
+  sparse_energy_plots[[sim_name]] <- p
+}
+
+sparse_energy_plots[[1]] + labs(subtitle = "T=50") +
+  sparse_energy_plots[[2]] + labs(subtitle = "T=300") +
+  plot_layout(ncol = 2, guides = "collect", axis_titles = "collect") & 
+  theme(legend.position = "bottom", legend.title = element_blank())
+```
+
+\vspace{1em}
+
+## Appendix: Description of Australian Geographical Hierarchy {#sec-tourism-supplement}
 
@@
 
+## Appendix: Reproducibility {#sec-reproducibility}
+
+This research project has 3 Github repositories:
+
+1.
+**ReconCov**: The main *R package* implements the studied covariance estimators, MinT reconciliation, general data generating process functions.
+Refer to [@ReconCov] for installation and documentation.
+
+2.
+**Recon_Honours_Analysis**: This repository contains all R codes to produce the simulation studies and empirical analysis in this thesis (including scripts to run on a high-performance cluster), along with wrangling, processing, and analysing the results data.
+It depends on the *ReconCov* package.
+The access is upon request.
+
+3. **Recon_Honours_Thesis**: This repository contains the thesis manuscript in Quarto format, along with all figures and tables. The access is upon request.
+
+Please contact the author via email `Vincent.Su@monash.edu` for access to the analysis and thesis repositories.
+
 \pagebreak

### 2025-10-23 thesis/paper_draft.qmd
minor change

@@
 {\huge \bfseries Enhancing Forecast Reconciliation:\\
-A Study of\\
-Alternative Covariance Estimators\par}
+A Study of Alternative Covariance Estimators\par}
 
@@
 
-```{r tourism-probscores}
+```{r}
 #| label: fig-tourism-probscores

### 2025-10-24 thesis/paper_draft.qmd
final check + changes

@@
 Furthermore, the covariance estimate is based on in-sample 1-step-ahead forecast errors, then proportionally scale it to approximate $h$-step-ahead covariance matrix.
-This leaves a question of whether this method is appropriate for all real-world applications.
+This raises the question of the method's appropriateness for all real-world applications.
 We study the shortcomings of current practice and assess the impacts of alternative covariance estimators on MinT, including the NOVELIST estimator (shrinkage towards a soft-thresholded target), PC-adjusted shrinkage (which utilises latent factor structures), and horizon-specific estimators that relax proportional scaling.
@@
 Shanika Wickramasuriya, for their invaluable guidance, support, and encouragement throughout my Honours research journey.
-Their directions and insights have been instrumental not only in shaping this thesis but also in my personal and academic growth, letting me understand what it means to be a researcher.
-Despite being busy with their own work, they always made time to assist me whenever I knocked on their doors.
-I also thank for the laughs and light-hearted moments during our weekly meetings, which made the process enjoyable.
+Their guidance and insights have been instrumental not only in shaping this thesis but also in my personal and academic development, helping me understand what it means to be a researcher.
+Despite their own commitments, they always made time to assist me whenever I knocked on their office doors.
+I also thank them for the laughs and light‑hearted moments during our weekly meetings, which made the process enjoyable.
 
@@
 Prof.
-Natalia Bailey for being a wonderful Honours coordinator, always ready to help and provide constructive comments after each milestone, and for organising the Honours seminar series throughout the year.
-She also provided great advice when I was considering pursuing Honours initially, and I would say I made the right choice.
-My gratitude should also go to Prof.
-Jiti Gao for his advice on my decision to pursue Honours.
-
-I am grateful to my friends and peers, and to the professors and lecturers for their insightful classes throughout my Honours studies.
+Natalia Bailey for being a wonderful Honours coordinator, always ready to help, providing constructive comments after each milestone, and organising the Honours seminar series throughout the year.
+She also provided valuable advice when I was initially considering pursuing Honours, and I believe I made the right choice.
+I am also grateful to Prof.
+Jiti Gao for his advice regarding my decision to pursue Honours.
+
+I am grateful to my friends and peers, as well as to the professors and lecturers, for their insightful classes throughout my Honours studies.
 I would also like to thank the Monash University Department of Econometrics and Business Statistics for providing a supportive and stimulating academic environment.
+
+Finally, I would like to thank my family for their unwavering support and encouragement throughout my academic journey, allowing me to focus on my studies and research.
 
@@
 As a result, there are over 50,000 sales series across all *aggregation levels*, and decision makers need forecasts at each level to manage inventory and plan marketing strategies effectively.
-The hierarchy can be even more complex if we consider the sales of different categories of products (e.g., beverage, food, etc.) and the sales of product within each category (e.g., latte, cappuccino, etc.) at each aggregation level.
+The hierarchy can become even more complex when considering the sales of different product categories (e.g., beverages, food) and the sales of products within each category (e.g., lattes, cappuccinos) at each aggregation level.
 In this case, the structure is called a *grouped structure*, where the aggregation paths are not unique.
-Such structures also arises in many other decision-making contexts, from supply chains [@Seaman2022-bb; @Angam2025-od] and energy planning [@Di-Modica2021-ad], to macroeconomics [@El-Gemayel2022-hx; @Li2019-vv] and tourism analysis [@Athanasopoulos2009-lp].
+Such structures also arise in many other decision-making contexts, from supply chains [@Seaman2022-bb; @Angam2025-od] and energy planning [@Di-Modica2021-ad], to macroeconomics [@El-Gemayel2022-hx; @Li2019-vv] and tourism analysis [@Athanasopoulos2009-lp].
 Stakeholders in these settings need forecasts at several aggregation levels to allocate resources and manage risk.
@@
 Forecast reconciliation, a post-processing step, utilises the information from the hierarchical structure and data to adjust the initially produced base forecasts, so that the resulting *reconciled forecasts* are *coherent* (i.e., respecting the aggregation constraints).
-It was first introduced by @Hyndman2011-jv, and later developed by @Van_Erven2015-nx, @Hyndman2016-ic, @Ben-Taieb2019-yx, @Wickramasuriya2019-xq, @Wickramasuriya2020-uk, and others that focus point forecast reconciliation (more accurately, reconciling the means of predictive distributions).
-Recognising the important of probabilistic forecasting, probabilistic forecast reconciliation was studied by @Shang2017-ux, @Jeon2019-wm, and @Ben-Taieb2021-bn, and later formally defined by @Panagiotelis2023-fs.
+It was first introduced by @Hyndman2011-jv, and later developed by @Van_Erven2015-nx, @Hyndman2016-ic, @Ben-Taieb2019-yx, @Wickramasuriya2019-xq, @Wickramasuriya2020-uk, and others who focus on point forecast reconciliation (more accurately, reconciling the means of predictive distributions).
+Recognising the importance of probabilistic forecasting, probabilistic forecast reconciliation was studied by @Shang2017-ux, @Jeon2019-wm, and @Ben-Taieb2021-bn, and later formally defined by @Panagiotelis2023-fs.
 @Athanasopoulos2024-as provide a comprehensive review of the literature on forecast reconciliation.
 
-Among the modern methods, the Min Trace (MinT) approach developed by @Wickramasuriya2019-xq is is widely used due to its strong theoretical properties for minimising total reconciled forecast error variance, computational efficiency, and robust empirical performance.
+Among the modern methods, the Min Trace (MinT) approach developed by @Wickramasuriya2019-xq is widely used due to its strong theoretical properties for minimising total reconciled forecast error variance, computational efficiency, and robust empirical performance.
 MinT is later extended to probabilistic reconciliation by @Wickramasuriya2024-mb, showing that it also minimises the negative log score of the reconciled distribution under Gaussian assumptions.
-@Wickramasuriya2019-xq also argued that modelling spatial autocorrelations directly from the start would be challenging as in this case of a large collection of time series.
-Post-processing reconciliation has the advantage to implicitly model this spatial autocorrelation structure, especially true for MinT.
+@Wickramasuriya2019-xq also argued that modeling spatial autocorrelations directly from the start would be challenging, as in this case of a large collection of time series.
+Post-processing reconciliation has the advantage of implicitly modeling this spatial autocorrelation structure, especially true for MinT.
 MinT is implemented in popular R and Python software ecosystems [@O-Hara-Wild2024-we; @Nixtla].
@@
 Building on shrinkage, @Huang2019-ua proposed NOVELIST, which shrinks toward a soft-thresholded target rather than a diagonal one, improving flexibility to sparse targets.
-Extending beyond linear shrinkage, @Ledoit2012-ga (and further developed by @Ledoit2020-xt) introduced nonlinear shrinkage that replaces sample eigenvalues with data-driven nonlinear transformation of themselves.
-In a complementary direction, @Fan2013-jz introduce factor-based estimators that explicitly preserve latent low-rank structure while thresholding the idiosyncratic component.
+Extending beyond linear shrinkage, @Ledoit2012-ga (and further developed by @Ledoit2020-xt) introduced nonlinear shrinkage that replaces sample eigenvalues with data-driven nonlinear transformations of themselves.
+In a complementary direction, @Fan2013-jz introduced factor-based estimators that explicitly preserve latent low-rank structure while thresholding the idiosyncratic component.
 Related thresholding approaches, including hard thresholding [@Bickel2008-nh], generalised thresholding [@Rothman2009-yf], and adaptive thresholding [@Cai2011-jf], have also been developed.
@@
 
-When setting $\boldsymbol{M} = \boldsymbol{I}_n$, the identity matrix, the $\boldsymbol{P}$ matrix reduces to the OLS reconciliation, which corresponds to an orthogonal projection onto $\mathfrak{s}$.
+When setting $\boldsymbol{M} = \boldsymbol{I}_n$, the $\boldsymbol{P}$ matrix reduces to the OLS reconciliation $\boldsymbol{S} (\boldsymbol{S}' \boldsymbol{S})^{-1} \boldsymbol{S}'$, which corresponds to an orthogonal projection onto $\mathfrak{s}$.
 For a general positive definite matrix $\boldsymbol{M}$, $\boldsymbol{P}$ is idempotent but not necessarily symmetric, resulting in an oblique projection.
@@
 This projection view extends seamlessly from point to probabilistic forecast reconciliation.
-Given an $h$-step-ahead base forecast density with mean $\hat{\boldsymbol{y}}_{t+h|t}$ and covariance $\boldsymbol{W}_h$, via the linear projection with $\boldsymbol{P}$, the resulting reconciled forecast density has mean $\tilde{\boldsymbol{y}}_{t+h|t} = \boldsymbol{P} \hat{\boldsymbol{y}}_{t+h|t}$ and covariance $\boldsymbol{P} \boldsymbol{W}_h \boldsymbol{P}'$.
+Given an $h$-step-ahead base forecast density with mean $\hat{\boldsymbol{y}}_{t+h|t}$ and covariance $\boldsymbol{W}_h$, via the linear projection, the resulting reconciled forecast density has mean $\tilde{\boldsymbol{y}}_{t+h|t} = \boldsymbol{P} \hat{\boldsymbol{y}}_{t+h|t}$ and covariance $\boldsymbol{P} \boldsymbol{W}_h \boldsymbol{P}'$.
 @fig-geometry schematically illustrates this projection.
@@
 @Wickramasuriya2019-xq showed that by setting $\boldsymbol{M} = \boldsymbol{W}_h = \mathbb{E}( \hat{\boldsymbol{e}}_{t+h|t} \; \hat{\boldsymbol{e}}_{t+h|t}' )$, the covariance matrix of the $h$-step-ahead base forecast errors $\hat{\boldsymbol{e}}_{t+h|t} = \boldsymbol{y}_{t+h} - \hat{\boldsymbol{y}}_{t+h|t}$, we essentially minimise the total variance of the reconciled forecast errors across all series.
-Equivalently, MinT is the unique linear-unbiased reconciler that minimises the trace of $\text{Var}[\boldsymbol{y}_{t+h} - \tilde{\boldsymbol{y}}_{t+h|t}] = \boldsymbol{S} \boldsymbol{G}_h \boldsymbol{W}_h \boldsymbol{G}_h' \boldsymbol{S}'$.
-This method is thus called Minimum Trace (MinT) reconciliation.
+Equivalently, MinT is the unique linear-unbiased reconciler that minimises the trace of $Var[\boldsymbol{y}_{t+h} - \tilde{\boldsymbol{y}}_{t+h|t}] = \boldsymbol{S} \boldsymbol{G}_h \boldsymbol{W}_h \boldsymbol{G}_h' \boldsymbol{S}'$.
+This method is thus called Minimum Trace reconciliation.
 The matrix $\boldsymbol{G}_h$ is thus given by:
@@
 
-where, $\hat{\boldsymbol{R}}_{1,\delta}$ is the thresholded correlation matrix, where each element is regularised by:
+where, $\hat{\boldsymbol{R}}_{1,\delta}$ is the thresholded correlation matrix. Each element is regularised by:
 
@@
 On the other hand, the optimal threshold $\hat \delta$ does not have a closed-form solution, and is obtained by rolling-window cross-validation procedure.
-The idea is to find the threshold $\delta^*$, with the corresponding $\lambda^*$ and $\hat{\boldsymbol{R}}^{N}_{1}(\delta^*, \lambda^*)$, that minimises the average out-of-sample 1-step-ahead mean squared reconciled forecast errors over all windows.
+The idea is to find the threshold $\hat \delta$, with the corresponding $\hat \lambda = \hat \lambda (\hat \delta)$ and $\hat{\boldsymbol{R}}^{N}_{1}(\hat \delta, \hat \lambda)$, that minimises the average out-of-sample 1-step-ahead mean squared reconciled forecast errors over all windows.
 The formal algorithm is given in Algorithm \ref{alg:novelist_cv}.
@@
     \State $\hat{\boldsymbol{W}}_1 = \frac{1}{v} \sum_{t=j}^{i} \hat{\boldsymbol{e}}_{t} \hat{\boldsymbol{e}}_{t}'$
-    \State $\hat{\boldsymbol{D}} = \text{diag}(\hat{\boldsymbol{W}}_1)$
-    \State $\hat{\boldsymbol{R}}_1 = \hat{\boldsymbol{D}}^{-1/2} \hat{\boldsymbol{W}}_1 \hat{\boldsymbol{D}}^{-1/2}$
+    \State $\hat{\boldsymbol{D}}_1 = \text{diag}(\hat{\boldsymbol{W}}_1)$
+    \State $\hat{\boldsymbol{R}}_1 = \hat{\boldsymbol{D}}_1^{-1/2} \hat{\boldsymbol{W}}_1 \hat{\boldsymbol{D}}_1^{-1/2}$
     \For{$\delta \in \Delta$}
@@
         \State Compute $\hat{\boldsymbol{R}}^{N}_{1,\delta}$ using Equation 1
-        \State $\hat{\boldsymbol{W}}^{N}_{1,\delta} = \hat{\boldsymbol{D}}^{1/2} \hat{\boldsymbol{R}}^{N}_{1,\delta} \hat{\boldsymbol{D}}^{1/2}$
+        \State $\hat{\boldsymbol{W}}^{N}_{1,\delta} = \hat{\boldsymbol{D}}_1^{1/2} \hat{\boldsymbol{R}}^{N}_{1,\delta} \hat{\boldsymbol{D}}_1^{1/2}$
         \State $\boldsymbol{G} = (\boldsymbol{S}' \hat{\boldsymbol{W}}^{N^{-1}}_{1,\delta} \boldsymbol{S})^{-1} \boldsymbol{S}' \hat{\boldsymbol{W}}^{N^{-1}}_{1,\delta}$
@@
 
-\State $\text{MSE}_{\delta} = \frac{1}{T-v} \sum_{i=v}^{T-1} (\tilde{\boldsymbol{e}}_{i+1, \delta})^2$ for each $\delta \in \Delta$
-\State $\hat{\delta} = \arg\min_{\delta \in \Delta} \text{MSE}_{\delta}$
+\State $MSE_{\delta} = \frac{1}{T-v} \sum_{i=v}^{T-1} (\tilde{\boldsymbol{e}}_{i+1, \delta})^2$ for each $\delta \in \Delta$
+\State $\hat{\delta} = \arg\min_{\delta \in \Delta} MSE_{\delta}$
 
@@
 
-The general design of data generating process for bottom-level series is a stationary VAR(1) process, with the following structure:
+The general design of data-generating process for bottom-level series is a stationary VAR(1) process, with the following structure:
 
@@
 These steps result in a positive covariance matrix $\boldsymbol{\Sigma}^+$ where all the elements are positive.
-To allow a mixture of positive and negative covariance, we randomly flip the signs of them but need to maintain the positive definiteness.
+To allow a mixture of positive and negative covariance, we randomly flip the signs of them, but need to maintain the positive definiteness.
 This can be done by pre- and post-multiplying $\boldsymbol{\Sigma}$ by a random diagonal matrix $\boldsymbol{V}$ with diagonal entries sampled from $\{-1, 1\}$ with equal probability.
@@
 
-For all hierarchies in our experiments, we simulate two panel lengths, $T=54$ ("short") and $T=304$ ("long"), reserving the final four observations as an out-of-sample test set.
+For all hierarchies in our experiments, we simulate two panel lengths, $T=50$ ("short") and $T=300$ ("long"), generating extra four observations ($T+1$ to $T+4$) to serve as an out-of-sample test set.
 We perform 500 Monte Carlo replications for each configuration.
 In each replication, we fit univariate ARIMA models to the training observations using the automatic AICc minimization algorithm of @Hyndman2008-rf, implemented in the *fabletools* package [@O-Hara-Wild2024-we], generating 1–4-step base forecasts.
-We then reconcile the base forecasts using different reconciliation methods, and evaluate their point and probabilistic forecast accuracy on the test set.
+We then reconcile the base forecasts using different reconciliation methods and evaluate their point and probabilistic forecast accuracy on the test set.
 
@@
 
-In our main experiments, we examine how MinT combined with the different estimators perform as the hierarchy expands.
+In our main experiments, we examine how MinT combined with the different estimators performs as the hierarchy expands.
 We generate synthetic data from the VAR(1) framework described earlier, varying the number of bottom‐level series, $n_b$, across two structures: a "small" structure with six groups of six bottom-level series ($n_b = 6$x$6 = 36$), and a "large" configuration with two groups of fifty bottom-level series ($n_b = 2$x$50 = 100$).
@@
 Second, the PC-adjusted variants (using one and two principal components) do not yield improvements over the vanilla versions.
-This is expected since the synthetic data generating process does not simulate from strong latent factors.
+This is expected since the synthetic data-generating process does not simulate from strong latent factors.
 Lastly, the scaled variance and $h$-step-ahead residual approaches do not enhance performance as the forecast horizon increases, suggesting that the proportionality assumption may not be severely violated in this VAR(1) setup.
@@
 
-## Other Data Generating Processes {#sec-sim-other}
+## Other Data-Generating Processes {#sec-sim-other}
 
@@
 
-These findings motivates our turn to empirical data in the next section, where latent structural features, regime shifts, and noisy, intermittent series may possibly reveal performance differences.
+These findings motivate our turn to empirical data in the next section, where latent structural features, regime shifts, and noisy, intermittent series may possibly reveal performance differences.
 
@@
 The domestic tourism flows in Australia exhibit a natural hierarchical and grouped structure, driven both by geography and by purpose of travel.
-At the top of this hierarchy lies the national total, which splits into the seven states and territories.
-Each state is further sub-divided into tourism zones, which in turn break down into 77 regions.
+At the top of this hierarchy lies the national total, which is split into the seven states and territories.
+Each state is further subdivided into tourism zones, which in turn break down into 77 regions.
 A complete illustration of this geographic hierarchy appears in Appendix @sec-tourism-supplement.
@@
 We quantify tourism demand via "visitor nights", the total number of nights spent by Australians away from home.
-The data is collected via the National Visitor Survey, managed by Tourism Research Australia, using computer assisted telephone interviews from nearly 120,000 Australian residents aged 15 years and over [@tourism-au].
+The data is collected via the National Visitor Survey, managed by Tourism Research Australia, using computer-assisted telephone interviews from nearly 120,000 Australian residents aged 15 years and over [@tourism-au].
 
 The data are monthly time series spanning from January 1998 to December 2016, resulting in 228 observations per series.
-This gives a challenging high-dimensional forecasting problem with number of series (560) doubled the number of observations, which is ideal for evaluating reconciliation approaches that rely on high-dimensional covariance estimation.
+This gives a challenging high-dimensional forecasting problem with a number of series (560) doubling the number of observations, which is ideal for evaluating reconciliation approaches that rely on high-dimensional covariance estimation.
 The extreme dimensionality over sample size mirrors many contemporary business problems, for instance, Starbucks Corporation sales.
-Tourism demand is also economically vital yet highly volatile, with geographical and purpose‑specific patterns create a realistic stress‑test for reconciliation algorithms.
+Tourism demand is also economically vital yet highly volatile, with geographical and purpose‑specific patterns creating a realistic stress‑test for reconciliation algorithms.
 
@@
 The 1- to 12-step-ahead base forecasts are then generated by these ARIMA models, and then reconciled using multiple approaches.
-To estimate the NOVELIST and its variants, we implement an extra cross-validation prodecure within this training window, as described in Algorithm \ref{alg:novelist_cv}.
+To estimate the NOVELIST and its variants, we implement an extra cross-validation procedure within this training window, as described in Algorithm \ref{alg:novelist_cv}.
 We then roll the training window forward by one month and refit all models, re-estimate reconciliation variants, and produce another batch of 1- to 12-step-ahead forecasts, repeating until the training set reaches December 2015.
@@
 Three patterns stand out.
-First, the vanilla *MinT-S* (solid green) slightly outperform *MinT-N* (solid purple) at most horizons, although the gap is small.
+First, the vanilla *MinT-S* (solid green) slightly outperforms *MinT-N* (solid purple) at most horizons, although the gap is small.
 Second, variants that alter the multi-step covariance, either via scaled variance or direct $h$-step residual covariances, underperform standard MinT, suggesting that extra estimation at horizon $h>1$ is not rewarded in this empirical analysis.
-Most importantly, incorporating a single dominant factor via PC-adjustment (dashed line with round points) delivers best performance among all methods.
+Most importantly, incorporating a single dominant factor via PC-adjustment (dashed line with round points) delivers the best performance among all methods.
 Both *MinT-S(PC1)* and *MinT-N(PC1)* consistently beat their unadjusted counterparts across horizons.
@@
 The heatmap highlights the pattern from the line plot: *MinT-N(PC1)* and *MinT-S(PC1)* exhibit the highest concentration of dark cells and are repeatedly marked as the best method across horizons (star markers).
-Overall, the results underscores the consistent gains from PC adjustment and highlights the diminishing returns from more complex covariance treatments (lighter cells for scaled variance and $h$-step residuals).
+Overall, the results underscore the consistent gains from PC adjustment and highlight the diminishing returns from more complex covariance treatments (lighter cells for scaled variance and $h$-step residuals).
 
@@
 From our inspection, this is due to the overly shrunk variances from the shrinkage estimator, leading to narrow prediction intervals.
-The right panel confirms this, as *MinT-S* has the lowest empirical coverage across all levels, indicating that its prediction intervals fail to capture the true observations adequately and resulting in poor Winkler scores.
+The right panel confirms this, as *MinT-S* has the lowest empirical coverage across all levels, indicating that its prediction intervals fail to capture the true observations adequately and result in poor Winkler scores.
 *MinT-N* has relatively good coverage and improved Winkler score at bottom levels, but still is inferior at higher levels.
@@
 We thus explored alternatives that address them individually and in combination.
-Specifically, we adopt NOVELIST to introduce a more flexible shrinkage target; develop PC-adjusted variants of shrinkage and NOVELIST to preserve dominant latent components in the data; and consider multi-step covariance constructions that relax proportional scaling assumption.
+Specifically, we adopt NOVELIST to introduce a more flexible shrinkage target; develop PC-adjusted variants of shrinkage and NOVELIST to preserve dominant latent components in the data; and consider multi-step covariance constructions that relax the proportional scaling assumption.
 
@@
 First, more intricate simulation designs are needed to further highlight the differences between shrinkage and NOVELIST, particularly under explicit factor models and varying sparsity/noise levels.
-Second, our current PC-adjusted implementations do not standardise series prior to factor extraction, which may lead to higher-level series largely influencing the principal components.
+Second, our current PC-adjusted implementations do not standardise series before factor extraction, which may lead to higher-level series largely influencing the principal components.
 Third, we can explore alternative methods that leverage cross-series information or eigenvalue/eigenvector structures.
@@
 1.
-**ReconCov**: The main *R package* implements the studied covariance estimators, MinT reconciliation, general data generating process functions.
+**ReconCov**: The main *R package* implements the studied covariance estimators, MinT reconciliation, general data-generating process functions.
 Refer to [@ReconCov] for installation and documentation.

### 2025-10-24 thesis/paper_draft.qmd
add links for reproducible

@@
 It depends on the *ReconCov* package.
+It is available at [https://github.com/lordtahdus/Recon_Honours_Analysis](https://github.com/lordtahdus/Recon_Honours_Analysis)
+
+3.
+**Recon_Honours_Thesis**: This repository contains the thesis manuscript in Quarto format, along with all figures and tables.
 The access is upon request.
-
-3. **Recon_Honours_Thesis**: This repository contains the thesis manuscript in Quarto format, along with all figures and tables. The access is upon request.
-
-Please contact the author via email `Vincent.Su@monash.edu` for access to the analysis and thesis repositories.
+[https://github.com/lordtahdus/Recon_Honours_Thesis](https://github.com/lordtahdus/Recon_Honours_Thesis)
+
+Please contact the author via email `Vincent.Su@monash.edu` for specific navigation or questions regarding the code repositories.
 

### 2025-10-24 thesis/archived/archived.qmd
reorganise and remove old files

--- 2025-10-13 thesis/archived.qmd
@@
 # Old literature review (archived)
+
+In time series forecasting, aggregation occurs in a variety of settings.
+While a formal definition of hierarchical time series can be found in @sec-hierarchy, we can think of Starbucks sales data as an illustrative example.
+Starbucks operates in many countries, and each country has multiple cities where they have outlets.
+The sales data is structured *hierarchically*: the top level is the total sales across all countries, the next level down is the sales per country, followed by sales per city within each country, and finally sales per outlet within each city.
+As a result, there are over 40,000 sales series across all *aggregation levels*.
+The hierarchy can be even more complex if we consider the sales of different kinds of drinks (e.g., coffees, teas, refreshers) at each aggregation level.
+
+Forecasting data from such hierarchical structures also arises in many other decision-making contexts, from supply chains [@Seaman2022-bb; @Angam2025-od] and energy planning [@Di-Modica2021-ad], to macroeconomics [@El-Gemayel2022-hx; @Li2019-vv] and tourism analysis [@Athanasopoulos2009-lp].
+Stakeholders in these settings need forecasts at several aggregation levels to allocate resources and manage risk.
+The impact of methods for forecasting hierarchical time series has not been limited to academia, with industry also showing a strong interest.
+Many companies and organisations have adopted these methods in practice, including Amazon, the International Monetary Fund, IBM, SAP, and more [@Athanasopoulos2024-as].
+
+<!-- ##########  Literature Review  ########## -->
+
+<!-- - Talk about the history and evolution of forecasting hierarchical time series, starting from the early heuristic methods to the modern statistical approaches.
+
+  - Single level methods
+  - Optimal combination methods (OLS, WLS)
+  - MinT
+  - Bayesian, Machine learning
+  - Probabilistic methods -->
+
+In practice, when forecasts are produced for all series (often called *base forecasts*), they typically violate the aggregation constraints observed in the data; such forecasts are *incoherent*.
+This can undermine downstream decisions that require internal consistency.
+
+Traditionally, forecasting these hierarchical time series has been done using single-level methods, such as bottom-up, top-down, and middle-out approaches.
+Bottom-up methods involve generating forecasts for the bottom-level series and aggregating them to higher levels.
+Top-down methods start with forecasts for the only top-level series and disaggregate them down.
+Middle-out methods combine both approaches by forecasting middle-level series and then aggregating or disaggregating as needed.
+Despite their simplicity, these methods only anchor forecasts to a single level, implying a large loss of information on the hierarchy's inherent correlation structure.
+Additionally, the most disaggregated series often are very noisy or even intermittent, and the higher-level data might be smoother due to the aggregation.
+Furthermore, as we saw from the Starbucks example considering the sales of different kinds of drinks at each aggregation level -- formally defined as grouped structure in @sec-hierarchy -- the disaggregation becomes more complex since the disaggregation paths are not unique.
+Consequently, these single-level methods tend to give poor results across other levels of the hierarchy.
+
+To overcome these issues, forecast reconciliation was introduced by @Hyndman2011-jv, and later developed by @Van_Erven2015-nx, @Hyndman2016-ic, @Ben-Taieb2019-yx, @Wickramasuriya2019-xq, and others to achieve coherency in point forecasts and enhance accuracy.
+Forecast reconciliation projects a collection of independent base forecasts into a set of coherent forecasts that respect the linear constraints defining a hierarchical or grouped time-series system.
+
+Among the modern reconciliation strategies, the Min Trace (MinT) approach developed by @Wickramasuriya2019-xq is widely used and perform significantly well under right conditions.
+Despite its properties, MinT relies on a good high-dimensional covariance estimator, which is positive-definite.
+However, not many researchers have explored this issue in the current literature, except @Carrara2025-rz, who introduced a new estimator for MinT--Double Shrinkage estimator.
+The current stage of this research piece is fairly early, leaving many potential gaps still not yet to explore.
+As a result, a study of alternative high-dimensional covariance estimators for MinT is timely and desirable.
+
+The remainder of the paper is organised as follows.
+@sec-pre-theory provides the basic theoretical framework for hierarchical time series and forecasting, and Min Trace approach, introducing notations, terminologies, and motivations for alternative estimators.
+@sec-cov-est walks through the main covariance estimators this paper explores, and argues their strengths and weaknesses.
+@sec-simulation covers the simulation design and currently explores the performance of NOVELIST on MinT.
+@sec-empirical shows a real-world application of MinT with NOVELIST, which produces results that did not occured in the simulation settings, suggesting further inspection and analysis.
+
+<!-- Include more details on the gaps for each of these methods, such as bias, high variance? -->
+
+<!-- - Discuss the interests in MinT and how it has become a standard method for forecast reconciliation.
+
+- Discuss the MinT's reliance on a good estimate of the covariance matrix of base forecast errors and other gaps
+
+  - Empirical evidence of MinT under perform
+  - Comparison with other methods
+
+- Explain why this paper focus on exploring alternative covariance estimators for MinT. And is there any paper talk about this.
+
+  - Is better estimate of W_h really lead to better performance?
+
+- Talk about the paper outline and what will be covered in the following sections. -->
+
+## Even older
 
